### 2019-04-23 markdown/paper.Rmd
First build

---
title: "Rethinking Felony Disenfranchisement: The Case of New York City and Miami"
author: "Kevin Morris"
date: "April 22, 2019"
output: pdf_document
bibliography: "test.bib"
link-citations: yes
---

<!-- --- -->
<!-- title: "Rethinking Felony Disenfranchisement: The Case of New York City and Miami" -->

<!-- # to produce blinded version set to 1 -->
<!-- blinded: 0 -->

<!-- authors: -->
<!-- - name: Kevin Morris -->
<!--   thanks: The author gratefully acknowledges Jacob Faber and Jeff Manza for their advisement of this project. He would also like to thank colleagues at the Brennan Center for Justice for their insights into this project. -->
<!--   affiliation: Robert F. Wagner School of Public Service, New York University -->

<!-- keywords: -->
<!-- - 3 to 6 keywords -->
<!-- - that do not appear in the title -->

<!-- abstract: | -->
<!--   The text of your abstract.  200 or fewer words. -->

<!-- output: rticles::asa_article -->
<!-- --- -->

```{r, echo = F}
knitr::opts_knit$set(root.dir = "../..")
```

"There is no Negro problem. There is no Southern problem. There is no Northern problem. There is only an American problem"

-*Lyndon Johnson, March 15, 1965*

#Introduction

Felony disenfranchisement has received much attention over the past few years. This is true in both the academic world (where researchers have dug into the political ramifications in such races as the 2000 presidential election) and in the political world as states such as Florida and Louisiana have gradually moved to dismantle their systems of disenfranchisement. Throughout this process, however, many researchers have focused on the *individual* effects of felony disenfranchisement. They have examined the effect of disenfranchisement on the political participation of individuals, including after they are no longer disenfranchised, but have generally neglected to examine the intersection of spatially concentrated policing and incarceration patterns with the political effects of disenfranchisement. This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the character of disenfranchisement in New York City and Miami-Dade County in Florida.

This paper also seeks to complicate the narratives surrounding the end of felony disenfranchisement. Although advocates are correct to push for an end to disenfranchisement, a change to the political system may not be sufficient to reincorporate the voices of formerly incarcerated individuals into the democratic process. In addition to laying out the spatial implications of felony disenfranchisement in two major American cities, this paper also explores the impact of two different policies intended to dismantle the disenfranchisement program. Specifically, I investigate the effect of Executive Order 181 (which ended the disenfranchisement of parolees in New York State) and  Amendment 4 (which restored the voting rights of Floridians who had completed all terms of their sentences). The effects of these policies make clear that, while the end of felony disenfranchisement is necessary, active steps must be taken to reincorporate the formerly disenfranchised.

#Background of Felony Disenfranchisement in the United States

In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time. In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated. In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws]. This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to encompass any felony offense at all. The Supreme Court, in cases such as *Richardson v. Ramirez*, has upheld states' right to do just that. Collectively, these laws disenfranchise as many as ```r 6.1 - 1.4``` million American citizens. Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]

In any discussion of felony disenfranchisement in the United States, it is imperative to acknowledge the central role played by race and white supremacy. As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were intended to" [@Burch2010]. The historical record is undeniable. Previous research established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003]. In Florida felony disenfranchisement was added to the state constitution in 1868. Afterwards, a lawmaker boasted that the effort had been made in order to prevent Florida from being "niggerized" [@florida_1868].

The racial imbalance of felony disenfranchisement laws are not, of course, confined to the 19th century. Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement. Indeed, as the United States has vastly increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown. In XX states, as many as 25 percent of Black men are currently disenfranchised.

In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Christopher Uggen and Jeff Manza [-@Uggen2002]. George W. Bush's margin of victory in Florida in 2000 was famously just 537 votes. In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past. They estimate that, if this group had been allowed to vote, they would have supported Al Gore. Their enfranchisement, Uggen and Manza posit, would have tipped the presidential contest and resulted in the election of Al Gore.

#Spatial Effects of Concentrated Policing and Incarceration

Previous literature has called attention to the racially disparate impacts of felony disenfranchisement in the United States. It has often failed, however, to explore the ways in which the disenfranchisement of individuals leads to a loss of political power for community members

##concentrated policing
\newpage
#References

### 2019-04-23 code/initialize.R

@@ -1,0 +1,5 @@
+library(tidyverse)
+library(data.table)
+library(sqldf)
+library(kevostools)
+library(rgdal)

### 2019-04-23 code/maps.R

@@ -1,0 +1,110 @@
+## geocode matched individuals
+
+matches <- readRDS("./temp/matched_ids_17.rds")
+e16 <- fread("D:/rolls/new_york/misc/2016_elections.csv")
+
+# ## NYS
+# db <- dbConnect(SQLite(), "D:/rolls.db")
+# nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, res_city, zip5, nys_id, voter_status, history
+#                   from nys_roll_0319") %>% 
+#   filter(nys_id %in% matches$nys_id) %>% 
+#   mutate(street = paste("-", res_house_number, res_pre_street, res_street_name, res_post_street_dir, "-", sep = "-"),
+#          street = gsub("\\s+", " ", str_trim(gsub("-", " ", gsub("-NA-", " ", street)))),
+#          city = res_city,
+#          zip = zip5,
+#          state = "NY") %>% 
+#   select(street, city, zip, state, nys_id, voter_status, history)
+# 
+# nys <- nys %>% 
+#   mutate(a = as.integer(voter_status == "ACTIVE"),
+#          b = as.integer(voter_status == "INACTIVE"),
+#          c = as.integer(voter_status == "PREREG"))
+# 
+# nys <- setorder(nys, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+# nys <- nys[!duplicated(nys$nys_id),]
+# nys <- select(nys, -a, -b, -c)
+# 
+# nys <- geocode(nys) %>% 
+#   filter(longitude != 0)
+# saveRDS(nys, "./temp/geocoded_17.rds")
+
+nys <- readRDS("./temp/geocoded_17.rds")
+
+
+nys <- cSplit(nys, "history", sep = ";", direction = "long", type.convert = F) %>% 
+  group_by(nys_id) %>% 
+  mutate(v16 = max(history %in% e16$election)) %>% 
+  filter(row_number() == 1)
+
+nys$voted_16 <- ifelse(nys$v16 == 1, "Cast Ballot in 2016", "Didn't Cast Ballot in 2016")
+
+
+council_districts <- readOGR("./raw_data/shapefiles/nycc_19a", "nycc")
+council_districts <- spTransform(council_districts, CRS("+proj=longlat +datum=WGS84"))
+pings  <- SpatialPoints(nys[c('longitude','latitude')], proj4string = council_districts@proj4string)
[…]

### 2019-04-23 code/merge_rolls_doccs.R

@@ -1,0 +1,37 @@
+## read voter file
+
+db <- dbConnect(SQLite(), "D:/rolls.db")
+nys_roll <- dbGetQuery(db, "select last_name, first_name, middle_name, dob, voter_status, nys_id
+                       from nys_roll_0319
+                       where (voter_status != 'PURGED' and registration_date < 20170000) or
+                       (voter_status == 'PURGED' and registration_date < 20170000 and date_purged >= 20170000)")
+
+nys_roll <- nys_roll %>% 
+  mutate(a = as.integer(voter_status == "ACTIVE"),
+         b = as.integer(voter_status == "INACTIVE"),
+         c = as.integer(voter_status == "PREREG"))
+
+nys_roll <- setorder(nys_roll, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+nys_roll <- nys_roll[!duplicated(nys_roll$nys_id),]
+nys_roll <- select(nys_roll, -a, -b, -c)
+
+
+nys_roll <- nys_roll %>%
+  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA))) %>%
+  mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
+
+## read doccs data
+
+doccs_to_rolls <- readRDS("./temp/incarcerated_2017.rds") %>% 
+  rename(first_name = first,
+         middle_name = middle,
+         last_name = last) %>% 
+  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)))
+
+merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
+
+small <- merge_list[[1]] %>% 
+  select(nys_id, din) %>% 
+  filter(!is.na(nys_id), !is.na(din))
+
+saveRDS(small, "./temp/matched_ids_17.rds")

### 2019-04-23 code/read_doccs_data.R

@@ -1,0 +1,69 @@
+######################################################################################################################################
+# READ IN DOCCS DATA
+
+names = read.csv("./raw_data/doccs_data/incarcerated_column_names.csv", header = F, stringsAsFactors=FALSE)$V1 # VARIABLE NAMES SET IN CSV 'CAUSE I'M LAZY
+
+nys_released = read_fwf(
+  file = "./raw_data/doccs_data/releasees/Released.010190.113018.txt",
+  fwf_widths(c(7, 18, 11, 1, 3, 8, 1, 8, 1, 1, 1, 8, 8, # PDF FROM FOIL REQUEST INCLUDES WIDTH OF EACH VARIABLE
+               3, 12, 01, 3, 1, 1, 4, 30, 1, 1, 3, 2, 
+               12, 14, 1, 1, 4, 30, 1, 1, 3, 2, 12, 14, 1, 
+               1, 4, 30, 1, 1, 3, 2, 12, 14, 8, 8, 6, 
+               4, 8, 8, 6, 4, 8, 8, 3, 27, 8))
+)
+
+colnames(nys_released) = names
+rm(names)
+
+nys_released <- nys_released %>%
+  mutate(dob = as.Date(as.character(dob), "%Y%m%d"),
+         original_reception_date = as.Date(as.character(original_reception_date), "%Y%m%d"),
+         latest_reception_date = as.Date(as.character(latest_reception_date), "%Y%m%d"),
+         parole_eligible_date = as.Date(as.character(parole_eligible_date), "%Y%m%d"),
+         release_date = as.Date(as.character(release_date), "%Y%m%d")) %>% 
+  select(-starts_with("filler"))
+
+#########################################
+######### NOW READ PEOPLE CURRENTLY INCARCERATED
+names = read.csv("./raw_data/doccs_data/currently_incarcerated/column_names.csv", header = F, stringsAsFactors=FALSE)$V1 # VARIABLE NAMES SET IN CSV 'CAUSE I'M LAZY
+
+incarcerated = read_fwf(
+  file = "./raw_data/doccs_data/currently_incarcerated/UnderCustody.120118.txt",
+  fwf_widths(c(7, 18, 11, 1, 3, 8, 1, 8, 1, 1, 1, 8, 8, # PDF FROM FOIL REQUEST INCLUDES WIDTH OF EACH VARIABLE
+               3, 12, 01, 3, 1, 1, 4, 30, 1, 1, 3, 2, 
[…]

### 2019-04-24 markdown/paper.Rmd
keep writing

@@
 knitr::opts_knit$set(root.dir = "../..")
+library(scales)
 ```
@@
 Ramirez*, has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as ```r 6.1 - 1.4``` million American citizens.
+Collectively, these laws disenfranchise as many as `r 6.1 - 1.4` million American citizens.
 Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
@@
 
-The racial imbalance of felony disenfranchisement laws are not, of course, confined to the 19th century.
+The racial imbalance of felony disenfranchisement laws are not confined to the 19th century.
 Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
 Indeed, as the United States has vastly increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-In XX states, as many as 25 percent of Black men are currently disenfranchised.
+As of 2016, more than 10 percent of African Americans were disenfranchised in 9 states.
+In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
+Although Black adults made up just `r percent(29932674 / 247219588, accuracy = 0.1)` of the voting age population in 2016, they accounted for `r percent(2228118 / 6106327, accuracy = 0.1)` of the disenfranchised population [@sentencing_2016].
+
+#Academic Literature and Felony Disenfranchisement
 
@@
 Their enfranchisement, Uggen and Manza posit, would have tipped the presidential contest and resulted in the election of Al Gore.
+They based their estimates on the voting patterns of *non*-disenfranchised individuals who were demographically similar to the disenfranchised population.
+Though much of the research conducted since their 2002 study has undermined some of their key assumptions (namely, that formerly incarcerated individuals turnout and cast ballots similarly to those who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
+In the years after the Uggen and Manza published their paper, scholars sought to investigate the effect of felony disenfranchisement on African American turnout [@Miles2004], young adult turnout [@Hjalmarsson2010], and other questions.
+Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004].
+Others used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+
+In a series of papers in between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
+Haselswerdt [-@Haselswerdt2009] matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
+Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patters of formerly disenfranchised individuals in a range of states.
+She uses release data from states' Departments of Corrections and their Registered Voter Files to estimate the number of formerly incarcerated individuals who went on to register to vote.
+Using the Registered Voter Files, she is also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and the turnout rate among these individuals.
+Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [xx].
+
+These investigations into the implications of felony disenfranchisement have shed great light on the impact that restricting access to the ballot box might have on election outcomes, and for understanding the political participation of some of the most marginalized members of society.
+The literature has explored the intersection of race and disenfranchisement and discussed in broad terms how disproportionately disenfranchising Black men is hugely problematic.
+The literature has not, however, spent a great deal of time exploring the political implications of felony disenfranchisement for *communities* of color.
 

### 2019-04-24 markdown/paper.Rmd
more writing and research

@@
 ---
-
 <!-- --- -->
@@
 <!-- --- -->
-
-```{r, echo = F}
-knitr::opts_knit$set(root.dir = "../..")
+```{r setup, echo = F, message = F}
+knitr::opts_knit$set(root.dir = "..")
 library(scales)
+library(ggplot2)
+library(extrafont)
 ```
-
 "There is no Negro problem. There is no Southern problem. There is no Northern problem. There is only an American problem"
@@
 In the years after the Uggen and Manza published their paper, scholars sought to investigate the effect of felony disenfranchisement on African American turnout [@Miles2004], young adult turnout [@Hjalmarsson2010], and other questions.
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004].
-Others used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
+Others have used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
 
@@
 Haselswerdt [-@Haselswerdt2009] matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patters of formerly disenfranchised individuals in a range of states.
+Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
 She uses release data from states' Departments of Corrections and their Registered Voter Files to estimate the number of formerly incarcerated individuals who went on to register to vote.
@@
 
-These investigations into the implications of felony disenfranchisement have shed great light on the impact that restricting access to the ballot box might have on election outcomes, and for understanding the political participation of some of the most marginalized members of society.
-The literature has explored the intersection of race and disenfranchisement and discussed in broad terms how disproportionately disenfranchising Black men is hugely problematic.
-The literature has not, however, spent a great deal of time exploring the political implications of felony disenfranchisement for *communities* of color.
+A number of papers have also explored the impact that felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+King and Erickson [-@King2016], for instance, leverage state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among African Americans.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to determine individual-level turnout.
+They include estimates of the share of African Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout." They conclude that disenfranchisement has spillover effects for African American voters: where more African Americans are disenfranchised, elibible African Americans are less likely to cast a ballot.
+These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
+As @Bowers2009 sum up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty."
+
+These findings are not suprising given the large body of research establishing the social nature of voting.
+Outside the felony disenfranchisement conversation, an enormous body of research has demonstrated that individuals are far more likely to cast a ballot if their peers, family members or neighbors are doing so.
+[Build this out]
+
+Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
+Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the types of neighborhoods in which these effects are concentrated, and the impact these laws might have at the local level.
 
@@
 ##concentrated policing
+
+#Research Design: Spatial Implications of Felony Disenfranchisement in New York City and Miami
+
+##Matching
+##Geocoding
+
+#Maps
+
 \newpage

### 2019-04-24 code/initialize.R

@@ -1,5 +1,4 @@
 library(tidyverse)
 library(data.table)
-library(sqldf)
 library(kevostools)
 library(rgdal)

### 2019-04-25 markdown/paper.Rmd
More writing. Incorporating code from elsewhere

@@
 Using the Registered Voter Files, she is also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and the turnout rate among these individuals.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [xx].
+Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [@Meredith2013; @Meredith2015].
 
@@
 #Maps
+![\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016](../output/citywide_map.png)
+
+As we can see in Figure \ref{fig:citymap}, these are not distributed evenly throughout the city.
 

### 2019-04-25 code/initialize.R

@@ -2,3 +2,7 @@ library(tidyverse)
 library(data.table)
 library(kevostools)
 library(rgdal)
+library(sqldf)
+
+
+db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-04-25 code/maps.R

@@ -1,7 +1,7 @@
 ## geocode matched individuals
 
 matches <- readRDS("./temp/matched_ids_17.rds")
-e16 <- fread("D:/rolls/new_york/misc/2016_elections.csv")
+elects <- fread("./temp/elects.csv")
 
 # ## NYS
 # db <- dbConnect(SQLite(), "D:/rolls.db")
@@ -31,21 +31,28 @@ e16 <- fread("D:/rolls/new_york/misc/2016_elections.csv")
 nys <- readRDS("./temp/geocoded_17.rds")
 
 
-nys <- cSplit(nys, "history", sep = ";", direction = "long", type.convert = F) %>% 
+nys <- cSplit(nys, "history", sep = ";", direction = "long", type.convert = F)
+
+nys <- left_join(nys, elects, by = "history") %>% 
   group_by(nys_id) %>% 
-  mutate(v16 = max(history %in% e16$election)) %>% 
-  filter(row_number() == 1)
+  mutate(voted = max(year > 2006 & !is.na(year))) %>% 
+  filter(row_number() == 1) %>% 
+  ungroup()
 
-nys$voted_16 <- ifelse(nys$v16 == 1, "Cast Ballot in 2016", "Didn't Cast Ballot in 2016")
+nys$voted <- ifelse(nys$voted == 1, "Cast Ballot in Past 10 Years", "Didn't Cast Ballot in Past 10 Years")
 
+
+# get borough boundaries for projection
+bb <- readOGR("./raw_data/shapefiles/Borough Boundaries", "geo_export_14dc9d2c-e65a-48c5-ac4d-1bcd90c6758d")
+#
 
 council_districts <- readOGR("./raw_data/shapefiles/nycc_19a", "nycc")
-council_districts <- spTransform(council_districts, CRS("+proj=longlat +datum=WGS84"))
+council_districts <- spTransform(council_districts, CRS(bb@proj4string@projargs))
 pings  <- SpatialPoints(nys[c('longitude','latitude')], proj4string = council_districts@proj4string)
 nys$district   <- over(pings, council_districts)$CounDist
 
 nys <- filter(nys, longitude != 0) %>% 
-  arrange(v16)
+  arrange(voted)
 
 dists <- fortify(council_districts)
 
@@ -61,12 +68,10 @@ ggplot() +
         legend.key=element_blank()) +
   geom_polygon(data = dists, aes(x = long, y = lat, group = group), fill = "#bfbfbf") +
[…]

### 2019-04-25 code/merge_rolls_doccs.R

@@ -1,9 +1,9 @@
 ## read voter file
-
-db <- dbConnect(SQLite(), "D:/rolls.db")
+# pull everyone who is active now (and registered before 1/1/17)
+# or who is purged now but was registered on 1/1/17
 nys_roll <- dbGetQuery(db, "select last_name, first_name, middle_name, dob, voter_status, nys_id
                        from nys_roll_0319
-                       where (voter_status != 'PURGED' and registration_date < 20170000) or
+                       where (voter_status != 'PURGED' and registration_date < 20170000) or 
                        (voter_status == 'PURGED' and registration_date < 20170000 and date_purged >= 20170000)")
 
 nys_roll <- nys_roll %>% 

### 2019-04-25 markdown/paper.Rmd
figure out figures

@@
 date: "April 22, 2019"
-output: pdf_document
+output:
+  pdf_document:
+    fig_caption: yes
 bibliography: "test.bib"
@@
 ```{r setup, echo = F, message = F}
-knitr::opts_knit$set(root.dir = "..")
+knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf"
+                     )
 library(scales)
@@
 #Maps
-![\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016](../output/citywide_map.png)
+```{r figs, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016"}
+j <- readRDS("./output/city_map.RDS")
+j
+```
 
-As we can see in Figure \ref{fig:citymap}, these are not distributed evenly throughout the city.
+As we can see in Figure \ref{fig:citymap}, all these other observations
 

### 2019-04-25 code/initialize.R

@@ -3,6 +3,7 @@ library(data.table)
 library(kevostools)
 library(rgdal)
 library(sqldf)
-
+library(splitstackshape)
+library(extrafont)
 
 db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-04-25 code/maps.R

@@ -56,23 +56,22 @@ nys <- filter(nys, longitude != 0) %>%
 
 dists <- fortify(council_districts)
 
-ggplot() +
+city_map <- ggplot() +
   theme(axis.ticks = element_blank(),
         axis.text = element_blank(),
         panel.background = element_blank(),
         panel.border = element_blank(),
         legend.position = "bottom",
-        text = element_text(family = "Gadugi"),
         plot.title = element_text(hjust = 0.5),
         legend.background = element_blank(),
         legend.key=element_blank()) +
   geom_polygon(data = dists, aes(x = long, y = lat, group = group), fill = "#bfbfbf") +
   geom_path(data = dists, aes(x = long, y = lat, group = group), color = "black") +
-  geom_point(data = filter(nys, !is.na(district), voted == "Cast Ballot in Past 10 Years"), aes(x = longitude, y = latitude), color = "red") +
+  geom_point(data = filter(nys, !is.na(district), voted == "Cast Ballot in Past 10 Years"), aes(x = longitude, y = latitude), shape = 21, color = "black", fill = "red") +
   coord_map() +
   labs(x = NULL, y = NULL, caption = "Sources: NYSBOE, NYSDOCCS")
 
-ggsave("./output/citywide_map.png")
+saveRDS(city_map, "./output/city_map.RDS")
 
 ggplot() +
   theme(axis.ticks = element_blank(),

### 2019-04-29 markdown/paper.Rmd
Figure out cross referencing of tables

@@
 output:
-  pdf_document:
-    fig_caption: yes
-bibliography: "test.bib"
+  bookdown::pdf_document2:
+    toc: no
+bibliography: "citations.bib"
 link-citations: yes
@@
 <!-- --- -->
-```{r setup, echo = F, message = F}
+```{r setup, echo = F, message = F, warning = F}
 knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf"
@@
 library(scales)
-library(ggplot2)
+library(tidyverse)
 library(extrafont)
+library(kableExtra)
+library(stargazer)
 ```
@@
 
-#Introduction
+#Introduction {-}
 
@@
 
-#Background of Felony Disenfranchisement in the United States
+#Background of Felony Disenfranchisement in the United States {-}
 
@@
 
-#Academic Literature and Felony Disenfranchisement
+#Academic Literature and Felony Disenfranchisement {-}
 
@@
 
-#Spatial Effects of Concentrated Policing and Incarceration
+#Spatial Effects of Concentrated Policing and Incarceration {-}
 
@@
 
-##concentrated policing
+##concentrated policing {-}
 
-#Research Design: Spatial Implications of Felony Disenfranchisement in New York City and Miami
+#Research Design: Spatial Implications of Felony Disenfranchisement in New York City and Miami {-}
 
-##Matching
-##Geocoding
+##Matching {-}
+##Geocoding {-}
 
-#Maps
-```{r figs, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016"}
+#Maps {-}
+```{r citywide-map, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016"}
 j <- readRDS("./output/city_map.RDS")
@@
 
+#Match Output {-}
+
+```{r tract-table-chunk, echo = F}
+t <- readRDS("./temp/match_table_tract.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:tract-table) Results of Tract-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+j
+```
+
+Reg table
+```{r tract-reg-chunk, echo = F, results = "asis"}
+stargazer(readRDS("./temp/match_reg_tract.rds"),
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Tract Lost a Voter)", "Constant"),
+          dep.var.labels = c("Turnout Rate"),
+          title = "(\\#tab:tract-reg) Census Tract Matching Regression",
+          table.placement = "H")
+```
+xxx
+```{r block-table-chunk, echo = F}
+t <- readRDS("./temp/match_table_block_group.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+j
+```
+
+Reg table
+```{r block-reg-chunk, echo = F, results = "asis"}
+stargazer(readRDS("./temp/match_reg_block_group.rds"),
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Block Group Lost a Voter)", "Constant"),
+          dep.var.labels = c("Turnout Rate"),
+          title = "(\\#tab:blockgroup-reg) Census Block Group Matching Regression",
+          table.placement = "H")
+```
+more text
+
 \newpage
-#References
+#References {-}

### 2019-04-29 code/geocode_nyc.R

@@ -139,6 +139,4 @@ saveRDS(block_groups, "./temp/block_group_pre_match.rds")
 
 
 
-#### regression
-summary(lm(to ~ arrests + median_income + latino + nh_black + nh_white + some_college + median_age + vap + share_dem,
-           data = block_groups))
+

### 2019-04-29 code/turnout_match_regression.R

@@ -1,15 +1,15 @@
 ####### match at bg and tract level
 
-for(geo in c("tract")){
+for(geo in c("block_group", "tract")){
 
   units <- readRDS(paste0("./temp/", geo, "_pre_match.rds"))
   
   match_data <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem)
   
-  genout <- GenMatch(Tr = units$treat, X = match_data,
-                     M = 3, replace = T, pop.size = 1000)
-  saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
+  # genout <- GenMatch(Tr = units$treat, X = match_data,
+  #                    M = 3, replace = T, pop.size = 1000)
+  # saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
   
   genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))
   
@@ -52,7 +52,9 @@ for(geo in c("tract")){
   
   reg <- inner_join(units, match_w, by = "GEOID")
   
-  lm(to ~ treat, data = reg, weights = weight)
+  reg_output <- lm(to ~ treat, data = reg, weights = weight)
+  saveRDS(reg_output, paste0("./temp/match_reg_", geo, ".rds"))
+  
   
   ###########
   load(paste0("./temp/mout_", geo, ".RData"))
@@ -112,8 +114,15 @@ for(geo in c("tract")){
   
   colnames(df) <- c("", "Treated", "Control", "Treated", "Control", "Mean Diff", "eQQ Med", "eQQ Mean", "eQQ Max")
   
+  saveRDS(df, paste0("./temp/match_table_", geo, ".rds"))
+  
   kable(df, escape = FALSE, align = c('l', rep('c', 99))) %>%
     add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>%
     kable_styling(font_size = 12, full_width = F) %>%
     save_kable(file = paste0("./output/matches_", geo, ".html"), self_contained = T)
-  }
+}
+
+
+#### regression
+summary(lm(to ~ arrests + median_income + latino + nh_black + nh_white + some_college + median_age + vap + share_dem,
+           data = units))

### 2019-05-01 code/01_read_doccs_data.R
change matching procedure from treated if someone entered prison in 2017 to if someone was in prison in 2017

@@ -60,10 +60,20 @@ saveRDS(all_doccs, "./temp/all_doccs.rds")
 rm(incarcerated, nys_released)
 
 
-in_2017 <- all_doccs %>% 
-  filter(substring(latest_reception_date, 1, 4) == "2017",
+entered_in_2017 <- all_doccs %>% 
+  filter(gsub("[-]", "", latest_reception_date) >= "20170000",
+         gsub("[-]", "", latest_reception_date) < "20171107",
          current_felon == T) %>% 
   group_by(din) %>% 
   filter(row_number() == 1)
 
-saveRDS(in_2017, "./temp/incarcerated_2017.rds")
+saveRDS(entered_in_2017, "./temp/admit_2017.rds")
+
+in_2017 <- all_doccs %>% 
+  filter(gsub("[-]", "", latest_reception_date) < "20171107",
+         (release_date > "2017-11-07" | is.na(release_date)),
+         current_felon == T) %>% 
+  group_by(din) %>% 
+  filter(row_number() == 1)
+
+saveRDS(in_2017, "./temp/in_2017.rds")

### 2019-05-01 code/02_merge_rolls_doccs.R

@@ -1,10 +1,10 @@
 ## read voter file
-# pull everyone who is active now (and registered before 1/1/17)
-# or who is purged now but was registered on 1/1/17
+# pull everyone who is active now (and registered before 2017 deadline)
+# or who is purged now but was registered prior to deadline and purged after election
 nys_roll <- dbGetQuery(db, "select last_name, first_name, middle_name, dob, voter_status, nys_id
                        from nys_roll_0319
-                       where (voter_status != 'PURGED' and registration_date < 20170000) or 
-                       (voter_status == 'PURGED' and registration_date < 20170000 and date_purged >= 20170000)")
+                       where (voter_status != 'PURGED' and registration_date <= 20171013) or 
+                       (voter_status == 'PURGED' and registration_date <= 20171013 and date_purged > 20171107)")
 
 nys_roll <- nys_roll %>% 
   mutate(a = as.integer(voter_status == "ACTIVE"),
@@ -20,9 +20,9 @@ nys_roll <- nys_roll %>%
   mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA))) %>%
   mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
 
-## read doccs data
+## read doccs data - entered in 2017
 
-doccs_to_rolls <- readRDS("./temp/incarcerated_2017.rds") %>% 
+doccs_to_rolls <- readRDS("./temp/admit_2017.rds") %>% 
   rename(first_name = first,
          middle_name = middle,
          last_name = last) %>% 
@@ -34,4 +34,20 @@ small <- merge_list[[1]] %>%
   select(nys_id, din) %>% 
   filter(!is.na(nys_id), !is.na(din))
 
-saveRDS(small, "./temp/matched_ids_17.rds")
+saveRDS(small, "./temp/matched_ids_enter_17.rds")
+
+## read doccs data - in at all in 2017
+
+doccs_to_rolls <- readRDS("./temp/in_2017.rds") %>% 
+  rename(first_name = first,
+         middle_name = middle,
+         last_name = last) %>% 
+  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)))
+
+merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
+
+small <- merge_list[[1]] %>% 
+  select(nys_id, din) %>% 
[…]

### 2019-05-01 code/03_lost_voters_maps.R

@@ -1,7 +1,7 @@
 ## geocode matched individuals
 
-matches <- readRDS("./temp/matched_ids_17.rds")
-elects <- fread("./temp/elects.csv")
+matches <- readRDS("./temp/matched_ids_in_17.rds")
+elects <- fread("./raw_data/misc/elects.csv")
 
 ## NYS
 db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-05-01 code/04_geocode_nyc.R

@@ -41,7 +41,7 @@ nyc <- readRDS("./temp/nyc.rds")
 
 ##### 2017 ballots bg
 # read election names
-elects <- fread("./temp/elects.csv")
+elects <- fread("./raw_data/misc/elects.csv")
 
 history <- cSplit(dplyr::select(nyc, nys_id, history), "history", sep = ";", direction = "long", type.convert = F)
 

### 2019-05-01 code/05_turnout_match_reg.R

@@ -7,9 +7,9 @@ for(geo in c("block_group", "tract")){
   match_data <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem)
   
-  # genout <- GenMatch(Tr = units$treat, X = match_data,
-  #                    M = 3, replace = T, pop.size = 1000)
-  # saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
+  genout <- GenMatch(Tr = units$treat, X = match_data,
+                     M = 3, replace = T, pop.size = 1000)
+  saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
   
   genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))
   

### 2019-05-01 markdown/paper.Rmd
change matching procedure from treated if someone entered prison in 2017 to if someone was in prison in 2017
keep writing

@@
 
-##concentrated policing {-}
+#Framing for Turnout Effects {-}
 
-#Research Design: Spatial Implications of Felony Disenfranchisement in New York City and Miami {-}
+Although understanding whether felony disenfranchisement has impacts on turnout among different social groups (namely, racial minorities) is of great importance, understanding the explicitly spatial dimension of the problem is also an important task.
+Many political decisions bear on the spatial distribution of resources (XX); funding for parks or schools, for instance, can impact all members of a neighborhood, regardless of their race or their income.
+As such, even a resident who is not directly affected by felony disenfranchisement policies may lose political power thanks to these policies; if many of my neighbors cannot vote, I am joined by fewer voters in advocating for neighborhood-specific investment and services.
+This is, of course, tied up closely with race; New York City remains highly racially segregated, and therefore disentangling neighborhood effects from race effects becomes difficult.
+Nevertheless, these effects cross race lines in integrated neighborhoods.
+
+Of course, despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
+In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with XX actively registered voters.
+Despite the lower share of residents who are disenfranchised through imprisonment, there is reason to believe the policy impacts more individuals than just those imprisoned.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher.
+Understanding the distribution of this effect is important for understanding its political implications.
+Are voters who live in the same neighborhood as the disenfranchised less likely to cast a ballot because their neighbors have been removed from the community?
+Or does felony disenfranchisement operate cast its chilling effect on turnout more broadly, crossing even into neighborhoods where voters have not been removed from the community?
+
+This paper begins by examining whether felony disenfranchisement is in fact associated with lower turnout rates in a given neighborhood.^[As discussed below, "neighborhood" is difficult to define.
+I explore turnout differentials at both the census block group and census tract level.] As discussed above, this depressive effect on turnout is expected to operate through two distinct mechanisms: the actual incarceration and disenfranchisment of convicted individuals, and the social effects on eligible voters identified in previous research.
+These turnout effects are estimated in various ways: by comparing neighborhoods where voters with a history of participating have been incarcerated with similar neighborhoods where voters have not been incarcerated; by using regression analysis to investigate whether neighborhood arrests are correlated with turnout rates;; and by using regression analysis to investigate the relationship between prison admissions and turnout at the zip-code level.
+
+#Data {-}
+##Criminal Justice Data {-}
+
+This paper employs multiple data sets to investigate the extent to which a neighborhood is directly impacted by felony disenfranchisement.
+The primary data set comes from a freedom of information request filed by the author to obtain individual-level incarceration records for individuals sentenced to incarceration in New York State since 1992.
+The data includes a host of information, including: first and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; and others.
+This analysis is limited to individuals incarcerated for felony offenses.
+Individuals convicted of misdemeanors are not disenfranchised in New York State.
+
+The City of New York also makes the location of all NYPD arrests since 2006 available publicly.
+This data includes the latitude and longitude of the arrest, the date of the offense, and the class of the alleged offense.
+This analysis uses only arrests for felony charges.
+
+The third dataset used to estimate the extent to which a neighborhood is directly impacted by felony disenfranchisement policies in New York State comes from another freedom of information request.
+The individual-level incarceration data referenced above does not include information about the neighborhood an incarcerated person lived in prior to their conviction.
+Although the State of New York does not make this information available at the individual level, the author has received prison-admission counts by zip-code for the years XX - XX.
+
+##Voter File Data {-}
+Most states in the United States are required to maintain files with information on all registered voters in the state.
+In New York, this information is publicly available from the Board of Elections.
+It includes information on all registered voters, including: first and last name; date of birth; home address; vote history; and other information.
+The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
+Because the voter file includes information on voters who have been purged because of a felony conviction, we are able to reconstruct the vote history of individuals who are no longer registered because of their conviction.
+
+#Method {-}
 
 ##Matching {-}
+Method.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
+Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.]
 ##Geocoding {-}

### 2019-05-01 code/01_read_doccs_data.R

@@ -38,6 +38,8 @@ incarcerated = read_fwf(
 colnames(incarcerated) = names
 rm(names)
 
+saveRDS(nrow(incarcerated), "./temp/number_incarcerated.rds")
+
 incarcerated <- incarcerated %>%
   mutate(dob = as.Date(as.character(dob), "%Y%m%d"),
          original_reception_date = as.Date(as.character(original_reception_date), "%Y%m%d"),

### 2019-05-02 markdown/kmorris_thesis.Rmd
baby let me write write write write write

--- 2019-05-01 markdown/paper.Rmd
@@
 
+#What This Paper Examines {-}
+This paper begins by examining the effect of felony disenfranchisement in the New York City Mayoral election of 2017.
+This paper is primarily interested in whether felony disenfranchisement has impacts on local political distribution.
+If, for instance, voters in particular marked neighborhoods turnout at lower rates because of disenfranchisement policies, this could have ramifications for city-wide policy.
+@Martin2013, for instance, documents the relationship between voter turnout and the responsiveness of Congressional representatives.
+"[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
+
+Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
+Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
+As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
+An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
+Insofar as neighborhoods vote cohesively on issues such as school and park funding, their turnout rates may influence city policy.
+
 #Data {-}
@@
 
-#Method {-}
+##Geocoding {-}
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+These latitudes and longitudes were then converted to census block groups using the statistical program R and shapefiles available from the United States Census Bureau.
+
+#Matching Methodology {-}
+In order to estimate the impact of felony disenfranchisement at the neighborhood level, I begin by indentifying voters who were inelibile to cast a ballot in the 2017 mayoral election in New York City.
+In order to identify incarcerated and paroled voters, matching records from the Department of Corrections and the Voter File is necessary.
+I here use the methodology first developed by Burch [-@Burch2010; -@Burch2011].
+I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
+Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.]
 

### 2019-05-07 markdown/kmorris_thesis.Rmd
Update kmorris_thesis.Rmd
write

@@
     toc: no
-bibliography: "citations.bib"
+    fig_caption: yes
+bibliography: "../research/citations.bib"
 link-citations: yes
@@
 Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the types of neighborhoods in which these effects are concentrated, and the impact these laws might have at the local level.
+@Burch2013 is an exception to this.
+In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
+She determines that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City and by introducing a new way of
 
@@
 
-Previous literature has called attention to the racially disparate impacts of felony disenfranchisement in the United States.
-It has often failed, however, to explore the ways in which the disenfranchisement of individuals leads to a loss of political power for community members
+XX
 
@@
 I explore turnout differentials at both the census block group and census tract level.] As discussed above, this depressive effect on turnout is expected to operate through two distinct mechanisms: the actual incarceration and disenfranchisment of convicted individuals, and the social effects on eligible voters identified in previous research.
-These turnout effects are estimated in various ways: by comparing neighborhoods where voters with a history of participating have been incarcerated with similar neighborhoods where voters have not been incarcerated; by using regression analysis to investigate whether neighborhood arrests are correlated with turnout rates;; and by using regression analysis to investigate the relationship between prison admissions and turnout at the zip-code level.
+These turnout effects are estimated in various ways: by comparing neighborhoods where voters with a history of participating have been incarcerated with similar neighborhoods where voters have not been incarcerated; by using regression analysis to investigate whether neighborhood arrests are correlated with turnout rates; and by using regression analysis to investigate the relationship between prison admissions and turnout at the zip-code level.
 
@@
 "[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
+@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not."
 
@@
 Insofar as neighborhoods vote cohesively on issues such as school and park funding, their turnout rates may influence city policy.
+
+After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo the impact of felony disenfranchisement.
+In April of 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
+Today, only those in prison are disenfranchised] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
+In theory, such a policy could increase turnout and registration for these individuals.
+To test whether the policy was successful, I compare the turnout rates in the 2018 general election of three groups who finished parole in 2018: those who finished parole immediately before the policy change, those who finished parole immediately after the policy change, and those who finished parole immediately after the election.
+These three groups allow me to explore whether the re-extension of voting rights while on parole has an impact on turnout (by comparing the first two groups) and whether being under formal supervision while being allowed to vote results in higher turnout than being allowed to vote but being under supervision (by comparing the third group to the first two).
 
@@
 
+The state does not make a unified database of parolees whose voting rights have been restored available to the public.
+However, the Department of Corrections and Community Supervision website's Parolee Lookup includes a flag indicating whether someone's voting rights haave been restored.
+By using identification numbers from the data obtained from the state to lookup parolees on the website, I was able to construct a list of the individuals who were on parole and had their rights restored.
+
 ##Voter File Data {-}
@@
 
-#Matching Methodology {-}
-In order to estimate the impact of felony disenfranchisement at the neighborhood level, I begin by indentifying voters who were inelibile to cast a ballot in the 2017 mayoral election in New York City.
-In order to identify incarcerated and paroled voters, matching records from the Department of Corrections and the Voter File is necessary.
-I here use the methodology first developed by Burch [-@Burch2010; -@Burch2011].
+#Effects of Felony Disenfranchisement on Neighborhood Turnout Levels {-}
+
+##Identification of Lost Voters {-}
+Because I am interested in the spatial effects of felony disenfranchisement at the local level, I explore the impact of disenfranchisement on turnout in the most recent mayoral election, which occured on November 7^th^, 2017.
+I begin by indentifying "lost voters," defined as individuals who were ineligible to cast a ballot in the 2017 (either because of incarceration of parole status) and had cast a ballot between 2007 and 2016.
+Lost voters are of particular interest because of their demonstrated propensity to vote.
+I assume that disenfranchisement is likely to have a bigger impact where it directly impacts residents with a relatively recent history of casting a ballot.
+
+These individuals are identified by matching Department of Corrections and Community Supervisions data to the voter file using the methodology developed by Burch [-@Burch2010; -@Burch2011].
 I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.]
+Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison (and city council districts).
 
-##Matching {-}
-Method.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.]
+```{r citywide-map, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016", out.extra = '', fig.pos = "H", fig.align = "center"}
+j <- readRDS("./output/city_map.RDS")
+j
+```
+The spatial concentration of lost
+
 ##Geocoding {-}
@@
 #Maps {-}
-```{r citywide-map, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016"}
-j <- readRDS("./output/city_map.RDS")
-j
-```
 

### 2019-05-07 code/initialize.R

@@ -10,6 +10,4 @@ library(scales)
 library(kableExtra)
 library(stargazer)
 
-ttt
-
 db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-05-07 markdown/kmorris_thesis.Rmd
add individuals on parole as of 2017 election (was previously just incarcerated people)

@@
     fig_caption: yes
+    latex_engine: xelatex
 bibliography: "../research/citations.bib"
@@
 library(stargazer)
+
 ```
@@
 
-```{r citywide-map, echo = FALSE, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016", out.extra = '', fig.pos = "H", fig.align = "center"}
+```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016", out.extra = '', fig.pos = "H", fig.align = "center"}
 j <- readRDS("./output/city_map.RDS")
-j
+print(j)
 ```

### 2019-05-07 code/01_read_doccs_data.R

@@ -71,11 +71,68 @@ entered_in_2017 <- all_doccs %>%
 
 saveRDS(entered_in_2017, "./temp/admit_2017.rds")
 
+
+### INDIVIDUALS WHO WERE INCARCERATED AS OF 2017 ELECTION
 in_2017 <- all_doccs %>% 
   filter(gsub("[-]", "", latest_reception_date) < "20171107",
          (release_date > "2017-11-07" | is.na(release_date)),
          current_felon == T) %>% 
   group_by(din) %>% 
-  filter(row_number() == 1)
+  filter(row_number() == 1) %>% 
+  select(din, last, first, middle, dob)
 
-saveRDS(in_2017, "./temp/in_2017.rds")
+### INDIVIDUALS ON PAROLE AS OF 2017 ELECTION
+names <- read.csv("./raw_data/doccs_data/parolee_column_names.csv", header = F, stringsAsFactors=FALSE)$V1 # VARIABLE NAMES SET IN CSV 'CAUSE I'M LAZY
+
+parolees <- read_fwf(
+  file = "./raw_data/doccs_data/parolee_data/Parolee.011419.txt",
+  fwf_widths(c(1, 9, 3, 10, 3, 7, 3, 30, 3, 6, 3, 10, 3, 15, 3,
+               19, 3, 11, 3, 1, 3, 22, 3, 10, 3, 15, 2))
+)
+
+colnames(parolees) = names
+rm(names)
+
+parolees <- select(parolees, -starts_with("filler"))
+
+###
+names <- read.csv("./raw_data/doccs_data/parolee_crimes_column_names.csv", header = F, stringsAsFactors=FALSE)$V1 # VARIABLE NAMES SET IN CSV 'CAUSE I'M LAZY
+
+parolee_crimes <- read_fwf(
+  file = "./raw_data/doccs_data/parolee_data/Parole.Crimes.011419.txt",
+  fwf_widths(c(1, 9, 3, 7, 2, 3, 2, 44, 3, 1, 3, 11, 2))
+)
+
+colnames(parolee_crimes) = names
+rm(names)
+
+parolee_crimes <- parolee_crimes[parolee_crimes$crime_class %in% c("A", "B", "C", "D", "E"), ]$din
+
+parolees_felons <- parolees %>% 
+  filter(din %in% parolee_crimes,
+         release_date <= "2017-11-07",
+         (status == "ACTIVE" | (status_date > "2017-11-07")))
+
[…]

### 2019-05-07 code/02_merge_rolls_doccs.R

@@ -20,21 +20,21 @@ nys_roll <- nys_roll %>%
   mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA))) %>%
   mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
 
-## read doccs data - entered in 2017
-
-doccs_to_rolls <- readRDS("./temp/admit_2017.rds") %>% 
-  rename(first_name = first,
-         middle_name = middle,
-         last_name = last) %>% 
-  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)))
-
-merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
-
-small <- merge_list[[1]] %>% 
-  select(nys_id, din) %>% 
-  filter(!is.na(nys_id), !is.na(din))
-
-saveRDS(small, "./temp/matched_ids_enter_17.rds")
+# ## read doccs data - entered in 2017
+# 
+# doccs_to_rolls <- readRDS("./temp/admit_2017.rds") %>% 
+#   rename(first_name = first,
+#          middle_name = middle,
+#          last_name = last) %>% 
+#   mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)))
+# 
+# merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
+# 
+# small <- merge_list[[1]] %>% 
+#   select(nys_id, din) %>% 
+#   filter(!is.na(nys_id), !is.na(din))
+# 
+# saveRDS(small, "./temp/matched_ids_enter_17.rds")
 
 ## read doccs data - in at all in 2017
 

### 2019-05-07 code/03_lost_voters_maps.R

@@ -4,32 +4,33 @@ matches <- readRDS("./temp/matched_ids_in_17.rds")
 elects <- fread("./raw_data/misc/elects.csv")
 
 ## NYS
-db <- dbConnect(SQLite(), "D:/rolls.db")
-nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, res_city, zip5, nys_id, voter_status, history
-                  from nys_roll_0319") %>%
-  filter(nys_id %in% matches$nys_id) %>%
-  mutate_at(vars(res_house_number, res_pre_street, res_street_name, res_post_street_dir), funs(ifelse(is.na(.), "", .))) %>% 
-  mutate(street = paste(res_house_number, res_pre_street, res_street_name, res_post_street_dir),
-         street = gsub("\\s+", " ", street),
-         city = res_city,
-         zip = zip5,
-         state = "NY") %>%
-  select(street, city, zip, state, nys_id, voter_status, history)
+# db <- dbConnect(SQLite(), "D:/rolls.db")
+# nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, res_city, zip5, nys_id, voter_status, history
+#                   from nys_roll_0319") %>%
+#   mutate_at(vars(res_house_number, res_pre_street, res_street_name, res_post_street_dir), funs(ifelse(is.na(.), "", .))) %>%
+#   mutate(street = paste(res_house_number, res_pre_street, res_street_name, res_post_street_dir),
+#          street = gsub("\\s+", " ", street),
+#          city = res_city,
+#          zip = zip5,
+#          state = "NY") %>%
+#   select(street, city, zip, state, nys_id, voter_status, history)
+# 
+# nys <- nys %>%
+#   mutate(a = as.integer(voter_status == "ACTIVE"),
+#          b = as.integer(voter_status == "INACTIVE"),
+#          c = as.integer(voter_status == "PREREG"))
+# 
+# nys <- setorder(nys, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+# nys <- nys[!duplicated(nys$nys_id),]
+# nys <- select(nys, -a, -b, -c)
+# 
+# nys <- geocode(nys) %>%
+#   filter(longitude != 0) %>%
[…]

### 2019-05-08 markdown/kmorris_thesis.Rmd
Map wasn't working for lost voters on parole. Correct, run through

@@
 Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison (and city council districts).
+There were ```r comma(sum(readRDS("./temp/disen_by_bg.rds")$lost_voters))``` such lost voters within New York City as of the 2017 general election, and ```r comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))``` statewide.
 
-```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Individuals Incarcerated in 2017 Who Voted Between 2007 and 2016", out.extra = '', fig.pos = "H", fig.align = "center"}
+```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
 j <- readRDS("./output/city_map.RDS")

### 2019-05-08 code/04_geocode_nyc.R

@@ -1,142 +1,0 @@
-## geocode nyc
-
-# nyc <- dbGetQuery(db, "select dob, res_house_number, res_pre_street, res_street_name, res_post_street_dir, res_city,
-#                        zip5, nys_id, voter_status, last_name, county_code, gender, political_party, history
-#                   from nys_roll_0319
-#                   where voter_status == 'ACTIVE' and
-#                   county_code in (24, 41, 31, 3, 43)") %>%
-#   mutate(age = floor(dob/10000),
-#          gender = ifelse(gender == "F", 0, ifelse(gender == "M", 1, 2)),
-#          party = ifelse(toupper(political_party) == "DEM", 0, ifelse(toupper(political_party) == "REP", 1, 2)))
-# 
-# nyc <- nyc %>%
-#   mutate(a = as.integer(voter_status == "ACTIVE"),
-#          b = as.integer(voter_status == "INACTIVE"),
-#          c = as.integer(voter_status == "PREREG"))
-# 
-# nyc <- setorder(nyc, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
-# nyc <- nyc[!duplicated(nyc$nys_id),]
-# nyc <- dplyr::select(nyc, -a, -b, -c)
-# 
-# nyc <- nyc %>%
-#   mutate_at(vars(res_house_number, res_pre_street, res_street_name, res_post_street_dir), funs(ifelse(is.na(.), "", .))) %>% 
-#   mutate(street = paste(res_house_number, res_pre_street, res_street_name, res_post_street_dir),
-#              street = gsub("\\s+", " ", street),
-#              city = res_city,
-#              zip = zip5,
-#              state = "NY") %>%
-#   select(nys_id, county_code, street, city, zip, state, party, gender, age, history)
-# 
-# 
-# 
-# nyc <- geocode(nyc)
-# 
-# bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
-# pings  <- SpatialPoints(nyc[c('longitude','latitude')], proj4string = bg_shp@proj4string)
-# nyc$bg <- over(pings, bg_shp)$GEOID
-# 
-# saveRDS(nyc, "./temp/nyc.rds")
-
-nyc <- readRDS("./temp/nyc.rds")
-
-##### 2017 ballots bg
-# read election names
[…]

### 2019-05-08 code/04_process_nyc_nhoods.R

@@ -1,0 +1,112 @@
+## spatial join nyc
+
+# nyc <- readRDS("./temp/nys_geocoded.rds") %>% 
+#   filter(county_code %in% c(24, 41, 31, 3, 43))
+# 
+# bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
+# pings  <- SpatialPoints(nyc[c('longitude','latitude')], proj4string = bg_shp@proj4string)
+# nyc$bg <- over(pings, bg_shp)$GEOID
+# 
+# saveRDS(nyc, "./temp/nyc.rds")
+
+nyc <- readRDS("./temp/nyc.rds")
+
+##### 2017 ballots bg
+# read election names
+elects <- fread("./raw_data/misc/elects.csv")
+
+history <- cSplit(dplyr::select(nyc, nys_id, history), "history", sep = ";", direction = "long", type.convert = F)
+
+history <- left_join(history, elects, by = "history")
+history <- filter(history, year == 2017, election_type == "general")
+nyc$v2017 <- nyc$nys_id %in% history$nys_id
+##### arrests
+arrests <- fread("./raw_data/NYPD_Arrests_Data__Historic_.csv") %>% 
+  filter(LAW_CAT_CD == "F")
+
+bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
+pings  <- SpatialPoints(arrests[c('Longitude','Latitude')], proj4string = bg_shp@proj4string)
+arrests$bg <- over(pings, bg_shp)$GEOID
+
+arrests_bg <- arrests %>% 
+  group_by(GEOID = bg) %>% 
+  summarize(arrests = n())
+
+arrests_tract <- arrests %>% 
+  group_by(GEOID = substring(bg, 1, 11)) %>% 
+  summarize(arrests = n())
+
+## share dem
+share_dem_bg <- nyc %>% 
+  group_by(GEOID = bg) %>% 
+  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+            v2017 = sum(v2017),
+            vcount = n(),
+            lost_voters = sum(lost_voter))
+
+share_dem_tract <- nyc %>% 
+  group_by(GEOID = substring(bg, 1, 11)) %>% 
+  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+            v2017 = sum(v2017),
+            vcount = n(),
+            lost_voters = sum(lost_voter))
+
[…]

### 2019-05-08 code/05_turnout_match_reg.R

@@ -1,4 +1,27 @@
-####### match at bg and tract level
+library(tidyverse)
+library(Matching)
+library(data.table)
+library(snow)
+library(parallel)
+library(scales)
+library(kableExtra)
+
+## this can be run locally or on NYU's HPC. Set option in next step
+## option allowed because of how long GenMatch can take
+
+on_nyu <- F
+
+if(on_nyu){
+  setwd("/scratch/km3815/felony_disenfranchisement")
+  
+  NodeFile = Sys.getenv("MY_HOSTFILE")
+  
+  
+  cl <- makeCluster(c(readLines(NodeFile)), type="SOCK")
+}else{
+  source("./code/misc/AutoCluster4.R")
+  cl <- NCPUS(detectCores() - 1)
+}
 
 for(geo in c("block_group", "tract")){
 
@@ -8,7 +31,7 @@ for(geo in c("block_group", "tract")){
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem)
   
   genout <- GenMatch(Tr = units$treat, X = match_data,
-                     M = 3, replace = T, pop.size = 1000)
+                     M = 3, replace = T, pop.size = 1000, cluster = cl)
   saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
   
   genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))

### 2019-05-08 code/initialize.R

@@ -1,13 +1,0 @@
-library(Matching)
-library(tidyverse)
-library(data.table)
-library(kevostools)
-library(rgdal)
-library(sqldf)
-library(splitstackshape)
-library(extrafont)
-library(scales)
-library(kableExtra)
-library(stargazer)
-
-db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-05-08 code/00_initialize.R
Map wasn't working for lost voters on parole. Correct, run through

@@ -1,0 +1,23 @@
+library(Matching)
+library(tidyverse)
+library(data.table)
+library(kevostools)
+library(rgdal)
+library(sqldf)
+library(splitstackshape)
+library(extrafont)
+library(scales)
+library(kableExtra)
+library(stargazer)
+
+db <- dbConnect(SQLite(), "D:/rolls.db")
+
+## QUICK FUNCTION TO CLEAR THE ENVIRONMENT OF MOST OBJECTS AFTER EACH STEP
+## DOING SO TO WORK AROUND MEMORY CONSTRAINTS
+save <- c("db", "cleanup", "theme_bc", "save")
+
+
+cleanup <- function(...){
+  save2 <- c(save, ...)
+  rm(list=ls(envir = .GlobalEnv)[! ls(envir = .GlobalEnv) %in% save2], envir = .GlobalEnv)
+}

### 2019-05-08 code/03_lost_voters_maps.R

@@ -3,17 +3,18 @@
 matches <- readRDS("./temp/matched_ids_in_17.rds")
 elects <- fread("./raw_data/misc/elects.csv")
 
-## NYS
+# NYS
 # db <- dbConnect(SQLite(), "D:/rolls.db")
-# nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, res_city, zip5, nys_id, voter_status, history
+# nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, dob,
+#                        res_city, zip5, nys_id, voter_status, history, county_code, gender, political_party, last_name
 #                   from nys_roll_0319") %>%
 #   mutate_at(vars(res_house_number, res_pre_street, res_street_name, res_post_street_dir), funs(ifelse(is.na(.), "", .))) %>%
 #   mutate(street = paste(res_house_number, res_pre_street, res_street_name, res_post_street_dir),
-#          street = gsub("\\s+", " ", street),
+#          street = gsub("\\s+", " ", trimws(street)),
 #          city = res_city,
 #          zip = zip5,
 #          state = "NY") %>%
-#   select(street, city, zip, state, nys_id, voter_status, history)
+#   select(street, city, zip, state, nys_id, voter_status, history, county_code, gender, political_party, last_name)
 # 
 # nys <- nys %>%
 #   mutate(a = as.integer(voter_status == "ACTIVE"),
@@ -24,12 +25,11 @@ elects <- fread("./raw_data/misc/elects.csv")
 # nys <- nys[!duplicated(nys$nys_id),]
 # nys <- select(nys, -a, -b, -c)
 # 
-# nys <- geocode(nys) %>%
-#   filter(longitude != 0) %>%
-#   mutate(lost_voter = nys_id %in% matches$nys_id) %>%
+# nys <- geocode(nys)
 # saveRDS(nys, "./temp/nys_geocoded.rds")
 
 nys <- readRDS("./temp/nys_geocoded.rds") %>% 
+  mutate(lost_voter = nys_id %in% matches$nys_id) %>% 
   filter(lost_voter)
 
 

### 2019-05-10 markdown/kmorris_thesis.Rmd
Map wasn't working for lost voters on parole. Correct, run through
Switch to old voter file

@@
 Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison (and city council districts).
-There were ```r comma(sum(readRDS("./temp/disen_by_bg.rds")$lost_voters))``` such lost voters within New York City as of the 2017 general election, and ```r comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))``` statewide.
+There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
 
@@
 ```
+
+Regression Using Basic Analysis - No matching
 more text

### 2019-05-10 code/005_geocode_nys.R

@@ -1,0 +1,34 @@
+### GEOCODE WHOLE STATE
+### APRIL 2018 voter file to start
+
+
+nys <- dbGetQuery(db, "select res_house_number, res_pre_street, res_street_name, res_post_street_dir, dob,
+                       res_city, zip5, nys_id, voter_status
+                       from nys_roll_0418") %>%
+  mutate_at(vars(res_house_number, res_pre_street, res_street_name, res_post_street_dir), funs(ifelse(is.na(.), "", .))) %>%
+  mutate(street = paste(res_house_number, res_pre_street, res_street_name, res_post_street_dir),
+         street = gsub("\\s+", " ", trimws(street)),
+         city = res_city,
+         zip = zip5,
+         state = "NY") %>%
+  select(street, city, zip, state, nys_id, voter_status)
+
+nys <- nys %>%
+  mutate(a = as.integer(voter_status == "ACTIVE"),
+         b = as.integer(voter_status == "INACTIVE"),
+         c = as.integer(voter_status == "PREREG"))
+
+nys <- setorder(nys, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+nys <- nys[!duplicated(nys$nys_id),]
+nys <- select(nys, -a, -b, -c)
+
+nys <- geocode(nys)
+
+bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
+pings  <- SpatialPoints(nys[c('longitude','latitude')], proj4string = bg_shp@proj4string)
+nys$bg <- over(pings, bg_shp)$GEOID
+
+nys <- nys %>% 
+  select(nys_id, longitude, latitude, bg)
+
+saveRDS(nys, "./temp/nys_0418_geocoded.rds")

### 2019-05-10 code/02_merge_rolls_doccs.R

@@ -1,10 +1,7 @@
 ## read voter file
-# pull everyone who is active now (and registered before 2017 deadline)
-# or who is purged now but was registered prior to deadline and purged after election
+
 nys_roll <- dbGetQuery(db, "select last_name, first_name, middle_name, dob, voter_status, nys_id
-                       from nys_roll_0319
-                       where (voter_status != 'PURGED' and registration_date <= 20171013) or 
-                       (voter_status == 'PURGED' and registration_date <= 20171013 and date_purged > 20171107)")
+                       from nys_roll_0418")
 
 nys_roll <- nys_roll %>% 
   mutate(a = as.integer(voter_status == "ACTIVE"),

### 2019-05-10 code/04_process_nyc_nhoods.R

@@ -1,16 +1,26 @@
 ## spatial join nyc
 
-# nyc <- readRDS("./temp/nys_geocoded.rds") %>% 
-#   filter(county_code %in% c(24, 41, 31, 3, 43))
+# nyc <- dbGetQuery(db, "select dob, nys_id, voter_status, history, county_code, gender, political_party, last_name
+#                        from nys_roll_0418 where county_code in (24, 41, 31, 3, 43)")
 # 
-# bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
-# pings  <- SpatialPoints(nyc[c('longitude','latitude')], proj4string = bg_shp@proj4string)
-# nyc$bg <- over(pings, bg_shp)$GEOID
+# nyc <- nyc %>%
+#   mutate(a = as.integer(voter_status == "ACTIVE"),
+#          b = as.integer(voter_status == "INACTIVE"),
+#          c = as.integer(voter_status == "PREREG"))
 # 
+# nyc <- setorder(nyc, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+# nyc <- nyc[!duplicated(nyc$nys_id),]
+# nyc <- select(nyc, -a, -b, -c)
+# 
+# nyc <- left_join(nyc, readRDS("./temp/nys_0418_geocoded.rds"), by = "nys_id")
 # saveRDS(nyc, "./temp/nyc.rds")
 
 nyc <- readRDS("./temp/nyc.rds")
 
+nyc$lost_voter <- nyc$nys_id %in% readRDS("./temp/ids_of_lost_voters.rds")$nys_id
+
+lost_nyc <- sum(nyc$lost_voter)
+saveRDS(lost_nyc, "./temp/lost_count_nyc.rds")
 ##### 2017 ballots bg
 # read election names
 elects <- fread("./raw_data/misc/elects.csv")

### 2019-05-10 code/05_turnout_match_reg.R

@@ -1,17 +1,17 @@
-library(tidyverse)
-library(Matching)
-library(data.table)
-library(snow)
-library(parallel)
-library(scales)
-library(kableExtra)
-
 ## this can be run locally or on NYU's HPC. Set option in next step
 ## option allowed because of how long GenMatch can take
 
-on_nyu <- F
+on_nyu <- T
 
 if(on_nyu){
+  library(Matching)
+  library(data.table)
+  library(snow)
+  library(parallel)
+  library(scales)
+  library(kableExtra)
+  library(tidyverse)
+  
   setwd("/scratch/km3815/felony_disenfranchisement")
   
   NodeFile = Sys.getenv("MY_HOSTFILE")
@@ -138,14 +138,5 @@ for(geo in c("block_group", "tract")){
   colnames(df) <- c("", "Treated", "Control", "Treated", "Control", "Mean Diff", "eQQ Med", "eQQ Mean", "eQQ Max")
   
   saveRDS(df, paste0("./temp/match_table_", geo, ".rds"))
-  
-  kable(df, escape = FALSE, align = c('l', rep('c', 99))) %>%
-    add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>%
-    kable_styling(font_size = 12, full_width = F) %>%
-    save_kable(file = paste0("./output/matches_", geo, ".html"), self_contained = T)
+
 }
-
-
-#### regression
-summary(lm(to ~ arrests + median_income + latino + nh_black + nh_white + some_college + median_age + vap + share_dem,
-           data = units))

### 2019-05-10 code/06_basic_regressions.R

@@ -1,0 +1,15 @@
+### run regressions not using matching - this allows us to measure intensity
+
+
+block_groups <- readRDS("./temp/block_group_pre_match.rds")
+
+block_groups$lv2 <- block_groups$lost_voters^2
+
+bg_model <- lm(to ~ lost_voters + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
+bg_model2 <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
+bg_model2 <- lm(to ~ arrests + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
+
+tracts <- readRDS("./temp/tract_pre_match.rds")
+tracts$lv2 <- tracts$lost_voters^2
+tract_model <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)
+tract_model2 <- lm(to ~ arrests + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)

### 2019-05-13 markdown/kmorris_thesis.Rmd
Map areas impacted (from reg coeffs) Refine regression

@@
 #Match Output {-}
+```{r block-table-chunk, echo = F}
+t <- readRDS("./temp/match_table_block_group.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+j
+```
 
@@
 ```
+At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like our treatment group.
+As discussed above, the types of neighborhoods targeted by felony disenfranchisement look far different than the average New York City neighborhood.
+They have far lower incomes and are much less White; they also have lower median ages and more residents are registered as Democrats.
 
 Reg table
-```{r tract-reg-chunk, echo = F, results = "asis"}
-stargazer(readRDS("./temp/match_reg_tract.rds"),
+```{r tract-reg-chunk, echo = F, results = "asis", message = F}
+tract_reg <- readRDS("./temp/match_reg_tract.rds")
+tract_reg_black <- readRDS("./temp/match_reg_tract_nhb.rds")
+bg_reg <- readRDS("./temp/match_reg_block_group.rds")
+bg_reg_black <- readRDS("./temp/match_reg_block_group_nhb.rds")
+
+stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
+          column.separate = c(2, 2),
+          column.labels = c("Block Group Level", "Block Group Level", "Tract Level", "Tract Level"),
           header = F,
           type = "latex",
-          covariate.labels = c("D(Tract Lost a Voter)", "Constant"),
+          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) * Share Black", "Constant"),
           dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:tract-reg) Census Tract Matching Regression",
-          table.placement = "H")
+          title = "(\\#tab:tract-reg) Matching Regression",
+          table.placement = "H",
+          style = "ajps",
+          omit.stat = c("f", "ser"))
 ```
-xxx
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-
-Reg table
-```{r block-reg-chunk, echo = F, results = "asis"}
-stargazer(readRDS("./temp/match_reg_block_group.rds"),
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Block Group Lost a Voter)", "Constant"),
-          dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:blockgroup-reg) Census Block Group Matching Regression",
-          table.placement = "H")
-```
+Although the removal of a voter is not generally associated with decreased turnout, we *do* see an impact on turnout in the block groups and census tracts
 

### 2019-05-13 code/04_process_nyc_nhoods.R

@@ -137,6 +137,8 @@ tract_shp$GEOID <- with(tract_shp, paste0("36", cc, CT2010))
 
 tract_shp <- left_join(tract_shp, tracts, by = "GEOID")
 
+tract_shp$dec <- with(tract_shp, lost_voters > 0 & nh_black > (0.45))
+
 ggplot() +
   theme(axis.ticks = element_blank(),
         axis.text = element_blank(),
@@ -146,7 +148,39 @@ ggplot() +
         plot.title = element_text(hjust = 0.5),
         legend.background = element_blank(),
         legend.key=element_blank()) +
-  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group, fill = arrests > 10000), color = "black") +
+  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group, fill = dec), color = "black") +
   coord_map() +
-  labs(x = NULL, y = NULL) + 
-  ggtitle("Lost Voters by Tract")
+  labs(x = NULL, y = NULL) + scale_fill_manual(values = c("gray", "red"), na.translate = T, na.value = "gray") +
+  guides(fill = F)
+
+## bg
+
+block_groups$decrease <- block_groups$lost_voters > 0 & block_groups$nh_black > (2/3)
+dec <- block_groups[block_groups$lost_voters > 0 & block_groups$nh_black >= 0.65, "GEOID"]
+
+bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
+bg_shp <- spTransform(bg_shp, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+bg_shp@data$id <- rownames(bg_shp@data)
+temp <- fortify(bg_shp)
+bg_shp <- inner_join(temp, bg_shp@data, by = "id")
+rm(temp)
+
+bg_shp$dec <- bg_shp$GEOID %in% dec
+
+dep_to <- ggplot() +
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        panel.background = element_blank(),
+        panel.border = element_blank(),
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        legend.background = element_blank(),
+        legend.key=element_blank()) +
+  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "gray", color = "black", size = 0.01) +
[…]

### 2019-05-13 code/05_turnout_match_reg.R

@@ -1,7 +1,7 @@
 ## this can be run locally or on NYU's HPC. Set option in next step
 ## option allowed because of how long GenMatch can take
 
-on_nyu <- T
+on_nyu <- F
 
 if(on_nyu){
   library(Matching)
@@ -48,8 +48,12 @@ for(geo in c("block_group", "tract")){
   
   save(mout, file = paste0("./temp/mout_", geo, ".RData"))
   
+  load(paste0("./temp/mout_", geo, ".RData"))
+  
+  
   matches <- data.frame("treated" = mout[["index.treated"]],
-                        "control" = mout[["index.control"]])
+                        "control" = mout[["index.control"]],
+                        "weight" = mout[["weights"]])
   
   units <- units %>% 
     mutate(id = row_number())
@@ -65,18 +69,20 @@ for(geo in c("block_group", "tract")){
   matches <- left_join(matches, treat_row, by = c("treated" = "id"))
   matches <- left_join(matches, untreat_row, by = c("control" = "id"))
   
+  matches <- dplyr::select(matches, -treated, -control)
   
-  matches_v <- as.data.frame(c(matches$GEOID, matches$control_GEOID))
-  colnames(matches_v) = "GEOID"
+  matches_v <- bind_rows(dplyr::select(matches, GEOID, weight), dplyr::select(matches, GEOID = control_GEOID, weight))
   
   match_w <- matches_v %>% 
     group_by(GEOID) %>% 
-    summarize(weight = n())
+    summarize(weight = sum(weight))
   
   reg <- inner_join(units, match_w, by = "GEOID")
   
   reg_output <- lm(to ~ treat, data = reg, weights = weight)
+  reg_output_nhblack <- lm(to ~ treat + treat * nh_black, data = reg, weights = weight)
   saveRDS(reg_output, paste0("./temp/match_reg_", geo, ".rds"))
+  saveRDS(reg_output_nhblack, paste0("./temp/match_reg_", geo, "_nhb.rds"))
   
   
   ###########

### 2019-05-14 markdown/kmorris_thesis.Rmd
start coding zipcodes keep writing

@@
 In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
-She determines that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City and by introducing a new way of
+She determines that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City and by introducing a new way of measuring the impact.
 
@@
 ##Identification of Lost Voters {-}
-Because I am interested in the spatial effects of felony disenfranchisement at the local level, I explore the impact of disenfranchisement on turnout in the most recent mayoral election, which occured on November 7^th^, 2017.
-I begin by indentifying "lost voters," defined as individuals who were ineligible to cast a ballot in the 2017 (either because of incarceration of parole status) and had cast a ballot between 2007 and 2016.
-Lost voters are of particular interest because of their demonstrated propensity to vote.
-I assume that disenfranchisement is likely to have a bigger impact where it directly impacts residents with a relatively recent history of casting a ballot.
+In this analysis, I offer a slightly different defintion of "lost voter" than much of the literature.
+Many recent papers have attempted to identify relationships between the number of *potentially* lost voters and turnout.
+Here, I explore whether the disenfranchisement of voters with a history of participating in elections is related to neighborhood turnout.
+In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
+To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent election for city-wide office in New York - the mayoral election which took place on November 7^th^, 2017.
+Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
 
@@
 I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison (and city council districts).
+Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
@@
 ```
-The spatial concentration of lost
+The spatial concentration of lost is readily apparent.
+In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
+In other communities, such as Harlem and Central Brooklyn, large numbers of individuals were not allowed to cast a ballot for mayor.
 
-##Geocoding {-}
+## Testing for Neighborhood Turnout Effects {-}
 
-#Maps {-}
+To test whether neighborhoods with lost voters had systematically lower turnout rates than similar neighborhoods without such disqualifications, I begin by using a matching model.
+Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no residents were disqualified from the election because of a felony conviction.
+I use a genetic match to match treated to untreated census tracts and block groups.
+I match on a series of demographic and political indicators: racial characteristics, median income, education, age, population, share noncitizen,^[The Census Bureau does not make noncitizen estimates available at the block group level.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.] and partisan composition.
+Each treated census tract is matched to 10 untreated census tract, while each treated block group is treated to 30 untreated block groups.
+Matching is done with replacement.
+Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
 
-As we can see in Figure \ref{fig:citymap}, all these other observations
-
-#Match Output {-}
+### Match Output {-}
 ```{r block-table-chunk, echo = F}
@@
 At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like our treatment group.
-As discussed above, the types of neighborhoods targeted by felony disenfranchisement look far different than the average New York City neighborhood.
-They have far lower incomes and are much less White; they also have lower median ages and more residents are registered as Democrats.
+These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
+Neighborhoods with lost voters are substantially less white, have far lower median incomes, and a larger share of voters are registered as Democrats.
 
-Reg table
+After matching neighborhoods, I use a simple regression to test whether turnout in the 2017 mayoral election was different in areas with lost voters.
+Census tract and block group level turnout rates are calculated using the geocoded voter file.
+Each voter's record indicates whether the voter participated in the 2017 general election,^[The New York State voter file is not without errors.
+For instance, the voter file erroneously does not record the author as participating in the 2017 mayoral election.
+However, insofar as these errors are not geographically correlated, there is little reason to believe they impact the results of this analysis.
+Unfortunately, testing the distribution of these errors is not possible.] which are then aggregated to estimate the number of ballots cast in each neighborhood.
+The number of ballots cast is divided by estimated voting age population in the neighborhood as reported by the Census Bureau.^[Voting age population is of course only an approximation of the number of eligible voters; notably, its use will underestimate turnout in neighborhoods with large noncitizen populations.
+However, treatment and matched control neighborhoods have comparable rates of noncitizenship, lessening any potential bias.]
+
+Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
+I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
+
+Table \ref{tab:match-reg} presents the results of these regression models.
 ```{r tract-reg-chunk, echo = F, results = "asis", message = F}
@@
           column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Block Group Level", "Tract Level", "Tract Level"),
+          column.labels = c("Block Group Level", "Tract Level"),
           header = F,
@@
           dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:tract-reg) Matching Regression",
+          title = "(\\#tab:match-reg) Matching Regression",
           table.placement = "H",

### 2019-05-14 code/00_initialize.R

@@ -10,8 +10,6 @@ library(scales)
 library(kableExtra)
 library(stargazer)
 
-db <- dbConnect(SQLite(), "D:/rolls.db")
-
 ## QUICK FUNCTION TO CLEAR THE ENVIRONMENT OF MOST OBJECTS AFTER EACH STEP
 ## DOING SO TO WORK AROUND MEMORY CONSTRAINTS
 save <- c("db", "cleanup", "theme_bc", "save")
@@ -21,3 +19,5 @@ cleanup <- function(...){
   save2 <- c(save, ...)
   rm(list=ls(envir = .GlobalEnv)[! ls(envir = .GlobalEnv) %in% save2], envir = .GlobalEnv)
 }
+
+db <- dbConnect(SQLite(), "D:/rolls.db")

### 2019-05-14 code/04_process_nyc_nhoods.R

@@ -1,6 +1,6 @@
 ## spatial join nyc
-
-# nyc <- dbGetQuery(db, "select dob, nys_id, voter_status, history, county_code, gender, political_party, last_name
+# 
+# nyc <- dbGetQuery(db, "select dob, nys_id, voter_status, history, county_code, gender, political_party, last_name, zip5
 #                        from nys_roll_0418 where county_code in (24, 41, 31, 3, 43)")
 # 
 # nyc <- nyc %>%
@@ -61,38 +61,60 @@ share_dem_tract <- nyc %>%
             vcount = n(),
             lost_voters = sum(lost_voter))
 
-### block group / tract level
-# geos <- lapply(c("tract", "block group"), function(var) {
-#   income <- rbindlist(lapply(c("KINGS", "QUEENS", "BRONX", "NEW YORK", "RICHMOND"), function(c){
-#     income <- census_income(var, state = "NY", year = 2017, county = c) %>% 
-#       dplyr::select(-NAME)
+share_dem_zip <- nyc %>% 
+  group_by(GEOID = zip5) %>% 
+  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+            v2017 = sum(v2017),
+            vcount = n(),
+            lost_voters = sum(lost_voter))
+
+#block group / tract level
+# geos <- lapply(c("tract", "block group", "zcta"), function(var) {
+#   units <- rbindlist(lapply(c("KINGS", "QUEENS", "BRONX", "NEW YORK", "RICHMOND"), function(c, run = run){
+#     if(var != "zcta" | c == "KINGS"){
+#       if(var == "zcta"){
+#         c = NULL
+#         s = NULL
+#       }else{
+#         s = "NY"
+#       }
+#       income <- census_income(var, state = s, year = 2017, county = c) %>%
+#         dplyr::select(-NAME)
+#   
+#       race <- census_race_ethnicity(var, state = s, year = 2017, county = c) %>%
+#         dplyr::select(-NAME)
+#   
+#       education <- census_education(var, state = s, year = 2017, county = c) %>%
+#         dplyr::select(-NAME)
+#   
+#       age <- census_median_age(var, state = s, year = 2017, county = c)
+#   
+#       vap <- census_vap(var, state = s, year = 2017, county = c)
+#       
[…]

### 2019-05-14 code/05_turnout_match_reg.R

@@ -1,7 +1,7 @@
 ## this can be run locally or on NYU's HPC. Set option in next step
 ## option allowed because of how long GenMatch can take
 
-on_nyu <- F
+on_nyu <- T
 
 if(on_nyu){
   library(Matching)
@@ -28,7 +28,7 @@ for(geo in c("block_group", "tract")){
   units <- readRDS(paste0("./temp/", geo, "_pre_match.rds"))
   
   match_data <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen)
   
   genout <- GenMatch(Tr = units$treat, X = match_data,
                      M = 3, replace = T, pop.size = 1000, cluster = cl)
@@ -39,7 +39,7 @@ for(geo in c("block_group", "tract")){
   treat <- units$treat
   
   X <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen)
   
   match_count <- ifelse(geo == "tract", 10, 30)
   
@@ -90,7 +90,7 @@ for(geo in c("block_group", "tract")){
   order <- fread("./raw_data/misc/var_orders.csv")
   
   balance <- MatchBalance(treat ~ median_income + latino + nh_black + nh_white +
-                            some_college + median_age + vap + share_dem, match.out = mout,
+                            some_college + median_age + vap + share_dem + share_non_citizen, match.out = mout,
                           data = units)
   TrMean <- c()
   PreMean <- c()
@@ -115,7 +115,7 @@ for(geo in c("block_group", "tract")){
     PostQQmax <- unlist(c(PostQQmax, balance$AfterMatching[[i]]$qqsummary[3]))
   }
   
-  varnames <- c("median_income", "latino", "nh_black", "nh_white", "some_college", "median_age", "vap", "share_dem")
+  varnames <- c("median_income", "latino", "nh_black", "nh_white", "some_college", "median_age", "vap", "share_dem", "share_non_citizen")
   
   
   df <- data.frame("TrMean" = TrMean,

### 2019-05-14 markdown/kmorris_thesis.Rmd
Add matching for high intensity of lost voters Add city council competitiveness to matching

@@
 I use a genetic match to match treated to untreated census tracts and block groups.
-I match on a series of demographic and political indicators: racial characteristics, median income, education, age, population, share noncitizen,^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] and partisan composition.
+I match on a series of demographic and political indicators.
+Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau.
+Party affiliation rates come from the geocoded voter file.
+I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.
+All else equal, a more competitive race is likely to result in higher turnout.^[Where neighborhoods cross council district lines, this measure is the weighted average of the competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
 Each treated census tract is matched to 10 untreated census tract, while each treated block group is treated to 30 untreated block groups.

### 2019-05-14 code/001_voteshare_districts.R

@@ -1,0 +1,36 @@
+
+
+
+### get city council race numbers
+
+dists <- fread("./raw_data/misc/council_dists.csv")
+
+url_1 <- "https://www.vote.nyc.ny.us/downloads/csv/election_results/2017/20171107General%20Election/00"
+url_2 <- "022000"
+url_3 <- "%20Member%20of%20the%20City%20Council%20"
+url_4 <- "%20Council%20District%20EDLevel.csv"
+
+dists$url <- with(dists, paste0(url_1, county_num, url_2, str_pad(as.character(district), width = 2, pad = "0", side = "left"), county, url_3, district, suffix, url_4))
+
+
+results <- rbindlist(lapply(dists$url, function(url){
+  j <- fread(url)
+}))
+
+colnames(results) <- gsub("[.]", "_", make.unique(make.names(colnames(results))))
+
+results <- results[grepl("\\(", results$Unit_Name), ]
+
+results$Unit_Name <- gsub("\\s*\\([^\\)]+\\)", "", results$Unit_Name)
+
+results_ll <- results %>% 
+  group_by(District_Key, Unit_Name) %>% 
+  summarize(vote_count = sum(Tally, na.rm = T)) %>% 
+  group_by(District_Key) %>% 
+  mutate(total_votes = sum(vote_count),
+         share_winner = vote_count / total_votes) %>% 
+  filter(vote_count == max(vote_count)) %>% 
+  dplyr::select(district = District_Key, share_winner)
+  
+
+saveRDS(results_ll, "./temp/council_competitiveness.rds")

### 2019-05-14 code/04_process_nyc_nhoods.R

@@ -13,14 +13,26 @@
 # nyc <- select(nyc, -a, -b, -c)
 # 
 # nyc <- left_join(nyc, readRDS("./temp/nys_0418_geocoded.rds"), by = "nys_id")
+# 
+# council_districts <- readOGR("./raw_data/shapefiles/nycc_19a", "nycc")
+# council_districts <- spTransform(council_districts, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+# pings  <- SpatialPoints(nyc[c('longitude','latitude')], proj4string = council_districts@proj4string)
+# nyc$cc_district <- over(pings, council_districts)$CounDist
 # saveRDS(nyc, "./temp/nyc.rds")
 
 nyc <- readRDS("./temp/nyc.rds")
 
+## find lost voters
 nyc$lost_voter <- nyc$nys_id %in% readRDS("./temp/ids_of_lost_voters.rds")$nys_id
 
 lost_nyc <- sum(nyc$lost_voter)
 saveRDS(lost_nyc, "./temp/lost_count_nyc.rds")
+
+## read in city council races to control for competitiveness
+races <- readRDS("./temp/council_competitiveness.rds")
+
+nyc <- left_join(nyc, races, by = c("cc_district" = "district"))
+
 ##### 2017 ballots bg
 # read election names
 elects <- fread("./raw_data/misc/elects.csv")
@@ -52,21 +64,24 @@ share_dem_bg <- nyc %>%
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
-            lost_voters = sum(lost_voter))
+            lost_voters = sum(lost_voter),
+            share_winner = mean(share_winner))
 
 share_dem_tract <- nyc %>% 
   group_by(GEOID = substring(bg, 1, 11)) %>% 
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
-            lost_voters = sum(lost_voter))
+            lost_voters = sum(lost_voter),
+            share_winner = mean(share_winner))
 
 share_dem_zip <- nyc %>% 
   group_by(GEOID = zip5) %>% 
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
-            lost_voters = sum(lost_voter))
[…]

### 2019-05-14 code/05b_turnout_match_reg_high_intensity.R
Add matching for high intensity of lost voters Add city council competitiveness to matching

@@ -24,31 +24,32 @@ if(on_nyu){
 }
 
 for(geo in c("block_group", "tract")){
-
-  units <- readRDS(paste0("./temp/", geo, "_pre_match.rds"))
+  
+  units <- readRDS(paste0("./temp/", geo, "_pre_match.rds")) %>% 
+    filter(lost_voters == 0 | lost_voters >= 3)
   
   match_data <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen)
   
   genout <- GenMatch(Tr = units$treat, X = match_data,
                      M = 3, replace = T, pop.size = 1000, cluster = cl)
-  saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
+  saveRDS(genout, paste0("./temp/genout_", geo, "_hi.rds"))
   
-  genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))
+  genout <- readRDS(paste0("./temp/genout_", geo, "_hi.rds"))
   
   treat <- units$treat
   
   X <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen)
   
   match_count <- ifelse(geo == "tract", 10, 30)
   
   mout <- Match(Tr = treat, X = X, estimand = "ATT", Weight.matrix = genout, version = "fast", M = match_count)
   summary(mout)
   
-  save(mout, file = paste0("./temp/mout_", geo, ".RData"))
+  save(mout, file = paste0("./temp/mout_", geo, "_hi.RData"))
   
-  load(paste0("./temp/mout_", geo, ".RData"))
+  load(paste0("./temp/mout_", geo, "_hi.RData"))
   
   
   matches <- data.frame("treated" = mout[["index.treated"]],
@@ -81,16 +82,16 @@ for(geo in c("block_group", "tract")){
   
   reg_output <- lm(to ~ treat, data = reg, weights = weight)
   reg_output_nhblack <- lm(to ~ treat + treat * nh_black, data = reg, weights = weight)
[…]

### 2019-05-14 code/06_basic_regressions.R

@@ -1,5 +1,5 @@
 ### run regressions not using matching - this allows us to measure intensity
-
+nyc_zips <- fread("./raw_data/misc/zips_in_nyc.csv")
 
 block_groups <- readRDS("./temp/block_group_pre_match.rds")
 
@@ -13,3 +13,7 @@ tracts <- readRDS("./temp/tract_pre_match.rds")
 tracts$lv2 <- tracts$lost_voters^2
 tract_model <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)
 tract_model2 <- lm(to ~ lost_voters + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)
+
+zips$lv2 <- zips$lost_voters^2
+zip_model <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen,
+                data = filter(zips, GEOID %in% nyc_zips$V1))

### 2019-05-14 markdown/kmorris_thesis.Rmd
incorporate standard regression

@@
 All else equal, a more competitive race is likely to result in higher turnout.^[Where neighborhoods cross council district lines, this measure is the weighted average of the competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
-Each treated census tract is matched to 10 untreated census tract, while each treated block group is treated to 30 untreated block groups.
+Each treated neighborhood is matched with 10 untreated neighborhoods.
 Matching is done with replacement.
@@
 Unfortunately, testing the distribution of these errors is not possible.] which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by estimated voting age population in the neighborhood as reported by the Census Bureau.^[Voting age population is of course only an approximation of the number of eligible voters; notably, its use will underestimate turnout in neighborhoods with large noncitizen populations.
-However, treatment and matched control neighborhoods have comparable rates of noncitizenship, lessening any potential bias.]
+The number of ballots cast is divided by estimated citizen voting age population in the neighborhood as reported by the Census Bureau.
 
@@
 I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-
 Table \ref{tab:match-reg} presents the results of these regression models.
@@
 ```
-Although the removal of a voter is not generally associated with decreased turnout, we *do* see an impact on turnout in the block groups and census tracts
+
+### Testing Intensity Effects {-}
+
+Matching methodologies, of course, only allow us to test the effect of being treated - here, losing a voter for the 2017 election.
+The model above does not allow for different effects on turnout based on *how many* voters a neighborhood lost.
+In Table \ref{tab:trad-reg} below, I adopt a standard regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
+
+```{r trad-reg-chunk, echo = F, results = "asis", message = F}
+tract_reg <- readRDS("./temp/tract_model_reg_nosquare.rds")
+tract_reg_sq <- readRDS("./temp/tract_model_reg_square.rds")
+bg_reg <- readRDS("./temp/bg_model_reg_nosquare.rds")
+bg_reg_sq <- readRDS("./temp/bg_model_reg_square.rds")
+
+stargazer(bg_reg, bg_reg_sq, tract_reg, tract_reg_sq,
+          column.separate = c(2, 2),
+          column.labels = c("Block Group Level", "Tract Level"),
+          header = F,
+          type = "latex",
+          covariate.labels = c("Lost Voters", "Lost Voters\\textsuperscript{2}", "Median Income", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+                               "Percent With Some College", "Median Age", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
+          dep.var.labels = c("Turnout Rate"),
+          title = "(\\#tab:trad-reg) Standard Regression",
+          table.placement = "H",
+          style = "ajps",
+          omit.stat = c("f", "ser"))
+```
 

### 2019-05-14 code/07_turnout_match_reg.R

@@ -30,9 +30,9 @@ for(geo in c("block_group", "tract")){
   match_data <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
   
-  genout <- GenMatch(Tr = units$treat, X = match_data,
-                     M = 3, replace = T, pop.size = 1000, cluster = cl)
-  saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
+  # genout <- GenMatch(Tr = units$treat, X = match_data,
+  #                    M = 3, replace = T, pop.size = 1000, cluster = cl)
+  # saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
   
   genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))
   
@@ -41,7 +41,7 @@ for(geo in c("block_group", "tract")){
   X <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
   
-  match_count <- ifelse(geo == "tract", 10, 30)
+  match_count <- ifelse(geo == "tract", 10, 10)
   
   mout <- Match(Tr = treat, X = X, estimand = "ATT", Weight.matrix = genout, version = "fast", M = match_count)
   summary(mout)

### 2019-05-14 code/07b_turnout_match_reg_high_intensity.R

@@ -42,7 +42,7 @@ for(geo in c("block_group", "tract")){
   X <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
   
-  match_count <- ifelse(geo == "tract", 10, 30)
+  match_count <- ifelse(geo == "tract", 10, 10)
   
   mout <- Match(Tr = treat, X = X, estimand = "ATT", Weight.matrix = genout, version = "fast", M = match_count)
   summary(mout)

### 2019-05-14 code/08_basic_regressions.R

@@ -1,19 +1,18 @@
 ### run regressions not using matching - this allows us to measure intensity
-nyc_zips <- fread("./raw_data/misc/zips_in_nyc.csv")
 
 block_groups <- readRDS("./temp/block_group_pre_match.rds")
 
 block_groups$lv2 <- block_groups$lost_voters^2
 
-bg_model <- lm(to ~ lost_voters + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
-bg_model2 <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
-bg_model3 <- lm(to ~ arrests + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = block_groups)
+bg_model  <- lm(to ~ lost_voters +       median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = block_groups)
+saveRDS(bg_model, "./temp/bg_model_reg_nosquare.rds")
+bg_model2 <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = block_groups)
+saveRDS(bg_model2, "./temp/bg_model_reg_square.rds")
 
 tracts <- readRDS("./temp/tract_pre_match.rds")
 tracts$lv2 <- tracts$lost_voters^2
-tract_model <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)
-tract_model2 <- lm(to ~ lost_voters + median_income + nh_black + nh_white + some_college + median_age + share_dem, data = tracts)
 
-zips$lv2 <- zips$lost_voters^2
-zip_model <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen,
-                data = filter(zips, GEOID %in% nyc_zips$V1))
+tract_model  <- lm(to ~ lost_voters +       median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = tracts)
+saveRDS(tract_model, "./temp/tract_model_reg_nosquare.rds")
[…]

### 2019-05-16 markdown/kmorris_thesis.Rmd
clean up turnout effects regs

@@
 ---
+
 <!-- --- -->
@@
 
-#Introduction {-}
+# Introduction {-}
 
@@
 
-#Background of Felony Disenfranchisement in the United States {-}
+# Background of Felony Disenfranchisement in the United States {-}
 
@@
 
-#Academic Literature and Felony Disenfranchisement {-}
+# Academic Literature and Felony Disenfranchisement {-}
 
@@
 
-#Spatial Effects of Concentrated Policing and Incarceration {-}
+# Spatial Effects of Concentrated Policing and Incarceration {-}
 
@@
 
-#Framing for Turnout Effects {-}
+# Framing for Turnout Effects {-}
 
@@
 
-#What This Paper Examines {-}
+# What This Paper Examines {-}
 This paper begins by examining the effect of felony disenfranchisement in the New York City Mayoral election of 2017.
@@
 
-#Data {-}
-##Criminal Justice Data {-}
+# Data {-}
+## Criminal Justice Data {-}
 
@@
 
-##Voter File Data {-}
+## Voter File Data {-}
 Most states in the United States are required to maintain files with information on all registered voters in the state.
@@
 
-##Geocoding {-}
+## Geocoding {-}
 Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
@@
 
-#Effects of Felony Disenfranchisement on Neighborhood Turnout Levels {-}
+# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels {-}
 
-##Identification of Lost Voters {-}
+## Identification of Lost Voters {-}
 In this analysis, I offer a slightly different defintion of "lost voter" than much of the literature.
@@
 Party affiliation rates come from the geocoded voter file.
+Registration rate is calculated by dividing the number of registered voters (from the voter file) by the voting age population.
 I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.
 All else equal, a more competitive race is likely to result in higher turnout.^[Where neighborhoods cross council district lines, this measure is the weighted average of the competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
-Each treated neighborhood is matched with 10 untreated neighborhoods.
+Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
 Matching is done with replacement.
@@
 Unfortunately, testing the distribution of these errors is not possible.] which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by estimated citizen voting age population in the neighborhood as reported by the Census Bureau.
+The number of ballots cast is divided the voting age population.
 
@@
 ```
+When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
+In block groups with lost voters, turnout was on average 0.8 percent lower than in comparable block groups without lost voters.
+This decrease, however, appears to be entirely concentrated within Black block groups.
+When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes insignificant (p = 0.24).
+The coefficient on the interaction between treatment and share Black indicates that neighborhoods that are largely Black saw turnout that was as much as *3 percent lower* than similar neighborhoods without lost voters.
+These variables are largely insignificant at the census tract level.
+This is not particularly surprising.
+If lost voters have the largest depressive effect on family members and close neighbors, it makes sense that the effect of losing a voter is locally concentrated.
 
@@
           type = "latex",
-          covariate.labels = c("Lost Voters", "Lost Voters\\textsuperscript{2}", "Median Income", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
+          covariate.labels = c("Lost Voters", "Lost Voters * Share Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
           dep.var.labels = c("Turnout Rate"),
@@
 
-Regression Using Basic Analysis - No matching
-more text
-
 \newpage
-#References {-}
+# References {-}

### 2019-05-16 code/07_turnout_match_reg.R

@@ -26,23 +26,24 @@ if(on_nyu){
 for(geo in c("block_group", "tract")){
 
   units <- readRDS(paste0("./temp/", geo, "_pre_match.rds"))
-  
+
   match_data <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, reg_rate, share_dem, share_non_citizen, share_winner)
   
-  # genout <- GenMatch(Tr = units$treat, X = match_data,
-  #                    M = 3, replace = T, pop.size = 1000, cluster = cl)
-  # saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
+  genout <- GenMatch(Tr = units$treat, X = match_data,
+                     M = 3, replace = T, pop.size = 1000, cluster = cl)
+  saveRDS(genout, paste0("./temp/genout_", geo, ".rds"))
   
   genout <- readRDS(paste0("./temp/genout_", geo, ".rds"))
   
   treat <- units$treat
   
   X <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, vap, share_dem, share_non_citizen, share_winner)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college,
+                  median_age, reg_rate, share_dem, share_non_citizen, share_winner)
   
-  match_count <- ifelse(geo == "tract", 10, 10)
-  
+  match_count <- ifelse(geo == "tract", 10, 30)
+
   mout <- Match(Tr = treat, X = X, estimand = "ATT", Weight.matrix = genout, version = "fast", M = match_count)
   summary(mout)
 
@@ -90,7 +91,7 @@ for(geo in c("block_group", "tract")){
   order <- fread("./raw_data/misc/var_orders.csv")
   
   balance <- MatchBalance(treat ~ median_income + latino + nh_black + nh_white +
-                            some_college + median_age + vap + share_dem + share_non_citizen + share_winner, match.out = mout,
+                            some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, match.out = mout,
                           data = units)
   TrMean <- c()
   PreMean <- c()
@@ -115,7 +116,7 @@ for(geo in c("block_group", "tract")){
[…]

### 2019-05-16 code/08_basic_regressions.R

@@ -2,17 +2,20 @@
 
 block_groups <- readRDS("./temp/block_group_pre_match.rds")
 
-block_groups$lv2 <- block_groups$lost_voters^2
+block_groups$median_income <- block_groups$median_income / 1000
 
-bg_model  <- lm(to ~ lost_voters +       median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = block_groups)
+block_groups$lost_voters_black <- block_groups$lost_voters * block_groups$nh_black
+
+bg_model  <- lm(to ~ lost_voters +                    median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
 saveRDS(bg_model, "./temp/bg_model_reg_nosquare.rds")
-bg_model2 <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = block_groups)
+bg_model2 <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
 saveRDS(bg_model2, "./temp/bg_model_reg_square.rds")
 
 tracts <- readRDS("./temp/tract_pre_match.rds")
-tracts$lv2 <- tracts$lost_voters^2
+tracts$median_income <- tracts$median_income / 1000
+tracts$lost_voters_black <- tracts$lost_voters * tracts$nh_black
 
-tract_model  <- lm(to ~ lost_voters +       median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = tracts)
+tract_model  <- lm(to ~ lost_voters +                     median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = tracts)
 saveRDS(tract_model, "./temp/tract_model_reg_nosquare.rds")
-tract_model2 <- lm(to ~ lost_voters + lv2 + median_income + nh_black + nh_white + some_college + median_age + share_dem + share_non_citizen + share_winner, data = tracts)
+tract_model2 <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = tracts)
 saveRDS(tract_model2, "./temp/tract_model_reg_square.rds")

### 2019-05-16 code/09_post_est_maps.R

@@ -1,0 +1,51 @@
+
+###### maps
+tracts <- readRDS("./temp/tract_pre_match.rds")
+block_groups <- readRDS("./temp/block_group_pre_match.rds")
+
+tract_shp <- readOGR("./raw_data/shapefiles/nyct2010_19a", "nyct2010")
+tract_shp <- spTransform(tract_shp, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+tract_shp@data$id <- rownames(tract_shp@data)
+temp <- fortify(tract_shp)
+tract_shp <- inner_join(temp, tract_shp@data, by = "id")
+rm(temp)
+
+codes <- data.frame("BoroCode" = as.character(c(1:5)), "cc" = c("061", "005", "047", "081", "085"))
+
+tract_shp <- left_join(tract_shp, codes, by = "BoroCode")
+
+tract_shp$GEOID <- with(tract_shp, paste0("36", cc, CT2010))
+
+tract_shp <- left_join(tract_shp, tracts, by = "GEOID")
+
+tract_shp$decrease <- -0.022 * tract_shp$lost_voters * tract_shp$nh_black
+
+
+
+bg_shp <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
+bg_shp <- spTransform(bg_shp, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+bg_shp@data$id <- rownames(bg_shp@data)
+temp <- fortify(bg_shp)
+bg_shp <- inner_join(temp, bg_shp@data, by = "id")
+rm(temp)
+
+bg_shp <- inner_join(bg_shp, block_groups, by = "GEOID")
+
+bg_shp$decrease <- -0.027 * bg_shp$lost_voters * bg_shp$nh_black
+
+
+
+ggplot() +
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        panel.background = element_blank(),
+        panel.border = element_blank(),
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        legend.background = element_blank(),
+        legend.key=element_blank()) +
+  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "gray") +
+  geom_polygon(data = bg_shp, aes(x = long, y = lat, group = group, fill = decrease)) +
+  coord_map() +
+  labs(x = NULL, y = NULL) + scale_fill_gradient(high = "gray", low = "red") +
+  guides(fill = F)

### 2019-05-16 markdown/kmorris_thesis.Rmd
post estimation maps

@@
           type = "latex",
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) * Share Black", "Constant"),
+          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) X Share Black", "Constant"),
           dep.var.labels = c("Turnout Rate"),
@@
           type = "latex",
-          covariate.labels = c("Lost Voters", "Lost Voters * Share Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+          covariate.labels = c("Lost Voters", "Lost Voters X Share Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
                                "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
@@
 
+The block groups where these depressive effects are not, of course, randomly distributed througout the city.
+They are highly spatially concentrated: in Central Brooklyn, Eastern Queens, and Harlem.
+Figure \ref{fig:postest} applies the coefficient on "Lost Voters * Share Black" from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
+
+```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Depressive Effect of Felony Disenfranchisement", out.extra = '', fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/dec_block_map.rds")
+print(j)
+```
+
 \newpage

### 2019-05-16 code/09_post_est_maps.R

@@ -35,7 +35,7 @@ bg_shp$decrease <- -0.027 * bg_shp$lost_voters * bg_shp$nh_black
 
 
 
-ggplot() +
+dec_map <- ggplot() +
   theme(axis.ticks = element_blank(),
         axis.text = element_blank(),
         panel.background = element_blank(),
@@ -43,9 +43,13 @@ ggplot() +
         legend.position = "bottom",
         plot.title = element_text(hjust = 0.5),
         legend.background = element_blank(),
-        legend.key=element_blank()) +
+        legend.key=element_blank(),
+        legend.key.width = unit(1.5, "cm")) +
   geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "gray") +
   geom_polygon(data = bg_shp, aes(x = long, y = lat, group = group, fill = decrease)) +
   coord_map() +
-  labs(x = NULL, y = NULL) + scale_fill_gradient(high = "gray", low = "red") +
-  guides(fill = F)
+  labs(x = NULL, y = NULL) + 
+  scale_fill_gradient(high = "gray", low = "red", labels = percent) +
+  guides(fill = guide_colorbar(title.hjust = .5, title.position = "top", title = "Decreased Turnout from Disenfranchisement"))
+
+saveRDS(dec_map, "./temp/dec_block_map.rds")

### 2019-05-17 markdown/kmorris_thesis.Rmd
Fix standard errors in turnout regressions

@@
                      )
+library(sandwich)
+library(lmtest)
 library(scales)
@@
 Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no residents were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups.
+I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011].
 I match on a series of demographic and political indicators.
@@
 I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
+Table \ref{tab:match-reg} presents the results of these regression models, with robust standard errors clustered at the level of the match [@Abadie2019].
 ```{r tract-reg-chunk, echo = F, results = "asis", message = F}
-tract_reg <- readRDS("./temp/match_reg_tract.rds")
-tract_reg_black <- readRDS("./temp/match_reg_tract_nhb.rds")
-bg_reg <- readRDS("./temp/match_reg_block_group.rds")
-bg_reg_black <- readRDS("./temp/match_reg_block_group_nhb.rds")
+load("./temp/match_reg_block_group.rdata")
+
+load("./temp/match_reg_block_group_nhb.rdata")
+bg_reg <- reg_output
+bg_reg_ses <- reg_output_ses
+
+bg_reg_black <- reg_output_nhblack
+bg_reg_black_ses <- reg_output_nhblack_ses
+
+load("./temp/match_reg_tract.rdata")
+
+load("./temp/match_reg_tract_nhb.rdata")
+tract_reg <- reg_output
+tract_reg_ses <- reg_output_ses
+
+tract_reg_black <- reg_output_nhblack
+tract_reg_black_ses <- reg_output_nhblack_ses
 
@@
           header = F,
+          se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
+          notes = "Robust standard errors (clustered at match level) in parentheses.",
           type = "latex",
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) X Share Black", "Constant"),
+          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) X Share Non-Hispanic Black", "Constant"),
           dep.var.labels = c("Turnout Rate"),
@@
 When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average 0.8 percent lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black block groups.
-When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes insignificant (p = 0.24).
-The coefficient on the interaction between treatment and share Black indicates that neighborhoods that are largely Black saw turnout that was as much as *3 percent lower* than similar neighborhoods without lost voters.
-These variables are largely insignificant at the census tract level.
-This is not particularly surprising.
+In block groups with lost voters, turnout was on average `r -1*round(bg_reg[["coefficients"]][["treatTRUE"]]*100, digits = 1)` percentage points lower than in comparable block groups without lost voters.
+This decrease, however, appears to be entirely concentrated within Black neighborhoods.
+When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes insignificant.
+The coefficient on the interaction between treatment and share Black indicates that neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
+Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))*100, digits = 1)` percent, this effect is alarmingly high.
+For every 100 votes cast in a predominantly Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
+Although these variables are largely insignificant at the census tract level, this is not particularly surprising.
 If lost voters have the largest depressive effect on family members and close neighbors, it makes sense that the effect of losing a voter is locally concentrated.
@@
 The model above does not allow for different effects on turnout based on *how many* voters a neighborhood lost.
-In Table \ref{tab:trad-reg} below, I adopt a standard regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
+In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
+This regression uses the same covariates that were used in the matching procedure described above.
 
 ```{r trad-reg-chunk, echo = F, results = "asis", message = F}
-tract_reg <- readRDS("./temp/tract_model_reg_nosquare.rds")
-tract_reg_sq <- readRDS("./temp/tract_model_reg_square.rds")
-bg_reg <- readRDS("./temp/bg_model_reg_nosquare.rds")
-bg_reg_sq <- readRDS("./temp/bg_model_reg_square.rds")
+load("./temp/tract_model_reg_ols.rdata")
+load("./temp/bg_model_reg_ols.rdata")
 
-stargazer(bg_reg, bg_reg_sq, tract_reg, tract_reg_sq,
+stargazer(bg_model, bg_model2, tract_model, tract_model2,
           column.separate = c(2, 2),
@@
           type = "latex",
-          covariate.labels = c("Lost Voters", "Lost Voters X Share Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+          se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
+          covariate.labels = c("Lost Voters", "Lost Voters X Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
                                "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
@@
           style = "ajps",
-          omit.stat = c("f", "ser"))
+          omit.stat = c("f", "ser"),
+          notes = "Robust standard errors in parentheses.")
 ```
+The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.^[Although the coefficients from the standard regressions are slightly lower than the coefficients from the matching model, this is expected.
+The average "treated" neighborhood lost more than one voter, so it makes sense that the treatment effect is larger than the marginal estimate from the standard regression.] Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
+In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease between `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` and `r -1*round(tract_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points (depending on neighborhood definition).
 
-The block groups where these depressive effects are not, of course, randomly distributed througout the city.
+The block groups where these depressive effects are not randomly distributed througout the city.
 They are highly spatially concentrated: in Central Brooklyn, Eastern Queens, and Harlem.
@@
 
+## Discussion {-}
+
+During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` from New York City neighborhoods.
+The spatial concentration of these lost voters is striking: as Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} make clear, these voters were removed from neighborhoods with significantly lower incomes than the rest of the city.
+They were also removed from neighborhoods that were far less white than the average neighborhood.
+The systematic removal of voters from these neighborhoods is troubling.
+However, more than 850,000 votes were cast in the general election of 2017.
+Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters - concentrated though they are - is likely to have large implications on its own.
+
+Felony disenfranchisement, however, reaches beyond the individuals who are incarcerated.
+Previous literature has established that felony disenfranchisement likely has impacts on African-American turnout at the state level.
+This analysis demonstrates that these demobilizing effects intersect with geographical space
+
 \newpage

### 2019-05-17 code/00_initialize.R

@@ -1,3 +1,5 @@
+library(sandwich)
+library(lmtest)
 library(Matching)
 library(tidyverse)
 library(data.table)

### 2019-05-17 code/06_process_nyc_nhoods.R

@@ -20,14 +20,25 @@
 # nyc$cc_district <- over(pings, council_districts)$CounDist
 # saveRDS(nyc, "./temp/nyc.rds")
 
-nyc <- readRDS("./temp/nyc.rds") %>% 
-  filter(voter_status != "PURGED")
+nyc <- readRDS("./temp/nyc.rds")
 
 ## find lost voters
 nyc$lost_voter <- nyc$nys_id %in% readRDS("./temp/ids_of_lost_voters.rds")$nys_id
 
 lost_nyc <- sum(nyc$lost_voter)
 saveRDS(lost_nyc, "./temp/lost_count_nyc.rds")
+
+lost_bg <- nyc %>% 
+  group_by(GEOID = bg) %>% 
+  summarize(lost_voters = sum(lost_voter))
+
+lost_tract <- nyc %>% 
+  group_by(GEOID = substring(bg, 1, 11)) %>% 
+  summarize(lost_voters = sum(lost_voter))
+
+lost_zip <- nyc %>% 
+  group_by(GEOID = zip5) %>% 
+  summarize(lost_voters = sum(lost_voter))
 
 ## read in city council races to control for competitiveness
 races <- readRDS("./temp/council_competitiveness.rds")
@@ -60,12 +71,14 @@ arrests_tract <- arrests %>%
   summarize(arrests = n())
 
 ## share dem
+nyc <- nyc %>% 
+  filter(voter_status != "PURGED")
+
 share_dem_bg <- nyc %>% 
   group_by(GEOID = bg) %>% 
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
-            lost_voters = sum(lost_voter),
             share_winner = mean(share_winner))
 
 share_dem_tract <- nyc %>% 
@@ -73,7 +86,6 @@ share_dem_tract <- nyc %>%
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
-            lost_voters = sum(lost_voter),
             share_winner = mean(share_winner))
 
 share_dem_zip <- nyc %>% 
@@ -81,7 +93,6 @@ share_dem_zip <- nyc %>%
   summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
[…]

### 2019-05-17 code/08_basic_regressions.R

@@ -5,17 +5,26 @@ block_groups <- readRDS("./temp/block_group_pre_match.rds")
 block_groups$median_income <- block_groups$median_income / 1000
 
 block_groups$lost_voters_black <- block_groups$lost_voters * block_groups$nh_black
+block_groups$boro <- substring(block_groups$GEOID, 1, 5)
 
-bg_model  <- lm(to ~ lost_voters +                    median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
-saveRDS(bg_model, "./temp/bg_model_reg_nosquare.rds")
-bg_model2 <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
-saveRDS(bg_model2, "./temp/bg_model_reg_square.rds")
+bg_model  <- lm(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
+bg_model_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups))$coefficients)[, 2]
+save(bg_model, bg_model_ses, file = "./temp/bg_model_reg_ols.rdata")
 
+
+bg_model2  <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
+bg_model2_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups))$coefficients)[, 2]
+save(bg_model, bg_model_ses, bg_model2, bg_model2_ses, file = "./temp/bg_model_reg_ols.rdata")
+
+#######
 tracts <- readRDS("./temp/tract_pre_match.rds")
 tracts$median_income <- tracts$median_income / 1000
 tracts$lost_voters_black <- tracts$lost_voters * tracts$nh_black
+tracts$boro <- substring(tracts$GEOID, 1, 5)
 
[…]

### 2019-05-17 code/09_post_est_maps.R

@@ -31,7 +31,7 @@ rm(temp)
 
 bg_shp <- inner_join(bg_shp, block_groups, by = "GEOID")
 
-bg_shp$decrease <- -0.027 * bg_shp$lost_voters * bg_shp$nh_black
+bg_shp$decrease <- (bg_shp$lost_voters >= 1)*-1
 
 
 

### 2019-05-17 markdown/kmorris_thesis.Rmd
clean up

@@
           omit.stat = c("f", "ser"),
-          notes = "Robust standard errors in parentheses.")
+          notes = "Robust standard errors in parentheses.",
+          omit = "boro")
 ```
-The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.^[Although the coefficients from the standard regressions are slightly lower than the coefficients from the matching model, this is expected.
-The average "treated" neighborhood lost more than one voter, so it makes sense that the treatment effect is larger than the marginal estimate from the standard regression.] Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
+The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.
+Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease between `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` and `r -1*round(tract_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points (depending on neighborhood definition).

### 2019-05-17 code/00_initialize.R

@@ -1,5 +1,4 @@
-library(sandwich)
-library(lmtest)
+library(estimatr)
 library(Matching)
 library(tidyverse)
 library(data.table)

### 2019-05-17 code/06_process_nyc_nhoods.R

@@ -75,36 +75,45 @@ nyc <- nyc %>%
   filter(voter_status != "PURGED")
 
 share_dem_bg <- nyc %>% 
-  group_by(GEOID = bg) %>% 
-  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+  rename(GEOID = bg) %>% 
+  group_by(GEOID, cc_district) %>% 
+  mutate(count_d = n()) %>% 
+  group_by(GEOID) %>% 
+  mutate(biggest_d = max(count_d),
+         biggest_district = ifelse(count_d == biggest_d, cc_district, 0)) %>% 
+  summarize(district = max(biggest_district),
+            share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
             share_winner = mean(share_winner))
 
 share_dem_tract <- nyc %>% 
-  group_by(GEOID = substring(bg, 1, 11)) %>% 
-  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+  mutate(GEOID = substring(bg, 1, 11)) %>% 
+  group_by(GEOID, cc_district) %>% 
+  mutate(count_d = n()) %>% 
+  group_by(GEOID) %>% 
+  mutate(biggest_d = max(count_d),
+         biggest_district = ifelse(count_d == biggest_d, cc_district, 0)) %>% 
+  summarize(district = max(biggest_district),
+            share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
             vcount = n(),
             share_winner = mean(share_winner))
 
 share_dem_zip <- nyc %>% 
-  group_by(GEOID = zip5) %>% 
-  summarize(share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+  rename(GEOID = zip5) %>% 
+  group_by(GEOID, cc_district) %>% 
+  mutate(count_d = n()) %>% 
+  group_by(GEOID) %>% 
+  mutate(biggest_d = max(count_d),
+         biggest_district = ifelse(count_d == biggest_d, cc_district, 0)) %>% 
+  summarize(district = max(biggest_district),
+            share_dem = mean(political_party == "DEM" & !is.na(political_party)),
             v2017 = sum(v2017),
[…]

### 2019-05-17 markdown/kmorris_thesis.Rmd
fix standard errors

@@
 Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-These latitudes and longitudes were then converted to census block groups using the statistical program R and shapefiles available from the United States Census Bureau.
+These latitudes and longitudes were then converted to census block groups, census tracts, and city council districts using the statistical program R.
 
@@
 I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models, with robust standard errors clustered at the level of the match [@Abadie2019].
+Table \ref{tab:match-reg} presents the results of these regression models.
+Each neighborhood is weighted by its adult population, and robust standard errors are clustered at the level of the match [@Abadie2019].
 ```{r tract-reg-chunk, echo = F, results = "asis", message = F}
@@
 This regression uses the same covariates that were used in the matching procedure described above.
+Each neighborhood is weighted by its voting age population, and robust standard errors are clustered by city council district.
 
@@
           omit.stat = c("f", "ser"),
-          notes = "Robust standard errors in parentheses.",
+          notes = "Robust standard errors (clustered by city council district) in parentheses.",
           omit = "boro")
@@
 The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.
-Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
+Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease between `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` and `r -1*round(tract_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points (depending on neighborhood definition).

### 2019-05-17 code/07_turnout_match_reg.R

@@ -80,15 +80,13 @@ for(geo in c("block_group", "tract")){
   reg <- left_join(matches_v, units, by = c("control_GEOID" = "GEOID"))
   reg$boro <- substring(reg$GEOID, 1, 5)
   
-  reg_output <- lm(to ~ treat, data = reg, weights = weight)
-  reg_output_ses <- data.frame(summary(lm_robust(to ~ treat, data = reg, weights = weight, cluster = GEOID))$coefficients)[, 2]
+  reg_output <- lm(to ~ treat, data = reg, weights = weight * vap)
+  reg_output_ses <- data.frame(summary(lm_robust(to ~ treat, data = reg, weights = weight * vap, cluster = GEOID))$coefficients)[, 2]
   save(reg_output, reg_output_ses, file = paste0("./temp/match_reg_", geo, ".rdata"))
   
-  reg_output_nhblack <- lm(to ~ treat + treat * nh_black, data = reg, weights = weight)
-  reg_output_nhblack_ses <- data.frame(summary(lm_robust(to ~ treat + treat * nh_black, data = reg, weights = weight, cluster = GEOID))$coefficients)[, 2]
+  reg_output_nhblack <- lm(to ~ treat + treat * nh_black, data = reg, weights = weight * vap)
+  reg_output_nhblack_ses <- data.frame(summary(lm_robust(to ~ treat + treat * nh_black, data = reg, weights = weight * vap, cluster = GEOID))$coefficients)[, 2]
   save(reg_output_nhblack, reg_output_nhblack_ses, file = paste0("./temp/match_reg_", geo, "_nhb.rdata"))
-  
-}
   
   ###########
   load(paste0("./temp/mout_", geo, ".RData"))

### 2019-05-17 code/08_basic_regressions.R

@@ -7,13 +7,13 @@ block_groups$median_income <- block_groups$median_income / 1000
 block_groups$lost_voters_black <- block_groups$lost_voters * block_groups$nh_black
 block_groups$boro <- substring(block_groups$GEOID, 1, 5)
 
-bg_model  <- lm(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
-bg_model_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups))$coefficients)[, 2]
+bg_model  <- lm(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap)
+bg_model_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap, clusters = district))$coefficients)[, 2]
 save(bg_model, bg_model_ses, file = "./temp/bg_model_reg_ols.rdata")
 
 
-bg_model2  <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups)
-bg_model2_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups))$coefficients)[, 2]
+bg_model2  <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap)
+bg_model2_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap, clusters = district))$coefficients)[, 2]
[…]

### 2019-05-17 markdown/kmorris_thesis.Rmd
clean up maps

@@
 ---
-title: "Rethinking Felony Disenfranchisement: The Case of New York City and Miami"
+title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
 author: "Kevin Morris"
@@
 They have examined the effect of disenfranchisement on the political participation of individuals, including after they are no longer disenfranchised, but have generally neglected to examine the intersection of spatially concentrated policing and incarceration patterns with the political effects of disenfranchisement.
-This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the character of disenfranchisement in New York City and Miami-Dade County in Florida.
+This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the character of disenfranchisement in New York City.
 
@@
 Of course, despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with XX actively registered voters.
+In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
 Despite the lower share of residents who are disenfranchised through imprisonment, there is reason to believe the policy impacts more individuals than just those imprisoned.
@@
 The individual-level incarceration data referenced above does not include information about the neighborhood an incarcerated person lived in prior to their conviction.
-Although the State of New York does not make this information available at the individual level, the author has received prison-admission counts by zip-code for the years XX - XX.
+Although the State of New York does not make this information available at the individual level, the author has received prison-admission counts by zip-code for the years XX -- XX.
 
@@
 In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent election for city-wide office in New York - the mayoral election which took place on November 7^th^, 2017.
+To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
 Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
@@
 Each voter's record indicates whether the voter participated in the 2017 general election,^[The New York State voter file is not without errors.
-For instance, the voter file erroneously does not record the author as participating in the 2017 mayoral election.
+For instance, the voter file erroneously does not record the author as participating in the 2017 primary election.
 However, insofar as these errors are not geographically correlated, there is little reason to believe they impact the results of this analysis.
@@
 
-Matching methodologies, of course, only allow us to test the effect of being treated - here, losing a voter for the 2017 election.
+Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing a voter for the 2017 election.
 The model above does not allow for different effects on turnout based on *how many* voters a neighborhood lost.
@@
 Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease between `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` and `r -1*round(tract_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points (depending on neighborhood definition).
+In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
 
@@
 j <- readRDS("./temp/dec_block_map.rds")
-print(j)
+j
 ```
@@
 
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` from New York City neighborhoods.
+During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` from New York City neighborhoods.
 The spatial concentration of these lost voters is striking: as Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} make clear, these voters were removed from neighborhoods with significantly lower incomes than the rest of the city.
@@
 However, more than 850,000 votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters - concentrated though they are - is likely to have large implications on its own.
+Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters --- concentrated though they are --- is likely to have large implications on its own.
 
@@
 Previous literature has established that felony disenfranchisement likely has impacts on African-American turnout at the state level.
-This analysis demonstrates that these demobilizing effects intersect with geographical space
+This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
+These communities are already home to some of the poorest and most marginalized voters in the city; they are home to residents who have the least ability to influence policy through other means such as large campaign contributions.
+The fact that felony disenfranchisement policies appear to be disincentivizing participation on election day --- that these policies are further weakening the ability of these neighborhoods to participate in the democratic process --- is cause for alarm.
 

### 2019-05-17 code/05_voteshare_districts.R

@@ -28,6 +28,48 @@ results_ll <- results %>%
   mutate(total_votes = sum(vote_count), ## total votes cast in district
          share_winner = vote_count / total_votes) %>% ## candidate share
   filter(vote_count == max(vote_count)) %>% ## only keep whoever won
-  dplyr::select(district = District_Key, share_winner) ## just keep district name, share won by winner
+  dplyr::select(district = District_Key, share_winner, winner = Unit_Name) ## just keep district name, share won by winner
 
 saveRDS(results_ll, "./temp/council_competitiveness.rds")
+
+
+######
+eds <- results %>% 
+  group_by(AD, ED, Unit_Name, District_Key) %>% 
+  summarize(tot = sum(Tally, na.rm = T)) %>% 
+  group_by(AD, ED, District_Key) %>% 
+  mutate(share = tot / sum(tot))
+
+eds <- left_join(eds, results_ll, by = c("District_Key" = "district")) %>% 
+  filter(Unit_Name == winner,
+         share_winner < 1)
+
+hist(eds$share, breaks = 50)
+
+
+
+may <- fread("https://www.vote.nyc.ny.us/downloads/csv/election_results/2018/20180626Primary%20Election/01302000009Kings%20Democratic%20Representative%20in%20Congress%209th%20Congressional%20District%20EDLevel.csv")
+
+may <- may %>% 
+  mutate(AD = V12,
+         ED = V13,
+         Unit_Name = V21,
+         Tally = as.integer(gsub(",", "", V22)))
+
+colnames(may) <- gsub("[.]", "_", make.unique(make.names(colnames(may))))
+
+#may2 <- may[grepl("\\(", may$Unit_Name), ] ## drop records that aren't candidates (all candidates have party in parentheses)
+may2 <- may %>% filter(Unit_Name %in% c("Adem Bunkeddeko", "Yvette D. Clarke"))
+
+## get rid of party names in parentheses - don't care about party
+may2$Unit_Name <- gsub("\\s*\\([^\\)]+\\)", "", may2$Unit_Name)
+
+eds_m <- may2 %>% 
+  group_by(AD, ED, Unit_Name) %>% 
+  summarize(votes = sum(Tally)) %>% 
+  group_by(AD, ED) %>% 
+  mutate(share_bdb = votes / sum(votes),
+         tot = sum(votes)) %>% 
+  filter(Unit_Name == "Adem Bunkeddeko")
+
+hist(filter(eds_m, tot > 10)$share_bdb, breaks = 40)

### 2019-05-17 code/08_basic_regressions.R

@@ -9,8 +9,6 @@ block_groups$boro <- substring(block_groups$GEOID, 1, 5)
 
 bg_model  <- lm(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap)
 bg_model_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap, clusters = district))$coefficients)[, 2]
-save(bg_model, bg_model_ses, file = "./temp/bg_model_reg_ols.rdata")
-
 
 bg_model2  <- lm(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap)
 bg_model2_ses  <- data.frame(summary(lm_robust(to ~ lost_voters + lost_voters_black + median_income + latino + nh_black + nh_white + some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, data = block_groups, weights = vap, clusters = district))$coefficients)[, 2]

### 2019-05-17 code/09_post_est_maps.R

@@ -31,9 +31,15 @@ rm(temp)
 
 bg_shp <- inner_join(bg_shp, block_groups, by = "GEOID")
 
-bg_shp$decrease <- (bg_shp$lost_voters >= 1)*-1
+load("./temp/bg_model_reg_ols.rdata")
+coef <- bg_model2[["coefficients"]][["lost_voters_black"]]
 
+bg_shp$decrease <- bg_shp$lost_voters * bg_shp$nh_black * coef
 
+## bbs 
+bbs <- readOGR("./raw_data/shapefiles/Borough Boundaries", "geo_export_14dc9d2c-e65a-48c5-ac4d-1bcd90c6758d")
+bbs <- spTransform(bbs, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+bbs <- fortify(bbs)
 
 dec_map <- ggplot() +
   theme(axis.ticks = element_blank(),
@@ -45,11 +51,18 @@ dec_map <- ggplot() +
         legend.background = element_blank(),
         legend.key=element_blank(),
         legend.key.width = unit(1.5, "cm")) +
-  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "gray") +
+  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "#bfbfbf") +
   geom_polygon(data = bg_shp, aes(x = long, y = lat, group = group, fill = decrease)) +
+  geom_path(data = bbs, aes(x = long, y = lat, group = group), color = "black") +
   coord_map() +
   labs(x = NULL, y = NULL) + 
-  scale_fill_gradient(high = "gray", low = "red", labels = percent) +
+  scale_fill_gradient(high = "#bfbfbf", low = "red", labels = percent_format(accuracy = 1), limits = c(-0.05, 0), oob = squish) +
   guides(fill = guide_colorbar(title.hjust = .5, title.position = "top", title = "Decreased Turnout from Disenfranchisement"))
 
 saveRDS(dec_map, "./temp/dec_block_map.rds")
+
+
+
+
+
+

### 2019-05-17 code/misc/active_voters_0319.R

@@ -1,0 +1,7 @@
+#### count active voters
+
+
+count <- dbGetQuery(db, "select count(*) from nys_roll_0319 where voter_status == 'ACTIVE'") %>% 
+  pull()
+
+saveRDS(count, "./temp/active_voters_0319.rds")

### 2019-05-18 markdown/kmorris_thesis.Rmd
more writing

@@
 ---
+<!-- --- -->
+<!-- title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City" -->
+<!-- author: "Kevin Morris" -->
+<!-- date: "April 22, 2019" -->
+<!-- output: word_document -->
+<!-- bibliography: "../research/citations.bib" -->
+<!-- link-citations: yes -->
+<!-- always_allow_html: yes -->
+<!-- --- -->
 
@@
 
-These findings are not suprising given the large body of research establishing the social nature of voting.
-Outside the felony disenfranchisement conversation, an enormous body of research has demonstrated that individuals are far more likely to cast a ballot if their peers, family members or neighbors are doing so.
-[Build this out]
-
 Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
@@
 
-# Framing for Turnout Effects {-}
+# What This Paper Examines {-}
+
+This paper begins by examining the effect of felony disenfranchisement in the New York City Mayoral election of 2017.
+I am primarily interested in whether felony disenfranchisement has impacts on local political distribution.
+As discussed above, policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods.
+It is possible, however, that these effects are even broader than the direct focus on policing has acknowledged.
+We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
+However, to the extend that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that disenfranchisement policies are powerful enough to materially reduce the representatition of certain parts of the city.
+Although turnout is not a perfect proxy for political representation, there is some evidence that policians are more responsive to groups that turnout at higher rates.
+@Martin2013, for instance, documents the relationship between voter turnout and the responsiveness of Congressional representatives.
+"[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
+@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not." If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most disadvantaged neighborhoods to begin with.
+
+After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo the impact of felony disenfranchisement.
+In April of 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
+Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
+In theory, such a policy could increase turnout and registration for parolees.
+To test whether the policy was successful, I compare the turnout rates in the 2018 general election of three groups who finished parole in 2018: those who finished parole immediately before the policy change, those who finished parole immediately after the policy change, and those who finished parole immediately after the election.
+These three groups allow me to explore whether the re-extension of voting rights while on parole has an impact on turnout (by comparing the first two groups) and whether being under formal supervision while being allowed to vote results in higher turnout than being allowed to vote but being under supervision (by comparing the third group to the first two).
+
+The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
+Rather thank think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on African-American turnout at the state level), I seek to introduce the importance of physical space into the conversation.
+This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
+Of course, understanding the impact of felony disenfranchisement is not enough: we must also be testing whether our efforts to undo its legacy are effective.
+The second part of this project seeks to do just that.
+
+## Framing for Turnout Effects {-}
 
@@
 
-Of course, despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
+Despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
 In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
-Despite the lower share of residents who are disenfranchised through imprisonment, there is reason to believe the policy impacts more individuals than just those imprisoned.
+Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
 As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher.
-Understanding the distribution of this effect is important for understanding its political implications.
-Are voters who live in the same neighborhood as the disenfranchised less likely to cast a ballot because their neighbors have been removed from the community?
-Or does felony disenfranchisement operate cast its chilling effect on turnout more broadly, crossing even into neighborhoods where voters have not been removed from the community?
-
-This paper begins by examining whether felony disenfranchisement is in fact associated with lower turnout rates in a given neighborhood.^[As discussed below, "neighborhood" is difficult to define.
-I explore turnout differentials at both the census block group and census tract level.] As discussed above, this depressive effect on turnout is expected to operate through two distinct mechanisms: the actual incarceration and disenfranchisment of convicted individuals, and the social effects on eligible voters identified in previous research.
-These turnout effects are estimated in various ways: by comparing neighborhoods where voters with a history of participating have been incarcerated with similar neighborhoods where voters have not been incarcerated; by using regression analysis to investigate whether neighborhood arrests are correlated with turnout rates; and by using regression analysis to investigate the relationship between prison admissions and turnout at the zip-code level.
-
-# What This Paper Examines {-}
-This paper begins by examining the effect of felony disenfranchisement in the New York City Mayoral election of 2017.
-This paper is primarily interested in whether felony disenfranchisement has impacts on local political distribution.
-If, for instance, voters in particular marked neighborhoods turnout at lower rates because of disenfranchisement policies, this could have ramifications for city-wide policy.
-@Martin2013, for instance, documents the relationship between voter turnout and the responsiveness of Congressional representatives.
-"[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
-@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not."
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
+The existence of such *de facto* disenfranchisement is bolstered by a large body of empirical work on the social nature of voting.
+Many scholars have established that being in a community where voting is common makes individuals more likely to vote.
+Children whose parents vote, for instance, are more likely to vote as adults [XX]; XX add more examples here.
+Many of these relationships are spatially bounded: family members are likely to live in the same home, or on the same block.
+Literature from urban sociology has established that social networks are largely spatially bounded.
+To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned, these policies are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
 
@@
 An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote cohesively on issues such as school and park funding, their turnout rates may influence city policy.
+Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
 
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo the impact of felony disenfranchisement.
-In April of 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those in prison are disenfranchised] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for these individuals.
-To test whether the policy was successful, I compare the turnout rates in the 2018 general election of three groups who finished parole in 2018: those who finished parole immediately before the policy change, those who finished parole immediately after the policy change, and those who finished parole immediately after the election.
-These three groups allow me to explore whether the re-extension of voting rights while on parole has an impact on turnout (by comparing the first two groups) and whether being under formal supervision while being allowed to vote results in higher turnout than being allowed to vote but being under supervision (by comparing the third group to the first two).
+## Previous Attempts to Re-Incorporate Formerly Disenfranchised Residents
 
@@
 However, the Department of Corrections and Community Supervision website's Parolee Lookup includes a flag indicating whether someone's voting rights haave been restored.
-By using identification numbers from the data obtained from the state to lookup parolees on the website, I was able to construct a list of the individuals who were on parole and had their rights restored.
+By using identification numbers from the data obtained from the state to lookup parolees on the website, I was able to construct a list of the individuals who were on parole and had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
 

### 2019-05-18 code/04_lost_voters_maps.R

@@ -59,7 +59,8 @@ city_map <- ggplot() +
         legend.key=element_blank()) +
   geom_polygon(data = dists, aes(x = long, y = lat, group = group), fill = "#bfbfbf") +
   geom_path(data = dists, aes(x = long, y = lat, group = group), color = "black") +
-  geom_point(data = filter(history, !is.na(district), voted == "Cast Ballot in Past 10 Years"), aes(x = longitude, y = latitude), shape = 21, color = "black", fill = "red") +
+  geom_point(data = filter(history, !is.na(district), voted == "Cast Ballot in Past 10 Years"),
+             aes(x = longitude, y = latitude), shape = 21, color = "black", fill = "red", alpha = 0.5) +
   coord_map() +
   labs(x = NULL, y = NULL)
 

### 2019-05-20 markdown/kmorris_thesis.Rmd
Show that ballot undercount is not associated with lost voters

@@
 I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Following @Meredith2013, I report estimates of the potential impact of these false positives in Appendix XX.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
@@
 Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election,^[The New York State voter file is not without errors.
-For instance, the voter file erroneously does not record the author as participating in the 2017 primary election.
-However, insofar as these errors are not geographically correlated, there is little reason to believe they impact the results of this analysis.
-Unfortunately, testing the distribution of these errors is not possible.] which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided the voting age population.
+Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.^[The New York registered voter file does not align exactly with results reported by the city.
+The voter file indicates that just XX voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that XX votes were cast.
+In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level, which means these reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided the voting age population.
 
@@
 # References {-}
+<div id="refs"></div>
+
+\newpage
+# Appendix A {-}
+
+As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
+According to the official results, XX ballots were cast in the 2017 mayoral election.
+The registered voter file reports that just XX voters cast a ballot.
+This is not necessarily evidence of poor management by the Board of Elections.
+Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).
+Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
+If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
+
+To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level.
+The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level.
+Because the census does not report statistics at the precinct level, I use block group information instead.
+Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group.
+The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.
+
+```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
+load("./temp/compare_ballots_off_results.RData")
+
+stargazer(ed_reg,
+          header = F,
+          type = "latex",
+          se = list(ed_comp_ses),
+          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
+          dep.var.labels = c("Undercount Rate"),
+          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
+          table.placement = "H",
+          style = "ajps",
+          omit.stat = c("f", "ser"),
+          notes = "Robust standard errors (clustered by assembly district) in parentheses.",
+          omit = "boro")
+```
+
+As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate.
+Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.

### 2019-05-20 code/XX_compare_results_ballots.R

@@ -1,0 +1,93 @@
+## try to understand over / undercount by ed
+
+election_districts <- dbGetQuery(db, "select nys_id, voter_status, election_district, assembly_district from nys_roll_0418")
+
+
+election_districts <- election_districts %>%
+  mutate(a = as.integer(voter_status == "ACTIVE"),
+         b = as.integer(voter_status == "INACTIVE"),
+         c = as.integer(voter_status == "PREREG"))
+
+election_districts <- setorder(election_districts, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+election_districts <- election_districts[!duplicated(election_districts$nys_id),]
+election_districts <- select(election_districts, -a, -b, -c)
+
+
+nyc <- readRDS("./temp/nyc.rds")
+
+
+nyc <- inner_join(nyc, election_districts)
+# read election names
+elects <- fread("./raw_data/misc/elects.csv")
+
+history <- cSplit(dplyr::select(nyc, nys_id, history), "history", sep = ";", direction = "long", type.convert = F)
+
+history <- left_join(history, elects, by = "history")
+history <- filter(history, year == 2017, election_type == "general")
+nyc$v2017 <- nyc$nys_id %in% history$nys_id
+
+## read in city council races to control for competitiveness
+races <- readRDS("./temp/council_competitiveness.rds")
+nyc <- left_join(nyc, races, by = c("cc_district" = "district"))
+
+### get census data
+load("./temp/census_data_nyc_bgs_tracts.RData")
+bgs <- geos[[2]]
+nyc <- left_join(nyc, bgs, by = c("bg" = "GEOID"))
+
+####
+
+ed_level <- nyc %>% 
+  filter(voter_status != "PURGED") %>% 
+  group_by(election_district, assembly_district) %>% 
+  summarize(median_income = mean(median_income, na.rm = T),
+            latino = mean(latino, na.rm = T),
+            nh_black = mean(nh_black, na.rm = T),
+            nh_white = mean(nh_white, na.rm = T),
+            some_college = mean(some_college, na.rm = T),
+            median_age = mean(median_age, na.rm = T),
+            share_dem = mean(political_party == "DEM" & !is.na(political_party)),
+            share_non_citizen = mean(share_non_citizen, na.rm = T),
+            share_winner = mean(share_winner),
+            vcount = n(),
[…]

### 2019-05-20 markdown/kmorris_thesis.Rmd
More writing

@@
 ```{r setup, echo = F, message = F, warning = F}
-knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf"
-                     )
+knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf")
 library(sandwich)
@@
 library(stargazer)
+loadfonts()
 
 ```
-"There is no Negro problem. There is no Southern problem. There is no Northern problem. There is only an American problem"
-
--*Lyndon Johnson, March 15, 1965*
-
 # Introduction {-}
@@
 
-## Previous Attempts to Re-Incorporate Formerly Disenfranchised Residents
+## Previous Attempts to Re-Incorporate Formerly Disenfranchised Residents {-}
 
@@
 Individuals convicted of misdemeanors are not disenfranchised in New York State.
+These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
 
@@
 I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these false positives in Appendix XX.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Following @Meredith2013, I report estimates of the potential impact of these errors in [Appendix B]] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
@@
 The voter file indicates that just XX voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that XX votes were cast.
-In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level, which means these reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided the voting age population.
+In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
+These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided the voting age population.
 
@@
 
-```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Depressive Effect of Felony Disenfranchisement", out.extra = '', fig.pos = "H", fig.align = "center"}
+```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
 j <- readRDS("./temp/dec_block_map.rds")
-j
+print(j)
 ```
@@
 This is not necessarily evidence of poor management by the Board of Elections.
-Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).
-Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
+Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).^[This is not to say that there is *not* evidence of mismanagement by the BOE.
+For instance, the author is incorrectly marked as not participating in the 2017 mayoral primary.] Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
 If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
@@
 Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
+
+\newpage
+# Appendix B {-}
+
+Matching across administrative data sets is not a perfect science.
+Despite having high-quality low-level data from both New York State Department of Corrections and Community Supervision (DOCCS) and the New York State Board of Elections, I cannot be sure that I am correctly identifying formerly incarcerated, registered voters.
+The results of the analyses in Table \ref{tab:match-reg} give some reason to believe that there are few false negatives.
+If there were many false negatives (lost voters who are not identified in this analysis), "treated" neighborhoods would be incorrectly categorized as "untreated" neighborhoods.
+If this were a widespread problem, we would expect to see little difference between treated and control neighborhoods.
+To the extent that lost voters go unidentified by this analysis, it is likely that the results presented in Tables \ref{tab:match-reg} and \ref{tab:trad-reg} are conservative by including treated neighborhoods in the control group.
+Directly testing for false negatives, however, is difficult.
+If a voter changes her name before being incarcerated, for instance, our analysis will fail to capture her.
+
+Testing for the presence of false positives is slightly easier.
+One way to do so is by constructing false records [@Meredith2013].
+By slightly changing the birthdates in one set of administrative data, I can estimate how frequently DOCCS records inaccurately match with voter records.
+To do so, I shift the dates of birth in the DOCCS data by 35 days.
+I then re-merge the DOCCS records (with "false" birth dates) against the registered voter file.
+
+Table \ref{tab:change-dobs} shows the results of these permutations.
+The first row shows the matches with the raw data.
+The second row shows the results of adding 35 days to the date of birth in each DOCCS record, and the final row shows the result of subtracting 35 days from these records.
+
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
+true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
+plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+Table \ref{tab:change-dobs} indicates that there are likely some false positives in our matches.
+It is not surprising that in a large state like New York, full names and dates of birth fail to uniquely identify individuals.
+However, the rate of false positives is quite low: this analysis indicates that there is a `r percent(as.integer(max(less_35, plus_35)) / as.integer(gsub(",", "", true)), accuracy = 0.01)`.
+It is unlikely that such a small incidence of false positives meaningfully impacts this analysis.

### 2019-05-20 code/09_post_est_maps.R

@@ -57,12 +57,6 @@ dec_map <- ggplot() +
   coord_map() +
   labs(x = NULL, y = NULL) + 
   scale_fill_gradient(high = "#bfbfbf", low = "red", labels = percent_format(accuracy = 1), limits = c(-0.05, 0), oob = squish) +
-  guides(fill = guide_colorbar(title.hjust = .5, title.position = "top", title = "Decreased Turnout from Disenfranchisement"))
+  guides(fill = guide_colorbar(title.hjust = .5, title = NULL))
 
 saveRDS(dec_map, "./temp/dec_block_map.rds")
-
-
-
-
-
-

### 2019-05-21 markdown/kmorris_thesis.Rmd
clean up written part

@@
 author: "Kevin Morris"
-date: "April 22, 2019"
+date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
@@
     fig_caption: yes
-    latex_engine: xelatex
 bibliography: "../research/citations.bib"
@@
 <!-- author: "Kevin Morris" -->
-<!-- date: "April 22, 2019" -->
+<!-- date: "`r format(Sys.time(), '%B %d, %Y')`" -->
 <!-- output: word_document -->
@@
 This is true in both the academic world (where researchers have dug into the political ramifications in such races as the 2000 presidential election) and in the political world as states such as Florida and Louisiana have gradually moved to dismantle their systems of disenfranchisement.
-Throughout this process, however, many researchers have focused on the *individual* effects of felony disenfranchisement.
-They have examined the effect of disenfranchisement on the political participation of individuals, including after they are no longer disenfranchised, but have generally neglected to examine the intersection of spatially concentrated policing and incarceration patterns with the political effects of disenfranchisement.
-This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the character of disenfranchisement in New York City.
+Many researchers, however, have focused on the effects of felony disenfranchisement on individuals, and on racial minorities as a group.
+They have examined the effect of disenfranchisement on the political participation of individuals, including after they are no longer disenfranchised, and the spillover effects disenfranchisement has on eligible voters of color.
+Researchers have generally neglected to examine the intersection of spatially concentrated policing and incarceration patterns with the political effects of disenfranchisement.
+This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the relationship between disenfranchisement and neighborhood turnout in New York City.
 
@@
 Although advocates are correct to push for an end to disenfranchisement, a change to the political system may not be sufficient to reincorporate the voices of formerly incarcerated individuals into the democratic process.
-In addition to laying out the spatial implications of felony disenfranchisement in two major American cities, this paper also explores the impact of two different policies intended to dismantle the disenfranchisement program.
-Specifically, I investigate the effect of Executive Order 181 (which ended the disenfranchisement of parolees in New York State) and  Amendment 4 (which restored the voting rights of Floridians who had completed all terms of their sentences).
-The effects of these policies make clear that, while the end of felony disenfranchisement is necessary, active steps must be taken to reincorporate the formerly disenfranchised.
+In addition to laying out the spatial implications of felony disenfranchisement in New York City, this paper also explores the impact of one policy intended to weaken the disenfranchisement program.
+Specifically, I investigate the effect of New York Governor Andrew Cuomo's Executive Order 181 which ended the disenfranchisement of parolees in New York State.
+The effect of these policies make clear that, while the end of felony disenfranchisement is necessary, it is not enough: Active steps must be taken to reincorporate the formerly disenfranchised.
 
@@
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to encompass any felony offense at all.
+This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchize citizens for any felony offense at all.
 The Supreme Court, in cases such as *Richardson v.
@@
 The historical record is undeniable.
-Previous research established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
+Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
 In Florida felony disenfranchisement was added to the state constitution in 1868.
@@
 Indeed, as the United States has vastly increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of African Americans were disenfranchised in 9 states.
+As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
 In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
 Although Black adults made up just `r percent(29932674 / 247219588, accuracy = 0.1)` of the voting age population in 2016, they accounted for `r percent(2228118 / 6106327, accuracy = 0.1)` of the disenfranchised population [@sentencing_2016].
+
+Recent data Florida make the victims of felony disenfranchisement laws especially clear.
+Prior to January, 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
+In November, 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
+In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate [@Morris2019].
 
@@
 In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that, if this group had been allowed to vote, they would have supported Al Gore.
-Their enfranchisement, Uggen and Manza posit, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of *non*-disenfranchised individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has undermined some of their key assumptions (namely, that formerly incarcerated individuals turnout and cast ballots similarly to those who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after the Uggen and Manza published their paper, scholars sought to investigate the effect of felony disenfranchisement on African American turnout [@Miles2004], young adult turnout [@Hjalmarsson2010], and other questions.
+They estimate that, if this group had been allowed to vote, they would have supported Al Gore by a wide margin.
+Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
+They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
+Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as those who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
+In the years after the Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
 Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
@@
 Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
-She uses release data from states' Departments of Corrections and their Registered Voter Files to estimate the number of formerly incarcerated individuals who went on to register to vote.
-Using the Registered Voter Files, she is also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and the turnout rate among these individuals.
+She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
+Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
 Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [@Meredith2013; @Meredith2015].
@@
 A number of papers have also explored the impact that felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson [-@King2016], for instance, leverage state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among African Americans.
+King and Erickson [-@King2016], for instance, leverage state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
 They use data from the 2004 Current Population Survey Voting and Registration Supplement to determine individual-level turnout.
-They include estimates of the share of African Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout." They conclude that disenfranchisement has spillover effects for African American voters: where more African Americans are disenfranchised, elibible African Americans are less likely to cast a ballot.
+They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout." They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, elibible Black voters are less likely to cast a ballot.
 These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
@@
 Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the types of neighborhoods in which these effects are concentrated, and the impact these laws might have at the local level.
+Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the impact these laws might have at the local level.
 @Burch2013 is an exception to this.
 In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
-She determines that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City and by introducing a new way of measuring the impact.
-
-# Spatial Effects of Concentrated Policing and Incarceration {-}
-
-XX
+Using a standard ordinary least squares regression determined that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definiton of "lost voters."
 
@@
 
-This paper begins by examining the effect of felony disenfranchisement in the New York City Mayoral election of 2017.
-I am primarily interested in whether felony disenfranchisement has impacts on local political distribution.
-As discussed above, policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods.
+This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017.
+Policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods [XX].
 It is possible, however, that these effects are even broader than the direct focus on policing has acknowledged.
 We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-However, to the extend that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that disenfranchisement policies are powerful enough to materially reduce the representatition of certain parts of the city.
-Although turnout is not a perfect proxy for political representation, there is some evidence that policians are more responsive to groups that turnout at higher rates.
+To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representatition of certain parts of the city.
+Although turnout is not a perfect proxy for political representation, there is some evidence that policians are more responsive to groups that turn out at higher rates.
 @Martin2013, for instance, documents the relationship between voter turnout and the responsiveness of Congressional representatives.
 "[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
-@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not." If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most disadvantaged neighborhoods to begin with.
+@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not."
 
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo the impact of felony disenfranchisement.
+Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
+Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
+As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
+An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
+Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
+If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most neglected neighborhoods to begin with.
+
+After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
 In April of 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
@@
 The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather thank think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on African-American turnout at the state level), I seek to introduce the importance of physical space into the conversation.
+Rather thank think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
 This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also be testing whether our efforts to undo its legacy are effective.
+Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
 The second part of this project seeks to do just that.
@@
 
-Although understanding whether felony disenfranchisement has impacts on turnout among different social groups (namely, racial minorities) is of great importance, understanding the explicitly spatial dimension of the problem is also an important task.
-Many political decisions bear on the spatial distribution of resources (XX); funding for parks or schools, for instance, can impact all members of a neighborhood, regardless of their race or their income.
+Although understanding whether felony disenfranchisement impacts turnout among different social groups is of great importance, understanding the explicitly spatial dimension of the problem is also an important task.
+Many political decisions bear on the spatial distribution of resources (XX); funding for parks or schools, for instance, can impact all members of a neighborhood, regardless of their race, income, or criminal justice involvement.
 As such, even a resident who is not directly affected by felony disenfranchisement policies may lose political power thanks to these policies; if many of my neighbors cannot vote, I am joined by fewer voters in advocating for neighborhood-specific investment and services.
-This is, of course, tied up closely with race; New York City remains highly racially segregated, and therefore disentangling neighborhood effects from race effects becomes difficult.
-Nevertheless, these effects cross race lines in integrated neighborhoods.
+This is, of course, tied up closely with race; New York City remains highly racially segregated.
+Thanks to spatially disparate policing patterns, any depressive effect is therefore likely to be concentrated among Black voters.
 
@@
 Children whose parents vote, for instance, are more likely to vote as adults [XX]; XX add more examples here.
-Many of these relationships are spatially bounded: family members are likely to live in the same home, or on the same block.
-Literature from urban sociology has established that social networks are largely spatially bounded.
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned, these policies are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
+Literature from urban sociology has established that social networks are largely spatially bounded [XX].
+To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
 
 ## Previous Attempts to Re-Incorporate Formerly Disenfranchised Residents {-}
+
+* Super low turnout
+* Immediate effects on future participation - Ariel's paper [@WHITE2019]
+* Changes in Notification Laws (M&M)
 
@@
 
-The City of New York also makes the location of all NYPD arrests since 2006 available publicly.
-This data includes the latitude and longitude of the arrest, the date of the offense, and the class of the alleged offense.
-This analysis uses only arrests for felony charges.
-
 The third dataset used to estimate the extent to which a neighborhood is directly impacted by felony disenfranchisement policies in New York State comes from another freedom of information request.
@@
 The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the Department of Corrections and Community Supervision website's Parolee Lookup includes a flag indicating whether someone's voting rights haave been restored.
-By using identification numbers from the data obtained from the state to lookup parolees on the website, I was able to construct a list of the individuals who were on parole and had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
+However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
+By using identification numbers from the parolee data obtained from the state, I constructed a list of the individuals who were on parole and had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
 
 ## Voter File Data {-}
-Most states in the United States are required to maintain files with information on all registered voters in the state.
+Most states in the United States are required to maintain files with information on all registered voters.
 In New York, this information is publicly available from the Board of Elections.
@@
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-Because the voter file includes information on voters who have been purged because of a felony conviction, we are able to reconstruct the vote history of individuals who are no longer registered because of their conviction.
+The first section of this paper uses a snapshot of the registered voter file from April 30^th^, 2018.
+The second section of this paper uses a snapshot from March 3^rd^, 2019.
 
@@
 Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-These latitudes and longitudes were then converted to census block groups, census tracts, and city council districts using the statistical program R.
+I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
 
@@
 In this analysis, I offer a slightly different defintion of "lost voter" than much of the literature.
-Many recent papers have attempted to identify relationships between the number of *potentially* lost voters and turnout.
-Here, I explore whether the disenfranchisement of voters with a history of participating in elections is related to neighborhood turnout.
+Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
+Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
 In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
+To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
 Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
 
-These individuals are identified by matching Department of Corrections and Community Supervisions data to the voter file using the methodology developed by Burch [-@Burch2010; -@Burch2011].
+These individuals are identified by matching NYSDOCCS data to the voter file.
 I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I report estimates of the potential impact of these errors in [Appendix B]] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Following @Meredith2013, I discuss the potential impact of these errors in [Appendix A].] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
@@
 In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals were not allowed to cast a ballot for mayor.
+In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
 
@@
 To test whether neighborhoods with lost voters had systematically lower turnout rates than similar neighborhoods without such disqualifications, I begin by using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no residents were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011].
+Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.^[It is worth noting that this definition of lost voter is highly conservative.
+For instance, some individuals who would have voted may have gone to prison after turning 18 but before their first general election.
+These "untreated" neighborhoods are likely home to individuals who would have voted if it were not for disenfranchisement policies.] I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011].
 I match on a series of demographic and political indicators.
@@
 Registration rate is calculated by dividing the number of registered voters (from the voter file) by the voting age population.
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.
-All else equal, a more competitive race is likely to result in higher turnout.^[Where neighborhoods cross council district lines, this measure is the weighted average of the competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
+I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
 Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
@@
 ```
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like our treatment group.
+At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
 These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are substantially less white, have far lower median incomes, and a larger share of voters are registered as Democrats.
+Neighborhoods with lost voters are far less white, have much lower median incomes, and a larger share of voters are registered as Democrats.
 
@@
 The voter file indicates that just XX voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that XX votes were cast.
-In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
-These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided the voting age population.
+In [Appendix B], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
+These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population as reported by the Census Bureau.
 
@@
 
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing a voter for the 2017 election.
+Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
 The model above does not allow for different effects on turnout based on *how many* voters a neighborhood lost.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates that were used in the matching procedure described above.
-Each neighborhood is weighted by its voting age population, and robust standard errors are clustered by city council district.
+This regression uses the same covariates used in the matching procedure described above.
+Each neighborhood is weighted by its voting age population, and robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purporses.]
 
@@
 The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.
-Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 make clear that this effect is concentrated in Black neighborhoods.
+Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
@@
 Felony disenfranchisement, however, reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement likely has impacts on African-American turnout at the state level.
+Previous literature has established that felony disenfranchisement likely has impacts on Black turnout at the state level.
 This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
@@
 # Appendix A {-}
-
-As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
-According to the official results, XX ballots were cast in the 2017 mayoral election.
-The registered voter file reports that just XX voters cast a ballot.
-This is not necessarily evidence of poor management by the Board of Elections.
-Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).^[This is not to say that there is *not* evidence of mismanagement by the BOE.
-For instance, the author is incorrectly marked as not participating in the 2017 mayoral primary.] Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
-If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
-
-To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level.
-The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level.
-Because the census does not report statistics at the precinct level, I use block group information instead.
-Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group.
-The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.
-
-```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/compare_ballots_off_results.RData")
-
-stargazer(ed_reg,
-          header = F,
-          type = "latex",
-          se = list(ed_comp_ses),
-          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Undercount Rate"),
-          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
-          table.placement = "H",
-          style = "ajps",
-          omit.stat = c("f", "ser"),
-          notes = "Robust standard errors (clustered by assembly district) in parentheses.",
-          omit = "boro")
-```
-
-As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate.
-Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
-
-\newpage
-# Appendix B {-}
 
@@
 It is unlikely that such a small incidence of false positives meaningfully impacts this analysis.
+
+\newpage
+# Appendix B {-}
+
+As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
+According to the official results, XX ballots were cast in the 2017 mayoral election.
+The registered voter file reports that just XX voters cast a ballot.
+This is not necessarily evidence of poor management by the Board of Elections.
+Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).^[This is not to say that there is *not* evidence of mismanagement by the BOE.
+For instance, the author is incorrectly marked as not participating in the 2017 mayoral primary.] Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
+If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
+
+To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level.
+The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level.
+Precincts are assigned the demographics of the block group in which they are situated.
+Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group.
+The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.
+Robust standard errors are clustered at the assembly district level.
+
+```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
+load("./temp/compare_ballots_off_results.RData")
+
+stargazer(ed_reg,
+          header = F,
+          type = "latex",
+          se = list(ed_comp_ses),
+          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
+          dep.var.labels = c("Undercount Rate"),
+          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
+          table.placement = "H",
+          style = "ajps",
+          omit.stat = c("f", "ser"),
+          notes = "Robust standard errors (clustered by assembly district) in parentheses.",
+          omit = "boro")
+```
+
+As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate.
+Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.

### 2019-05-21 code/04_lost_voters_maps.R

@@ -66,26 +66,17 @@ city_map <- ggplot() +
 
 saveRDS(city_map, "./output/city_map.RDS")
 
-#### tract map
 
-tract_shp <- readOGR("./raw_data/shapefiles/nyct2010_19a", "nyct2010")
-tract_shp <- spTransform(tract_shp, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
-tract_shp@data$id <- rownames(tract_shp@data)
-temp <- fortify(tract_shp)
-tract_shp <- inner_join(temp, tract_shp@data, by = "id")
-rm(temp)
+### bg map
 
-count_by_tract <- count_by_bg %>% 
-  group_by(tract = substring(bg, 1, 11)) %>% 
-  summarize(lost_voters = sum(lost_voters, na.rm = T))
+bgs <- readOGR("./raw_data/shapefiles/tl_2018_36_bg", "tl_2018_36_bg")
 
-codes <- data.frame("BoroCode" = as.character(c(1:5)), "cc" = c("061", "005", "047", "081", "085"))
+bgs@data$id <- rownames(bgs@data)
+t <- fortify(bgs)
 
-tract_shp <- left_join(tract_shp, codes, by = "BoroCode")
+bgs <- inner_join(bgs@data, t)
 
-tract_shp$GEOID <- with(tract_shp, paste0("36", cc, CT2010))
-tract_shp <- left_join(tract_shp, count_by_tract, by = c("GEOID" = "tract"))
-tract_shp$lost_voters <- ifelse(is.na(tract_shp$lost_voters), 0, tract_shp$lost_voters)
+bgs <- filter(bgs, GEOID %in% filter(history, !is.na(district))$bg)
 
 ggplot() +
   theme(axis.ticks = element_blank(),
@@ -96,52 +87,4 @@ ggplot() +
         plot.title = element_text(hjust = 0.5),
         legend.background = element_blank(),
         legend.key=element_blank()) +
-  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group, fill = lost_voters), color = "black") +
-  coord_map() +
-  labs(x = NULL, y = NULL) + 
-  scale_fill_gradient2() + ggtitle("Lost Voters by Tract")
-
-########
-
-
-############### old maps below
-# 
-# ggplot() +
-#   theme(axis.ticks = element_blank(),
-#         axis.text = element_blank(),
-#         panel.background = element_blank(),
-#         panel.border = element_blank(),
[…]

### 2019-05-21 code/09_post_est_maps.R

@@ -34,7 +34,7 @@ bg_shp <- inner_join(bg_shp, block_groups, by = "GEOID")
 load("./temp/bg_model_reg_ols.rdata")
 coef <- bg_model2[["coefficients"]][["lost_voters_black"]]
 
-bg_shp$decrease <- bg_shp$lost_voters * bg_shp$nh_black * coef
+bg_shp$decrease <- bg_shp$lost_voters
 
 ## bbs 
 bbs <- readOGR("./raw_data/shapefiles/Borough Boundaries", "geo_export_14dc9d2c-e65a-48c5-ac4d-1bcd90c6758d")

### 2019-05-21 code/XX_compare_results_ballots.R

@@ -10,7 +10,7 @@ election_districts <- election_districts %>%
 
 election_districts <- setorder(election_districts, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
 election_districts <- election_districts[!duplicated(election_districts$nys_id),]
-election_districts <- select(election_districts, -a, -b, -c)
+election_districts <- dplyr::select(election_districts, -a, -b, -c)
 
 
 nyc <- readRDS("./temp/nyc.rds")

### 2019-05-29 markdown/kmorris_thesis.Rmd
Time trends

@@
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchize citizens for any felony offense at all.
+This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense at all.
 The Supreme Court, in cases such as *Richardson v.
@@
 
+# Executive Order 181 and Turnout in 2018 {-}
+
+Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
+For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
+New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
+
+On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
+The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].^[I've also talked with folks on probation whose probation officers told them they were disenfranchised.
+Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the XX voters living in the community and to clarify the rules about who is eligible to vote.
+
+Of course, re-extending the right to vote is only the first step.
+As previous literature has established, interactions with the criminal justice system leaves residents far less likely to vote in the future [@WHITE2019; @XX].
+As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are vanishingly low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
+To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change risks leaving much of the system of effective disenfranchisement undisturbed.
+
+There is reason to believe that the policy change may increase the political participation of formerly disenfranchised individuals.
+Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
+Under the Executive Order, however, parole officers were required to explicitly inform their supervisees of their newly restored voting rights.
+It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly eligible voters.
+
+## Identifying Parolees Whose Rights Were Restored {-}
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
+By using the DIN number provided from the parolee public records request, I was able to identify individuals who did and did not have their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
+For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
+
+Although Governor Cuomo signed the Executive Order in April of 2018, an examination of the individuals whose rights were ultimately restored indicates that the program may not have gone into full effect until later in May.
+As Figure \ref{fig:restoration-chart} demonstrates, individuals who finished their parole supervision in early May did not have their voting rights formally restored until they finished their term of supervision.
+It is not until May 21^st^ that the majority of individuals being discharged from parole had their voting rights restored prior to their discharge.
+However, even after the policy change went into effect, not all parolees had their voting rights restored prior to being discharged from parole.
+These are likely individuals who are not citizens, and therefore ineligible to register to vote.
+```{r restoration-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/restoration_plot.rds")
+print(j)
+```
+Figure \ref{fig:restoration-chart} complicates our ability to cleanly identify the impact of the policy change.
+If no parolees discharged prior to May 21^st^ had had their rights restored, but all parolees discharged on or after May 21^st^ had, identifying those impacted by the policy would be straightforward.
+That is not the case: some select individuals seem to have had their rights restored before the policy change was fully implemented.
+Similarly, we would ideally exclude all individuals who did not have their rights restored after the policy change as well as the individuals who were discharged before the change who would not have had their rights restored if they had been discharged from parole later.
+However, identifying individuals whose rights would not have been restored is not possible.
+
+In the analyses that follow, I consider all parolees discharged on or after May 21^st^, 2018, beneficiaries of the policy change, whether their rights were restored or not.
+I assume that the share of individuals discharged after May 21^st^ whose rights were not restored is comparable to the share of individuals discharged prior to May 21^st^ whose rights would not have been restored if they had been discharged later.
+I also remove from the analyses that follow the 16 individuals discharged from parole prior to May 21^st^, 2018, whose rights were restored prior to their discharge.
+
+## Trends in Turnout {-}
+Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
+It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier.
+If individuals who were discharged from parole prior to May 21^st^, 2016, were less likely to vote than those discharged between May 21^st^ and Election Day, disentangling the impact of the policy change in 2018 from the expected uptick would be more complicated.
+However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier in the year.
+The longer that individuals have been off of parole, the more likely they are to cast a ballot.
+For instance, of the individuals last discharged from parole in 2010, 6.7 percent cast a ballot in the 2016 election, while just 4.3 percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included in the denominator only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
+A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
+This curve is fit on monthly data running from January, 2010 through May, 2016, and extended through October, 2016.
+
+```{r to-16-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-16}Turnout in 2016 Presidential Election", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/to16_chart.rds")
+print(j)
+```
+Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
+Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
+This curve is fit on monthly data running from January, 2012 through May, 2018, and extended through October, 2018.
+```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/to18_chart.rds")
+print(j)
+```
+Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
+Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- New Yorkers whose rights were restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
+
 \newpage

### 2019-05-29 code/part_2/05_to_18.R

@@ -1,63 +1,128 @@
+# 
+# elects <- fread("./raw_data/misc/elects.csv")
+# 
+# ############# READ IN NYS VOTER FILES #############
+# db <- dbConnect(SQLite(), "D:/rolls.db")
+# nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id from nys_roll_0319")
+# 
+# nys_roll <- nys_roll %>% 
+#   mutate(a = as.integer(voter_status == "ACTIVE"),
+#          b = as.integer(voter_status == "INACTIVE"),
+#          c = as.integer(voter_status == "PREREG"))
+# 
+# nys_roll <- setorder(nys_roll, nys_id, -a, -b, -c) ## KEEP ONE RECORD FOR EVERY VOTER
+# nys_roll <- nys_roll[!duplicated(nys_roll$nys_id),]
+# nys_roll <- select(nys_roll, -a, -b, -c)
+# 
+# 
+# 
+# nys_roll <- nys_roll[nys_roll$nys_id %in% readRDS("./temp/din_nys_parolees.rds")$nys_id, ]
+# nys_roll <- left_join(nys_roll, readRDS("./temp/din_nys_parolees.rds"), by = "nys_id")
+# 
+# nys_roll <- cSplit(nys_roll, "history", sep = ";", direction = "long", type.convert = F)
+# 
+# nys_roll <- left_join(nys_roll, elects, by = "history")
+# 
+# 
+# nys_roll <- nys_roll %>% 
+#   group_by(nys_id) %>% 
+#   mutate(year = ifelse(is.na(year), 0, year),
+#          v2018 = max(year == 2018 & election_type == "general"),
+#          v2016 = max(year == 2016 & election_type == "general")) %>% 
+#   filter(row_number() == 1)
+# 
+# saveRDS(nys_roll, "./temp/parolee_to_18.rds")
 
-elects <- fread("./raw_data/misc/elects.csv")
-
-############# READ IN NYS VOTER FILES #############
-db <- dbConnect(SQLite(), "D:/rolls.db")
-nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id from nys_roll_0319")
-
-nys_roll <- nys_roll %>% 
-  mutate(a = as.integer(voter_status == "ACTIVE"),
-         b = as.integer(voter_status == "INACTIVE"),
-         c = as.integer(voter_status == "PREREG"))
-
[…]

### 2019-05-29 code/part_2/05_turnout_trends.R
don't hardcode numbers dummy

@@ -41,37 +41,7 @@ parolees <- left_join(parolees, nys_roll, by = "din")
 parolees$v2018 <- ifelse(is.na(parolees$v2018), 0, parolees$v2018)
 parolees$v2016 <- ifelse(is.na(parolees$v2016), 0, parolees$v2016)
 
-ll <- parolees %>% 
-  filter(parole_status == "DISCHARGED") %>% 
-  mutate(parole_status_date = as.Date(parole_status_date, "%m/%d/%Y"),
-         month_done = make_date(year = year(parole_status_date), month = 4 * (ceiling(month(parole_status_date) / 4)), day = 1)) %>% 
-  group_by(month_done) %>% 
-  summarize(to = mean(v2018),
-            count = n())
-
-
-ggplot(filter(ll, year(month_done) >= 2010), aes(x = month_done, y = to)) + geom_line()
-
-ll2 <- parolees %>% 
-  filter(parole_status == "DISCHARGED") %>% 
-  mutate(parole_status_date = as.Date(parole_status_date, "%m/%d/%Y"),
-         month_done = make_date(year = 1900, month = month(parole_status_date), day = 1),
-         year = year(parole_status_date)) %>% 
-  filter(year %in% c(2016, 2018)) %>% 
-  group_by(month_done, year) %>% 
-  summarize(to16 = mean(v2016), 
-            to18 = mean(v2018),
-            count = n())
-
-ll2$to <- ifelse(ll2$year == 2016, ll2$to16, ll2$to18)
-
-ggplot(filter(ll2), aes(x = month_done, y = to, color = as.factor(year))) + geom_line() +
-  theme_minimal() + scale_y_continuous(labels = scales::percent) + labs(x = "Month Finished With Parole",
-                                                                        y = "Turnout") +
-  scale_x_date(labels = date_format("%B")) +
-  scale_color_manual(name = "Year Finished With Parole", values = c("red", "blue"))
-
-ll3 <- parolees %>% 
+low_level <- parolees %>% 
   filter(parole_status == "DISCHARGED") %>% 
   mutate(parole_status_date = as.Date(parole_status_date, "%m/%d/%Y"),
          month_done = make_date(year = year(parole_status_date), month(parole_status_date), day = 1)) %>% 
@@ -81,36 +51,45 @@ ll3 <- parolees %>%
             count = n())
 
 
[…]

### 2019-05-29 markdown/kmorris_thesis.Rmd
don't hardcode numbers dummy

@@
 ## Identifying Parolees Whose Rights Were Restored {-}
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the DIN number provided from the parolee public records request, I was able to identify individuals who did and did not have their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.^[Only individuals sentenced to prison for felony offenses lose their voting rights in New York.
+As such, I remove individuals sentenced to prison for misdemeanors.
+Incarceration likely also has a *de facto* disenfranchising effect on turnout among individuals incarcerated for misdemeanors, though exploring that effect is beyond the scope of this project.] By using the DIN number provided from the parolee public records request, I was able to identify individuals who did and did not have their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
@@
 I assume that the share of individuals discharged after May 21^st^ whose rights were not restored is comparable to the share of individuals discharged prior to May 21^st^ whose rights would not have been restored if they had been discharged later.
-I also remove from the analyses that follow the 16 individuals discharged from parole prior to May 21^st^, 2018, whose rights were restored prior to their discharge.
+I also remove from the analyses that follow the `r readRDS("./temp/count_restored_before_may21.rds")` individuals discharged from parole prior to May 21^st^, 2018, whose rights were restored prior to their discharge.
 
@@
 The longer that individuals have been off of parole, the more likely they are to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.7 percent cast a ballot in the 2016 election, while just 4.3 percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
 Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included in the denominator only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
@@
 Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- New Yorkers whose rights were restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
+However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy.
+In the section that follows, I develop a logit model to explore whether, after controlling for the available characteristics, individuals discharged from parole after May 21^st^, 2018, were more likely to cast a ballot in the 2018 midterm elections.
+
+## Individual-Level Turnout Regressions {-}
 

### 2019-05-29 code/part_2/03_process_scraped_data.R

@@ -30,7 +30,7 @@ full <- full %>%
   mutate(Release_to_parole_supervision = as.Date(Release_to_parole_supervision, "%m/%d/%Y"),
          restored = tolower(Voting_pardon_issued) == "yes") %>% 
   select(din = DIN, restored, prison_release = Release_to_parole_supervision,
-         status = Parole_status, effective_date = Effective_date, race = Race_ethnicity)
+         status = Parole_status, effective_date = Effective_date)
 
 rm(m, ids, i)
 #### READ IN PAROLEE DATA
@@ -47,9 +47,16 @@ colnames(parolees) <- names
 rm(names)
 
 parolees <- parolees %>% 
-  select(din, parole_status = status, parole_status_date = status_date,
-         release_date_parole = release_date, dob_parole = dob,
-         parole_name = name, -starts_with("filler")) %>% 
+  select(din,
+         parole_status = status,
+         parole_status_date = status_date,
+         release_date_parole = release_date,
+         dob_parole = dob,
+         parole_name = name,
+         -starts_with("filler"),
+         county,
+         sex,
+         race) %>% 
   group_by(din) %>% 
   arrange(desc(parole_status_date)) %>% 
   filter(row_number() == 1) %>% 
@@ -62,9 +69,29 @@ parolees <- cSplit(parolees, "first_name", sep = " ", direction = "wide", drop =
 parolees <- parolees[ , 1:(x+1)]
 parolees <- rename(parolees, first_name = first_name_1, middle_name = first_name_2)
 
-parolees_with_restoration <- full_join(parolees, full)
+parolees_with_restoration <- left_join(parolees, full, by = "din")
+
+
+
+##### read in parolee crimes
+
+names <- fread("./raw_data/doccs_data/parolee_crimes_column_names.csv", header = F)$V1
+
+parolee_crimes <- read_fwf(
+  file = "./raw_data/doccs_data/parolee_data/Parole.Crimes.011419.txt",
+  fwf_widths(c(1, 9, 3, 7, 2, 3, 2, 44, 3, 1, 3, 11, 2))
+)
+
+colnames(parolee_crimes) <- names
+parolee_crimes <- select(parolee_crimes, -starts_with("filler")) %>% 
+  mutate(current_felon = crime_class %in% c("1", "2", "3", "B", "C", "D", "E")) %>% 
+  group_by(din) %>% 
+  summarize(current_felon = max(current_felon),
+            counts = n())
+
[…]

### 2019-05-29 code/part_2/05_turnout_trends.R

@@ -94,7 +94,8 @@ to16_chart <- ggplot(filter(ll3, year(month_done) >= 2010, month_done < "2016-11
   geom_vline(xintercept = as.Date("2016-05-21"), color = "red") +
   scale_x_date(labels = date_format("%b-%Y")) +
   annotate(geom = "text", x = as.Date("2016-05-21"), y = 0.08, label = "May 21, 2016", hjust = 1.05, family = "LM Roman 10") +
-  theme(text = element_text(family = "LM Roman 10"))
+  theme(text = element_text(family = "LM Roman 10"),
+        panel.grid = element_blank())
 
 saveRDS(to16_chart, "./temp/to16_chart.rds")
 
@@ -105,10 +106,24 @@ to18_chart <- ggplot(filter(ll3, year(month_done) >= 2012, month_done < "2018-11
   geom_vline(xintercept = as.Date("2018-05-21"), color = "red") +
   scale_x_date(labels = date_format("%b-%Y")) +
   annotate(geom = "text", x = as.Date("2018-05-21"), y = 0.06, label = "May 21, 2018", hjust = 1.05, family = "LM Roman 10") +
-  theme(text = element_text(family = "LM Roman 10"))
+  theme(text = element_text(family = "LM Roman 10"),
+        panel.grid = element_blank())
 
 saveRDS(to18_chart, "./temp/to18_chart.rds")
+### turnout stats
 
+to_16_10 <- parolees %>% 
+  filter(parole_status == "DISCHARGED",
+         year(as.Date(parole_status_date, "%m/%d/%Y")) == 2010) %>% 
+  summarize(mean(v2016)) %>% 
+  pull()
+saveRDS(to_16_10, "./temp/turnout_in_16_finished_10.rds")
+to_16_15 <- parolees %>% 
+  filter(parole_status == "DISCHARGED",
+         year(as.Date(parole_status_date, "%m/%d/%Y")) == 2015) %>% 
+  summarize(mean(v2016)) %>% 
+  pull()
+saveRDS(to_16_15, "./temp/turnout_in_16_finished_15.rds")
 
 ### restoration
 tt <- parolees %>% 
@@ -130,3 +145,7 @@ restoration_plot <- ggplot(filter(tt, parole_status_date >= "2018-04-15", parole
   theme(text = element_text(family = "LM Roman 10"))
 
 saveRDS(restoration_plot, "./temp/restoration_plot.rds")
+
[…]

### 2019-05-30 markdown/kmorris_thesis.Rmd
more writing

@@
 In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I compare the turnout rates in the 2018 general election of three groups who finished parole in 2018: those who finished parole immediately before the policy change, those who finished parole immediately after the policy change, and those who finished parole immediately after the election.
-These three groups allow me to explore whether the re-extension of voting rights while on parole has an impact on turnout (by comparing the first two groups) and whether being under formal supervision while being allowed to vote results in higher turnout than being allowed to vote but being under supervision (by comparing the third group to the first two).
+To test whether the policy was successful, I build an individual-level logit model to explore whether turnout rates among individuals discharged from parole after Executive Order 181 went into effect cast ballots at higher rates than those discharged previously.
 
@@
 Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.^[The New York registered voter file does not align exactly with results reported by the city.
-The voter file indicates that just XX voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that XX votes were cast.
+The voter file indicates that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that `r comma(readRDS("./temp/mayoral_ballots.rds"))` votes were cast.
 In [Appendix B], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
@@
 This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes insignificant.
+When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
 The coefficient on the interaction between treatment and share Black indicates that neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
@@
 For every 100 votes cast in a predominantly Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
-Although these variables are largely insignificant at the census tract level, this is not particularly surprising.
+Although these variables are largely nonsignificant at the census tract level, this is not particularly surprising.
 If lost voters have the largest depressive effect on family members and close neighbors, it makes sense that the effect of losing a voter is locally concentrated.
@@
 The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].^[I've also talked with folks on probation whose probation officers told them they were disenfranchised.
-Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the XX voters living in the community and to clarify the rules about who is eligible to vote.
+Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
@@
 Under the Executive Order, however, parole officers were required to explicitly inform their supervisees of their newly restored voting rights.
-It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly eligible voters.
+It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly re-eligible voters to cast a ballot.
 
@@
 
+Figure \ref{fig:to-18} displays the share of individuals who were discharged from parole in each month and cast a ballot in the 2018 elections.
+It does not control for any sort of individual level characterisitcs.
+In \ref{tab:to-18-logit}, I present the results of an individual-level regression exploring the impact of Executive Order 181 on turnout in 2018.
+Because turnout is a binary dependent variable, this is a logistic regression.
+It includes all individuals discharged from parole between January 1^st^, 2012, through October 10^th^, 2018 (the registration deadline in New York State).^[This of course ignores the individuals who were still on parole and cast a ballot in the 2018 election.
+However, for this exercise, I limit the analysis group to individuals who would have been eligible to vote on election day even if Governor Cuomo had not signed Executive Order 181.]
+
+```{r logit-reg-chunk, echo = F, results = "asis", message = F}
+load("./temp/individual_turnout_18.rdata")
+
+stargazer(model1, model2, model3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2018)", "Days Since Discharged",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
+          table.placement = "H",
+          style = "ajps",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+
+Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual was discharged on or after May 21^st^, 2018, and the number of days between the individual's discharge date and November 6^th^, 2018.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county, and race.
+Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class of felony for which they were convicted.
+In each successive model, the AIC decreases substantially, indicating that including these controls are warranted.
+Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+
+Each model also indicates that Executive Order 181 was successful at increasing turnout among formerly incarcerated New Yorkers.^[These models do not explicitly control for the potentiality that individuals discharged from parole shortly before an election always turnout at higher rates.
+In [Appendix C], I run the models from Table \ref{tab:to-18-logit} on turnout in the 2016 presidential election.
+[Appendix C] shows that in 2016, individuals discharged from parole in the months leading up to the election were no more likely to participate in that election] Exponentiating the coefficients on D(Discharged on or after May 21^st^, 2018) indicates that individuals who were discharged after Executive Order 181 went into effect were between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote.
+
 \newpage
@@
 As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
-According to the official results, XX ballots were cast in the 2017 mayoral election.
-The registered voter file reports that just XX voters cast a ballot.
+According to the official results, `r comma(readRDS("./temp/mayoral_ballots.rds"))` ballots were cast in the 2017 mayoral election.
+The registered voter file reports that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot.
 This is not necessarily evidence of poor management by the Board of Elections.
@@
 Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
+
+\newpage
+# Appendix C {-}
+
+Table \ref{tab:to-18-logit} shows that individuals who were discharged from parole after Governor Cuomo signed Executive Order 181 were substantially more likely to cast a ballot in the 2018 election than individuals who were discharged prior to the policy change.
+It is, of course, possible that individuals discharged from parole immediately before an election are always more likely to vote, due perhaps to the recency of their interation with state systems and a desire to have their voices heard.
+In that case, claiming that Executive Order 181 caused turnout to increase would be a mistake.
+
+To test whether this is true --- that individuals discharged shortly before an election turn out at higher rates --- I fit the models from Table \ref{tab:to-18-logit} on 2016 turnout.
+Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2010, and October 14^th^, 2010.
+```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
+load("./temp/individual_turnout_16.rdata")
+
+stargazer(model1, model2, model3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2016 Election"),
+          title = "(\\#tab:to-16-logit) Individual-Level Logit Model",
+          table.placement = "H",
+          style = "ajps",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that individuals discharged in the summer of 2016 were more likely to cast a ballot than individuals who finished parole earlier (the p-value on the coefficient of interest exceeds 0.5 in each model).
+The nonsignificant results from 2016 provides strong corroboration for causal claims regarding the efficacy of Executive Order 181 at boosting turnout.

### 2019-06-03 markdown/kmorris_thesis.Rmd
more writing

@@
 Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class of felony for which they were convicted.
+Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
 In each successive model, the AIC decreases substantially, indicating that including these controls are warranted.

### 2019-06-04 markdown/kmorris_thesis.Rmd
write 6/4/19

@@
 loadfonts()
-
 ```
@@
 
+As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
+@King2016; @Bowers2009].
+With the exception of Burch [-@Burch2013], however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
+The degree to which these effects are concentrated in particular neighborhoods has important implications for our understanding of the problem of felony disenfranchisement.
+If the depressive effects are widely spread throughout a given state, the effects of felony disenfranchisement are also likely to be widely spread.
+In such a world, disenfranchisement is likely to undermine political support for br
+
 Although understanding whether felony disenfranchisement impacts turnout among different social groups is of great importance, understanding the explicitly spatial dimension of the problem is also an important task.
@@
 
-## Previous Attempts to Re-Incorporate Formerly Disenfranchised Residents {-}
+## Turnout Among Formerly Disenfranchised Individuals {-}
 
-* Super low turnout
-* Immediate effects on future participation - Ariel's paper [@WHITE2019]
-* Changes in Notification Laws (M&M)
+Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
+The effect of incarceration on participation is the subject of some debate within the field.
+Individuals who go to prison share many characterstics with lower propensity voters generally.
+Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
+In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
+[-@Gerber2017] uses administrative data from Pennsylvania to estimate turnout rates prior to incarceration.
+They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects onsubsequent participation."
+
+White [-@White2019], however, indicates that interaction with the criminal justice system for individuals in the context of arrests for misdemeanor charges may have depressive effects on turnout.
+This finding does not necessarily conflict with Gerber et al.
+[-@Gerber2017]; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
+The findings in White [-@White2019] agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
+Weaver and Lerman [-@Weaver2010] argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual."
+
+Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
+A criminal justice system that inflicts dire consequences on a population that has relatively little political voice is problematic.
+Whether or not incarceration *causes* low turnout, however, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout.
+
+Some research has been done in this area.
+Meredith and Morse [-@Meredith2015] examines the impact of ending permanent disenfranchisement in Iowa.
+They find that individuals who received letters explicitly informing of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
+Meredith and Morse [-@Meredith2013], however, examines states where so-called notification laws went into effect.
+Although rules about eligibility did not change in these states, new policies required Departments of Corrections to formally inform formerly disenfranchised individuals of their re-instated voting rights.
+Meredith and Morse [-@Meredith2013] finds no effect from notification in the absense of eligibility changes.
+
+Gerber et al.
+[-@Gerber2014] conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+They find that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome." It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
+
+Executive Order 181 is expected to have increased turnout among formerly incarcerated individuals through two primary mechanisms.
+The first addresses the impact of negative experiences with the state.
+To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
+Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel explicitly invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
+
+Secondly, Executive Order 181 is expected to dispel confusion about eligibilty and serve as a reminder to vote.
+As Drucker and Barreras [-@Drucker2005] and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
+If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.
+Even if confusion does not lead to lower turnout, there is some reason to believe that reminding formerly incarcerated individuals of their rights is a successful intervention for boosting turnout [@Meredith2015; @Gerber2014].
 
@@
           se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
-          notes = "Robust standard errors (clustered at match level) in parentheses.",
+          notes = c("Robust standard errors (clustered at", "match level) in parentheses."),
           type = "latex",
@@
           table.placement = "H",
-          style = "ajps",
           omit.stat = c("f", "ser"))
@@
           table.placement = "H",
-          style = "ajps",
           omit.stat = c("f", "ser"),
-          notes = "Robust standard errors (clustered by city council district) in parentheses.",
+          notes = c("Robust standard errors (clustered by city",  "council district) in parentheses."),
           omit = "boro")
@@
 Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents far less likely to vote in the future [@WHITE2019; @XX].
+As previous literature has established, interactions with the criminal justice system leaves residents far less likely to vote in the future [@White2019; @XX].
 As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are vanishingly low.
@@
 These are likely individuals who are not citizens, and therefore ineligible to register to vote.
-```{r restoration-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
+```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
 j <- readRDS("./temp/restoration_plot.rds")
@@
           table.placement = "H",
-          style = "ajps",
           omit = c("county", "race", "felony"),
@@
           table.placement = "H",
-          style = "ajps",
           omit.stat = c("f", "ser"),
-          notes = "Robust standard errors (clustered by assembly district) in parentheses.",
+          notes = c("Robust standard errors (clustered by", "assembly district) in parentheses."),
           omit = "boro")

### 2019-06-05 markdown/kmorris_thesis.Rmd
more writing

@@
 
-<!-- --- -->
-<!-- title: "Rethinking Felony Disenfranchisement: The Case of New York City and Miami" -->
-
-<!-- # to produce blinded version set to 1 -->
-<!-- blinded: 0 -->
-
-<!-- authors: -->
-<!-- - name: Kevin Morris -->
-<!--   thanks: The author gratefully acknowledges Jacob Faber and Jeff Manza for their advisement of this project.
-He would also like to thank colleagues at the Brennan Center for Justice for their insights into this project.
--->
-<!--   affiliation: Robert F. Wagner School of Public Service, New York University -->
-
-<!-- keywords: -->
-<!-- - 3 to 6 keywords -->
-<!-- - that do not appear in the title -->
-
-<!-- abstract: | -->
-<!--   The text of your abstract.  200 or fewer words. -->
-
-<!-- output: rticles::asa_article -->
-<!-- --- -->
 ```{r setup, echo = F, message = F, warning = F}
@@
 
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Christopher Uggen and Jeff Manza [-@Uggen2002].
+In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
 George W.
@@
 In a series of papers in between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-Haselswerdt [-@Haselswerdt2009] matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
+@Haselswerdt2009 matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
 Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
@@
 A number of papers have also explored the impact that felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson [-@King2016], for instance, leverage state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
+@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
 They use data from the 2004 Current Population Survey Voting and Registration Supplement to determine individual-level turnout.
@@
 These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
-As @Bowers2009 sum up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
+As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
 [Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty."
@@
 To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representatition of certain parts of the city.
-Although turnout is not a perfect proxy for political representation, there is some evidence that policians are more responsive to groups that turn out at higher rates.
-@Martin2013, for instance, documents the relationship between voter turnout and the responsiveness of Congressional representatives.
-"[H]igher citizen participation is rewarded," they argue, "with enhanced policy responsiveness." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
-@Griffin2005 find similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not."
-
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
-If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most neglected neighborhoods to begin with.
 
@@
 @King2016; @Bowers2009].
-With the exception of Burch [-@Burch2013], however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
+With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
 The degree to which these effects are concentrated in particular neighborhoods has important implications for our understanding of the problem of felony disenfranchisement.
-If the depressive effects are widely spread throughout a given state, the effects of felony disenfranchisement are also likely to be widely spread.
-In such a world, disenfranchisement is likely to undermine political support for br
+Although literature examining the degree to which geographic polarization effects voting patterns is scarce, there is some evidence to suggest that neighborhoods do vote similarly.
+@Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
+They find evidence that, over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
+Insofar as neighborhoods show distinct preferences for politicians, spatially concentrated depressive effects from felony disenfranchisement could influence election outcomes.
 
-Although understanding whether felony disenfranchisement impacts turnout among different social groups is of great importance, understanding the explicitly spatial dimension of the problem is also an important task.
-Many political decisions bear on the spatial distribution of resources (XX); funding for parks or schools, for instance, can impact all members of a neighborhood, regardless of their race, income, or criminal justice involvement.
-As such, even a resident who is not directly affected by felony disenfranchisement policies may lose political power thanks to these policies; if many of my neighbors cannot vote, I am joined by fewer voters in advocating for neighborhood-specific investment and services.
-This is, of course, tied up closely with race; New York City remains highly racially segregated.
-Thanks to spatially disparate policing patterns, any depressive effect is therefore likely to be concentrated among Black voters.
+Even if neighborhoods to not have unique preferences for presidential candidates, lower turnout in impacted communities may reduce their allocation of public goods.
+At the Congressional level, for instance, representatives direct resources to areas within their districts that provide the greatest political benefit [@Martin2003].
+Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
+"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness." @Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
 
-Despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
+Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
+Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
+As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
+An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
+Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
+If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most neglected neighborhoods to begin with.
+
+Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
 In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
@@
 This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-The existence of such *de facto* disenfranchisement is bolstered by a large body of empirical work on the social nature of voting.
-Many scholars have established that being in a community where voting is common makes individuals more likely to vote.
-Children whose parents vote, for instance, are more likely to vote as adults [XX]; XX add more examples here.
+
+This *de facto* disenfranchisement is likely to be concentrated within the neighborhoods home to formally disenfranchised residents.
+Voting is a social act, and social networks play an enormous role in predicting political participation [e.g.
+@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
 Literature from urban sociology has established that social networks are largely spatially bounded [XX].
@@
 Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
-[-@Gerber2017] uses administrative data from Pennsylvania to estimate turnout rates prior to incarceration.
+In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
 They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects onsubsequent participation."
 
-White [-@White2019], however, indicates that interaction with the criminal justice system for individuals in the context of arrests for misdemeanor charges may have depressive effects on turnout.
-This finding does not necessarily conflict with Gerber et al.
-[-@Gerber2017]; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+@White2019, however, indicates that interaction with the criminal justice system for individuals in the context of arrests for misdemeanor charges may have depressive effects on turnout.
+This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
 Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-The findings in White [-@White2019] agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-Weaver and Lerman [-@Weaver2010] argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual."
+The findings in @White2019 agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
+@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual."
 
@@
 Some research has been done in this area.
-Meredith and Morse [-@Meredith2015] examines the impact of ending permanent disenfranchisement in Iowa.
+@Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
 They find that individuals who received letters explicitly informing of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-Meredith and Morse [-@Meredith2013], however, examines states where so-called notification laws went into effect.
+@Meredith2013, however, examines states where so-called notification laws went into effect.
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to formally inform formerly disenfranchised individuals of their re-instated voting rights.
-Meredith and Morse [-@Meredith2013] finds no effect from notification in the absense of eligibility changes.
+@Meredith2013 finds no effect from notification in the absense of eligibility changes.
 
-Gerber et al.
-[-@Gerber2014] conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election.
 They find that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
@@
 Secondly, Executive Order 181 is expected to dispel confusion about eligibilty and serve as a reminder to vote.
-As Drucker and Barreras [-@Drucker2005] and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
+As @Drucker2005 and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
 If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.

### 2019-06-05 markdown/kmorris_thesis.Rmd
start writing IV section

@@
 
+Of course, in these models we are *not* directly testing the impact of having ones rights restored on propensity to vote.
+In the next section, I employ an instrumental variables approach to confirm that Executive Order was in fact successful at raising turnout.
+
+## Instrumenting Rights Restoration {-}
+
+In the previous section, I explored whether individuals who were discharged from parole after Executive Order 181 went into effect were more likely to vote, whether or not their voting rights were actually restored.
+There is reason to believe that whether an individual received the treatment (had their rights restored) is correlated with their propensity to vote.
+For instance, parolees who were arrested did not have their rights restored; a recent arrest is also likely to impact turnout.
+Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also correlated with propensity to vote.
+Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the effect of Executive Order 181.
+This calls for an instrumental variables approach.
+
+Although \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is far less signficant in the short run.
+Figure \ref{fig:to-18-recent} shows that, for individuals discharged from parole in 2017 and 2018, there does not appear to be a relationship between time off parole and turnout rates.
+
+```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Presidential Election", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/to18_chart_recent.rds")
+print(j)
+```
+
+Formalizing this chart into an individual-level logit model demonstrates that this relationship does not hold over the short run.
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and May 20^th^, 2018.
+
+```{r logit-reg-chunk-short, echo = F, results = "asis", message = F}
+load("./temp/short_term_to.rdata")
+
+stargazer(date_to_model_notime, date_to_model_time,
+          header = F,
+          type = "latex",
+          covariate.labels = c("Days Since Discharged",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "(\\#tab:to-18-logit-short) Individual-Level Logit Model",
+          table.placement = "H",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "X", "X"),
+                         c("Felony Class Fixed Effects", "X", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+
+Adding controls for time in Table \ref{tab:to-18-logit-short} increases the AIC, indicating that the inclusion of these variables is not warranted.
+A Chi-squared test confirms that the model is not improved when controls for time are included.
+
+Although date of discharge is not correlated with turnout over the 2017 -- 2018 period, it is clear from Figure \ref{fig:restoration-chart} that the date of discharge from parole *is* correlated with formal rights restoration.
+Although not all individuals released from parole after May 21^st^ had their rights formally restored, many did.
+A dummy variable indicating whether an individual was released on or after May 21^st^, 2018, therefore, is a good instrument for formal rights restoration.
+
 \newpage

### 2019-06-06 markdown/kmorris_thesis.Rmd
continue incorporating IV

@@
 knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf")
+library(systemfit)
+library(AER)
 library(sandwich)
@@
 As such, I remove individuals sentenced to prison for misdemeanors.
-Incarceration likely also has a *de facto* disenfranchising effect on turnout among individuals incarcerated for misdemeanors, though exploring that effect is beyond the scope of this project.] By using the DIN number provided from the parolee public records request, I was able to identify individuals who did and did not have their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+Incarceration likely also has a *de facto* disenfranchising effect on turnout among individuals incarcerated for misdemeanors [@White2019], though exploring that effect is beyond the scope of this project.] By using the DIN number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
@@
 
-Although Governor Cuomo signed the Executive Order in April of 2018, an examination of the individuals whose rights were ultimately restored indicates that the program may not have gone into full effect until later in May.
-As Figure \ref{fig:restoration-chart} demonstrates, individuals who finished their parole supervision in early May did not have their voting rights formally restored until they finished their term of supervision.
-It is not until May 21^st^ that the majority of individuals being discharged from parole had their voting rights restored prior to their discharge.
-However, even after the policy change went into effect, not all parolees had their voting rights restored prior to being discharged from parole.
-These are likely individuals who are not citizens, and therefore ineligible to register to vote.
-```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/restoration_plot.rds")
-print(j)
-```
-Figure \ref{fig:restoration-chart} complicates our ability to cleanly identify the impact of the policy change.
-If no parolees discharged prior to May 21^st^ had had their rights restored, but all parolees discharged on or after May 21^st^ had, identifying those impacted by the policy would be straightforward.
-That is not the case: some select individuals seem to have had their rights restored before the policy change was fully implemented.
-Similarly, we would ideally exclude all individuals who did not have their rights restored after the policy change as well as the individuals who were discharged before the change who would not have had their rights restored if they had been discharged from parole later.
-However, identifying individuals whose rights would not have been restored is not possible.
-
-In the analyses that follow, I consider all parolees discharged on or after May 21^st^, 2018, beneficiaries of the policy change, whether their rights were restored or not.
-I assume that the share of individuals discharged after May 21^st^ whose rights were not restored is comparable to the share of individuals discharged prior to May 21^st^ whose rights would not have been restored if they had been discharged later.
-I also remove from the analyses that follow the `r readRDS("./temp/count_restored_before_may21.rds")` individuals discharged from parole prior to May 21^st^, 2018, whose rights were restored prior to their discharge.
-
 ## Trends in Turnout {-}
 Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier.
-If individuals who were discharged from parole prior to May 21^st^, 2016, were less likely to vote than those discharged between May 21^st^ and Election Day, disentangling the impact of the policy change in 2018 from the expected uptick would be more complicated.
+It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
 However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier in the year.
@@
 A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2010 through May, 2016, and extended through October, 2016.
+This curve is fit on monthly data running from January, 2010 through April, 2016, and extended through October, 2016.
 
@@
 Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2012 through May, 2018, and extended through October, 2018.
+This curve is fit on monthly data running from January, 2012 through April, 2018, and extended through October, 2018.
 ```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
@@
 Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- New Yorkers whose rights were restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
+Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
 However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy.
-In the section that follows, I develop a logit model to explore whether, after controlling for the available characteristics, individuals discharged from parole after May 21^st^, 2018, were more likely to cast a ballot in the 2018 midterm elections.
+In the section that follows, I develop a logit model to explore whether, after controlling for the available characteristics, individuals whose rights were restored prior to being discharged were more likely to cast a ballot in the 2018 midterm elections.
 
@@
 It does not control for any sort of individual level characterisitcs.
-In \ref{tab:to-18-logit}, I present the results of an individual-level regression exploring the impact of Executive Order 181 on turnout in 2018.
+In \ref{tab:to-18-logit}, I present the results of an individual-level regression exploring the turnout in 2018 among individuals whose rights were restored.
 Because turnout is a binary dependent variable, this is a logistic regression.
@@
           type = "latex",
-          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2018)", "Days Since Discharged",
+          covariate.labels = c("D(Rights Restored)", "Days Since Discharged",
                                "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
@@
 
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual was discharged on or after May 21^st^, 2018, and the number of days between the individual's discharge date and November 6^th^, 2018.
+Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual had his rights restored before being discharged from parole, and the number of days between the individual's discharge date and November 6^th^, 2018.
 Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county, and race.
@@
 
-Each model also indicates that Executive Order 181 was successful at increasing turnout among formerly incarcerated New Yorkers.^[These models do not explicitly control for the potentiality that individuals discharged from parole shortly before an election always turnout at higher rates.
-In [Appendix C], I run the models from Table \ref{tab:to-18-logit} on turnout in the 2016 presidential election.
-[Appendix C] shows that in 2016, individuals discharged from parole in the months leading up to the election were no more likely to participate in that election] Exponentiating the coefficients on D(Discharged on or after May 21^st^, 2018) indicates that individuals who were discharged after Executive Order 181 went into effect were between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote.
+Each model also indicates that individuals whose rights were restored prior to discharge were more likely to cast a ballot than those whose rights were not restored.
+Exponentiating the coefficients on D(Rights Restored) indicates that individuals whose rights were restored due to Executive Order 181 were between `r round((exp(model1[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote than those who did not.
 
-Of course, in these models we are *not* directly testing the impact of having ones rights restored on propensity to vote.
-In the next section, I employ an instrumental variables approach to confirm that Executive Order was in fact successful at raising turnout.
+Table \ref{tab:to-18-logit} provides some indication that Executive Order 181 was successful, but does not establish causality.
+There is reason to believe that individuals who had their rights restored were more likely to vote even had they not been "treated" by EO 181.
+In the section that follows, I employ an instrumental variables approach to control for this possibility.
 
@@
 
-In the previous section, I explored whether individuals who were discharged from parole after Executive Order 181 went into effect were more likely to vote, whether or not their voting rights were actually restored.
+In the previous section, I explored whether individuals who had their rights formally restored prior to parole discharge were more likely to vote in the 2018 midterm election.
 There is reason to believe that whether an individual received the treatment (had their rights restored) is correlated with their propensity to vote.
-For instance, parolees who were arrested did not have their rights restored; a recent arrest is also likely to impact turnout.
-Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also correlated with propensity to vote.
+For instance, parolees who were arrested while on parole did not have their rights restored; a recent arrest is also likely to impact turnout.
+Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also of course correlated with propensity to vote.
 Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the effect of Executive Order 181.
@@
 
+The timing of Executive Order 181 serves as a useful instrument for formal rights restoration.
 Although \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is far less signficant in the short run.
@@
 
-Formalizing this chart into an individual-level logit model demonstrates that this relationship does not hold over the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and May 20^th^, 2018.
+Formalizing this chart into an individual-level logit model demonstrates that time since discharge is not correlated with turnout in the short run.
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
 
@@
 Adding controls for time in Table \ref{tab:to-18-logit-short} increases the AIC, indicating that the inclusion of these variables is not warranted.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix C] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.]
 
-Although date of discharge is not correlated with turnout over the 2017 -- 2018 period, it is clear from Figure \ref{fig:restoration-chart} that the date of discharge from parole *is* correlated with formal rights restoration.
-Although not all individuals released from parole after May 21^st^ had their rights formally restored, many did.
-A dummy variable indicating whether an individual was released on or after May 21^st^, 2018, therefore, is a good instrument for formal rights restoration.
+Although date of discharge from parole is not correlated with turnout in the short run, it is correlated with likelihood of being having one's rights restored while still on parole.
+Governor Cuomo signed Executive Order 181 in April of 2018, but an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
+As Figure \ref{fig:restoration-chart} demonstrates, individuals who finished their parole supervision in early May did not have their voting rights restored before they finished their term of supervision.
+It is not until May 21^st^ that the majority of parolees had their voting rights restored prior to their discharge.
+```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/restoration_plot.rds")
+print(j)
+```
+There is no evidence that date of discharged influences propensity to vote through channels other than formal rights restoration in the short run.
+A dummy variable indicating whether an individual was discharged on or after May 21^st^ can therefore effectively be used to instrument rights restoration.
+Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using OLS.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
+
+```{r iv-reg-chunk-short, echo = F, results = "asis", message = F}
+load("./temp/iv_individual_turnout_18.rdata")
+
+summ.fit1 <- summary(iv1, diagnostics = T)
+summ.fit2 <- summary(iv2, diagnostics = T)
+summ.fit3 <- summary(iv3, diagnostics = T)
+
+stargazer(iv1, iv2, iv3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Rights Restored)", 
+                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "(\\#tab:iv-models) Second-Stage Regression",
+          table.placement = "H",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X"),
+                         c("\\hline \\\\[-1.8ex]"),
+                         c(rownames(summ.fit1$diagnostics)[1],
+                           round(summ.fit1$diagnostics[1, "p-value"], 2),
+                           round(summ.fit2$diagnostics[1, "p-value"], 2),
+                           round(summ.fit3$diagnostics[1, "p-value"], 2)),
+                         c(rownames(summ.fit1$diagnostics)[2],
+                           round(summ.fit1$diagnostics[2, "p-value"], 2),
+                           round(summ.fit2$diagnostics[2, "p-value"], 2),
+                           round(summ.fit3$diagnostics[2, "p-value"], 2))),
+          omit.stat = "ser",
+          table.layout = "-dm#-t-as-n",
+          notes = c("Table reports p-values for Wu-Hausman and",
+                    "Weak Instruments tests.",
+                    "D(Discharged from Parole on or after May",
+                    "21, 2018) instrument for D(Rights Restored)"))
+```
+
+Table \ref{tab:iv-models} indicates that using D(Discharged from Parole on or After May 21) is likely warranted.
+The p-values of the Wu-Hausman tests hover around 0.05, while the Weak Instruments test is highly significant in each model.
+
+Model 3 in Table \ref{tab:iv-models} indicates that having one's rights restored caused an individual's propensity to vote to increase by `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage points.
+Approximately `r round(readRDS("./temp/to_restored.rds") * 100, digits = 2)` percent of individuals who had their rights restored cast a ballot in 2018, indicating that Executive Order 181 increased the propensity to vote by around `r round(summary(iv3)[["coefficients"]][2,1] / (readRDS("./temp/to_restored.rds") - summary(iv3)[["coefficients"]][2,1]) * 100, digits = 1)` percent.
+This estimate is much lower than the estimates reported in Table \ref{tab:to-18-logit}, but that is not surprising; as discussed above, individuals whose rights were restored were more likely to vote whether or not they received restoration before discharge.
+The `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage point increase reported in Table \ref{tab:iv-models} accounts only for the extent to which rights restoration prior to discharge from parole caused higher turnout.
 
@@
 
-Table \ref{tab:to-18-logit} shows that individuals who were discharged from parole after Governor Cuomo signed Executive Order 181 were substantially more likely to cast a ballot in the 2018 election than individuals who were discharged prior to the policy change.
-It is, of course, possible that individuals discharged from parole immediately before an election are always more likely to vote, due perhaps to the recency of their interation with state systems and a desire to have their voices heard.
-In that case, claiming that Executive Order 181 caused turnout to increase would be a mistake.
-
-To test whether this is true --- that individuals discharged shortly before an election turn out at higher rates --- I fit the models from Table \ref{tab:to-18-logit} on 2016 turnout.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2010, and October 14^th^, 2010.
+In the [Instrumenting Rights Restoration] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 21^st^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
 ```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
@@
 ```
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that individuals discharged in the summer of 2016 were more likely to cast a ballot than individuals who finished parole earlier (the p-value on the coefficient of interest exceeds 0.5 in each model).
-The nonsignificant results from 2016 provides strong corroboration for causal claims regarding the efficacy of Executive Order 181 at boosting turnout.
+Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot (the p-value on the coefficient of interest exceeds 0.5 in each model).
+The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.

### 2019-06-06 markdown/kmorris_thesis.Rmd
intent to treat

@@
 There is no evidence that date of discharged influences propensity to vote through channels other than formal rights restoration in the short run.
-A dummy variable indicating whether an individual was discharged on or after May 21^st^ can therefore effectively be used to instrument rights restoration.
+A dummy variable indicating whether an individual was discharged on or after May 21^st^ is therefore analogous to an intent-to-treat, with a dummy indicating whether one's rights were restored measuring imperfect compliance.
 Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using OLS.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]

### 2019-06-10 markdown/kmorris_thesis.Rmd
re-work introduction (more to come)

@@
   bookdown::pdf_document2:
-    toc: no
+    toc: yes
     fig_caption: yes
@@
 
-Felony disenfranchisement has received much attention over the past few years.
-This is true in both the academic world (where researchers have dug into the political ramifications in such races as the 2000 presidential election) and in the political world as states such as Florida and Louisiana have gradually moved to dismantle their systems of disenfranchisement.
-Many researchers, however, have focused on the effects of felony disenfranchisement on individuals, and on racial minorities as a group.
-They have examined the effect of disenfranchisement on the political participation of individuals, including after they are no longer disenfranchised, and the spillover effects disenfranchisement has on eligible voters of color.
-Researchers have generally neglected to examine the intersection of spatially concentrated policing and incarceration patterns with the political effects of disenfranchisement.
-This paper seeks to increase our understanding of the spatial implications of felony disenfranchisement by examining the relationship between disenfranchisement and neighborhood turnout in New York City.
+The political history of the United States has been characterized by a general trend toward universal sufferage.
+At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
+Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
+And yet, despite the United State's purported march toward ever-more-inclusive systems of democracy, one large group of American citizens is formally barred from voting.
+In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
+Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised today [@sentencing_2016].
+These disenfranchising policies provide a stark reminder that our democracy is not as inclusive as we might like to tell ourselves.
 
-This paper also seeks to complicate the narratives surrounding the end of felony disenfranchisement.
-Although advocates are correct to push for an end to disenfranchisement, a change to the political system may not be sufficient to reincorporate the voices of formerly incarcerated individuals into the democratic process.
-In addition to laying out the spatial implications of felony disenfranchisement in New York City, this paper also explores the impact of one policy intended to weaken the disenfranchisement program.
-Specifically, I investigate the effect of New York Governor Andrew Cuomo's Executive Order 181 which ended the disenfranchisement of parolees in New York State.
-The effect of these policies make clear that, while the end of felony disenfranchisement is necessary, it is not enough: Active steps must be taken to reincorporate the formerly disenfranchised.
+The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
+As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
+This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
+This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing.
+As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+
+Due to economic and racial segregation, these effects are also spatially concentrated.
+Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
+Many scholars have detailed the impact of living in areas with high levels of police activity.
+Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
+The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
+Concentrated policing has also credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+
+Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
+The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges stemming from targeted policing.
+The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Burch2011; @Gerber2014].
+Others have looked at the impact of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
+With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
+
+This paper seeks to extend the extant literature on felony disenfranchisement by considering the implications of disenfranchisement on local political participation.
+I begin by looking at participation rates in New York City neighborhoods where individuals with a demonstrated history of participation have been disenfranchised.
+This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
+The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if these already-disadvantaged communities suffer *de facto* disenfranchisement from felony disenfranchisement, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
+
+Undoing the negative effects of felony disenfranchisement on individual and neighborhood representation requires crafting policy aimed at increasing turnout among the formerly disenfranchised.
+New York State has taken a step in that direction: in April of 2018, Governor Andrew Cuomo signed Executive Order 181, which came close to ending disenfranchisement for New Yorkers on parole.
+After examining the neighborhood effects of felony disenfranchisement, I test whether having one's rights restored while still on parole increased turnout among formerly disenfranchised individuals in the 2018 midterm elections.
 
@@
 The timing of Executive Order 181 serves as a useful instrument for formal rights restoration.
-Although \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is far less signficant in the short run.
+Although Figure \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is far less signficant in the short run.
 Figure \ref{fig:to-18-recent} shows that, for individuals discharged from parole in 2017 and 2018, there does not appear to be a relationship between time off parole and turnout rates.

### 2019-06-11 markdown/kmorris_thesis.Rmd
write

@@
 link-citations: yes
+
 ---
@@
 ```
+\pagenumbering{gobble}
+\newpage
+\pagenumbering{arabic}
 # Introduction {-}
@@
 
-Due to economic and racial segregation, these effects are also spatially concentrated.
+Due to economic and racial segregation, these effects are highly spatially concentrated.
 Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
@@
 As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were intended to" [@Burch2010].
-The historical record is undeniable.
 Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
-In Florida felony disenfranchisement was added to the state constitution in 1868.
+In Florida, felony disenfranchisement was added to the state constitution in 1868.
 Afterwards, a lawmaker boasted that the effort had been made in order to prevent Florida from being "niggerized" [@florida_1868].
@@
 Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has vastly increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
+Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
 As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
@@
 
-Recent data Florida make the victims of felony disenfranchisement laws especially clear.
+Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
 Prior to January, 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
@@
 
-In a series of papers in between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
+In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
 @Haselswerdt2009 matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
@@
 This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017.
-Policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods [XX].
+Policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods [e.g.
+@Sewell2016; @Clear2008; @Lerman2013].
 It is possible, however, that these effects are even broader than the direct focus on policing has acknowledged.
@@
 Insofar as neighborhoods show distinct preferences for politicians, spatially concentrated depressive effects from felony disenfranchisement could influence election outcomes.
+Previous research on felony disenfranchisement has been helpful for thinking through the potential implications of felony disenfranchisement on explored the potential impact of
 
@@
 This *de facto* disenfranchisement is likely to be concentrated within the neighborhoods home to formally disenfranchised residents.
-Voting is a social act, and social networks play an enormous role in predicting political participation [e.g.
+Voting is a social act, and social networks play an important role in predicting political participation [e.g.
 @Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
-Literature from urban sociology has established that social networks are largely spatially bounded [XX].
+Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 
-This paper employs multiple data sets to investigate the extent to which a neighborhood is directly impacted by felony disenfranchisement.
-The primary data set comes from a freedom of information request filed by the author to obtain individual-level incarceration records for individuals sentenced to incarceration in New York State since 1992.
-The data includes a host of information, including: first and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; and others.
+The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
 This analysis is limited to individuals incarcerated for felony offenses.
@@
 These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-
-The third dataset used to estimate the extent to which a neighborhood is directly impacted by felony disenfranchisement policies in New York State comes from another freedom of information request.
-The individual-level incarceration data referenced above does not include information about the neighborhood an incarcerated person lived in prior to their conviction.
-Although the State of New York does not make this information available at the individual level, the author has received prison-admission counts by zip-code for the years XX -- XX.
+These data are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
 
@@
 In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first and last name; date of birth; home address; vote history; and other information.
+It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
@@
 ## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.^[Because the criminal justice data does not include individuals' addresses, incarcerated individuals who have never registered to vote cannot be geocoded.] I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
 
@@
 Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
+Such an approach is informative for understanding the impact of disenfranchisement.
+Many young men, for instance, are admitted to prison each year.
+If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
+This approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
+Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
+
+Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
+As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration.
+If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
+At the local level where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition provides insight only into the effect of incarceration on turnout, not felony disenfranchisement.
+
 Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
 In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
+As discussed above, voting is a highly social act.
+Individuals whose family members and neighbors cast ballots are more likely to cast a ballot.
+Identifying the effects of felony disenfranchisement on individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into the extent to which felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
+
 To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
@@
 Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents far less likely to vote in the future [@White2019; @XX].
-As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are vanishingly low.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
+As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
 Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
@@
 
+## Discussion {-}
+As discussed above, the effect of incarceration on individuals' propensity to vote is an open question for the literature.
+What is known, however, is that individuals who have been to prison were likely to vote at very low rates prior to incarceration, and continue to vote at very low rates after incarceration.
+Whether incarceration reduces political participation or not, increasing the propensity to vote of individuals who have been to prison is a laudable goal.
+Prisoners overwhelmingly come from marginalized communities with much to gain from policy.
+Executive Order 181 appears to have substantially boosted turnout among formerly incarcerated individuals through two mechanisms.
+The first is obvious: parolees who would not have been eligible to vote absent the Executive Order were allowed to cast ballots in the midterm elections.
+The second mechanism is less obvious.
+Individuals who finished parole between May 21^st^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
+The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
+The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 40 percent.
+
+The precise mechanism through which the rules change increased turnout among these individuals is not clear.
+It is possibly a social mechanism: having an officer of the state affirm one's eligibility (and therefore reunion with the body politic) could be responsible for the increase.
+It could be due to better information: individuals whose rights were restored were likely far more confident of their eligibility to vote than other formerly incarcerated individuals.
+It could also be an issue of timing: individuals whose rights were restored in May of 2018 but were not discharged from parole until early October had four additional months to register to vote than they would have absent the Executive Order.
+In reality, the turnout boost is likely due to a combination of different factors.
+
+Despite the substantial increase in turnout thanks to Executive Order 181, turnout among formerly incarcerated individuals remained stubbornly low in the 2018 midterm elections.
+Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline.
+Although the successes of Executive Order 181 should be celebrated, more must be done to encourage formerly incarcerated individuals to participate in the political process.
+
 \newpage

### 2019-06-12 markdown/kmorris_thesis.Rmd
write

@@
 link-citations: yes
+header-includes:
+    - \usepackage{setspace}\doublespacing
 
@@
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense at all.
+This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as *Richardson v.
@@
 Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
-In Florida, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the effort had been made in order to prevent Florida from being "niggerized" [@florida_1868].
+In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
+Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868].
 
-The racial imbalance of felony disenfranchisement laws are not confined to the 19th century.
+The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
 Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
@@
 ## Testing for Neighborhood Turnout Effects {-}
+Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
+Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting to felony disenfranchisement.
+More than 77 percent of block groups that lost a voter lost only one voter, while 58 percent of census tracts with a lost voter lost only one.
 
-To test whether neighborhoods with lost voters had systematically lower turnout rates than similar neighborhoods without such disqualifications, I begin by using a matching model.
+```{r distribution-lost, echo = F, message = F, warning = F}
+t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
+  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
+
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
+                  align = "c") %>% 
+  row_spec(11, hline_after = T) %>% 
+  kable_styling(latex_options = c("HOLD_position"))
+j
+```
+
+Because the vast majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
 Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.^[It is worth noting that this definition of lost voter is highly conservative.
@@
           type = "latex",
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) X Share Non-Hispanic Black", "Constant"),
+          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
           dep.var.labels = c("Turnout Rate"),
@@
 Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The model above does not allow for different effects on turnout based on *how many* voters a neighborhood lost.
+The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfrancisement in the most hard-hit communities.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
@@
           se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters X Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
                                "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
@@
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
+The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
 
 The block groups where these depressive effects are not randomly distributed througout the city.
-They are highly spatially concentrated: in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on "Lost Voters * Share Black" from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
+They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
+Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
+The estimated depressive effect is $-0.017\times Lost\ Voters \times Share\ Black$
 
@@
 A dummy variable indicating whether an individual was discharged on or after May 21^st^ is therefore analogous to an intent-to-treat, with a dummy indicating whether one's rights were restored measuring imperfect compliance.
-Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using OLS.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
+Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using an OLS sepcification.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
 
@@
           table.layout = "-dm#-t-as-n",
-          notes = c("Table reports p-values for Wu-Hausman and",
-                    "Weak Instruments tests.",
+          notes = c("Table reports p-values for Weak Instruments",
+                    "and Wu-Hausman tests.",
                     "D(Discharged from Parole on or after May",
-                    "21, 2018) instrument for D(Rights Restored)"))
+                    "21, 2018) instrument for D(Rights Restored)."))
 ```
 
-Table \ref{tab:iv-models} indicates that using D(Discharged from Parole on or After May 21) is likely warranted.
+Table \ref{tab:iv-models} indicates that using discharge date of before or after May 21^st^ as an instrument for rights restoration is likely warranted.
 The p-values of the Wu-Hausman tests hover around 0.05, while the Weak Instruments test is highly significant in each model.

### 2019-06-17 markdown/kmorris_thesis.Rmd
write

@@
 bibliography: "../research/citations.bib"
+biblio-style: chicago
 link-citations: yes
+fontsize: 12pt
 header-includes:
     - \usepackage{setspace}\doublespacing
-
 ---
@@
 \pagenumbering{gobble}
+\pagenumbering{gobble}
 \newpage
@@
 This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
-This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing.
+This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
 As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
@@
 Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges stemming from targeted policing.
+The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges stemming from targeted policing and economic marginalization.
 The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
 A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Burch2011; @Gerber2014].
-Others have looked at the impact of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
+Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
@@
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if these already-disadvantaged communities suffer *de facto* disenfranchisement from felony disenfranchisement, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
+The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among elgibile voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
 
@@
 @Sewell2016; @Clear2008; @Lerman2013].
-It is possible, however, that these effects are even broader than the direct focus on policing has acknowledged.
+It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
 We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
@@
 After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April of 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
+In April, 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
 Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
 In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I build an individual-level logit model to explore whether turnout rates among individuals discharged from parole after Executive Order 181 went into effect cast ballots at higher rates than those discharged previously.
+To test whether the policy was successful, I build individual-level models to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
 
@@
 @Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-They find evidence that, over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
+Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
 Insofar as neighborhoods show distinct preferences for politicians, spatially concentrated depressive effects from felony disenfranchisement could influence election outcomes.
-Previous research on felony disenfranchisement has been helpful for thinking through the potential implications of felony disenfranchisement on explored the potential impact of
 
@@
 An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a bloc, their turnout rates may influence city policy.
-If felony disenfranchisement has a depressive effect on turnout, it is likely that it is responsible for politicians being less responsive to the particular preferences of voters in neighborhoods where voters are removed --- some of the most neglected neighborhoods to begin with.
+Insofar as neighborhoods vote as a bloc, their relative turnout rates may influence city policy.
+If felony disenfranchisement has a depressive effect on turnout, politicians are likely less responsive to the particular preferences of voters in these neighborhoods  --- some of the most neglected neighborhoods to begin with.
 
@@
 Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
-The effect of incarceration on participation is the subject of some debate within the field.
+The causal effect of incarceration on participation is the subject of some debate within the field.
 Individuals who go to prison share many characterstics with lower propensity voters generally.
@@
 In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects onsubsequent participation."
+They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation."
 
@@
 Whether or not incarceration *causes* low turnout, however, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
 
@@
 @Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
+They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
 @Meredith2013, however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to formally inform formerly disenfranchised individuals of their re-instated voting rights.
+Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
 @Meredith2013 finds no effect from notification in the absense of eligibility changes.
 
-@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election.
-They find that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
 "Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome." It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
@@
 ## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.^[Because the criminal justice data does not include individuals' addresses, incarcerated individuals who have never registered to vote cannot be geocoded.] I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
 
@@
 ## Identification of Lost Voters {-}
-In this analysis, I offer a slightly different defintion of "lost voter" than much of the literature.
+In this analysis, I offer a different defintion of "lost voter" than much of the literature.
 Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
@@
 If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-This approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
+This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
 Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
@@
 In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-As discussed above, voting is a highly social act.
-Individuals whose family members and neighbors cast ballots are more likely to cast a ballot.
-Identifying the effects of felony disenfranchisement on individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into the extent to which felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
+Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote [@Gerber2003].
+As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
+Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
 
@@
 ## Testing for Neighborhood Turnout Effects {-}
+```{r table-table, , echo = F, message = F, warning = F}
+
+```
 Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
 Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting to felony disenfranchisement.
-More than 77 percent of block groups that lost a voter lost only one voter, while 58 percent of census tracts with a lost voter lost only one.
+More than two-thirds of block groups that lost a voter lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
 
@@
                   align = "c") %>% 
-  row_spec(11, hline_after = T) %>% 
+  row_spec(15, hline_after = T) %>% 
   kable_styling(latex_options = c("HOLD_position"))
@@
 
-Because the vast majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
+Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
 Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.^[It is worth noting that this definition of lost voter is highly conservative.
 For instance, some individuals who would have voted may have gone to prison after turning 18 but before their first general election.
-These "untreated" neighborhoods are likely home to individuals who would have voted if it were not for disenfranchisement policies.] I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011].
-I match on a series of demographic and political indicators.
+These "untreated" neighborhoods are likely home to individuals who would have voted if it were not for disenfranchisement policies.] I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
 Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
@@
 Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (from the voter file) by the voting age population.
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] These data come from the New York City Board of Elections.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
+Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the voting age population (estimated by the Census Bureau).
+I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
 Matching is done with replacement.
@@
 In [Appendix B], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
-These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population as reported by the Census Bureau.
+These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
 
@@
 When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
+The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
 Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))*100, digits = 1)` percent, this effect is alarmingly high.
@@
 ```
-The results presented in Table \ref{tab:trad-reg} align very closely with the estimated effect from the matching model.
-Once again, lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
+The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
+Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
@@
 Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.017\times Lost\ Voters \times Share\ Black$
+The estimated depressive effect is $-0.017\times Lost\ Voters \times Share\ Black$.
 
@@
 The systematic removal of voters from these neighborhoods is troubling.
-However, more than 850,000 votes were cast in the general election of 2017.
+However, more than 900,000 votes were cast in the general election of 2017.
 Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters --- concentrated though they are --- is likely to have large implications on its own.
 
-Felony disenfranchisement, however, reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement likely has impacts on Black turnout at the state level.
+As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
+Previous literature has established that felony disenfranchisement likely impacts Black turnout at the state level.
 This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
@@
 
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
+These are the individuals who would have been eligible to cast a ballot whether or not the Executive Order was signed.
+This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
+However, by focusing on individuals who would have been eligible to vote either way, we can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
+
 ## Identifying Parolees Whose Rights Were Restored {-}
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.^[Only individuals sentenced to prison for felony offenses lose their voting rights in New York.
-As such, I remove individuals sentenced to prison for misdemeanors.
-Incarceration likely also has a *de facto* disenfranchising effect on turnout among individuals incarcerated for misdemeanors [@White2019], though exploring that effect is beyond the scope of this project.] By using the DIN number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
+By using the identification number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
@@
 It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier in the year.
-The longer that individuals have been off of parole, the more likely they are to cast a ballot.
+However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
+The longer an individual has been off of parole, the more likely he is to cast a ballot.
 For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included in the denominator only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
+Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
 A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
@@
 It does not control for any sort of individual level characterisitcs.
-In \ref{tab:to-18-logit}, I present the results of an individual-level regression exploring the turnout in 2018 among individuals whose rights were restored.
-Because turnout is a binary dependent variable, this is a logistic regression.
-It includes all individuals discharged from parole between January 1^st^, 2012, through October 10^th^, 2018 (the registration deadline in New York State).^[This of course ignores the individuals who were still on parole and cast a ballot in the 2018 election.
-However, for this exercise, I limit the analysis group to individuals who would have been eligible to vote on election day even if Governor Cuomo had not signed Executive Order 181.]
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals whose rights were restored.
+It includes all individuals discharged from parole between January 1^st^, 2012, through October 10^th^, 2018 (the registration deadline in New York State).
 
@@
 Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual had his rights restored before being discharged from parole, and the number of days between the individual's discharge date and November 6^th^, 2018.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county, and race.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
 Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
-In each successive model, the AIC decreases substantially, indicating that including these controls are warranted.
+In each successive model, the AIC decreases substantially, indicating that these controls are warranted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
@@
 Each model also indicates that individuals whose rights were restored prior to discharge were more likely to cast a ballot than those whose rights were not restored.
-Exponentiating the coefficients on D(Rights Restored) indicates that individuals whose rights were restored due to Executive Order 181 were between `r round((exp(model1[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote than those who did not.
+Exponentiating the coefficients on D(Rights Restored) indicates that individuals whose rights were restored due to Executive Order 181 were between `r round((exp(model1[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote than those whose rights were not.
 
@@
 Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also of course correlated with propensity to vote.
-Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the effect of Executive Order 181.
+Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the causal effect of Executive Order 181.
 This calls for an instrumental variables approach.
@@
 The timing of Executive Order 181 serves as a useful instrument for formal rights restoration.
-Although Figure \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is far less signficant in the short run.
+Although Figure \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is nonsignficant in the short run.
 Figure \ref{fig:to-18-recent} shows that, for individuals discharged from parole in 2017 and 2018, there does not appear to be a relationship between time off parole and turnout rates.
@@
 
-Although date of discharge from parole is not correlated with turnout in the short run, it is correlated with likelihood of being having one's rights restored while still on parole.
+Because the date of parole discharge is not correlated with turnout, we can conceptualize the timing of Executive Order 181 as a randomly assigned treatment variable.
+An individual who was released after the EO was signed is part of the intend-to-treat group, and individuals who actually had their rights restored comply with the treatment.
+As with true randomized control trials, compliance (restoration) is correlated with the outcome of interest (turnout) and therefore must be controlled for.
+
 Governor Cuomo signed Executive Order 181 in April of 2018, but an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
@@
 
-Table \ref{tab:iv-models} indicates that using discharge date of before or after May 21^st^ as an instrument for rights restoration is likely warranted.
+Table \ref{tab:iv-models} indicates that using discharge date of before or after May 21^st^ as an instrument for rights restoration is warranted.
 The p-values of the Wu-Hausman tests hover around 0.05, while the Weak Instruments test is highly significant in each model.
@@
 It is not surprising that in a large state like New York, full names and dates of birth fail to uniquely identify individuals.
-However, the rate of false positives is quite low: this analysis indicates that there is a `r percent(as.integer(max(less_35, plus_35)) / as.integer(gsub(",", "", true)), accuracy = 0.01)`.
+However, the rate of false positives is quite low: this analysis indicates that there is a `r percent(as.integer(max(less_35, plus_35)) / as.integer(gsub(",", "", true)), accuracy = 0.01)` rate of false positives.
 It is unlikely that such a small incidence of false positives meaningfully impacts this analysis.

### 2019-06-19 markdown/felony_disenfranchisement_nys.Rmd
rename

--- 2019-06-17 markdown/kmorris_thesis.Rmd
@@
   bookdown::pdf_document2:
-    toc: yes
+    toc: no
     fig_caption: yes
@@
 bibliography: "../research/citations.bib"
-biblio-style: chicago
+biblio-style: apalike
 link-citations: yes
-fontsize: 12pt
 header-includes:
@@
 <!-- --- -->
-
+\pagenumbering{gobble}
+\tableofcontents
+\pagebreak
 ```{r setup, echo = F, message = F, warning = F}
@@
 ```
-\pagenumbering{gobble}
-\pagenumbering{gobble}
-\newpage
 \pagenumbering{arabic}
@@
 
+\pagebreak
+
 # Executive Order 181 and Turnout in 2018 {-}
@@
 
-Adding controls for time in Table \ref{tab:to-18-logit-short} increases the AIC, indicating that the inclusion of these variables is not warranted.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix C] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.]
+The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix C] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] Over the short term, the length of time an individual has been off parole is not associated with his propensity to vote.
 
-Because the date of parole discharge is not correlated with turnout, we can conceptualize the timing of Executive Order 181 as a randomly assigned treatment variable.
-An individual who was released after the EO was signed is part of the intend-to-treat group, and individuals who actually had their rights restored comply with the treatment.
-As with true randomized control trials, compliance (restoration) is correlated with the outcome of interest (turnout) and therefore must be controlled for.
+Because the date of parole discharge is not correlated with turnout, we can conceptualize the timing of Executive Order 181 as a randomly assigned treatment.
+An individual who was released after the Executive Order went into effect is part of the intend-to-treat group, and individuals who actually had their rights restored "comply" with the treatment.
+As with many true randomized control trials, compliance (restoration) is correlated with the outcome of interest (turnout) and therefore must be controlled for.
 

### 2019-06-24 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 
-The political history of the United States has been characterized by a general trend toward universal sufferage.
+The political history of the United States has been characterized by a general, if unlinear, trend toward universal sufferage [see, for instance, @Keyssar2009].
 At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
@@
 The Supreme Court, in cases such as *Richardson v.
-Ramirez*, has upheld states' right to do just that.
+Ramirez* (1974), has upheld states' right to do just that.
 Collectively, these laws disenfranchise as many as `r 6.1 - 1.4` million American citizens.
@@
 In any discussion of felony disenfranchisement in the United States, it is imperative to acknowledge the central role played by race and white supremacy.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were intended to" [@Burch2010].
+As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, p.
+4; emphasis in the original].
 Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
 In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868].
+Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, p.
+374].
 
@@
 They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout." They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, elibible Black voters are less likely to cast a ballot.
+Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout" (p.
+800).
+They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, elibible Black voters are less likely to cast a ballot.
 These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
 As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty."
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
+724).
 
@@
 In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
-Using a standard ordinary least squares regression determined that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods." This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definiton of "lost voters."
+Using a standard ordinary least squares regression, she determined that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (p.
+185).
+This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definiton of "lost voters."
 
@@
 Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness." @Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not." If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
+"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (p.
+59).
+@Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (p.
+1206).
+If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
 
@@
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
+It should be no suprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation."
+They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (p.
+1144).
 
@@
 The findings in @White2019 agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual."
+@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (p.
+829).
 
@@
 @Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome." It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
+"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (p.
+924).
+It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
 

### 2019-06-24 markdown/part2_standalone.Rmd
breakout separate paper

--- 2019-06-24 markdown/felony_disenfranchisement_nys.Rmd
@@
 
-The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
-This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
-
-Due to economic and racial segregation, these effects are highly spatially concentrated.
-Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
-Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
-Concentrated policing has also credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
-
-Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges stemming from targeted policing and economic marginalization.
-The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Burch2011; @Gerber2014].
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
-With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-This paper seeks to extend the extant literature on felony disenfranchisement by considering the implications of disenfranchisement on local political participation.
-I begin by looking at participation rates in New York City neighborhoods where individuals with a demonstrated history of participation have been disenfranchised.
-This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among elgibile voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
-
-Undoing the negative effects of felony disenfranchisement on individual and neighborhood representation requires crafting policy aimed at increasing turnout among the formerly disenfranchised.
-New York State has taken a step in that direction: in April of 2018, Governor Andrew Cuomo signed Executive Order 181, which came close to ending disenfranchisement for New Yorkers on parole.
-After examining the neighborhood effects of felony disenfranchisement, I test whether having one's rights restored while still on parole increased turnout among formerly disenfranchised individuals in the 2018 midterm elections.
-
-# Background of Felony Disenfranchisement in the United States {-}
-
 In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
@@
 
-In any discussion of felony disenfranchisement in the United States, it is imperative to acknowledge the central role played by race and white supremacy.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, p.
-4; emphasis in the original].
-Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
-In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, p.
-374].
+The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
+As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
+This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
+This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
 
-The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
-Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
-In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just `r percent(29932674 / 247219588, accuracy = 0.1)` of the voting age population in 2016, they accounted for `r percent(2228118 / 6106327, accuracy = 0.1)` of the disenfranchised population [@sentencing_2016].
+Although felony disenfranchisement is highly disempowering for the individuals impacted by the policy, the political consequences of these laws is not entirely clear.
+The number of incarcerated individuals is relatively low compared to the number of voters.
+In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
+Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those formally disenfranchised.
+Previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
 
-Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January, 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November, 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate [@Morris2019].
+These spillover effects have been identified using a wide variety of empirical methods.
+@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans, finding that "African American disenfranchisement plays a unique role in predicting African American voter turnout" (p.
+800).
+@Ochs2006, @Bowers2009, and @Walker2014 similarly exploit state-level differences to estimate the (de)mobilizing effect of felony disenfranchise on eligible voters.
+As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
+724).
+@Burch2013 used neighborhood-level data to further investigate how felony disenfranchisement operates the sub-state level.
+She finds that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (p.
+185).
 
-# Academic Literature and Felony Disenfranchisement {-}
+It is clear that although felony disenfranchisement directly impacts relatively few potential voters, its demobilizing character reaches many citizens who are not formally disenfranchised.
+Considering the characteristics of the neighborhoods that incarcerated individuals call home --- neighborhoods that are among the most marginalized --- felony disenfranchisement is a large problem.
+
+In this paper, I examine the effect of Executive Order 181 in New York State.
+Signed in April of 2018, the Executive Order restored voting rights to many individuals on parole.
+Prior to Executive Order 181, individuals were disenfranchised while incarcerated or on parole.
+Most individuals who were on parole on Election Day in 2018 were allowed to vote, a major step forward for democracy in the Empire State.
+In this paper, however, I focus on individuals who were discharged from parole prior to Election Day.
+These individuals, therefore, would not have been directly disenfranchised by felony disenfranchisement policies even in the absense of the Executive Order.
+I hypothesize that Executive Order 181 increased turnout among this population, even though their eligibility was not directly impacted by the policy change.
+
+# Turnout Among the Formerly Disenfranchised {-}
 
@@
 They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as those who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after the Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
-Others have used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+Uggen and Manza demonstrated that felony disenfranchisement can have material political consequences.
 
+In subsequent years, new research called into question Uggen and Manza's assumption that disenfranchised individuals would vote at rates comparable to their sociodemographic peers.
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
@@
 
-A number of papers have also explored the impact that felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to determine individual-level turnout.
-They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout" (p.
-800).
-They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, elibible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
-As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
-724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the impact these laws might have at the local level.
-@Burch2013 is an exception to this.
-In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
-Using a standard ordinary least squares regression, she determined that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (p.
-185).
-This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definiton of "lost voters."
-
-# What This Paper Examines {-}
-
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017.
-Policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods [e.g.
-@Sewell2016; @Clear2008; @Lerman2013].
-It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representatition of certain parts of the city.
-
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April, 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I build individual-level models to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
-
-The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather thank think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
-This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
-The second part of this project seeks to do just that.
-
-## Framing for Turnout Effects {-}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
-@King2016; @Bowers2009].
-With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-The degree to which these effects are concentrated in particular neighborhoods has important implications for our understanding of the problem of felony disenfranchisement.
-Although literature examining the degree to which geographic polarization effects voting patterns is scarce, there is some evidence to suggest that neighborhoods do vote similarly.
-@Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-Insofar as neighborhoods show distinct preferences for politicians, spatially concentrated depressive effects from felony disenfranchisement could influence election outcomes.
-
-Even if neighborhoods to not have unique preferences for presidential candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level, for instance, representatives direct resources to areas within their districts that provide the greatest political benefit [@Martin2003].
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (p.
-59).
-@Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (p.
-1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
-
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a bloc, their relative turnout rates may influence city policy.
-If felony disenfranchisement has a depressive effect on turnout, politicians are likely less responsive to the particular preferences of voters in these neighborhoods  --- some of the most neglected neighborhoods to begin with.
-
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-
-This *de facto* disenfranchisement is likely to be concentrated within the neighborhoods home to formally disenfranchised residents.
-Voting is a social act, and social networks play an important role in predicting political participation [e.g.
-@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no suprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-
-## Turnout Among Formerly Disenfranchised Individuals {-}
-
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
 The causal effect of incarceration on participation is the subject of some debate within the field.
@@
 829).
-
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts dire consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout, however, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
 
@@
 
-Secondly, Executive Order 181 is expected to dispel confusion about eligibilty and serve as a reminder to vote.
-As @Drucker2005 and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
-If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.
-Even if confusion does not lead to lower turnout, there is some reason to believe that reminding formerly incarcerated individuals of their rights is a successful intervention for boosting turnout [@Meredith2015; @Gerber2014].
+Secondly, Executive Order 181 is expected to dispel confusion about eligibilty.
+Since @Downs1957, political scientists have argued that an individual's propensity to vote is decided at least in part by comparing the expected costs of casting a ballot with the expected benefits of doing so.
+The costs of voting while disenfranchised are exceedingly high, as high-profile cases in states such as Texas have made clear in recent years [@Flynn2018].
+If an individual assessment of the cost of voting is determined in part by the product of their uncertainty about their eligibility and the cost of casting a ballot while ineligible, formerly disenfranchised individals are unlikely to participate unless they are fully certain of their eligibility.
+As @Drucker2005 and others have detailed, however, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
+By dispelling uncertainty around eligibility, I hypothesize that Executive Order 181 increased turnout among the formerly disenfranchised.
+
+# Research Design {-}
+
+Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
+For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
+New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
+
+On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+While the Executive Order is obviously beneficial for individuals who are on parole on Election Day (and therefore would have been disenfranchised but for the policy change), the Executive Order may have benefits even for the individuals whose eligibility was not directly impacted.
+Under the Executive Order, parole officers were required to inform their parolees of their new status under the law.
+In the analysis that follows, I look only at individuals who were discharged before October 10^th^, 2018 (the registration deadline in New York State).
+Although Executive Order 181 did not change the eligibility of these individuals to cast a ballot in November, there is reason to believe that it might still have impacted their turnout propensity.
+
+In order to estimate the impact of Executive Order 181 on turnout, I begin by exploring whether the turnout rate among individuals whose rights were restored prior to discharge from parole systematically differed from those whose rights were not restored.
+This analysis uses an individual-level logistic regression that estimates individuals' turnout rates and controls for various individual-level characteristics.
+
+However, not all individuals who were discharged after the Executive Order was signed had their rights restored.
+Individuals who were arrested or were not citizens were not granted voting rights.
+Because there is reason to believe that voting propensity is positively correlated with the likelihood that an individual's rights were restored, I employ an instrumental variables approach to more precisely estimate the impact of the executive order on turnout.
 
@@
 
-The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
+The criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
@@
 However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the state, I constructed a list of the individuals who were on parole and had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
+By using identification numbers from the parolee data obtained from the state, I constructed a list of the individuals who were on parole and had their rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
+For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
 
@@
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-The first section of this paper uses a snapshot of the registered voter file from April 30^th^, 2018.
-The second section of this paper uses a snapshot from March 3^rd^, 2019.
+I use a snapshot of the registered voter file from March 3^rd^, 2019.
 
-## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+## Matching {-}
+Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
+I match individuals in each dataset using first name, middle name, last name, and date of birth.
+Of course, matching using administrative data is not a perfect science.
+Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
+Following @Meredith2013, I discuss the potential impact of these errors in [Appendix A].
+As Appendix A makes clear, however, there is little reason to believe that false matches systematically impact the results of the analyses that follow.
 
-# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels {-}
-
-## Identification of Lost Voters {-}
-In this analysis, I offer a different defintion of "lost voter" than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
-Such an approach is informative for understanding the impact of disenfranchisement.
-Many young men, for instance, are admitted to prison each year.
-If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
-Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
-
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration.
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
-At the local level where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition provides insight only into the effect of incarceration on turnout, not felony disenfranchisement.
-
-Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote [@Gerber2003].
-As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
-
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
-
-These individuals are identified by matching NYSDOCCS data to the voter file.
-I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I discuss the potential impact of these errors in [Appendix A].] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
-
-```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
-j <- readRDS("./output/city_map.RDS")
-print(j)
-```
-The spatial concentration of lost is readily apparent.
-In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
-
-## Testing for Neighborhood Turnout Effects {-}
-```{r table-table, , echo = F, message = F, warning = F}
-
-```
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting to felony disenfranchisement.
-More than two-thirds of block groups that lost a voter lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
-
-```{r distribution-lost, echo = F, message = F, warning = F}
-t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
-  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
-
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
-                  align = "c") %>% 
-  row_spec(15, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"))
-j
-```
-
-Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.^[It is worth noting that this definition of lost voter is highly conservative.
-For instance, some individuals who would have voted may have gone to prison after turning 18 but before their first general election.
-These "untreated" neighborhoods are likely home to individuals who would have voted if it were not for disenfranchisement policies.] I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the voting age population (estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
-Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
-
-### Match Output {-}
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-
-```{r tract-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_tract.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:tract-table) Results of Tract-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less white, have much lower median incomes, and a larger share of voters are registered as Democrats.
-
-After matching neighborhoods, I use a simple regression to test whether turnout in the 2017 mayoral election was different in areas with lost voters.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.^[The New York registered voter file does not align exactly with results reported by the city.
-The voter file indicates that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that `r comma(readRDS("./temp/mayoral_ballots.rds"))` votes were cast.
-In [Appendix B], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
-These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Each neighborhood is weighted by its adult population, and robust standard errors are clustered at the level of the match [@Abadie2019].
-```{r tract-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/match_reg_block_group.rdata")
-
-load("./temp/match_reg_block_group_nhb.rdata")
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-bg_reg_black <- reg_output_nhblack
-bg_reg_black_ses <- reg_output_nhblack_ses
-
-load("./temp/match_reg_tract.rdata")
-
-load("./temp/match_reg_tract_nhb.rdata")
-tract_reg <- reg_output
-tract_reg_ses <- reg_output_ses
-
-tract_reg_black <- reg_output_nhblack
-tract_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
-          header = F,
-          se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
-          notes = c("Robust standard errors (clustered at", "match level) in parentheses."),
-          type = "latex",
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:match-reg) Matching Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"))
-```
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average `r -1*round(bg_reg[["coefficients"]][["treatTRUE"]]*100, digits = 1)` percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))*100, digits = 1)` percent, this effect is alarmingly high.
-For every 100 votes cast in a predominantly Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
-Although these variables are largely nonsignificant at the census tract level, this is not particularly surprising.
-If lost voters have the largest depressive effect on family members and close neighbors, it makes sense that the effect of losing a voter is locally concentrated.
-
-### Testing Intensity Effects {-}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfrancisement in the most hard-hit communities.
-In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
-Each neighborhood is weighted by its voting age population, and robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purporses.]
-
-```{r trad-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/tract_model_reg_ols.rdata")
-load("./temp/bg_model_reg_ols.rdata")
-
-stargazer(bg_model, bg_model2, tract_model, tract_model2,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
-          header = F,
-          type = "latex",
-          se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:trad-reg) Standard Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = c("Robust standard errors (clustered by city",  "council district) in parentheses."),
-          omit = "boro")
-```
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
-The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
-
-The block groups where these depressive effects are not randomly distributed througout the city.
-They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.017\times Lost\ Voters \times Share\ Black$.
-
-```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/dec_block_map.rds")
-print(j)
-```
-
-## Discussion {-}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` from New York City neighborhoods.
-The spatial concentration of these lost voters is striking: as Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} make clear, these voters were removed from neighborhoods with significantly lower incomes than the rest of the city.
-They were also removed from neighborhoods that were far less white than the average neighborhood.
-The systematic removal of voters from these neighborhoods is troubling.
-However, more than 900,000 votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters --- concentrated though they are --- is likely to have large implications on its own.
-
-As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement likely impacts Black turnout at the state level.
-This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-These communities are already home to some of the poorest and most marginalized voters in the city; they are home to residents who have the least ability to influence policy through other means such as large campaign contributions.
-The fact that felony disenfranchisement policies appear to be disincentivizing participation on election day --- that these policies are further weakening the ability of these neighborhoods to participate in the democratic process --- is cause for alarm.
-
-\pagebreak
-
-# Executive Order 181 and Turnout in 2018 {-}
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].^[I've also talked with folks on probation whose probation officers told them they were disenfranchised.
-Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
-As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
-To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the policy change may increase the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-Under the Executive Order, however, parole officers were required to explicitly inform their supervisees of their newly restored voting rights.
-It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly re-eligible voters to cast a ballot.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
-These are the individuals who would have been eligible to cast a ballot whether or not the Executive Order was signed.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, we can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
-
-## Identifying Parolees Whose Rights Were Restored {-}
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
+# Results {-}
 
@@
 Figure \ref{fig:to-18} displays the share of individuals who were discharged from parole in each month and cast a ballot in the 2018 elections.
-It does not control for any sort of individual level characterisitcs.
+It does not control for any individual level characteristics.
 In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals whose rights were restored.
@@
 <div id="refs"></div>
-
-\newpage
-# Appendix A {-}
-
-Matching across administrative data sets is not a perfect science.
-Despite having high-quality low-level data from both New York State Department of Corrections and Community Supervision (DOCCS) and the New York State Board of Elections, I cannot be sure that I am correctly identifying formerly incarcerated, registered voters.
-The results of the analyses in Table \ref{tab:match-reg} give some reason to believe that there are few false negatives.
-If there were many false negatives (lost voters who are not identified in this analysis), "treated" neighborhoods would be incorrectly categorized as "untreated" neighborhoods.
-If this were a widespread problem, we would expect to see little difference between treated and control neighborhoods.
-To the extent that lost voters go unidentified by this analysis, it is likely that the results presented in Tables \ref{tab:match-reg} and \ref{tab:trad-reg} are conservative by including treated neighborhoods in the control group.
-Directly testing for false negatives, however, is difficult.
-If a voter changes her name before being incarcerated, for instance, our analysis will fail to capture her.
-
-Testing for the presence of false positives is slightly easier.
-One way to do so is by constructing false records [@Meredith2013].
-By slightly changing the birthdates in one set of administrative data, I can estimate how frequently DOCCS records inaccurately match with voter records.
-To do so, I shift the dates of birth in the DOCCS data by 35 days.
-I then re-merge the DOCCS records (with "false" birth dates) against the registered voter file.
-
-Table \ref{tab:change-dobs} shows the results of these permutations.
-The first row shows the matches with the raw data.
-The second row shows the results of adding 35 days to the date of birth in each DOCCS record, and the final row shows the result of subtracting 35 days from these records.
-
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-Table \ref{tab:change-dobs} indicates that there are likely some false positives in our matches.
-It is not surprising that in a large state like New York, full names and dates of birth fail to uniquely identify individuals.
-However, the rate of false positives is quite low: this analysis indicates that there is a `r percent(as.integer(max(less_35, plus_35)) / as.integer(gsub(",", "", true)), accuracy = 0.01)` rate of false positives.
-It is unlikely that such a small incidence of false positives meaningfully impacts this analysis.
-
-\newpage
-# Appendix B {-}
-
-As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
-According to the official results, `r comma(readRDS("./temp/mayoral_ballots.rds"))` ballots were cast in the 2017 mayoral election.
-The registered voter file reports that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot.
-This is not necessarily evidence of poor management by the Board of Elections.
-Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).^[This is not to say that there is *not* evidence of mismanagement by the BOE.
-For instance, the author is incorrectly marked as not participating in the 2017 mayoral primary.] Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
-If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
-
-To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level.
-The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level.
-Precincts are assigned the demographics of the block group in which they are situated.
-Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group.
-The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.
-Robust standard errors are clustered at the assembly district level.
-
-```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/compare_ballots_off_results.RData")
-
-stargazer(ed_reg,
-          header = F,
-          type = "latex",
-          se = list(ed_comp_ses),
-          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Undercount Rate"),
-          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = c("Robust standard errors (clustered by", "assembly district) in parentheses."),
-          omit = "boro")
-```
-
-As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate.
-Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
-
-\newpage
-# Appendix C {-}
-
-In the [Instrumenting Rights Restoration] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 21^st^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
-load("./temp/individual_turnout_16.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "(\\#tab:to-16-logit) Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot (the p-value on the coefficient of interest exceeds 0.5 in each model).
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.

### 2019-07-02 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 bibliography: "../research/citations.bib"
-biblio-style: apalike
 link-citations: yes

### 2019-07-02 markdown/part2_standalone.Rmd
write

@@
 Most individuals who were on parole on Election Day in 2018 were allowed to vote, a major step forward for democracy in the Empire State.
-In this paper, however, I focus on individuals who were discharged from parole prior to Election Day.
-These individuals, therefore, would not have been directly disenfranchised by felony disenfranchisement policies even in the absense of the Executive Order.
+In this paper, however, I focus on individuals who were discharged from parole prior to Election Day: individuals who would not have been directly disenfranchised by felony disenfranchisement policies even in the absense of the Executive Order.
 I hypothesize that Executive Order 181 increased turnout among this population, even though their eligibility was not directly impacted by the policy change.
@@
 Since @Downs1957, political scientists have argued that an individual's propensity to vote is decided at least in part by comparing the expected costs of casting a ballot with the expected benefits of doing so.
-The costs of voting while disenfranchised are exceedingly high, as high-profile cases in states such as Texas have made clear in recent years [@Flynn2018].
+The costs of voting while disenfranchised (intentionally or otherwise) are exceedingly high, as high-profile cases in states such as Texas have made clear in recent years [@Flynn2018].
 If an individual assessment of the cost of voting is determined in part by the product of their uncertainty about their eligibility and the cost of casting a ballot while ineligible, formerly disenfranchised individals are unlikely to participate unless they are fully certain of their eligibility.
@@
 In the analysis that follows, I look only at individuals who were discharged before October 10^th^, 2018 (the registration deadline in New York State).
-Although Executive Order 181 did not change the eligibility of these individuals to cast a ballot in November, there is reason to believe that it might still have impacted their turnout propensity.
+Limiting the analysis to individuals whose eligibility to cast a ballot was not directly impacted by the policy change allows me to explore whether having one's rights restored while still under formal supervision is effective at raising that individual's propensity to vote even once supervision is ended.
 
-In order to estimate the impact of Executive Order 181 on turnout, I begin by exploring whether the turnout rate among individuals whose rights were restored prior to discharge from parole systematically differed from those whose rights were not restored.
+In order to estimate the impact of Executive Order 181 on turnout, I begin by exploring whether the turnout rate among individuals discharged from parole after the policy went into effect systematically differed from those discharged earlier.
 This analysis uses an individual-level logistic regression that estimates individuals' turnout rates and controls for various individual-level characteristics.
@@
 In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
+It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
@@
 I match individuals in each dataset using first name, middle name, last name, and date of birth.
-Of course, matching using administrative data is not a perfect science.
-Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I discuss the potential impact of these errors in [Appendix A].
-As Appendix A makes clear, however, there is little reason to believe that false matches systematically impact the results of the analyses that follow.
+To be considered a "match," records must have the exact same birth date.
+The first and last names must also be exact matches (conditional on the adjustments discussed below).
+The middle names must meet one of the following conditions in order to qualify:
+
+* Middle names are identical. If neither set of records includes a middle name, this condition is met.
+* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
+* A middle name or middle initial in one set of records, and a missing middle name in the other set.
+
+Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andew Doe" and "John Anthony Doe" would not.
+
+There are two types of potential error in this methdology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
+
+Testing for the presence of false positive matches is fairly straightforward.
+@Meredith2013 offers one way to test their prevalence using placebo matching.
+I slightly alter the date of birth reported in the parole discharge dataset to create false records.
+Comparing the number of matches between these "fake" discharge records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
+This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
+true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
+plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+Testing for false negatives is more challenging.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
+To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
+A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
+Such standardizations, however, will miss individuals who change their names entirely.
+For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
 
 # Results {-}
+
+Below, I employ a logistic regression to investigate the impact of Executive Order 181 on turnout among all parolees discharged after the policy went into effect.
+Because not all individuals had their rights restored, I follow this analysis with an instrumental-variables approach to determine the effect of rights restoration on individuals whose rights were in fact formally restored prior to discharge from parole.
 
@@
 Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy.
-In the section that follows, I develop a logit model to explore whether, after controlling for the available characteristics, individuals whose rights were restored prior to being discharged were more likely to cast a ballot in the 2018 midterm elections.
+However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that Executive Order increased turnout.
 
@@
 
-Figure \ref{fig:to-18} displays the share of individuals who were discharged from parole in each month and cast a ballot in the 2018 elections.
-It does not control for any individual level characteristics.
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals whose rights were restored.
-It includes all individuals discharged from parole between January 1^st^, 2012, through October 10^th^, 2018 (the registration deadline in New York State).
-
-```{r logit-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/individual_turnout_18.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Rights Restored)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual had his rights restored before being discharged from parole, and the number of days between the individual's discharge date and November 6^th^, 2018.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
-In each successive model, the AIC decreases substantially, indicating that these controls are warranted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Each model also indicates that individuals whose rights were restored prior to discharge were more likely to cast a ballot than those whose rights were not restored.
-Exponentiating the coefficients on D(Rights Restored) indicates that individuals whose rights were restored due to Executive Order 181 were between `r round((exp(model1[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote than those whose rights were not.
-
-Table \ref{tab:to-18-logit} provides some indication that Executive Order 181 was successful, but does not establish causality.
-There is reason to believe that individuals who had their rights restored were more likely to vote even had they not been "treated" by EO 181.
-In the section that follows, I employ an instrumental variables approach to control for this possibility.
-
-## Instrumenting Rights Restoration {-}
-
-In the previous section, I explored whether individuals who had their rights formally restored prior to parole discharge were more likely to vote in the 2018 midterm election.
-There is reason to believe that whether an individual received the treatment (had their rights restored) is correlated with their propensity to vote.
-For instance, parolees who were arrested while on parole did not have their rights restored; a recent arrest is also likely to impact turnout.
-Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also of course correlated with propensity to vote.
-Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the causal effect of Executive Order 181.
-This calls for an instrumental variables approach.
-
-The timing of Executive Order 181 serves as a useful instrument for formal rights restoration.
-Although Figure \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is nonsignficant in the short run.
-Figure \ref{fig:to-18-recent} shows that, for individuals discharged from parole in 2017 and 2018, there does not appear to be a relationship between time off parole and turnout rates.
+Although Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that the length of time an individual has been off parole is correlated with her propensity to vote, that relationship is far weaker in the short term.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
 
@@
 ```
-
 Formalizing this chart into an individual-level logit model demonstrates that time since discharge is not correlated with turnout in the short run.
@@
 The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix C] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] Over the short term, the length of time an individual has been off parole is not associated with his propensity to vote.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] Over the short term, the length of time an individual has been off parole is not associated with his propensity to vote.
 
-Because the date of parole discharge is not correlated with turnout, we can conceptualize the timing of Executive Order 181 as a randomly assigned treatment.
-An individual who was released after the Executive Order went into effect is part of the intend-to-treat group, and individuals who actually had their rights restored "comply" with the treatment.
-As with many true randomized control trials, compliance (restoration) is correlated with the outcome of interest (turnout) and therefore must be controlled for.
+This lack of correlation between discharge date and turnout indicates that we can conceptualize Executive Order as a natural experiment.
+We can expect that, in the absence of Executive Order 181, individuals discharged from parole before and after the policy went into effect would have turned out at comparable rates.
+Any observed difference in turnout rates, therefore, can be attributed to the implementation of the policy.
 
-Governor Cuomo signed Executive Order 181 in April of 2018, but an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-As Figure \ref{fig:restoration-chart} demonstrates, individuals who finished their parole supervision in early May did not have their voting rights restored before they finished their term of supervision.
-It is not until May 21^st^ that the majority of parolees had their voting rights restored prior to their discharge.
+Although Governor Cuomo signed the Executive Order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
+Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
+It is not until May 21^st^ that the majority of discharged parolees had their voting rights restored prior to their discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 ```
-There is no evidence that date of discharged influences propensity to vote through channels other than formal rights restoration in the short run.
-A dummy variable indicating whether an individual was discharged on or after May 21^st^ is therefore analogous to an intent-to-treat, with a dummy indicating whether one's rights were restored measuring imperfect compliance.
+
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals who were discharged on or after May 2121^st^, 2018, the date when the policy went into full effect.
+It includes all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
+
+```{r logit-reg-chunk, echo = F, results = "asis", message = F}
+load("./temp/individual_turnout_18_itt.rdata")
+
+stargazer(model1, model2, model3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
+          table.placement = "H",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+
+Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
+Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
+In each successive model, the AIC decreases substantially, indicating that these controls are warranted.
+Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+
+Each model also indicates that individuals discharged from parole after the Executive Order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+
+Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
+The above analysis, however, likely understates the impact of the Executive Order on individuals who were in fact eligible to cast a ballot.
+As Figure \ref{fig:restoration-chart} indicates, not all individuals discharged after the Executive Order went into effect had their rights restored.
+Individuals who were discharged after the policy was changed but but did not have their rights restored can be conceptualized as "non-compliers" --- although they were "assigned" to the treatment group (were discharged after the policy changed), they did not receive the treatment (have their rights formally restored).
+Some of these non-compliers were likely non-citizens, who could not be granted voting rights.
+Others may have been re-arrested, which disqualified them from having rights restored.
+Both non-citizens and individuals who have been recently arrested are unlikely to vote.
+"Compliance," therefore, is almost certainly correlated with with an individual's likelihood of participating in the 2018 election.
+As such, the estimated treatment effect on the treated --- on individuals who were likely eligible to cast a ballot on election day --- requires an instrumental variables approach.
+
+In the models below, a dummy indicating whether an individual was discharged from parole after the Executive Order went into effect (i.e., was in the "intend-to-treat" group) is used as an instrument for whether an individual's rights were actually restored (i.e., "complied" with the treatment).
 Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using an OLS sepcification.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
@@
 
-Table \ref{tab:iv-models} indicates that using discharge date of before or after May 21^st^ as an instrument for rights restoration is warranted.
-The p-values of the Wu-Hausman tests hover around 0.05, while the Weak Instruments test is highly significant in each model.
+Table \ref{tab:iv-models} indicates that using discharge date of on or after May 21^st^ as an instrument for rights restoration is warranted.
+The p-values of the Wu-Hausman tests hover just below 0.05, while the Weak Instruments test is highly significant in each model.
 
@@
 Approximately `r round(readRDS("./temp/to_restored.rds") * 100, digits = 2)` percent of individuals who had their rights restored cast a ballot in 2018, indicating that Executive Order 181 increased the propensity to vote by around `r round(summary(iv3)[["coefficients"]][2,1] / (readRDS("./temp/to_restored.rds") - summary(iv3)[["coefficients"]][2,1]) * 100, digits = 1)` percent.
-This estimate is much lower than the estimates reported in Table \ref{tab:to-18-logit}, but that is not surprising; as discussed above, individuals whose rights were restored were more likely to vote whether or not they received restoration before discharge.
-The `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage point increase reported in Table \ref{tab:iv-models} accounts only for the extent to which rights restoration prior to discharge from parole caused higher turnout.
+Unsurprisingly, this estimate is higher than the estimates reported in Table \ref{tab:to-18-logit}.
 
@@
 The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
-The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 40 percent.
+The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 35 percent, and raised the turnout of formerly disenfranchised individuals as a whole by
 
@@
 <div id="refs"></div>
+
+# Appendix A {-}
+
+In the [Individual-Level Turnout Regressions] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 21^st^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
+```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
+load("./temp/individual_turnout_16.rdata")
+
+stargazer(model1, model2, model3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2016 Election"),
+          title = "(\\#tab:to-16-logit) Individual-Level Logit Model",
+          table.placement = "H",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot (the p-value on the coefficient of interest exceeds 0.5 in each model).
+The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.

### 2019-07-05 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 
-Figure \ref{fig:to-18} displays the share of individuals who were discharged from parole in each month and cast a ballot in the 2018 elections.
-It does not control for any sort of individual level characterisitcs.
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals whose rights were restored.
-It includes all individuals discharged from parole between January 1^st^, 2012, through October 10^th^, 2018 (the registration deadline in New York State).
+When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
+The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
+For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the Executive Order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
+Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
 
-```{r logit-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/individual_turnout_18.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Rights Restored)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18} by controlling only for whether an individual had his rights restored before being discharged from parole, and the number of days between the individual's discharge date and November 6^th^, 2018.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
-In each successive model, the AIC decreases substantially, indicating that these controls are warranted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Each model also indicates that individuals whose rights were restored prior to discharge were more likely to cast a ballot than those whose rights were not restored.
-Exponentiating the coefficients on D(Rights Restored) indicates that individuals whose rights were restored due to Executive Order 181 were between `r round((exp(model1[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["restoredTRUE"]]) - 1) * 100, digits = 1)` percent more likely to vote than those whose rights were not.
-
-Table \ref{tab:to-18-logit} provides some indication that Executive Order 181 was successful, but does not establish causality.
-There is reason to believe that individuals who had their rights restored were more likely to vote even had they not been "treated" by EO 181.
-In the section that follows, I employ an instrumental variables approach to control for this possibility.
-
-## Instrumenting Rights Restoration {-}
-
-In the previous section, I explored whether individuals who had their rights formally restored prior to parole discharge were more likely to vote in the 2018 midterm election.
-There is reason to believe that whether an individual received the treatment (had their rights restored) is correlated with their propensity to vote.
-For instance, parolees who were arrested while on parole did not have their rights restored; a recent arrest is also likely to impact turnout.
-Similarly, noncitizens did not have voting rights restored / granted; such citizenship status is also of course correlated with propensity to vote.
-Because the treatment was not randomly assigned, direct comparison of turnout rates among individuals who did and did not have their rights restored is not probitive to the causal effect of Executive Order 181.
-This calls for an instrumental variables approach.
-
-The timing of Executive Order 181 serves as a useful instrument for formal rights restoration.
-Although Figure \ref{fig:to-18} shows that there is a relationship between how long an individual has been off parole and their propensity to vote over the long run, this relationship is nonsignficant in the short run.
-Figure \ref{fig:to-18-recent} shows that, for individuals discharged from parole in 2017 and 2018, there does not appear to be a relationship between time off parole and turnout rates.
+However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
 
@@
           covariate.labels = c("Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)",
                                "Time Spent on Parole (Years)"),
@@
 The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix C] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] Over the short term, the length of time an individual has been off parole is not associated with his propensity to vote.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
 
-Because the date of parole discharge is not correlated with turnout, we can conceptualize the timing of Executive Order 181 as a randomly assigned treatment.
-An individual who was released after the Executive Order went into effect is part of the intend-to-treat group, and individuals who actually had their rights restored "comply" with the treatment.
-As with many true randomized control trials, compliance (restoration) is correlated with the outcome of interest (turnout) and therefore must be controlled for.
-
-Governor Cuomo signed Executive Order 181 in April of 2018, but an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-As Figure \ref{fig:restoration-chart} demonstrates, individuals who finished their parole supervision in early May did not have their voting rights restored before they finished their term of supervision.
-It is not until May 21^st^ that the majority of parolees had their voting rights restored prior to their discharge.
+Although Governor Cuomo signed the Executive Order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
+Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
+It is not until May 21^st^ that the majority of discharged parolees had their voting rights restored prior to their discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 ```
-There is no evidence that date of discharged influences propensity to vote through channels other than formal rights restoration in the short run.
-A dummy variable indicating whether an individual was discharged on or after May 21^st^ is therefore analogous to an intent-to-treat, with a dummy indicating whether one's rights were restored measuring imperfect compliance.
-Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using an OLS sepcification.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
+
+Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
+Whether an individual was discharged from parole before or after May 21^st^, 2018, is akin to a randomly assigned treatment.
+Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
+
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 21^st^, 2018, turned out at higher rates than those discharged earlier.
+The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
+
+```{r logit-reg-chunk, echo = F, results = "asis", message = F}
+load("./temp/individual_turnout_18_itt.rdata")
+
+stargazer(model1, model2, model3,
+          header = F,
+          type = "latex",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+                               "Time Spent on Parole (Years)"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
+          table.placement = "H",
+          omit = c("county", "race", "felony"),
+          add.lines=list(c("County Fixed Effects", "", "X", "X"),
+                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
+                         c("Felony Class Fixed Effects", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n")
+```
+
+Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
+Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
+Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+
+Each model also indicates that individuals discharged from parole after the Executive Order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+
+## Instrumental Variables Approach {-}
+
+Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
+Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
+To answer that question, we must specifically control for whether an individual did have his rights restored before he was discharged from parole.
+
+An individual's propensity to vote can be expressed using the following equation:
+$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + \epsilon_{i} $$
+Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
+Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
+Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote.
+Neither of these assumptions are valid: an individual's eligibility to have her rights restored is almost certainly correlated with whether or not they actually were restored.
+Moreover, eligiblity for rights restoration is likely correlated with propensity to vote.
+Non-citizens, for instance, were not eligible for rights restoration under EO181 because non-citizens cannot vote in New York State.
+Similarly, individuals who were re-arrested while on parole were not eligible to have their rights restored.
+Citizenship status and recency of arrest are both highly correlated with propensity to vote.
+
+Our variable of interest, therefore, is likely correlated with our regression error.
+Such a correlation inhibits our ability to simply regress turnout on a dummy indicating whether an individual's rights were restored.
+The standard way of dealing with such a problem is an instrumental variables approach.
+Such an approach has been widely used in the context of voter turnout [e.g.
+@Gerber2000; @Green2012].
+A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
+
+In this case, such a variable is readily at hand.
+The likelihood that an individual had his voting rights restored is a function of whether he was discharged from parole after Executive Order 181 took effect.
+As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
+A dummy variable indicating whether an individual was discharged after the Executive Order, therefore, satisfies the criteria for an instrumental variable.
+
+Table \ref{tab:iv-models} presents the two-stage least-squares regression estimates exploring the turnout impact of rights restoration.
+As discussed above, the instrumental variable used in these regressions measure whether an individual was discharged from parole after Executive Order 181 went into effect.
+These models are estimated using an OLS sepcification.
+Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.
 
@@
           covariate.labels = c("D(Rights Restored)", 
-                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "D(Male)", "Age (Years)",
                                "Time Spent on Parole (Years)"),
@@
                     "and Wu-Hausman tests.",
-                    "D(Discharged from Parole on or after May",
-                    "21, 2018) instrument for D(Rights Restored)."))
+                    "First-stage equations use dummy indicating",
+                    "whether individual was discharged after the",
+                    "Exectutive Order went into effect. The",
+                    "first-stage equations also include",
+                    "covariates in models 2 and 3."))
 ```
 
-Table \ref{tab:iv-models} indicates that using discharge date of before or after May 21^st^ as an instrument for rights restoration is warranted.
-The p-values of the Wu-Hausman tests hover around 0.05, while the Weak Instruments test is highly significant in each model.
+Table \ref{tab:iv-models} indicates that using discharge date of on or after May 21^st^ as an instrument for rights restoration is warranted.
+The p-values of the Wu-Hausman tests hover just below 0.05, while the Weak Instruments test is highly significant in each model.
 
@@
 Approximately `r round(readRDS("./temp/to_restored.rds") * 100, digits = 2)` percent of individuals who had their rights restored cast a ballot in 2018, indicating that Executive Order 181 increased the propensity to vote by around `r round(summary(iv3)[["coefficients"]][2,1] / (readRDS("./temp/to_restored.rds") - summary(iv3)[["coefficients"]][2,1]) * 100, digits = 1)` percent.
-This estimate is much lower than the estimates reported in Table \ref{tab:to-18-logit}, but that is not surprising; as discussed above, individuals whose rights were restored were more likely to vote whether or not they received restoration before discharge.
-The `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage point increase reported in Table \ref{tab:iv-models} accounts only for the extent to which rights restoration prior to discharge from parole caused higher turnout.
+Unsurprisingly, this estimate is higher than the estimates reported in Table \ref{tab:to-18-logit}.
+Because Executive Order 181 theoretically only impacts individuals whose rights were actually restored prior to discharge, the magnitude of the estimated effect on this subgroup must be larger than the group taken as a whole.
 

### 2019-07-05 markdown/part2_standalone.Rmd
write

@@
 This analysis uses an individual-level logistic regression that estimates individuals' turnout rates and controls for various individual-level characteristics.
+Such an analysis sheds light on the impact of rights restoration on the turnout rate of the formerly disenfranchised population as a whole.
 
-However, not all individuals who were discharged after the Executive Order was signed had their rights restored.
-Individuals who were arrested or were not citizens were not granted voting rights.
-Because there is reason to believe that voting propensity is positively correlated with the likelihood that an individual's rights were restored, I employ an instrumental variables approach to more precisely estimate the impact of the executive order on turnout.
+Such an approach has some benefits: supposing that policymakers care about the political representation of formerly incarcerated residents as a group, then examining the group turnout rate can yield important insights.
+This approach does not, however, address the impact having one's rights restored has on individual propensity to vote.
+To interrogate whether an individual's propensity to vote is increased if she has her rights restored prior to discharge from parole, I test whether individuals who had their rights restored prior to discharge from parole turned out at higher rates in midterm election.
+Because the restoration of voting rights is not randomly assigned, I employ an instrumental variables approach, as discussed in greater detail below.
 
@@
 
-Although Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that the length of time an individual has been off parole is correlated with her propensity to vote, that relationship is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
+When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
+The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
+For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the Executive Order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
+Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
+
+However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
 
@@
 The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] Over the short term, the length of time an individual has been off parole is not associated with his propensity to vote.
-
-This lack of correlation between discharge date and turnout indicates that we can conceptualize Executive Order as a natural experiment.
-We can expect that, in the absence of Executive Order 181, individuals discharged from parole before and after the policy went into effect would have turned out at comparable rates.
-Any observed difference in turnout rates, therefore, can be attributed to the implementation of the policy.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
 
@@
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring turnout in 2018 among individuals who were discharged on or after May 2121^st^, 2018, the date when the policy went into full effect.
-It includes all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
+Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
+Whether an individual was discharged from parole before or after May 21^st^, 2018, is akin to a randomly assigned treatment.
+Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
+
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 21^st^, 2018, turned out at higher rates than those discharged earlier.
+The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
 
@@
           type = "latex",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
                                "Time Spent on Parole (Years)"),
@@
 Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
-In each successive model, the AIC decreases substantially, indicating that these controls are warranted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
@@
 
+## Instrumental Variables Approach {-}
+
 Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
-The above analysis, however, likely understates the impact of the Executive Order on individuals who were in fact eligible to cast a ballot.
-As Figure \ref{fig:restoration-chart} indicates, not all individuals discharged after the Executive Order went into effect had their rights restored.
-Individuals who were discharged after the policy was changed but but did not have their rights restored can be conceptualized as "non-compliers" --- although they were "assigned" to the treatment group (were discharged after the policy changed), they did not receive the treatment (have their rights formally restored).
-Some of these non-compliers were likely non-citizens, who could not be granted voting rights.
-Others may have been re-arrested, which disqualified them from having rights restored.
-Both non-citizens and individuals who have been recently arrested are unlikely to vote.
-"Compliance," therefore, is almost certainly correlated with with an individual's likelihood of participating in the 2018 election.
-As such, the estimated treatment effect on the treated --- on individuals who were likely eligible to cast a ballot on election day --- requires an instrumental variables approach.
+Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
+To answer that question, we must specifically control for whether an individual did have his rights restored before he was discharged from parole.
 
-In the models below, a dummy indicating whether an individual was discharged from parole after the Executive Order went into effect (i.e., was in the "intend-to-treat" group) is used as an instrument for whether an individual's rights were actually restored (i.e., "complied" with the treatment).
-Table \ref{tab:iv-models} shows the second-stage estimation results, estimated using an OLS sepcification.^[Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.]
+An individual's propensity to vote can be expressed using the following equation:
+$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + \epsilon_{i} $$
+Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
+Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
+Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote.
+Neither of these assumptions are valid: an individual's eligibility to have her rights restored is almost certainly correlated with whether or not they actually were restored.
+Moreover, eligiblity for rights restoration is likely correlated with propensity to vote.
+Non-citizens, for instance, were not eligible for rights restoration under EO181 because non-citizens cannot vote in New York State.
+Similarly, individuals who were re-arrested while on parole were not eligible to have their rights restored.
+Citizenship status and recency of arrest are both highly correlated with propensity to vote.
+
+Our variable of interest, therefore, is likely correlated with our regression error.
+Such a correlation inhibits our ability to simply regress turnout on a dummy indicating whether an individual's rights were restored.
+The standard way of dealing with such a problem is an instrumental variables approach.
+Such an approach has been widely used in the context of voter turnout [e.g.
+@Gerber2000; @Green2012].
+A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
+
+In this case, such a variable is readily at hand.
+The likelihood that an individual had his voting rights restored is a function of whether he was discharged from parole after Executive Order 181 took effect.
+As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
+A dummy variable indicating whether an individual was discharged after the Executive Order, therefore, satisfies the criteria for an instrumental variable.
+
+Table \ref{tab:iv-models} presents the two-stage least-squares regression estimates exploring the turnout impact of rights restoration.
+As discussed above, the instrumental variable used in these regressions measure whether an individual was discharged from parole after Executive Order 181 went into effect.
+These models are estimated using an OLS sepcification.
+Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.
 
@@
                     "and Wu-Hausman tests.",
-                    "D(Discharged from Parole on or after May",
-                    "21, 2018) instrument for D(Rights Restored)."))
+                    "First-stage equations use dummy indicating",
+                    "whether individual was discharged after the",
+                    "Exectutive Order went into effect. The",
+                    "first-stage equations also include",
+                    "covariates in models 2 and 3."))
 ```
@@
 Unsurprisingly, this estimate is higher than the estimates reported in Table \ref{tab:to-18-logit}.
+Because Executive Order 181 theoretically only impacts individuals whose rights were actually restored prior to discharge, the magnitude of the estimated effect on this subgroup must be larger than the group taken as a whole.
 
-## Discussion {-}
+# Discussion {-}
 As discussed above, the effect of incarceration on individuals' propensity to vote is an open question for the literature.
@@
 Individuals who finished parole between May 21^st^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
-The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
-The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 35 percent, and raised the turnout of formerly disenfranchised individuals as a whole by
 
@@
 
+\newpage
 # Appendix A {-}
@@
           covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)",
                                "Time Spent on Parole (Years)"),

### 2019-07-05 code/part_2/03_process_scraped_data.R

@@ -118,7 +118,7 @@ parole_officers <- read_fwf(
 
 colnames(parole_officers) <- names
 
-parole_officers <- select(parole_officers, parole_office, din) %>% 
+parole_officers <- select(parole_officers, parole_office, din, parole_officer_name) %>% 
   group_by(din) %>% 
   filter(row_number() == 1)
 

### 2019-07-05 code/part_2/06_individual_turnout.R

@@ -29,7 +29,7 @@ model2 <- glm(v2018 ~ restored + days_since_done + days2 +
 
 model3 <- glm(v2018 ~ restored + days_since_done + days2 +
                 as.factor(county) + as.factor(race) + as.factor(sex) + age +
-                felony_a + felony_b + felony_c + felony_d + felony_e + counts + parole_time,
+                felony_a + felony_b + felony_c + felony_d + felony_e + parole_time,
               family = "binomial", data = parolees)
 
 save(model1, model2, model3, file = "./temp/individual_turnout_18.rdata")
@@ -45,7 +45,7 @@ model2 <- glm(v2018 ~ finished_post +
 
 model3 <- glm(v2018 ~ finished_post +
                 as.factor(county) + as.factor(race) + as.factor(sex) + age +
-                felony_a + felony_b + felony_c + felony_d + felony_e + counts + parole_time,
+                felony_a + felony_b + felony_c + felony_d + felony_e + parole_time,
               family = "binomial", data = filter(parolees, year(parole_status_date) >= 2017))
 
 save(model1, model2, model3, file = "./temp/individual_turnout_18_itt.rdata")
@@ -55,13 +55,13 @@ save(model1, model2, model3, file = "./temp/individual_turnout_18_itt.rdata")
 
 date_to_model_notime <- glm(v2018 ~ 
                        as.factor(county) + as.factor(race) + as.factor(sex) + age +
-                       felony_a + felony_b + felony_c + felony_d + felony_e + counts + parole_time,
+                       felony_a + felony_b + felony_c + felony_d + felony_e + parole_time,
                      family = "binomial",
                      data = filter(parolees, year(parole_status_date) >= 2017, parole_status_date < "2018-04-18"))
 
 date_to_model_time <- glm(v2018 ~ days_since_done + days2 +
                               as.factor(county) + as.factor(race) + as.factor(sex) + age +
-                              felony_a + felony_b + felony_c + felony_d + felony_e + counts + parole_time,
+                              felony_a + felony_b + felony_c + felony_d + felony_e + parole_time,
                             family = "binomial",
                             data = filter(parolees, year(parole_status_date) >= 2017, parole_status_date < "2018-04-18"))
 
@@ -80,15 +80,16 @@ iv2 <- ivreg(v2018 ~ restored +
 
 iv3 <- ivreg(v2018 ~ restored + 
[…]

### 2019-07-24 markdown/felony_disenfranchisement_nys.Rmd
write

@@
     - \usepackage{setspace}\doublespacing
+    - \usepackage{rotating}
 ---
@@
 Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
-And yet, despite the United State's purported march toward ever-more-inclusive systems of democracy, one large group of American citizens is formally barred from voting.
+Today, despite the United State's march toward ever-more-inclusive systems of democracy, just one large group of American citizens is formally barred from voting.
 In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised today [@sentencing_2016].
-These disenfranchising policies provide a stark reminder that our democracy is not as inclusive as we might like to tell ourselves.
 
-The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
-This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+The disenfranchisement of citizens convicted of felony offenses is intersects with the racialized and place-based patterns of policing and incarceration in the United States.
+As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
+In states such as New York, for instance, the overrepresentation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also overrepresented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
 
@@
 Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges stemming from targeted policing and economic marginalization.
+The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges thanks to social and economic marginalization.
 The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
@@
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
+Filling this gap in the literature is of great importance.
 
@@
 
-In any discussion of felony disenfranchisement in the United States, it is imperative to acknowledge the central role played by race and white supremacy.
+In any discussion of felony disenfranchisement in the United States, it is important to acknowledge the central role played by race.
 As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, p.
@@
 
-The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
+The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
-This analysis is limited to individuals incarcerated for felony offenses.
-Individuals convicted of misdemeanors are not disenfranchised in New York State.
-These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These data are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
+These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS) and are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
+
+The state makes records available only for individuals who have been incarcerated for felony offenses.
+It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
+Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
+It also only includes individuals who are held in state prisons.
 
@@
 It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
 The first section of this paper uses a snapshot of the registered voter file from April 30^th^, 2018.
 The second section of this paper uses a snapshot from March 3^rd^, 2019.
+
+The New York State Voter File is unique in its treatment of "purged" voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
+I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
 
@@
 I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
+The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
+Voters who were not successfully geocoded are dropped from the analysis; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+
+## Matching {-}
+Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
+I match individuals in each dataset using first name, middle name, last name, and date of birth.
+To be considered a "match," records must have the exact same birth date.
+The first and last names must also be exact matches (conditional on the adjustments discussed below).
+The middle names must meet one of the following conditions in order to qualify:
+
+* Middle names are identical. If neither set of records includes a middle name, this condition is met.
+* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
+* A middle name or middle initial in one set of records, and a missing middle name in the other set.
+
+Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andew Doe" and "John Anthony Doe" would not.
+
+There are two types of potential error in this methdology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
+
+Testing for the presence of false positive matches is fairly straightforward.
+@Meredith2013 offers one way to test their prevalence using placebo matching.
+I slightly alter the date of birth reported in the parole discharge dataset to create false records.
+Comparing the number of matches between these "fake" discharge records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
+This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
+true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
+plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+Testing for false negatives is more challenging.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
+To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
+A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
+Such standardizations, however, will miss individuals who change their names entirely.
+For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
 
@@
 Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
+Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
 
@@
 
-These individuals are identified by matching NYSDOCCS data to the voter file.
-I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York.
-Following @Meredith2013, I discuss the potential impact of these errors in [Appendix A].] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
@@
 Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the voting age population (estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race.^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
+I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
+Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
 Matching is done with replacement.
@@
 Table \ref{tab:match-reg} presents the results of these regression models.
-Each neighborhood is weighted by its adult population, and robust standard errors are clustered at the level of the match [@Abadie2019].
+Robust standard errors are clustered at the level of the match [@Abadie2019].^[The observations in these models are not weighted by voting age population.
+Such weighting materially impacts neither the size or statistical significance of the coefficients of interest.]
 ```{r tract-reg-chunk, echo = F, results = "asis", message = F}
@@
 
-stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
+stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black, tract_reg_black,
           column.separate = c(2, 2),
@@
 Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))*100, digits = 1)` percent, this effect is alarmingly high.
-For every 100 votes cast in a predominantly Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
-Although these variables are largely nonsignificant at the census tract level, this is not particularly surprising.
-If lost voters have the largest depressive effect on family members and close neighbors, it makes sense that the effect of losing a voter is locally concentrated.
+For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
+The effects are slightly larger when neighborhoods are measured as block groups than as census tracts.
+This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more effected.
 
@@
 This regression uses the same covariates used in the matching procedure described above.
-Each neighborhood is weighted by its voting age population, and robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purporses.]
+Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purporses.
+The observations in these models are not weighted by voting age population.
+Such weighting materially impacts neither the size or statistical significance of the coefficients of interest.]
 
@@
 Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.017\times Lost\ Voters \times Share\ Black$.
+The estimated depressive effect is $-0.018\times Lost\ Voters \times Share\ Black$.
 
@@
 During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` from New York City neighborhoods.
-The spatial concentration of these lost voters is striking: as Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} make clear, these voters were removed from neighborhoods with significantly lower incomes than the rest of the city.
-They were also removed from neighborhoods that were far less white than the average neighborhood.
+The spatial concentration of these lost voters is striking, as demonstrated in \ref{fig:citywide-map}.
 The systematic removal of voters from these neighborhoods is troubling.
-However, more than 900,000 votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, it is unlikely that the removal of such a small number of voters --- concentrated though they are --- is likely to have large implications on its own.
+However, more than one million votes were cast in the general election of 2017.
+Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
 
@@
 This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-These communities are already home to some of the poorest and most marginalized voters in the city; they are home to residents who have the least ability to influence policy through other means such as large campaign contributions.
-The fact that felony disenfranchisement policies appear to be disincentivizing participation on election day --- that these policies are further weakening the ability of these neighborhoods to participate in the democratic process --- is cause for alarm.
+As discussed above, these neighborhoods have far lower incomes than the rest of the city.
+The median income of a block groups without lost voters is more than 40 percent higher than the median income of block groups with lost voters.
+Similarly, just 17 percent of individuals in the average block group with a lost voter were white, compared with 40 percent of the population in other block groups.
+Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
+The effects are concentrated in neighborhoods where the most marginalized members of society live.
+These highly concentrated spillover effects are cause for major concern.
+
+Of equal concern is the apparent concentration of these effects in Black neighborhoods.
+Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is Non-Hispanic Black, the lost voter indicator is either insignificant or significant and positive.
+This means that lost voters have no spillover effects in neighborhoods with small Black populations.
+Understanding the magnitude of this finding is important.
+In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population.
+On average, the voting age population in these neighborhoods was `r round(readRDS("./temp/vap_gt_90_black"))`.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a voting-age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+
+The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood `r round(0.9 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+
+Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
+Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future.
+@Lerman2013, for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
+@Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their governments and their identities as citizens.
+It is possible that Black communities understand the loss of voters differently than non-Black communities, and therefore respond differently.
+
+Individuals who live in neighborhoods where police activity is relatively limited may interpret the incarceration of a neighbor as a largely individual phenomenon.
+If it is understood as an isolated or individual event, voters in these neighborhoods who are not incarcerated are not likely to update their view of the state.
+They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
+
+In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interepreted so individualistically.
+It may, rather, be interpreted as another reminder of the government's unfairness.
+If a would-be voter finds herself soured on political participation because of her neighbor's incarceration, she may be less likely to cast a ballot.
+
+Interestingly, this finding mirrors @White2019, which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This is perhaps unsurprising.
+A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans [e.g.
+@Browning1992; @Henderson1997; @Wu2009].
+It should come as no surprise that experience with the criminal justice system less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
+
+Whether this finding is surprising or not, it is troubling.
 
@@
 Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 21^st^ that the majority of discharged parolees had their voting rights restored prior to their discharge.
+It is not until May 16^th^ that there were consistently parolees who had their voting rights restored prior to discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 21^st^, 2018, is akin to a randomly assigned treatment.
+Whether an individual was discharged from parole before or after May 16^th^, 2018, is akin to a randomly assigned treatment.
 Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 21^st^, 2018, turned out at higher rates than those discharged earlier.
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 16^th^, 2018, turned out at higher rates than those discharged earlier.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
@@
           type = "latex",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Voted in 2016)", "D(Male)", "Age (Years)",
                                "Time Spent on Parole (Years)"),
@@
 Such an approach has been widely used in the context of voter turnout [e.g.
-@Gerber2000; @Green2012].
+@Ansolabehere1999; @Gerber2000; @Milligan2004].
 A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
@@
 As discussed above, the instrumental variable used in these regressions measure whether an individual was discharged from parole after Executive Order 181 went into effect.
-These models are estimated using an OLS sepcification.
-Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.
+Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2001 recommends using a standard two-stage least squares model in instrumental variables analyses even under such conditions.
+Models 1 --- 3 use a two-stage least squares specification, while Model 4 uses a two-stage probit model.
 
-```{r iv-reg-chunk-short, echo = F, results = "asis", message = F}
+\input{"../temp/iv_tables.tex"}
+
+```{r nameless-chunk, echo = F, results = "asis", message = F}
 load("./temp/iv_individual_turnout_18.rdata")
-
-summ.fit1 <- summary(iv1, diagnostics = T)
-summ.fit2 <- summary(iv2, diagnostics = T)
-summ.fit3 <- summary(iv3, diagnostics = T)
-
-stargazer(iv1, iv2, iv3,
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Rights Restored)", 
-                               "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:iv-models) Second-Stage Regression",
-          table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X"),
-                         c("\\hline \\\\[-1.8ex]"),
-                         c(rownames(summ.fit1$diagnostics)[1],
-                           round(summ.fit1$diagnostics[1, "p-value"], 2),
-                           round(summ.fit2$diagnostics[1, "p-value"], 2),
-                           round(summ.fit3$diagnostics[1, "p-value"], 2)),
-                         c(rownames(summ.fit1$diagnostics)[2],
-                           round(summ.fit1$diagnostics[2, "p-value"], 2),
-                           round(summ.fit2$diagnostics[2, "p-value"], 2),
-                           round(summ.fit3$diagnostics[2, "p-value"], 2))),
-          omit.stat = "ser",
-          table.layout = "-dm#-t-as-n",
-          notes = c("Table reports p-values for Weak Instruments",
-                    "and Wu-Hausman tests.",
-                    "First-stage equations use dummy indicating",
-                    "whether individual was discharged after the",
-                    "Exectutive Order went into effect. The",
-                    "first-stage equations also include",
-                    "covariates in models 2 and 3."))
 ```
 
-Table \ref{tab:iv-models} indicates that using discharge date of on or after May 21^st^ as an instrument for rights restoration is warranted.
+Table \ref{tab:iv-models} indicates that using discharge date of on or after May 16^th^ as an instrument for rights restoration is warranted.
 The p-values of the Wu-Hausman tests hover just below 0.05, while the Weak Instruments test is highly significant in each model.
@@
 Because Executive Order 181 theoretically only impacts individuals whose rights were actually restored prior to discharge, the magnitude of the estimated effect on this subgroup must be larger than the group taken as a whole.
-
 ## Discussion {-}
@@
 The second mechanism is less obvious.
-Individuals who finished parole between May 21^st^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
+Individuals who finished parole between May 16^th^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
 The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
@@
 <div id="refs"></div>
-
-\newpage
-# Appendix A {-}
-
-Matching across administrative data sets is not a perfect science.
-Despite having high-quality low-level data from both New York State Department of Corrections and Community Supervision (DOCCS) and the New York State Board of Elections, I cannot be sure that I am correctly identifying formerly incarcerated, registered voters.
-The results of the analyses in Table \ref{tab:match-reg} give some reason to believe that there are few false negatives.
-If there were many false negatives (lost voters who are not identified in this analysis), "treated" neighborhoods would be incorrectly categorized as "untreated" neighborhoods.
-If this were a widespread problem, we would expect to see little difference between treated and control neighborhoods.
-To the extent that lost voters go unidentified by this analysis, it is likely that the results presented in Tables \ref{tab:match-reg} and \ref{tab:trad-reg} are conservative by including treated neighborhoods in the control group.
-Directly testing for false negatives, however, is difficult.
-If a voter changes her name before being incarcerated, for instance, our analysis will fail to capture her.
-
-Testing for the presence of false positives is slightly easier.
-One way to do so is by constructing false records [@Meredith2013].
-By slightly changing the birthdates in one set of administrative data, I can estimate how frequently DOCCS records inaccurately match with voter records.
-To do so, I shift the dates of birth in the DOCCS data by 35 days.
-I then re-merge the DOCCS records (with "false" birth dates) against the registered voter file.
-
-Table \ref{tab:change-dobs} shows the results of these permutations.
-The first row shows the matches with the raw data.
-The second row shows the results of adding 35 days to the date of birth in each DOCCS record, and the final row shows the result of subtracting 35 days from these records.
-
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-Table \ref{tab:change-dobs} indicates that there are likely some false positives in our matches.
-It is not surprising that in a large state like New York, full names and dates of birth fail to uniquely identify individuals.
-However, the rate of false positives is quite low: this analysis indicates that there is a `r percent(as.integer(max(less_35, plus_35)) / as.integer(gsub(",", "", true)), accuracy = 0.01)` rate of false positives.
-It is unlikely that such a small incidence of false positives meaningfully impacts this analysis.
 
@@
 In the [Instrumenting Rights Restoration] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 21^st^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 16^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
 Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
@@
           type = "latex",
-          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
+          covariate.labels = c("D(Discharged on or after May 16\\textsuperscript{th}, 2016)", 
+                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
                                "Time Spent on Parole (Years)"),

### 2019-07-24 markdown/part2_standalone.Rmd
write

@@
 Below, I employ a logistic regression to investigate the impact of Executive Order 181 on turnout among all parolees discharged after the policy went into effect.
-Because not all individuals had their rights restored, I follow this analysis with an instrumental-variables approach to determine the effect of rights restoration on individuals whose rights were in fact formally restored prior to discharge from parole.
+However, not all individuals discharged after the Executive Order went into effect had their rights restored.
+Because an individual's likelihood of having his rights restored prior to discharge from parole may be correlated with his propensity to vote, I also employ an instrumental-variables approach to estimate the effect of the Executive Order on turnout.
 
@@
 Whether an individual was discharged from parole before or after May 21^st^, 2018, is akin to a randomly assigned treatment.
+There is no evidence that individuals discharged from parole after this date would have turned out at different rates in the absence of the Executive Order.
 Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 21^st^, 2018, turned out at higher rates than those discharged earlier.
+Table \ref{tab:to-18-logit} presents models estimating individual-level turnout.
+The dependent variable is binary --- it equals 1 if an individual cast a ballot, and 0 if they did not.
+Because of the binary nature of the dependent variable, I fit the models using a logistic regression.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
@@
 Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
+Model 2 adds individual-level demographic characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
+Model 3 adds sentence-specific information to Model 2: the amount of time an individual spent on parole and the class(es) of felony for which they were convicted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
@@
 Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
-To answer that question, we must specifically control for whether an individual did have his rights restored before he was discharged from parole.
+Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a class, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
+To answer that question, we must look specificially at whether an individual did have his rights restored before he was discharged from parole.
 

### 2019-07-30 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 \pagenumbering{arabic}
-# Introduction {-}
+# Introduction
 
@@
 
-This paper seeks to extend the extant literature on felony disenfranchisement by considering the implications of disenfranchisement on local political participation.
-I begin by looking at participation rates in New York City neighborhoods where individuals with a demonstrated history of participation have been disenfranchised.
+The rest of this paper is structured as follows.
+**Section 2** discusses the historical context of felony disenfranchisement in the United States, and **Section 3** provides an overview of the scholarship conducted in the field over the past two decades.
+**Section 4** provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
+**Section 5** presents the data sources used throughout the project.
+
+**Section 6** proposes a new definition of "lost voters" in the context of felony disenfranchisement.
+I define lost voters as individuals who are not only disenfranchised, but are disenfranchised and have a history of participating in elections.
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
+Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
 The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among elgibile voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
+A discussion of the results follows.
 
-Undoing the negative effects of felony disenfranchisement on individual and neighborhood representation requires crafting policy aimed at increasing turnout among the formerly disenfranchised.
-New York State has taken a step in that direction: in April of 2018, Governor Andrew Cuomo signed Executive Order 181, which came close to ending disenfranchisement for New Yorkers on parole.
-After examining the neighborhood effects of felony disenfranchisement, I test whether having one's rights restored while still on parole increased turnout among formerly disenfranchised individuals in the 2018 midterm elections.
+**Section 7** turns from the 2017 mayoral election to the 2018 midterm elections.
+In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to many individuals in the state who were on parole.
+I employ an instrumental variables approach and a bivariate probit approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
+I focus specifically on individuals who would have been eligible to cast a ballot even if the Executive Order had not gone into effect.
+A discussion of the results follows.
 
-# Background of Felony Disenfranchisement in the United States {-}
+**Section 8** sythesizes the findings presented in the rest of the paper and provides concluding remarks.
+
+# Background of Felony Disenfranchisement in the United States
 
@@
 
-# Academic Literature and Felony Disenfranchisement {-}
+# Academic Literature and Felony Disenfranchisement
 
@@
 
-# What This Paper Examines {-}
+# What This Paper Examines
 
@@
 
-# Data {-}
+# Data
 ## Criminal Justice Data {-}
@@
 The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
 These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS) and are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
@@
 This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
 
@@
 
-# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels {-}
+# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels
 
@@
 The voter file indicates that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that `r comma(readRDS("./temp/mayoral_ballots.rds"))` votes were cast.
-In [Appendix B], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
+In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
 These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
@@
 
-stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black, tract_reg_black,
+stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
           column.separate = c(2, 2),
@@
           notes = c("Robust standard errors (clustered at", "match level) in parentheses."),
-          type = "latex",
+          type = "latex", notes.align = "l",
           covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
@@
           header = F,
-          type = "latex",
+          type = "latex", notes.align = "l",
           se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
@@
 
-Whether this finding is surprising or not, it is troubling.
+Whether this finding is surprising or not, it is troubling. Build8 this out more
 
@@
 
-# Executive Order 181 and Turnout in 2018 {-}
+# Executive Order 181 and Turnout in 2018
 
@@
 
-```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Presidential Election", fig.pos = "H", fig.align = "center"}
+```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
 j <- readRDS("./temp/to18_chart_recent.rds")
@@
           header = F,
-          type = "latex",
-          covariate.labels = c("Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
+          type = "latex", notes.align = "l",
+          covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
+                               "D(Male)", "Age (Years)",
+                               "Years on Parole"),
           dep.var.labels = c("Cast Ballot in 2018 Election"),
@@
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "X", "X"),
-                         c("Felony Class Fixed Effects", "X", "X")),
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
+                         c("Felony Class FE", "X", "X")),
           table.layout = "-dm#-t-a-s-n")
@@
 The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix B] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
 
@@
 Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 16^th^ that there were consistently parolees who had their voting rights restored prior to discharge.
+It is not until May 18^th^ that there were consistently parolees who had their voting rights restored prior to discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 16^th^, 2018, is akin to a randomly assigned treatment.
+Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
+Individuals who were discharged from parole prior to the implementation of the Executive Order are part of the "control group", while individuals discharged after are "treated" by the policy change.
 Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 16^th^, 2018, turned out at higher rates than those discharged earlier.
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
@@
           header = F,
-          type = "latex",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Voted in 2016)", "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
+          type = "latex", notes.align = "l",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+                               "Years on Parole"),
           dep.var.labels = c("Cast Ballot in 2018 Election"),
@@
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
+                         c("Felony Class FE", "", "", "X")),
           table.layout = "-dm#-t-a-s-n")
@@
 Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
 Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
@@
 Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
-To answer that question, we must specifically control for whether an individual did have his rights restored before he was discharged from parole.
+To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole.
+
+Investigating the causal effect of rights restoration using the natural experiment conceptualizes introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied."
 
 An individual's propensity to vote can be expressed using the following equation:
-$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + \epsilon_{i} $$
+$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} $$
 Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
+*Z* is a vector of other factors (such as age and race) known to influence voter turnout.
 Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote.
-Neither of these assumptions are valid: an individual's eligibility to have her rights restored is almost certainly correlated with whether or not they actually were restored.
-Moreover, eligiblity for rights restoration is likely correlated with propensity to vote.
-Non-citizens, for instance, were not eligible for rights restoration under EO181 because non-citizens cannot vote in New York State.
-Similarly, individuals who were re-arrested while on parole were not eligible to have their rights restored.
-Citizenship status and recency of arrest are both highly correlated with propensity to vote.
+Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
+Neither of these assumptions are valid: an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored: the Executive Order required that all eligible individuals have their voting rights restored.
 
-Our variable of interest, therefore, is likely correlated with our regression error.
-Such a correlation inhibits our ability to simply regress turnout on a dummy indicating whether an individual's rights were restored.
-The standard way of dealing with such a problem is an instrumental variables approach.
+There is also reason to believe that an individual's likelihood of having his rights restored is correlated with his natural propensity to vote.
+Individuals who were discharged from parole but are not U.S.
+citizens, for instance, did not having voting rights restored (or, in this case, granted) after the Executive Order went into effect.
+This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
+The available data does not indicate whether individuals were eligible to have their rights restored.
+We can, however, examine the demographic differences between individuals who had their rights restored and those who did not.
+
+```{r restoree-demo-chunk, echo = F}
+t <- readRDS("./temp/demos_restorees.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
+                  align = c("l", "c", "c")) %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F) %>% 
+  footnote(c("* Difference is significant at 95% confidence level.",
+             "Shows demographics for individuals discharged from parole",
+             "between May 18th, 2018, and October 12th, 2018."))
+j
+```
+
+Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
+Men, for instance, were less likely to have their rights restored than women.
+Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
+Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
+This is likely a reflection of citizenship status.
+Table \ref{tab:demos-restorees} indicates that *b~1~* is probably correlated with *Y*.
+
+An ordinary least squares (or one-stage logistic) approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate.
+Such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
+
+The standard way of dealing with such a problem is an instrumental variables approach [@Angrist1996].
 Such an approach has been widely used in the context of voter turnout [e.g.
-@Ansolabehere1999; @Gerber2000; @Milligan2004].
+@Ansolabehere1999; @Gerber2000; @Milligan2004;  @Lassen2004; @Sondheimer2010].
 A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
@@
 In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is a function of whether he was discharged from parole after Executive Order 181 took effect.
+The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
 As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
@@
 
-Table \ref{tab:iv-models} presents the two-stage least-squares regression estimates exploring the turnout impact of rights restoration.
-As discussed above, the instrumental variable used in these regressions measure whether an individual was discharged from parole after Executive Order 181 went into effect.
-Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2001 recommends using a standard two-stage least squares model in instrumental variables analyses even under such conditions.
-Models 1 --- 3 use a two-stage least squares specification, while Model 4 uses a two-stage probit model.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the "local average treatment effect." The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
+Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
+Some research [e.g.
+@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (IV probit) model to avoid the constraints of the nonparametric framework.
+The IV probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable.
 
-\input{"../temp/iv_tables.tex"}
+The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
+This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
+
+Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
+Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
+Model 3 uses the IV probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
+For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
+
+\input{"../temp/iv_clean.tex"}
+
+Although these models demonstratke that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
+In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of parolees.
+These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
+Once again, the marginal effects in the biprobit models are calculated at the means of the other variables.
+
+\input{"../temp/iv_clean_race.tex"}
 
@@
 
-Table \ref{tab:iv-models} indicates that using discharge date of on or after May 16^th^ as an instrument for rights restoration is warranted.
-The p-values of the Wu-Hausman tests hover just below 0.05, while the Weak Instruments test is highly significant in each model.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
+We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
 
-Model 3 in Table \ref{tab:iv-models} indicates that having one's rights restored caused an individual's propensity to vote to increase by `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage points.
-Approximately `r round(readRDS("./temp/to_restored.rds") * 100, digits = 2)` percent of individuals who had their rights restored cast a ballot in 2018, indicating that Executive Order 181 increased the propensity to vote by around `r round(summary(iv3)[["coefficients"]][2,1] / (readRDS("./temp/to_restored.rds") - summary(iv3)[["coefficients"]][2,1]) * 100, digits = 1)` percent.
-Unsurprisingly, this estimate is higher than the estimates reported in Table \ref{tab:to-18-logit}.
-Because Executive Order 181 theoretically only impacts individuals whose rights were actually restored prior to discharge, the magnitude of the estimated effect on this subgroup must be larger than the group taken as a whole.
+Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
+Some of this may be explained by different propensities to vote.
+@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (pg.
+321).
+The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote, they may be less "susceptible" to a policy intervention of this sort.
+
+There is reason to believe this may be the case.
+Table \ref{tab:black-to} presents a series of logit models estimating turnout rates of all individuals discharged from parole prior to the Executive Order going into effect, and the turnout rates of all individuals who had their rights restored.
+Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate), these individuals have been excluded from Models 2 and 4.
+
+```{r logit-black-to-chunk, echo = F, results = "asis", message = F}
+load("./temp/black_to.rdata")
+
+stargazer(black_to1, black_to2, black_to3, black_to4,
+          column.separate = c(2, 2),
+          column.labels = c("Discharged Prior to EO", "Had Rights Restored"),
+          header = F,
+          type = "latex", notes.align = "l",
+          covariate.labels = c("D(Black)", "D(Male)", "Age (Years)",
+                               "Years on Parole"),
+          dep.var.labels = c("Voted in 2018"),
+          title = "(\\#tab:black-to) Turnout Rate Among Former Parolees",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          notes = c("Standard errors in parentheses.",
+                    "Latino individuals excluded",
+                    "from Models 2 and 4."),
+          omit = "felony")
+```
+
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, 6,066 (or 43 percent) were Black.
+These Black voters were more than 50 percent more likely to vote than the rest of the population.
+Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
+In both specifications, this higher turnout rate is signficant at the 95 percent level of confidence.
+
+Table \ref{tab:black-to} also shows, however, that the turnout rate among Black individuals who had their rights restored were not statistically significantly higher (or lower) than other individuals who had their rights restored.
+The Executive Order, therefore, seems to have had a leveling effect --- Black parolees were not longer substantially more likely to turnout than other individuals discharged from parole in New York State.
+
+**EVERYTHING BELOW HERE NEEDS TO BE RE-WORKED. IT DOES NOT DEAL WITH RACIAL DIFFERENCES IN TURNOUT EFFECTS**
+
 ## Discussion {-}
@@
 The second mechanism is less obvious.
-Individuals who finished parole between May 16^th^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
+Individuals who finished parole between May 18^th^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
 The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
@@
 
+# Conclusion
+
 \newpage
@@
 \newpage
-# Appendix B {-}
+# Appendix A {-}
 
@@
           header = F,
-          type = "latex",
+          type = "latex", notes.align = "l",
           se = list(ed_comp_ses),
@@
 \newpage
-# Appendix C {-}
+# Appendix B {-}
 
-In the [Instrumenting Rights Restoration] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 16^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
 Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
@@
           header = F,
-          type = "latex",
-          covariate.labels = c("D(Discharged on or after May 16\\textsuperscript{th}, 2016)", 
+          type = "latex", notes.align = "l",
+          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
                                "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
@@
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
+                         c("Felony Class FE", "", "", "X")),
           table.layout = "-dm#-t-a-s-n")
 ```
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot (the p-value on the coefficient of interest exceeds 0.5 in each model).
+Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
 The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.

### 2019-07-30 markdown/felony_disenfranchisement_nys.Rmd
spell check

@@
 
-The political history of the United States has been characterized by a general, if unlinear, trend toward universal sufferage [see, for instance, @Keyssar2009].
+The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage [see, for instance, @Keyssar2009].
 At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
@@
 As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the overrepresentation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also overrepresented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
 As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
@@
 Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantanges thanks to social and economic marginalization.
+The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
 The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
@@
 Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among elgibile voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
+The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
 A discussion of the results follows.
@@
 
-**Section 8** sythesizes the findings presented in the rest of the paper and provides concluding remarks.
+**Section 8** synthesizes the findings presented in the rest of the paper and provides concluding remarks.
 
@@
 800).
-They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, elibible Black voters are less likely to cast a ballot.
+They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
 These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
@@
 185).
-This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definiton of "lost voters."
+This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
 
@@
 This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017.
-Policing and incarceration patterns have historically targeted communities of color, wreaking havok on the social fabric of these neighborhoods [e.g.
+Policing and incarceration patterns have historically targeted communities of color, wreaking havoc on the social fabric of these neighborhoods [e.g.
 @Sewell2016; @Clear2008; @Lerman2013].
@@
 We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representatition of certain parts of the city.
+To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
 
@@
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no suprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
+It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characterstics with lower propensity voters generally.
+Individuals who go to prison share many characteristics with lower propensity voters generally.
 Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
@@
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect from notification in the absense of eligibility changes.
+@Meredith2013 finds no effect from notification in the absence of eligibility changes.
 
@@
 
-Secondly, Executive Order 181 is expected to dispel confusion about eligibilty and serve as a reminder to vote.
+Secondly, Executive Order 181 is expected to dispel confusion about eligibility and serve as a reminder to vote.
 As @Drucker2005 and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
@@
 
-Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andew Doe" and "John Anthony Doe" would not.
+Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andrew Doe" and "John Anthony Doe" would not.
 
-There are two types of potential error in this methdology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
 False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
@@
 ## Identification of Lost Voters {-}
-In this analysis, I offer a different defintion of "lost voter" than much of the literature.
+In this analysis, I offer a different definition of "lost voter" than much of the literature.
 Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
@@
 In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
-These reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
+These reporting anomalies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
 
@@
 Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfrancisement in the most hard-hit communities.
+The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
 This regression uses the same covariates used in the matching procedure described above.
-Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purporses.
+Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.
 The observations in these models are not weighted by voting age population.
@@
 
-The block groups where these depressive effects are not randomly distributed througout the city.
+The block groups where these depressive effects are not randomly distributed throughout the city.
 They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
@@
 
-In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interepreted so individualistically.
+In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
 It may, rather, be interpreted as another reminder of the government's unfairness.
@@
 
-Although these models demonstratke that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
+Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
 In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of parolees.
@@
 Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
-In both specifications, this higher turnout rate is signficant at the 95 percent level of confidence.
+In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 

### 2019-09-04 markdown/felony_disenfranchisement_nys.Rmd
write much bro?

@@
 link-citations: yes
+fontsize: 12pt
 header-includes:
@@
 Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
-Today, despite the United State's march toward ever-more-inclusive systems of democracy, just one large group of American citizens is formally barred from voting.
+Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
 In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised today [@sentencing_2016].
+Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised [@sentencing_2016].
 
-The disenfranchisement of citizens convicted of felony offenses is intersects with the racialized and place-based patterns of policing and incarceration in the United States.
+The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
 As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
@@
 The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Burch2011; @Gerber2014].
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2014; @Burch2011].
 Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
@@
 **Section 7** turns from the 2017 mayoral election to the 2018 midterm elections.
-In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to many individuals in the state who were on parole.
-I employ an instrumental variables approach and a bivariate probit approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
+In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
+I employ an instrumental variables approach test whether rights restoration boosted turnout among formerly incarcerated individuals.
 I focus specifically on individuals who would have been eligible to cast a ballot even if the Executive Order had not gone into effect.
@@
 
-In any discussion of felony disenfranchisement in the United States, it is important to acknowledge the central role played by race.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, p.
-4; emphasis in the original].
+Any discussion of felony disenfranchisement in the United States must center the role played by race.
+As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (Burch 2010, 4; emphasis in the original).
 Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
@@
 They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as those who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after the Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
+Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
+In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
 Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
@@
 
-A number of papers have also explored the impact that felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
 @King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to determine individual-level turnout.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates.
 They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-Ultimately, they argue that "African American disenfranchisement plays a unique role in predicting African American voter turnout" (p.
-800).
-They conclude that disenfranchisement has spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
+They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
 These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
 As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
-724).
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (724).
 
@@
 @Burch2013 is an exception to this.
-In this paper, Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system.
-Using a standard ordinary least squares regression, she determined that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (p.
-185).
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
 This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
@@
 
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017.
+This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
 Policing and incarceration patterns have historically targeted communities of color, wreaking havoc on the social fabric of these neighborhoods [e.g.
@@
 We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
+To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
 
@@
 In April, 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms the parolees under their supervision and inform them of their voting rights.
+Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
 In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I build individual-level models to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
+To test whether the policy was successful, I use administrative data to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
 
 The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather thank think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
+Rather than think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
 This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
@@
 With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-The degree to which these effects are concentrated in particular neighborhoods has important implications for our understanding of the problem of felony disenfranchisement.
-Although literature examining the degree to which geographic polarization effects voting patterns is scarce, there is some evidence to suggest that neighborhoods do vote similarly.
-@Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
+For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
+There is some evidence that spatial segregation leads to unique voting preferences by neighborhoods: @Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
 Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-Insofar as neighborhoods show distinct preferences for politicians, spatially concentrated depressive effects from felony disenfranchisement could influence election outcomes.
 
 Even if neighborhoods to not have unique preferences for presidential candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level, for instance, representatives direct resources to areas within their districts that provide the greatest political benefit [@Martin2003].
+At the Congressional level, for instance, representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
 Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (p.
-59).
-@Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (p.
-1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs.
+"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (59).
+@Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
+If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
 
@@
 
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
+It should be noted that, despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
 In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
@@
 A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
+Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
+This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from incarceration.
+Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
 
@@
 
+The bivariate probit model returns modestly more conservative estimates than the two-stage least squares model.
+After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored prior to discharge by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
+Though these numbers may seem small, they represent relatively large gains.
+Just 3.0 percent of individuals who had their rights restored prior to parole discharge cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
+
+### Racial Variation {-}
+
 Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
@@
 The Executive Order, therefore, seems to have had a leveling effect --- Black parolees were not longer substantially more likely to turnout than other individuals discharged from parole in New York State.
+
+### Variable Treatment Intensity {-}
+
+There is some reason to believe that using a constant treatment variable understates the true impact of Executive Order 181.
+Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
+This may be due to a number of factors: parole officers may have had longer to understand and communicate the new rules.
+Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
+Table \ref{tab:iv-models-time} interacts an individual's restoration status with the number of months they spent on parole after the policy went into effect ("Months Restored").
+This variable is once again instrumented using a variable indicating whether the individual was discharged before or after the Executive Order went into effect.
+Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in the Table \ref{tab:iv-models-time}.
+
+\input{"../temp/iv_clean_months.tex"}
+
+Table \ref{tab:iv-models-time} indicate that there may be variable treatment intensity --- that individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
+The two-stage least squares model indicates that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.28 percentage points (in the 2SLS model) and 0.30 percentage points (in the 2S Probit model).
+Put another way: these models indicate that rights restoration increased the propensity to vote for individuals released within one month of the policy being implemented by `r round((0.28 / (3.13589 - 0.28)) * 100, digits = 1)` percent.
+It increased the propensity to vote among those released five months after the Executive Order went into effect by `r round(((0.28 * 5) / (3.7702 - (0.28 * 5))) * 100, digits = 1)` percent.
+
+Of course, with very few months between the implementation of the Executive Order and the registration deadline for the 2018 midterms, it is impossible to know the true impact more months on parole after having one's rights restored has.
+One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
+Incorporation of data after the 2020 presidential election will likely shed more light on the Executive Order's impact on turnout among formerly disenfrnachised individuals.
 

### 2019-09-06 markdown/felony_disenfranchisement_nys.Rmd
writing, add citations

@@
   bookdown::pdf_document2:
+    keep_tex: true
     toc: no
@@
 link-citations: yes
-fontsize: 12pt
+# fontsize: 12pt
 header-includes:
@@
 library(stargazer)
+library(data.table)
 loadfonts()
@@
 Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (Burch 2010, 4; emphasis in the original).
+As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, 4; emphasis in the original].
 Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
 In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, p.
-374].
+Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, 374].
 
@@
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006; @Walker2014].
+These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006].
 As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
@@
 This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, wreaking havoc on the social fabric of these neighborhoods [e.g.
+Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods [e.g.
 @Sewell2016; @Clear2008; @Lerman2013].
 It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes voters from certain neighborhoods, but it is not clear whether enough voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on *eligible* voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
+We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear whether enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
+To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
+
+I also offer a refinement of our understanding of who lost voters are.
+Previous studies exploring the relationship between felony disenfranchisement and spill-over effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
+In this analysis, I use administrative data to identify neighborhoods where individuals with a demonstrated history of participation are disenfranchised.
 
@@
 
+Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
+In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
+Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
+
+Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
+@Burch2013, for one, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
+Moreover, voting is a social act, and social networks play an important role in predicting political participation [e.g.
+@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
+Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
+It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
+To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
+
+Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spill-over effects of lost voters.
+@Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact", [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
+These effects are particularly pronounced for women of color.
+
+### Theoretical Implications for Reduced Turnout {-}
+
 As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
@@
 For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences by neighborhoods: @Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
+There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
 Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
@@
 Even if neighborhoods to not have unique preferences for presidential candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level, for instance, representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
+At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
 Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
 "[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (59).
-@Griffin2005 finds similar effects in the United States Senate, reporting that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
+@Griffin2005 finds similar effects in the United States Senate, demonstrating that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
 If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
+This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
 
 Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group." American cities are highly segregated by race and by class, but less so by age.
+Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group" (1).
+American cities are highly segregated by race and by class, but less so by age.
 As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
 An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a bloc, their relative turnout rates may influence city policy.
+Insofar as neighborhoods vote as a "cohesive" bloc, their relative turnout rates may influence city policy.
 If felony disenfranchisement has a depressive effect on turnout, politicians are likely less responsive to the particular preferences of voters in these neighborhoods --- some of the most neglected neighborhoods to begin with.
-
-It should be noted that, despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-
-This *de facto* disenfranchisement is likely to be concentrated within the neighborhoods home to formally disenfranchised residents.
-Voting is a social act, and social networks play an important role in predicting political participation [e.g.
-@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
 
@@
 In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (p.
-1144).
-
-@White2019, however, indicates that interaction with the criminal justice system for individuals in the context of arrests for misdemeanor charges may have depressive effects on turnout.
+They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
+
+@White2019, however, comes to a different conclusion.
+Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
 This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
 Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-The findings in @White2019 agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (p.
-829).
+The findings in @White2019 agree with previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
+@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
 
@@
 @Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (p.
-924).
+"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
 It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
@@
 Secondly, Executive Order 181 is expected to dispel confusion about eligibility and serve as a reminder to vote.
-As @Drucker2005 and others have detailed, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
+Many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot [@Drucker2005; @locked_out].
 If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.
@@
 
-The primary criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
+The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
+They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS) and are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
+These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and when they finished their parole supervision.
 
@@
 However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the state, I constructed a list of the individuals who were on parole and had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
+By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
 
@@
 Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration.
+As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration [e.g.
+@Gerber2017].
 If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
@@
                   align = "c") %>% 
-  row_spec(15, hline_after = T) %>% 
+  row_spec(14, hline_after = T) %>% 
   kable_styling(latex_options = c("HOLD_position"))
@@
 Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.^[The New York registered voter file does not align exactly with results reported by the city.
-The voter file indicates that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that `r comma(readRDS("./temp/mayoral_ballots.rds"))` votes were cast.
-In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level.
-These reporting anomalies are unlikely to impact this analysis.] The number of ballots cast is divided by the neighborhood's voting age population.
+Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
+The number of ballots cast is divided by the neighborhood's voting age population.
 
@@
 Such weighting materially impacts neither the size or statistical significance of the coefficients of interest.]
-```{r tract-reg-chunk, echo = F, results = "asis", message = F}
+```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
 load("./temp/match_reg_block_group.rdata")
@@
           se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
-          notes = c("Robust standard errors (clustered at", "match level) in parentheses."),
+          notes = c("TO REPLACE"),
           type = "latex", notes.align = "l",
+          notes.append = T,
           covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:match-reg) Matching Regression",
+          dep.var.labels = c("Turnout Rate in 2017 General Election"),
+          title = "\\label{tab:match-reg} Matching Regression",
           table.placement = "H",
-          omit.stat = c("f", "ser"))
-```
+          omit.stat = c("f", "ser"),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F)
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
+
+j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table1.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table1.tex"}
+
 When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
@@
 
-```{r trad-reg-chunk, echo = F, results = "asis", message = F}
+```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
 load("./temp/tract_model_reg_ols.rdata")
@@
           header = F,
-          type = "latex", notes.align = "l",
+          type = "latex",
+          notes.align = "l",
+          notes.append = T,
           se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
@@
                                "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Turnout Rate"),
-          title = "(\\#tab:trad-reg) Standard Regression",
+          dep.var.labels = c("Turnout Rate in 2017 General Election"),
+          title = "\\label{tab:trad-reg} Standard Regression",
           table.placement = "H",
           omit.stat = c("f", "ser"),
-          notes = c("Robust standard errors (clustered by city",  "council district) in parentheses."),
-          omit = "boro")
-```
+          notes = "TO REPLACE",
+          omit = "boro",
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F)
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
+
+j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table2.tex"}
+
 The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
@@
 As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement likely impacts Black turnout at the state level.
+Previous literature has established that felony disenfranchisement impacts Black turnout at the state level.
 This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
@@
 They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
-
 In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
@@
 
-Interestingly, this finding mirrors @White2019, which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This finding mirrors @White2019, which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
 This is perhaps unsurprising.
 A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans [e.g.
-@Browning1992; @Henderson1997; @Wu2009].
-It should come as no surprise that experience with the criminal justice system less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
-
-Whether this finding is surprising or not, it is troubling. Build8 this out more
+@Browning1992; @Hurwitz2005; @Henderson1997; @Wu2009].
+Experience with the criminal justice system less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
+
+For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods [@Wilson1990].
+More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
+Much of this work, however, has focused on either the impacts of living in marginalized communities (by measuring health and economic impacts) or has not accounted for the importance of physical space (by focusing only on proximal social, and not geographic, contact with directly impacted individuals).
+This analysis allows us to understand the implications for living in a neighborhood where individuals likely to cast a ballot are not allowed to because of a felony conviction.
+I find that neighborhoods that are home to lost voters --- and particularly neighborhoods with large Black populations --- systematically turn out for local elections at lower rates than otherwise similar neighborhoods.
+
+This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
+In the case of the 2017, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
+This does not mean, however, that the depressed turnout in these neighborhoods had no impact on their representation.
+City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
+
+Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
+Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and nonwhite Americans.
+As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
 
@@
 
-```{r logit-reg-chunk-short, echo = F, results = "asis", message = F}
+```{r logit-reg-chunk-short, echo = F, results = "hide", message = F}
 load("./temp/short_term_to.rdata")
@@
           dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit-short) Individual-Level Logit Model",
+          title = "\\label{tab:to-18-logit-short} Individual-Level Logit Models",
           table.placement = "H",
@@
                          c("Felony Class FE", "X", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
-
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(
+            sqrt(diag(vcovHC(date_to_model_notime, type = "HC0"))),
+            sqrt(diag(vcovHC(date_to_model_time, type = "HC0"))))
+          )
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+write.table(j, "./temp/table3.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../temp/table3.tex"}
 The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix B] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
+A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
 
@@
 
-```{r logit-reg-chunk, echo = F, results = "asis", message = F}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F}
 load("./temp/individual_turnout_18_itt.rdata")
@@
           dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
+          title = "\\label{tab:to-18-logit} Individual-Level Logit Models",
           table.placement = "H",
@@
                          c("Felony Class FE", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(
+            sqrt(diag(vcovHC(model1, type = "HC0"))),
+            sqrt(diag(vcovHC(model2, type = "HC0"))),
+            sqrt(diag(vcovHC(model3, type = "HC0")))))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+write.table(j, "./temp/table4.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../temp/table4.tex"}
 
@@
 
-```{r restoree-demo-chunk, echo = F}
+```{r restoree-demo-chunk, echo = F, warning = F}
 t <- readRDS("./temp/demos_restorees.rds")
 j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
-                  align = c("l", "c", "c")) %>% 
+                  align = c("l", "c", "c"), "latex") %>% 
   kable_styling(latex_options = c("HOLD_position"),
-                full_width = F) %>% 
-  footnote(c("* Difference is significant at 95% confidence level.",
-             "Shows demographics for individuals discharged from parole",
-             "between May 18th, 2018, and October 12th, 2018."))
-j
-```
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
+\\\\Shows demographics for individuals discharged from parole \\\\between May 18th, 2018, and October 12th, 2018.}}}"
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex), n = 16.5)) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[3] <- lab
+
+write.table(j, "./temp/table_whatever.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table_whatever.tex"}
 
@@
 Table \ref{tab:demos-restorees} indicates that *b~1~* is probably correlated with *Y*.
-
-An ordinary least squares (or one-stage logistic) approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate.
-Such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
+A one-stage approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
 
@@
 @White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (pg.
-321).
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
 The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote, they may be less "susceptible" to a policy intervention of this sort.
@@
 
-```{r logit-black-to-chunk, echo = F, results = "asis", message = F}
+```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
 load("./temp/black_to.rdata")
@@
                                "Years on Parole"),
-          dep.var.labels = c("Voted in 2018"),
-          title = "(\\#tab:black-to) Turnout Rate Among Former Parolees",
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "\\label{tab:black-to} Turnout Rate Among Former Parolees",
           table.placement = "H",
           omit.stat = c("f", "ser"),
-          notes = c("Standard errors in parentheses.",
-                    "Latino individuals excluded",
-                    "from Models 2 and 4."),
-          omit = "felony")
-```
-
+          omit = "felony",
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F,
+          notes = "TO REPLACE",
+          se = list(
+            sqrt(diag(vcovHC(black_to1, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to2, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to3, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to4, type = "HC0")))
+          ))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors in parentheses.
+\\\\Latino individuals excluded from Models 2 and 4.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
+
+j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table5.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table5.tex"}
 Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, 6,066 (or 43 percent) were Black.
@@
 
-As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections.
-According to the official results, `r comma(readRDS("./temp/mayoral_ballots.rds"))` ballots were cast in the 2017 mayoral election.
-The registered voter file reports that just `r comma(readRDS("./temp/vf_ballots_17.rds"))` voters cast a ballot.
-This is not necessarily evidence of poor management by the Board of Elections.
-Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court).^[This is not to say that there is *not* evidence of mismanagement by the BOE.
-For instance, the author is incorrectly marked as not participating in the 2017 mayoral primary.] Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower.
-If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.
-
-To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level.
-The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level.
-Precincts are assigned the demographics of the block group in which they are situated.
-Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group.
-The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.
-Robust standard errors are clustered at the assembly district level.
-
-```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
-load("./temp/compare_ballots_off_results.RData")
-
-stargazer(ed_reg,
-          header = F,
-          type = "latex", notes.align = "l",
-          se = list(ed_comp_ses),
-          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Undercount Rate"),
-          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = c("Robust standard errors (clustered by", "assembly district) in parentheses."),
-          omit = "boro")
-```
-
-As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate.
-Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
-
-\newpage
-# Appendix B {-}
-
 In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
@@
 Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
+
+```{r logit-reg-chunk-16, echo = F, results = "hide", message = F}
 load("./temp/individual_turnout_16.rdata")
@@
           dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "(\\#tab:to-16-logit) Individual-Level Logit Model",
+          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
           table.placement = "H",
@@
                          c("Felony Class FE", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
-```
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F,
+          notes = "TO REPLACE",
+          se = list(
+            sqrt(diag(vcovHC(model1, type = "HC0"))),
+            sqrt(diag(vcovHC(model2, type = "HC0"))),
+            sqrt(diag(vcovHC(model3, type = "HC0")))
+          ))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+write.table(j, "./temp/table55.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table55.tex"}
+
 Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.

### 2019-09-06 markdown/felony_disenfranchisement_nys.Rmd
write

@@
   bookdown::pdf_document2:
-    keep_tex: true
     toc: no
@@
     latex_engine: xelatex
+    # includes: 
+    #   in_header: "../code/misc/preamble.tex"
+    keep_tex: false
 bibliography: "../research/citations.bib"
 link-citations: yes
-# fontsize: 12pt
+
 header-includes:
@@
     - \usepackage{rotating}
+
 ---
@@
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-Filling this gap in the literature is of great importance.
+Filling this gap in the literature is of great importance considering the spatial concentration of policing networks.
 
@@
 @Browning1992; @Hurwitz2005; @Henderson1997; @Wu2009].
-Experience with the criminal justice system less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
+Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
 
@@
 Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
+Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration *nearly doubled* these individuals' likelihood of casting a ballot.
 We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
@@
 
-**EVERYTHING BELOW HERE NEEDS TO BE RE-WORKED. IT DOES NOT DEAL WITH RACIAL DIFFERENCES IN TURNOUT EFFECTS**
-
 ## Discussion {-}
-As discussed above, the effect of incarceration on individuals' propensity to vote is an open question for the literature.
-What is known, however, is that individuals who have been to prison were likely to vote at very low rates prior to incarceration, and continue to vote at very low rates after incarceration.
-Whether incarceration reduces political participation or not, increasing the propensity to vote of individuals who have been to prison is a laudable goal.
-Prisoners overwhelmingly come from marginalized communities with much to gain from policy.
-Executive Order 181 appears to have substantially boosted turnout among formerly incarcerated individuals through two mechanisms.
-The first is obvious: parolees who would not have been eligible to vote absent the Executive Order were allowed to cast ballots in the midterm elections.
-The second mechanism is less obvious.
-Individuals who finished parole between May 18^th^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
-The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
-The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 40 percent.
-
-The precise mechanism through which the rules change increased turnout among these individuals is not clear.
+
+Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both formal and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
+Prior to the Executive Order, parolees were required to wait until they finished their parole term to register to vote.
+That changed in 2018.
+On October 12^th^ --- the registration deadline --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` parolees who had their rights restored.
+Without the Executive Order, every single one of these individuals would have been barred from participating.
+Though turnout was low (just `r comma(readRDS("./temp/new_votes_18.rds"))`, or `r round((readRDS("./temp/new_votes_18.rds")/readRDS("./temp/new_eligible_18.rds"))*100, 2)` percent, of these individuals successfully cast a ballot), their reenfranchisement marks an important milestone for New York State.
+
+As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who were formally reenfranchised.
+In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
+In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` percent.
+This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
+Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
+
+The mechanism through which the rules change increased turnout among these individuals is not clear.
 It is possibly a social mechanism: having an officer of the state affirm one's eligibility (and therefore reunion with the body politic) could be responsible for the increase.
-It could be due to better information: individuals whose rights were restored were likely far more confident of their eligibility to vote than other formerly incarcerated individuals.
-It could also be an issue of timing: individuals whose rights were restored in May of 2018 but were not discharged from parole until early October had four additional months to register to vote than they would have absent the Executive Order.
+Because parole officers were required to provide parolees with a physical voter registration form, the Executive Order likely made registering far more convenient for parolees.
+It could be due to better information: individuals whose rights were restored prior to discharge could be more confident of their eligibility to vote than other formerly incarcerated individuals.
 In reality, the turnout boost is likely due to a combination of different factors.
 
+Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
+Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
+The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
+This discrepancy could be caused by a number of different factors.
+Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
+As discussed above, Black participation among individuals discharged from parole prior to the Executive Order voted at significantly higher rates.
+The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
+It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals just below the cusp.
+
+These differences may arise from differences in the characteristics of Black and White parolees, but the evidence on this front is mixed.
+Some research [@HUEBNER2008] has found that Black incarerated individuals spend longer in prison before being released to parole, though others [@Anwar2015] have failed to find evidence of discrimination in parole board hearings.
+If Black parolees have spent more time in prison relative to similar White individuals, it may reduce their susceptibility to the intervention.
+
+We cannot, however, rule out the possibility that racial discrimination among parole officers plays some role in the effectiveness of the intervention.
+Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
+Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners [@Hemmens2000].
+Recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
+@WHITE2014, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2016].
+Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
+This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
+Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences (although a slightly higher percentage of individuals whose rights were restored under the Executive Order registered as Republicans than individuals discharged from parole earlier, the difference is not statistically significant).
+
 Despite the substantial increase in turnout thanks to Executive Order 181, turnout among formerly incarcerated individuals remained stubbornly low in the 2018 midterm elections.
-Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline.
+Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline turned out.
 Although the successes of Executive Order 181 should be celebrated, more must be done to encourage formerly incarcerated individuals to participate in the political process.
+Future research should pay particular attention to the ways in which the racial identity of formerly incarcerated individuals interacts with interventions meant to increase participation, and ensure that interventions are successful for broad swaths of the formerly disenfranchised population.
 

### 2019-09-25 markdown/p1.Rmd
start part 1

--- 2019-09-06 markdown/felony_disenfranchisement_nys.Rmd
@@
 ---
-title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
+title: "Neighborhoods and Felony Disenfranchisement: The Case of New York CityXX"
 author: "Kevin Morris"
@@
     latex_engine: xelatex
-bibliography: "../research/citations.bib"
+    # includes: 
+    #   in_header: "../code/misc/preamble.tex"
+    keep_tex: false
+bibliography: "citations.bib"
 link-citations: yes
-# fontsize: 12pt
+
 header-includes:
@@
 At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
+Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
 Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
@@
 The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
+Michelle Alexander [-@Alexander2012] and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
 In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
@@
 The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
-Concentrated policing has also credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+Concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
 
-Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
+Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
 The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
@@
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-Filling this gap in the literature is of great importance.
 
-The rest of this paper is structured as follows.
-**Section 2** discusses the historical context of felony disenfranchisement in the United States, and **Section 3** provides an overview of the scholarship conducted in the field over the past two decades.
-**Section 4** provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
-**Section 5** presents the data sources used throughout the project.
+Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
+If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
+If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
+It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
+This is not merely of theoretical concern: research indicates that turnout differentials at the city level has material consequences.
+As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
 
-**Section 6** proposes a new definition of "lost voters" in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised and have a history of participating in elections.
-This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
-A discussion of the results follows.
-
-**Section 7** turns from the 2017 mayoral election to the 2018 midterm elections.
-In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
-I employ an instrumental variables approach test whether rights restoration boosted turnout among formerly incarcerated individuals.
-I focus specifically on individuals who would have been eligible to cast a ballot even if the Executive Order had not gone into effect.
-A discussion of the results follows.
-
-**Section 8** synthesizes the findings presented in the rest of the paper and provides concluding remarks.
-
-# Background of Felony Disenfranchisement in the United States
-
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as *Richardson v.
-Ramirez* (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as `r 6.1 - 1.4` million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
-
-Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, 4; emphasis in the original].
-Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
-In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, 374].
-
-The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
-Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
-In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just `r percent(29932674 / 247219588, accuracy = 0.1)` of the voting age population in 2016, they accounted for `r percent(2228118 / 6106327, accuracy = 0.1)` of the disenfranchised population [@sentencing_2016].
-
-Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January, 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November, 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate [@Morris2019].
-
-# Academic Literature and Felony Disenfranchisement
+## Background
 
@@
 In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that, if this group had been allowed to vote, they would have supported Al Gore by a wide margin.
+They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
 Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
@@
 
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-@Haselswerdt2009 matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [@Meredith2013; @Meredith2015].
-
 A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
 @King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates.
-They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
@@
 Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the impact these laws might have at the local level.
+Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
 @Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
-This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
-
-# What This Paper Examines
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
+This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
 
@@
 It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear whether enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
+We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
 To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
 
-I also offer a refinement of our understanding of who lost voters are.
+I begin by offering a refinement of our understanding of who lost voters are.
 Previous studies exploring the relationship between felony disenfranchisement and spill-over effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
 In this analysis, I use administrative data to identify neighborhoods where individuals with a demonstrated history of participation are disenfranchised.
-
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April, 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I use administrative data to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
-
-The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather than think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
-This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
-The second part of this project seeks to do just that.
+By matching "treated" neighborhoods (neighborhoods with lost voters) to "untreated" neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
 
@@
 Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
 
 Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-@Burch2013, for one, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
+@Burch2013, for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
 Moreover, voting is a social act, and social networks play an important role in predicting political participation [e.g.
@@
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [Gimpel2004; @Cho2006].
+It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
+I therefore hypothesize that felony disenfranchisement does have locally bounded spillover effects, and that turnout in neighborhoods with lost voters systematically vote at a lower rate than others.
 
 Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spill-over effects of lost voters.
-@Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact", [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
+@Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact" [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
 These effects are particularly pronounced for women of color.
+Although Walker and others demonstrate that these actions largely take place outside of the voting booth, they may moderate spillover effects on turnout.
 
@@
 With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
+Understanding how these spillover effects are distributed is of practical importance.
 For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015, for instance, examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
+There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015 examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
 Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
@@
 As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a "cohesive" bloc, their relative turnout rates may influence city policy.
-If felony disenfranchisement has a depressive effect on turnout, politicians are likely less responsive to the particular preferences of voters in these neighborhoods --- some of the most neglected neighborhoods to begin with.
 
-## Turnout Among Formerly Disenfranchised Individuals {-}
-
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
-
-@White2019, however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-The findings in @White2019 agree with previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
-
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts dire consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout, however, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
-
-Some research has been done in this area.
-@Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-@Meredith2013, however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect from notification in the absence of eligibility changes.
-
-@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
-It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
-
-Executive Order 181 is expected to have increased turnout among formerly incarcerated individuals through two primary mechanisms.
-The first addresses the impact of negative experiences with the state.
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel explicitly invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-
-Secondly, Executive Order 181 is expected to dispel confusion about eligibility and serve as a reminder to vote.
-Many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot [@Drucker2005; @locked_out].
-If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.
-Even if confusion does not lead to lower turnout, there is some reason to believe that reminding formerly incarcerated individuals of their rights is a successful intervention for boosting turnout [@Meredith2015; @Gerber2014].
+@Hajnal2005, however, shows that racial variation in turnout has real consequences for local elections and local political power.
+They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
+Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
+In the nation's ten largest cities, @Hajnal2005 argues that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
+Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
+@Hajnal2005 and @Anzia2019 both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
 
@@
 Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It also only includes individuals who are held in state prisons.
+It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
 
@@
 However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.^[This list was compiled by using a webscraper written in the Python language.]
+By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
 
@@
 The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the analysis; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
 
@@
 
-There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
 False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
@@
 @Meredith2013 offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the parole discharge dataset to create false records.
-Comparing the number of matches between these "fake" discharge records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
-This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
+Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
+This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
 The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
@@
 Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
-This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from incarceration.
+This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
 Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
@@
 If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
-At the local level where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition provides insight only into the effect of incarceration on turnout, not felony disenfranchisement.
+At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
 
@@
 As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout through socialization mechanisms.
+Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
 Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
@@
 ```
-The spatial concentration of lost is readily apparent.
+The spatial concentration of lost voters is readily apparent.
 In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
@@
 Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.^[It is worth noting that this definition of lost voter is highly conservative.
-For instance, some individuals who would have voted may have gone to prison after turning 18 but before their first general election.
-These "untreated" neighborhoods are likely home to individuals who would have voted if it were not for disenfranchisement policies.] I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
+Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
+I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
 Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
@@
 Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the voting age population (estimated by the Census Bureau).
+Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
 I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
@@
 
-After matching neighborhoods, I use a simple regression to test whether turnout in the 2017 mayoral election was different in areas with lost voters.
+After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
 Census tract and block group level turnout rates are calculated using the geocoded voter file.
 Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's voting age population.
+The number of ballots cast is divided by the neighborhood's CVAP.
 
@@
 Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match [@Abadie2019].^[The observations in these models are not weighted by voting age population.
-Such weighting materially impacts neither the size or statistical significance of the coefficients of interest.]
+Robust standard errors are clustered at the level of the match [@Abadie2019].
 ```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the dummy identifying neighborhoods with lost voters is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
+When the treatment indicator is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
 The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
@@
 For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more effected.
+The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
+This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
 
@@
 This regression uses the same covariates used in the matching procedure described above.
-Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.
-The observations in these models are not weighted by voting age population.
-Such weighting materially impacts neither the size or statistical significance of the coefficients of interest.]
+Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
 
@@
 Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.018\times Lost\ Voters \times Share\ Black$.
+The estimated depressive effect is $-0.019\times Lost\ Voters \times Share\ Black$.
 
@@
 
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in \ref{fig:citywide-map}.
-The systematic removal of voters from these neighborhoods is troubling.
+During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` voters from New York City neighborhoods.
+The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
 However, more than one million votes were cast in the general election of 2017.
@@
 As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement impacts Black turnout at the state level.
-This analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city.
-The median income of a block groups without lost voters is more than 40 percent higher than the median income of block groups with lost voters.
-Similarly, just 17 percent of individuals in the average block group with a lost voter were white, compared with 40 percent of the population in other block groups.
+Previous literature has established that felony disenfranchisement impacts Black turnout at the state level, but this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
+As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
+Similarly, just 17 percent of individuals in the average block group with a lost voter were White, compared with 40 percent of the population in other block groups.
 Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
@@
 Understanding the magnitude of this finding is important.
-In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population.
-On average, the voting age population in these neighborhoods was `r round(readRDS("./temp/vap_gt_90_black"))`.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a voting-age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population, with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
 Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
@@
 @Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their governments and their identities as citizens.
-It is possible that Black communities understand the loss of voters differently than non-Black communities, and therefore respond differently.
+It is possible that Black communities interpret the loss of voters differently than non-Black communities, and therefore respond differently.
 
@@
 This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
-In the case of the 2017, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-This does not mean, however, that the depressed turnout in these neighborhoods had no impact on their representation.
+In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
+A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement has no implications for who is elected to local office.
+As @Hajnal2005 shows, racial turnout differentials can have real consequences for city politics.
+
+Moreover, we cannot conclude that depressed turnout in these neighborhoods in 2017 had no impact on their representation.
 City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
+Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals *held* and *used* their power.
 
@@
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-
-\pagebreak
-
-# Executive Order 181 and Turnout in 2018
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].^[I've also talked with folks on probation whose probation officers told them they were disenfranchised.
-Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
-As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
-To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the policy change may increase the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-Under the Executive Order, however, parole officers were required to explicitly inform their supervisees of their newly restored voting rights.
-It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly re-eligible voters to cast a ballot.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
-These are the individuals who would have been eligible to cast a ballot whether or not the Executive Order was signed.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, we can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
-
-## Identifying Parolees Whose Rights Were Restored {-}
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
-
-## Trends in Turnout {-}
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
-A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2010 through April, 2016, and extended through October, 2016.
-
-```{r to-16-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-16}Turnout in 2016 Presidential Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to16_chart.rds")
-print(j)
-```
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2012 through April, 2018, and extended through October, 2018.
-```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart.rds")
-print(j)
-```
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that Executive Order increased turnout.
-
-## Individual-Level Turnout Regressions {-}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the Executive Order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
-
-```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart_recent.rds")
-print(j)
-```
-Formalizing this chart into an individual-level logit model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
-
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F}
-load("./temp/short_term_to.rdata")
-
-stargazer(date_to_model_notime, date_to_model_time,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
-                               "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit-short} Individual-Level Logit Models",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
-                         c("Felony Class FE", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(date_to_model_notime, type = "HC0"))),
-            sqrt(diag(vcovHC(date_to_model_time, type = "HC0"))))
-          )
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-write.table(j, "./temp/table3.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../temp/table3.tex"}
-The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
-
-Although Governor Cuomo signed the Executive Order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18^th^ that there were consistently parolees who had their voting rights restored prior to discharge.
-```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/restoration_plot.rds")
-print(j)
-```
-
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the Executive Order are part of the "control group", while individuals discharged after are "treated" by the policy change.
-Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
-
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
-
-```{r logit-reg-chunk, echo = F, results = "hide", message = F}
-load("./temp/individual_turnout_18_itt.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit} Individual-Level Logit Models",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../temp/table4.tex"}
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Each model also indicates that individuals discharged from parole after the Executive Order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
-
-## Instrumental Variables Approach {-}
-
-Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole.
-
-Investigating the causal effect of rights restoration using the natural experiment conceptualizes introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied."
-
-An individual's propensity to vote can be expressed using the following equation:
-$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} $$
-Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
-*Z* is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
-Neither of these assumptions are valid: an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored: the Executive Order required that all eligible individuals have their voting rights restored.
-
-There is also reason to believe that an individual's likelihood of having his rights restored is correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the Executive Order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-The available data does not indicate whether individuals were eligible to have their rights restored.
-We can, however, examine the demographic differences between individuals who had their rights restored and those who did not.
-
-```{r restoree-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_restorees.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
-\\\\Shows demographics for individuals discharged from parole \\\\between May 18th, 2018, and October 12th, 2018.}}}"
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex), n = 16.5)) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[3] <- lab
-
-write.table(j, "./temp/table_whatever.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women.
-Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that *b~1~* is probably correlated with *Y*.
-A one-stage approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach [@Angrist1996].
-Such an approach has been widely used in the context of voter turnout [e.g.
-@Ansolabehere1999; @Gerber2000; @Milligan2004;  @Lassen2004; @Sondheimer2010].
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the Executive Order, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the "local average treatment effect." The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
-Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research [e.g.
-@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (IV probit) model to avoid the constraints of the nonparametric framework.
-The IV probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable.
-
-The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the IV probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
-\input{"../temp/iv_clean.tex"}
-
-The bivariate probit model returns modestly more conservative estimates than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored prior to discharge by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored prior to parole discharge cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
-
-### Racial Variation {-}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of parolees.
-These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the biprobit models are calculated at the means of the other variables.
-
-\input{"../temp/iv_clean_race.tex"}
-
-```{r nameless-chunk, echo = F, results = "asis", message = F}
-load("./temp/iv_individual_turnout_18.rdata")
-```
-
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
-
-Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
-Some of this may be explained by different propensities to vote.
-@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote, they may be less "susceptible" to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logit models estimating turnout rates of all individuals discharged from parole prior to the Executive Order going into effect, and the turnout rates of all individuals who had their rights restored.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate), these individuals have been excluded from Models 2 and 4.
-
-```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/black_to.rdata")
-
-stargazer(black_to1, black_to2, black_to3, black_to4,
-          column.separate = c(2, 2),
-          column.labels = c("Discharged Prior to EO", "Had Rights Restored"),
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Black)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:black-to} Turnout Rate Among Former Parolees",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          omit = "felony",
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(black_to1, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to2, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to3, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to4, type = "HC0")))
-          ))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.
-\\\\Latino individuals excluded from Models 2 and 4.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
-
-j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table5.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, 6,066 (or 43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
-
-Table \ref{tab:black-to} also shows, however, that the turnout rate among Black individuals who had their rights restored were not statistically significantly higher (or lower) than other individuals who had their rights restored.
-The Executive Order, therefore, seems to have had a leveling effect --- Black parolees were not longer substantially more likely to turnout than other individuals discharged from parole in New York State.
-
-### Variable Treatment Intensity {-}
-
-There is some reason to believe that using a constant treatment variable understates the true impact of Executive Order 181.
-Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
-This may be due to a number of factors: parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} interacts an individual's restoration status with the number of months they spent on parole after the policy went into effect ("Months Restored").
-This variable is once again instrumented using a variable indicating whether the individual was discharged before or after the Executive Order went into effect.
-Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in the Table \ref{tab:iv-models-time}.
-
-\input{"../temp/iv_clean_months.tex"}
-
-Table \ref{tab:iv-models-time} indicate that there may be variable treatment intensity --- that individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-The two-stage least squares model indicates that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.28 percentage points (in the 2SLS model) and 0.30 percentage points (in the 2S Probit model).
-Put another way: these models indicate that rights restoration increased the propensity to vote for individuals released within one month of the policy being implemented by `r round((0.28 / (3.13589 - 0.28)) * 100, digits = 1)` percent.
-It increased the propensity to vote among those released five months after the Executive Order went into effect by `r round(((0.28 * 5) / (3.7702 - (0.28 * 5))) * 100, digits = 1)` percent.
-
-Of course, with very few months between the implementation of the Executive Order and the registration deadline for the 2018 midterms, it is impossible to know the true impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
-Incorporation of data after the 2020 presidential election will likely shed more light on the Executive Order's impact on turnout among formerly disenfrnachised individuals.
-
-**EVERYTHING BELOW HERE NEEDS TO BE RE-WORKED. IT DOES NOT DEAL WITH RACIAL DIFFERENCES IN TURNOUT EFFECTS**
-
-## Discussion {-}
-As discussed above, the effect of incarceration on individuals' propensity to vote is an open question for the literature.
-What is known, however, is that individuals who have been to prison were likely to vote at very low rates prior to incarceration, and continue to vote at very low rates after incarceration.
-Whether incarceration reduces political participation or not, increasing the propensity to vote of individuals who have been to prison is a laudable goal.
-Prisoners overwhelmingly come from marginalized communities with much to gain from policy.
-Executive Order 181 appears to have substantially boosted turnout among formerly incarcerated individuals through two mechanisms.
-The first is obvious: parolees who would not have been eligible to vote absent the Executive Order were allowed to cast ballots in the midterm elections.
-The second mechanism is less obvious.
-Individuals who finished parole between May 18^th^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
-The EO 181, however, appears to have boosted turnout even among individuals whose eligibility to vote was not directly impacted by the rules change.
-The increase in turnout was substantial: Executive Order 181 increased turnout among individuals who had their rights restored by more than 40 percent.
-
-The precise mechanism through which the rules change increased turnout among these individuals is not clear.
-It is possibly a social mechanism: having an officer of the state affirm one's eligibility (and therefore reunion with the body politic) could be responsible for the increase.
-It could be due to better information: individuals whose rights were restored were likely far more confident of their eligibility to vote than other formerly incarcerated individuals.
-It could also be an issue of timing: individuals whose rights were restored in May of 2018 but were not discharged from parole until early October had four additional months to register to vote than they would have absent the Executive Order.
-In reality, the turnout boost is likely due to a combination of different factors.
-
-Despite the substantial increase in turnout thanks to Executive Order 181, turnout among formerly incarcerated individuals remained stubbornly low in the 2018 midterm elections.
-Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline.
-Although the successes of Executive Order 181 should be celebrated, more must be done to encourage formerly incarcerated individuals to participate in the political process.
-
-# Conclusion
-
-\newpage
-# References {-}
-<div id="refs"></div>
-
-\newpage
-# Appendix A {-}
-
-In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-
-```{r logit-reg-chunk-16, echo = F, results = "hide", message = F}
-load("./temp/individual_turnout_16.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
-                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))
-          ))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-write.table(j, "./temp/table55.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../temp/table55.tex"}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
+The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchise all combine to undermine the political power of marginalized communities in New York City.

### 2019-09-25 markdown/felony_disenfranchisement_nys.Rmd
write

@@
     keep_tex: false
-bibliography: "../research/citations.bib"
+bibliography: "citations.bib"
 link-citations: yes
@@
 The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
-Concentrated policing has also credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
-
-Felony disenfranchisement policies are part of an interlocking criminal justice system that disproportionately impacts Black Americans living in certain communities.
+Concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+
+Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
 The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
@@
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-Filling this gap in the literature is of great importance considering the spatial concentration of policing networks.
-
-The rest of this paper is structured as follows.
-**Section 2** discusses the historical context of felony disenfranchisement in the United States, and **Section 3** provides an overview of the scholarship conducted in the field over the past two decades.
+
+Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
+If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
+If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
+It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
+This is not merely of theoretical concern: research indicates that turnout differentials at the city level has material consequences.
+As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
+
+The rest of this paper is structured as follows: **Section 2** discusses the historical context of felony disenfranchisement in the United States, and **Section 3** provides an overview of the scholarship conducted in the field over the past two decades.
 **Section 4** provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
@@
 In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
-I employ an instrumental variables approach test whether rights restoration boosted turnout among formerly incarcerated individuals.
+I employ an instrumental variables approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
 I focus specifically on individuals who would have been eligible to cast a ballot even if the Executive Order had not gone into effect.
@@
 Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" [@Burch2010, 4; emphasis in the original].
-Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies and that these policies were often adopted during Jim Crow to reduce the political power of Black Americans [@Behrens2003].
-In Florida, for instance, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been added to prevent the state from being "niggerized" [@florida_1868, 374].
+As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (Burch 2010, 4; emphasis in the original).
+Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans [@Behrens2003].
+In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
+Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being "niggerized" [@florida_1868, 374].
 
@@
 In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just `r percent(29932674 / 247219588, accuracy = 0.1)` of the voting age population in 2016, they accounted for `r percent(2228118 / 6106327, accuracy = 0.1)` of the disenfranchised population [@sentencing_2016].
+Although Black adults made up just `r round(29932674 * 100 / 247219588, 1)` percent of the voting age population in 2016, they accounted for `r round(2228118 * 100 / 6106327, 1)` percent of the disenfranchised population [@sentencing_2016].
 
@@
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-@Haselswerdt2009 matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
+@Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
 Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
@@
 Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [@Meredith2013; @Meredith2015].
+Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [e.g.
+@Meredith2013; @Meredith2015].
 
@@
 @King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates.
-They include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
@@
 Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, less work has been done exploring the impact these laws might have at the local level.
+Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
 @Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block-group level turnout and involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
-This paper seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
+This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
 
@@
 It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear whether enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
+We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
 To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
 
-I also offer a refinement of our understanding of who lost voters are.
+I begin by offering a refinement of our understanding of who lost voters are.
 Previous studies exploring the relationship between felony disenfranchisement and spill-over effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
 In this analysis, I use administrative data to identify neighborhoods where individuals with a demonstrated history of participation are disenfranchised.
+By matching "treated" neighborhoods (neighborhoods with lost voters) to "untreated" neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
 
 After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April, 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
+In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
+Today, only those currently in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
 In theory, such a policy could increase turnout and registration for parolees.
@@
 As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-An extension of her paper, however, implies that what she finds might be probitive to the effect of neighborhood turnout rates.
-Insofar as neighborhoods vote as a "cohesive" bloc, their relative turnout rates may influence city policy.
-If felony disenfranchisement has a depressive effect on turnout, politicians are likely less responsive to the particular preferences of voters in these neighborhoods --- some of the most neglected neighborhoods to begin with.
+
+@Hajnal2005, however, shows that racial variation in turnout has real consequences for local elections and local political power.
+They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
+Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
+In the nation's ten largest cities, @Hajnal2005 argues that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
+Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
+@Hajnal2005 and @Anzia2019 both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
 
@@
 This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
 Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-The findings in @White2019 agree with previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
 
 Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts dire consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout, however, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
+Whether or not incarceration *causes* low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
 Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
 
-Some research has been done in this area.
+The literature on turnout among formerly incarcerated individuals leads me to believe that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
+I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
+
+Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
+Some research has been done indicating that dispelling misinformation boosts turnout among the formerly disenfranchised.
 @Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
@@
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect from notification in the absence of eligibility changes.
-
+@Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
 @Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
 "Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
-It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
-
-Executive Order 181 is expected to have increased turnout among formerly incarcerated individuals through two primary mechanisms.
-The first addresses the impact of negative experiences with the state.
+Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
+
+Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
+@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
 To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel explicitly invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-
-Secondly, Executive Order 181 is expected to dispel confusion about eligibility and serve as a reminder to vote.
-Many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot [@Drucker2005; @locked_out].
-If confusion about eligibility leads to lower turnout, parole officers informing their parolees of their rights is likely to increase turnout.
-Even if confusion does not lead to lower turnout, there is some reason to believe that reminding formerly incarcerated individuals of their rights is a successful intervention for boosting turnout [@Meredith2015; @Gerber2014].
+Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
+Such repairs may lead to increased propensity to participate in elections in the future.
 
@@
 
+### Toward a Causal Estimation {-}
+
+As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
+The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
+If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
+As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
+
+I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
+This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
+These individuals home neighborhoods are removed from the subsequent analysis.
+I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
+These neighborhoods, therefore, were *newly* home to lost voters.
+If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
+In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and included here.
+The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
+
+Table \ref{tab:did-match} presents the results of matching on this subset of block groups. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+
+```{r block-table-chunk-did, echo = F}
+t <- readRDS("./temp/match_table_block_group_did.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+j
+```
+
+In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups.
+D(2017) indicates that, across the board, turnout in 2017 was far lower than in 2016.
+D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
+The coefficient is positive and statistically significant, indicating that prior to treatment, neighborhoods with lost voters had *higher* turnout rates than other neighborhoods.
+D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
+This coefficient is statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
+Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.2 percentage points.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
+These neighborhoods may differ from neighborhoods that frequently have lost voters.
+
+```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
+load("./temp/match_reg_block_group_did.rdata")
+
+bg_reg <- reg_output
+bg_reg_ses <- reg_output_ses
+
+load("./temp/match_reg_tract_did.rdata")
+
+tract_reg <- reg_output
+tract_reg_ses <- reg_output_ses
+
+stargazer(bg_reg,
+          header = F,
+          se = list(bg_reg_ses),
+          notes = c("TO REPLACE"),
+          type = "latex", notes.align = "l",
+          notes.append = T,
+          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
+          omit = "district",
+          dep.var.labels = c("Turnout Rate"),
+          title = "\\label{tab:did-reg} Block Group Turnout in 2016 and 2017 General Elections",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test_abc.tex",
+          out.header = F,
+          add.lines=list(c("Congressional District FE", "X")))
+
+j <- fread("./temp/test_abc.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+j <- j %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table1111.tex"}
+
+Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
+If political participation in neighborhoods that are home to lost voters systematically differed from turnout in other neighborhoods, they would have been unlikely to turnout at the same rate as other neighborhoods in 2016.
+Clearly, this is not the case: turnout in treated neighborhoods was not lower relative to other neighborhoods prior to the removal of a voter.
+
 ## Discussion {-}
 
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in \ref{fig:citywide-map}.
-The systematic removal of voters from these neighborhoods is troubling.
+During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` voters from New York City neighborhoods.
+The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
 However, more than one million votes were cast in the general election of 2017.
@@
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
+The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchise all combine to undermine the political power of marginalized communities in New York City.
 
@@
 Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].^[I've also talked with folks on probation whose probation officers told them they were disenfranchised.
-Not sure if it's appropriate to mention anecdotes like these, or to just cite to the literature] The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
@@
 As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
-To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the policy change may increase the political participation of formerly disenfranchised individuals.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters [see, for instance, @Fraga2018].
+To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
+
+There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-Under the Executive Order, however, parole officers were required to explicitly inform their supervisees of their newly restored voting rights.
-It is possible that having a representative of the government tell parolees of their rights was an effective encouragement for these newly re-eligible voters to cast a ballot.
+New York's correction code provides no more guidance other than that "upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration" (Section 75 of the New York State Correction Law).
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[Link: https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
+In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" (emphasis in the original).
+Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
 
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
-These are the individuals who would have been eligible to cast a ballot whether or not the Executive Order was signed.
+This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
+senatorial contests (as well as other down-ballot races).
+
+I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
 This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, we can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
+However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
 
@@
 It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
+However, as Figure \ref{fig:to-16} (which plots turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
 The longer an individual has been off of parole, the more likely he is to cast a ballot.
 For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
-A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2010 through April, 2016, and extended through October, 2016.
+Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
+This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
 
@@
 Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2012 through April, 2018, and extended through October, 2018.
+This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
 ```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
@@
 Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- were more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that Executive Order increased turnout.
+Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- *were* more likely to participate than those discharged earlier in the year.
+However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
 
@@
 The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the Executive Order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
+For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
 Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
@@
 \input{"../temp/table3.tex"}
-The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
-
-Although Governor Cuomo signed the Executive Order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
+The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
+A Chi-squared test confirms that the model is not improved when controls for time are included.
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.]
+
+Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
 Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18^th^ that there were consistently parolees who had their voting rights restored prior to discharge.
+It is not until May 18^th^ that parolees consistently had their voting rights restored prior to discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the Executive Order are part of the "control group", while individuals discharged after are "treated" by the policy change.
-Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
+Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are "treated" by the policy change.
+Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
@@
 Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
+Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-Each model also indicates that individuals discharged from parole after the Executive Order went into effect were more likely to cast a ballot than those discharged earlier.
+Each model also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
 Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
@@
 
-Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole.
-
-Investigating the causal effect of rights restoration using the natural experiment conceptualizes introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied."
+Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
+Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote specifically for the individuals who received the treatment.
+To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether they were discharged after the policy change.
+
+Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied." Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
 
@@
 Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
-Neither of these assumptions are valid: an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored: the Executive Order required that all eligible individuals have their voting rights restored.
-
-There is also reason to believe that an individual's likelihood of having his rights restored is correlated with his natural propensity to vote.
+Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
+
+Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
+Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
 Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the Executive Order went into effect.
+citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
 This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-The available data does not indicate whether individuals were eligible to have their rights restored.
-We can, however, examine the demographic differences between individuals who had their rights restored and those who did not.
+Although the available data does not indicate whether individuals were eligible to have their rights restored, we can examine the demographic differences between individuals who had their rights restored and those who did not.
 
@@
 Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women.
-Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
+Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
 Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
 This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that *b~1~* is probably correlated with *Y*.
-A one-stage approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
+Table \ref{tab:demos-restorees} indicates that *b~1~* is very probably correlated with *Y*.
+An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
 
@@
 @Ansolabehere1999; @Gerber2000; @Milligan2004;  @Lassen2004; @Sondheimer2010].
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
+A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
 
@@
 Some research [e.g.
-@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (IV probit) model to avoid the constraints of the nonparametric framework.
-The IV probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable.
-
+@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
+The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2014].
 The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
@@
 Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the IV probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
+Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
 For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
@@
 
-The bivariate probit model returns modestly more conservative estimates than the two-stage least squares model.
+The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
 After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored prior to discharge by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
@@
 Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of parolees.
+In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of former parolees.
 These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
@@
 She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote, they may be less "susceptible" to a policy intervention of this sort.
+The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logit models estimating turnout rates of all individuals discharged from parole prior to the Executive Order going into effect, and the turnout rates of all individuals who had their rights restored.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate), these individuals have been excluded from Models 2 and 4.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole prior to the executive order going into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
 
@@
 
-stargazer(black_to1, black_to2, black_to3, black_to4,
-          column.separate = c(2, 2),
-          column.labels = c("Discharged Prior to EO", "Had Rights Restored"),
+stargazer(black_to1, black_to2, black_to1b, black_to2b, black_to3, black_to4,
+          column.separate = c(2, 2, 2),
+          column.labels = c("Discharged Prior to EO", "ITT Group", "Treatment Group"),
           header = F,
@@
             sqrt(diag(vcovHC(black_to2, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to1b, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to2b, type = "HC0"))),
             sqrt(diag(vcovHC(black_to3, type = "HC0"))),
@@
 
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+note.latex <- "\\multicolumn{7}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
 \\\\Robust standard errors in parentheses.
-\\\\Latino individuals excluded from Models 2 and 4.}}}"
+\\\\Latino individuals excluded from Models 2, 4, and 6.}}}"
 
@@
 
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}  \\cmidrule(l{3pt}r{3pt}){6-7}"
 
@@
 \input{"../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, 6,066 (or 43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote than the rest of the population.
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43) were Black.
+These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
 Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
@@
 
-Table \ref{tab:black-to} also shows, however, that the turnout rate among Black individuals who had their rights restored were not statistically significantly higher (or lower) than other individuals who had their rights restored.
-The Executive Order, therefore, seems to have had a leveling effect --- Black parolees were not longer substantially more likely to turnout than other individuals discharged from parole in New York State.
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+The intervention, therefore, appears to have a leveling effect.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
 
@@
 
-There is some reason to believe that using a constant treatment variable understates the true impact of Executive Order 181.
+There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
 Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
-This may be due to a number of factors: parole officers may have had longer to understand and communicate the new rules.
+This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
 Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} interacts an individual's restoration status with the number of months they spent on parole after the policy went into effect ("Months Restored").
-This variable is once again instrumented using a variable indicating whether the individual was discharged before or after the Executive Order went into effect.
-Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in the Table \ref{tab:iv-models-time}.
+Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
+If the individual did not have voting rights restored, this variable takes the value 0.
+This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
+It is coded as 0 for individuals discharged prior to the implementation of the executive order.^[Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.]
 
@@
 
-Table \ref{tab:iv-models-time} indicate that there may be variable treatment intensity --- that individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-The two-stage least squares model indicates that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.28 percentage points (in the 2SLS model) and 0.30 percentage points (in the 2S Probit model).
-Put another way: these models indicate that rights restoration increased the propensity to vote for individuals released within one month of the policy being implemented by `r round((0.28 / (3.13589 - 0.28)) * 100, digits = 1)` percent.
-It increased the propensity to vote among those released five months after the Executive Order went into effect by `r round(((0.28 * 5) / (3.7702 - (0.28 * 5))) * 100, digits = 1)` percent.
+Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- that is to say, that individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
+The two-stage least squares model indicates that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
+It is worth noting that the *R^2^* in Model 1 is higher than the *R^2^* in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
 
@@
 One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
-Incorporation of data after the 2020 presidential election will likely shed more light on the Executive Order's impact on turnout among formerly disenfrnachised individuals.
 
@@
 Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both formal and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
-Prior to the Executive Order, parolees were required to wait until they finished their parole term to register to vote.
+Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
 That changed in 2018.
-On October 12^th^ --- the registration deadline --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` parolees who had their rights restored.
+On October 12^th^ --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
 Without the Executive Order, every single one of these individuals would have been barred from participating.
@@
 The mechanism through which the rules change increased turnout among these individuals is not clear.
-It is possibly a social mechanism: having an officer of the state affirm one's eligibility (and therefore reunion with the body politic) could be responsible for the increase.
-Because parole officers were required to provide parolees with a physical voter registration form, the Executive Order likely made registering far more convenient for parolees.
-It could be due to better information: individuals whose rights were restored prior to discharge could be more confident of their eligibility to vote than other formerly incarcerated individuals.
-In reality, the turnout boost is likely due to a combination of different factors.
+@Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
+Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
+On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+Such interactions may undo some of the negative socialization identified by @Weaver2010.
+
+It could also be a story of better information.
+As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation.
+Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
+In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
 
@@
 The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals just below the cusp.
-
-These differences may arise from differences in the characteristics of Black and White parolees, but the evidence on this front is mixed.
-Some research [@HUEBNER2008] has found that Black incarerated individuals spend longer in prison before being released to parole, though others [@Anwar2015] have failed to find evidence of discrimination in parole board hearings.
-If Black parolees have spent more time in prison relative to similar White individuals, it may reduce their susceptibility to the intervention.
-
-We cannot, however, rule out the possibility that racial discrimination among parole officers plays some role in the effectiveness of the intervention.
+It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
+
+We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
 Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
 Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners [@Hemmens2000].
-Recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
+Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
 @WHITE2014, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
 There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2016].
+
 Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
 This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences (although a slightly higher percentage of individuals whose rights were restored under the Executive Order registered as Republicans than individuals discharged from parole earlier, the difference is not statistically significant).
-
-Despite the substantial increase in turnout thanks to Executive Order 181, turnout among formerly incarcerated individuals remained stubbornly low in the 2018 midterm elections.
-Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline turned out.
-Although the successes of Executive Order 181 should be celebrated, more must be done to encourage formerly incarcerated individuals to participate in the political process.
-Future research should pay particular attention to the ways in which the racial identity of formerly incarcerated individuals interacts with interventions meant to increase participation, and ensure that interventions are successful for broad swaths of the formerly disenfranchised population.
+Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
+In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
+Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
+
+Re-enfranchising voters while they are still under formal supervision is obviously beneficial to the individuals who are on parole on election day; such policies allow them to make their voices heard.
+The case of Executive Order 181 also indicates that restoring voting rights prior to parole discharge has further benefits.
+In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
+The state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
+By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
 
@@
 
-\newpage
-# References {-}
-<div id="refs"></div>
+Felony disenfranchisement was originally implemented to limit the franchise of Black Americans.
+It continues to serve that purpose today, as Black residents of New York and states around the country are disproportionately convicted of felony offenses and imprisoned.
+Although we know that many of these individuals were unlikely to cast a ballot even if they had not gone to prison, the targeted disenfranchisement of a single segment of the population is deeply troubling.
+
+As this project has demonstrated, however, the racialized aspects of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
+The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these rules.
+By both redefining "lost voters" as individuals who have a demonstrated propensity of participation and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
+What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, but this effect is concentrated in Black neighborhoods.
+Moreover, there is reason to believe that this is not driven by unobservable differences between neighborhoods; I also show that turnout decreased more between 2016 and 2017 in neighborhoods that newly lost voters than in neighborhoods unaffected by disenfranchisement policies.
+
+I also offer an early investigation of the effect of Executive Order 181, which granted voting rights to most citizens on parole in New York State.
+This research indicates that individuals who have their rights restored prior to being discharged from parole are more likely to participate in elections, even after they are discharged.
+This is a welcome finding: increasing the political participation of marginalized individuals brings us toward a fuller democracy.
+As with neighborhood spillover effects, however, this effect depends in great deal on the race of the parolee: rights restoration, at least in the first five months of the program in New York State, did not increase future participation among non-Whites.
+More research needs to be done to investigate whether this is true over a longer time period in New York, and whether it is true in other states as well.
+
+America is becoming ever-more-diverse, and will likely be less than 50 percent non-Hispanic White by the middle of the century.
+As Fraga [-@Fraga2018] has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
+How districts are drawn, population patterns, and rules about who can fundraise how will continue to shape the politicians we elect and the policies they enact.
+Breaking down *de jure* (like formal disenfranchisement) and *de facto* (like neighborhood spillover effects) barriers to the ballot box is of great importance.
+The experience in New York State over the past few years offers a better understanding of just how big this problem is, but it also offers evidence of a way forward.
 
@@
 The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
+
+\newpage
+# References {-}
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}

### 2019-09-26 markdown/felony_disenfranchisement_nys.Rmd
final edits (pre-sending)

@@
 ---
-title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
+title: "In the Shadow of Jim Crow"
+subtitle: "Felony Disenfranchisement and New York State"
 author: "Kevin Morris"
@@
     #   in_header: "../code/misc/preamble.tex"
-    keep_tex: false
+    keep_tex: true
 bibliography: "citations.bib"
 link-citations: yes
-
+fontsize: 12pt
 header-includes:
@@
 \pagenumbering{gobble}
+
 \tableofcontents
@@
 Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
+Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised [@sentencing_2016].
@@
 Michelle Alexander [-@Alexander2012] and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
 
@@
 Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
-Concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008], while concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
 
@@
 The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
-The neighborhood-specific implications of felony disenfranchisement, however, largely remain unstudied.
+The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
 A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2014; @Burch2011].
@@
 It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level has material consequences.
+This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
 As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
@@
 **Section 6** proposes a new definition of "lost voters" in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised and have a history of participating in elections.
+I define lost voters as individuals who are not only disenfranchised, but are disenfranchised *and* have a history of participating in elections.
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
@@
 I employ an instrumental variables approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
-I focus specifically on individuals who would have been eligible to cast a ballot even if the Executive Order had not gone into effect.
+I focus specifically on individuals who would have been eligible to cast a ballot even if the executive order had not gone into effect.
 A discussion of the results follows.
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
+In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government [@bcj_laws].
 This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Traci Burch has explained, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (Burch 2010, 4; emphasis in the original).
-Previous research has established that the presence of nonwhite potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans [@Behrens2003].
+As @Burch2010 explains, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (4; emphasis in the original).
+Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans [@Behrens2003].
 In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
@@
 Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January, 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November, 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
+Prior to January 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
+In November 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
 In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate [@Morris2019].
@@
 A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
@@
 @Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
 This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
@@
 
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
+This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City mayoral election of 2017 using individual-level administrative data.
 Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods [e.g.
@@
 I begin by offering a refinement of our understanding of who lost voters are.
-Previous studies exploring the relationship between felony disenfranchisement and spill-over effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a demonstrated history of participation are disenfranchised.
+Previous studies exploring the relationship between felony disenfranchisement and spillover effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
+In this analysis, I use administrative data to identify neighborhoods where individuals with a *demonstrated history* of participation are disenfranchised.
 By matching "treated" neighborhoods (neighborhoods with lost voters) to "untreated" neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
@@
 
-Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
+Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
 @Burch2013, for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
@@
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
+It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spill-over effects of lost voters.
+Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
 @Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact" [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
@@
 
-Even if neighborhoods to not have unique preferences for presidential candidates, lower turnout in impacted communities may reduce their allocation of public goods.
+Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
 At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
@@
 
-@Hajnal2005, however, shows that racial variation in turnout has real consequences for local elections and local political power.
+@Hajnal2005 and @Hajnal2009, however, show that racial variation in turnout has real consequences for local elections and local political power.
 They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
 Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-In the nation's ten largest cities, @Hajnal2005 argues that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
+@Hajnal2005 tells us that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
 Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
@@
 In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
+They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
 
@@
 
-The literature on turnout among formerly incarcerated individuals leads me to believe that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
+The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
 I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
@@
 Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
-Some research has been done indicating that dispelling misinformation boosts turnout among the formerly disenfranchised.
+Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
 @Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
@@
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and when they finished their parole supervision.
+These data are used to determine when individuals were incarcerated or on parole and demographic characteristics.
 
 The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
+It does not make information about individuals sentenced to probation or incarcerated for misdemeanors available.
 Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
@@
 ## Voter File Data {-}
-Most states in the United States are required to maintain files with information on all registered voters.
+Most states in the United States are required to maintain files with information on all registered and recently purged voters.
 In New York, this information is publicly available from the Board of Elections.
@@
 The New York State Voter File is unique in its treatment of "purged" voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
+I can therefore identify voters who were registered (and participated) in the past but have since been purged due to a felony conviction.
 
@@
 Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
 This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
@@
   kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
+                position = "center",
+                font_size = 10)
 j
@@
 Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting to felony disenfranchisement.
-More than two-thirds of block groups that lost a voter lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
+Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
+More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
 
@@
   row_spec(14, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"))
+  kable_styling(latex_options = c("HOLD_position"),
+                font_size = 10)
 j
@@
 These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less white, have much lower median incomes, and a larger share of voters are registered as Democrats.
+Neighborhoods with lost voters are far less White,^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
+Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] have much lower median incomes, and a larger share of voters are registered as Democrats.
 
@@
 insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
-
-j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is Non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
+When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
 The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))*100, digits = 1)` percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$vap))), digits = 1)` votes went uncast.
+Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))*100, digits = 1)` percent, this effect is alarmingly high.
+For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))), digits = 1)` votes went uncast.
 The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
@@
 This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
-These individuals home neighborhoods are removed from the subsequent analysis.
+These individuals' home neighborhoods are removed from the subsequent analysis.
 I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
@@
 If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and included here.
+In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
 The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
 
-Table \ref{tab:did-match} presents the results of matching on this subset of block groups. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
+As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
 
@@
   add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+  kable_styling(latex_options = c("scale_down", "HOLD_position"),
+                font_size = 10)
 j
@@
 
-In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups.
-D(2017) indicates that, across the board, turnout in 2017 was far lower than in 2016.
+```{r tract-chunk-did, echo = F}
+t <- readRDS("./temp/match_table_tract_did.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match-tract) Results of Tract-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"),
+                font_size = 10)
+j
+```
+
+In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups and tracts.
+In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
 D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-The coefficient is positive and statistically significant, indicating that prior to treatment, neighborhoods with lost voters had *higher* turnout rates than other neighborhoods.
+In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly *higher* turnout prior to intervention.
 D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-This coefficient is statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.2 percentage points.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
-These neighborhoods may differ from neighborhoods that frequently have lost voters.
+In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
+Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
+In Model 2, D(Treat) $\times$ D(2017) is negative but not statistically significant.
+However, because treated tracts had significantly higher turnout than control tracts in 2016, this nonsignificance implies treatment --- the loss of a voter --- did decrease turnout to a significant degree relative to the control tracts.
 
@@
 
-stargazer(bg_reg,
+stargazer(bg_reg, tract_reg,
           header = F,
-          se = list(bg_reg_ses),
+          column.labels = c("Block Group Level", "Tract Level"),
+          se = list(bg_reg_ses, tract_reg_ses),
           notes = c("TO REPLACE"),
@@
           dep.var.labels = c("Turnout Rate"),
-          title = "\\label{tab:did-reg} Block Group Turnout in 2016 and 2017 General Elections",
+          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
           table.placement = "H",
@@
           out.header = F,
-          add.lines=list(c("Congressional District FE", "X")))
+          add.lines=list(c("Congressional District FE", "X", "X")))
 
@@
 
-note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
 
@@
 
-j <- j %>% 
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-2} \\cmidrule(l{3pt}r{3pt}){3-3}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-If political participation in neighborhoods that are home to lost voters systematically differed from turnout in other neighborhoods, they would have been unlikely to turnout at the same rate as other neighborhoods in 2016.
-Clearly, this is not the case: turnout in treated neighborhoods was not lower relative to other neighborhoods prior to the removal of a voter.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
+These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
+Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Unfortunately, because many neighborhoods consistently have lost voters, testing the causal effect of the first lost voter for all neighborhoods is not possible.
+Nevertheless, this analysis lends credence to the causal nature of lost voters on depressed turnout.
 
@@
 However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have enormous implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
+Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
 
@@
 Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is Non-Hispanic Black, the lost voter indicator is either insignificant or significant and positive.
+Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
 This means that lost voters have no spillover effects in neighborhoods with small Black populations.
@@
 Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future.
+Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large "interpretive effects" [@Pierson1993].
 @Lerman2013, for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-@Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their governments and their identities as citizens.
-It is possible that Black communities interpret the loss of voters differently than non-Black communities, and therefore respond differently.
+@Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
+Similarly, @Weaver2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
+It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
 
@@
 @Browning1992; @Hurwitz2005; @Henderson1997; @Wu2009].
-Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to effect voting patterns) in Black neighborhoods.
+Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
+There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
 
@@
 In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement has no implications for who is elected to local office.
+A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
 As @Hajnal2005 shows, racial turnout differentials can have real consequences for city politics.
@@
 Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and nonwhite Americans.
+Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchise all combine to undermine the political power of marginalized communities in New York City.
+The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
 
@@
 The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
-The Executive Order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
@@
 As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
-As Traci Burch [-@Burch2011] and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
 Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters [see, for instance, @Fraga2018].
@@
 Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" (emphasis in the original).
+In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" [emphasis in the original].
 Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
@@
 This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement to cast a ballot for this population.
+However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
 
@@
 Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
@@
 It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
+However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
 The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+For instance, of the individuals last discharged from parole in 2010, `r round(readRDS("./temp/turnout_in_16_finished_10.rds") * 100, digits = 1)` percent cast a ballot in the 2016 election, while just `r round(readRDS("./temp/turnout_in_16_finished_15.rds") * 100, digits = 1)` percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
 Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
@@
 ```
-Formalizing this chart into an individual-level logit model demonstrates that time since discharge is not correlated with turnout in the short run.
+Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
 The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
 
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F}
+```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
 load("./temp/short_term_to.rdata")
@@
          V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
 
@@
 
-```{r logit-reg-chunk, echo = F, results = "hide", message = F}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
 load("./temp/individual_turnout_18_itt.rdata")
@@
 
-Each model also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model2[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
 
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote specifically for the individuals who received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether they were discharged after the policy change.
+Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
 
@@
 lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex), n = 16.5)) %>% 
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 
-j$V1[3] <- lab
+j$V1[4] <- lab
 
@@
 As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the Executive Order, therefore, satisfies the criteria for an instrumental variable.
+A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
 
@@
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored prior to discharge by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
+After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored prior to parole discharge cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
+Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
 
@@
 Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and biprobit models on subsets of former parolees.
+In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
 These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the biprobit models are calculated at the means of the other variables.
+Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
 
@@
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole prior to the executive order going into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
 Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
@@
 insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}  \\cmidrule(l{3pt}r{3pt}){6-7}"
-
-j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 \input{"../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43) were Black.
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
 These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
@@
 
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- that is to say, that individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-The two-stage least squares model indicates that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
+Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
+Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
 It is worth noting that the *R^2^* in Model 1 is higher than the *R^2^* in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
 
-Of course, with very few months between the implementation of the Executive Order and the registration deadline for the 2018 midterms, it is impossible to know the true impact more months on parole after having one's rights restored has.
+Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
 One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
@@
 
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both formal and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
+Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both *de jure* and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
 Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
@@
 On October 12^th^ --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
-Without the Executive Order, every single one of these individuals would have been barred from participating.
-Though turnout was low (just `r comma(readRDS("./temp/new_votes_18.rds"))`, or `r round((readRDS("./temp/new_votes_18.rds")/readRDS("./temp/new_eligible_18.rds"))*100, 2)` percent, of these individuals successfully cast a ballot), their reenfranchisement marks an important milestone for New York State.
-
-As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who were formally reenfranchised.
+Without the executive order, every single one of these individuals would have been barred from participating.
+Though turnout among this group was low (just `r comma(readRDS("./temp/new_votes_18.rds"))`, or `r round((readRDS("./temp/new_votes_18.rds")/readRDS("./temp/new_eligible_18.rds"))*100, 2)` percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
+
+As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
 In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
@@
 It could also be a story of better information.
-As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation.
+As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
 Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
@@
 Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the Executive Order voted at significantly higher rates.
+As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
 The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
@@
 In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
-The state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
+This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
 By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
@@
 
-As this project has demonstrated, however, the racialized aspects of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
-The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these rules.
-By both redefining "lost voters" as individuals who have a demonstrated propensity of participation and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
-What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, but this effect is concentrated in Black neighborhoods.
+As this project has demonstrated, however, the racialized penalties of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
+The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these policies.
+By both redefining "lost voters" as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
+What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, and this effect is concentrated in Black neighborhoods.
 Moreover, there is reason to believe that this is not driven by unobservable differences between neighborhoods; I also show that turnout decreased more between 2016 and 2017 in neighborhoods that newly lost voters than in neighborhoods unaffected by disenfranchisement policies.
@@
 This research indicates that individuals who have their rights restored prior to being discharged from parole are more likely to participate in elections, even after they are discharged.
-This is a welcome finding: increasing the political participation of marginalized individuals brings us toward a fuller democracy.
+This is a welcome finding, as increasing the political participation of marginalized individuals brings us toward a fuller democracy.
 As with neighborhood spillover effects, however, this effect depends in great deal on the race of the parolee: rights restoration, at least in the first five months of the program in New York State, did not increase future participation among non-Whites.
@@
 
-```{r logit-reg-chunk-16, echo = F, results = "hide", message = F}
+```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
 load("./temp/individual_turnout_16.rdata")
@@
 \setlength{\leftskip}{0.5in}
+\noindent

### 2019-09-26 markdown/felony_disenfranchisement_nys.tex
final edits (pre-sending)

--- 2019-09-26 markdown/felony_disenfranchisement_nys.Rmd
@@
----
-title: "In the Shadow of Jim Crow"
-subtitle: "Felony Disenfranchisement and New York State"
-author: "Kevin Morris"
-date: "`r format(Sys.time(), '%B %d, %Y')`"
-output:
-  bookdown::pdf_document2:
-    toc: no
-    fig_caption: yes
-    latex_engine: xelatex
-    # includes: 
-    #   in_header: "../code/misc/preamble.tex"
-    keep_tex: true
-bibliography: "citations.bib"
-link-citations: yes
-fontsize: 12pt
-header-includes:
-    - \usepackage{setspace}\doublespacing
-    - \usepackage{rotating}
-
----
-<!-- --- -->
-<!-- title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City" -->
-<!-- author: "Kevin Morris" -->
-<!-- date: "`r format(Sys.time(), '%B %d, %Y')`" -->
-<!-- output: word_document -->
-<!-- bibliography: "../research/citations.bib" -->
-<!-- link-citations: yes -->
-<!-- always_allow_html: yes -->
-<!-- --- -->
+\documentclass[12pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={In the Shadow of Jim Crow},
+            pdfauthor={Kevin Morris},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{In the Shadow of Jim Crow}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \subtitle{Felony Disenfranchisement and New York State}
+  \author{Kevin Morris}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{September 26, 2019}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\usepackage{setspace}\doublespacing
+\usepackage{rotating}
+
+\begin{document}
+\maketitle
+
 \pagenumbering{gobble}
@@
 \pagebreak
-```{r setup, echo = F, message = F, warning = F}
-knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf")
-library(systemfit)
-library(AER)
-library(sandwich)
-library(lmtest)
-library(scales)
-library(tidyverse)
-library(extrafont)
-library(kableExtra)
-library(stargazer)
-library(data.table)
-loadfonts()
-```
+
 \pagenumbering{arabic}
-# Introduction
-
-The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage [see, for instance, @Keyssar2009].
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
 At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
 Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised [@sentencing_2016].
+Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
 
 The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-Michelle Alexander [-@Alexander2012] and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
+Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
+In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
 
@@
 Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008], while concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+Residents of such neighborhoods suffer from worse physical health (Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}) and are more likely to suffer from anxiety and exhibit symptoms of trauma (Geller et al.
+\protect\hyperlink{ref-Geller2014}{2014}).
+The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted (Clear \protect\hyperlink{ref-Clear2008}{2008}), while concentrated policing has also been credited with having a ``chilling effect'' on neighborhoods' willingness to reach out for help to local governments (Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
 
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2014; @Burch2011].
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
-With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}; Gerber et al.
+\protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
+Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
+With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
 
@@
 This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
-
-The rest of this paper is structured as follows: **Section 2** discusses the historical context of felony disenfranchisement in the United States, and **Section 3** provides an overview of the scholarship conducted in the field over the past two decades.
-**Section 4** provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
-**Section 5** presents the data sources used throughout the project.
-
-**Section 6** proposes a new definition of "lost voters" in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised *and* have a history of participating in elections.
+As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (8).
+
+The rest of this paper is structured as follows: \textbf{Section 2} discusses the historical context of felony disenfranchisement in the United States, and \textbf{Section 3} provides an overview of the scholarship conducted in the field over the past two decades.
+\textbf{Section 4} provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
+\textbf{Section 5} presents the data sources used throughout the project.
+
+\textbf{Section 6} proposes a new definition of ``lost voters'' in the context of felony disenfranchisement.
+I define lost voters as individuals who are not only disenfranchised, but are disenfranchised \emph{and} have a history of participating in elections.
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
 Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
+The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely \emph{understated} the degree to which they are impacted by spatially concentrated patterns of criminal justice.
 A discussion of the results follows.
 
-**Section 7** turns from the 2017 mayoral election to the 2018 midterm elections.
+\textbf{Section 7} turns from the 2017 mayoral election to the 2018 midterm elections.
 In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
@@
 
-**Section 8** synthesizes the findings presented in the rest of the paper and provides concluding remarks.
-
-# Background of Felony Disenfranchisement in the United States
+\textbf{Section 8} synthesizes the findings presented in the rest of the paper and provides concluding remarks.
+
+\hypertarget{background-of-felony-disenfranchisement-in-the-united-states}{%
+\section{Background of Felony Disenfranchisement in the United States}\label{background-of-felony-disenfranchisement-in-the-united-states}}
 
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as *Richardson v.
-Ramirez* (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as `r 6.1 - 1.4` million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
+In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
+The Supreme Court, in cases such as \emph{Richardson v.
+Ramirez} (1974), has upheld states' right to do just that.
+Collectively, these laws disenfranchise as many as 4.7 million American citizens.
+Of these, the majority are no longer incarcerated, but are living and working in their communities (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
 
 Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As @Burch2010 explains, "If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were *intended* to" (4; emphasis in the original).
-Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans [@Behrens2003].
+As Burch (\protect\hyperlink{ref-Burch2010}{2010}) explains, ``If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were \emph{intended} to'' (4; emphasis in the original).
+Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans (Behrens, Uggen, and Manza \protect\hyperlink{ref-Behrens2003}{2003}).
 In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being "niggerized" [@florida_1868, 374].
+Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being ``niggerized'' (Shofner \protect\hyperlink{ref-florida_1868}{1963}, 374).
 
@@
 In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just `r round(29932674 * 100 / 247219588, 1)` percent of the voting age population in 2016, they accounted for `r round(2228118 * 100 / 6106327, 1)` percent of the disenfranchised population [@sentencing_2016].
+Although Black adults made up just 12.1 percent of the voting age population in 2016, they accounted for 36.5 percent of the disenfranchised population (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
 
@@
 In November 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate [@Morris2019].
-
-# Academic Literature and Felony Disenfranchisement
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
+In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate (Morris \protect\hyperlink{ref-Morris2019}{2019}).
+
+\hypertarget{academic-literature-and-felony-disenfranchisement}{%
+\section{Academic Literature and Felony Disenfranchisement}\label{academic-literature-and-felony-disenfranchisement}}
+
+In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
 George W.
@@
 Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
-Others have used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout (Miles \protect\hyperlink{ref-Miles2004}{2004}; Hjalmarsson and Lopez \protect\hyperlink{ref-Hjalmarsson2010}{2010}).
+Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects (Miles \protect\hyperlink{ref-Miles2004}{2004}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
+Others have used survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
 
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-@Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
+Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
+Traci Burch (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}) expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
 She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
 Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [e.g.
-@Meredith2013; @Meredith2015].
+Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
 
 A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
+King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006].
-As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
+These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
+As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
+{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (724).
+
+Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the \emph{state} level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
 Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-@Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
-
-# What This Paper Examines
+Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
+This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
+
+\hypertarget{what-this-paper-examines}{%
+\section{What This Paper Examines}\label{what-this-paper-examines}}
 
 This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods [e.g.
-@Sewell2016; @Clear2008; @Lerman2013].
+Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods (e.g.~Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}; Clear \protect\hyperlink{ref-Clear2008}{2008}; Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
 It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
@@
 Previous studies exploring the relationship between felony disenfranchisement and spillover effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a *demonstrated history* of participation are disenfranchised.
-By matching "treated" neighborhoods (neighborhoods with lost voters) to "untreated" neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
+In this analysis, I use administrative data to identify neighborhoods where individuals with a \emph{demonstrated history} of participation are disenfranchised.
+By matching ``treated'' neighborhoods (neighborhoods with lost voters) to ``untreated'' neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
 
 After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.^[Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those currently in prison for felony offenses are disenfranchised.] Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
+In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.\footnote{Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
+Today, only those currently in prison for felony offenses are disenfranchised.} Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
 In theory, such a policy could increase turnout and registration for parolees.
@@
 
-## Framing for Turnout Effects {-}
+\hypertarget{framing-for-turnout-effects}{%
+\subsection*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
+\addcontentsline{toc}{subsection}{Framing for Turnout Effects}
 
 Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
+In New York, for instance, 46,232 individuals were imprisoned in New York State in early 2019, compared with 11.6 million actively registered voters.
 Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-
-Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-@Burch2013, for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
-Moreover, voting is a social act, and social networks play an important role in predicting political participation [e.g.
-@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
+
+Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
+Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
+Moreover, voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
+Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
+It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
-@Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact" [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
+Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
 These effects are particularly pronounced for women of color.
@@
 
-### Theoretical Implications for Reduced Turnout {-}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
-@King2016; @Bowers2009].
-With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
+\hypertarget{theoretical-implications-for-reduced-turnout}{%
+\subsubsection*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
+\addcontentsline{toc}{subsubsection}{Theoretical Implications for Reduced Turnout}
+
+As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community (e.g.~King and Erickson \protect\hyperlink{ref-King2016}{2016}; Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}).
+With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
 Understanding how these spillover effects are distributed is of practical importance.
 For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015 examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
+There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: Kinsella, McTague, and Raleigh (\protect\hyperlink{ref-Kinsella2015}{2015}) examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
 Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
@@
 Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
+At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes (Martin \protect\hyperlink{ref-Martin2003}{2003}).
 Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (59).
-@Griffin2005 finds similar effects in the United States Senate, demonstrating that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
+``{[}H{]}igher citizen participation is rewarded,'' Martin and Claibourn (\protect\hyperlink{ref-Martin2013}{2013}) concludes, ``with enhanced policy responsiveness'' (59).
+Griffin and Newman (\protect\hyperlink{ref-Griffin2005}{2005}) finds similar effects in the United States Senate, demonstrating that ``voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not'' (1206).
 If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
@@
 
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group" (1).
+Although research on the impact of local turnout on city-wide policy is scarce, Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) examines the impact of senior turnout on ``senior-friendly'' policy at the city level.
+Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials \emph{are} responsive to senior turnout when seniors ``are a more cohesive, meaningful group'' (1).
 American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-
-@Hajnal2005 and @Hajnal2009, however, show that racial variation in turnout has real consequences for local elections and local political power.
+As such, Anzia's study does not speak directly to the impact of \emph{neighborhood} turnout rates on city policy.
+
+Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}), however, show that racial variation in turnout has real consequences for local elections and local political power.
 They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
 Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-@Hajnal2005 tells us that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
+Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) tells us that ``Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils'' (518).
 Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-@Hajnal2005 and @Anzia2019 both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
-
-## Turnout Among Formerly Disenfranchised Individuals {-}
-
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
+Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
+
+\hypertarget{turnout-among-formerly-disenfranchised-individuals}{%
+\subsection*{Turnout Among Formerly Disenfranchised Individuals}\label{turnout-among-formerly-disenfranchised-individuals}}
+\addcontentsline{toc}{subsection}{Turnout Among Formerly Disenfranchised Individuals}
+
+Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so (Haselswerdt \protect\hyperlink{ref-Haselswerdt2009}{2009}; Burch \protect\hyperlink{ref-Burch2011}{2011}; Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}).
 The causal effect of incarceration on participation is the subject of some debate within the field.
@@
 Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
-
-@White2019, however, comes to a different conclusion.
+In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
+(\protect\hyperlink{ref-Gerber2017}{2017}) uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
+They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that ``it appears that spending time in prison does not have large negative effects on subsequent participation'' (1144).
+
+White (\protect\hyperlink{ref-White2019}{2019}), however, comes to a different conclusion.
 Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+This finding does not necessarily conflict with Gerber et al.
+(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
 By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
@@
 A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Whether or not incarceration \emph{causes} low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
 Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
@@
 
-Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
+Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2006}).
 Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
-@Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
 They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-@Meredith2013, however, examines states where so-called notification laws went into effect.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
-@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
+Gerber et al.
+(\protect\hyperlink{ref-Gerber2014}{2014}) conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+``Whatever the participatory consequences of incarceration,'' they conclude, ``they are not in large part impossible to overcome'' (924).
 Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
 
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
+Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
+Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that ``contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual'' (829).
 To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
@@
 
-# Data
-## Criminal Justice Data {-}
+\hypertarget{data}{%
+\section{Data}\label{data}}
+
+\hypertarget{criminal-justice-data}{%
+\subsection*{Criminal Justice Data}\label{criminal-justice-data}}
+\addcontentsline{toc}{subsection}{Criminal Justice Data}
 
@@
 
-## Voter File Data {-}
+\hypertarget{voter-file-data}{%
+\subsection*{Voter File Data}\label{voter-file-data}}
+\addcontentsline{toc}{subsection}{Voter File Data}
+
 Most states in the United States are required to maintain files with information on all registered and recently purged voters.
@@
 It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The first section of this paper uses a snapshot of the registered voter file from April 30^th^, 2018.
-The second section of this paper uses a snapshot from March 3^rd^, 2019.
-
-The New York State Voter File is unique in its treatment of "purged" voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
+The first section of this paper uses a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
+The second section of this paper uses a snapshot from March 3\textsuperscript{rd}, 2019.
+
+The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
 I can therefore identify voters who were registered (and participated) in the past but have since been purged due to a felony conviction.
 
-## Geocoding {-}
+\hypertarget{geocoding}{%
+\subsection*{Geocoding}\label{geocoding}}
+\addcontentsline{toc}{subsection}{Geocoding}
+
 Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
 I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
+This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
+The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6 percent of these individuals were not geocoded.
 Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
 
-## Matching {-}
+\hypertarget{matching}{%
+\subsection*{Matching}\label{matching}}
+\addcontentsline{toc}{subsection}{Matching}
+
 Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
 I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a "match," records must have the exact same birth date.
+To be considered a ``match,'' records must have the exact same birth date.
 The first and last names must also be exact matches (conditional on the adjustments discussed below).
@@
 
-* Middle names are identical. If neither set of records includes a middle name, this condition is met.
-* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-* A middle name or middle initial in one set of records, and a missing middle name in the other set.
-
-Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andrew Doe" and "John Anthony Doe" would not.
+\begin{itemize}
+\tightlist
+\item
+  Middle names are identical. If neither set of records includes a middle name, this condition is met.
+\item
+  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
+\item
+  A middle name or middle initial in one set of records, and a missing middle name in the other set.
+\end{itemize}
+
+Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
 
@@
 Testing for the presence of false positive matches is fairly straightforward.
-@Meredith2013 offers one way to test their prevalence using placebo matching.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
 I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
+Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
 Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
@@
 
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center",
-                font_size = 10)
-j
-
-```
+\begin{table}[H]
+
+\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
+\centering
+\fontsize{10}{12}\selectfont
+\begin{tabular}{cc}
+\toprule
+Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
+\midrule
+Actual Birthdate & 20,955\\
+Birthdate + 35 Days & 105\\
+Birthdate - 35 Days & 92\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Testing for false negatives is more challenging.
 If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
+Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
 To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
+A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
 Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
 This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
-Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
-
-# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels
-
-## Identification of Lost Voters {-}
-In this analysis, I offer a different definition of "lost voter" than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
+Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
+
+\hypertarget{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}{%
+\section{Effects of Felony Disenfranchisement on Neighborhood Turnout Levels}\label{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}}
+
+\hypertarget{identification-of-lost-voters}{%
+\subsection*{Identification of Lost Voters}\label{identification-of-lost-voters}}
+\addcontentsline{toc}{subsection}{Identification of Lost Voters}
+
+In this analysis, I offer a different definition of ``lost voter'' than much of the literature.
+Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- \emph{potentially} lost voters --- and turnout.
 Such an approach is informative for understanding the impact of disenfranchisement.
@@
 Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration [e.g.
-@Gerber2017].
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
+As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration (e.g.~Gerber et al.
+\protect\hyperlink{ref-Gerber2017}{2017}).
+If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement \emph{per se} but rather to their incarceration.
 At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
@@
 Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote [@Gerber2003].
+In this analysis, ``lost voters'' are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
+Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote (Gerber, Green, and Shachar \protect\hyperlink{ref-Gerber2003}{2003}).
 As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
@@
 
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
+To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
+Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
 
 Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
-
-```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
-j <- readRDS("./output/city_map.RDS")
-print(j)
-```
+There were 2,518 such lost voters within New York City as of the 2017 general election, and 6,166 statewide.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/citywide-map-1}
+
+}
+
+\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
+\end{figure}
+
 The spatial concentration of lost voters is readily apparent.
@@
 
-## Testing for Neighborhood Turnout Effects {-}
-```{r table-table, , echo = F, message = F, warning = F}
-
-```
+\hypertarget{testing-for-neighborhood-turnout-effects}{%
+\subsection*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
+\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Effects}
+
 Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
@@
 
-```{r distribution-lost, echo = F, message = F, warning = F}
-t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
-  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
-
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
-                                                 align = "c") %>% 
-  row_spec(14, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                font_size = 10)
-j
-```
+\begin{table}[H]
+
+\caption{\label{tab:distribution-lost}\label{tab:nhood-dist} Distribution of Lost Voters}
+\centering
+\fontsize{10}{12}\selectfont
+\begin{tabular}{ccc}
+\toprule
+Number of Lost Voters & Count of Block Groups & Count of Tracts\\
+\midrule
+0 & 4,272 & 1,008\\
+1 & 1,023 & 416\\
+2 & 301 & 212\\
+3 & 107 & 138\\
+4 & 30 & 51\\
+5 & 19 & 46\\
+6 & 9 & 23\\
+7 & 3 & 15\\
+8 & 5 & 13\\
+9 & 0 & 7\\
+10 & 0 & 2\\
+11 & 0 & 2\\
+12 & 0 & 2\\
+14 & 0 & 2\\
+\hline
+Total & 5,769 & 1,937\\
+\bottomrule
+\end{tabular}
+\end{table}
 
 Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau.
+Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
+I use a genetic match to match treated to untreated census tracts and block groups (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}), using a series of demographic and political indicators.
+Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau.
 Party affiliation rates come from the geocoded voter file.
 Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
+I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
 Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
@@
 
-### Match Output {-}
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-
-```{r tract-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_tract.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:tract-table) Results of Tract-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
+\hypertarget{match-output}{%
+\subsubsection*{Match Output}\label{match-output}}
+\addcontentsline{toc}{subsubsection}{Match Output}
+
+\begin{table}[H]
+
+\caption{\label{tab:block-table-chunk}\label{tab:blockgroup-table} Results of Block Group-Level Matching}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllllrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
+ & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
+\midrule
+\% Latino & 0.35 & 0.25 & 0.35 & 0.34 & 93.02 & 89.13 & 85.79 & 79.01\\
+\% Non-Hispanic Black & 0.38 & 0.16 & 0.38 & 0.37 & 97.93 & 94.46 & 94.41 & 92.77\\
+\% Non-Hispanic White & 0.17 & 0.40 & 0.17 & 0.18 & 96.37 & 96.06 & 95.56 & 92.99\\
+Median Income & 51,792.52 & 72,612.36 & 51,792.52 & 52,768.97 & 95.31 & 93.47 & 88.62 & 74.49\\
+\% With Some College & 0.64 & 0.70 & 0.64 & 0.64 & 90.75 & 94.14 & 90.29 & 80.70\\
+Median Age & 35.99 & 38.34 & 35.99 & 36.25 & 89.18 & 90.84 & 85.93 & 78.56\\
+Registration Rate & 1.00 & 0.94 & 1.00 & 0.98 & 59.77 & 79.89 & 72.28 & 60.42\\
+\% Democrats & 0.75 & 0.65 & 0.75 & 0.75 & 96.22 & 96.52 & 94.92 & 91.02\\
+\% Noncitizen & 0.16 & 0.16 & 0.16 & 0.16 & 16.58 & 14.80 & -0.06 & -60.28\\
+\% Won by City Council Representative & 0.85 & 0.81 & 0.85 & 0.86 & 89.44 & 83.40 & 79.88 & 70.11\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:tract-table-chunk}\label{tab:tract-table} Results of Tract-Level Matching}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllllrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
+ & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
+\midrule
+\% Latino & 0.32 & 0.22 & 0.32 & 0.33 & 98.55 & 85.12 & 82.84 & 73.01\\
+\% Non-Hispanic Black & 0.36 & 0.14 & 0.36 & 0.36 & 99.77 & 88.74 & 88.36 & 83.30\\
+\% Non-Hispanic White & 0.19 & 0.42 & 0.19 & 0.19 & 97.50 & 96.32 & 91.29 & 79.95\\
+Median Income & 53,409.42 & 71,754.50 & 53,409.42 & 55,422.41 & 89.03 & 90.83 & 84.86 & 62.84\\
+\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 90.19 & 93.52 & 88.89 & 73.01\\
+Median Age & 35.76 & 38.04 & 35.76 & 35.80 & 98.12 & 74.51 & 80.56 & 71.91\\
+Registration Rate & 0.94 & 0.90 & 0.94 & 0.93 & 76.63 & 86.16 & 84.24 & 71.28\\
+\% Democrats & 0.74 & 0.62 & 0.74 & 0.73 & 94.58 & 94.93 & 92.05 & 82.37\\
+\% Noncitizen & 0.16 & 0.17 & 0.16 & 0.17 & -123.57 & 22.48 & 8.85 & -38.79\\
+\% Won by City Council Representative & 0.85 & 0.79 & 0.85 & 0.85 & 93.03 & 81.57 & 79.27 & 67.47\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
 At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
 These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less White,^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] have much lower median incomes, and a larger share of voters are registered as Democrats.
+Neighborhoods with lost voters are far less White,\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
+Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} have much lower median incomes, and a larger share of voters are registered as Democrats.
 
@@
 Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match [@Abadie2019].
-```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group.rdata")
-
-load("./temp/match_reg_block_group_nhb.rdata")
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-bg_reg_black <- reg_output_nhblack
-bg_reg_black_ses <- reg_output_nhblack_ses
-
-load("./temp/match_reg_tract.rdata")
-
-load("./temp/match_reg_tract_nhb.rdata")
-tract_reg <- reg_output
-tract_reg_ses <- reg_output_ses
-
-tract_reg_black <- reg_output_nhblack
-tract_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
-          header = F,
-          se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          title = "\\label{tab:match-reg} Matching Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F)
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1.tex", quote = F, col.names = F,
-            row.names = F)
-```
+Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
+
 \input{"../temp/table1.tex"}
@@
 When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average `r -1*round(bg_reg[["coefficients"]][["treatTRUE"]]*100, digits = 1)` percentage points lower than in comparable block groups without lost voters.
+In block groups with lost voters, turnout was on average 0.9 percentage points lower than in comparable block groups without lost voters.
 This decrease, however, appears to be entirely concentrated within Black neighborhoods.
 When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))*100, digits = 1)` percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))), digits = 1)` votes went uncast.
+The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as 2.7 percentage points lower than similar neighborhoods without lost voters.
+Considering that the overall turnout rate in block groups with a lost voter was just 14.6 percent, this effect is alarmingly high.
+For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
 The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
@@
 
-### Testing Intensity Effects {-}
+\hypertarget{testing-intensity-effects}{%
+\subsubsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
+\addcontentsline{toc}{subsubsection}{Testing Intensity Effects}
 
 Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
+The models above do not allow for different effects on turnout based on \emph{how many} voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
 This regression uses the same covariates used in the matching procedure described above.
-Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
-
-```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/tract_model_reg_ols.rdata")
-load("./temp/bg_model_reg_ols.rdata")
-
-stargazer(bg_model, bg_model2, tract_model, tract_model2,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
-          header = F,
-          type = "latex",
-          notes.align = "l",
-          notes.append = T,
-          se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          title = "\\label{tab:trad-reg} Standard Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = "TO REPLACE",
-          omit = "boro",
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F)
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table2.tex", quote = F, col.names = F,
-            row.names = F)
-```
+Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
+
 \input{"../temp/table2.tex"}
@@
 The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
+Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
+In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.87 percentage points.
 The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
@@
 They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.019\times Lost\ Voters \times Share\ Black$.
-
-```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/dec_block_map.rds")
-print(j)
-```
-
-### Toward a Causal Estimation {-}
+Figure \ref{fig:postest} applies the coefficient on \(Lost\ Voters \times Share\ Black\) from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
+The estimated depressive effect is \(-0.019\times Lost\ Voters \times Share\ Black\).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/postest-map-1}
+
+}
+
+\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
+\end{figure}
+
+\hypertarget{toward-a-causal-estimation}{%
+\subsubsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
+\addcontentsline{toc}{subsubsection}{Toward a Causal Estimation}
 
@@
 I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
+This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
 These individuals' home neighborhoods are removed from the subsequent analysis.
 I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were *newly* home to lost voters.
+These neighborhoods, therefore, were \emph{newly} home to lost voters.
 If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
 In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
+The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 \emph{did} have a lost voter in 2016, and are thus excluded from this analysis.
 
@@
 
-```{r block-table-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_block_group_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
-
-```{r tract-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_tract_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match-tract) Results of Tract-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
+\begin{table}[H]
+
+\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{10}{12}\selectfont
+\begin{tabular}{lllllrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
+ & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
+\midrule
+\% Latino & 0.35 & 0.25 & 0.35 & 0.35 & 99.00 & 96.26 & 94.92 & 85.60\\
+\% Non-Hispanic Black & 0.33 & 0.16 & 0.33 & 0.33 & 99.02 & 96.57 & 95.67 & 90.10\\
+\% Non-Hispanic White & 0.21 & 0.41 & 0.21 & 0.21 & 99.61 & 94.66 & 92.61 & 82.94\\
+Median Income & 55,852.62 & 73,433.49 & 55,852.62 & 56,287.76 & 97.52 & 87.85 & 86.56 & 72.89\\
+\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 99.31 & 92.25 & 87.30 & 71.15\\
+Median Age & 36.17 & 38.47 & 36.17 & 36.04 & 94.37 & 62.35 & 72.27 & 69.84\\
+Registration Rate & 0.98 & 0.94 & 0.98 & 0.95 & 25.09 & 46.49 & 42.16 & 33.06\\
+\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 95.94 & 94.99 & 93.62 & 86.68\\
+\% Noncitizen & 0.17 & 0.16 & 0.17 & 0.17 & 63.46 & 75.01 & 66.85 & 45.71\\
+\% Won by City Council Representative & 0.85 & 0.80 & 0.85 & 0.85 & 88.19 & 83.99 & 78.23 & 67.54\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\begin{table}[H]
+
+\caption{\label{tab:tract-chunk-did}\label{tab:did-match-tract} Results of Tract-Level Matching}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{10}{12}\selectfont
+\begin{tabular}{lllllrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
+ & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
+\midrule
+\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 96.18 & 85.46 & 85.00 & 76.98\\
+\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.24 & 97.15 & 88.52 & 85.34 & 77.02\\
+\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.20 & 87.45 & 79.41 & 64.95\\
+Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,711.89 & 86.31 & 76.32 & 76.77 & 65.49\\
+\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 95.15 & 88.35 & 84.14 & 75.44\\
+Median Age & 36.36 & 38.23 & 36.36 & 36.38 & 98.90 & 69.64 & 65.76 & 40.69\\
+Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 8.97 & 63.21 & 60.34 & 50.97\\
+\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.01 & 87.99 & 85.70 & 73.48\\
+\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 83.92 & 77.72 & 65.69 & 47.73\\
+\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.99 & 77.50 & 67.39 & 41.31\\
+\bottomrule
+\end{tabular}}
+\end{table}
 
@@
 D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly *higher* turnout prior to intervention.
-D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
+In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly \emph{higher} turnout prior to intervention.
+D(Treat) \(\times\) D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
 In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
 Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-In Model 2, D(Treat) $\times$ D(2017) is negative but not statistically significant.
+In Model 2, D(Treat) \(\times\) D(2017) is negative but not statistically significant.
 However, because treated tracts had significantly higher turnout than control tracts in 2016, this nonsignificance implies treatment --- the loss of a voter --- did decrease turnout to a significant degree relative to the control tracts.
 
-```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group_did.rdata")
-
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-load("./temp/match_reg_tract_did.rdata")
-
-tract_reg <- reg_output
-tract_reg_ses <- reg_output_ses
-
-stargazer(bg_reg, tract_reg,
-          header = F,
-          column.labels = c("Block Group Level", "Tract Level"),
-          se = list(bg_reg_ses, tract_reg_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
-          omit = "district",
-          dep.var.labels = c("Turnout Rate"),
-          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test_abc.tex",
-          out.header = F,
-          add.lines=list(c("Congressional District FE", "X", "X")))
-
-j <- fread("./temp/test_abc.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-2} \\cmidrule(l{3pt}r{3pt}){3-3}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
-            row.names = F)
-```
 \input{"../temp/table1111.tex"}
@@
 Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Unfortunately, because many neighborhoods consistently have lost voters, testing the causal effect of the first lost voter for all neighborhoods is not possible.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Unfortunately, because many neighborhoods consistently have lost voters, testing the causal effect of the first lost voter for all neighborhoods is not possible.
 Nevertheless, this analysis lends credence to the causal nature of lost voters on depressed turnout.
 
-## Discussion {-}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` voters from New York City neighborhoods.
+\hypertarget{discussion}{%
+\subsection*{Discussion}\label{discussion}}
+\addcontentsline{toc}{subsection}{Discussion}
+
+During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated 2,518 voters from New York City neighborhoods.
 The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
@@
 Understanding the magnitude of this finding is important.
-In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population, with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood `r round(0.9 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+In New York City, there are 165 block groups where the Black community makes up more than 90 percent of the population, with an average of 821 voting age citizens.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of 821, having any lost voter reduced turnout by 20 ballots.
+Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just 2.2 ballots.
+
+The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood 13.9 ballots.
+In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just 1.5 ballots --- hardly more votes than the lost voter herself.
 
 Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large "interpretive effects" [@Pierson1993].
-@Lerman2013, for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-@Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, @Weaver2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
+Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large ``interpretive effects'' (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
+Lerman and Weaver (\protect\hyperlink{ref-Lerman2013}{2013}), for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
+Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
+Similarly, Weaver and Lerman (\protect\hyperlink{ref-Weaver2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
 It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
@@
 
-This finding mirrors @White2019, which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
 This is perhaps unsurprising.
-A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans [e.g.
-@Browning1992; @Hurwitz2005; @Henderson1997; @Wu2009].
+A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans (e.g.~Browning and Cao \protect\hyperlink{ref-Browning1992}{1992}; Hurwitz and Peffley \protect\hyperlink{ref-Hurwitz2005}{2005}; Henderson et al.
+\protect\hyperlink{ref-Henderson1997}{1997}; Wu, Sun, and Triplett \protect\hyperlink{ref-Wu2009}{2009}).
 Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
@@
 
-For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods [@Wilson1990].
+For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods (Wilson \protect\hyperlink{ref-Wilson1990}{1990}).
 More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
@@
 A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
-As @Hajnal2005 shows, racial turnout differentials can have real consequences for city politics.
+As Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) shows, racial turnout differentials can have real consequences for city politics.
 
@@
 City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
-Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals *held* and *used* their power.
+Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals \emph{held} and \emph{used} their power.
 
@@
 
-# Executive Order 181 and Turnout in 2018
+\hypertarget{executive-order-181-and-turnout-in-2018}{%
+\section{Executive Order 181 and Turnout in 2018}\label{executive-order-181-and-turnout-in-2018}}
 
@@
 
-On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
 Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
 The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
@@
 Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
-As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters [see, for instance, @Fraga2018].
-To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
+As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
+To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
 
@@
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that "upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration" (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[Link: https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
 Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" [emphasis in the original].
+In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
 Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018.
 This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
@@
 
-## Identifying Parolees Whose Rights Were Restored {-}
+\hypertarget{identifying-parolees-whose-rights-were-restored}{%
+\subsection*{Identifying Parolees Whose Rights Were Restored}\label{identifying-parolees-whose-rights-were-restored}}
+\addcontentsline{toc}{subsection}{Identifying Parolees Whose Rights Were Restored}
+
 Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
-
-## Trends in Turnout {-}
+By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
+For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
+For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.}
+
+\hypertarget{trends-in-turnout}{%
+\subsection*{Trends in Turnout}\label{trends-in-turnout}}
+\addcontentsline{toc}{subsection}{Trends in Turnout}
+
 Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
@@
 The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, `r round(readRDS("./temp/turnout_in_16_finished_10.rds") * 100, digits = 1)` percent cast a ballot in the 2016 election, while just `r round(readRDS("./temp/turnout_in_16_finished_15.rds") * 100, digits = 1)` percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
+For instance, of the individuals last discharged from parole in 2010, 6.6 percent cast a ballot in the 2016 election, while just 4.2 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
 This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
 
-```{r to-16-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-16}Turnout in 2016 Presidential Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to16_chart.rds")
-print(j)
-```
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-16-chart-1}
+
+}
+
+\caption{\label{fig:to-16}Turnout in 2016 Presidential Election}\label{fig:to-16-chart}
+\end{figure}
+
 Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
@@
 This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
-```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart.rds")
-print(j)
-```
+
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-chart-1}
+
+}
+
+\caption{\label{fig:to-18}Turnout in 2018 Midterm Election}\label{fig:to-18-chart}
+\end{figure}
+
 Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- *were* more likely to participate than those discharged earlier in the year.
+Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- \emph{were} more likely to participate than those discharged earlier in the year.
 However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
 
-## Individual-Level Turnout Regressions {-}
+\hypertarget{individual-level-turnout-regressions}{%
+\subsection*{Individual-Level Turnout Regressions}\label{individual-level-turnout-regressions}}
+\addcontentsline{toc}{subsection}{Individual-Level Turnout Regressions}
 
@@
 The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
+For a true natural experiment to hold, an individual's probability of being ``assigned'' to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
 Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
@@
 
-```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart_recent.rds")
-print(j)
-```
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-recent-chart-1}
+
+}
+
+\caption{\label{fig:to-18-recent}Turnout in 2018 Midterm Election}\label{fig:to-18-recent-chart}
+\end{figure}
+
 Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
-
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
-load("./temp/short_term_to.rdata")
-
-stargazer(date_to_model_notime, date_to_model_time,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
-                               "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit-short} Individual-Level Logit Models",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
-                         c("Felony Class FE", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(date_to_model_notime, type = "HC0"))),
-            sqrt(diag(vcovHC(date_to_model_time, type = "HC0"))))
-          )
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table3.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018.
+
 \input{"../temp/table3.tex"}
+
 The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.]
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.}
 
@@
 Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18^th^ that parolees consistently had their voting rights restored prior to discharge.
-```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/restoration_plot.rds")
-print(j)
-```
+It is not until May 18\textsuperscript{th} that parolees consistently had their voting rights restored prior to discharge.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/restoration-chart-chunk-1}
+
+}
+
+\caption{\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge}\label{fig:restoration-chart-chunk}
+\end{figure}
 
 Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are "treated" by the policy change.
+Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
+Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are ``treated'' by the policy change.
 Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
-
-```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit} Individual-Level Logit Models",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18\textsuperscript{th}, 2018, turned out at higher rates than those discharged earlier.
+The models include all individuals discharged from parole between January 1\textsuperscript{st}, 2017, through October 10\textsuperscript{th}, 2018 (the registration deadline in New York State).
+
 \input{"../temp/table4.tex"}
@@
 Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
+Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
 Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
@@
 Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model2[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
-
-## Instrumental Variables Approach {-}
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 31.2 and 31.8 percent.
+
+\hypertarget{instrumental-variables-approach}{%
+\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
+\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
 
@@
 
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied." Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
+Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
 
 An individual's propensity to vote can be expressed using the following equation:
-$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} $$
-Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
-*Z* is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
+\[ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} \]
+Where \emph{Y\textsubscript{i}} is 1 if individual \emph{i} cast a ballot, \emph{X\textsubscript{1i}} is the probability that individual \emph{i} was eligible to have his rights restored prior to discharge from parole, and \emph{X\textsubscript{2i}} is 1 if individual \emph{i} actually had his rights restored prior to discharge from parole.
+\emph{Z} is a vector of other factors (such as age and race) known to influence voter turnout.
+Given that we cannot observe \emph{X\textsubscript{1i}}, we could simply choose to ignore it.
+Doing so, however, will result in consistent regression coefficients only if \emph{X\textsubscript{1i}} and \emph{X\textsubscript{2i}} (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if \emph{b\textsubscript{1}} is uncorrelated with an individual's propensity to vote (\emph{Y\textsubscript{i}}).
 
@@
 
-```{r restoree-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_restorees.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
-\\\\Shows demographics for individuals discharged from parole \\\\between May 18th, 2018, and October 12th, 2018.}}}"
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever.tex", quote = F, col.names = F,
-            row.names = F)
-```
 \input{"../temp/table_whatever.tex"}
 
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
+Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
 Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
@@
 This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that *b~1~* is very probably correlated with *Y*.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach [@Angrist1996].
-Such an approach has been widely used in the context of voter turnout [e.g.
-@Ansolabehere1999; @Gerber2000; @Milligan2004;  @Lassen2004; @Sondheimer2010].
+Table \ref{tab:demos-restorees} indicates that \emph{b\textsubscript{1}} is very probably correlated with \emph{Y}.
+An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration \emph{caused} turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
+
+The standard way of dealing with such a problem is an instrumental variables approach (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
+Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
 A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
@@
 
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the "local average treatment effect." The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
-Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research [e.g.
-@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2014].
-The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
+Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
+Some research (e.g.~Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
+The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
+The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2007}{2007}).
+This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
 
@@
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
-
-### Racial Variation {-}
+Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between 32.7 and 39.5 percent.
+
+\hypertarget{racial-variation}{%
+\subsubsection*{Racial Variation}\label{racial-variation}}
+\addcontentsline{toc}{subsubsection}{Racial Variation}
 
@@
 
-```{r nameless-chunk, echo = F, results = "asis", message = F}
-load("./temp/iv_individual_turnout_18.rdata")
-```
-
 Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration *nearly doubled* these individuals' likelihood of casting a ballot.
+Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
 We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
@@
 Some of this may be explained by different propensities to vote.
-@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
+White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
 The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
@@
 
-```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/black_to.rdata")
-
-stargazer(black_to1, black_to2, black_to1b, black_to2b, black_to3, black_to4,
-          column.separate = c(2, 2, 2),
-          column.labels = c("Discharged Prior to EO", "ITT Group", "Treatment Group"),
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Black)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:black-to} Turnout Rate Among Former Parolees",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          omit = "felony",
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(black_to1, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to2, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to1b, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to2b, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to3, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to4, type = "HC0")))
-          ))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{7}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.
-\\\\Latino individuals excluded from Models 2, 4, and 6.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}  \\cmidrule(l{3pt}r{3pt}){6-7}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table5.tex", quote = F, col.names = F,
-            row.names = F)
-```
 \input{"../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
+
+Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
 These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
@@
 
-### Variable Treatment Intensity {-}
+\hypertarget{variable-treatment-intensity}{%
+\subsubsection*{Variable Treatment Intensity}\label{variable-treatment-intensity}}
+\addcontentsline{toc}{subsubsection}{Variable Treatment Intensity}
 
@@
 This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.^[Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.]
+It is coded as 0 for individuals discharged prior to the implementation of the executive order.\footnote{Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.}
 
@@
 Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
-It is worth noting that the *R^2^* in Model 1 is higher than the *R^2^* in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
+It is worth noting that the \emph{R\textsuperscript{2}} in Model 1 is higher than the \emph{R\textsuperscript{2}} in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
 
@@
 
-## Discussion {-}
-
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both *de jure* and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
+\hypertarget{discussion-1}{%
+\subsection*{Discussion}\label{discussion-1}}
+\addcontentsline{toc}{subsection}{Discussion}
+
+Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
 Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
 That changed in 2018.
-On October 12^th^ --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
+On October 12\textsuperscript{th} --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
 Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just `r comma(readRDS("./temp/new_votes_18.rds"))`, or `r round((readRDS("./temp/new_votes_18.rds")/readRDS("./temp/new_eligible_18.rds"))*100, 2)` percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
+Though turnout among this group was low (just 791, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
 
@@
 In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` percent.
+In this project I estimate that rights restoration increased individuals' propensity to vote by at least 32.7 percent.
 This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
@@
 The mechanism through which the rules change increased turnout among these individuals is not clear.
-@Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
+Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
 Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
 On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by @Weaver2010.
+Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
 
 It could also be a story of better information.
-As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
-Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
+(\protect\hyperlink{ref-Gerber2014}{2014}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
+Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
 When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
@@
 Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners [@Hemmens2000].
+Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
 Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-@WHITE2014, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2016].
+White, Nathan, and Faller (\protect\hyperlink{ref-WHITE2014}{2014}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2016}{2016}).
 
@@
 
-# Conclusion
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
@@
 The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these policies.
-By both redefining "lost voters" as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
+By both redefining ``lost voters'' as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
 What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, and this effect is concentrated in Black neighborhoods.
@@
 America is becoming ever-more-diverse, and will likely be less than 50 percent non-Hispanic White by the middle of the century.
-As Fraga [-@Fraga2018] has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
+As Fraga (\protect\hyperlink{ref-Fraga2018}{2018}) has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
 How districts are drawn, population patterns, and rules about who can fundraise how will continue to shape the politicians we elect and the policies they enact.
-Breaking down *de jure* (like formal disenfranchisement) and *de facto* (like neighborhood spillover effects) barriers to the ballot box is of great importance.
+Breaking down \emph{de jure} (like formal disenfranchisement) and \emph{de facto} (like neighborhood spillover effects) barriers to the ballot box is of great importance.
 The experience in New York State over the past few years offers a better understanding of just how big this problem is, but it also offers evidence of a way forward.
@@
 \newpage
-# Appendix A {-}
-
-In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-
-```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_16.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
-                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))
-          ))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table55.tex", quote = F, col.names = F,
-            row.names = F)
-```
+
+\hypertarget{appendix-a}{%
+\section*{Appendix A}\label{appendix-a}}
+\addcontentsline{toc}{section}{Appendix A}
+
+In the \protect\hyperlink{instrumental-variables-approach}{Instrumental Variables Approach} section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1\textsuperscript{st}, 2015, and October 14\textsuperscript{th}, 2016.
+
 \input{"../temp/table55.tex"}
@@
 \newpage
-# References {-}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
 \setlength{\parindent}{-0.5in}
 \setlength{\leftskip}{0.5in}
+
 \noindent
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Abadie2019}{}%
+Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \url{https://scholar.harvard.edu/spiess/publications/robust-post-matching-inference.}
+
+\leavevmode\hypertarget{ref-Alexander2012}{}%
+Alexander, Michelle. 2012. \emph{The New Jim Crow}. New York: The New Press.
+
+\leavevmode\hypertarget{ref-Angrist2001}{}%
+Angrist, Joshua D.
+2001.
+``Estimation of Limited Dependent Variable Models with Dummy Endogenous Regressors.'' \emph{Journal of Business \& Economic Statistics} 19 (1): 2--28.
+\url{https://doi.org/10.1198/07350010152472571}.
+
+\leavevmode\hypertarget{ref-Angrist1996}{}%
+Angrist, Joshua D., Guido W.
+Imbens, and Donald B.
+Rubin.
+1996.
+``Identification of Causal Effects Using Instrumental Variables.'' \emph{Journal of the American Statistical Association} 91 (434): 444--55.
+\url{https://doi.org/10.1080/01621459.1996.10476902}.
+
+\leavevmode\hypertarget{ref-Angrist2008}{}%
+Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
+
+\leavevmode\hypertarget{ref-Ansolabehere1999}{}%
+Ansolabehere, Stephen D., Shanto Iyengar, and Adam Simon.
+1999.
+``Replicating Experiments Using Aggregate and Survey Data: The Case of Negative Advertising and Turnout.'' \emph{American Political Science Review} 93 (4): 901--9.
+\url{https://doi.org/10.2307/2586120}.
+
+\leavevmode\hypertarget{ref-Anzia2019}{}%
+Anzia, Sarah F.
+2019.
+``When Does a Group of Citizens Influence Policy?
+Evidence from Senior Citizen Participation in City Politics.'' \emph{The Journal of Politics} 81 (1): 1--14.
+\url{https://doi.org/10.1086/699916}.
+
+\leavevmode\hypertarget{ref-Behrens2003}{}%
+Behrens, Angela, Christopher Uggen, and Jeff Manza.
+2003.
+``Ballot Manipulation and the ``Menace of Negro Domination'': Racial Threat and Felon Disenfranchisement in the United States, 18502002.'' \emph{American Journal of Sociology} 109 (3): 559--605.
+\url{https://doi.org/10.1086/378647}.
+
+\leavevmode\hypertarget{ref-Bowers2009}{}%
+Bowers, Melanie, and Robert R.
+Preuhs.
+2009.
+``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons.'' \emph{Social Science Quarterly} 90 (3): 722--43.
+\url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
+
+\leavevmode\hypertarget{ref-bcj_laws}{}%
+Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
+
+\leavevmode\hypertarget{ref-Browning1992}{}%
+Browning, Sandra Lee, and Liqun Cao. 1992. ``The Impact of Race on Criminal Justice Ideology.'' \emph{Justice Quarterly} 9 (4): 685--701. \url{https://doi.org/10.1080/07418829200091611}.
+
+\leavevmode\hypertarget{ref-Burch2010}{}%
+Burch, Traci.
+2010.
+``Did Disfranchisement Laws Help Elect President Bush?
+New Evidence on the Turnout Rates and Candidate Preferences of Florida's Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
+\url{https://doi.org/10.1007/s11109-010-9150-9}.
+
+\leavevmode\hypertarget{ref-Burch2011}{}%
+---------.
+2011.
+``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
+\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
+
+\leavevmode\hypertarget{ref-Burch2013}{}%
+---------.
+2013.
+``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina.'' Edited by Christopher Wildeman, Jacob S.
+Hacker, and Vesla M.
+Weaver.
+\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 184--201.
+\url{https://doi.org/10.1177/0002716213503093}.
+
+\leavevmode\hypertarget{ref-Butler2011}{}%
+Butler, Daniel M., and David E.
+Broockman.
+2011.
+``Do Politicians Racially Discriminate Against Constituents?
+A Field Experiment on State Legislators.'' \emph{American Journal of Political Science} 55 (3): 463--77.
+\url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
+
+\leavevmode\hypertarget{ref-Cho2006}{}%
+Cho, Wendy K.
+Tam, James G.
+Gimpel, and Joshua J.
+Dyck.
+2006.
+``Residential Concentration, Political Socialization, and Voter Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
+\url{https://doi.org/10.1111/j.1468-2508.2006.00377.x}.
+
+\leavevmode\hypertarget{ref-Clear2008}{}%
+Clear, Todd R. 2008. ``The Effects of High Imprisonment Rates on Communities.'' \emph{Crime and Justice} 37 (1): 97--132. \url{https://doi.org/10.1086/522360}.
+
+\leavevmode\hypertarget{ref-Dawkins2006}{}%
+Dawkins, Casey J. 2006. ``Are Social Networks the Ties That Bind Families to Neighborhoods?'' \emph{Housing Studies} 21 (6): 867--81. \url{https://doi.org/10.1080/02673030600917776}.
+
+\leavevmode\hypertarget{ref-Dong2014}{}%
+Dong, Yingying, and Arthur Lewbel.
+2014.
+``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
+\url{https://doi.org/10.1080/07474938.2014.944470}.
+
+\leavevmode\hypertarget{ref-Drucker2005}{}%
+Drucker, Ernest, and Ricardo Barreras.
+2005.
+``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
+Sentencing Project.
+\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
+
+\leavevmode\hypertarget{ref-Einstein2016}{}%
+Einstein, Katherine Levine, and David M.
+Glick.
+2016.
+``Does Race Affect Access to Government Services?
+An Experiment Exploring Street-Level Bureaucrats and Access to Public Housing.'' \emph{American Journal of Political Science} 61 (1): 100--116.
+\url{https://doi.org/10.1111/ajps.12252}.
+
+\leavevmode\hypertarget{ref-Foladare1968}{}%
+Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
+
+\leavevmode\hypertarget{ref-Fraga2018}{}%
+Fraga, Bernard L. 2018. \emph{The Turnout Gap}. Cambridge, United Kingdom: Cambridge University Press. \url{https://doi.org/10.1017/9781108566483}.
+
+\leavevmode\hypertarget{ref-Geller2014}{}%
+Geller, Amanda, Jeffrey Fagan, Tom Tyler, and Bruce G.
+Link.
+2014.
+``Aggressive Policing and the Mental Health of Young Urban Men.'' \emph{American Journal of Public Health} 104 (12): 2321--7.
+\url{https://doi.org/10.2105/ajph.2014.302046}.
+
+\leavevmode\hypertarget{ref-Gelman2007}{}%
+Gelman, Andrew, Jeffrey Fagan, and Alex Kiss.
+2007.
+``An Analysis of the New York City Police Departments ``Stop-and-Frisk'' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
+\url{https://doi.org/10.1198/016214506000001040}.
+
+\leavevmode\hypertarget{ref-Gerber2000}{}%
+Gerber, Alan S., and Donald P.
+Green.
+2000.
+``The Effects of Canvassing, Telephone Calls, and Direct Mail on Voter Turnout: A Field Experiment.'' \emph{American Political Science Review} 94 (3): 653--63.
+\url{https://doi.org/10.2307/2585837}.
+
+\leavevmode\hypertarget{ref-Gerber2003}{}%
+Gerber, Alan S., Donald P.
+Green, and Ron Shachar.
+2003.
+``Voting May Be Habit-Forming: Evidence from a Randomized Field Experiment.'' \emph{American Journal of Political Science} 47 (3): 540--50.
+\url{https://doi.org/10.1111/1540-5907.00038}.
+
+\leavevmode\hypertarget{ref-Gerber2014}{}%
+Gerber, Alan S., Gregory A.
+Huber, Marc Meredith, Daniel R.
+Biggers, and David J.
+Hendry.
+2014.
+``Can Incarcerated Felons Be (Re)integrated into the Political System?
+Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
+\url{https://doi.org/10.1111/ajps.12166}.
+
+\leavevmode\hypertarget{ref-Gerber2017}{}%
+---------.
+2017.
+``Does Incarceration Reduce Voting?
+Evidence About the Political Consequences of Spending Time in Prison.'' \emph{The Journal of Politics} 79 (4): 1130--46.
+\url{https://doi.org/10.1086/692670}.
+
+\leavevmode\hypertarget{ref-Gimpel2004}{}%
+Gimpel, James G., Joshua J.
+Dyck, and Daron R.
+Shaw.
+2004.
+``Registrants, Voters, and Turnout Variability Across Neighborhoods.'' \emph{Political Behavior} 26 (4): 343--75.
+\url{https://doi.org/10.1007/s11109-004-0900-4}.
+
+\leavevmode\hypertarget{ref-Green2000}{}%
+Green, Donald P., and Ron Shachar.
+2000.
+``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
+\url{https://doi.org/10.1017/s0007123400000247}.
+
+\leavevmode\hypertarget{ref-Griffin2005}{}%
+Griffin, John D., and Brian Newman. 2005. ``Are Voters Better Represented?'' \emph{The Journal of Politics} 67 (4): 1206--27. \url{https://doi.org/10.1111/j.1468-2508.2005.00357.x}.
+
+\leavevmode\hypertarget{ref-Guest1999}{}%
+Guest, Avery M., and Susan K. Wierzbicki. 1999. ``Social Ties at the Neighborhood Level.'' \emph{Urban Affairs Review} 35 (1): 92--111. \url{https://doi.org/10.1177/10780879922184301}.
+
+\leavevmode\hypertarget{ref-Hajnal2009}{}%
+Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
+
+\leavevmode\hypertarget{ref-Hajnal2005}{}%
+Hajnal, Zoltan, and Jessica Trounstine.
+2005.
+``Where Turnout Matters: The Consequences of Uneven Turnout in City Politics.'' \emph{The Journal of Politics} 67 (2): 515--35.
+\url{https://doi.org/10.1111/j.1468-2508.2005.00327.x}.
+
+\leavevmode\hypertarget{ref-Haselswerdt2009}{}%
+Haselswerdt, Michael V.
+2009.
+``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data.'' \emph{Social Science Quarterly} 90 (2): 262--73.
+\url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
+
+\leavevmode\hypertarget{ref-Hemmens2000}{}%
+Hemmens, Craig, and James W.
+Marquart.
+2000.
+``Friend or Foe?
+Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
+\url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
+
+\leavevmode\hypertarget{ref-Henderson1997}{}%
+Henderson, Martha L., Francis T.
+Cullen, Liqun Cao, Sandra Lee Browning, and Renee Kopache.
+1997.
+``The Impact of Race on Perceptions of Criminal Injustice.'' \emph{Journal of Criminal Justice} 25 (6): 447--62.
+\url{https://doi.org/10.1016/s0047-2352(97)00032-9}.
+
+\leavevmode\hypertarget{ref-Hjalmarsson2010}{}%
+Hjalmarsson, Randi, and Mark Lopez.
+2010.
+``The Voting Behavior of Young Disenfranchised Felons: Would They Vote If They Could?'' \emph{American Law and Economics Review} 12 (2): 356--93.
+\url{https://doi.org/10.1093/aler/ahq004}.
+
+\leavevmode\hypertarget{ref-Huckfeldt1979}{}%
+Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
+
+\leavevmode\hypertarget{ref-Hurwitz2005}{}%
+Hurwitz, Jon, and Mark Peffley.
+2005.
+``Explaining the Great Racial Divide: Perceptions of Fairness in the U.s.
+Criminal Justice System.'' \emph{The Journal of Politics} 67 (3): 762--83.
+\url{https://doi.org/10.1111/j.1468-2508.2005.00338.x}.
+
+\leavevmode\hypertarget{ref-Kenny1992}{}%
+Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259. \url{https://doi.org/10.2307/2111432}.
+
+\leavevmode\hypertarget{ref-Keyssar2009}{}%
+Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. New York: Basic Books.
+
+\leavevmode\hypertarget{ref-King2016}{}%
+King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821. \url{https://doi.org/10.1177/0021934716659195}.
+
+\leavevmode\hypertarget{ref-Kinsella2015}{}%
+Kinsella, Chad, Colleen McTague, and Kevin N.
+Raleigh.
+2015.
+``Unmasking Geographic Polarization and Clustering: A Micro-Scalar Analysis of Partisan Voting Behavior.'' \emph{Applied Geography} 62 (August): 404--19.
+\url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
+
+\leavevmode\hypertarget{ref-Lassen2004}{}%
+Lassen, David Dreyer. 2004. ``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' \emph{SSRN Electronic Journal}. \url{https://doi.org/10.2139/ssrn.475821}.
+
+\leavevmode\hypertarget{ref-Lerman2013}{}%
+Lerman, Amy E., and Vesla Weaver.
+2013.
+``Staying Out of Sight?
+Concentrated Policing and Local Political Action.'' Edited by Christopher Wildeman, Jacob S.
+Hacker, and Vesla M.
+Weaver.
+\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 202--19.
+\url{https://doi.org/10.1177/0002716213503085}.
+
+\leavevmode\hypertarget{ref-locked_out}{}%
+Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
+
+\leavevmode\hypertarget{ref-Martin2003}{}%
+Martin, Paul S.
+2003.
+``Votings Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
+\url{https://doi.org/10.1111/1540-5907.00008}.
+
+\leavevmode\hypertarget{ref-Martin2013}{}%
+Martin, Paul S., and Michele P.
+Claibourn.
+2013.
+``Citizen Participation and Congressional Responsiveness: New Evidence That Participation Matters.'' \emph{Legislative Studies Quarterly} 38 (1): 59--81.
+\url{https://doi.org/10.1111/lsq.12003}.
+
+\leavevmode\hypertarget{ref-Meredith2013}{}%
+Meredith, Marc, and Michael Morse.
+2013.
+``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?'' Edited by Christopher Wildeman, Jacob S.
+Hacker, and Vesla M.
+Weaver.
+\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 220--49.
+\url{https://doi.org/10.1177/0002716213502931}.
+
+\leavevmode\hypertarget{ref-Meredith2015}{}%
+---------.
+2015.
+``The Politics of the Restoration of Ex-Felon Voting Rights: The Case of Iowa.'' \emph{Quarterly Journal of Political Science} 10 (1): 41--100.
+\url{https://doi.org/10.1561/100.00013026}.
+
+\leavevmode\hypertarget{ref-Miles2004}{}%
+Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.'' \emph{The Journal of Legal Studies} 33 (1): 85--129. \url{https://doi.org/10.1086/381290}.
+
+\leavevmode\hypertarget{ref-Milligan2004}{}%
+Milligan, Kevin, Enrico Moretti, and Philip Oreopoulos.
+2004.
+``Does Education Improve Citizenship?
+Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9-10): 1667--95.
+\url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
+
+\leavevmode\hypertarget{ref-Morris2019}{}%
+Morris, Kevin. 2019. ``Thwarting Amendment 4.'' May 9, 2019. \url{https://www.brennancenter.org/analysis/thwarting-amendment-4}.
+
+\leavevmode\hypertarget{ref-Mutz2002}{}%
+Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838. \url{https://doi.org/10.2307/3088437}.
+
+\leavevmode\hypertarget{ref-Ochs2006}{}%
+Ochs, Holona Leanne.
+2006.
+``'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
+\url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
+
+\leavevmode\hypertarget{ref-Pierson1993}{}%
+Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
+
+\leavevmode\hypertarget{ref-Sekhon2011}{}%
+Sekhon, Jasjeet S.
+2011.
+``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (7).
+\url{https://doi.org/10.18637/jss.v042.i07}.
+
+\leavevmode\hypertarget{ref-Sewell2016}{}%
+Sewell, Abigail A., and Kevin A.
+Jefferson.
+2016.
+``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (S1): 42--67.
+\url{https://doi.org/10.1007/s11524-015-0016-7}.
+
+\leavevmode\hypertarget{ref-florida_1868}{}%
+Shofner, Jerrell H. 1963. ``The Constitution of 1868.'' \emph{The Florida Historical Quarterly} 41 (4): 356--74. \url{http://www.jstor.org/stable/30139965}.
+
+\leavevmode\hypertarget{ref-Sondheimer2010}{}%
+Sondheimer, Rachel Milstein, and Donald P.
+Green.
+2010.
+``Using Experiments to Estimate the Effects of Education on Voter Turnout.'' \emph{American Journal of Political Science} 54 (1): 174--89.
+\url{https://doi.org/10.1111/j.1540-5907.2009.00425.x}.
+
+\leavevmode\hypertarget{ref-Terza2007}{}%
+Terza, Joseph V., W.
+David Bradford, and Clara E.
+Dismuke.
+2007.
+``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
+\url{https://doi.org/10.1111/j.1475-6773.2007.00807.x}.
+
+\leavevmode\hypertarget{ref-sentencing_2016}{}%
+Uggen, Christopher, Ryan Larson, and Sarah Shannon.
+2016.
+``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
+Sentencing Project.
+\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
+
+\leavevmode\hypertarget{ref-Uggen2002}{}%
+Uggen, Christopher, and Jeff Manza.
+2002.
+``Democratic Contraction?
+Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
+\url{https://doi.org/10.2307/3088970}.
+
+\leavevmode\hypertarget{ref-Uggen2004}{}%
+---------.
+2004.
+``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
+New York: Russell Sage Foundation.
+
+\leavevmode\hypertarget{ref-Walker2014}{}%
+Walker, Hannah L.
+2014.
+``Extending the Effects of the Carceral State.
+Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly} 67 (4): 809--22.
+\url{https://doi.org/10.1177/1065912914542522}.
+
+\leavevmode\hypertarget{ref-Walker2017}{}%
+Walker, Hannah L., and Marcela Garcı́a-Castañon.
+2017.
+``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (04): 541--68.
+\url{https://doi.org/10.1017/s1743923x17000198}.
+
+\leavevmode\hypertarget{ref-Weaver2014}{}%
+Weaver, Vesla, and Amy E. Lerman. 2014. \emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}. Chicago: University of Chicago Press.
+
+\leavevmode\hypertarget{ref-Weaver2010}{}%
+Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
+
+\leavevmode\hypertarget{ref-White2019}{}%
+White, Ariel.
+2019.
+``Misdemeanor Disenfranchisement?
+The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
+\url{https://doi.org/10.1017/s000305541800093x}.
+
+\leavevmode\hypertarget{ref-WHITE2014}{}%
+White, Ariel R., Noah L.
+Nathan, and Julie K.
+Faller.
+2014.
+``What Do I Need to Vote?
+Bureaucratic Discretion and Discrimination by Local Election Officials.'' \emph{American Political Science Review} 109 (1): 129--42.
+\url{https://doi.org/10.1017/s0003055414000562}.
+
+\leavevmode\hypertarget{ref-Wilson1990}{}%
+Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
+
+\leavevmode\hypertarget{ref-Wooldridge2010}{}%
+Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
+
+\leavevmode\hypertarget{ref-Wu2009}{}%
+Wu, Yuning, Ivan Y.
+Sun, and Ruth A.
+Triplett.
+2009.
+``Race, Class or Neighborhood Context: Which Matters More in Measuring Satisfaction with Police?'' \emph{Justice Quarterly} 26 (1): 125--56.
+\url{https://doi.org/10.1080/07418820802119950}.
+
+\end{document}

### 2019-10-18 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 
+Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
+It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
+Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
+
 Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
@@
 Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-In Model 2, D(Treat) $\times$ D(2017) is negative but not statistically significant.
-However, because treated tracts had significantly higher turnout than control tracts in 2016, this nonsignificance implies treatment --- the loss of a voter --- did decrease turnout to a significant degree relative to the control tracts.
+The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
 
@@
 Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Unfortunately, because many neighborhoods consistently have lost voters, testing the causal effect of the first lost voter for all neighborhoods is not possible.
-Nevertheless, this analysis lends credence to the causal nature of lost voters on depressed turnout.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
+
+Of particular note is the lower share of the population in these neighborhoods that is Black.
+The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
+The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
+Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
 
@@
 Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
 
@@
 Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are "treated" by the policy change.
-Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
+Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are assigned "treatment" by the policy change.
+Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
+
+To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
+With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
+To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is biased against finding a significant increase due to the executive order.
+
+```{r rd-demo-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_rd_time.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of \"Control\" and \"Intend-to-Treat\" Groups", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
+\\\\Shows demographics for individuals discharged from parole \\\\between January 1st, 2017, and October 12th, 2018.}}}"
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Control\" and \"Intend-to-Treat\" Groups}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table_whatever2.tex"}
 

### 2019-10-18 markdown/felony_disenfranchisement_nys.tex
write

@@
   \postdate{\par}
-    \date{September 26, 2019}
+    \date{October 17, 2019}
 
@@
 \midrule
-\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 96.18 & 85.46 & 85.00 & 76.98\\
-\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.24 & 97.15 & 88.52 & 85.34 & 77.02\\
-\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.20 & 87.45 & 79.41 & 64.95\\
-Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,711.89 & 86.31 & 76.32 & 76.77 & 65.49\\
-\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 95.15 & 88.35 & 84.14 & 75.44\\
-Median Age & 36.36 & 38.23 & 36.36 & 36.38 & 98.90 & 69.64 & 65.76 & 40.69\\
-Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 8.97 & 63.21 & 60.34 & 50.97\\
-\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.01 & 87.99 & 85.70 & 73.48\\
-\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 83.92 & 77.72 & 65.69 & 47.73\\
-\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.99 & 77.50 & 67.39 & 41.31\\
+\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 94.34 & 86.10 & 84.77 & 75.82\\
+\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.25 & 99.28 & 89.75 & 86.62 & 77.90\\
+\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.02 & 87.45 & 77.73 & 61.67\\
+Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,973.89 & 84.23 & 78.35 & 76.98 & 66.57\\
+\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 96.86 & 86.51 & 83.73 & 76.55\\
+Median Age & 36.36 & 38.23 & 36.36 & 36.39 & 98.60 & 73.15 & 69.00 & 42.11\\
+Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 10.68 & 64.27 & 63.46 & 56.35\\
+\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.23 & 90.06 & 87.35 & 76.31\\
+\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 78.09 & 73.55 & 62.03 & 48.36\\
+\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.91 & 75.59 & 65.98 & 38.91\\
 \bottomrule
@@
 Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-In Model 2, D(Treat) \(\times\) D(2017) is negative but not statistically significant.
-However, because treated tracts had significantly higher turnout than control tracts in 2016, this nonsignificance implies treatment --- the loss of a voter --- did decrease turnout to a significant degree relative to the control tracts.
+The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
 
@@
 Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Unfortunately, because many neighborhoods consistently have lost voters, testing the causal effect of the first lost voter for all neighborhoods is not possible.
-Nevertheless, this analysis lends credence to the causal nature of lost voters on depressed turnout.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
+
+Of particular note is the lower share of the population in these neighborhoods that is Black.
+The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
+The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
+Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
 
@@
 Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018.
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
 
@@
 Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are ``treated'' by the policy change.
-Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
+Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
+
+To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
+With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
+To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is biased against finding a significant increase due to the executive order.
+
+\input{"../temp/table_whatever2.tex"}
 
@@
 Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just 791, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
+Though turnout among this group was low (just 792, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
 

### 2019-11-06 markdown/felony_disenfranchisement_nys.Rmd
write

@@
 In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
+As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific arrest rates.
 
@@
 Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
+If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
 If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
@@
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
+Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods through both direct (the formal disenfranchisement of would-be voters) and indirect (through proximal contact with disenfranchised individuals) effects.
 The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely *understated* the degree to which they are impacted by spatially concentrated patterns of criminal justice.
@@
 
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
+In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
+This variation in laws has been defended with language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as *Richardson v.
@@
 Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended.
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
 
@@
 
+Similarly, @White2019a finds that incarceration has only moderate indirect effects on voting.
+She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (607).
+This study examines only the indirect effects of proximal contact with the criminal justice system.
+It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect depressive effect of incarceration more generally.
+Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
+
 ### Theoretical Implications for Reduced Turnout {-}
@@
 Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
 By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
@@
 By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
+Individuals who were (A) citizens, (B) not in violation of any terms of their parole, and (C) had not been rearrested were eligible for rights restoration.
 
@@
 ## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+Voters' home addresses were converted to latitudes and longitudes using a geocoding services from address-verification company SmartyStreets.
 I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
@@
 ## Identification of Lost Voters {-}
+
+Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
+By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
+Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
+
 In this analysis, I offer a different definition of "lost voter" than much of the literature.
@@
 If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
+Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have "aged in" to voting but for their incarceration, they will not will have no history of voting.
 This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
@@
 
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
+Examining the impact of all disenfranchised voters, however, limits our ability to understand the specific direct and indirect causal effects of felony disenfranchisement.
 As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration [e.g.
 @Gerber2017].
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement *per se* but rather to their incarceration.
-At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
+Research showing lower turnout where more individuals are formally disenfranchised might only be identifying the indirect effects of incarceration on turnout.
+Estimation of the effects of felony disenfranchisement above-and-beyond the effects of incarceration must account for the low propensity to vote among the formally disenfranchised.
 
@@
 Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote [@Gerber2003].
-As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
+As such, these are the individuals most likely to have been directly impacted not only by incarceration but specifically by felony disenfranchisement.
 Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
@@
 
+Of course, the incarceration of neighbors and family members who have *not* voted in the past might also have indirect effects on neighborhood turnout rates.
+Even if an eligible New Yorker's incarcerated brother would not have voted, for instance, the eligible voter might not cast a ballot due to a soured impression of the political process.
+Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
+I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
+These higher indirect (depressive) effects likely arise through three mechanisms.
+
+* Firstly, lost voters are likely members of families and communities where voting is valued.
+Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
+Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
+Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
+
+* Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
+The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
+
+* Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
+Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
+
 To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
 Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
+In many cases, these "lost voters" will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
+However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
 
 Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.
+There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election.
 
@@
 ```
+Although `r comma(nrow(readRDS("./temp/in_2017.rds")))` individuals were formally disenfranchised as of the 2017 election, there were just `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` disenfranchised voters statewide who had participated in the past 10 years.^[Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.] Just `r round(100 * nrow(readRDS("./temp/ids_of_lost_voters.rds")) / nrow(readRDS("./temp/in_2017.rds")), digits = 1)` percent of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
+
 The spatial concentration of lost voters is readily apparent.
@@
 
-Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
+I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
 Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
@@
 Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau's 5-year American Community Survey estimates.
 Party affiliation rates come from the geocoded voter file.
@@
 At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less White,^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] have much lower median incomes, and a larger share of voters are registered as Democrats.
+
+In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
+On average, the non-Hispanic Black share of a block group with a lost voter is more than *double* the share in block groups without lost voters.
+These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
+Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] The median income in neighborhoods without lost voters is more than 40 percent higher.
+Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
 
@@
 
-The block groups where these depressive effects are not randomly distributed throughout the city.
+The block groups where these depressive effects are concentrated are not randomly distributed throughout the city.
 They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
@@
 
-In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups and tracts.
+In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups and tracts.
 In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
@@
 Understanding the magnitude of this finding is important.
-In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population, with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood `r round(0.9 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.16 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots.
+
+The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
 
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018.
-This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
-senatorial contests (as well as other down-ballot races).
-
-I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
 This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
@@
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
+For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018 before the executive order was enacted.
 
@@
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.]
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
 
@@
 The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is biased against finding a significant increase due to the executive order.
+To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
 
@@
 
-lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Control\" and \"Intend-to-Treat\" Groups}"
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of \"Control\" and \"Intend-to-Treat\" Groups}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"

### 2019-11-06 markdown/felony_disenfranchisement_nys.tex
write

@@
   \postdate{\par}
-    \date{October 17, 2019}
+    \date{November 05, 2019}
 
@@
 In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
+As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific arrest rates.
 
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}; Gerber et al.
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (A.
+White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}; Gerber et al.
 \protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
@@
 This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
+Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods through both direct (the formal disenfranchisement of would-be voters) and indirect (through proximal contact with disenfranchised individuals) effects.
 The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely \emph{understated} the degree to which they are impacted by spatially concentrated patterns of criminal justice.
@@
 
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
+In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
+This variation in laws has been defended with language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as \emph{Richardson v.
@@
 Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended.
+This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
 
@@
 
+Similarly, A.
+White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
+She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (607).
+This study examines only the indirect effects of proximal contact with the criminal justice system.
+It is possible that the indirect effects of the incarceration or jailing of an individual who would have voted but not for their contact with the criminal justice system would have larger indirect effects.
+Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
+
 \hypertarget{theoretical-implications-for-reduced-turnout}{%
@@
 
-White (\protect\hyperlink{ref-White2019}{2019}), however, comes to a different conclusion.
+A.
+White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), however, comes to a different conclusion.
 Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
 This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
 By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
@@
 
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+Voters' home addresses were converted to latitudes and longitudes using a geocoding services from address-verification company SmartyStreets.
 I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
@@
 
+Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
+By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
+Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
+
 In this analysis, I offer a different definition of ``lost voter'' than much of the literature.
@@
 If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
+Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have ``aged in'' to voting but for their incarceration, they will not will have no history of voting.
 This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
@@
 
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
+Examining the impact of all disenfranchised voters, however, limits our ability to understand the specific direct and indirect causal effects of felony disenfranchisement.
 As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration (e.g.~Gerber et al.
 \protect\hyperlink{ref-Gerber2017}{2017}).
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement \emph{per se} but rather to their incarceration.
-At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
+Research showing lower turnout where more individuals are formally disenfranchised might only be identifying the indirect effects of incarceration on turnout.
+Estimation of the effects of felony disenfranchisement above-and-beyond the effects of incarceration must account for the low propensity to vote among the formally disenfranchised.
 
@@
 Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote (Gerber, Green, and Shachar \protect\hyperlink{ref-Gerber2003}{2003}).
-As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
+As such, these are the individuals most likely to have been directly impacted not only by incarceration but specifically by felony disenfranchisement.
 Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
@@
 
+Of course, the incarceration of neighbors and family members who have \emph{not} voted in the past might also have indirect effects on neighborhood turnout rates.
+Even if an eligible New Yorker's incarcerated brother would not have voted, for instance, the eligible voter might not cast a ballot due to a soured impression of the political process.
+Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
+I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
+These higher indirect (depressive) effects likely arise through three mechanisms.
+* Firstly, lost voters are likely members of families and communities where voting is valued.
+Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
+Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
+Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
+* Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
+The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
+* Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
+Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
+
 To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
 Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
+In many cases, these ``lost voters'' will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
+However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
 
 Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were 2,518 such lost voters within New York City as of the 2017 general election, and 6,166 statewide.
+There were 2,518 such lost voters within New York City as of the 2017 general election.
 
@@
 \end{figure}
+
+Although 73,223 individuals were formally disenfranchised as of the 2017 election, there were just 6,166 disenfranchised voters statewide who had participated in the past 10 years.\footnote{Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.} Just 8.4 percent of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
 
@@
 
-Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
+I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
 Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
@@
 Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau's 5-year American Community Survey estimates.
 Party affiliation rates come from the geocoded voter file.
@@
 At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less White,\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} have much lower median incomes, and a larger share of voters are registered as Democrats.
+
+In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
+On average, the non-Hispanic Black share of a block group with a lost voter is more than \emph{double} the share in block groups without lost voters.
+These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
+Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} The median income in neighborhoods without lost voters is more than 40 percent higher.
+Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
 
@@
 Understanding the magnitude of this finding is important.
-In New York City, there are 165 block groups where the Black community makes up more than 90 percent of the population, with an average of 821 voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of 821, having any lost voter reduced turnout by 20 ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just 2.2 ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood 13.9 ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just 1.5 ballots --- hardly more votes than the lost voter herself.
+In New York City, there are 1386 block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of 912 voting age citizens.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of 912, having any lost voter reduced turnout by 9.4 ballots.
+Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just 3.8 ballots.
+
+The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood 6.5 ballots.
+In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just 1.6 ballots --- hardly more votes than the lost voter herself.
 
@@
 
-This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This finding mirrors A.
+White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
 This is perhaps unsurprising.
@@
 Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (A.
+White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}).
 As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
@@
 Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are ``treated'' by the policy change.
+Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are assigned ``treatment'' by the policy change.
 Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
@@
 Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
+A.
+White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
 She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
@@
 White, Ariel.
-2019.
+2019a.
 ``Misdemeanor Disenfranchisement?
@@
 \url{https://doi.org/10.1017/s000305541800093x}.
+
+\leavevmode\hypertarget{ref-White2019a}{}%
+---------.
+2019b.
+``Family Matters?
+Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
+\url{https://doi.org/10.1017/s0003055418000862}.
 

### 2019-11-06 markdown/part2_standalone.Rmd
write

--- 2019-07-05 markdown/part2_standalone.Rmd
@@
 ---
-title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
+title: "Voting Rights Restoration and Post-Supervision Turnout"
 author: "Kevin Morris"
@@
     latex_engine: xelatex
-bibliography: "../research/citations.bib"
+    # includes: 
+    #   in_header: "../code/misc/preamble.tex"
+    keep_tex: true
+bibliography: "citations.bib"
 link-citations: yes
+fontsize: 12pt
 header-includes:
     - \usepackage{setspace}\doublespacing
+    - \usepackage{rotating}
+
 ---
@@
 library(kableExtra)
+library(data.table)
 library(stargazer)
@@
 
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
+In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
@@
 
-Although felony disenfranchisement is highly disempowering for the individuals impacted by the policy, the political consequences of these laws is not entirely clear.
-The number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those formally disenfranchised.
-Previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-
-These spillover effects have been identified using a wide variety of empirical methods.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans, finding that "African American disenfranchisement plays a unique role in predicting African American voter turnout" (p.
-800).
-@Ochs2006, @Bowers2009, and @Walker2014 similarly exploit state-level differences to estimate the (de)mobilizing effect of felony disenfranchise on eligible voters.
-As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
-724).
-@Burch2013 used neighborhood-level data to further investigate how felony disenfranchisement operates the sub-state level.
-She finds that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (p.
-185).
-
-It is clear that although felony disenfranchisement directly impacts relatively few potential voters, its demobilizing character reaches many citizens who are not formally disenfranchised.
-Considering the characteristics of the neighborhoods that incarcerated individuals call home --- neighborhoods that are among the most marginalized --- felony disenfranchisement is a large problem.
-
-In this paper, I examine the effect of Executive Order 181 in New York State.
-Signed in April of 2018, the Executive Order restored voting rights to many individuals on parole.
-Prior to Executive Order 181, individuals were disenfranchised while incarcerated or on parole.
-Most individuals who were on parole on Election Day in 2018 were allowed to vote, a major step forward for democracy in the Empire State.
-In this paper, however, I focus on individuals who were discharged from parole prior to Election Day: individuals who would not have been directly disenfranchised by felony disenfranchisement policies even in the absense of the Executive Order.
-I hypothesize that Executive Order 181 increased turnout among this population, even though their eligibility was not directly impacted by the policy change.
-
-# Turnout Among the Formerly Disenfranchised {-}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that, if this group had been allowed to vote, they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Uggen and Manza demonstrated that felony disenfranchisement can have material political consequences.
-
-In subsequent years, new research called into question Uggen and Manza's assumption that disenfranchised individuals would vote at rates comparable to their sociodemographic peers.
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-@Haselswerdt2009 matched release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [@Meredith2013; @Meredith2015].
-
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characterstics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the vast majority of low turnout among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (p.
-1144).
-
-@White2019, however, indicates that interaction with the criminal justice system for individuals in the context of arrests for misdemeanor charges may have depressive effects on turnout.
-This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-The findings in @White2019 agree with much previous research which shows that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (p.
-829).
-
-Some research has been done in this area.
-@Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-@Meredith2013, however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect from notification in the absense of eligibility changes.
-
-@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (p.
-924).
-It is not clear whether this increase in turnout is *undoing* the depressive effect of incarceration or boosting the participation of individuals whose (low) propensity to vote was unaffected by incarceration.
-
-Executive Order 181 is expected to have increased turnout among formerly incarcerated individuals through two primary mechanisms.
-The first addresses the impact of negative experiences with the state.
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel explicitly invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-
-Secondly, Executive Order 181 is expected to dispel confusion about eligibilty.
-Since @Downs1957, political scientists have argued that an individual's propensity to vote is decided at least in part by comparing the expected costs of casting a ballot with the expected benefits of doing so.
-The costs of voting while disenfranchised (intentionally or otherwise) are exceedingly high, as high-profile cases in states such as Texas have made clear in recent years [@Flynn2018].
-If an individual assessment of the cost of voting is determined in part by the product of their uncertainty about their eligibility and the cost of casting a ballot while ineligible, formerly disenfranchised individals are unlikely to participate unless they are fully certain of their eligibility.
-As @Drucker2005 and others have detailed, however, many formerly incarcerated individuals are misinformed about their eligibility to cast a ballot.
-By dispelling uncertainty around eligibility, I hypothesize that Executive Order 181 increased turnout among the formerly disenfranchised.
-
-# Research Design {-}
-
 Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
@@
 On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-While the Executive Order is obviously beneficial for individuals who are on parole on Election Day (and therefore would have been disenfranchised but for the policy change), the Executive Order may have benefits even for the individuals whose eligibility was not directly impacted.
-Under the Executive Order, parole officers were required to inform their parolees of their new status under the law.
-In the analysis that follows, I look only at individuals who were discharged before October 10^th^, 2018 (the registration deadline in New York State).
-Limiting the analysis to individuals whose eligibility to cast a ballot was not directly impacted by the policy change allows me to explore whether having one's rights restored while still under formal supervision is effective at raising that individual's propensity to vote even once supervision is ended.
+Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
+The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
-In order to estimate the impact of Executive Order 181 on turnout, I begin by exploring whether the turnout rate among individuals discharged from parole after the policy went into effect systematically differed from those discharged earlier.
-This analysis uses an individual-level logistic regression that estimates individuals' turnout rates and controls for various individual-level characteristics.
-Such an analysis sheds light on the impact of rights restoration on the turnout rate of the formerly disenfranchised population as a whole.
+Of course, re-extending the right to vote is only the first step.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
+As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters [see, for instance, @Fraga2018].
+To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
 
-Such an approach has some benefits: supposing that policymakers care about the political representation of formerly incarcerated residents as a group, then examining the group turnout rate can yield important insights.
-This approach does not, however, address the impact having one's rights restored has on individual propensity to vote.
-To interrogate whether an individual's propensity to vote is increased if she has her rights restored prior to discharge from parole, I test whether individuals who had their rights restored prior to discharge from parole turned out at higher rates in midterm election.
-Because the restoration of voting rights is not randomly assigned, I employ an instrumental variables approach, as discussed in greater detail below.
+There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
+Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
+New York's correction code provides no more guidance other than that "upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration" (Section 75 of the New York State Correction Law).
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[Link: https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
+In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" [emphasis in the original].
+Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
+
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
+This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
+However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
+
+# Turnout Among the Formerly Disenfranchised {-}
+
+In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
+@Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
+Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
+She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
+Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
+Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [e.g.
+@Meredith2013; @Meredith2015].
+
+Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
+The causal effect of incarceration on participation is the subject of some debate within the field.
+Individuals who go to prison share many characteristics with lower propensity voters generally.
+Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
+In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
+They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
+
+@White2019, however, comes to a different conclusion.
+Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
+This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
+By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
+Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
+
+Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
+A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
+Whether or not incarceration *causes* low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
+
+The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
+I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
+
+Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
+Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
+@Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
+They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
+@Meredith2013, however, examines states where so-called notification laws went into effect.
+Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
+@Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
+@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+"Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
+Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
+
+Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
+@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
+To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
+Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
+Such repairs may lead to increased propensity to participate in elections in the future.
 
@@
 These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These data are used to determine when individuals were incarcerated or on parole, and when they finished their parole supervision.
+These data are used to determine when individuals were incarcerated or on parole, when they finished their parole supervision, and demographic information such as age and race.
 
-The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the state, I constructed a list of the individuals who were on parole and had their rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
+By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.]
+For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 
-Below, I employ a logistic regression to investigate the impact of Executive Order 181 on turnout among all parolees discharged after the policy went into effect.
-Because not all individuals had their rights restored, I follow this analysis with an instrumental-variables approach to determine the effect of rights restoration on individuals whose rights were in fact formally restored prior to discharge from parole.
-
-## Trends in Turnout {-}
 Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
 It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
+However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
 The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, `r percent(readRDS("./temp/turnout_in_16_finished_10.rds"), accuracy = 0.1)` cast a ballot in the 2016 election, while just `r percent(readRDS("./temp/turnout_in_16_finished_15.rds"), accuracy = 0.1)` of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] Figure \ref{fig:to-16} plots turnout rates by month of parole discharge.
-A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January, 2010 through April, 2016, and extended through October, 2016.
+For instance, of the individuals last discharged from parole in 2010, `r round(readRDS("./temp/turnout_in_16_finished_10.rds") * 100, digits = 1)` percent cast a ballot in the 2016 election, while just `r round(readRDS("./temp/turnout_in_16_finished_15.rds") * 100, digits = 1)` percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
+This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
 
@@
 However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated forindividuals discharged in 2017 or 2018 before the executive order was enacted.
 
-```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Presidential Election", fig.pos = "H", fig.align = "center"}
+```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
 j <- readRDS("./temp/to18_chart_recent.rds")
@@
 ```
-Formalizing this chart into an individual-level logit model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018.
+Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
+The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
 
-```{r logit-reg-chunk-short, echo = F, results = "asis", message = F}
+```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
 load("./temp/short_term_to.rdata")
@@
           header = F,
-          type = "latex",
-          covariate.labels = c("Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
+          type = "latex", notes.align = "l",
+          covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
+                               "D(Male)", "Age (Years)",
+                               "Years on Parole"),
           dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit-short) Individual-Level Logit Model",
+          title = "\\label{tab:to-18-logit-short} Individual-Level Logit Models",
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "X", "X"),
-                         c("Felony Class Fixed Effects", "X", "X")),
-          table.layout = "-dm#-t-a-s-n")
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
+                         c("Felony Class FE", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(
+            sqrt(diag(vcovHC(date_to_model_notime, type = "HC0"))),
+            sqrt(diag(vcovHC(date_to_model_time, type = "HC0"))))
+          )
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table3.tex", quote = F, col.names = F,
+            row.names = F)
+
 ```
+\input{"../temp/table3.tex"}
+The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
+A Chi-squared test confirms that the model is not improved when controls for time are included.
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
 
-The inclusion of time controls in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.
-
-Although Governor Cuomo signed the Executive Order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
+Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
 Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 21^st^ that the majority of discharged parolees had their voting rights restored prior to their discharge.
+It is not until May 18^th^ that parolees consistently had their voting rights restored prior to discharge.
 ```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
@@
 Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 21^st^, 2018, is akin to a randomly assigned treatment.
-Any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the Executive Order.
+Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
+Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are assigned "treatment" by the policy change.
+Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 21^st^, 2018, turned out at higher rates than those discharged earlier.
+To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
+With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
+To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
+
+```{r rd-demo-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_rd_time.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of \"Control\" and \"Intend-to-Treat\" Groups", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
+\\\\Shows demographics for individuals discharged from parole \\\\between January 1st, 2017, and October 12th, 2018.}}}"
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of \"Control\" and \"Intend-to-Treat\" Groups}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table_whatever2.tex"}
+
+In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
 
-```{r logit-reg-chunk, echo = F, results = "asis", message = F}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
 load("./temp/individual_turnout_18_itt.rdata")
@@
           header = F,
-          type = "latex",
+          type = "latex", notes.align = "l",
           covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
+                               "Years on Parole"),
           dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:to-18-logit) Individual-Level Logit Model",
+          title = "\\label{tab:to-18-logit} Individual-Level Logit Models",
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
+                         c("Felony Class FE", "", "", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(
+            sqrt(diag(vcovHC(model1, type = "HC0"))),
+            sqrt(diag(vcovHC(model2, type = "HC0"))),
+            sqrt(diag(vcovHC(model3, type = "HC0")))))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table4.tex", quote = F, col.names = F,
+            row.names = F)
+
 ```
+\input{"../temp/table4.tex"}
 
 Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, county of incarceration, and race.
-Model 3 adds sentence-specific information to Model 2: the number of counts in the individual's most recent sentence, the amount of time they spent on parole, and the class(es) of felony for which they were convicted.
+Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
+Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-Each model also indicates that individuals discharged from parole after the Executive Order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model2[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+The turnout rate for the control group was `r format(round(100*readRDS("./temp/control_to_18.rds"), digits = 1), nsmall = 1)` percent.
 
@@
 
-Table \ref{tab:to-18-logit} indicates that Executive Order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which having one's voting rights restored prior to discharge from parole impacts one's propensity to vote.
-To answer that question, we must specifically control for whether an individual did have his rights restored before he was discharged from parole.
+Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
+Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+
+Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied." Noncitizens and individuals who violated their terms of parole were not eligible for rights restoration, and therefore did not "comply" with the treatment.
+Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
 
 An individual's propensity to vote can be expressed using the following equation:
-$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + \epsilon_{i} $$
+$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} $$
 Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
+*Z* is a vector of other factors (such as age and race) known to influence voter turnout.
 Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote.
-Neither of these assumptions are valid: an individual's eligibility to have her rights restored is almost certainly correlated with whether or not they actually were restored.
-Moreover, eligiblity for rights restoration is likely correlated with propensity to vote.
-Non-citizens, for instance, were not eligible for rights restoration under EO181 because non-citizens cannot vote in New York State.
-Similarly, individuals who were re-arrested while on parole were not eligible to have their rights restored.
-Citizenship status and recency of arrest are both highly correlated with propensity to vote.
+Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
 
-Our variable of interest, therefore, is likely correlated with our regression error.
-Such a correlation inhibits our ability to simply regress turnout on a dummy indicating whether an individual's rights were restored.
-The standard way of dealing with such a problem is an instrumental variables approach.
+Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
+Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
+Individuals who were discharged from parole but are not U.S.
+citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
+This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
+Table \ref{tab:demos-restorees} presents demographic differences between individuals who had their rights restored and those who did not among the intend-to-treat group (we cannot tell who would and would not have complied among the control group).
+
+```{r restoree-demo-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_restorees.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
+\\\\Shows demographics for individuals discharged from parole \\\\between May 18th, 2018, and October 12th, 2018.}}}"
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table_whatever.tex"}
+
+Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
+Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
+Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
+This is likely a reflection of citizenship status.
+Table \ref{tab:demos-restorees} indicates that *b~1~* is very probably correlated with *Y*.
+An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
+
+The standard way of dealing with such a problem is an instrumental variables approach [@Angrist1996].
 Such an approach has been widely used in the context of voter turnout [e.g.
-@Gerber2000; @Green2012].
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with the error of the regression.
+@Ansolabehere1999; @Gerber2000; @Milligan2004;  @Lassen2004; @Sondheimer2010].
+A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
 
 In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is a function of whether he was discharged from parole after Executive Order 181 took effect.
+The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
 As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the Executive Order, therefore, satisfies the criteria for an instrumental variable.
+A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
 
-Table \ref{tab:iv-models} presents the two-stage least-squares regression estimates exploring the turnout impact of rights restoration.
-As discussed above, the instrumental variable used in these regressions measure whether an individual was discharged from parole after Executive Order 181 went into effect.
-These models are estimated using an OLS sepcification.
-Although the dependent variable remains dichotomous (as do both the independent variable of interest and the instrument), @Angrist2008 recommends using OLS in instrumental variables analyses even under such conditions.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the "local average treatment effect." The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
+Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
+Some research [e.g.
+@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
+The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2014].
+The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
+This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
 
-```{r iv-reg-chunk-short, echo = F, results = "asis", message = F}
+Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
+Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
+Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
+For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
+
+\input{"../temp/iv_clean.tex"}
+
+The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
+After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
+Though these numbers may seem small, they represent relatively large gains.
+Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
+
+### Racial Variation {-}
+
+Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
+In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
+These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
+Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
+
+\input{"../temp/iv_clean_race.tex"}
+
+```{r nameless-chunk, echo = F, results = "asis", message = F}
 load("./temp/iv_individual_turnout_18.rdata")
-
-summ.fit1 <- summary(iv1, diagnostics = T)
-summ.fit2 <- summary(iv2, diagnostics = T)
-summ.fit3 <- summary(iv3, diagnostics = T)
-
-stargazer(iv1, iv2, iv3,
-          header = F,
-          type = "latex",
-          covariate.labels = c("D(Rights Restored)", 
-                               "D(Male)", "Age (Years)",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "(\\#tab:iv-models) Second-Stage Regression",
-          table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X"),
-                         c("\\hline \\\\[-1.8ex]"),
-                         c(rownames(summ.fit1$diagnostics)[1],
-                           round(summ.fit1$diagnostics[1, "p-value"], 2),
-                           round(summ.fit2$diagnostics[1, "p-value"], 2),
-                           round(summ.fit3$diagnostics[1, "p-value"], 2)),
-                         c(rownames(summ.fit1$diagnostics)[2],
-                           round(summ.fit1$diagnostics[2, "p-value"], 2),
-                           round(summ.fit2$diagnostics[2, "p-value"], 2),
-                           round(summ.fit3$diagnostics[2, "p-value"], 2))),
-          omit.stat = "ser",
-          table.layout = "-dm#-t-as-n",
-          notes = c("Table reports p-values for Weak Instruments",
-                    "and Wu-Hausman tests.",
-                    "First-stage equations use dummy indicating",
-                    "whether individual was discharged after the",
-                    "Exectutive Order went into effect. The",
-                    "first-stage equations also include",
-                    "covariates in models 2 and 3."))
 ```
 
-Table \ref{tab:iv-models} indicates that using discharge date of on or after May 21^st^ as an instrument for rights restoration is warranted.
-The p-values of the Wu-Hausman tests hover just below 0.05, while the Weak Instruments test is highly significant in each model.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
+Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration *nearly doubled* these individuals' likelihood of casting a ballot.
+We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
 
-Model 3 in Table \ref{tab:iv-models} indicates that having one's rights restored caused an individual's propensity to vote to increase by `r round(summary(iv3)[["coefficients"]][2,1] * 100, digits = 2)` percentage points.
-Approximately `r round(readRDS("./temp/to_restored.rds") * 100, digits = 2)` percent of individuals who had their rights restored cast a ballot in 2018, indicating that Executive Order 181 increased the propensity to vote by around `r round(summary(iv3)[["coefficients"]][2,1] / (readRDS("./temp/to_restored.rds") - summary(iv3)[["coefficients"]][2,1]) * 100, digits = 1)` percent.
-Unsurprisingly, this estimate is higher than the estimates reported in Table \ref{tab:to-18-logit}.
-Because Executive Order 181 theoretically only impacts individuals whose rights were actually restored prior to discharge, the magnitude of the estimated effect on this subgroup must be larger than the group taken as a whole.
+Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
+Some of this may be explained by different propensities to vote.
+@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
+The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
-# Discussion {-}
-As discussed above, the effect of incarceration on individuals' propensity to vote is an open question for the literature.
-What is known, however, is that individuals who have been to prison were likely to vote at very low rates prior to incarceration, and continue to vote at very low rates after incarceration.
-Whether incarceration reduces political participation or not, increasing the propensity to vote of individuals who have been to prison is a laudable goal.
-Prisoners overwhelmingly come from marginalized communities with much to gain from policy.
-Executive Order 181 appears to have substantially boosted turnout among formerly incarcerated individuals through two mechanisms.
-The first is obvious: parolees who would not have been eligible to vote absent the Executive Order were allowed to cast ballots in the midterm elections.
-The second mechanism is less obvious.
-Individuals who finished parole between May 21^st^, 2018, and  October 10^th^, 2018, would have been eligible to participate in the election even if the governor had not signed the Executive Order.
+There is reason to believe this may be the case.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
 
-The precise mechanism through which the rules change increased turnout among these individuals is not clear.
-It is possibly a social mechanism: having an officer of the state affirm one's eligibility (and therefore reunion with the body politic) could be responsible for the increase.
-It could be due to better information: individuals whose rights were restored were likely far more confident of their eligibility to vote than other formerly incarcerated individuals.
-It could also be an issue of timing: individuals whose rights were restored in May of 2018 but were not discharged from parole until early October had four additional months to register to vote than they would have absent the Executive Order.
-In reality, the turnout boost is likely due to a combination of different factors.
+```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/black_to.rdata")
 
-Despite the substantial increase in turnout thanks to Executive Order 181, turnout among formerly incarcerated individuals remained stubbornly low in the 2018 midterm elections.
-Just 3.2 percent of individuals who had their rights formally restored and finished parole before the registration deadline.
-Although the successes of Executive Order 181 should be celebrated, more must be done to encourage formerly incarcerated individuals to participate in the political process.
+stargazer(black_to1, black_to2, black_to1b, black_to2b, black_to3, black_to4,
+          column.separate = c(2, 2, 2),
+          column.labels = c("Discharged Prior to EO", "ITT Group", "Treatment Group"),
+          header = F,
+          type = "latex", notes.align = "l",
+          covariate.labels = c("D(Black)", "D(Male)", "Age (Years)",
+                               "Years on Parole"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "\\label{tab:black-to} Turnout Rate Among Former Parolees",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          omit = "felony",
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F,
+          notes = "TO REPLACE",
+          se = list(
+            sqrt(diag(vcovHC(black_to1, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to2, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to1b, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to2b, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to3, type = "HC0"))),
+            sqrt(diag(vcovHC(black_to4, type = "HC0")))
+          ))
 
-\newpage
-# References {-}
-<div id="refs"></div>
+j <- fread("./temp/test.tex", header = F, sep = "+")
 
-\newpage
+note.latex <- "\\multicolumn{7}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors in parentheses.
+\\\\Latino individuals excluded from Models 2, 4, and 6.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}  \\cmidrule(l{3pt}r{3pt}){6-7}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table5.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table5.tex"}
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
+These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
+Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
+In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
+
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+The intervention, therefore, appears to have a leveling effect.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
+
+### Variable Treatment Intensity {-}
+
+There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
+Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
+This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
+Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
+Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
+If the individual did not have voting rights restored, this variable takes the value 0.
+This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
+It is coded as 0 for individuals discharged prior to the implementation of the executive order.^[Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.]
+
+\input{"../temp/iv_clean_months.tex"}
+
+Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
+Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
+It is worth noting that the *R^2^* in Model 1 is higher than the *R^2^* in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
+
+Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
+One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
+
+## Discussion {-}
+
+Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both *de jure* and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
+Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
+That changed in 2018.
+On October 12^th^ --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
+Without the executive order, every single one of these individuals would have been barred from participating.
+Though turnout among this group was low (just `r comma(readRDS("./temp/new_votes_18.rds"))`, or `r round((readRDS("./temp/new_votes_18.rds")/readRDS("./temp/new_eligible_18.rds"))*100, 2)` percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
+
+As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
+In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
+In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` percent.
+This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
+Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
+
+The mechanism through which the rules change increased turnout among these individuals is not clear.
+@Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
+Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
+On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+Such interactions may undo some of the negative socialization identified by @Weaver2010.
+
+It could also be a story of better information.
+As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
+Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
+In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
+
+Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
+Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
+The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
+This discrepancy could be caused by a number of different factors.
+Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
+As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
+The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
+It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
+
+We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
+Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
+Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners [@Hemmens2000].
+Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
+@WHITE2014, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2016].
+
+Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
+This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
+Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
+In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
+Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
+
+Re-enfranchising voters while they are still under formal supervision is obviously beneficial to the individuals who are on parole on election day; such policies allow them to make their voices heard.
+The case of Executive Order 181 also indicates that restoring voting rights prior to parole discharge has further benefits.
+In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
+This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
+By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
+
 # Appendix A {-}
 
-In the [Individual-Level Turnout Regressions] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 21^st^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
+In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
 Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-```{r logit-reg-chunk-16, echo = F, results = "asis", message = F}
+
+```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
 load("./temp/individual_turnout_16.rdata")
@@
           header = F,
-          type = "latex",
-          covariate.labels = c("D(Discharged on or after May 21\\textsuperscript{st}, 2016)", "Days Since Discharged",
-                               "Days Since Discharged\\textsuperscript{2}", "D(Male)", "Age (Years)",
+          type = "latex", notes.align = "l",
+          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
+                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
                                "Time Spent on Parole (Years)"),
           dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "(\\#tab:to-16-logit) Individual-Level Logit Model",
+          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
           table.placement = "H",
-          omit = c("county", "race", "felony"),
-          add.lines=list(c("County Fixed Effects", "", "X", "X"),
-                         c("Race / Ethnicity Fixed Effects", "", "X", "X"),
-                         c("Felony Class Fixed Effects", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n")
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
+                         c("Felony Class FE", "", "", "X")),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F,
+          notes = "TO REPLACE",
+          se = list(
+            sqrt(diag(vcovHC(model1, type = "HC0"))),
+            sqrt(diag(vcovHC(model2, type = "HC0"))),
+            sqrt(diag(vcovHC(model3, type = "HC0")))
+          ))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table55.tex", quote = F, col.names = F,
+            row.names = F)
 ```
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot (the p-value on the coefficient of interest exceeds 0.5 in each model).
+\input{"../temp/table55.tex"}
+
+Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
 The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
+
+\newpage
+# References {-}
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\noindent

### 2019-11-06 markdown/part2_standalone.tex
write

--- 2019-10-18 markdown/felony_disenfranchisement_nys.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={In the Shadow of Jim Crow},
+            pdftitle={Voting Rights Restoration and Post-Supervision Turnout},
             pdfauthor={Kevin Morris},
@@
 
-  \title{In the Shadow of Jim Crow}
+  \title{Voting Rights Restoration and Post-Supervision Turnout}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \subtitle{Felony Disenfranchisement and New York State}
     \author{Kevin Morris}
@@
   \postdate{\par}
-    \date{October 17, 2019}
+    \date{November 06, 2019}
 
@@
 \pagenumbering{gobble}
-
 \tableofcontents
@@
 \hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
+\section*{Introduction}\label{introduction}}
+\addcontentsline{toc}{section}{Introduction}
+
+The political history of the United States has been characterized by a general, if unlinear, trend toward universal sufferage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
 At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
-
-Due to economic and racial segregation, these effects are highly spatially concentrated.
-Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
-Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health (Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}) and are more likely to suffer from anxiety and exhibit symptoms of trauma (Geller et al.
-\protect\hyperlink{ref-Geller2014}{2014}).
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted (Clear \protect\hyperlink{ref-Clear2008}{2008}), while concentrated policing has also been credited with having a ``chilling effect'' on neighborhoods' willingness to reach out for help to local governments (Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-
-Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
-The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}; Gerber et al.
-\protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (8).
-
-The rest of this paper is structured as follows: \textbf{Section 2} discusses the historical context of felony disenfranchisement in the United States, and \textbf{Section 3} provides an overview of the scholarship conducted in the field over the past two decades.
-\textbf{Section 4} provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
-\textbf{Section 5} presents the data sources used throughout the project.
-
-\textbf{Section 6} proposes a new definition of ``lost voters'' in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised \emph{and} have a history of participating in elections.
-This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely \emph{understated} the degree to which they are impacted by spatially concentrated patterns of criminal justice.
-A discussion of the results follows.
-
-\textbf{Section 7} turns from the 2017 mayoral election to the 2018 midterm elections.
-In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
-I employ an instrumental variables approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
-I focus specifically on individuals who would have been eligible to cast a ballot even if the executive order had not gone into effect.
-A discussion of the results follows.
-
-\textbf{Section 8} synthesizes the findings presented in the rest of the paper and provides concluding remarks.
-
-\hypertarget{background-of-felony-disenfranchisement-in-the-united-states}{%
-\section{Background of Felony Disenfranchisement in the United States}\label{background-of-felony-disenfranchisement-in-the-united-states}}
-
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
+Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
+And yet, despite the United State's purported march toward ever-more-inclusive systems of democracy, one large group of American citizens is formally barred from voting.
+In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised today (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
+These disenfranchising policies provide a stark reminder that our democracy is not as inclusive as we might like to tell ourselves.
+
+In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
 This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 
-Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Burch (\protect\hyperlink{ref-Burch2010}{2010}) explains, ``If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were \emph{intended} to'' (4; emphasis in the original).
-Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans (Behrens, Uggen, and Manza \protect\hyperlink{ref-Behrens2003}{2003}).
-In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being ``niggerized'' (Shofner \protect\hyperlink{ref-florida_1868}{1963}, 374).
-
-The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
-Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
-In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just 12.1 percent of the voting age population in 2016, they accounted for 36.5 percent of the disenfranchised population (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate (Morris \protect\hyperlink{ref-Morris2019}{2019}).
-
-\hypertarget{academic-literature-and-felony-disenfranchisement}{%
-\section{Academic Literature and Felony Disenfranchisement}\label{academic-literature-and-felony-disenfranchisement}}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout (Miles \protect\hyperlink{ref-Miles2004}{2004}; Hjalmarsson and Lopez \protect\hyperlink{ref-Hjalmarsson2010}{2010}).
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects (Miles \protect\hyperlink{ref-Miles2004}{2004}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-Others have used survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
+As Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
+This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
+This disparity is likely due \emph{not} to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+
+Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
+For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
+New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
+
+On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
+The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+
+Of course, re-extending the right to vote is only the first step.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
+As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
+To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
+
+There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
+Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
+New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
+In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
+Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
+
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
+This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
+However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
+
+\hypertarget{turnout-among-the-formerly-disenfranchised}{%
+\section*{Turnout Among the Formerly Disenfranchised}\label{turnout-among-the-formerly-disenfranchised}}
+\addcontentsline{toc}{section}{Turnout Among the Formerly Disenfranchised}
 
@@
 Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
-
-A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
-{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the \emph{state} level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
-
-\hypertarget{what-this-paper-examines}{%
-\section{What This Paper Examines}\label{what-this-paper-examines}}
-
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods (e.g.~Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}; Clear \protect\hyperlink{ref-Clear2008}{2008}; Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
-
-I begin by offering a refinement of our understanding of who lost voters are.
-Previous studies exploring the relationship between felony disenfranchisement and spillover effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a \emph{demonstrated history} of participation are disenfranchised.
-By matching ``treated'' neighborhoods (neighborhoods with lost voters) to ``untreated'' neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
-
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.\footnote{Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those currently in prison for felony offenses are disenfranchised.} Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I use administrative data to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
-
-The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather than think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
-This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
-The second part of this project seeks to do just that.
-
-\hypertarget{framing-for-turnout-effects}{%
-\subsection*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
-\addcontentsline{toc}{subsection}{Framing for Turnout Effects}
-
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, 46,232 individuals were imprisoned in New York State in early 2019, compared with 11.6 million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
-
-Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
-Moreover, voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
-It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-I therefore hypothesize that felony disenfranchisement does have locally bounded spillover effects, and that turnout in neighborhoods with lost voters systematically vote at a lower rate than others.
-
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
-Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
-These effects are particularly pronounced for women of color.
-Although Walker and others demonstrate that these actions largely take place outside of the voting booth, they may moderate spillover effects on turnout.
-
-\hypertarget{theoretical-implications-for-reduced-turnout}{%
-\subsubsection*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
-\addcontentsline{toc}{subsubsection}{Theoretical Implications for Reduced Turnout}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community (e.g.~King and Erickson \protect\hyperlink{ref-King2016}{2016}; Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: Kinsella, McTague, and Raleigh (\protect\hyperlink{ref-Kinsella2015}{2015}) examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes (Martin \protect\hyperlink{ref-Martin2003}{2003}).
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-``{[}H{]}igher citizen participation is rewarded,'' Martin and Claibourn (\protect\hyperlink{ref-Martin2013}{2013}) concludes, ``with enhanced policy responsiveness'' (59).
-Griffin and Newman (\protect\hyperlink{ref-Griffin2005}{2005}) finds similar effects in the United States Senate, demonstrating that ``voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not'' (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
-
-Although research on the impact of local turnout on city-wide policy is scarce, Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) examines the impact of senior turnout on ``senior-friendly'' policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials \emph{are} responsive to senior turnout when seniors ``are a more cohesive, meaningful group'' (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of \emph{neighborhood} turnout rates on city policy.
-
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}), however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) tells us that ``Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils'' (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
-
-\hypertarget{turnout-among-formerly-disenfranchised-individuals}{%
-\subsection*{Turnout Among Formerly Disenfranchised Individuals}\label{turnout-among-formerly-disenfranchised-individuals}}
-\addcontentsline{toc}{subsection}{Turnout Among Formerly Disenfranchised Individuals}
 
@@
 This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
+(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
 By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
@@
 \hypertarget{data}{%
-\section{Data}\label{data}}
+\section*{Data}\label{data}}
+\addcontentsline{toc}{section}{Data}
 
@@
 
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole and demographic characteristics.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors available.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
+The criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
+This analysis is limited to individuals incarcerated for felony offenses.
+Individuals convicted of misdemeanors are not disenfranchised in New York State.
+These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
+These data are used to determine when individuals were incarcerated or on parole, when they finished their parole supervision, and demographic information such as age and race.
+
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
+By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
+For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
+For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 
-Most states in the United States are required to maintain files with information on all registered and recently purged voters.
+Most states in the United States are required to maintain files with information on all registered voters.
 In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The first section of this paper uses a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
-The second section of this paper uses a snapshot from March 3\textsuperscript{rd}, 2019.
-
-The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered (and participated) in the past but have since been purged due to a felony conviction.
-
-\hypertarget{geocoding}{%
-\subsection*{Geocoding}\label{geocoding}}
-\addcontentsline{toc}{subsection}{Geocoding}
-
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6 percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
+The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
+I use a snapshot of the registered voter file from March 3\textsuperscript{rd}, 2019.
 
@@
 
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
+Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
 I match individuals in each dataset using first name, middle name, last name, and date of birth.
@@
 
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
-
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andew Doe'' and ``John Anthony Doe'' would not.
+
+There are two types of potential error in this methdology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
 False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
@@
 Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
+I slightly alter the date of birth reported in the parole discharge dataset to create false records.
+Comparing the number of matches between these ``fake'' discharge records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
+This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
 
@@
 \centering
-\fontsize{10}{12}\selectfont
 \begin{tabular}{cc}
@@
 \midrule
-Actual Birthdate & 20,955\\
-Birthdate + 35 Days & 105\\
-Birthdate - 35 Days & 92\\
+Actual Birthdate & 68,340\\
+Birthdate + 35 Days & 480\\
+Birthdate - 35 Days & 419\\
 \bottomrule
@@
 A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
-This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
-Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-
-\hypertarget{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}{%
-\section{Effects of Felony Disenfranchisement on Neighborhood Turnout Levels}\label{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}}
-
-\hypertarget{identification-of-lost-voters}{%
-\subsection*{Identification of Lost Voters}\label{identification-of-lost-voters}}
-\addcontentsline{toc}{subsection}{Identification of Lost Voters}
-
-In this analysis, I offer a different definition of ``lost voter'' than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- \emph{potentially} lost voters --- and turnout.
-Such an approach is informative for understanding the impact of disenfranchisement.
-Many young men, for instance, are admitted to prison each year.
-If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
-Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
-
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration (e.g.~Gerber et al.
-\protect\hyperlink{ref-Gerber2017}{2017}).
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement \emph{per se} but rather to their incarceration.
-At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
-
-Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, ``lost voters'' are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote (Gerber, Green, and Shachar \protect\hyperlink{ref-Gerber2003}{2003}).
-As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
-Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
-
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
-
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were 2,518 such lost voters within New York City as of the 2017 general election, and 6,166 statewide.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/citywide-map-1}
-
-}
-
-\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
-\end{figure}
-
-The spatial concentration of lost voters is readily apparent.
-In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
-
-\hypertarget{testing-for-neighborhood-turnout-effects}{%
-\subsection*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
-\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Effects}
-
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
-
-\begin{table}[H]
-
-\caption{\label{tab:distribution-lost}\label{tab:nhood-dist} Distribution of Lost Voters}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}{ccc}
-\toprule
-Number of Lost Voters & Count of Block Groups & Count of Tracts\\
-\midrule
-0 & 4,272 & 1,008\\
-1 & 1,023 & 416\\
-2 & 301 & 212\\
-3 & 107 & 138\\
-4 & 30 & 51\\
-5 & 19 & 46\\
-6 & 9 & 23\\
-7 & 3 & 15\\
-8 & 5 & 13\\
-9 & 0 & 7\\
-10 & 0 & 2\\
-11 & 0 & 2\\
-12 & 0 & 2\\
-14 & 0 & 2\\
-\hline
-Total & 5,769 & 1,937\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}), using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
-Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
-
-\hypertarget{match-output}{%
-\subsubsection*{Match Output}\label{match-output}}
-\addcontentsline{toc}{subsubsection}{Match Output}
-
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk}\label{tab:blockgroup-table} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.34 & 93.02 & 89.13 & 85.79 & 79.01\\
-\% Non-Hispanic Black & 0.38 & 0.16 & 0.38 & 0.37 & 97.93 & 94.46 & 94.41 & 92.77\\
-\% Non-Hispanic White & 0.17 & 0.40 & 0.17 & 0.18 & 96.37 & 96.06 & 95.56 & 92.99\\
-Median Income & 51,792.52 & 72,612.36 & 51,792.52 & 52,768.97 & 95.31 & 93.47 & 88.62 & 74.49\\
-\% With Some College & 0.64 & 0.70 & 0.64 & 0.64 & 90.75 & 94.14 & 90.29 & 80.70\\
-Median Age & 35.99 & 38.34 & 35.99 & 36.25 & 89.18 & 90.84 & 85.93 & 78.56\\
-Registration Rate & 1.00 & 0.94 & 1.00 & 0.98 & 59.77 & 79.89 & 72.28 & 60.42\\
-\% Democrats & 0.75 & 0.65 & 0.75 & 0.75 & 96.22 & 96.52 & 94.92 & 91.02\\
-\% Noncitizen & 0.16 & 0.16 & 0.16 & 0.16 & 16.58 & 14.80 & -0.06 & -60.28\\
-\% Won by City Council Representative & 0.85 & 0.81 & 0.85 & 0.86 & 89.44 & 83.40 & 79.88 & 70.11\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-table-chunk}\label{tab:tract-table} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.32 & 0.22 & 0.32 & 0.33 & 98.55 & 85.12 & 82.84 & 73.01\\
-\% Non-Hispanic Black & 0.36 & 0.14 & 0.36 & 0.36 & 99.77 & 88.74 & 88.36 & 83.30\\
-\% Non-Hispanic White & 0.19 & 0.42 & 0.19 & 0.19 & 97.50 & 96.32 & 91.29 & 79.95\\
-Median Income & 53,409.42 & 71,754.50 & 53,409.42 & 55,422.41 & 89.03 & 90.83 & 84.86 & 62.84\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 90.19 & 93.52 & 88.89 & 73.01\\
-Median Age & 35.76 & 38.04 & 35.76 & 35.80 & 98.12 & 74.51 & 80.56 & 71.91\\
-Registration Rate & 0.94 & 0.90 & 0.94 & 0.93 & 76.63 & 86.16 & 84.24 & 71.28\\
-\% Democrats & 0.74 & 0.62 & 0.74 & 0.73 & 94.58 & 94.93 & 92.05 & 82.37\\
-\% Noncitizen & 0.16 & 0.17 & 0.16 & 0.17 & -123.57 & 22.48 & 8.85 & -38.79\\
-\% Won by City Council Representative & 0.85 & 0.79 & 0.85 & 0.85 & 93.03 & 81.57 & 79.27 & 67.47\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less White,\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} have much lower median incomes, and a larger share of voters are registered as Democrats.
-
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
-
-\input{"../temp/table1.tex"}
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average 0.9 percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as 2.7 percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just 14.6 percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
-
-\hypertarget{testing-intensity-effects}{%
-\subsubsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
-\addcontentsline{toc}{subsubsection}{Testing Intensity Effects}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on \emph{how many} voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
-In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
-Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
-
-\input{"../temp/table2.tex"}
-
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.87 percentage points.
-The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
-
-The block groups where these depressive effects are not randomly distributed throughout the city.
-They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on \(Lost\ Voters \times Share\ Black\) from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is \(-0.019\times Lost\ Voters \times Share\ Black\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/postest-map-1}
-
-}
-
-\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
-\end{figure}
-
-\hypertarget{toward-a-causal-estimation}{%
-\subsubsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
-\addcontentsline{toc}{subsubsection}{Toward a Causal Estimation}
-
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
-
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were \emph{newly} home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 \emph{did} have a lost voter in 2016, and are thus excluded from this analysis.
-
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
-
-Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
-As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
-
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.35 & 99.00 & 96.26 & 94.92 & 85.60\\
-\% Non-Hispanic Black & 0.33 & 0.16 & 0.33 & 0.33 & 99.02 & 96.57 & 95.67 & 90.10\\
-\% Non-Hispanic White & 0.21 & 0.41 & 0.21 & 0.21 & 99.61 & 94.66 & 92.61 & 82.94\\
-Median Income & 55,852.62 & 73,433.49 & 55,852.62 & 56,287.76 & 97.52 & 87.85 & 86.56 & 72.89\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 99.31 & 92.25 & 87.30 & 71.15\\
-Median Age & 36.17 & 38.47 & 36.17 & 36.04 & 94.37 & 62.35 & 72.27 & 69.84\\
-Registration Rate & 0.98 & 0.94 & 0.98 & 0.95 & 25.09 & 46.49 & 42.16 & 33.06\\
-\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 95.94 & 94.99 & 93.62 & 86.68\\
-\% Noncitizen & 0.17 & 0.16 & 0.17 & 0.17 & 63.46 & 75.01 & 66.85 & 45.71\\
-\% Won by City Council Representative & 0.85 & 0.80 & 0.85 & 0.85 & 88.19 & 83.99 & 78.23 & 67.54\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-chunk-did}\label{tab:did-match-tract} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 94.34 & 86.10 & 84.77 & 75.82\\
-\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.25 & 99.28 & 89.75 & 86.62 & 77.90\\
-\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.02 & 87.45 & 77.73 & 61.67\\
-Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,973.89 & 84.23 & 78.35 & 76.98 & 66.57\\
-\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 96.86 & 86.51 & 83.73 & 76.55\\
-Median Age & 36.36 & 38.23 & 36.36 & 36.39 & 98.60 & 73.15 & 69.00 & 42.11\\
-Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 10.68 & 64.27 & 63.46 & 56.35\\
-\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.23 & 90.06 & 87.35 & 76.31\\
-\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 78.09 & 73.55 & 62.03 & 48.36\\
-\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.91 & 75.59 & 65.98 & 38.91\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups and tracts.
-In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly \emph{higher} turnout prior to intervention.
-D(Treat) \(\times\) D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
-
-\input{"../temp/table1111.tex"}
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated 2,518 voters from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
-However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
-
-As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement impacts Black turnout at the state level, but this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
-Similarly, just 17 percent of individuals in the average block group with a lost voter were White, compared with 40 percent of the population in other block groups.
-Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
-The effects are concentrated in neighborhoods where the most marginalized members of society live.
-These highly concentrated spillover effects are cause for major concern.
-
-Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
-This means that lost voters have no spillover effects in neighborhoods with small Black populations.
-Understanding the magnitude of this finding is important.
-In New York City, there are 165 block groups where the Black community makes up more than 90 percent of the population, with an average of 821 voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of 821, having any lost voter reduced turnout by 20 ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just 2.2 ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood 13.9 ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just 1.5 ballots --- hardly more votes than the lost voter herself.
-
-Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large ``interpretive effects'' (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Lerman and Weaver (\protect\hyperlink{ref-Lerman2013}{2013}), for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, Weaver and Lerman (\protect\hyperlink{ref-Weaver2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
-
-Individuals who live in neighborhoods where police activity is relatively limited may interpret the incarceration of a neighbor as a largely individual phenomenon.
-If it is understood as an isolated or individual event, voters in these neighborhoods who are not incarcerated are not likely to update their view of the state.
-They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
-In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
-It may, rather, be interpreted as another reminder of the government's unfairness.
-If a would-be voter finds herself soured on political participation because of her neighbor's incarceration, she may be less likely to cast a ballot.
-
-This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
-This is perhaps unsurprising.
-A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans (e.g.~Browning and Cao \protect\hyperlink{ref-Browning1992}{1992}; Hurwitz and Peffley \protect\hyperlink{ref-Hurwitz2005}{2005}; Henderson et al.
-\protect\hyperlink{ref-Henderson1997}{1997}; Wu, Sun, and Triplett \protect\hyperlink{ref-Wu2009}{2009}).
-Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
-There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
-
-For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods (Wilson \protect\hyperlink{ref-Wilson1990}{1990}).
-More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
-Much of this work, however, has focused on either the impacts of living in marginalized communities (by measuring health and economic impacts) or has not accounted for the importance of physical space (by focusing only on proximal social, and not geographic, contact with directly impacted individuals).
-This analysis allows us to understand the implications for living in a neighborhood where individuals likely to cast a ballot are not allowed to because of a felony conviction.
-I find that neighborhoods that are home to lost voters --- and particularly neighborhoods with large Black populations --- systematically turn out for local elections at lower rates than otherwise similar neighborhoods.
-
-This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
-In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
-As Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) shows, racial turnout differentials can have real consequences for city politics.
-
-Moreover, we cannot conclude that depressed turnout in these neighborhoods in 2017 had no impact on their representation.
-City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
-Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals \emph{held} and \emph{used} their power.
-
-Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
-As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
-
-\pagebreak
-
-\hypertarget{executive-order-181-and-turnout-in-2018}{%
-\section{Executive Order 181 and Turnout in 2018}\label{executive-order-181-and-turnout-in-2018}}
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
-To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
-Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018.
-This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
-senatorial contests (as well as other down-ballot races).
-
-I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
-
-\hypertarget{identifying-parolees-whose-rights-were-restored}{%
-\subsection*{Identifying Parolees Whose Rights Were Restored}\label{identifying-parolees-whose-rights-were-restored}}
-\addcontentsline{toc}{subsection}{Identifying Parolees Whose Rights Were Restored}
-
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.}
-
-\hypertarget{trends-in-turnout}{%
-\subsection*{Trends in Turnout}\label{trends-in-turnout}}
-\addcontentsline{toc}{subsection}{Trends in Turnout}
+Such standardizations, however, will miss individuals who change their names entirely.
+For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
+
+\hypertarget{results}{%
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-16-chart-1}
+{\centering \includegraphics{part2_standalone_files/figure-latex/to-16-chart-1}
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-chart-1}
+{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-chart-1}
 
@@
 However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated forindividuals discharged in 2017 or 2018 before the executive order was enacted.
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-recent-chart-1}
+{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-recent-chart-1}
 
@@
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.}
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/restoration-chart-chunk-1}
+{\centering \includegraphics{part2_standalone_files/figure-latex/restoration-chart-chunk-1}
 
@@
 Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 31.2 and 31.8 percent.
+The turnout rate for the control group was 2.0 percent.
 
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
 To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
 
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
+Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Noncitizens and individuals who violated their terms of parole were not eligible for rights restoration, and therefore did not ``comply'' with the treatment.
+Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
 
@@
 This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Although the available data does not indicate whether individuals were eligible to have their rights restored, we can examine the demographic differences between individuals who had their rights restored and those who did not.
+Table \ref{tab:demos-restorees} presents demographic differences between individuals who had their rights restored and those who did not among the intend-to-treat group (we cannot tell who would and would not have complied among the control group).
 
@@
 
-\hypertarget{discussion-1}{%
-\subsection*{Discussion}\label{discussion-1}}
+\hypertarget{discussion}{%
+\subsection*{Discussion}\label{discussion}}
 \addcontentsline{toc}{subsection}{Discussion}
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-Felony disenfranchisement was originally implemented to limit the franchise of Black Americans.
-It continues to serve that purpose today, as Black residents of New York and states around the country are disproportionately convicted of felony offenses and imprisoned.
-Although we know that many of these individuals were unlikely to cast a ballot even if they had not gone to prison, the targeted disenfranchisement of a single segment of the population is deeply troubling.
-
-As this project has demonstrated, however, the racialized penalties of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
-The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these policies.
-By both redefining ``lost voters'' as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
-What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, and this effect is concentrated in Black neighborhoods.
-Moreover, there is reason to believe that this is not driven by unobservable differences between neighborhoods; I also show that turnout decreased more between 2016 and 2017 in neighborhoods that newly lost voters than in neighborhoods unaffected by disenfranchisement policies.
-
-I also offer an early investigation of the effect of Executive Order 181, which granted voting rights to most citizens on parole in New York State.
-This research indicates that individuals who have their rights restored prior to being discharged from parole are more likely to participate in elections, even after they are discharged.
-This is a welcome finding, as increasing the political participation of marginalized individuals brings us toward a fuller democracy.
-As with neighborhood spillover effects, however, this effect depends in great deal on the race of the parolee: rights restoration, at least in the first five months of the program in New York State, did not increase future participation among non-Whites.
-More research needs to be done to investigate whether this is true over a longer time period in New York, and whether it is true in other states as well.
-
-America is becoming ever-more-diverse, and will likely be less than 50 percent non-Hispanic White by the middle of the century.
-As Fraga (\protect\hyperlink{ref-Fraga2018}{2018}) has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
-How districts are drawn, population patterns, and rules about who can fundraise how will continue to shape the politicians we elect and the policies they enact.
-Breaking down \emph{de jure} (like formal disenfranchisement) and \emph{de facto} (like neighborhood spillover effects) barriers to the ballot box is of great importance.
-The experience in New York State over the past few years offers a better understanding of just how big this problem is, but it also offers evidence of a way forward.
-
-\newpage
-
 \hypertarget{appendix-a}{%
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abadie2019}{}%
-Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \url{https://scholar.harvard.edu/spiess/publications/robust-post-matching-inference.}
-
 \leavevmode\hypertarget{ref-Alexander2012}{}%
@@
 
-\leavevmode\hypertarget{ref-Anzia2019}{}%
-Anzia, Sarah F.
-2019.
-``When Does a Group of Citizens Influence Policy?
-Evidence from Senior Citizen Participation in City Politics.'' \emph{The Journal of Politics} 81 (1): 1--14.
-\url{https://doi.org/10.1086/699916}.
-
-\leavevmode\hypertarget{ref-Behrens2003}{}%
-Behrens, Angela, Christopher Uggen, and Jeff Manza.
-2003.
-``Ballot Manipulation and the ``Menace of Negro Domination'': Racial Threat and Felon Disenfranchisement in the United States, 18502002.'' \emph{American Journal of Sociology} 109 (3): 559--605.
-\url{https://doi.org/10.1086/378647}.
-
-\leavevmode\hypertarget{ref-Bowers2009}{}%
-Bowers, Melanie, and Robert R.
-Preuhs.
-2009.
-``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons.'' \emph{Social Science Quarterly} 90 (3): 722--43.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
-
 \leavevmode\hypertarget{ref-bcj_laws}{}%
 Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
-
-\leavevmode\hypertarget{ref-Browning1992}{}%
-Browning, Sandra Lee, and Liqun Cao. 1992. ``The Impact of Race on Criminal Justice Ideology.'' \emph{Justice Quarterly} 9 (4): 685--701. \url{https://doi.org/10.1080/07418829200091611}.
 
@@
 \url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
-
-\leavevmode\hypertarget{ref-Burch2013}{}%
----------.
-2013.
-``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 184--201.
-\url{https://doi.org/10.1177/0002716213503093}.
 
@@
 
-\leavevmode\hypertarget{ref-Cho2006}{}%
-Cho, Wendy K.
-Tam, James G.
-Gimpel, and Joshua J.
-Dyck.
-2006.
-``Residential Concentration, Political Socialization, and Voter Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
-\url{https://doi.org/10.1111/j.1468-2508.2006.00377.x}.
-
-\leavevmode\hypertarget{ref-Clear2008}{}%
-Clear, Todd R. 2008. ``The Effects of High Imprisonment Rates on Communities.'' \emph{Crime and Justice} 37 (1): 97--132. \url{https://doi.org/10.1086/522360}.
-
-\leavevmode\hypertarget{ref-Dawkins2006}{}%
-Dawkins, Casey J. 2006. ``Are Social Networks the Ties That Bind Families to Neighborhoods?'' \emph{Housing Studies} 21 (6): 867--81. \url{https://doi.org/10.1080/02673030600917776}.
-
 \leavevmode\hypertarget{ref-Dong2014}{}%
@@
 
-\leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
-
 \leavevmode\hypertarget{ref-Fraga2018}{}%
 Fraga, Bernard L. 2018. \emph{The Turnout Gap}. Cambridge, United Kingdom: Cambridge University Press. \url{https://doi.org/10.1017/9781108566483}.
-
-\leavevmode\hypertarget{ref-Geller2014}{}%
-Geller, Amanda, Jeffrey Fagan, Tom Tyler, and Bruce G.
-Link.
-2014.
-``Aggressive Policing and the Mental Health of Young Urban Men.'' \emph{American Journal of Public Health} 104 (12): 2321--7.
-\url{https://doi.org/10.2105/ajph.2014.302046}.
 
@@
 \url{https://doi.org/10.2307/2585837}.
-
-\leavevmode\hypertarget{ref-Gerber2003}{}%
-Gerber, Alan S., Donald P.
-Green, and Ron Shachar.
-2003.
-``Voting May Be Habit-Forming: Evidence from a Randomized Field Experiment.'' \emph{American Journal of Political Science} 47 (3): 540--50.
-\url{https://doi.org/10.1111/1540-5907.00038}.
 
@@
 
-\leavevmode\hypertarget{ref-Gimpel2004}{}%
-Gimpel, James G., Joshua J.
-Dyck, and Daron R.
-Shaw.
-2004.
-``Registrants, Voters, and Turnout Variability Across Neighborhoods.'' \emph{Political Behavior} 26 (4): 343--75.
-\url{https://doi.org/10.1007/s11109-004-0900-4}.
-
 \leavevmode\hypertarget{ref-Green2000}{}%
@@
 \url{https://doi.org/10.1017/s0007123400000247}.
-
-\leavevmode\hypertarget{ref-Griffin2005}{}%
-Griffin, John D., and Brian Newman. 2005. ``Are Voters Better Represented?'' \emph{The Journal of Politics} 67 (4): 1206--27. \url{https://doi.org/10.1111/j.1468-2508.2005.00357.x}.
-
-\leavevmode\hypertarget{ref-Guest1999}{}%
-Guest, Avery M., and Susan K. Wierzbicki. 1999. ``Social Ties at the Neighborhood Level.'' \emph{Urban Affairs Review} 35 (1): 92--111. \url{https://doi.org/10.1177/10780879922184301}.
-
-\leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
-
-\leavevmode\hypertarget{ref-Hajnal2005}{}%
-Hajnal, Zoltan, and Jessica Trounstine.
-2005.
-``Where Turnout Matters: The Consequences of Uneven Turnout in City Politics.'' \emph{The Journal of Politics} 67 (2): 515--35.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00327.x}.
 
@@
 
-\leavevmode\hypertarget{ref-Henderson1997}{}%
-Henderson, Martha L., Francis T.
-Cullen, Liqun Cao, Sandra Lee Browning, and Renee Kopache.
-1997.
-``The Impact of Race on Perceptions of Criminal Injustice.'' \emph{Journal of Criminal Justice} 25 (6): 447--62.
-\url{https://doi.org/10.1016/s0047-2352(97)00032-9}.
-
-\leavevmode\hypertarget{ref-Hjalmarsson2010}{}%
-Hjalmarsson, Randi, and Mark Lopez.
-2010.
-``The Voting Behavior of Young Disenfranchised Felons: Would They Vote If They Could?'' \emph{American Law and Economics Review} 12 (2): 356--93.
-\url{https://doi.org/10.1093/aler/ahq004}.
-
-\leavevmode\hypertarget{ref-Huckfeldt1979}{}%
-Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
-
-\leavevmode\hypertarget{ref-Hurwitz2005}{}%
-Hurwitz, Jon, and Mark Peffley.
-2005.
-``Explaining the Great Racial Divide: Perceptions of Fairness in the U.s.
-Criminal Justice System.'' \emph{The Journal of Politics} 67 (3): 762--83.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00338.x}.
-
-\leavevmode\hypertarget{ref-Kenny1992}{}%
-Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259. \url{https://doi.org/10.2307/2111432}.
-
 \leavevmode\hypertarget{ref-Keyssar2009}{}%
@@
 
-\leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821. \url{https://doi.org/10.1177/0021934716659195}.
-
-\leavevmode\hypertarget{ref-Kinsella2015}{}%
-Kinsella, Chad, Colleen McTague, and Kevin N.
-Raleigh.
-2015.
-``Unmasking Geographic Polarization and Clustering: A Micro-Scalar Analysis of Partisan Voting Behavior.'' \emph{Applied Geography} 62 (August): 404--19.
-\url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
-
 \leavevmode\hypertarget{ref-Lassen2004}{}%
@@
 
-\leavevmode\hypertarget{ref-Lerman2013}{}%
-Lerman, Amy E., and Vesla Weaver.
-2013.
-``Staying Out of Sight?
-Concentrated Policing and Local Political Action.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 202--19.
-\url{https://doi.org/10.1177/0002716213503085}.
-
 \leavevmode\hypertarget{ref-locked_out}{}%
 Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
-
-\leavevmode\hypertarget{ref-Martin2003}{}%
-Martin, Paul S.
-2003.
-``Votings Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
-\url{https://doi.org/10.1111/1540-5907.00008}.
-
-\leavevmode\hypertarget{ref-Martin2013}{}%
-Martin, Paul S., and Michele P.
-Claibourn.
-2013.
-``Citizen Participation and Congressional Responsiveness: New Evidence That Participation Matters.'' \emph{Legislative Studies Quarterly} 38 (1): 59--81.
-\url{https://doi.org/10.1111/lsq.12003}.
 
@@
 
-\leavevmode\hypertarget{ref-Miles2004}{}%
-Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.'' \emph{The Journal of Legal Studies} 33 (1): 85--129. \url{https://doi.org/10.1086/381290}.
-
 \leavevmode\hypertarget{ref-Milligan2004}{}%
@@
 
-\leavevmode\hypertarget{ref-Morris2019}{}%
-Morris, Kevin. 2019. ``Thwarting Amendment 4.'' May 9, 2019. \url{https://www.brennancenter.org/analysis/thwarting-amendment-4}.
-
-\leavevmode\hypertarget{ref-Mutz2002}{}%
-Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838. \url{https://doi.org/10.2307/3088437}.
-
-\leavevmode\hypertarget{ref-Ochs2006}{}%
-Ochs, Holona Leanne.
-2006.
-``'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
-\url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
-
 \leavevmode\hypertarget{ref-Pierson1993}{}%
 Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
-
-\leavevmode\hypertarget{ref-Sekhon2011}{}%
-Sekhon, Jasjeet S.
-2011.
-``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (7).
-\url{https://doi.org/10.18637/jss.v042.i07}.
-
-\leavevmode\hypertarget{ref-Sewell2016}{}%
-Sewell, Abigail A., and Kevin A.
-Jefferson.
-2016.
-``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (S1): 42--67.
-\url{https://doi.org/10.1007/s11524-015-0016-7}.
-
-\leavevmode\hypertarget{ref-florida_1868}{}%
-Shofner, Jerrell H. 1963. ``The Constitution of 1868.'' \emph{The Florida Historical Quarterly} 41 (4): 356--74. \url{http://www.jstor.org/stable/30139965}.
 
@@
 \url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
-
-\leavevmode\hypertarget{ref-Uggen2002}{}%
-Uggen, Christopher, and Jeff Manza.
-2002.
-``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
-\url{https://doi.org/10.2307/3088970}.
-
-\leavevmode\hypertarget{ref-Uggen2004}{}%
----------.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
-
-\leavevmode\hypertarget{ref-Walker2014}{}%
-Walker, Hannah L.
-2014.
-``Extending the Effects of the Carceral State.
-Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly} 67 (4): 809--22.
-\url{https://doi.org/10.1177/1065912914542522}.
-
-\leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela Garcı́a-Castañon.
-2017.
-``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (04): 541--68.
-\url{https://doi.org/10.1017/s1743923x17000198}.
-
-\leavevmode\hypertarget{ref-Weaver2014}{}%
-Weaver, Vesla, and Amy E. Lerman. 2014. \emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}. Chicago: University of Chicago Press.
 
@@
 
-\leavevmode\hypertarget{ref-Wilson1990}{}%
-Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
-
 \leavevmode\hypertarget{ref-Wooldridge2010}{}%
@@
 
-\leavevmode\hypertarget{ref-Wu2009}{}%
-Wu, Yuning, Ivan Y.
-Sun, and Ruth A.
-Triplett.
-2009.
-``Race, Class or Neighborhood Context: Which Matters More in Measuring Satisfaction with Police?'' \emph{Justice Quarterly} 26 (1): 125--56.
-\url{https://doi.org/10.1080/07418820802119950}.
-
 \end{document}

### 2019-11-12 markdown/part2_standalone.Rmd
write

@@
 ---
-title: "Voting Rights Restoration and Post-Supervision Turnout"
-author: "Kevin Morris"
+title: "Welcome Home --- Now Vote!"
+subtitle: "Voting Rights Restoration and Post-Supervision Participation"
+author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
+institute: "Brennan Center for Justice"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 header-includes:
-    - \usepackage{setspace}\doublespacing
     - \usepackage{rotating}
-
+    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+abstract: |
+  This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
+In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
+By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to parolees increases their post-supervision propensity to cast a ballot.
+I demonstrate that individuals whose rights were restored while still on parole were more likely to participate than those whose rights were automatically restored upon completion of parole.
+This group-level effect, however, masks race-specific effects.
+Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
+This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure black Americans’ view of the democratic process.
 ---
@@
 \pagenumbering{gobble}
-\tableofcontents
 \pagebreak
@@
 
-The political history of the United States has been characterized by a general, if unlinear, trend toward universal sufferage [see, for instance, @Keyssar2009].
-At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
-And yet, despite the United State's purported march toward ever-more-inclusive systems of democracy, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised today [@sentencing_2016].
-These disenfranchising policies provide a stark reminder that our democracy is not as inclusive as we might like to tell ourselves.
-
 In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
@@
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
+This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as *Richardson v.
 Ramirez* (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as `r 6.1 - 1.4` million American citizens.
+Collectively, these laws disenfranchise as many as 4.7 million American citizens.
 Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
@@
 The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle Alexander [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
+As Michelle [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly black --- Americans.
+This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is black.
 This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+As @Gelman2007 shows, for instance, New York's ``stop-and-frisk'' policy targeted black and Latino New Yorkers at rates far higher than white residents, even after controlling for neighborhood variability and race-specific arrest rates.
+
+There is some evidence, however, that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019; but see @Gerber2017].
+As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
+The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
+To address only the formal laws contributing to disenfranchisement without also interrogating efforts to boost post-supervision participation risks leaving much of the system of effective disenfranchisement undisturbed.
+New York State offers us the opportunity to test how the timing of the re-instation of voting rights structures post-supervision participation.
 
@@
 On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
+Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
 The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
@@
 
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019].
-As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters [see, for instance, @Fraga2018].
-To address *only* the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
+There is reason to believe that the executive order may also have increased the political participation of formerly disenfranchised individuals.
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
@@
 These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
+This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
 However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
@@
 
+In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
+George W.
+Bush's margin of victory in Florida in 2000 was famously just 537 votes.
+In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
+They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
+Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
+They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
+Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
+In the years after Uggen and Manza published their paper, scholars sought to investigate the number of individuals who would have voted if they were not disenfranchised, often using survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
 @Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch [-@Burch2010; -@Burch2011] expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
+Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states [-@Burch2010; -@Burch2011].
 She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
+Using the registered voter files, she estimated the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
 Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [e.g.
@@
 
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so [@Haselswerdt2009; @Burch2011; @Meredith2015].
+Much of this literature has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
 The causal effect of incarceration on participation is the subject of some debate within the field.
@@
 The criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of committment; and others.
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of commitment; and others.
 This analysis is limited to individuals incarcerated for felony offenses.
@@
 However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated forindividuals discharged in 2017 or 2018 before the executive order was enacted.
+Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated for individuals discharged in 2017 or 2018 before the executive order was enacted.
 
@@
 \input{"../temp/table3.tex"}
+
 The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
@@
 \input{"../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights ("compliance") is strongly correlated with certain demographics that are also linked to propensity to vote.
+Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
 Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
@@
 
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the "local average treatment effect." The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
 Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research [e.g.
-@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
+Some research [@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
 The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2014].
@@
 \input{"../temp/iv_clean.tex"}
-
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
@@
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
+Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
 
@@
 In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
+These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only black former parolees (Models 3 and 6).
 Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
@@
 
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration *nearly doubled* these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.6 percentage points.
+Given that just 3.3 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
+We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
 
-Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
+Why would the intervention have increased turnout among white individuals and have had no effect on black individuals?
 Some of this may be explained by different propensities to vote.
-@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
+@White2019, for instance, shows that brief periods of incarceration decreases black individuals' propensity to vote by substantially more than white individuals.
+She demonstrates that, prior to incarceration, black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
+The inverse may hold true here: if black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether black individuals turned out at higher rates.
+Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of black individuals to vote.
 
@@
 \input{"../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were black.
+These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
+Even when Latinos are excluded from this group, black individuals were still 36 percent more likely to cast a ballot.
 In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for black individuals, but that does not mean they are less likely to vote than other former parolees.
 
@@
 Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
-It is worth noting that the *R^2^* in Model 1 is higher than the *R^2^* in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
 
@@
 
+\newpage
+\beginsupplement
 # Appendix A {-}

### 2019-11-12 markdown/part2_standalone.tex
write

@@
 \hypersetup{unicode=true,
-            pdftitle={Voting Rights Restoration and Post-Supervision Turnout},
+            pdftitle={Welcome Home --- Now Vote!},
             pdfauthor={Kevin Morris},
@@
 
-  \title{Voting Rights Restoration and Post-Supervision Turnout}
+  \title{Welcome Home --- Now Vote!}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-    \author{Kevin Morris}
+  \subtitle{Voting Rights Restoration and Post-Supervision Participation}
+  \author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
     \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-    \date{November 06, 2019}
+    \date{November 12, 2019}
 
@@
 
-\usepackage{setspace}\doublespacing
 \usepackage{rotating}
+\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
 
@@
 \maketitle
+\begin{abstract}
+This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
+In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
+By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to parolees increases their post-supervision propensity to cast a ballot.
+I demonstrate that individuals whose rights were restored while still on parole were more likely to participate than those whose rights were automatically restored upon completion of parole.
+This group-level effect, however, masks race-specific effects.
+Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
+This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure black Americans' view of the democratic process.
+\end{abstract}
 
 \pagenumbering{gobble}
-\tableofcontents
 \pagebreak
@@
 
-The political history of the United States has been characterized by a general, if unlinear, trend toward universal sufferage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
-At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship in a liberal democracy.
-And yet, despite the United State's purported march toward ever-more-inclusive systems of democracy, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised today (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-These disenfranchising policies provide a stark reminder that our democracy is not as inclusive as we might like to tell ourselves.
-
 In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
@@
 In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
+This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as \emph{Richardson v.
@@
 The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly Black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were Black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is Black.
+As Michelle (\protect\hyperlink{ref-Alexander2012}{2012}) and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly black --- Americans.
+This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is black.
 This disparity is likely due \emph{not} to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy targeted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
+As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy targeted black and Latino New Yorkers at rates far higher than white residents, even after controlling for neighborhood variability and race-specific arrest rates.
+
+There is some evidence, however, that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}; but see Gerber et al.
+\protect\hyperlink{ref-Gerber2017}{2017}).
+As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
+The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
+To address only the formal laws contributing to disenfranchisement without also interrogating efforts to boost post-supervision participation risks leaving much of the system of effective disenfranchisement undisturbed.
+New York State offers us the opportunity to test how the timing of the re-instation of voting rights structures post-supervision participation.
 
@@
 On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
+Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
 The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
@@
 
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
-To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
+There is reason to believe that the executive order may also have increased the political participation of formerly disenfranchised individuals.
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
@@
 
+In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
+George W.
+Bush's margin of victory in Florida in 2000 was famously just 537 votes.
+In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
+They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
+Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
+They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
+Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
+In the years after Uggen and Manza published their paper, scholars sought to investigate the number of individuals who would have voted if they were not disenfranchised, often using survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
 Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}) expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
+Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}).
 She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
+Using the registered voter files, she estimated the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
 Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
 
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so (Haselswerdt \protect\hyperlink{ref-Haselswerdt2009}{2009}; Burch \protect\hyperlink{ref-Burch2011}{2011}; Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}).
+Much of this literature has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
 The causal effect of incarceration on participation is the subject of some debate within the field.
@@
 
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andew Doe'' and ``John Anthony Doe'' would not.
+Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
 
-There are two types of potential error in this methdology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
+There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
 False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
@@
 \centering
-\begin{tabular}{cc}
+\begin{tabular}[t]{cc}
 \toprule
@@
 Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the \{{[}\}0, 1\{{]}\} interval.
 Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research (e.g.~Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
+Some research (Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
 The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
@@
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between 32.7 and 39.5 percent.
+Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 32.7 and 39.5 percent.
 
@@
 
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.6 percentage points.
+Given that just 3.3 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
+We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
 
-Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
+Why would the intervention have increased turnout among white individuals and have had no effect on black individuals?
 Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
+White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases black individuals' propensity to vote by substantially more than white individuals.
+She demonstrates that, prior to incarceration, black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
+The inverse may hold true here: if black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether black individuals turned out at higher rates.
+Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of black individuals to vote.
 
@@
 
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
+Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were black.
+These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
+Even when Latinos are excluded from this group, black individuals were still 36 percent more likely to cast a ballot.
 In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for black individuals, but that does not mean they are less likely to vote than other former parolees.
 
@@
 Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
-It is worth noting that the \emph{R\textsuperscript{2}} in Model 1 is higher than the \emph{R\textsuperscript{2}} in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
 
@@
 By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
+
+\newpage
+\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}
 
@@
 
-\leavevmode\hypertarget{ref-Fraga2018}{}%
-Fraga, Bernard L. 2018. \emph{The Turnout Gap}. Cambridge, United Kingdom: Cambridge University Press. \url{https://doi.org/10.1017/9781108566483}.
-
 \leavevmode\hypertarget{ref-Gelman2007}{}%
@@
 \url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
-
-\leavevmode\hypertarget{ref-Keyssar2009}{}%
-Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. New York: Basic Books.
 
@@
 
+\leavevmode\hypertarget{ref-Uggen2002}{}%
+Uggen, Christopher, and Jeff Manza.
+2002.
+``Democratic Contraction?
+Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
+\url{https://doi.org/10.2307/3088970}.
+
+\leavevmode\hypertarget{ref-Uggen2004}{}%
+---------.
+2004.
+``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
+New York: Russell Sage Foundation.
+
 \leavevmode\hypertarget{ref-Weaver2010}{}%

### 2019-11-12 markdown/part1/p1.Rmd
change folder structure

--- 2019-09-25 markdown/p1.Rmd
@@
 ---
-title: "Neighborhoods and Felony Disenfranchisement: The Case of New York CityXX"
+title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
 author: "Kevin Morris"
@@
 To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
-
-I begin by offering a refinement of our understanding of who lost voters are.
-Previous studies exploring the relationship between felony disenfranchisement and spill-over effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a demonstrated history of participation are disenfranchised.
-By matching "treated" neighborhoods (neighborhoods with lost voters) to "untreated" neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
 
@@
 
+### Toward a Causal Estimation {-}
+
+As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
+The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
+If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
+As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
+
+I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
+This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
+These individuals' home neighborhoods are removed from the subsequent analysis.
+I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
+These neighborhoods, therefore, were *newly* home to lost voters.
+If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
+In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
+The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
+
+Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
+It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
+Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
+
+Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
+As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+
+```{r block-table-chunk-did, echo = F}
+t <- readRDS("./temp/match_table_block_group_did.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"),
+                font_size = 10)
+j
+```
+
+```{r tract-chunk-did, echo = F}
+t <- readRDS("./temp/match_table_tract_did.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match-tract) Results of Tract-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"),
+                font_size = 10)
+j
+```
+
+In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups and tracts.
+In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
+D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
+In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly *higher* turnout prior to intervention.
+D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
+In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
+Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
+The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
+
+```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
+load("./temp/match_reg_block_group_did.rdata")
+
+bg_reg <- reg_output
+bg_reg_ses <- reg_output_ses
+
+load("./temp/match_reg_tract_did.rdata")
+
+tract_reg <- reg_output
+tract_reg_ses <- reg_output_ses
+
+stargazer(bg_reg, tract_reg,
+          header = F,
+          column.labels = c("Block Group Level", "Tract Level"),
+          se = list(bg_reg_ses, tract_reg_ses),
+          notes = c("TO REPLACE"),
+          type = "latex", notes.align = "l",
+          notes.append = T,
+          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
+          omit = "district",
+          dep.var.labels = c("Turnout Rate"),
+          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test_abc.tex",
+          out.header = F,
+          add.lines=list(c("Congressional District FE", "X", "X")))
+
+j <- fread("./temp/test_abc.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert <- "\\cmidrule(l{3pt}r{3pt}){2-2} \\cmidrule(l{3pt}r{3pt}){3-3}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../temp/table1111.tex"}
+
+Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
+These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
+Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
+
+Of particular note is the lower share of the population in these neighborhoods that is Black.
+The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
+The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
+Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
+
 ## Discussion {-}
@@
 Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and nonwhite Americans.
+Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchise all combine to undermine the political power of marginalized communities in New York City.
+The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.

### 2019-11-12 markdown/part1/p1.Rmd
write

@@
 title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
-author: "Kevin Morris"
+author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
+institute: "Brennan Center for Justice"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
     latex_engine: xelatex
-    # includes: 
-    #   in_header: "../code/misc/preamble.tex"
-    keep_tex: false
-bibliography: "citations.bib"
+    keep_tex: true
+bibliography: "../citations.bib"
 link-citations: yes
-
+fontsize: 12pt
 header-includes:
-    - \usepackage{setspace}\doublespacing
     - \usepackage{rotating}
-
+    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+abstract: |
+  Over the past two decades, scholars have sought to estimate the direct and indirect effects of felony disenfranchisement on political representation.
+This literature, however, has often overlooked both the geographic concentration of communities impacted by over-incarceration and the low propensity to vote exhibited by individuals convicted of felony crimes.
+In this paper, I redefine “lost voters” as disenfranchised individuals with a history of participating in elections.
+I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods.
+I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
 ---
@@
 \pagenumbering{gobble}
-\tableofcontents
 \pagebreak
 ```{r setup, echo = F, message = F, warning = F}
-knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf")
+knitr::opts_knit$set(root.dir = "../..", dev = "cairo_pdf")
 library(systemfit)
@@
 \pagenumbering{arabic}
-# Introduction
+# Introduction {-}
 
@@
 Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008].
-Concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008], while concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
+Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
 
@@
 Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
+If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
 If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
 It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level has material consequences.
+This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
 As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
 
-## Background
+# Background {-}
 
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout [@Miles2004; @Hjalmarsson2010].
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects [@Miles2004; @Ochs2006].
-Others have used survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
+Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
+@Uggen2002 produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for for disenfranchisment in Florida.
+Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
+Since @Uggen2002, other scholars have also investigated the direct effect of felony disenfranchisement [@Miles2004;@Uggen2004; @Drucker2005; @Ochs2006].
 
-A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
 @King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
@@
 
-## Framing for Turnout Effects {-}
+# Framing for Turnout Effects {-}
 
@@
 
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spill-over effects of lost voters.
+Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the indirect neighborhood effects of lost voters.
 @Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact" [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
@@
 
-### Theoretical Implications for Reduced Turnout {-}
+Similarly, @White2019a finds that incarceration has only moderate indirect effects on voting.
+She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (607).
+This study does not, however, interrogate whether the incarceration of a would-be voter has different indirect effects than the incarceration of an individual who would not have voted either way.
+It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect effects arising from other individuals.
+Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
+
+# Theoretical Implications for Reduced Turnout {-}
 
@@
 
-# Data
+# Data {-}
+
 ## Criminal Justice Data {-}
@@
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and when they finished their parole supervision.
+These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and other demographic information.
 
@@
 
-The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
-
 ## Voter File Data {-}
@@
 In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The first section of this paper uses a snapshot of the registered voter file from April 30^th^, 2018.
-The second section of this paper uses a snapshot from March 3^rd^, 2019.
+It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
+The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
+I use a snapshot of the registered voter file from April 30^th^, 2018.
 
@@
 A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
-This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
-Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+Such standardizations, however, will miss individuals who change their names entirely.
+For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
 
-# Effects of Felony Disenfranchisement on Neighborhood Turnout Levels
+# Identification of Lost Voters {-}
 
-## Identification of Lost Voters {-}
+Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
+By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
+Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
+
 In this analysis, I offer a different definition of "lost voter" than much of the literature.
@@
 
-## Testing for Neighborhood Turnout Effects {-}
+# Testing for Neighborhood Turnout Effects {-}
 ```{r table-table, , echo = F, message = F, warning = F}
@@
 
-### Match Output {-}
+## Match Output {-}
 ```{r block-table-chunk, echo = F}
@@
 At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less white, have much lower median incomes, and a larger share of voters are registered as Democrats.
+
+In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
+On average, the non-Hispanic Black share of a block group with a lost voter is more than *double* the share in block groups without lost voters.
+These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
+Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] The median income in neighborhoods without lost voters is more than 40 percent higher.
+Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
 
@@
 insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
 
-j <- bind_rows(j, data.frame(V1 = c(insert), n = 7.5)) %>% 
+j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 ```
-\input{"../temp/table1.tex"}
+\input{"../../temp/table1.tex"}
 
@@
 
-### Testing Intensity Effects {-}
+## Testing Intensity Effects {-}
 
@@
 ```
-\input{"../temp/table2.tex"}
+\input{"../../temp/table2.tex"}
 
@@
 
-### Toward a Causal Estimation {-}
+## Toward a Causal Estimation {-}
 
@@
 ```
-\input{"../temp/table1111.tex"}
+\input{"../../temp/table1111.tex"}
 
@@
 
-## Discussion {-}
+# Discussion {-}
 
@@
 Understanding the magnitude of this finding is important.
-In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 90 percent of the population, with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.9 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
+Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.16 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots.
 
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood `r round(0.9 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
 
@@
 The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
+
+\newpage
+# References {-}
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\noindent

### 2019-11-12 markdown/part1/p1.tex
write

--- 2019-11-06 markdown/felony_disenfranchisement_nys.tex
@@
 \hypersetup{unicode=true,
-            pdftitle={In the Shadow of Jim Crow},
+            pdftitle={Neighborhoods and Felony Disenfranchisement: The Case of New York City},
             pdfauthor={Kevin Morris},
@@
 
-  \title{In the Shadow of Jim Crow}
+  \title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
     \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \subtitle{Felony Disenfranchisement and New York State}
-  \author{Kevin Morris}
+    \author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
     \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-    \date{November 05, 2019}
+    \date{November 12, 2019}
 
@@
 \maketitle
+\begin{abstract}
+Over the past two decades, scholars have sought to estimate the direct and indirect effects of felony disenfranchisement on political representation.
+This literature, however, has often overlooked both the geographic concentration of communities impacted by over-incarceration and the low propensity to vote exhibited by individuals convicted of felony crimes.
+In this paper, I redefine ``lost voters'' as disenfranchised individuals with a history of participating in elections.
+I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods.
+I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
+\end{abstract}
 
 \pagenumbering{gobble}
-
-\tableofcontents
 \pagebreak
@@
 \hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section*{Introduction}\label{introduction}}
+\addcontentsline{toc}{section}{Introduction}
 
@@
 Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
+In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
@@
 Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific arrest rates.
+In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
+As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
 
@@
 The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted (Clear \protect\hyperlink{ref-Clear2008}{2008}), while concentrated policing has also been credited with having a ``chilling effect'' on neighborhoods' willingness to reach out for help to local governments (Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
+Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
 
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (A.
-White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}; Gerber et al.
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}; Gerber et al.
 \protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
@@
 
-The rest of this paper is structured as follows: \textbf{Section 2} discusses the historical context of felony disenfranchisement in the United States, and \textbf{Section 3} provides an overview of the scholarship conducted in the field over the past two decades.
-\textbf{Section 4} provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
-\textbf{Section 5} presents the data sources used throughout the project.
-
-\textbf{Section 6} proposes a new definition of ``lost voters'' in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised \emph{and} have a history of participating in elections.
-This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods through both direct (the formal disenfranchisement of would-be voters) and indirect (through proximal contact with disenfranchised individuals) effects.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely \emph{understated} the degree to which they are impacted by spatially concentrated patterns of criminal justice.
-A discussion of the results follows.
-
-\textbf{Section 7} turns from the 2017 mayoral election to the 2018 midterm elections.
-In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
-I employ an instrumental variables approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
-I focus specifically on individuals who would have been eligible to cast a ballot even if the executive order had not gone into effect.
-A discussion of the results follows.
-
-\textbf{Section 8} synthesizes the findings presented in the rest of the paper and provides concluding remarks.
-
-\hypertarget{background-of-felony-disenfranchisement-in-the-united-states}{%
-\section{Background of Felony Disenfranchisement in the United States}\label{background-of-felony-disenfranchisement-in-the-united-states}}
-
-In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-This variation in laws has been defended with language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as \emph{Richardson v.
-Ramirez} (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
-
-Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Burch (\protect\hyperlink{ref-Burch2010}{2010}) explains, ``If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were \emph{intended} to'' (4; emphasis in the original).
-Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans (Behrens, Uggen, and Manza \protect\hyperlink{ref-Behrens2003}{2003}).
-In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being ``niggerized'' (Shofner \protect\hyperlink{ref-florida_1868}{1963}, 374).
-
-The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
-Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
-In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just 12.1 percent of the voting age population in 2016, they accounted for 36.5 percent of the disenfranchised population (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate (Morris \protect\hyperlink{ref-Morris2019}{2019}).
-
-\hypertarget{academic-literature-and-felony-disenfranchisement}{%
-\section{Academic Literature and Felony Disenfranchisement}\label{academic-literature-and-felony-disenfranchisement}}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout (Miles \protect\hyperlink{ref-Miles2004}{2004}; Hjalmarsson and Lopez \protect\hyperlink{ref-Hjalmarsson2010}{2010}).
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects (Miles \protect\hyperlink{ref-Miles2004}{2004}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-Others have used survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}) expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
-
-A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
+\hypertarget{background}{%
+\section*{Background}\label{background}}
+\addcontentsline{toc}{section}{Background}
+
+Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
+Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}) produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for for disenfranchisment in Florida.
+Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
+Since Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}), other scholars have also investigated the direct effect of felony disenfranchisement (Miles \protect\hyperlink{ref-Miles2004}{2004}; Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
+
+A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
@@
 Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
+Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
 This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
 
-\hypertarget{what-this-paper-examines}{%
-\section{What This Paper Examines}\label{what-this-paper-examines}}
-
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City mayoral election of 2017 using individual-level administrative data.
+This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
 Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods (e.g.~Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}; Clear \protect\hyperlink{ref-Clear2008}{2008}; Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
@@
 
-I begin by offering a refinement of our understanding of who lost voters are.
-Previous studies exploring the relationship between felony disenfranchisement and spillover effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a \emph{demonstrated history} of participation are disenfranchised.
-By matching ``treated'' neighborhoods (neighborhoods with lost voters) to ``untreated'' neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
-
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.\footnote{Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those currently in prison for felony offenses are disenfranchised.} Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I use administrative data to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
-
-The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather than think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
-This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
-The second part of this project seeks to do just that.
-
 \hypertarget{framing-for-turnout-effects}{%
-\subsection*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
-\addcontentsline{toc}{subsection}{Framing for Turnout Effects}
+\section*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
+\addcontentsline{toc}{section}{Framing for Turnout Effects}
 
@@
 Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended.
-This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
-
-Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
+As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
+
+Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
 Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
@@
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
-It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
+It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
 To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
@@
 
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
+Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the indirect neighborhood effects of lost voters.
 Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
@@
 
-Similarly, A.
-White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
+Similarly, White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
 She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (607).
-This study examines only the indirect effects of proximal contact with the criminal justice system.
-It is possible that the indirect effects of the incarceration or jailing of an individual who would have voted but not for their contact with the criminal justice system would have larger indirect effects.
+This study does not, however, interrogate whether the incarceration of a would-be voter has different indirect effects than the incarceration of an individual who would not have voted either way.
+It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect effects arising from other individuals.
 Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
@@
 \hypertarget{theoretical-implications-for-reduced-turnout}{%
-\subsubsection*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
-\addcontentsline{toc}{subsubsection}{Theoretical Implications for Reduced Turnout}
+\section*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
+\addcontentsline{toc}{section}{Theoretical Implications for Reduced Turnout}
 
@@
 
-\hypertarget{turnout-among-formerly-disenfranchised-individuals}{%
-\subsection*{Turnout Among Formerly Disenfranchised Individuals}\label{turnout-among-formerly-disenfranchised-individuals}}
-\addcontentsline{toc}{subsection}{Turnout Among Formerly Disenfranchised Individuals}
-
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so (Haselswerdt \protect\hyperlink{ref-Haselswerdt2009}{2009}; Burch \protect\hyperlink{ref-Burch2011}{2011}; Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}).
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}) uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that ``it appears that spending time in prison does not have large negative effects on subsequent participation'' (1144).
-
-A.
-White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
-By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration \emph{causes} low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
-
-The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
-I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
-
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2006}).
-Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
-Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-``Whatever the participatory consequences of incarceration,'' they conclude, ``they are not in large part impossible to overcome'' (924).
-Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
-
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that ``contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual'' (829).
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-Such repairs may lead to increased propensity to participate in elections in the future.
-
 \hypertarget{data}{%
-\section{Data}\label{data}}
+\section*{Data}\label{data}}
+\addcontentsline{toc}{section}{Data}
 
@@
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole and demographic characteristics.
+These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and other demographic information.
 
 The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors available.
+It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
 Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
 It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
-Individuals who were (A) citizens, (B) not in violation of any terms of their parole, and (C) had not been rearrested were eligible for rights restoration.
 
@@
 
-Most states in the United States are required to maintain files with information on all registered and recently purged voters.
+Most states in the United States are required to maintain files with information on all registered voters.
 In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The first section of this paper uses a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
-The second section of this paper uses a snapshot from March 3\textsuperscript{rd}, 2019.
+It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
+The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
+I use a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
 
 The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered (and participated) in the past but have since been purged due to a felony conviction.
+I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
 
@@
 
-Voters' home addresses were converted to latitudes and longitudes using a geocoding services from address-verification company SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
 This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
@@
 \centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}{cc}
+\begin{tabular}[t]{cc}
 \toprule
@@
 A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
-This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
-Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-
-\hypertarget{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}{%
-\section{Effects of Felony Disenfranchisement on Neighborhood Turnout Levels}\label{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}}
+Such standardizations, however, will miss individuals who change their names entirely.
+For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
 
 \hypertarget{identification-of-lost-voters}{%
-\subsection*{Identification of Lost Voters}\label{identification-of-lost-voters}}
-\addcontentsline{toc}{subsection}{Identification of Lost Voters}
+\section*{Identification of Lost Voters}\label{identification-of-lost-voters}}
+\addcontentsline{toc}{section}{Identification of Lost Voters}
 
@@
 These higher indirect (depressive) effects likely arise through three mechanisms.
-* Firstly, lost voters are likely members of families and communities where voting is valued.
+
+\begin{itemize}
+\item
+  Firstly, lost voters are likely members of families and communities where voting is valued.
 Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
@@
 Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
-* Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
+\item
+  Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
 The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
-* Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
+\item
+  Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
 Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
+\end{itemize}
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/citywide-map-1}
+{\centering \includegraphics{p1_files/figure-latex/citywide-map-1}
 
@@
 \hypertarget{testing-for-neighborhood-turnout-effects}{%
-\subsection*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
-\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Effects}
+\section*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
+\addcontentsline{toc}{section}{Testing for Neighborhood Turnout Effects}
 
@@
 \fontsize{10}{12}\selectfont
-\begin{tabular}{ccc}
+\begin{tabular}[t]{ccc}
 \toprule
@@
 \hypertarget{match-output}{%
-\subsubsection*{Match Output}\label{match-output}}
-\addcontentsline{toc}{subsubsection}{Match Output}
+\subsection*{Match Output}\label{match-output}}
+\addcontentsline{toc}{subsection}{Match Output}
 
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{lllllrrrr}
+\begin{tabular}[t]{lllllrrrr}
 \toprule
@@
 
-\input{"../temp/table1.tex"}
+\input{"../../temp/table1.tex"}
 
@@
 \hypertarget{testing-intensity-effects}{%
-\subsubsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
-\addcontentsline{toc}{subsubsection}{Testing Intensity Effects}
+\subsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
+\addcontentsline{toc}{subsection}{Testing Intensity Effects}
 
@@
 
-\input{"../temp/table2.tex"}
+\input{"../../temp/table2.tex"}
 
@@
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/postest-map-1}
+{\centering \includegraphics{p1_files/figure-latex/postest-map-1}
 
@@
 \hypertarget{toward-a-causal-estimation}{%
-\subsubsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
-\addcontentsline{toc}{subsubsection}{Toward a Causal Estimation}
+\subsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
+\addcontentsline{toc}{subsection}{Toward a Causal Estimation}
 
@@
 \fontsize{10}{12}\selectfont
-\begin{tabular}{lllllrrrr}
+\begin{tabular}[t]{lllllrrrr}
 \toprule
@@
 
-\input{"../temp/table1111.tex"}
+\input{"../../temp/table1111.tex"}
 
@@
 \hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
+\section*{Discussion}\label{discussion}}
+\addcontentsline{toc}{section}{Discussion}
 
@@
 
-This finding mirrors A.
-White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
 This is perhaps unsurprising.
@@
 
-\pagebreak
-
-\hypertarget{executive-order-181-and-turnout-in-2018}{%
-\section{Executive Order 181 and Turnout in 2018}\label{executive-order-181-and-turnout-in-2018}}
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (A.
-White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
-To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
-Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018.
-This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
-senatorial contests (as well as other down-ballot races).
-
-I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
-
-\hypertarget{identifying-parolees-whose-rights-were-restored}{%
-\subsection*{Identifying Parolees Whose Rights Were Restored}\label{identifying-parolees-whose-rights-were-restored}}
-\addcontentsline{toc}{subsection}{Identifying Parolees Whose Rights Were Restored}
-
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.}
-
-\hypertarget{trends-in-turnout}{%
-\subsection*{Trends in Turnout}\label{trends-in-turnout}}
-\addcontentsline{toc}{subsection}{Trends in Turnout}
-
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.6 percent cast a ballot in the 2016 election, while just 4.2 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-16-chart-1}
-
-}
-
-\caption{\label{fig:to-16}Turnout in 2016 Presidential Election}\label{fig:to-16-chart}
-\end{figure}
-
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-chart-1}
-
-}
-
-\caption{\label{fig:to-18}Turnout in 2018 Midterm Election}\label{fig:to-18-chart}
-\end{figure}
-
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- \emph{were} more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
-
-\hypertarget{individual-level-turnout-regressions}{%
-\subsection*{Individual-Level Turnout Regressions}\label{individual-level-turnout-regressions}}
-\addcontentsline{toc}{subsection}{Individual-Level Turnout Regressions}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being ``assigned'' to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-recent-chart-1}
-
-}
-
-\caption{\label{fig:to-18-recent}Turnout in 2018 Midterm Election}\label{fig:to-18-recent-chart}
-\end{figure}
-
-Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
-
-\input{"../temp/table3.tex"}
-
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.}
-
-Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18\textsuperscript{th} that parolees consistently had their voting rights restored prior to discharge.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/restoration-chart-chunk-1}
-
-}
-
-\caption{\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge}\label{fig:restoration-chart-chunk}
-\end{figure}
-
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are assigned ``treatment'' by the policy change.
-Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is biased against finding a significant increase due to the executive order.
-
-\input{"../temp/table_whatever2.tex"}
-
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18\textsuperscript{th}, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1\textsuperscript{st}, 2017, through October 10\textsuperscript{th}, 2018 (the registration deadline in New York State).
-
-\input{"../temp/table4.tex"}
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 31.2 and 31.8 percent.
-
-\hypertarget{instrumental-variables-approach}{%
-\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
-\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
-
-Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
-
-An individual's propensity to vote can be expressed using the following equation:
-\[ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} \]
-Where \emph{Y\textsubscript{i}} is 1 if individual \emph{i} cast a ballot, \emph{X\textsubscript{1i}} is the probability that individual \emph{i} was eligible to have his rights restored prior to discharge from parole, and \emph{X\textsubscript{2i}} is 1 if individual \emph{i} actually had his rights restored prior to discharge from parole.
-\emph{Z} is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe \emph{X\textsubscript{1i}}, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if \emph{X\textsubscript{1i}} and \emph{X\textsubscript{2i}} (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if \emph{b\textsubscript{1}} is uncorrelated with an individual's propensity to vote (\emph{Y\textsubscript{i}}).
-
-Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
-Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Although the available data does not indicate whether individuals were eligible to have their rights restored, we can examine the demographic differences between individuals who had their rights restored and those who did not.
-
-\input{"../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that \emph{b\textsubscript{1}} is very probably correlated with \emph{Y}.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration \emph{caused} turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
-Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research (e.g.~Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
-The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2007}{2007}).
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
-\input{"../temp/iv_clean.tex"}
-
-The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between 32.7 and 39.5 percent.
-
-\hypertarget{racial-variation}{%
-\subsubsection*{Racial Variation}\label{racial-variation}}
-\addcontentsline{toc}{subsubsection}{Racial Variation}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
-
-\input{"../temp/iv_clean_race.tex"}
-
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
-
-Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
-Some of this may be explained by different propensities to vote.
-A.
-White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
-
-\input{"../temp/table5.tex"}
-
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
-
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
-The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
-
-\hypertarget{variable-treatment-intensity}{%
-\subsubsection*{Variable Treatment Intensity}\label{variable-treatment-intensity}}
-\addcontentsline{toc}{subsubsection}{Variable Treatment Intensity}
-
-There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
-This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
-If the individual did not have voting rights restored, this variable takes the value 0.
-This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.\footnote{Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.}
-
-\input{"../temp/iv_clean_months.tex"}
-
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
-It is worth noting that the \emph{R\textsuperscript{2}} in Model 1 is higher than the \emph{R\textsuperscript{2}} in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
-
-Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
-
-\hypertarget{discussion-1}{%
-\subsection*{Discussion}\label{discussion-1}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
-Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
-That changed in 2018.
-On October 12\textsuperscript{th} --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
-Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just 792, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
-
-As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least 32.7 percent.
-This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
-Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
-
-The mechanism through which the rules change increased turnout among these individuals is not clear.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
-Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
-
-It could also be a story of better information.
-As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
-Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
-In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
-
-Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
-Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
-The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
-This discrepancy could be caused by a number of different factors.
-Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
-The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
-
-We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
-Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
-Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-WHITE2014}{2014}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2016}{2016}).
-
-Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
-This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
-
-Re-enfranchising voters while they are still under formal supervision is obviously beneficial to the individuals who are on parole on election day; such policies allow them to make their voices heard.
-The case of Executive Order 181 also indicates that restoring voting rights prior to parole discharge has further benefits.
-In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
-This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
-By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-Felony disenfranchisement was originally implemented to limit the franchise of Black Americans.
-It continues to serve that purpose today, as Black residents of New York and states around the country are disproportionately convicted of felony offenses and imprisoned.
-Although we know that many of these individuals were unlikely to cast a ballot even if they had not gone to prison, the targeted disenfranchisement of a single segment of the population is deeply troubling.
-
-As this project has demonstrated, however, the racialized penalties of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
-The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these policies.
-By both redefining ``lost voters'' as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
-What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, and this effect is concentrated in Black neighborhoods.
-Moreover, there is reason to believe that this is not driven by unobservable differences between neighborhoods; I also show that turnout decreased more between 2016 and 2017 in neighborhoods that newly lost voters than in neighborhoods unaffected by disenfranchisement policies.
-
-I also offer an early investigation of the effect of Executive Order 181, which granted voting rights to most citizens on parole in New York State.
-This research indicates that individuals who have their rights restored prior to being discharged from parole are more likely to participate in elections, even after they are discharged.
-This is a welcome finding, as increasing the political participation of marginalized individuals brings us toward a fuller democracy.
-As with neighborhood spillover effects, however, this effect depends in great deal on the race of the parolee: rights restoration, at least in the first five months of the program in New York State, did not increase future participation among non-Whites.
-More research needs to be done to investigate whether this is true over a longer time period in New York, and whether it is true in other states as well.
-
-America is becoming ever-more-diverse, and will likely be less than 50 percent non-Hispanic White by the middle of the century.
-As Fraga (\protect\hyperlink{ref-Fraga2018}{2018}) has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
-How districts are drawn, population patterns, and rules about who can fundraise how will continue to shape the politicians we elect and the policies they enact.
-Breaking down \emph{de jure} (like formal disenfranchisement) and \emph{de facto} (like neighborhood spillover effects) barriers to the ballot box is of great importance.
-The experience in New York State over the past few years offers a better understanding of just how big this problem is, but it also offers evidence of a way forward.
-
-\newpage
-
-\hypertarget{appendix-a}{%
-\section*{Appendix A}\label{appendix-a}}
-\addcontentsline{toc}{section}{Appendix A}
-
-In the \protect\hyperlink{instrumental-variables-approach}{Instrumental Variables Approach} section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1\textsuperscript{st}, 2015, and October 14\textsuperscript{th}, 2016.
-
-\input{"../temp/table55.tex"}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
-
 \newpage
@@
 Alexander, Michelle. 2012. \emph{The New Jim Crow}. New York: The New Press.
-
-\leavevmode\hypertarget{ref-Angrist2001}{}%
-Angrist, Joshua D.
-2001.
-``Estimation of Limited Dependent Variable Models with Dummy Endogenous Regressors.'' \emph{Journal of Business \& Economic Statistics} 19 (1): 2--28.
-\url{https://doi.org/10.1198/07350010152472571}.
-
-\leavevmode\hypertarget{ref-Angrist1996}{}%
-Angrist, Joshua D., Guido W.
-Imbens, and Donald B.
-Rubin.
-1996.
-``Identification of Causal Effects Using Instrumental Variables.'' \emph{Journal of the American Statistical Association} 91 (434): 444--55.
-\url{https://doi.org/10.1080/01621459.1996.10476902}.
-
-\leavevmode\hypertarget{ref-Angrist2008}{}%
-Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
-
-\leavevmode\hypertarget{ref-Ansolabehere1999}{}%
-Ansolabehere, Stephen D., Shanto Iyengar, and Adam Simon.
-1999.
-``Replicating Experiments Using Aggregate and Survey Data: The Case of Negative Advertising and Turnout.'' \emph{American Political Science Review} 93 (4): 901--9.
-\url{https://doi.org/10.2307/2586120}.
 
@@
 
-\leavevmode\hypertarget{ref-Behrens2003}{}%
-Behrens, Angela, Christopher Uggen, and Jeff Manza.
-2003.
-``Ballot Manipulation and the ``Menace of Negro Domination'': Racial Threat and Felon Disenfranchisement in the United States, 18502002.'' \emph{American Journal of Sociology} 109 (3): 559--605.
-\url{https://doi.org/10.1086/378647}.
-
 \leavevmode\hypertarget{ref-Bowers2009}{}%
@@
 
-\leavevmode\hypertarget{ref-Burch2010}{}%
+\leavevmode\hypertarget{ref-Burch2011}{}%
 Burch, Traci.
-2010.
-``Did Disfranchisement Laws Help Elect President Bush?
-New Evidence on the Turnout Rates and Candidate Preferences of Florida's Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
-\url{https://doi.org/10.1007/s11109-010-9150-9}.
-
-\leavevmode\hypertarget{ref-Burch2011}{}%
----------.
 2011.
@@
 
-\leavevmode\hypertarget{ref-Butler2011}{}%
-Butler, Daniel M., and David E.
-Broockman.
-2011.
-``Do Politicians Racially Discriminate Against Constituents?
-A Field Experiment on State Legislators.'' \emph{American Journal of Political Science} 55 (3): 463--77.
-\url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
-
 \leavevmode\hypertarget{ref-Cho2006}{}%
@@
 
-\leavevmode\hypertarget{ref-Dong2014}{}%
-Dong, Yingying, and Arthur Lewbel.
-2014.
-``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
-\url{https://doi.org/10.1080/07474938.2014.944470}.
-
 \leavevmode\hypertarget{ref-Drucker2005}{}%
@@
 
-\leavevmode\hypertarget{ref-Einstein2016}{}%
-Einstein, Katherine Levine, and David M.
-Glick.
-2016.
-``Does Race Affect Access to Government Services?
-An Experiment Exploring Street-Level Bureaucrats and Access to Public Housing.'' \emph{American Journal of Political Science} 61 (1): 100--116.
-\url{https://doi.org/10.1111/ajps.12252}.
-
 \leavevmode\hypertarget{ref-Foladare1968}{}%
 Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
-
-\leavevmode\hypertarget{ref-Fraga2018}{}%
-Fraga, Bernard L. 2018. \emph{The Turnout Gap}. Cambridge, United Kingdom: Cambridge University Press. \url{https://doi.org/10.1017/9781108566483}.
 
@@
 \url{https://doi.org/10.1198/016214506000001040}.
-
-\leavevmode\hypertarget{ref-Gerber2000}{}%
-Gerber, Alan S., and Donald P.
-Green.
-2000.
-``The Effects of Canvassing, Telephone Calls, and Direct Mail on Voter Turnout: A Field Experiment.'' \emph{American Political Science Review} 94 (3): 653--63.
-\url{https://doi.org/10.2307/2585837}.
 
@@
 
-\leavevmode\hypertarget{ref-Green2000}{}%
-Green, Donald P., and Ron Shachar.
-2000.
-``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
-\url{https://doi.org/10.1017/s0007123400000247}.
-
 \leavevmode\hypertarget{ref-Griffin2005}{}%
@@
 \url{https://doi.org/10.1111/j.1468-2508.2005.00327.x}.
-
-\leavevmode\hypertarget{ref-Haselswerdt2009}{}%
-Haselswerdt, Michael V.
-2009.
-``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data.'' \emph{Social Science Quarterly} 90 (2): 262--73.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
-
-\leavevmode\hypertarget{ref-Hemmens2000}{}%
-Hemmens, Craig, and James W.
-Marquart.
-2000.
-``Friend or Foe?
-Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
-\url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
 
@@
 
-\leavevmode\hypertarget{ref-Hjalmarsson2010}{}%
-Hjalmarsson, Randi, and Mark Lopez.
-2010.
-``The Voting Behavior of Young Disenfranchised Felons: Would They Vote If They Could?'' \emph{American Law and Economics Review} 12 (2): 356--93.
-\url{https://doi.org/10.1093/aler/ahq004}.
-
 \leavevmode\hypertarget{ref-Huckfeldt1979}{}%
@@
 \url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
-
-\leavevmode\hypertarget{ref-Lassen2004}{}%
-Lassen, David Dreyer. 2004. ``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' \emph{SSRN Electronic Journal}. \url{https://doi.org/10.2139/ssrn.475821}.
 
@@
 
-\leavevmode\hypertarget{ref-Meredith2015}{}%
----------.
-2015.
-``The Politics of the Restoration of Ex-Felon Voting Rights: The Case of Iowa.'' \emph{Quarterly Journal of Political Science} 10 (1): 41--100.
-\url{https://doi.org/10.1561/100.00013026}.
-
 \leavevmode\hypertarget{ref-Miles2004}{}%
 Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.'' \emph{The Journal of Legal Studies} 33 (1): 85--129. \url{https://doi.org/10.1086/381290}.
-
-\leavevmode\hypertarget{ref-Milligan2004}{}%
-Milligan, Kevin, Enrico Moretti, and Philip Oreopoulos.
-2004.
-``Does Education Improve Citizenship?
-Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9-10): 1667--95.
-\url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
-
-\leavevmode\hypertarget{ref-Morris2019}{}%
-Morris, Kevin. 2019. ``Thwarting Amendment 4.'' May 9, 2019. \url{https://www.brennancenter.org/analysis/thwarting-amendment-4}.
 
@@
 
-\leavevmode\hypertarget{ref-florida_1868}{}%
-Shofner, Jerrell H. 1963. ``The Constitution of 1868.'' \emph{The Florida Historical Quarterly} 41 (4): 356--74. \url{http://www.jstor.org/stable/30139965}.
-
-\leavevmode\hypertarget{ref-Sondheimer2010}{}%
-Sondheimer, Rachel Milstein, and Donald P.
-Green.
-2010.
-``Using Experiments to Estimate the Effects of Education on Voter Turnout.'' \emph{American Journal of Political Science} 54 (1): 174--89.
-\url{https://doi.org/10.1111/j.1540-5907.2009.00425.x}.
-
-\leavevmode\hypertarget{ref-Terza2007}{}%
-Terza, Joseph V., W.
-David Bradford, and Clara E.
-Dismuke.
-2007.
-``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
-\url{https://doi.org/10.1111/j.1475-6773.2007.00807.x}.
-
 \leavevmode\hypertarget{ref-sentencing_2016}{}%
@@
 
-\leavevmode\hypertarget{ref-WHITE2014}{}%
-White, Ariel R., Noah L.
-Nathan, and Julie K.
-Faller.
-2014.
-``What Do I Need to Vote?
-Bureaucratic Discretion and Discrimination by Local Election Officials.'' \emph{American Political Science Review} 109 (1): 129--42.
-\url{https://doi.org/10.1017/s0003055414000562}.
-
 \leavevmode\hypertarget{ref-Wilson1990}{}%
 Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
-
-\leavevmode\hypertarget{ref-Wooldridge2010}{}%
-Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
 

### 2019-11-12 markdown/part2/part2_standalone.Rmd
write

--- 2019-11-12 markdown/part2_standalone.Rmd
@@
     latex_engine: xelatex
-    # includes: 
-    #   in_header: "../code/misc/preamble.tex"
     keep_tex: true
-bibliography: "citations.bib"
+bibliography: "../citations.bib"
 link-citations: yes
@@
 ```{r setup, echo = F, message = F, warning = F}
-knitr::opts_knit$set(root.dir = "..", dev = "cairo_pdf")
+knitr::opts_knit$set(root.dir = "../..", dev = "cairo_pdf")
 library(systemfit)
@@
 ```
-\input{"../temp/table3.tex"}
+\input{"../../temp/table3.tex"}
 
@@
 ```
-\input{"../temp/table_whatever2.tex"}
+\input{"../../temp/table_whatever2.tex"}
 
@@
 ```
-\input{"../temp/table4.tex"}
+\input{"../../temp/table4.tex"}
 
@@
 ```
-\input{"../temp/table_whatever.tex"}
+\input{"../../temp/table_whatever.tex"}
 Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
@@
 
-\input{"../temp/iv_clean.tex"}
+\input{"../../temp/iv_clean.tex"}
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
@@
 
-\input{"../temp/iv_clean_race.tex"}
+\input{"../../temp/iv_clean_race.tex"}
 
@@
 ```
-\input{"../temp/table5.tex"}
+\input{"../../temp/table5.tex"}
 Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were black.
@@
 
-\input{"../temp/iv_clean_months.tex"}
+\input{"../../temp/iv_clean_months.tex"}
 
@@
 ```
-\input{"../temp/table55.tex"}
+\input{"../../temp/table55.tex"}
 

### 2019-11-12 markdown/part1/p1.Rmd
write

@@
 
-The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-Michelle Alexander [-@Alexander2012] and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.] This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
-
 Due to economic and racial segregation, these effects are highly spatially concentrated.
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
+The result is an incarcerated population that looks far different than the rest of th state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.]
 

### 2019-11-12 markdown/part1/p1.tex
write

@@
 
-The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is likely due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific criminal propensity.
-
 Due to economic and racial segregation, these effects are highly spatially concentrated.
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
+The result is an incarcerated population that looks far different than the rest of th state: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.}
 
@@
 
-\leavevmode\hypertarget{ref-Alexander2012}{}%
-Alexander, Michelle. 2012. \emph{The New Jim Crow}. New York: The New Press.
-
 \leavevmode\hypertarget{ref-Anzia2019}{}%

### 2020-01-28 markdown/felony_disenfranchisement_nys.Rmd
too many changes but need to commit

@@
 subtitle: "Felony Disenfranchisement and New York State"
-author: "Kevin Morris"
+author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
+Despite the United States' march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised [@sentencing_2016].
@@
 If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have "aged in" to voting but for their incarceration, they will not will have no history of voting.
+Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have "aged in" to voting but for their incarceration, they will have no history of voting.
 This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
@@
 In many cases, these "lost voters" will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
+However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effect in its wake.
 
@@
 For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))), digits = 1)` votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
+The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters when treatment is interacted with racial demographics.
 This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
@@
 Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
+The relationship does not hold for tracts (surprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
 
@@
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
+The interlocking nature of racioeconomic segregation, policing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
 
@@
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
 This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.

### 2020-01-28 markdown/felony_disenfranchisement_nys.tex
too many changes but need to commit

@@
-\documentclass[12pt,]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  12pt,
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={In the Shadow of Jim Crow},
   pdfauthor={Kevin Morris},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{In the Shadow of Jim Crow}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{Felony Disenfranchisement and New York State}
-  \author{Kevin Morris}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{November 05, 2019}
-
+\usepackage{setspace}\doublespacing
+\usepackage{rotating}
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
-
-\usepackage{setspace}\doublespacing
-\usepackage{rotating}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
+\title{In the Shadow of Jim Crow}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Felony Disenfranchisement and New York State}
+\author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
+\date{January 22, 2020}
 
@@
 Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+Despite the United States' march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
@@
 Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
-Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
+Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and García-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and García-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
 These effects are particularly pronounced for women of color.
@@
 
-Similarly, A.
-White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
+Similarly, White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
 She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (607).
 This study examines only the indirect effects of proximal contact with the criminal justice system.
-It is possible that the indirect effects of the incarceration or jailing of an individual who would have voted but not for their contact with the criminal justice system would have larger indirect effects.
+It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect depressive effect of incarceration more generally.
 Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
@@
 
-A.
 White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), however, comes to a different conclusion.
@@
 \fontsize{10}{12}\selectfont
-\begin{tabular}{cc}
+\begin{tabular}[t]{cc}
 \toprule
@@
 If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have ``aged in'' to voting but for their incarceration, they will not will have no history of voting.
+Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have ``aged in'' to voting but for their incarceration, they will have no history of voting.
 This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
@@
 In many cases, these ``lost voters'' will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
+However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effect in its wake.
 
@@
 For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
+The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters when treatment is interacted with racial demographics.
 This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
@@
 Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (A.
-White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}).
+As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}).
 As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018.
-This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
-senatorial contests (as well as other down-ballot races).
-
-I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
 This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
@@
 For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.}
+For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 Some of this may be explained by different propensities to vote.
-A.
 White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Abadie2019}{}%
@@
 \url{https://doi.org/10.1007/s11109-010-9150-9}.
-
-\leavevmode\hypertarget{ref-Burch2011}{}%
----------.
-2011.
-``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
-\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
 
@@
 \url{https://doi.org/10.1177/0002716213503093}.
+
+\leavevmode\hypertarget{ref-Burch2011}{}%
+---------.
+2011.
+``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
+\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
 
@@
 2006.
-``'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
+``\,'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
 \url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
@@
 
+\leavevmode\hypertarget{ref-Uggen2004}{}%
+Uggen, Christopher, and Jeff Manza.
+2004.
+``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
+New York: Russell Sage Foundation.
+
 \leavevmode\hypertarget{ref-Uggen2002}{}%
-Uggen, Christopher, and Jeff Manza.
+---------.
 2002.
@@
 \url{https://doi.org/10.2307/3088970}.
-
-\leavevmode\hypertarget{ref-Uggen2004}{}%
----------.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
 
@@
 \leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela Garcı́a-Castañon.
+Walker, Hannah L., and Marcela García-Castañon.
 2017.
@@
 \leavevmode\hypertarget{ref-WHITE2014}{}%
-White, Ariel R., Noah L.
-Nathan, and Julie K.
-Faller.
+White, Ariel, Noah Nathan, and Julie Faller.
 2014.
@@
 \url{https://doi.org/10.1080/07418820802119950}.
+\end{cslreferences}
 

### 2020-01-28 markdown/part1/p1_no_tract.Rmd
too many changes but need to commit

--- 2019-11-12 markdown/part1/p1.Rmd
@@
 title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
-author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
+thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
+author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
 institute: "Brennan Center for Justice"
@@
     latex_engine: xelatex
-    keep_tex: true
+    keep_tex: false
 bibliography: "../citations.bib"
@@
     - \usepackage{rotating}
-    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
 abstract: |
@@
 In this paper, I redefine “lost voters” as disenfranchised individuals with a history of participating in elections.
-I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods.
+I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
 I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
@@
 ---
-<!-- --- -->
-<!-- title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City" -->
-<!-- author: "Kevin Morris" -->
-<!-- date: "`r format(Sys.time(), '%B %d, %Y')`" -->
-<!-- output: word_document -->
-<!-- bibliography: "../research/citations.bib" -->
-<!-- link-citations: yes -->
-<!-- always_allow_html: yes -->
-<!-- --- -->
 \pagenumbering{gobble}
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is an incarcerated population that looks far different than the rest of th state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.]
+The result is an incarcerated population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.]
 
@@
 Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
-@Uggen2002 produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for for disenfranchisment in Florida.
+@Uggen2002 produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for disenfranchisement in Florida.
 Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
@@
 Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
+More than two-thirds of block groups that lost any voters lost only one voter.
 
@@
   mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
+
+t <- t[, c(1,2)]
 
@@
 I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups [@Sekhon2011], using a series of demographic and political indicators.
+Neighborhoods are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
+I use a genetic match to match treated to untreated census block groups [@Sekhon2011], using a series of demographic and political indicators.
 Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
@@
 I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
+Each treated block group is matched to 30 untreated block groups.
 Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
+Table \ref{tab:blockgroup-table} presents the results of these matches.
 
 ## Match Output {-}
+As Table \ref{tab:blockgroup-table} demonstrates, the matching procedure results in an untreated group of neighborhoods that looks substantially like the treatment group.
+
 ```{r block-table-chunk, echo = F}
@@
 
-```{r tract-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_tract.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:tract-table) Results of Tract-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-
 In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
@@
 Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] The median income in neighborhoods without lost voters is more than 40 percent higher.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
+Table \ref{tab:blockgroup-table} throws into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
 
 After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
+Census block group level turnout rates are calculated using the geocoded voter file.
 Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
@@
 Robust standard errors are clustered at the level of the match [@Abadie2019].
+Models 2 and 4 include the covariates on which the matching was done.
 ```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 
-load("./temp/match_reg_tract.rdata")
-
-load("./temp/match_reg_tract_nhb.rdata")
-tract_reg <- reg_output
-tract_reg_ses <- reg_output_ses
-
-tract_reg_black <- reg_output_nhblack
-tract_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, bg_reg_black, tract_reg, tract_reg_black,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
+stargazer(bg_reg, reg_output2, bg_reg_black, reg_output2_nhblack,
           header = F,
-          se = list(bg_reg_ses, bg_reg_black_ses, tract_reg_ses, tract_reg_black_ses),
+          se = list(bg_reg_ses, reg_output2_ses, bg_reg_black_ses, reg_output2_nhblack_ses),
           notes = c("TO REPLACE"),
@@
           dep.var.labels = c("Turnout Rate in 2017 General Election"),
+          omit = c("median_income", "latino", "nh_white", "some_college",
+                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner"),
           title = "\\label{tab:match-reg} Matching Regression",
@@
           out = "./temp/test.tex",
-          out.header = F)
+          out.header = F,
+          add.lines=list(c("Controls Included" , " ", "X", " ", "X")))
 
@@
 
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
 
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
   arrange(n) %>% 
@@
 When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*100, digits = 1)` percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just `r round((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))*100, digits = 1)` percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as `r round(100 - ((100 * ((sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap)) + bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]])) / (sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$v2017) / sum(filter(readRDS("./temp/block_group_pre_match.rds"), treat == T)$cvap))), digits = 1)` votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
+The average treated neighborhood was 38 percent Black, which means that the average block group with a lost voter saw turnout that was `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*38, digits = 2)` percentage points lower than similar neighborhoods without lost voters.
+This is very close to the average treatment effect presented in Models 1 and 2; Models 3 and 4, however, makes clear that this effect is largely concentrated in Black communities.
+This disparate effect could have troubling ramifications for these neighborhoods' political representation.
 
@@
 ```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/tract_model_reg_ols.rdata")
 load("./temp/bg_model_reg_ols.rdata")
 
-stargazer(bg_model, bg_model2, tract_model, tract_model2,
-          column.separate = c(2, 2),
-          column.labels = c("Block Group Level", "Tract Level"),
+stargazer(bg_model, bg_model2,
           header = F,
@@
           notes.append = T,
-          se = list(bg_model_ses, bg_model2_ses, tract_model_ses, tract_model2_ses),
+          se = list(bg_model_ses, bg_model2_ses),
           covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
@@
 
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
 \\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
@@
 
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
+The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model, though they are slightly smaller.
+This is unsurprising, because the average treated neighborhood had more than one lost voter.
+Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Model 2 again makes clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
@@
 
-Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
-As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
 
@@
 ```
-
-```{r tract-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_tract_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match-tract) Results of Tract-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
-
-In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups and tracts.
-In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly *higher* turnout prior to intervention.
+In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups.
+The variable D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
+D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election, a difference that was not statistically significant.
 D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
+This coefficient is negative and statistically significant, indicating that in the first election after the removal of a voter a neighborhood's turnout rate decreased by 2.6 percentage points.
 
@@
 
-load("./temp/match_reg_tract_did.rdata")
-
-tract_reg <- reg_output
-tract_reg_ses <- reg_output_ses
-
-stargazer(bg_reg, tract_reg,
+stargazer(bg_reg, reg_output2,
           header = F,
-          column.labels = c("Block Group Level", "Tract Level"),
-          se = list(bg_reg_ses, tract_reg_ses),
+          se = list(bg_reg_ses, reg_output2_ses),
           notes = c("TO REPLACE"),
@@
           covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
-          omit = "district",
+
           dep.var.labels = c("Turnout Rate"),
@@
           out = "./temp/test_abc.tex",
+          omit = c("median_income", "latino", "nh_white", "some_college", "district",
+                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black"),
           out.header = F,
-          add.lines=list(c("Congressional District FE", "X", "X")))
+          add.lines=list(c("Congressional District FE", "X", "X"),
+                         c("Controls Included", " ", "X")))
 
@@
 
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-2} \\cmidrule(l{3pt}r{3pt}){3-3}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
 
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
   arrange(n) %>% 
@@
 Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.] Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.]
 
@@
 # References {-}
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-\noindent

### 2020-01-28 markdown/part2/part2_standalone.Rmd
too many changes but need to commit

@@
 subtitle: "Voting Rights Restoration and Post-Supervision Participation"
-author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
+thanks: Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
+The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
+All errors are my responsibility.
+author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
 institute: "Brennan Center for Justice"
@@
 
-The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle [-@Alexander2012] and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is black.
-This disparity is likely due *not* to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As @Gelman2007 shows, for instance, New York's ``stop-and-frisk'' policy targeted black and Latino New Yorkers at rates far higher than white residents, even after controlling for neighborhood variability and race-specific arrest rates.
-
-There is some evidence, however, that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019; but see @Gerber2017].
@@
 
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
+The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
 A Chi-squared test confirms that the model is not improved when controls for time are included.
@@
 Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18^th^ that parolees consistently had their voting rights restored prior to discharge.
-```{r restoration-chart-chunk, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/restoration_plot.rds")
-print(j)
-```
+According to the rights restoration data, it is not until May 18^th^ that the majority of individuals discharged from parole had their rights restored prior to discharge.
 
@@
 Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
 Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
@@
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
+After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.84 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.96 percentage points.
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` and `r round((0.85 / (3.0 - 0.85)) * 100, digits = 1)` percent.
+Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between `r round((0.84 / (3.1684 - 0.84)) * 100, digits = 1)` and `r round((0.96 / (3.1684 - 0.96)) * 100, digits = 1)` percent.
 
@@
 
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.6 percentage points.
-Given that just 3.3 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.8 percentage points.
+Given that just 3.5 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{doubled} these individuals' likelihood of casting a ballot.
 We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
@@
 These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, black individuals were still 36 percent more likely to cast a ballot.
+Even when Latinos are excluded from this group, black individuals were still a third more likely to cast a ballot.
 In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
@@
 Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
+Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.29 percentage points (in the 2SLS model) and 0.50 percentage points (in the 2S probit model).
 
@@
 In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.74 / (3.0 - 0.74)) * 100, digits = 1)` percent.
+In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.84 / (3.1684 - 0.84)) * 100, digits = 1)` percent.
 This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.

### 2020-01-28 markdown/part2/part2_standalone.tex
write
too many changes but need to commit

--- 2019-11-12 markdown/part2_standalone.tex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={Welcome Home --- Now Vote!},
   pdfauthor={Kevin Morris},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Welcome Home --- Now Vote!}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{Voting Rights Restoration and Post-Supervision Participation}
-  \author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{November 12, 2019}
-
+\usepackage{rotating}
+\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
-\usepackage{rotating}
-\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+\title{Welcome Home --- Now Vote!\thanks{Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
+The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
+All errors are my responsibility.}}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Voting Rights Restoration and Post-Supervision Participation}
+\author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
+\date{January 23, 2020}
 
@@
 
-The disenfranchisement of citizens convicted of felony offenses is particularly troubling given the racialized and place-based patterns of policing and incarceration in the United States.
-As Michelle (\protect\hyperlink{ref-Alexander2012}{2012}) and others have shown, mass incarceration in the post-Civil Rights era has been a tool used by the state to exert control over minority --- and particularly black --- Americans.
-This is clear in the case of states such as New York: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were black, although the Census Bureau estimates that just 14.6 percent of the citizen voting age population in the state is black.
-This disparity is likely due \emph{not} to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy targeted black and Latino New Yorkers at rates far higher than white residents, even after controlling for neighborhood variability and race-specific arrest rates.
-
-There is some evidence, however, that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}; but see Gerber et al.
@@
 \midrule
-Actual Birthdate & 68,340\\
-Birthdate + 35 Days & 480\\
-Birthdate - 35 Days & 419\\
+Actual Birthdate & 69,644\\
+Birthdate + 35 Days & 502\\
+Birthdate - 35 Days & 426\\
 \bottomrule
@@
 The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.6 percent cast a ballot in the 2016 election, while just 4.2 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
+For instance, of the individuals last discharged from parole in 2010, 6.8 percent cast a ballot in the 2016 election, while just 4.3 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
 Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
@@
 
-\input{"../temp/table3.tex"}
+\input{"../../temp/table3.tex"}
 
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
+The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
 A Chi-squared test confirms that the model is not improved when controls for time are included.
@@
 Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18\textsuperscript{th} that parolees consistently had their voting rights restored prior to discharge.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/restoration-chart-chunk-1}
-
-}
-
-\caption{\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge}\label{fig:restoration-chart-chunk}
-\end{figure}
+According to the rights restoration data, it is not until May 18\textsuperscript{th} that the majority of individuals discharged from parole had their rights restored prior to discharge.
 
@@
 
-\input{"../temp/table_whatever2.tex"}
+\input{"../../temp/table_whatever2.tex"}
 
@@
 
-\input{"../temp/table4.tex"}
+\input{"../../temp/table4.tex"}
 
@@
 Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 31.2 and 31.8 percent.
-The turnout rate for the control group was 2.0 percent.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 34.5 and 35 percent.
+The turnout rate for the control group was 2.1 percent.
 
@@
 
-\input{"../temp/table_whatever.tex"}
+\input{"../../temp/table_whatever.tex"}
 
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
+Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'\,') is strongly correlated with certain demographics that are also linked to propensity to vote.
 Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
@@
 
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the \{{[}\}0, 1\{{]}\} interval.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'\,' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
 Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
@@
 
-\input{"../temp/iv_clean.tex"}
+\input{"../../temp/iv_clean.tex"}
 
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
+After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.84 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.96 percentage points.
 Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 32.7 and 39.5 percent.
+Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 36.1 and 43.5 percent.
 
@@
 
-\input{"../temp/iv_clean_race.tex"}
+\input{"../../temp/iv_clean_race.tex"}
 
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.6 percentage points.
-Given that just 3.3 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
+Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.8 percentage points.
+Given that just 3.5 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{doubled} these individuals' likelihood of casting a ballot.
 We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
@@
 
-\input{"../temp/table5.tex"}
+\input{"../../temp/table5.tex"}
 
@@
 
-\input{"../temp/iv_clean_months.tex"}
+\input{"../../temp/iv_clean_months.tex"}
 
 Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
+Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.29 percentage points (in the 2SLS model) and 0.50 percentage points (in the 2S probit model).
 
@@
 Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just 792, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
+Though turnout among this group was low (just 832, or 3.81 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
 
@@
 In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least 32.7 percent.
+In this project I estimate that rights restoration increased individuals' propensity to vote by at least 36.1 percent.
 This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
@@
 
-\input{"../temp/table55.tex"}
+\input{"../../temp/table55.tex"}
 
@@
 
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\noindent
-
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Alexander2012}{}%
-Alexander, Michelle. 2012. \emph{The New Jim Crow}. New York: The New Press.
-
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Angrist2001}{}%
@@
 \url{https://doi.org/10.1111/ajps.12252}.
-
-\leavevmode\hypertarget{ref-Gelman2007}{}%
-Gelman, Andrew, Jeffrey Fagan, and Alex Kiss.
-2007.
-``An Analysis of the New York City Police Departments ``Stop-and-Frisk'' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
-\url{https://doi.org/10.1198/016214506000001040}.
 
@@
 Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
+\end{cslreferences}
 

### 2020-01-28 presentations/spsa.Rmd
too many changes but need to commit

---
title: "Felony Disenfranchisement and Neighborhood Turnout"
subtitle: "The Case of New York City"
shorttitle: "Other"
author: Kevin Morris
abstract: kevin.morris@nyu.edu
institute: Brennan Center for Justice
date: Southern Political Science Association, 2020
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "Berlin"
    keep_tex: true
bibliography: "../markdown/citations.bib"
---
```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(dev = "cairo_pdf")
library(systemfit)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(tidyverse)
library(extrafont)
library(kableExtra)
library(stargazer)
library(data.table)
loadfonts()
```

## Outline

>* Introducing space into the conversation

>* Redefining "lost voters"

>* Identifying neighborhoods with lost voters

>* Testing turnout effects

## Introducing Space

>* Most of the existing literature on felony disenfranchisement and indirect turnout effects looks for effects at the state level [@Miles2004; @Bowers2009; @King2016]

>* But we *know* that incarceration patterns aren't uniformly distributed throughout a given state [see, for instance, @Gelman2007]

>* Turnout effects are likely to be socially mediated - and therefore show up in the neighborhoods where disenfranchised individuals live [e.g. @Foladare1968; @Huckfeldt1979; @Cho2006]

## Redefining Lost Voters

>* Much of the existing literature cannot identify individuals who *would have voted* if not for their disenfranchisement [@Burch2013]

>* This undermines our ability to distinguish the causal effects of incarceration / probation from the effects of disenfranchisement

>* This is especially likely given the low turnout propensity among the formerly disenfranchised [e.g. @Meredith2015; @Gerber2017; @White2019]

## This Project Addresses Both Problems

>* Administrative data from the NYS DOC allows us to conduct individual-level analysis

>* Purge records in the NYS voter file allow us to construct vote histories for disenfranchised individuals

>* Lost voters are all formally disenfranchised individuals who have voted in the past 10 years

## Identifying Neighborhoods with Lost Voters

```{r ttt, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../temp/lost_voters_map.png")
```

## Methodology

* A genetic match algorithm [@Sekhon2011] is used to match treated census block groups to untreated ones.

* Each block group is matched to 30 untreated block groups; matches are done with replacement.

## Matching Results

```{r, echo=FALSE,out.width="49%", out.height="49%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/perc_white_spsa.png", "../temp/perc_black_spsa.png"))
```

## Matching Results

```{r, echo=FALSE,out.width="49%", out.height="49%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/perc_latino_spsa.png", "../temp/income_spsa.png"))
```

## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot1.png"))
```
## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot2.png"))
```

## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot3b.png"))
```

## Testing Turnout Effects

```{r post-est, echo=FALSE, out.width = '100%'}
knitr::include_graphics(c("../output/dep_map.png"))
```

## Conclusions

>* The indirect turnout effects identified in past research appear to be geographically concentrated where the lost voters live.

>* @Hajnal2009 and others have demonstrated that turnout differentials can have real political implications, especially in low-turnout contests.

>* Felony disenfranchisement undermines Black political representation at the local level.

## We Made It!

Thanks!

kevin.morris@nyu.edu

## References {.allowframebreaks}

### 2020-01-28 presentations/spsa.tex
too many changes but need to commit

--- 2019-10-18 markdown/felony_disenfranchisement_nys.tex
@@
-\documentclass[12pt,]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  ignorenonframetext,
+]{beamer}
+\usepackage{pgfpages}
+\setbeamertemplate{caption}[numbered]
+\setbeamertemplate{caption label separator}{: }
+\setbeamercolor{caption name}{fg=normal text.fg}
+\beamertemplatenavigationsymbolsempty
+% Prevent slide breaks in the middle of a paragraph
+\widowpenalties 1 10000
+\raggedbottom
+\setbeamertemplate{part page}{
+  \centering
+  \begin{beamercolorbox}[sep=16pt,center]{part title}
+    \usebeamerfont{part title}\insertpart\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{section page}{
+  \centering
+  \begin{beamercolorbox}[sep=12pt,center]{part title}
+    \usebeamerfont{section title}\insertsection\par
+  \end{beamercolorbox}
+}
+\setbeamertemplate{subsection page}{
+  \centering
+  \begin{beamercolorbox}[sep=8pt,center]{part title}
+    \usebeamerfont{subsection title}\insertsubsection\par
+  \end{beamercolorbox}
+}
+\AtBeginPart{
+  \frame{\partpage}
+}
+\AtBeginSection{
+  \ifbibliography
+  \else
+    \frame{\sectionpage}
+  \fi
+}
+\AtBeginSubsection{
+  \frame{\subsectionpage}
+}
 \usepackage{lmodern}
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+\usetheme[]{Berlin}
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={In the Shadow of Jim Crow},
-            pdfauthor={Kevin Morris},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Felony Disenfranchisement and Neighborhood Turnout},
+  pdfauthor={Kevin Morris},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\newif\ifbibliography
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{In the Shadow of Jim Crow}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{Felony Disenfranchisement and New York State}
+\title{Felony Disenfranchisement and Neighborhood Turnout}
+\subtitle{The Case of New York City}
 \author{Kevin Morris}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{October 17, 2019}
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
-\usepackage{setspace}\doublespacing
-\usepackage{rotating}
+\date{Southern Political Science Association, 2020}
+\institute{Brennan Center for Justice}
 
 \begin{document}
-\maketitle
+\frame{\titlepage}
+\begin{abstract}
+\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}}
+\end{abstract}
 
-\pagenumbering{gobble}
-
-\tableofcontents
-\pagebreak
-
-\pagenumbering{arabic}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
-At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting: in most of the country citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-The disenfranchisement of citizens convicted of felony offenses intersects with the racialized and place-based patterns of policing and incarceration in the United States.
-Michelle Alexander (\protect\hyperlink{ref-Alexander2012}{2012}) and others have argued that mass incarceration in the post-Civil Rights era has been used to exert control over minority --- and particularly Black --- Americans.
-In states such as New York, for instance, the over-representation of minorities among the incarcerated is striking: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.} This disparity is due not to any inherent differential propensity to commit crimes among different racial groups, but rather to systems of policing and concentrated poverty.
-As Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood characteristics and race-specific criminal propensity.
-
-Due to economic and racial segregation, these effects are highly spatially concentrated.
-Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
-Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health (Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}) and are more likely to suffer from anxiety and exhibit symptoms of trauma (Geller et al.
-\protect\hyperlink{ref-Geller2014}{2014}).
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted (Clear \protect\hyperlink{ref-Clear2008}{2008}), while concentrated policing has also been credited with having a ``chilling effect'' on neighborhoods' willingness to reach out for help to local governments (Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-
-Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
-The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}; Gerber et al.
-\protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely minimal.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (8).
-
-The rest of this paper is structured as follows: \textbf{Section 2} discusses the historical context of felony disenfranchisement in the United States, and \textbf{Section 3} provides an overview of the scholarship conducted in the field over the past two decades.
-\textbf{Section 4} provides an in-depth discussion of the insight extant literature provides for the questions discussed in this project.
-\textbf{Section 5} presents the data sources used throughout the project.
-
-\textbf{Section 6} proposes a new definition of ``lost voters'' in the context of felony disenfranchisement.
-I define lost voters as individuals who are not only disenfranchised, but are disenfranchised \emph{and} have a history of participating in elections.
-This serves as a refinement on past work, which considered only which individuals are formally disenfranchised, regardless of whether they had cast a ballot in the past.
-Using both a matching model and a standard regression specification, I investigate whether neighborhoods with lost voters saw depressed turnout in the 2017 mayoral election relative to other neighborhoods.
-The disadvantaged state of neighborhoods with high levels of incarceration has been firmly established; if felony disenfranchisement results in lower turnout among eligible voters in these communities, the literature has likely \emph{understated} the degree to which they are impacted by spatially concentrated patterns of criminal justice.
-A discussion of the results follows.
-
-\textbf{Section 7} turns from the 2017 mayoral election to the 2018 midterm elections.
-In early 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most individuals in the state who were on parole.
-I employ an instrumental variables approach to test whether rights restoration boosted turnout among formerly incarcerated individuals.
-I focus specifically on individuals who would have been eligible to cast a ballot even if the executive order had not gone into effect.
-A discussion of the results follows.
-
-\textbf{Section 8} synthesizes the findings presented in the rest of the paper and provides concluding remarks.
-
-\hypertarget{background-of-felony-disenfranchisement-in-the-united-states}{%
-\section{Background of Felony Disenfranchisement in the United States}\label{background-of-felony-disenfranchisement-in-the-united-states}}
-
-In all but two states (Maine and Vermont), felony disenfranchisement laws ensure that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state government (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-This variation in laws flows directly from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as \emph{Richardson v.
-Ramirez} (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
-
-Any discussion of felony disenfranchisement in the United States must center the role played by race.
-As Burch (\protect\hyperlink{ref-Burch2010}{2010}) explains, ``If policies restricting the voting rights of offenders disparately affect one racial group or party, it is because such policies were \emph{intended} to'' (4; emphasis in the original).
-Previous research has established that the presence of non-White potential voters is associated with the implementation of felony disenfranchisement policies, and that these policies were often adopted during Jim Crow explicitly to reduce the political power of Black Americans (Behrens, Uggen, and Manza \protect\hyperlink{ref-Behrens2003}{2003}).
-In Florida, to take just one example, felony disenfranchisement was added to the state constitution in 1868.
-Afterwards, a lawmaker boasted that the amendment had been adopted to prevent the state from being ``niggerized'' (Shofner \protect\hyperlink{ref-florida_1868}{1963}, 374).
-
-The racially unbalanced effects of felony disenfranchisement laws were not confined to the 19th century.
-Although the Voting Rights Act of 1965 did much to improve access to the ballot box for minorities, it did nothing to undermine the explicitly racialized system of disenfranchisement.
-Indeed, as the United States has increased the reach of the carceral system in the post-Civil Rights era, the implications of felony disenfranchisement have only grown.
-As of 2016, more than 10 percent of Black Americans were disenfranchised in 9 states.
-In Kentucky, the state with the highest level of disenfranchised Black residents, more than one in four Black adults are barred from casting a ballot.
-Although Black adults made up just 12.1 percent of the voting age population in 2016, they accounted for 36.5 percent of the disenfranchised population (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-Recent data from Florida make the victims of felony disenfranchisement laws especially clear.
-Prior to January 2019, Floridians convicted of felony offenses were permanently disenfranchised unless they were individually pardoned by a clemency board.
-In November 2018, voters passed a ballot initiative amending the state constitution to end permanent disenfranchisement.
-In the first three months after permanent disenfranchisement was ended, 44 percent of formerly incarcerated Floridians who registered to vote were Black --- compared with just 13 percent of the statewide electorate (Morris \protect\hyperlink{ref-Morris2019}{2019}).
-
-\hypertarget{academic-literature-and-felony-disenfranchisement}{%
-\section{Academic Literature and Felony Disenfranchisement}\label{academic-literature-and-felony-disenfranchisement}}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the relationship between felony disenfranchisement and Black and youth turnout (Miles \protect\hyperlink{ref-Miles2004}{2004}; Hjalmarsson and Lopez \protect\hyperlink{ref-Hjalmarsson2010}{2010}).
-Some of this research compared states and regions with differing disenfranchisement regimes to estimate these effects (Miles \protect\hyperlink{ref-Miles2004}{2004}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-Others have used survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}) expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states.
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she was also able to estimate the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
-
-A number of papers have also explored the impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among eligible Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates by race, and include estimates of the share of Black citizens who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
-{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the \emph{state} level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining Census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
-
-\hypertarget{what-this-paper-examines}{%
-\section{What This Paper Examines}\label{what-this-paper-examines}}
-
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods (e.g.~Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}; Clear \protect\hyperlink{ref-Clear2008}{2008}; Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
-
-I begin by offering a refinement of our understanding of who lost voters are.
-Previous studies exploring the relationship between felony disenfranchisement and spillover effects have not adequately accounted for the fact that many of the disenfranchised are low-propensity voters to begin with.
-In this analysis, I use administrative data to identify neighborhoods where individuals with a \emph{demonstrated history} of participation are disenfranchised.
-By matching ``treated'' neighborhoods (neighborhoods with lost voters) to ``untreated'' neighborhoods, I estimate the causal effect of felony disenfranchisement on neighborhood turnout.
-
-After examining the effect of felony disenfranchisement on voter turnout at the neighborhood level, I consider the efficacy of a program intended to undo one piece of the felony disenfranchisement apparatus.
-In April 2018, Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to New Yorkers on parole.\footnote{Prior to Executive Order 181, residents convicted of felonies in prison and on parole were barred from voting, while individuals sentenced to probation did not have their voting rights revoked.
-Today, only those currently in prison for felony offenses are disenfranchised.} Parole officers were required to provide registration forms to the parolees under their supervision and inform them of their voting rights.
-In theory, such a policy could increase turnout and registration for parolees.
-To test whether the policy was successful, I use administrative data to explore whether turnout rates in the 2018 midterm elections among individuals discharged from parole were impacted by the implementation of Executive Order 181.
-
-The first part of this project seeks to refine how we think about felony disenfranchisement in the United States.
-Rather than think about it purely through the prism of those directly impacted (residents who are actively disenfranchised) or through a prism of race (by examining the impact of disenfranchisement on Black turnout at the state level), I seek to introduce the importance of physical space into the conversation.
-This will allow us to both better understand the mechanisms through which the depressive effects operate and also to understand more deeply where these effects are concentrated.
-Of course, understanding the impact of felony disenfranchisement is not enough: we must also test whether our efforts to undo its legacy are effective.
-The second part of this project seeks to do just that.
-
-\hypertarget{framing-for-turnout-effects}{%
-\subsection*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
-\addcontentsline{toc}{subsection}{Framing for Turnout Effects}
-
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, 46,232 individuals were imprisoned in New York State in early 2019, compared with 11.6 million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
-
-Though previous studies have focused largely on the state-level spillover effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
-Moreover, voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
-It should be no surprise, then, that neighborhood characteristics have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-I therefore hypothesize that felony disenfranchisement does have locally bounded spillover effects, and that turnout in neighborhoods with lost voters systematically vote at a lower rate than others.
-
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the neighborhood spillover effects of lost voters.
-Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
-These effects are particularly pronounced for women of color.
-Although Walker and others demonstrate that these actions largely take place outside of the voting booth, they may moderate spillover effects on turnout.
-
-\hypertarget{theoretical-implications-for-reduced-turnout}{%
-\subsubsection*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
-\addcontentsline{toc}{subsubsection}{Theoretical Implications for Reduced Turnout}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community (e.g.~King and Erickson \protect\hyperlink{ref-King2016}{2016}; Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: Kinsella, McTague, and Raleigh (\protect\hyperlink{ref-Kinsella2015}{2015}) examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes (Martin \protect\hyperlink{ref-Martin2003}{2003}).
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-``{[}H{]}igher citizen participation is rewarded,'' Martin and Claibourn (\protect\hyperlink{ref-Martin2013}{2013}) concludes, ``with enhanced policy responsiveness'' (59).
-Griffin and Newman (\protect\hyperlink{ref-Griffin2005}{2005}) finds similar effects in the United States Senate, demonstrating that ``voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not'' (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
-
-Although research on the impact of local turnout on city-wide policy is scarce, Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) examines the impact of senior turnout on ``senior-friendly'' policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials \emph{are} responsive to senior turnout when seniors ``are a more cohesive, meaningful group'' (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of \emph{neighborhood} turnout rates on city policy.
-
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}), however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) tells us that ``Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils'' (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
-
-\hypertarget{turnout-among-formerly-disenfranchised-individuals}{%
-\subsection*{Turnout Among Formerly Disenfranchised Individuals}\label{turnout-among-formerly-disenfranchised-individuals}}
-\addcontentsline{toc}{subsection}{Turnout Among Formerly Disenfranchised Individuals}
-
-Much of the literature discussed above has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so (Haselswerdt \protect\hyperlink{ref-Haselswerdt2009}{2009}; Burch \protect\hyperlink{ref-Burch2011}{2011}; Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}).
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}) uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that ``it appears that spending time in prison does not have large negative effects on subsequent participation'' (1144).
-
-White (\protect\hyperlink{ref-White2019}{2019}), however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, incarceration often occurs after many other interactions with the criminal justice system.
-By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration \emph{causes} low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
-
-The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
-I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
-
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2006}).
-Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
-Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-``Whatever the participatory consequences of incarceration,'' they conclude, ``they are not in large part impossible to overcome'' (924).
-Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
-
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that ``contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual'' (829).
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-Such repairs may lead to increased propensity to participate in elections in the future.
-
-\hypertarget{data}{%
-\section{Data}\label{data}}
-
-\hypertarget{criminal-justice-data}{%
-\subsection*{Criminal Justice Data}\label{criminal-justice-data}}
-\addcontentsline{toc}{subsection}{Criminal Justice Data}
-
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole and demographic characteristics.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors available.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-The state does not make a unified database of parolees whose voting rights have been restored available to the public.
-However, the NYSDOCCS Parolee Lookup website includes a flag indicating whether someone's voting rights have been restored.
-By using identification numbers from the parolee data obtained from the public records request, I constructed a list of individuals who have had their rights restored.
-
-\hypertarget{voter-file-data}{%
-\subsection*{Voter File Data}\label{voter-file-data}}
-\addcontentsline{toc}{subsection}{Voter File Data}
-
-Most states in the United States are required to maintain files with information on all registered and recently purged voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; home address; vote history; and other information.
-The first section of this paper uses a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
-The second section of this paper uses a snapshot from March 3\textsuperscript{rd}, 2019.
-
-The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered (and participated) in the past but have since been purged due to a felony conviction.
-
-\hypertarget{geocoding}{%
-\subsection*{Geocoding}\label{geocoding}}
-\addcontentsline{toc}{subsection}{Geocoding}
-
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to Census block groups, Census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6 percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
-
-\hypertarget{matching}{%
-\subsection*{Matching}\label{matching}}
-\addcontentsline{toc}{subsection}{Matching}
-
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a ``match,'' records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
-
-\begin{itemize}
+\begin{frame}{Outline}
+\protect\hypertarget{outline}{}
+\begin{itemize}[<+->]
 \tightlist
 \item
-  Middle names are identical. If neither set of records includes a middle name, this condition is met.
-\item
-  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-\item
-  A middle name or middle initial in one set of records, and a missing middle name in the other set.
+  Introducing space into the conversation
 \end{itemize}
 
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Redefining ``lost voters''
+\end{itemize}
 
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Identifying neighborhoods with lost voters
+\end{itemize}
 
-Testing for the presence of false positive matches is fairly straightforward.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Testing turnout effects
+\end{itemize}
+\end{frame}
 
-\begin{table}[H]
+\begin{frame}{Introducing Space}
+\protect\hypertarget{introducing-space}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Most of the existing literature on felony disenfranchisement and
+  indirect turnout effects looks for effects at the state level (Miles
+  2004; Bowers and Preuhs 2009; King and Erickson 2016)
+\end{itemize}
 
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 20,955\\
-Birthdate + 35 Days & 105\\
-Birthdate - 35 Days & 92\\
-\bottomrule
-\end{tabular}
-\end{table}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  But we \emph{know} that incarceration patterns aren't uniformly
+  distributed throughout a given state (see, for instance, Gelman,
+  Fagan, and Kiss 2007)
+\end{itemize}
 
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely or use different names in different contexts.
-This is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been released from incarceration in New York State.
-Secondly, because both Department of Corrections and voter registration data are legal records, individuals are likely to be recorded using their full names in each (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Turnout effects are likely to be socially mediated - and therefore
+  show up in the neighborhoods where disenfranchised individuals live
+  (e.g.~Foladare 1968; Huckfeldt 1979; Cho, Gimpel, and Dyck 2006)
+\end{itemize}
+\end{frame}
 
-\hypertarget{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}{%
-\section{Effects of Felony Disenfranchisement on Neighborhood Turnout Levels}\label{effects-of-felony-disenfranchisement-on-neighborhood-turnout-levels}}
+\begin{frame}{Redefining Lost Voters}
+\protect\hypertarget{redefining-lost-voters}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Much of the existing literature cannot identify individuals who
+  \emph{would have voted} if not for their disenfranchisement (Burch
+  2013)
+\end{itemize}
 
-\hypertarget{identification-of-lost-voters}{%
-\subsection*{Identification of Lost Voters}\label{identification-of-lost-voters}}
-\addcontentsline{toc}{subsection}{Identification of Lost Voters}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  This undermines our ability to distinguish the causal effects of
+  incarceration / probation from the effects of disenfranchisement
+\end{itemize}
 
-In this analysis, I offer a different definition of ``lost voter'' than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- \emph{potentially} lost voters --- and turnout.
-Such an approach is informative for understanding the impact of disenfranchisement.
-Many young men, for instance, are admitted to prison each year.
-If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
-Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  This is especially likely given the low turnout propensity among the
+  formerly disenfranchised (e.g.~Meredith and Morse 2015; Gerber et al.
+  2017; White 2019)
+\end{itemize}
+\end{frame}
 
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the true effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration (e.g.~Gerber et al.
-\protect\hyperlink{ref-Gerber2017}{2017}).
-If incarcerated individuals would not have cast a ballot if they were not incarcerated, any effect of their incarceration on neighborhood turnout cannot be attributed to felony disenfranchisement \emph{per se} but rather to their incarceration.
-At the local level, where felony disenfranchisement rules do not vary from neighborhood to neighborhood, such a definition does not allow us to disentangle the depressive effect of having a neighbor go to prison from the effects of felony disenfranchisement.
+\begin{frame}{This Project Addresses Both Problems}
+\protect\hypertarget{this-project-addresses-both-problems}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Administrative data from the NYS DOC allows us to conduct
+  individual-level analysis
+\end{itemize}
 
-Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, ``lost voters'' are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote (Gerber, Green, and Shachar \protect\hyperlink{ref-Gerber2003}{2003}).
-As such, these are the individuals most likely to have been impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
-Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Purge records in the NYS voter file allow us to construct vote
+  histories for disenfranchised individuals
+\end{itemize}
 
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Lost voters are all formally disenfranchised individuals who have
+  voted in the past 10 years
+\end{itemize}
+\end{frame}
 
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were 2,518 such lost voters within New York City as of the 2017 general election, and 6,166 statewide.
+\begin{frame}{Identifying Neighborhoods with Lost Voters}
+\protect\hypertarget{identifying-neighborhoods-with-lost-voters}{}
+\includegraphics[width=1\linewidth]{../temp/lost_voters_map}
+\end{frame}
 
-\begin{figure}[H]
+\begin{frame}{Methodology}
+\protect\hypertarget{methodology}{}
+\begin{itemize}
+\item
+  A genetic match algorithm (Sekhon 2011) is used to match treated
+  census block groups to untreated ones.
+\item
+  Each block group is matched to 30 untreated block groups; matches are
+  done with replacement.
+\end{itemize}
+\end{frame}
 
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/citywide-map-1}
+\begin{frame}{Matching Results}
+\protect\hypertarget{matching-results}{}
+\begin{center}\includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_white_spsa} \includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_black_spsa} \end{center}
+\end{frame}
 
-}
+\begin{frame}{Matching Results}
+\protect\hypertarget{matching-results-1}{}
+\begin{center}\includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_latino_spsa} \includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/income_spsa} \end{center}
+\end{frame}
 
-\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
-\end{figure}
+\begin{frame}{Testing Turnout Effects}
+\protect\hypertarget{testing-turnout-effects}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot1} \end{center}
+\end{frame}
 
-The spatial concentration of lost voters is readily apparent.
-In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
+\begin{frame}{Testing Turnout Effects}
+\protect\hypertarget{testing-turnout-effects-1}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot2} \end{center}
+\end{frame}
 
-\hypertarget{testing-for-neighborhood-turnout-effects}{%
-\subsection*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
-\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Effects}
+\begin{frame}{Testing Turnout Effects}
+\protect\hypertarget{testing-turnout-effects-2}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot3b} \end{center}
+\end{frame}
 
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
+\begin{frame}{Testing Turnout Effects}
+\protect\hypertarget{testing-turnout-effects-3}{}
+\includegraphics[width=1\linewidth]{../output/dep_map}
+\end{frame}
 
-\begin{table}[H]
+\begin{frame}{Conclusions}
+\protect\hypertarget{conclusions}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  The indirect turnout effects identified in past research appear to be
+  geographically concentrated where the lost voters live.
+\end{itemize}
 
-\caption{\label{tab:distribution-lost}\label{tab:nhood-dist} Distribution of Lost Voters}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}{ccc}
-\toprule
-Number of Lost Voters & Count of Block Groups & Count of Tracts\\
-\midrule
-0 & 4,272 & 1,008\\
-1 & 1,023 & 416\\
-2 & 301 & 212\\
-3 & 107 & 138\\
-4 & 30 & 51\\
-5 & 19 & 46\\
-6 & 9 & 23\\
-7 & 3 & 15\\
-8 & 5 & 13\\
-9 & 0 & 7\\
-10 & 0 & 2\\
-11 & 0 & 2\\
-12 & 0 & 2\\
-14 & 0 & 2\\
-\hline
-Total & 5,769 & 1,937\\
-\bottomrule
-\end{tabular}
-\end{table}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Hajnal (2009) and others have demonstrated that turnout differentials
+  can have real political implications, especially in low-turnout
+  contests.
+\end{itemize}
 
-Because the majority of neighborhoods have either zero or one lost voter, I begin testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}), using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
-Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
+\begin{itemize}[<+->]
+\tightlist
+\item
+  Felony disenfranchisement undermines Black political representation at
+  the local level.
+\end{itemize}
+\end{frame}
 
-\hypertarget{match-output}{%
-\subsubsection*{Match Output}\label{match-output}}
-\addcontentsline{toc}{subsubsection}{Match Output}
+\begin{frame}{We Made It!}
+\protect\hypertarget{we-made-it}{}
+Thanks!
 
-\begin{table}[H]
+\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}}
+\end{frame}
 
-\caption{\label{tab:block-table-chunk}\label{tab:blockgroup-table} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.34 & 93.02 & 89.13 & 85.79 & 79.01\\
-\% Non-Hispanic Black & 0.38 & 0.16 & 0.38 & 0.37 & 97.93 & 94.46 & 94.41 & 92.77\\
-\% Non-Hispanic White & 0.17 & 0.40 & 0.17 & 0.18 & 96.37 & 96.06 & 95.56 & 92.99\\
-Median Income & 51,792.52 & 72,612.36 & 51,792.52 & 52,768.97 & 95.31 & 93.47 & 88.62 & 74.49\\
-\% With Some College & 0.64 & 0.70 & 0.64 & 0.64 & 90.75 & 94.14 & 90.29 & 80.70\\
-Median Age & 35.99 & 38.34 & 35.99 & 36.25 & 89.18 & 90.84 & 85.93 & 78.56\\
-Registration Rate & 1.00 & 0.94 & 1.00 & 0.98 & 59.77 & 79.89 & 72.28 & 60.42\\
-\% Democrats & 0.75 & 0.65 & 0.75 & 0.75 & 96.22 & 96.52 & 94.92 & 91.02\\
-\% Noncitizen & 0.16 & 0.16 & 0.16 & 0.16 & 16.58 & 14.80 & -0.06 & -60.28\\
-\% Won by City Council Representative & 0.85 & 0.81 & 0.85 & 0.86 & 89.44 & 83.40 & 79.88 & 70.11\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-table-chunk}\label{tab:tract-table} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.32 & 0.22 & 0.32 & 0.33 & 98.55 & 85.12 & 82.84 & 73.01\\
-\% Non-Hispanic Black & 0.36 & 0.14 & 0.36 & 0.36 & 99.77 & 88.74 & 88.36 & 83.30\\
-\% Non-Hispanic White & 0.19 & 0.42 & 0.19 & 0.19 & 97.50 & 96.32 & 91.29 & 79.95\\
-Median Income & 53,409.42 & 71,754.50 & 53,409.42 & 55,422.41 & 89.03 & 90.83 & 84.86 & 62.84\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 90.19 & 93.52 & 88.89 & 73.01\\
-Median Age & 35.76 & 38.04 & 35.76 & 35.80 & 98.12 & 74.51 & 80.56 & 71.91\\
-Registration Rate & 0.94 & 0.90 & 0.94 & 0.93 & 76.63 & 86.16 & 84.24 & 71.28\\
-\% Democrats & 0.74 & 0.62 & 0.74 & 0.73 & 94.58 & 94.93 & 92.05 & 82.37\\
-\% Noncitizen & 0.16 & 0.17 & 0.16 & 0.17 & -123.57 & 22.48 & 8.85 & -38.79\\
-\% Won by City Council Representative & 0.85 & 0.79 & 0.85 & 0.85 & 93.03 & 81.57 & 79.27 & 67.47\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-These tables also demonstrate the striking extent to which neighborhoods with lost voters differ from the average neighborhood in New York City.
-Neighborhoods with lost voters are far less White,\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} have much lower median incomes, and a larger share of voters are registered as Democrats.
-
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
-
-\input{"../temp/table1.tex"}
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average 0.9 percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as 2.7 percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just 14.6 percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
-
-\hypertarget{testing-intensity-effects}{%
-\subsubsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
-\addcontentsline{toc}{subsubsection}{Testing Intensity Effects}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on \emph{how many} voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
-In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
-Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
-
-\input{"../temp/table2.tex"}
-
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.87 percentage points.
-The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
-
-The block groups where these depressive effects are not randomly distributed throughout the city.
-They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on \(Lost\ Voters \times Share\ Black\) from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is \(-0.019\times Lost\ Voters \times Share\ Black\).
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/postest-map-1}
-
-}
-
-\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
-\end{figure}
-
-\hypertarget{toward-a-causal-estimation}{%
-\subsubsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
-\addcontentsline{toc}{subsubsection}{Toward a Causal Estimation}
-
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
-
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were \emph{newly} home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 \emph{did} have a lost voter in 2016, and are thus excluded from this analysis.
-
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
-
-Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
-As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
-
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.35 & 99.00 & 96.26 & 94.92 & 85.60\\
-\% Non-Hispanic Black & 0.33 & 0.16 & 0.33 & 0.33 & 99.02 & 96.57 & 95.67 & 90.10\\
-\% Non-Hispanic White & 0.21 & 0.41 & 0.21 & 0.21 & 99.61 & 94.66 & 92.61 & 82.94\\
-Median Income & 55,852.62 & 73,433.49 & 55,852.62 & 56,287.76 & 97.52 & 87.85 & 86.56 & 72.89\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 99.31 & 92.25 & 87.30 & 71.15\\
-Median Age & 36.17 & 38.47 & 36.17 & 36.04 & 94.37 & 62.35 & 72.27 & 69.84\\
-Registration Rate & 0.98 & 0.94 & 0.98 & 0.95 & 25.09 & 46.49 & 42.16 & 33.06\\
-\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 95.94 & 94.99 & 93.62 & 86.68\\
-\% Noncitizen & 0.17 & 0.16 & 0.17 & 0.17 & 63.46 & 75.01 & 66.85 & 45.71\\
-\% Won by City Council Representative & 0.85 & 0.80 & 0.85 & 0.85 & 88.19 & 83.99 & 78.23 & 67.54\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-chunk-did}\label{tab:did-match-tract} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 94.34 & 86.10 & 84.77 & 75.82\\
-\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.25 & 99.28 & 89.75 & 86.62 & 77.90\\
-\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.02 & 87.45 & 77.73 & 61.67\\
-Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,973.89 & 84.23 & 78.35 & 76.98 & 66.57\\
-\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 96.86 & 86.51 & 83.73 & 76.55\\
-Median Age & 36.36 & 38.23 & 36.36 & 36.39 & 98.60 & 73.15 & 69.00 & 42.11\\
-Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 10.68 & 64.27 & 63.46 & 56.35\\
-\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.23 & 90.06 & 87.35 & 76.31\\
-\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 78.09 & 73.55 & 62.03 & 48.36\\
-\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.91 & 75.59 & 65.98 & 38.91\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-In Table \ref{tab:did-reg}, I present the out come of a simple least squares regression on this subset of block groups and tracts.
-In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly \emph{higher} turnout prior to intervention.
-D(Treat) \(\times\) D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
-
-\input{"../temp/table1111.tex"}
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated 2,518 voters from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
-However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
-
-As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement impacts Black turnout at the state level, but this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
-Similarly, just 17 percent of individuals in the average block group with a lost voter were White, compared with 40 percent of the population in other block groups.
-Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
-The effects are concentrated in neighborhoods where the most marginalized members of society live.
-These highly concentrated spillover effects are cause for major concern.
-
-Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
-This means that lost voters have no spillover effects in neighborhoods with small Black populations.
-Understanding the magnitude of this finding is important.
-In New York City, there are 165 block groups where the Black community makes up more than 90 percent of the population, with an average of 821 voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 90 percent Black with a citizen voting age population of 821, having any lost voter reduced turnout by 20 ballots.
-Conversely, in a block group with the same voting age population that is just 10 percent Black, the loss of any voter decreased turnout by just 2.2 ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 90 percent Black, each lost voter cost the neighborhood 13.9 ballots.
-In block groups of the same size where the population is just 10 percent Black, lost voters cost the neighborhood just 1.5 ballots --- hardly more votes than the lost voter herself.
-
-Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large ``interpretive effects'' (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Lerman and Weaver (\protect\hyperlink{ref-Lerman2013}{2013}), for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, Weaver and Lerman (\protect\hyperlink{ref-Weaver2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
-
-Individuals who live in neighborhoods where police activity is relatively limited may interpret the incarceration of a neighbor as a largely individual phenomenon.
-If it is understood as an isolated or individual event, voters in these neighborhoods who are not incarcerated are not likely to update their view of the state.
-They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
-In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
-It may, rather, be interpreted as another reminder of the government's unfairness.
-If a would-be voter finds herself soured on political participation because of her neighbor's incarceration, she may be less likely to cast a ballot.
-
-This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
-This is perhaps unsurprising.
-A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans (e.g.~Browning and Cao \protect\hyperlink{ref-Browning1992}{1992}; Hurwitz and Peffley \protect\hyperlink{ref-Hurwitz2005}{2005}; Henderson et al.
-\protect\hyperlink{ref-Henderson1997}{1997}; Wu, Sun, and Triplett \protect\hyperlink{ref-Wu2009}{2009}).
-Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
-There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
-
-For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods (Wilson \protect\hyperlink{ref-Wilson1990}{1990}).
-More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
-Much of this work, however, has focused on either the impacts of living in marginalized communities (by measuring health and economic impacts) or has not accounted for the importance of physical space (by focusing only on proximal social, and not geographic, contact with directly impacted individuals).
-This analysis allows us to understand the implications for living in a neighborhood where individuals likely to cast a ballot are not allowed to because of a felony conviction.
-I find that neighborhoods that are home to lost voters --- and particularly neighborhoods with large Black populations --- systematically turn out for local elections at lower rates than otherwise similar neighborhoods.
-
-This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
-In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
-As Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) shows, racial turnout differentials can have real consequences for city politics.
-
-Moreover, we cannot conclude that depressed turnout in these neighborhoods in 2017 had no impact on their representation.
-City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
-Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals \emph{held} and \emph{used} their power.
-
-Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
-As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
-
-\pagebreak
-
-\hypertarget{executive-order-181-and-turnout-in-2018}{%
-\section{Executive Order 181 and Turnout in 2018}\label{executive-order-181-and-turnout-in-2018}}
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for the communities in which parolees live; as discussed above, the disenfranchisement of voters has large spillover effects in the neighborhoods in which these lost voters are concentrated.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-Of course, re-extending the right to vote is only the first step.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters (see, for instance, Fraga \protect\hyperlink{ref-Fraga2018}{2018}).
-To address \emph{only} the formal laws contributing to disenfranchisement without also interrogating the efficacy of any policy change to boost participation risks leaving much of the system of effective disenfranchisement undisturbed.
-
-There is reason to believe that the executive order may have increased the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
-Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018.
-This marked the registration deadline for the 2018 general election, in which New Yorkers cast ballots in both gubernatorial and U.S.
-senatorial contests (as well as other down-ballot races).
-
-I focus only on the individuals who would have been eligible to cast a ballot whether or not the executive order was signed.
-This is, of course, only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on Election Day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
-
-\hypertarget{identifying-parolees-whose-rights-were-restored}{%
-\subsection*{Identifying Parolees Whose Rights Were Restored}\label{identifying-parolees-whose-rights-were-restored}}
-\addcontentsline{toc}{subsection}{Identifying Parolees Whose Rights Were Restored}
-
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.}
-
-\hypertarget{trends-in-turnout}{%
-\subsection*{Trends in Turnout}\label{trends-in-turnout}}
-\addcontentsline{toc}{subsection}{Trends in Turnout}
-
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.6 percent cast a ballot in the 2016 election, while just 4.2 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-16-chart-1}
-
-}
-
-\caption{\label{fig:to-16}Turnout in 2016 Presidential Election}\label{fig:to-16-chart}
-\end{figure}
-
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-chart-1}
-
-}
-
-\caption{\label{fig:to-18}Turnout in 2018 Midterm Election}\label{fig:to-18-chart}
-\end{figure}
-
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- \emph{were} more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
-
-\hypertarget{individual-level-turnout-regressions}{%
-\subsection*{Individual-Level Turnout Regressions}\label{individual-level-turnout-regressions}}
-\addcontentsline{toc}{subsection}{Individual-Level Turnout Regressions}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being ``assigned'' to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated when we limit the analysis to individuals discharged in 2017 or 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/to-18-recent-chart-1}
-
-}
-
-\caption{\label{fig:to-18-recent}Turnout in 2018 Midterm Election}\label{fig:to-18-recent-chart}
-\end{figure}
-
-Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
-
-\input{"../temp/table3.tex"}
-
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} increases the AIC.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.}
-
-Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-Figure \ref{fig:restoration-chart} plots the share of individuals discharged from parole each day in the spring of 2018.
-It is not until May 18\textsuperscript{th} that parolees consistently had their voting rights restored prior to discharge.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{felony_disenfranchisement_nys_files/figure-latex/restoration-chart-chunk-1}
-
-}
-
-\caption{\label{fig:restoration-chart}Share of Discharged Parolees Whose Voting Rights Were Restored Prior to Discharge}\label{fig:restoration-chart-chunk}
-\end{figure}
-
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are ``treated'' by the policy change.
-Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is biased against finding a significant increase due to the executive order.
-
-\input{"../temp/table_whatever2.tex"}
-
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18\textsuperscript{th}, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1\textsuperscript{st}, 2017, through October 10\textsuperscript{th}, 2018 (the registration deadline in New York State).
-
-\input{"../temp/table4.tex"}
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 31.2 and 31.8 percent.
-
-\hypertarget{instrumental-variables-approach}{%
-\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
-\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
-
-Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
-
-An individual's propensity to vote can be expressed using the following equation:
-\[ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} \]
-Where \emph{Y\textsubscript{i}} is 1 if individual \emph{i} cast a ballot, \emph{X\textsubscript{1i}} is the probability that individual \emph{i} was eligible to have his rights restored prior to discharge from parole, and \emph{X\textsubscript{2i}} is 1 if individual \emph{i} actually had his rights restored prior to discharge from parole.
-\emph{Z} is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe \emph{X\textsubscript{1i}}, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if \emph{X\textsubscript{1i}} and \emph{X\textsubscript{2i}} (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if \emph{b\textsubscript{1}} is uncorrelated with an individual's propensity to vote (\emph{Y\textsubscript{i}}).
-
-Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
-Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Although the available data does not indicate whether individuals were eligible to have their rights restored, we can examine the demographic differences between individuals who had their rights restored and those who did not.
-
-\input{"../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that \emph{b\textsubscript{1}} is very probably correlated with \emph{Y}.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration \emph{caused} turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
-Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research (e.g.~Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
-The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2007}{2007}).
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
-\input{"../temp/iv_clean.tex"}
-
-The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.74 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.85 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.0 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the Executive Order increased turnout by between 32.7 and 39.5 percent.
-
-\hypertarget{racial-variation}{%
-\subsubsection*{Racial Variation}\label{racial-variation}}
-\addcontentsline{toc}{subsubsection}{Racial Variation}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only White former parolees (Models 1 and 4), all non-White former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
-
-\input{"../temp/iv_clean_race.tex"}
-
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among White individuals by around 1.6 percentage points.
-Given that just 3.3 percent of White former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{nearly doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-White individuals: the coefficient on these estimates are small and not statistically significant.
-
-Why would the intervention have increased turnout among White individuals and have had no effect on Black individuals?
-Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than White individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many Black defendants into court, including some voters (so they can be deterred), while lower arrest rates among Whites mean that the White defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
-
-\input{"../temp/table5.tex"}
-
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still 36 percent more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
-
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
-The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
-
-\hypertarget{variable-treatment-intensity}{%
-\subsubsection*{Variable Treatment Intensity}\label{variable-treatment-intensity}}
-\addcontentsline{toc}{subsubsection}{Variable Treatment Intensity}
-
-There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
-This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
-If the individual did not have voting rights restored, this variable takes the value 0.
-This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.\footnote{Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.}
-
-\input{"../temp/iv_clean_months.tex"}
-
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.25 percentage points (in the 2SLS model) and 0.48 percentage points (in the 2S probit model).
-It is worth noting that the \emph{R\textsuperscript{2}} in Model 1 is higher than the \emph{R\textsuperscript{2}} in Model 2 in Table \ref{tab:iv-models}, indicating that allowing for a time-variant treatment effect is warranted.
-
-Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
-
-\hypertarget{discussion-1}{%
-\subsection*{Discussion}\label{discussion-1}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
-Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
-That changed in 2018.
-On October 12\textsuperscript{th} --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
-Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just 792, or 3.62 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
-
-As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least 32.7 percent.
-This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
-Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
-
-The mechanism through which the rules change increased turnout among these individuals is not clear.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
-Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
-
-It could also be a story of better information.
-As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
-Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
-In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
-
-Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
-Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
-The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
-This discrepancy could be caused by a number of different factors.
-Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
-The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
-
-We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
-Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
-Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-WHITE2014}{2014}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2016}{2016}).
-
-Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
-This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
-
-Re-enfranchising voters while they are still under formal supervision is obviously beneficial to the individuals who are on parole on election day; such policies allow them to make their voices heard.
-The case of Executive Order 181 also indicates that restoring voting rights prior to parole discharge has further benefits.
-In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
-This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
-By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-Felony disenfranchisement was originally implemented to limit the franchise of Black Americans.
-It continues to serve that purpose today, as Black residents of New York and states around the country are disproportionately convicted of felony offenses and imprisoned.
-Although we know that many of these individuals were unlikely to cast a ballot even if they had not gone to prison, the targeted disenfranchisement of a single segment of the population is deeply troubling.
-
-As this project has demonstrated, however, the racialized penalties of felony disenfranchisement do not disappear after we control for low propensity to vote among individuals in New York City and State who are going to prison.
-The first half of this project attempted to refine how we think about the spatial aspects of felony disenfranchisement, and also how we identify the individuals most likely to be directly impacted by these policies.
-By both redefining ``lost voters'' as individuals who have a demonstrated propensity to participate and examining turnout in the physical neighborhoods they left behind, I offer a more presise accounting of the impact of these policies.
-What I find is troubling: neighborhoods that were home to disenfranchised individuals turn out at systematically lower rates, and this effect is concentrated in Black neighborhoods.
-Moreover, there is reason to believe that this is not driven by unobservable differences between neighborhoods; I also show that turnout decreased more between 2016 and 2017 in neighborhoods that newly lost voters than in neighborhoods unaffected by disenfranchisement policies.
-
-I also offer an early investigation of the effect of Executive Order 181, which granted voting rights to most citizens on parole in New York State.
-This research indicates that individuals who have their rights restored prior to being discharged from parole are more likely to participate in elections, even after they are discharged.
-This is a welcome finding, as increasing the political participation of marginalized individuals brings us toward a fuller democracy.
-As with neighborhood spillover effects, however, this effect depends in great deal on the race of the parolee: rights restoration, at least in the first five months of the program in New York State, did not increase future participation among non-Whites.
-More research needs to be done to investigate whether this is true over a longer time period in New York, and whether it is true in other states as well.
-
-America is becoming ever-more-diverse, and will likely be less than 50 percent non-Hispanic White by the middle of the century.
-As Fraga (\protect\hyperlink{ref-Fraga2018}{2018}) has documented, however, this does not necessarily mean that minorities will enjoy political power in line with their share of the population.
-How districts are drawn, population patterns, and rules about who can fundraise how will continue to shape the politicians we elect and the policies they enact.
-Breaking down \emph{de jure} (like formal disenfranchisement) and \emph{de facto} (like neighborhood spillover effects) barriers to the ballot box is of great importance.
-The experience in New York State over the past few years offers a better understanding of just how big this problem is, but it also offers evidence of a way forward.
-
-\newpage
-
-\hypertarget{appendix-a}{%
-\section*{Appendix A}\label{appendix-a}}
-\addcontentsline{toc}{section}{Appendix A}
-
-In the \protect\hyperlink{instrumental-variables-approach}{Instrumental Variables Approach} section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1\textsuperscript{st}, 2015, and October 14\textsuperscript{th}, 2016.
-
-\input{"../temp/table55.tex"}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\noindent
-
+\begin{frame}[allowframebreaks,allowframebreaks]{References}
+\protect\hypertarget{references}{}
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abadie2019}{}%
-Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \url{https://scholar.harvard.edu/spiess/publications/robust-post-matching-inference.}
-
-\leavevmode\hypertarget{ref-Alexander2012}{}%
-Alexander, Michelle. 2012. \emph{The New Jim Crow}. New York: The New Press.
-
-\leavevmode\hypertarget{ref-Angrist2001}{}%
-Angrist, Joshua D.
-2001.
-``Estimation of Limited Dependent Variable Models with Dummy Endogenous Regressors.'' \emph{Journal of Business \& Economic Statistics} 19 (1): 2--28.
-\url{https://doi.org/10.1198/07350010152472571}.
-
-\leavevmode\hypertarget{ref-Angrist1996}{}%
-Angrist, Joshua D., Guido W.
-Imbens, and Donald B.
-Rubin.
-1996.
-``Identification of Causal Effects Using Instrumental Variables.'' \emph{Journal of the American Statistical Association} 91 (434): 444--55.
-\url{https://doi.org/10.1080/01621459.1996.10476902}.
-
-\leavevmode\hypertarget{ref-Angrist2008}{}%
-Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
-
-\leavevmode\hypertarget{ref-Ansolabehere1999}{}%
-Ansolabehere, Stephen D., Shanto Iyengar, and Adam Simon.
-1999.
-``Replicating Experiments Using Aggregate and Survey Data: The Case of Negative Advertising and Turnout.'' \emph{American Political Science Review} 93 (4): 901--9.
-\url{https://doi.org/10.2307/2586120}.
-
-\leavevmode\hypertarget{ref-Anzia2019}{}%
-Anzia, Sarah F.
-2019.
-``When Does a Group of Citizens Influence Policy?
-Evidence from Senior Citizen Participation in City Politics.'' \emph{The Journal of Politics} 81 (1): 1--14.
-\url{https://doi.org/10.1086/699916}.
-
-\leavevmode\hypertarget{ref-Behrens2003}{}%
-Behrens, Angela, Christopher Uggen, and Jeff Manza.
-2003.
-``Ballot Manipulation and the ``Menace of Negro Domination'': Racial Threat and Felon Disenfranchisement in the United States, 18502002.'' \emph{American Journal of Sociology} 109 (3): 559--605.
-\url{https://doi.org/10.1086/378647}.
-
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Bowers2009}{}%
-Bowers, Melanie, and Robert R.
-Preuhs.
-2009.
-``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons.'' \emph{Social Science Quarterly} 90 (3): 722--43.
+Bowers, Melanie, and Robert R. Preuhs. 2009. ``Collateral Consequences
+of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement
+Laws on the Political Participation of Nonfelons.'' \emph{Social Science
+Quarterly} 90 (3): 722--43.
 \url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
 
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
-
-\leavevmode\hypertarget{ref-Browning1992}{}%
-Browning, Sandra Lee, and Liqun Cao. 1992. ``The Impact of Race on Criminal Justice Ideology.'' \emph{Justice Quarterly} 9 (4): 685--701. \url{https://doi.org/10.1080/07418829200091611}.
-
-\leavevmode\hypertarget{ref-Burch2010}{}%
-Burch, Traci.
-2010.
-``Did Disfranchisement Laws Help Elect President Bush?
-New Evidence on the Turnout Rates and Candidate Preferences of Florida's Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
-\url{https://doi.org/10.1007/s11109-010-9150-9}.
-
-\leavevmode\hypertarget{ref-Burch2011}{}%
----------.
-2011.
-``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
-\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
-
 \leavevmode\hypertarget{ref-Burch2013}{}%
----------.
-2013.
-``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 184--201.
-\url{https://doi.org/10.1177/0002716213503093}.
-
-\leavevmode\hypertarget{ref-Butler2011}{}%
-Butler, Daniel M., and David E.
-Broockman.
-2011.
-``Do Politicians Racially Discriminate Against Constituents?
-A Field Experiment on State Legislators.'' \emph{American Journal of Political Science} 55 (3): 463--77.
-\url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
+Burch, Traci. 2013. ``Effects of Imprisonment and Community Supervision
+on Neighborhood Political Participation in North Carolina.'' Edited by
+Christopher Wildeman, Jacob S. Hacker, and Vesla M. Weaver. \emph{The
+ANNALS of the American Academy of Political and Social Science} 651 (1):
+184--201. \url{https://doi.org/10.1177/0002716213503093}.
 
 \leavevmode\hypertarget{ref-Cho2006}{}%
-Cho, Wendy K.
-Tam, James G.
-Gimpel, and Joshua J.
-Dyck.
-2006.
-``Residential Concentration, Political Socialization, and Voter Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
+Cho, Wendy K. Tam, James G. Gimpel, and Joshua J. Dyck. 2006.
+``Residential Concentration, Political Socialization, and Voter
+Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
 \url{https://doi.org/10.1111/j.1468-2508.2006.00377.x}.
 
-\leavevmode\hypertarget{ref-Clear2008}{}%
-Clear, Todd R. 2008. ``The Effects of High Imprisonment Rates on Communities.'' \emph{Crime and Justice} 37 (1): 97--132. \url{https://doi.org/10.1086/522360}.
-
-\leavevmode\hypertarget{ref-Dawkins2006}{}%
-Dawkins, Casey J. 2006. ``Are Social Networks the Ties That Bind Families to Neighborhoods?'' \emph{Housing Studies} 21 (6): 867--81. \url{https://doi.org/10.1080/02673030600917776}.
-
-\leavevmode\hypertarget{ref-Dong2014}{}%
-Dong, Yingying, and Arthur Lewbel.
-2014.
-``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
-\url{https://doi.org/10.1080/07474938.2014.944470}.
-
-\leavevmode\hypertarget{ref-Drucker2005}{}%
-Drucker, Ernest, and Ricardo Barreras.
-2005.
-``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
-Sentencing Project.
-\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
-
-\leavevmode\hypertarget{ref-Einstein2016}{}%
-Einstein, Katherine Levine, and David M.
-Glick.
-2016.
-``Does Race Affect Access to Government Services?
-An Experiment Exploring Street-Level Bureaucrats and Access to Public Housing.'' \emph{American Journal of Political Science} 61 (1): 100--116.
-\url{https://doi.org/10.1111/ajps.12252}.
-
 \leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
-
-\leavevmode\hypertarget{ref-Fraga2018}{}%
-Fraga, Bernard L. 2018. \emph{The Turnout Gap}. Cambridge, United Kingdom: Cambridge University Press. \url{https://doi.org/10.1017/9781108566483}.
-
-\leavevmode\hypertarget{ref-Geller2014}{}%
-Geller, Amanda, Jeffrey Fagan, Tom Tyler, and Bruce G.
-Link.
-2014.
-``Aggressive Policing and the Mental Health of Young Urban Men.'' \emph{American Journal of Public Health} 104 (12): 2321--7.
-\url{https://doi.org/10.2105/ajph.2014.302046}.
+Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting
+Behavior.'' \emph{Political Science Quarterly} 83 (4): 516.
+\url{https://doi.org/10.2307/2146812}.
 
 \leavevmode\hypertarget{ref-Gelman2007}{}%
-Gelman, Andrew, Jeffrey Fagan, and Alex Kiss.
-2007.
-``An Analysis of the New York City Police Departments ``Stop-and-Frisk'' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
+Gelman, Andrew, Jeffrey Fagan, and Alex Kiss. 2007. ``An Analysis of the
+New York City Police Departments ``Stop-and-Frisk'' Policy in the
+Context of Claims of Racial Bias.'' \emph{Journal of the American
+Statistical Association} 102 (479): 813--23.
 \url{https://doi.org/10.1198/016214506000001040}.
 
-\leavevmode\hypertarget{ref-Gerber2000}{}%
-Gerber, Alan S., and Donald P.
-Green.
-2000.
-``The Effects of Canvassing, Telephone Calls, and Direct Mail on Voter Turnout: A Field Experiment.'' \emph{American Political Science Review} 94 (3): 653--63.
-\url{https://doi.org/10.2307/2585837}.
-
-\leavevmode\hypertarget{ref-Gerber2003}{}%
-Gerber, Alan S., Donald P.
-Green, and Ron Shachar.
-2003.
-``Voting May Be Habit-Forming: Evidence from a Randomized Field Experiment.'' \emph{American Journal of Political Science} 47 (3): 540--50.
-\url{https://doi.org/10.1111/1540-5907.00038}.
-
-\leavevmode\hypertarget{ref-Gerber2014}{}%
-Gerber, Alan S., Gregory A.
-Huber, Marc Meredith, Daniel R.
-Biggers, and David J.
-Hendry.
-2014.
-``Can Incarcerated Felons Be (Re)integrated into the Political System?
-Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
-\url{https://doi.org/10.1111/ajps.12166}.
-
 \leavevmode\hypertarget{ref-Gerber2017}{}%
----------.
-2017.
-``Does Incarceration Reduce Voting?
-Evidence About the Political Consequences of Spending Time in Prison.'' \emph{The Journal of Politics} 79 (4): 1130--46.
+Gerber, Alan S., Gregory A. Huber, Marc Meredith, Daniel R. Biggers, and
+David J. Hendry. 2017. ``Does Incarceration Reduce Voting? Evidence
+About the Political Consequences of Spending Time in Prison.'' \emph{The
+Journal of Politics} 79 (4): 1130--46.
 \url{https://doi.org/10.1086/692670}.
 
-\leavevmode\hypertarget{ref-Gimpel2004}{}%
-Gimpel, James G., Joshua J.
-Dyck, and Daron R.
-Shaw.
-2004.
-``Registrants, Voters, and Turnout Variability Across Neighborhoods.'' \emph{Political Behavior} 26 (4): 343--75.
-\url{https://doi.org/10.1007/s11109-004-0900-4}.
-
-\leavevmode\hypertarget{ref-Green2000}{}%
-Green, Donald P., and Ron Shachar.
-2000.
-``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
-\url{https://doi.org/10.1017/s0007123400000247}.
-
-\leavevmode\hypertarget{ref-Griffin2005}{}%
-Griffin, John D., and Brian Newman. 2005. ``Are Voters Better Represented?'' \emph{The Journal of Politics} 67 (4): 1206--27. \url{https://doi.org/10.1111/j.1468-2508.2005.00357.x}.
-
-\leavevmode\hypertarget{ref-Guest1999}{}%
-Guest, Avery M., and Susan K. Wierzbicki. 1999. ``Social Ties at the Neighborhood Level.'' \emph{Urban Affairs Review} 35 (1): 92--111. \url{https://doi.org/10.1177/10780879922184301}.
-
 \leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
-
-\leavevmode\hypertarget{ref-Hajnal2005}{}%
-Hajnal, Zoltan, and Jessica Trounstine.
-2005.
-``Where Turnout Matters: The Consequences of Uneven Turnout in City Politics.'' \emph{The Journal of Politics} 67 (2): 515--35.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00327.x}.
-
-\leavevmode\hypertarget{ref-Haselswerdt2009}{}%
-Haselswerdt, Michael V.
-2009.
-``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data.'' \emph{Social Science Quarterly} 90 (2): 262--73.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
-
-\leavevmode\hypertarget{ref-Hemmens2000}{}%
-Hemmens, Craig, and James W.
-Marquart.
-2000.
-``Friend or Foe?
-Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
-\url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
-
-\leavevmode\hypertarget{ref-Henderson1997}{}%
-Henderson, Martha L., Francis T.
-Cullen, Liqun Cao, Sandra Lee Browning, and Renee Kopache.
-1997.
-``The Impact of Race on Perceptions of Criminal Injustice.'' \emph{Journal of Criminal Justice} 25 (6): 447--62.
-\url{https://doi.org/10.1016/s0047-2352(97)00032-9}.
-
-\leavevmode\hypertarget{ref-Hjalmarsson2010}{}%
-Hjalmarsson, Randi, and Mark Lopez.
-2010.
-``The Voting Behavior of Young Disenfranchised Felons: Would They Vote If They Could?'' \emph{American Law and Economics Review} 12 (2): 356--93.
-\url{https://doi.org/10.1093/aler/ahq004}.
+Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge
+University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
 
 \leavevmode\hypertarget{ref-Huckfeldt1979}{}%
-Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
-
-\leavevmode\hypertarget{ref-Hurwitz2005}{}%
-Hurwitz, Jon, and Mark Peffley.
-2005.
-``Explaining the Great Racial Divide: Perceptions of Fairness in the U.s.
-Criminal Justice System.'' \emph{The Journal of Politics} 67 (3): 762--83.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00338.x}.
-
-\leavevmode\hypertarget{ref-Kenny1992}{}%
-Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259. \url{https://doi.org/10.2307/2111432}.
-
-\leavevmode\hypertarget{ref-Keyssar2009}{}%
-Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. New York: Basic Books.
+Huckfeldt, R. Robert. 1979. ``Political Participation and the
+Neighborhood Social Context.'' \emph{American Journal of Political
+Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
 
 \leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821. \url{https://doi.org/10.1177/0021934716659195}.
-
-\leavevmode\hypertarget{ref-Kinsella2015}{}%
-Kinsella, Chad, Colleen McTague, and Kevin N.
-Raleigh.
-2015.
-``Unmasking Geographic Polarization and Clustering: A Micro-Scalar Analysis of Partisan Voting Behavior.'' \emph{Applied Geography} 62 (August): 404--19.
-\url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
-
-\leavevmode\hypertarget{ref-Lassen2004}{}%
-Lassen, David Dreyer. 2004. ``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' \emph{SSRN Electronic Journal}. \url{https://doi.org/10.2139/ssrn.475821}.
-
-\leavevmode\hypertarget{ref-Lerman2013}{}%
-Lerman, Amy E., and Vesla Weaver.
-2013.
-``Staying Out of Sight?
-Concentrated Policing and Local Political Action.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 202--19.
-\url{https://doi.org/10.1177/0002716213503085}.
-
-\leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
-
-\leavevmode\hypertarget{ref-Martin2003}{}%
-Martin, Paul S.
-2003.
-``Votings Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
-\url{https://doi.org/10.1111/1540-5907.00008}.
-
-\leavevmode\hypertarget{ref-Martin2013}{}%
-Martin, Paul S., and Michele P.
-Claibourn.
-2013.
-``Citizen Participation and Congressional Responsiveness: New Evidence That Participation Matters.'' \emph{Legislative Studies Quarterly} 38 (1): 59--81.
-\url{https://doi.org/10.1111/lsq.12003}.
-
-\leavevmode\hypertarget{ref-Meredith2013}{}%
-Meredith, Marc, and Michael Morse.
-2013.
-``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 220--49.
-\url{https://doi.org/10.1177/0002716213502931}.
+King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the
+Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821.
+\url{https://doi.org/10.1177/0021934716659195}.
 
 \leavevmode\hypertarget{ref-Meredith2015}{}%
----------.
-2015.
-``The Politics of the Restoration of Ex-Felon Voting Rights: The Case of Iowa.'' \emph{Quarterly Journal of Political Science} 10 (1): 41--100.
+Meredith, Marc, and Michael Morse. 2015. ``The Politics of the
+Restoration of Ex-Felon Voting Rights: The Case of Iowa.''
+\emph{Quarterly Journal of Political Science} 10 (1): 41--100.
 \url{https://doi.org/10.1561/100.00013026}.
@@
 \leavevmode\hypertarget{ref-Miles2004}{}%
-Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.'' \emph{The Journal of Legal Studies} 33 (1): 85--129. \url{https://doi.org/10.1086/381290}.
-
-\leavevmode\hypertarget{ref-Milligan2004}{}%
-Milligan, Kevin, Enrico Moretti, and Philip Oreopoulos.
-2004.
-``Does Education Improve Citizenship?
-Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9-10): 1667--95.
-\url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
-
-\leavevmode\hypertarget{ref-Morris2019}{}%
-Morris, Kevin. 2019. ``Thwarting Amendment 4.'' May 9, 2019. \url{https://www.brennancenter.org/analysis/thwarting-amendment-4}.
-
-\leavevmode\hypertarget{ref-Mutz2002}{}%
-Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838. \url{https://doi.org/10.2307/3088437}.
-
-\leavevmode\hypertarget{ref-Ochs2006}{}%
-Ochs, Holona Leanne.
-2006.
-``'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
-\url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
-
-\leavevmode\hypertarget{ref-Pierson1993}{}%
-Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
+Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.''
+\emph{The Journal of Legal Studies} 33 (1): 85--129.
+\url{https://doi.org/10.1086/381290}.
 
 \leavevmode\hypertarget{ref-Sekhon2011}{}%
-Sekhon, Jasjeet S.
-2011.
-``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (7).
+Sekhon, Jasjeet S. 2011. ``Multivariate and Propensity Score Matching
+Software with Automated Balance Optimization: The Matching Package for
+R.'' \emph{Journal of Statistical Software} 42 (7).
 \url{https://doi.org/10.18637/jss.v042.i07}.
 
-\leavevmode\hypertarget{ref-Sewell2016}{}%
-Sewell, Abigail A., and Kevin A.
-Jefferson.
-2016.
-``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (S1): 42--67.
-\url{https://doi.org/10.1007/s11524-015-0016-7}.
-
-\leavevmode\hypertarget{ref-florida_1868}{}%
-Shofner, Jerrell H. 1963. ``The Constitution of 1868.'' \emph{The Florida Historical Quarterly} 41 (4): 356--74. \url{http://www.jstor.org/stable/30139965}.
-
-\leavevmode\hypertarget{ref-Sondheimer2010}{}%
-Sondheimer, Rachel Milstein, and Donald P.
-Green.
-2010.
-``Using Experiments to Estimate the Effects of Education on Voter Turnout.'' \emph{American Journal of Political Science} 54 (1): 174--89.
-\url{https://doi.org/10.1111/j.1540-5907.2009.00425.x}.
-
-\leavevmode\hypertarget{ref-Terza2007}{}%
-Terza, Joseph V., W.
-David Bradford, and Clara E.
-Dismuke.
-2007.
-``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
-\url{https://doi.org/10.1111/j.1475-6773.2007.00807.x}.
-
-\leavevmode\hypertarget{ref-sentencing_2016}{}%
-Uggen, Christopher, Ryan Larson, and Sarah Shannon.
-2016.
-``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-Sentencing Project.
-\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
-
-\leavevmode\hypertarget{ref-Uggen2002}{}%
-Uggen, Christopher, and Jeff Manza.
-2002.
-``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
-\url{https://doi.org/10.2307/3088970}.
-
-\leavevmode\hypertarget{ref-Uggen2004}{}%
----------.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
-
-\leavevmode\hypertarget{ref-Walker2014}{}%
-Walker, Hannah L.
-2014.
-``Extending the Effects of the Carceral State.
-Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly} 67 (4): 809--22.
-\url{https://doi.org/10.1177/1065912914542522}.
-
-\leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela Garcı́a-Castañon.
-2017.
-``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (04): 541--68.
-\url{https://doi.org/10.1017/s1743923x17000198}.
-
-\leavevmode\hypertarget{ref-Weaver2014}{}%
-Weaver, Vesla, and Amy E. Lerman. 2014. \emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}. Chicago: University of Chicago Press.
-
-\leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
-
 \leavevmode\hypertarget{ref-White2019}{}%
-White, Ariel.
-2019.
-``Misdemeanor Disenfranchisement?
-The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
+White, Ariel. 2019. ``Misdemeanor Disenfranchisement? The Demobilizing
+Effects of Brief Jail Spells on Potential Voters.'' \emph{American
+Political Science Review} 113 (2): 311--24.
 \url{https://doi.org/10.1017/s000305541800093x}.
-
-\leavevmode\hypertarget{ref-WHITE2014}{}%
-White, Ariel R., Noah L.
-Nathan, and Julie K.
-Faller.
-2014.
-``What Do I Need to Vote?
-Bureaucratic Discretion and Discrimination by Local Election Officials.'' \emph{American Political Science Review} 109 (1): 129--42.
-\url{https://doi.org/10.1017/s0003055414000562}.
-
-\leavevmode\hypertarget{ref-Wilson1990}{}%
-Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
-
-\leavevmode\hypertarget{ref-Wooldridge2010}{}%
-Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
-
-\leavevmode\hypertarget{ref-Wu2009}{}%
-Wu, Yuning, Ivan Y.
-Sun, and Ruth A.
-Triplett.
-2009.
-``Race, Class or Neighborhood Context: Which Matters More in Measuring Satisfaction with Police?'' \emph{Justice Quarterly} 26 (1): 125--56.
-\url{https://doi.org/10.1080/07418820802119950}.
+\end{cslreferences}
+\end{frame}
 

### 2020-01-28 code/00_initialize.R

@@ -1,4 +1,5 @@
 library(AER)
+library(jtools)
 library(lubridate)
 library(RStata)
 library(estimatr)
@@ -14,8 +15,33 @@ library(scales)
 library(kableExtra)
 library(stargazer)
 
+
+
 ## QUICK FUNCTION TO CLEAR THE ENVIRONMENT OF MOST OBJECTS AFTER EACH STEP
 ## DOING SO TO WORK AROUND MEMORY CONSTRAINTS
+
+
+theme_bc <- function(){
+  library(ggplot2)
+  library(extrafont)
+  theme(plot.title = element_text(family = "Roboto Black", 
+                                  color = "black", hjust = 0.5, margin = margin(0, 0, 10, 0)),
+        axis.title = element_text(family = "Roboto", 
+                                  color = "#525353"),
+        axis.line = element_line(colour = "black"), 
+        text = element_text(family = "Roboto", color = "#525353"), 
+        plot.caption = element_text(family = "Roboto", size = 8),
+        panel.border = element_blank(),
+        panel.grid.major = element_blank(), 
+        panel.background = element_rect(fill = "#dddddb"),
+        plot.background = element_rect(fill = "#dddddb"), 
+        panel.grid.minor = element_blank(),
+        legend.key = element_blank(), 
+        legend.background = element_rect(color = "black", fill = "#dddddb"), 
+        plot.subtitle = element_text(hjust = 0.5, size = 10), 
+        plot.margin = margin(10, 10, 10, 10))
+}
+
 save <- c("db", "cleanup", "theme_bc", "save")
 
 

### 2020-01-28 code/part_1/02_merge_rolls_doccs.R

@@ -14,7 +14,8 @@ nys_roll <- select(nys_roll, -a, -b, -c)
 
 
 nys_roll <- nys_roll %>%
-  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA))) %>%
+  mutate_at(vars(last_name, first_name, middle_name),
+            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
   mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
 
 # ## read doccs data - entered in 2017
@@ -39,7 +40,8 @@ doccs_to_rolls <- readRDS("./temp/in_2017.rds") %>%
   rename(first_name = first,
          middle_name = middle,
          last_name = last) %>% 
-  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)))
+  mutate_at(vars(last_name, first_name, middle_name),
+            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .)))
 
 merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
 

### 2020-01-28 code/part_1/04_lost_voters_maps.R

@@ -70,4 +70,6 @@ city_map <- ggplot() +
 
 saveRDS(city_map, "./output/city_map.RDS")
 
+ggsave(city_map, file = "./temp/lost_voters_map.png")
+
 

### 2020-01-28 code/part_1/06_process_nyc_nhoods.R

@@ -171,21 +171,21 @@ share_dem_zip <- nyc %>%
 #       age <- census_median_age(var, state = s, year = 2017, county = c)
 # 
 #       if(var == "tract"){
-#         cvap <- fread("./raw_data/CVAP_2013-2017_ACS_csv_files/Tract.csv") %>% 
-#           filter(lntitle == "Total") %>% 
-#           mutate(GEOID = substring(geoid, 8)) %>% 
+#         cvap <- fread("./raw_data/CVAP_2013-2017_ACS_csv_files/Tract.csv") %>%
+#           filter(lntitle == "Total") %>%
+#           mutate(GEOID = substring(geoid, 8)) %>%
 #           select(GEOID, cvap = CVAP_EST)
 #       }
 #       if(var == "block group"){
-#         cvap <- fread("./raw_data/CVAP_2013-2017_ACS_csv_files/BlockGr.csv") %>% 
-#           filter(lntitle == "Total") %>% 
-#           mutate(GEOID = substring(geoid, 8)) %>% 
+#         cvap <- fread("./raw_data/CVAP_2013-2017_ACS_csv_files/BlockGr.csv") %>%
+#           filter(lntitle == "Total") %>%
+#           mutate(GEOID = substring(geoid, 8)) %>%
 #           select(GEOID, cvap = CVAP_EST)
 #       }
 #       if(var == "zcta"){
 #         cvap <- census_vap(var, state = s, year = 2017, county = c)
 #       }
-#       
+# 
 #       noncit <- census_non_citizen(var, state = s, year = 2017, county = c)
 # 
 #       units <- left_join(income,
@@ -199,18 +199,18 @@ share_dem_zip <- nyc %>%
 #   }))
 #   return(units)
 #   })
-# 
-# #### noncit not available at block group level
-# noncit <- rbindlist(lapply(c("KINGS", "QUEENS", "BRONX", "NEW YORK", "RICHMOND"), function(c){
-#   noncit <- census_non_citizen("tract", state = "NY", year = 2017, county = c)
-# }))
-# 
-# geos[[2]]$tract <- substring(geos[[2]]$GEOID, 1, 11)
-# geos[[2]] <- dplyr::select(geos[[2]], -share_non_citizen)
-# geos[[2]] <- left_join(geos[[2]], noncit, by = c("tract" = "GEOID"))
-# 
-# 
[…]

### 2020-01-28 code/part_1/08_basic_regressions.R

@@ -36,6 +36,66 @@ bg_model2_ses  <- data.frame(
 
 save(bg_model, bg_model_ses, bg_model2, bg_model2_ses, file = "./temp/bg_model_reg_ols.rdata")
 
+
+bg1 <- lm_robust(to ~ lost_voters + nh_black +
+                   median_income + latino + nh_white +
+                   some_college + median_age + reg_rate + share_dem +
+                   share_non_citizen + share_winner,
+                 data = block_groups, clusters = district, se_type = "stata")
+
+bg2 <- lm_robust(to ~ lost_voters * nh_black +
+                   median_income + latino + nh_white +
+                   some_college + median_age + reg_rate + share_dem +
+                   share_non_citizen + share_winner,
+                 data = block_groups, clusters = district, se_type = "stata")
+
+start <- plot_summs(bg1, bg2,
+                    coefs = c("Lost Voter Count" = "lost_voters",
+                              "(Lost Voters)×(% Black)" = "lost_voters:nh_black"))
+
+
+start + scale_x_continuous(labels = percent_format(accuracy = 0.1)) + theme_bc() +
+  theme(axis.text = element_text(size = 23),
+        axis.text.y = element_text(size = 15),
+        legend.text = element_text(size = 15),
+        legend.title = element_text(size = 15),
+        axis.title = element_text(size = 23),
+        text = element_text(size = 15),
+        plot.title = element_text(size = 23)) +
+  ggtitle("Effect of Lost Voters on Neighborhood Turnout")
+
+
+ggsave("./temp/coef_plot3.png", width = 11, height = 7.25, units = "in")
+
+start <- plot_summs(bg1, bg2,
+                    coefs = c("Lost Voter Count" = "lost_voters",
+                              "(Lost Voters)×(% Black)" = "lost_voters:nh_black",
+                              "% Non-Hispanic Black" = "nh_black",
+                              "% Latino" = "latino",
+                              "% Non-Hispanic White" = "nh_white",
+                              "Median Income (Thousands)" = "median_income",
+                              "Median Age" = "median_age",
+                              "% With Some College" = "some_college",
+                              "% Non-Citizens" = "share_non_citizen",
+                              "Registration Rate" = "reg_rate",
+                              "% Democrats" = "share_dem",
+                              "% Won by Council Rep." = "share_winner"))
+
+
[…]

### 2020-01-28 code/part_1/09_post_est_maps.R

@@ -58,10 +58,38 @@ dec_map <- ggplot() +
   geom_path(data = bbs, aes(x = long, y = lat, group = group), color = "black") +
   coord_map() +
   labs(x = NULL, y = NULL) + 
-  scale_fill_gradient(high = "#DCDCDC", low = "red", labels = percent_format(accuracy = 1), limits = c(-0.05, 0), oob = squish) +
+  scale_fill_gradient(high = "#DCDCDC", low = "red",
+                      labels = percent_format(accuracy = 1),
+                      limits = c(-0.05, 0), oob = squish,
+                      guide = guide_legend(title.position = "top")) +
   guides(fill = guide_colorbar(title.hjust = .5, title = NULL))
 
 saveRDS(dec_map, "./temp/dec_block_map.rds")
+
+dec_map <- ggplot() +
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        panel.background = element_blank(),
+        panel.border = element_blank(),
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        plot.subtitle = element_text(hjust = 0.5),
+        legend.background = element_blank(),
+        legend.key=element_blank(),
+        legend.key.width = unit(1.5, "cm"),
+        text = element_text(family = "LM Roman 10")) +
+  geom_polygon(data = tract_shp, aes(x = long, y = lat, group = group), fill = "#DCDCDC") +
+  geom_polygon(data = bg_shp, aes(x = long, y = lat, group = group, fill = decrease)) +
+  geom_path(data = bbs, aes(x = long, y = lat, group = group), color = "black") +
+  coord_map() +
+  labs(x = NULL, y = NULL) + 
+  scale_fill_gradient(low = "red", high = "#DCDCDC",
+                     limits = c(-0.05, 0), oob = squish,
+                     name = "Estimated Effect (Percentage Points)",
+                     labels = percent_format(accuracy = 1),
+                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5))
+
+ggsave("./output/dep_map.png", plot = dec_map)
 
 ### interactive
 

### 2020-01-28 code/part_1/did/05_match_did.R

@@ -23,7 +23,7 @@ if(on_nyu){
   cl <- NCPUS(detectCores() - 1)
 }
 
-for(geo in c("tract")){
+for(geo in c("block_group")){
   
   units <- readRDS(paste0("./temp/", geo, "_pre_match.rds"))
   units2 <- readRDS(paste0("./temp/", geo, "_pre_match_16.rds")) %>% 
@@ -34,7 +34,7 @@ for(geo in c("tract")){
   
 
   match_data <- units %>% 
-    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, reg_rate, share_dem, share_non_citizen, share_winner)
+    dplyr::select(median_income, latino, nh_black, nh_white, some_college, median_age, reg_rate, share_dem, share_non_citizen)
   
   genout <- GenMatch(Tr = units$treat, X = match_data,
                      M = 3, replace = T, pop.size = 1000, cluster = cl)
@@ -46,7 +46,7 @@ for(geo in c("tract")){
   
   X <- units %>% 
     dplyr::select(median_income, latino, nh_black, nh_white, some_college,
-                  median_age, reg_rate, share_dem, share_non_citizen, share_winner)
+                  median_age, reg_rate, share_dem, share_non_citizen)
   
   match_count <- ifelse(geo == "tract", 10, 30)
 
@@ -90,7 +90,7 @@ for(geo in c("tract")){
   order <- fread("./raw_data/misc/var_orders.csv")
   
   balance <- MatchBalance(treat ~ median_income + latino + nh_black + nh_white +
-                            some_college + median_age + reg_rate + share_dem + share_non_citizen + share_winner, match.out = mout,
+                            some_college + median_age + reg_rate + share_dem + share_non_citizen, match.out = mout,
                           data = units)
   TrMean <- c()
   PreMean <- c()
@@ -115,7 +115,7 @@ for(geo in c("tract")){
     PostQQmax <- unlist(c(PostQQmax, balance$AfterMatching[[i]]$qqsummary[3]))
   }
   
-  varnames <- c("median_income", "latino", "nh_black", "nh_white", "some_college", "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner")
+  varnames <- c("median_income", "latino", "nh_black", "nh_white", "some_college", "median_age", "reg_rate", "share_dem", "share_non_citizen")
   
   
   df <- data.frame("TrMean" = TrMean,
@@ -166,9 +166,21 @@ for(geo in c("tract")){
   
[…]

### 2020-01-28 code/part_1/xx_arrest_map.R

@@ -1,0 +1,40 @@
+
+
+arrests <- fread("./raw_data/NYPD_Arrests_Data__Historic_.csv") %>% 
+  filter(LAW_CAT_CD == "F",
+         substring(ARREST_DATE, nchar(ARREST_DATE) - 3) == "2017") %>%
+  group_by(ARREST_PRECINCT) %>% 
+  tally() %>% 
+  rename(arrests = n)
+  
+
+ps <- readOGR("./raw_data/shapefiles/Police Precincts", "geo_export_3c2a739c-9600-4c5a-b7e1-a9b2010ac598")
+ps <- spTransform(ps, CRS("+proj=longlat +ellps=WGS84 +no_defs"))
+ps@data$id <- rownames(ps@data)
+dists <- fortify(ps)
+
+dists <- left_join(dists, ps@data)
+
+dists <- left_join(dists, arrests, by = c("precinct" = "ARREST_PRECINCT"))
+
+city_map <- ggplot() +
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank(),
+        panel.background = element_blank(),
+        panel.border = element_blank(),
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        legend.background = element_blank(),
+        legend.key=element_blank()) +
+  geom_polygon(data = dists, aes(x = long, y = lat, group = group, fill = arrests)) +
+  geom_path(data = dists, aes(x = long, y = lat, group = group), color = "black") +
+  coord_map() +
+  labs(x = NULL, y = NULL) +
+  scale_fill_gradient2(low = "#ffeda0", mid = "#feb24c", high = "red", limits = c(0, 2000),
+                       oob = squish, name = "Felony Arrests in 2017",
+                       labels = comma,
+                       guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = 10))
+
+saveRDS(city_map, "./temp/arrest_map.rds")
+
+ggsave(city_map, file = "./temp/arrest_map.png")

### 2020-01-28 code/part_2/04_match_parolees_voter.R

@@ -15,14 +15,16 @@ nys_roll <- select(nys_roll, -a, -b, -c)
 
 
 nys_roll <- nys_roll %>%
-  mutate(gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA)),
-         dob = as.Date(as.character(dob), "%Y%m%d"))
+  mutate_at(vars(last_name, first_name, middle_name),
+            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
+  mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
 
 #### MATCH DOC TO VOTER FILE ####
 
 doccs_to_rolls <- readRDS("./temp/parolees_with_restoration.rds")
 doccs_to_rolls <- doccs_to_rolls %>% 
-  mutate(middle_name = gsub("[[:punct:]]", "", ifelse(middle_name != "", middle_name, NA))) %>% 
+  mutate_at(vars(last_name, first_name, middle_name),
+            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
   rename(dob = dob_parole)
 
 merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)

### 2020-01-28 code/presentations/compare_neighborhoods.R

@@ -1,0 +1,68 @@
+theme_bc <- function(){theme(plot.title = element_text(family = "Roboto Black", size = 23,
+                                            color = "black", hjust = 0.5),
+                  axis.title = element_text(family = "Roboto", 
+                                            color = "#525353", size = 23),
+                  axis.line = element_line(colour = "black"), 
+                  text = element_text(family = "Roboto", color = "#525353"), 
+                  plot.caption = element_text(family = "Roboto", size = 8), 
+                  panel.border = element_blank(), panel.grid.major = element_blank(), 
+                  panel.background = element_rect(fill = "#dddddb"), plot.background = element_rect(fill = "#dddddb"), 
+                  panel.grid.minor = element_blank(), legend.key = element_blank(), 
+                  legend.background = element_rect(color = "black", fill = "#dddddb"), 
+                  plot.subtitle = element_text(hjust = 0.5), 
+                  plot.margin = margin(10, 10, 10, 10),
+                  axis.text = element_text(family = "Roboto", color = "#525353", size = 23))}
+
+
+
+table <- readRDS("./temp/match_table_block_group.rds")
+
+colnames(table) <- make.unique(make.names(colnames(table)))
+
+table <- select(table, X, Treated, Untreated = Control, Control = Control.1)
+
+table <- melt(table, id.vars = "X", value.vars = c("Treated", "Untreated", "Control"))
+
+table$variable <- factor(table$variable, levels = c("Untreated", "Treated", "Control"))
+
+table$value <- as.numeric(gsub(",", "", table$value))
+
+ggplot(filter(table, X == "% Non-Hispanic White"), aes(x = variable, y = value, label = percent(value))) +
+  geom_col(color = "black", fill = "#ED1C24") +
+  theme_bc +
+  scale_y_continuous(labels = percent) + labs(y = "Share Non-Hispanic White", x = NULL) +
+  geom_text(size = 6, position = position_dodge(width = 1), vjust = -0.25,
+            family = "Roboto", color = "#525353") +
+  ggtitle("Neighborhood Share Non-Hispanic White")
+
+ggsave("./temp/perc_white_spsa.png", width = 11, height = 7.25, units = "in")
+
[…]

### 2020-03-04 markdown/part1/UAR_submission.Rmd
write

--- 2020-01-28 markdown/part1/p1_no_tract.Rmd
@@
 title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
-thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
-author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
-institute: "Brennan Center for Justice"
+# thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
+# author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
+# institute: "Brennan Center for Justice"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
     keep_tex: false
-bibliography: "../citations.bib"
+bibliography: "../nyc_neighborhoods.bib"
+mainfont: Times New Roman
 link-citations: yes
@@
     - \usepackage{rotating}
+    - \usepackage{setspace}\doublespacing
 abstract: |
@@
 \pagenumbering{gobble}
-\pagebreak
+\newpage
+\pagenumbering{arabic}
 ```{r setup, echo = F, message = F, warning = F}
@@
 ```
+\begin{singlespace}
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
+true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
+plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
+j <- readRDS("./output/city_map.RDS")
+print(j)
+```
+```{r distribution-lost, echo = F, message = F, warning = F}
+t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
+  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
+
+t <- t[1:11, c(1,2)]
+
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
+                                                 align = "c") %>% 
+  row_spec(14, hline_after = T) %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                font_size = 10)
+j
+```
+```{r block-table-chunk, echo = F}
+t <- readRDS("./temp/match_table_block_group.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"))
+j
+```
+```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/match_reg_block_group.rdata")
+
+load("./temp/match_reg_block_group_nhb.rdata")
+bg_reg <- reg_output
+bg_reg_ses <- reg_output_ses
+
+bg_reg_black <- reg_output_nhblack
+bg_reg_black_ses <- reg_output_nhblack_ses
+
+stargazer(bg_reg, reg_output2, bg_reg_black, reg_output2_nhblack,
+          header = F,
+          se = list(bg_reg_ses, reg_output2_ses, bg_reg_black_ses, reg_output2_nhblack_ses),
+          notes = c("TO REPLACE"),
+          type = "latex", notes.align = "l",
+          notes.append = T,
+          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
+          dep.var.labels = c("Turnout Rate in 2017 General Election"),
+          omit = c("median_income", "latino", "nh_white", "some_college",
+                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner"),
+          title = "\\label{tab:match-reg} Matching Regression",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F,
+          add.lines=list(c("Controls Included" , " ", "X", " ", "X")))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table1.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table1.tex"}
+```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/bg_model_reg_ols.rdata")
+
+stargazer(bg_model, bg_model2,
+          header = F,
+          type = "latex",
+          notes.align = "l",
+          notes.append = T,
+          se = list(bg_model_ses, bg_model2_ses),
+          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
+          dep.var.labels = c("Turnout Rate in 2017 General Election"),
+          title = "\\label{tab:trad-reg} Standard Regression",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          notes = "TO REPLACE",
+          omit = "boro",
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test.tex",
+          out.header = F)
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table2.tex"}
+```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/dec_block_map.rds")
+print(j)
+```
+```{r block-table-chunk-did, echo = F}
+t <- readRDS("./temp/match_table_block_group_did.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
+  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
+  kable_styling(latex_options = c("scale_down", "HOLD_position"),
+                font_size = 10)
+j
+```
+```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
+load("./temp/match_reg_block_group_did.rdata")
+
+bg_reg <- reg_output
+bg_reg_ses <- reg_output_ses
+
+stargazer(bg_reg, reg_output2,
+          header = F,
+          se = list(bg_reg_ses, reg_output2_ses),
+          notes = c("TO REPLACE"),
+          type = "latex", notes.align = "l",
+          notes.append = T,
+          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
+
+          dep.var.labels = c("Turnout Rate"),
+          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
+          table.placement = "H",
+          omit.stat = c("f", "ser"),
+          table.layout = "-dcm#-t-a-s-n",
+          out = "./temp/test_abc.tex",
+          omit = c("median_income", "latino", "nh_white", "some_college", "district",
+                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black"),
+          out.header = F,
+          add.lines=list(c("Congressional District FE", "X", "X"),
+                         c("Controls Included", " ", "X")))
+
+j <- fread("./temp/test_abc.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+
+j <- j %>% 
+  mutate(n = row_number(),
+         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1))
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table1111.tex"}
+\end{singlespace}
+\pagebreak
 \pagenumbering{arabic}
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2014; @Burch2011].
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2015; @Burch2011].
 Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
@@
 
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
+TABLE 1 ABOUT HERE
 
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
 Testing for false negatives is more challenging.
@@
 
-```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
-j <- readRDS("./output/city_map.RDS")
-print(j)
-```
+FIGURE 1 ABOUT HERE
+
 Although `r comma(nrow(readRDS("./temp/in_2017.rds")))` individuals were formally disenfranchised as of the 2017 election, there were just `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` disenfranchised voters statewide who had participated in the past 10 years.^[Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.] Just `r round(100 * nrow(readRDS("./temp/ids_of_lost_voters.rds")) / nrow(readRDS("./temp/in_2017.rds")), digits = 1)` percent of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
@@
 # Testing for Neighborhood Turnout Effects {-}
-```{r table-table, , echo = F, message = F, warning = F}
 
-```
 Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
@@
 
-```{r distribution-lost, echo = F, message = F, warning = F}
-t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
-  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
-
-t <- t[, c(1,2)]
-
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
-                                                 align = "c") %>% 
-  row_spec(14, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                font_size = 10)
-j
-```
+TABLE 2 ABOUT HERE
 
@@
 
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
+TABLE 3 ABOUT HERE
 
@@
 
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
+After matching neighborhoods, I use an ordinary least squares regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
 Census block group level turnout rates are calculated using the geocoded voter file.
@@
 Models 2 and 4 include the covariates on which the matching was done.
-```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group.rdata")
 
-load("./temp/match_reg_block_group_nhb.rdata")
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-bg_reg_black <- reg_output_nhblack
-bg_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, reg_output2, bg_reg_black, reg_output2_nhblack,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses, bg_reg_black_ses, reg_output2_nhblack_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          omit = c("median_income", "latino", "nh_white", "some_college",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner"),
-          title = "\\label{tab:match-reg} Matching Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          add.lines=list(c("Controls Included" , " ", "X", " ", "X")))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table1.tex"}
+TABLE 4 ABOUT HERE
 
@@
 
-```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/bg_model_reg_ols.rdata")
-
-stargazer(bg_model, bg_model2,
-          header = F,
-          type = "latex",
-          notes.align = "l",
-          notes.append = T,
-          se = list(bg_model_ses, bg_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          title = "\\label{tab:trad-reg} Standard Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = "TO REPLACE",
-          omit = "boro",
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F)
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table2.tex"}
+TABLE 5 ABOUT HERE
 
@@
 
-```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/dec_block_map.rds")
-print(j)
-```
+FIGURE 2 ABOUT HERE
 
@@
 
-```{r block-table-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_block_group_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
-In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups.
+TABLE 6 ABOUT HERE
+
+In Table \ref{tab:did-reg}, I present the outcome of an OLS regression on this subset of block groups.
 The variable D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
@@
 
-```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group_did.rdata")
-
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-stargazer(bg_reg, reg_output2,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
-
-          dep.var.labels = c("Turnout Rate"),
-          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test_abc.tex",
-          omit = c("median_income", "latino", "nh_white", "some_college", "district",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black"),
-          out.header = F,
-          add.lines=list(c("Congressional District FE", "X", "X"),
-                         c("Controls Included", " ", "X")))
-
-j <- fread("./temp/test_abc.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table1111.tex"}
+TABLE 7 ABOUT HERE
 
@@
 @Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, @Weaver2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
+Similarly, @Lerman2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
 It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.

### 2020-03-04 markdown/part1/for_brennan.Rmd
write

--- 2020-01-28 markdown/part1/p1_no_tract.Rmd
@@
 title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
-thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
-author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
+author: Kevin Morris^[The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.]
 institute: "Brennan Center for Justice"
@@
     latex_engine: xelatex
-    keep_tex: false
-bibliography: "../citations.bib"
+    keep_tex: true
+bibliography: "../nyc_neighborhoods.bib"
 link-citations: yes
@@
     - \usepackage{rotating}
+    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+    - \usepackage{setspace}
 abstract: |
@@
 In this paper, I redefine “lost voters” as disenfranchised individuals with a history of participating in elections.
-I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
+I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods.
 I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
@@
 ---
+<!-- --- -->
+<!-- title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City" -->
+<!-- author: "Kevin Morris" -->
+<!-- date: "`r format(Sys.time(), '%B %d, %Y')`" -->
+<!-- output: word_document -->
+<!-- bibliography: "../research/citations.bib" -->
+<!-- link-citations: yes -->
+<!-- always_allow_html: yes -->
+<!-- --- -->
 \pagenumbering{gobble}
@@
 \pagenumbering{arabic}
+\doublespacing
 # Introduction {-}
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2014; @Burch2011].
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2015; @Burch2011].
 Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
 With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
 
@@
 Since @Uggen2002, other scholars have also investigated the direct effect of felony disenfranchisement [@Miles2004;@Uggen2004; @Drucker2005; @Ochs2006].
-
-A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006].
-As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-@Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
 
@@
 
-# Theoretical Implications for Reduced Turnout {-}
+# Methodology and Data {-}
 
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
-@King2016; @Bowers2009].
-With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015 examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
+To identify lost voters, and to determine whether lost voters reduce neighborhood turnout, I use data provided directly by the State of New York.
 
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (59).
-@Griffin2005 finds similar effects in the United States Senate, demonstrating that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
+The first dataset is from the Department of Corrections and Community Supervision (DOCCS).
+This data includes information on every individual who was incarcerated in state prison, or on parole after serving time in state prison.
+These data include individuals' first, middle, and last names, as well as their date of birth, the dates they spent incarcerated or on parole, and other information.
+I limit the records in the data to include all individuals who were either in prison or on parole as of the mayoral election in 2017.
 
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group" (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
+The second primary dataset is the registered voter file from New York State.
+The registered voter file includes information on every voter registered in the state such as name, address, vote history, and gender.
+It also includes the records of everyone who has been purged --- in other words, whose registration has been cancelled.
+The records are marked as "purged," and allow me to identify the vote history even of individuals who have been purged for a felony conviction.
 
-@Hajnal2005 and @Hajnal2009, however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-@Hajnal2005 tells us that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-@Hajnal2005 and @Anzia2019 both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
+By matching these two datasets using first, middle, and last names, and date of birth, I can identify all individuals who were disenfranchised in the City of New York on election day in 2017 *who had a history of voting.* Specifically, a "lost voter" is someone who cast a ballot between January 1, 2007 and December 31, 2016 and was imprisoned or on parole for a felony offense on November 7^th^, 2017.
 
-# Data {-}
-
-## Criminal Justice Data {-}
-
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and other demographic information.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-## Voter File Data {-}
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from April 30^th^, 2018.
-
-The New York State Voter File is unique in its treatment of "purged" voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
-
-## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
-
-## Matching {-}
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a "match," records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
-
-* Middle names are identical. If neither set of records includes a middle name, this condition is met.
-* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-* A middle name or middle initial in one set of records, and a missing middle name in the other set.
-
-Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andrew Doe" and "John Anthony Doe" would not.
-
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-@Meredith2013 offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
-
-# Identification of Lost Voters {-}
-
-Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
-By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
-Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
+# Why Do We Care About "Lost Voters"? {-}
 
@@
 Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
-I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
-These higher indirect (depressive) effects likely arise through three mechanisms.
+Here, I am isolating the direct effect of felony disenfranchisement.
 
-* Firstly, lost voters are likely members of families and communities where voting is valued.
-Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
-Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
-Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
+# Geocoding
 
-* Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
-The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
-
-* Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
-Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
-
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
-In many cases, these "lost voters" will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effect in its wake.
-
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election.
+After the lost voters have been identified, the full registered voter file is geocoded by SmartyStreets, which successfully geocoded 99 percent of all voters.
+Figure \ref{fig:citywide-map} shows the geographical distribution of the lost voters.
 
@@
 
-# Testing for Neighborhood Turnout Effects {-}
-```{r table-table, , echo = F, message = F, warning = F}
+# Results {-}
 
-```
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter.
-
-```{r distribution-lost, echo = F, message = F, warning = F}
-t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
-  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
-
-t <- t[, c(1,2)]
-
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
-                                                 align = "c") %>% 
-  row_spec(14, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                font_size = 10)
-j
-```
-
-I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census block groups [@Sekhon2011], using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau's 5-year American Community Survey estimates.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups.
-Matching is done with replacement.
-Table \ref{tab:blockgroup-table} presents the results of these matches.
-
-## Match Output {-}
-As Table \ref{tab:blockgroup-table} demonstrates, the matching procedure results in an untreated group of neighborhoods that looks substantially like the treatment group.
-
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-
-In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
-On average, the non-Hispanic Black share of a block group with a lost voter is more than *double* the share in block groups without lost voters.
-These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] The median income in neighborhoods without lost voters is more than 40 percent higher.
-Table \ref{tab:blockgroup-table} throws into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
-
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match [@Abadie2019].
-Models 2 and 4 include the covariates on which the matching was done.
-```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group.rdata")
-
-load("./temp/match_reg_block_group_nhb.rdata")
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-bg_reg_black <- reg_output_nhblack
-bg_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, reg_output2, bg_reg_black, reg_output2_nhblack,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses, bg_reg_black_ses, reg_output2_nhblack_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          omit = c("median_income", "latino", "nh_white", "some_college",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner"),
-          title = "\\label{tab:match-reg} Matching Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          add.lines=list(c("Controls Included" , " ", "X", " ", "X")))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table1.tex"}
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average `r -1*round(bg_reg[["coefficients"]][["treatTRUE"]]*100, digits = 1)` percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The average treated neighborhood was 38 percent Black, which means that the average block group with a lost voter saw turnout that was `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*38, digits = 2)` percentage points lower than similar neighborhoods without lost voters.
-This is very close to the average treatment effect presented in Models 1 and 2; Models 3 and 4, however, makes clear that this effect is largely concentrated in Black communities.
-This disparate effect could have troubling ramifications for these neighborhoods' political representation.
-
-## Testing Intensity Effects {-}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
+After determining lost voters' geographic coordinates, I determine the Census block group in which they live.
+I then aggregate these numbers up to determine the total number of lost voters residing in each block group in New York City.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
+In addition to the number of lost voters a neighborhood has, I include other sociodemographic characteristics from the Census Bureau's American Community Survey for the 5 years ending in 2017.
+Neighborhood-level turnout is calculated by aggregating the number of votes cast according to the geocoded voter file, and dividing that number by the citizen voting age population.
 Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
 
+Model 1 includes an un-interacted estimate of the number of lost voters as the primary dependent variable.
+Model 2, however, asks whether the effect of lost voters operates differently in Black communities.
+Prior research indicates that this might be the case, so the primary independent variable --- the number of lost voters --- is interacted in Model 2 with the share of the neighborhood that is non-Hispanic Black.
+
+\begin{singlespace}
 ```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table2.tex"}
+\end{singlespace}
 
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model, though they are slightly smaller.
-This is unsurprising, because the average treated neighborhood had more than one lost voter.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Model 2 again makes clear that this effect is concentrated in Black neighborhoods.
+As Table \ref{tab:trad-reg} shows, lost voters are generally associated with lower turnout.
+Each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points.
+Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
@@
 ```
-
-## Toward a Causal Estimation {-}
-
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
-
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were *newly* home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
-
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
-
-Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
-
-```{r block-table-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_block_group_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
-In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups.
-The variable D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election, a difference that was not statistically significant.
-D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-This coefficient is negative and statistically significant, indicating that in the first election after the removal of a voter a neighborhood's turnout rate decreased by 2.6 percentage points.
-
-```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group_did.rdata")
-
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-stargazer(bg_reg, reg_output2,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
-
-          dep.var.labels = c("Turnout Rate"),
-          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test_abc.tex",
-          omit = c("median_income", "latino", "nh_white", "some_college", "district",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black"),
-          out.header = F,
-          add.lines=list(c("Congressional District FE", "X", "X"),
-                         c("Controls Included", " ", "X")))
-
-j <- fread("./temp/test_abc.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table1111.tex"}
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.]
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
 
@@
 Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
+Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
 This means that lost voters have no spillover effects in neighborhoods with small Black populations.
@@
 In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.16 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+In an average block group that is 38 percent Black, each lost voter cost the neighborhood `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
 In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.

### 2020-03-04 markdown/part1/for_brennan.tex
write

--- 2019-11-12 markdown/part1/p1.tex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={Neighborhoods and Felony Disenfranchisement: The Case of New York City},
   pdfauthor={Kevin Morris},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{November 12, 2019}
-
+\usepackage{rotating}
+\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+\usepackage{setspace}\doublespacing
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
-\usepackage{rotating}
-\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+\title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
+\author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
+\date{March 03, 2020}
 
@@
 Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is an incarcerated population that looks far different than the rest of th state: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.}
+The result is an incarcerated population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.}
 
@@
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}; Gerber et al.
-\protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{b}; Gerber et al.
+\protect\hyperlink{ref-Gerber2015}{2015}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
 Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
 With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (8).
 
@@
 Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
-Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}) produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for for disenfranchisment in Florida.
+Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}) produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for disenfranchisement in Florida.
 Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
 Since Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}), other scholars have also investigated the direct effect of felony disenfranchisement (Miles \protect\hyperlink{ref-Miles2004}{2004}; Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-
-A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
-King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
-{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (724).
-
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the \emph{state} level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
 
@@
 Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the indirect neighborhood effects of lost voters.
-Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
+Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and García-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and García-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
 These effects are particularly pronounced for women of color.
@@
 
-Similarly, White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
+Similarly, White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{a}) finds that incarceration has only moderate indirect effects on voting.
 She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (607).
@@
 
-\hypertarget{theoretical-implications-for-reduced-turnout}{%
-\section*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
-\addcontentsline{toc}{section}{Theoretical Implications for Reduced Turnout}
+\hypertarget{methodology-and-data}{%
+\section*{Methodology and Data}\label{methodology-and-data}}
+\addcontentsline{toc}{section}{Methodology and Data}
 
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community (e.g.~King and Erickson \protect\hyperlink{ref-King2016}{2016}; Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: Kinsella, McTague, and Raleigh (\protect\hyperlink{ref-Kinsella2015}{2015}) examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
+To identify lost voters, and to determine whether lost voters reduce neighborhood turnout, I use data provided directly by the State of New York.
 
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes (Martin \protect\hyperlink{ref-Martin2003}{2003}).
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-``{[}H{]}igher citizen participation is rewarded,'' Martin and Claibourn (\protect\hyperlink{ref-Martin2013}{2013}) concludes, ``with enhanced policy responsiveness'' (59).
-Griffin and Newman (\protect\hyperlink{ref-Griffin2005}{2005}) finds similar effects in the United States Senate, demonstrating that ``voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not'' (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
+The first dataset is from the Department of Corrections and Community Supervision (DOCCS).
+This data includes information on every individual who was incarcerated in state prison, or on parole after serving time in state prison.
+These data include individuals' first, middle, and last names, as well as their date of birth, the dates they spent incarcerated or on parole, and other information.
+I limit the records in the data to include all individuals who were either in prison or on parole as of the mayoral election in 2017.
 
-Although research on the impact of local turnout on city-wide policy is scarce, Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) examines the impact of senior turnout on ``senior-friendly'' policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials \emph{are} responsive to senior turnout when seniors ``are a more cohesive, meaningful group'' (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of \emph{neighborhood} turnout rates on city policy.
+The second primary dataset is the registered voter file from New York State.
+The registered voter file includes information on every voter registered in the state such as name, address, vote history, and gender.
+It also includes the records of everyone who has been purged --- in other words, whose registration has been cancelled.
+The records are marked as ``purged,'' and allow me to identify the vote history even of individuals who have been purged for a felony conviction.
 
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}), however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) tells us that ``Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils'' (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
+By matching these two datasets using first, middle, and last names, and date of birth, I can identify all individuals who were disenfranchised in the City of New York on election day in 2017 \emph{who had a history of voting.} Specifically, a ``lost voter'' is someone who cast a ballot between January 1, 2007 and December 31, 2016 and was imprisoned or on parole for a felony offense on November 7\textsuperscript{th}, 2017.
 
-\hypertarget{data}{%
-\section*{Data}\label{data}}
-\addcontentsline{toc}{section}{Data}
-
-\hypertarget{criminal-justice-data}{%
-\subsection*{Criminal Justice Data}\label{criminal-justice-data}}
-\addcontentsline{toc}{subsection}{Criminal Justice Data}
-
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and other demographic information.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-\hypertarget{voter-file-data}{%
-\subsection*{Voter File Data}\label{voter-file-data}}
-\addcontentsline{toc}{subsection}{Voter File Data}
-
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
-
-The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
-
-\hypertarget{geocoding}{%
-\subsection*{Geocoding}\label{geocoding}}
-\addcontentsline{toc}{subsection}{Geocoding}
-
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6 percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
-
-\hypertarget{matching}{%
-\subsection*{Matching}\label{matching}}
-\addcontentsline{toc}{subsection}{Matching}
-
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a ``match,'' records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
-
-\begin{itemize}
-\tightlist
-\item
-  Middle names are identical. If neither set of records includes a middle name, this condition is met.
-\item
-  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-\item
-  A middle name or middle initial in one set of records, and a missing middle name in the other set.
-\end{itemize}
-
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
-
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 20,955\\
-Birthdate + 35 Days & 105\\
-Birthdate - 35 Days & 92\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
-
-\hypertarget{identification-of-lost-voters}{%
-\section*{Identification of Lost Voters}\label{identification-of-lost-voters}}
-\addcontentsline{toc}{section}{Identification of Lost Voters}
-
-Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
-By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
-Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
+\hypertarget{why-do-we-care-about-lost-voters}{%
+\section*{Why Do We Care About ``Lost Voters''?}\label{why-do-we-care-about-lost-voters}}
+\addcontentsline{toc}{section}{Why Do We Care About ``Lost Voters''?}
 
@@
 Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
-I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
-These higher indirect (depressive) effects likely arise through three mechanisms.
+Here, I am isolating the direct effect of felony disenfranchisement.
 
-\begin{itemize}
-\item
-  Firstly, lost voters are likely members of families and communities where voting is valued.
-Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
-Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
-Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
-\item
-  Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
-The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
-\item
-  Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
-Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
-\end{itemize}
+\hypertarget{geocoding}{%
+\section{Geocoding}\label{geocoding}}
 
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
-In many cases, these ``lost voters'' will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
-
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
-There were 2,518 such lost voters within New York City as of the 2017 general election.
+After the lost voters have been identified, the full registered voter file is geocoded by SmartyStreets, which successfully geocoded 99 percent of all voters.
+Figure \ref{fig:citywide-map} shows the geographical distribution of the lost voters.
 
@@
 
-{\centering \includegraphics{p1_files/figure-latex/citywide-map-1}
+{\centering \includegraphics{for_brennan_files/figure-latex/citywide-map-1}
 
@@
 
-\hypertarget{testing-for-neighborhood-turnout-effects}{%
-\section*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
-\addcontentsline{toc}{section}{Testing for Neighborhood Turnout Effects}
+\hypertarget{results}{%
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
 
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
-
-\begin{table}[H]
-
-\caption{\label{tab:distribution-lost}\label{tab:nhood-dist} Distribution of Lost Voters}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{ccc}
-\toprule
-Number of Lost Voters & Count of Block Groups & Count of Tracts\\
-\midrule
-0 & 4,272 & 1,008\\
-1 & 1,023 & 416\\
-2 & 301 & 212\\
-3 & 107 & 138\\
-4 & 30 & 51\\
-5 & 19 & 46\\
-6 & 9 & 23\\
-7 & 3 & 15\\
-8 & 5 & 13\\
-9 & 0 & 7\\
-10 & 0 & 2\\
-11 & 0 & 2\\
-12 & 0 & 2\\
-14 & 0 & 2\\
-\hline
-Total & 5,769 & 1,937\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}), using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau's 5-year American Community Survey estimates.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
-Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
-
-\hypertarget{match-output}{%
-\subsection*{Match Output}\label{match-output}}
-\addcontentsline{toc}{subsection}{Match Output}
-
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk}\label{tab:blockgroup-table} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.34 & 93.02 & 89.13 & 85.79 & 79.01\\
-\% Non-Hispanic Black & 0.38 & 0.16 & 0.38 & 0.37 & 97.93 & 94.46 & 94.41 & 92.77\\
-\% Non-Hispanic White & 0.17 & 0.40 & 0.17 & 0.18 & 96.37 & 96.06 & 95.56 & 92.99\\
-Median Income & 51,792.52 & 72,612.36 & 51,792.52 & 52,768.97 & 95.31 & 93.47 & 88.62 & 74.49\\
-\% With Some College & 0.64 & 0.70 & 0.64 & 0.64 & 90.75 & 94.14 & 90.29 & 80.70\\
-Median Age & 35.99 & 38.34 & 35.99 & 36.25 & 89.18 & 90.84 & 85.93 & 78.56\\
-Registration Rate & 1.00 & 0.94 & 1.00 & 0.98 & 59.77 & 79.89 & 72.28 & 60.42\\
-\% Democrats & 0.75 & 0.65 & 0.75 & 0.75 & 96.22 & 96.52 & 94.92 & 91.02\\
-\% Noncitizen & 0.16 & 0.16 & 0.16 & 0.16 & 16.58 & 14.80 & -0.06 & -60.28\\
-\% Won by City Council Representative & 0.85 & 0.81 & 0.85 & 0.86 & 89.44 & 83.40 & 79.88 & 70.11\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-table-chunk}\label{tab:tract-table} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.32 & 0.22 & 0.32 & 0.33 & 98.55 & 85.12 & 82.84 & 73.01\\
-\% Non-Hispanic Black & 0.36 & 0.14 & 0.36 & 0.36 & 99.77 & 88.74 & 88.36 & 83.30\\
-\% Non-Hispanic White & 0.19 & 0.42 & 0.19 & 0.19 & 97.50 & 96.32 & 91.29 & 79.95\\
-Median Income & 53,409.42 & 71,754.50 & 53,409.42 & 55,422.41 & 89.03 & 90.83 & 84.86 & 62.84\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 90.19 & 93.52 & 88.89 & 73.01\\
-Median Age & 35.76 & 38.04 & 35.76 & 35.80 & 98.12 & 74.51 & 80.56 & 71.91\\
-Registration Rate & 0.94 & 0.90 & 0.94 & 0.93 & 76.63 & 86.16 & 84.24 & 71.28\\
-\% Democrats & 0.74 & 0.62 & 0.74 & 0.73 & 94.58 & 94.93 & 92.05 & 82.37\\
-\% Noncitizen & 0.16 & 0.17 & 0.16 & 0.17 & -123.57 & 22.48 & 8.85 & -38.79\\
-\% Won by City Council Representative & 0.85 & 0.79 & 0.85 & 0.85 & 93.03 & 81.57 & 79.27 & 67.47\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-
-In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
-On average, the non-Hispanic Black share of a block group with a lost voter is more than \emph{double} the share in block groups without lost voters.
-These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} The median income in neighborhoods without lost voters is more than 40 percent higher.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
-
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
-
-\input{"../../temp/table1.tex"}
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average 0.9 percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as 2.7 percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just 14.6 percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
-
-\hypertarget{testing-intensity-effects}{%
-\subsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
-\addcontentsline{toc}{subsection}{Testing Intensity Effects}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on \emph{how many} voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
+After determining lost voters' geographic coordinates, I determine the Census block group in which they live.
+I then aggregate these numbers up to determine the total number of lost voters residing in each block group in New York City.
 In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
+In addition to the number of lost voters a neighborhood has, I include other sociodemographic characteristics from the Census Bureau's American Community Survey for the 5 years ending in 2017.
+Neighborhood-level turnout is calculated by aggregating the number of votes cast according to the geocoded voter file, and dividing that number by the citizen voting age population.
 Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
 
+Model 1 includes an un-interacted estimate of the number of lost voters as the primary dependent variable.
+Model 2, however, asks whether the effect of lost voters operates differently in Black communities.
+Prior research indicates that this might be the case, so the primary independent variable --- the number of lost voters --- is interacted in Model 2 with the share of the neighborhood that is non-Hispanic Black.
+
+\begin{singlespace}
+
 \input{"../../temp/table2.tex"}
+\end{singlespace}
 
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
+As Table \ref{tab:trad-reg} shows, lost voters are generally associated with lower turnout.
+Each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points.
+Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.87 percentage points.
@@
 
-{\centering \includegraphics{p1_files/figure-latex/postest-map-1}
+{\centering \includegraphics{for_brennan_files/figure-latex/postest-map-1}
 
@@
 \end{figure}
-
-\hypertarget{toward-a-causal-estimation}{%
-\subsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
-\addcontentsline{toc}{subsection}{Toward a Causal Estimation}
-
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
-
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were \emph{newly} home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 \emph{did} have a lost voter in 2016, and are thus excluded from this analysis.
-
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
-
-Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
-As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
-
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.35 & 99.00 & 96.26 & 94.92 & 85.60\\
-\% Non-Hispanic Black & 0.33 & 0.16 & 0.33 & 0.33 & 99.02 & 96.57 & 95.67 & 90.10\\
-\% Non-Hispanic White & 0.21 & 0.41 & 0.21 & 0.21 & 99.61 & 94.66 & 92.61 & 82.94\\
-Median Income & 55,852.62 & 73,433.49 & 55,852.62 & 56,287.76 & 97.52 & 87.85 & 86.56 & 72.89\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 99.31 & 92.25 & 87.30 & 71.15\\
-Median Age & 36.17 & 38.47 & 36.17 & 36.04 & 94.37 & 62.35 & 72.27 & 69.84\\
-Registration Rate & 0.98 & 0.94 & 0.98 & 0.95 & 25.09 & 46.49 & 42.16 & 33.06\\
-\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 95.94 & 94.99 & 93.62 & 86.68\\
-\% Noncitizen & 0.17 & 0.16 & 0.17 & 0.17 & 63.46 & 75.01 & 66.85 & 45.71\\
-\% Won by City Council Representative & 0.85 & 0.80 & 0.85 & 0.85 & 88.19 & 83.99 & 78.23 & 67.54\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-chunk-did}\label{tab:did-match-tract} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 94.34 & 86.10 & 84.77 & 75.82\\
-\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.25 & 99.28 & 89.75 & 86.62 & 77.90\\
-\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.02 & 87.45 & 77.73 & 61.67\\
-Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,973.89 & 84.23 & 78.35 & 76.98 & 66.57\\
-\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 96.86 & 86.51 & 83.73 & 76.55\\
-Median Age & 36.36 & 38.23 & 36.36 & 36.39 & 98.60 & 73.15 & 69.00 & 42.11\\
-Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 10.68 & 64.27 & 63.46 & 56.35\\
-\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.23 & 90.06 & 87.35 & 76.31\\
-\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 78.09 & 73.55 & 62.03 & 48.36\\
-\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.91 & 75.59 & 65.98 & 38.91\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups and tracts.
-In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly \emph{higher} turnout prior to intervention.
-D(Treat) \(\times\) D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
-
-\input{"../../temp/table1111.tex"}
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
 
@@
 In New York City, there are 1386 block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of 912 voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of 912, having any lost voter reduced turnout by 9.4 ballots.
-Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just 3.8 ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood 6.5 ballots.
+In an average block group that is 38 percent Black, each lost voter cost the neighborhood 6.5 ballots.
 In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just 1.6 ballots --- hardly more votes than the lost voter herself.
@@
 Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, Weaver and Lerman (\protect\hyperlink{ref-Weaver2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
+Similarly, Lerman and Weaver (\protect\hyperlink{ref-Lerman2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
 It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
@@
 
-This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
+This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{b}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
 This is perhaps unsurprising.
@@
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\noindent
+The interlocking nature of racioeconomic segregation, policing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abadie2019}{}%
-Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \url{https://scholar.harvard.edu/spiess/publications/robust-post-matching-inference.}
-
-\leavevmode\hypertarget{ref-Anzia2019}{}%
-Anzia, Sarah F.
-2019.
-``When Does a Group of Citizens Influence Policy?
-Evidence from Senior Citizen Participation in City Politics.'' \emph{The Journal of Politics} 81 (1): 1--14.
-\url{https://doi.org/10.1086/699916}.
-
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Bowers2009}{}%
@@
 2009.
-``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons.'' \emph{Social Science Quarterly} 90 (3): 722--43.
+``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons*.'' \emph{Social Science Quarterly} 90 (3): 722--43.
 \url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
@@
 \leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
+Brennan Center for Justice.
+2019.
+``Criminal Disenfranchisement Laws Across the United States.'' https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states.
 
@@
 2011.
-``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
+``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law \& Society Review} 45 (3): 699--730.
 \url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
@@
 \leavevmode\hypertarget{ref-Burch2013}{}%
----------.
+Burch, Traci R.
 2013.
-``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 184--201.
+``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
 \url{https://doi.org/10.1177/0002716213503093}.
@@
 2005.
-``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
-Sentencing Project.
-\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
+``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research Report.
+The Sentencing Project.
 
 \leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
+Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516--29. \url{https://doi.org/10.2307/2146812}.
 
@@
 ``Aggressive Policing and the Mental Health of Young Urban Men.'' \emph{American Journal of Public Health} 104 (12): 2321--7.
-\url{https://doi.org/10.2105/ajph.2014.302046}.
+\url{https://doi.org/10.2105/AJPH.2014.302046}.
 
@@
 2007.
-``An Analysis of the New York City Police Departments ``Stop-and-Frisk'' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
+``An Analysis of the New York City Police Department's `Stop-and-Frisk' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
 \url{https://doi.org/10.1198/016214506000001040}.
@@
 
-\leavevmode\hypertarget{ref-Gerber2014}{}%
+\leavevmode\hypertarget{ref-Gerber2015}{}%
 Gerber, Alan S., Gregory A.
@@
 Hendry.
-2014.
-``Can Incarcerated Felons Be (Re)integrated into the Political System?
+2015.
+``Can Incarcerated Felons Be (Re)Integrated into the Political System?
 Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
@@
 
-\leavevmode\hypertarget{ref-Griffin2005}{}%
-Griffin, John D., and Brian Newman. 2005. ``Are Voters Better Represented?'' \emph{The Journal of Politics} 67 (4): 1206--27. \url{https://doi.org/10.1111/j.1468-2508.2005.00357.x}.
-
 \leavevmode\hypertarget{ref-Guest1999}{}%
-Guest, Avery M., and Susan K. Wierzbicki. 1999. ``Social Ties at the Neighborhood Level.'' \emph{Urban Affairs Review} 35 (1): 92--111. \url{https://doi.org/10.1177/10780879922184301}.
-
-\leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
+Guest, Avery M., and Susan K.
+Wierzbicki.
+1999.
+``Social Ties at the Neighborhood Level: Two Decades of GSS Evidence.'' \emph{Urban Affairs Review}, September.
+\url{https://doi.org/10.1177/10780879922184301}.
 
@@
 ``The Impact of Race on Perceptions of Criminal Injustice.'' \emph{Journal of Criminal Justice} 25 (6): 447--62.
-\url{https://doi.org/10.1016/s0047-2352(97)00032-9}.
+\url{https://doi.org/10.1016/S0047-2352(97)00032-9}.
 
 \leavevmode\hypertarget{ref-Huckfeldt1979}{}%
-Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
+Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579--92. \url{https://doi.org/10.2307/2111030}.
 
@@
 2005.
-``Explaining the Great Racial Divide: Perceptions of Fairness in the U.s.
+``Explaining the Great Racial Divide: Perceptions of Fairness in the U.S.
 Criminal Justice System.'' \emph{The Journal of Politics} 67 (3): 762--83.
@@
 \leavevmode\hypertarget{ref-Kenny1992}{}%
-Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259. \url{https://doi.org/10.2307/2111432}.
+Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259--67. \url{https://doi.org/10.2307/2111432}.
 
 \leavevmode\hypertarget{ref-Keyssar2009}{}%
-Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. New York: Basic Books.
+Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. Revised edition. New York: Basic Books.
 
 \leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821. \url{https://doi.org/10.1177/0021934716659195}.
-
-\leavevmode\hypertarget{ref-Kinsella2015}{}%
-Kinsella, Chad, Colleen McTague, and Kevin N.
-Raleigh.
-2015.
-``Unmasking Geographic Polarization and Clustering: A Micro-Scalar Analysis of Partisan Voting Behavior.'' \emph{Applied Geography} 62 (August): 404--19.
-\url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
+King, Bridgett A., and Laura Erickson.
+2016.
+``Disenfranchising the Enfranchised: Exploring the Relationship Between Felony Disenfranchisement and African American Voter Turnout.'' \emph{Journal of Black Studies}, July.
+\url{https://doi.org/10.1177/0021934716659195}.
 
@@
 ``Staying Out of Sight?
-Concentrated Policing and Local Political Action.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 202--19.
+Concentrated Policing and Local Political Action:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
 \url{https://doi.org/10.1177/0002716213503085}.
 
-\leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
-
-\leavevmode\hypertarget{ref-Martin2003}{}%
-Martin, Paul S.
-2003.
-``Votings Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
-\url{https://doi.org/10.1111/1540-5907.00008}.
-
-\leavevmode\hypertarget{ref-Martin2013}{}%
-Martin, Paul S., and Michele P.
-Claibourn.
-2013.
-``Citizen Participation and Congressional Responsiveness: New Evidence That Participation Matters.'' \emph{Legislative Studies Quarterly} 38 (1): 59--81.
-\url{https://doi.org/10.1111/lsq.12003}.
-
-\leavevmode\hypertarget{ref-Meredith2013}{}%
-Meredith, Marc, and Michael Morse.
-2013.
-``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
+\leavevmode\hypertarget{ref-Lerman2014}{}%
+Lerman, Amy E., and Vesla M.
 Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 220--49.
-\url{https://doi.org/10.1177/0002716213502931}.
+2014.
+\emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}.
+Chicago Studies in American Politics.
+Chicago ; London: The University of Chicago Press.
 
@@
 \leavevmode\hypertarget{ref-Mutz2002}{}%
-Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838. \url{https://doi.org/10.2307/3088437}.
+Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838--55. \url{https://doi.org/10.2307/3088437}.
 
@@
 2006.
-``'Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
+``\,`Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
 \url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
@@
 
-\leavevmode\hypertarget{ref-Sekhon2011}{}%
-Sekhon, Jasjeet S.
-2011.
-``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (7).
-\url{https://doi.org/10.18637/jss.v042.i07}.
-
 \leavevmode\hypertarget{ref-Sewell2016}{}%
@@
 2016.
-``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (S1): 42--67.
+``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (1): 42--67.
 \url{https://doi.org/10.1007/s11524-015-0016-7}.
@@
 \leavevmode\hypertarget{ref-sentencing_2016}{}%
-Uggen, Christopher, Ryan Larson, and Sarah Shannon.
-2016.
-``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-Sentencing Project.
-\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
+Uggen, Christopher, Ryan Larson, and Sarah Shannon. 2016. ``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research Report. The Sentencing Project.
 
@@
 ``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
+Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777--803.
 \url{https://doi.org/10.2307/3088970}.
@@
 \leavevmode\hypertarget{ref-Uggen2004}{}%
----------.
+Uggen, Christopher, and Jeffrey Manza.
 2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
+``Lost Voices: The Civic and Political Views of Disenfranchised Felons.'' In \emph{Imprisoning America: The Social Impact of Mass Incarceration}, 165--204.
+Russell Sage Foundation.
 
@@
 2014.
-``Extending the Effects of the Carceral State.
-Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly} 67 (4): 809--22.
+``Extending the Effects of the Carceral State: Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly}, July.
 \url{https://doi.org/10.1177/1065912914542522}.
@@
 \leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela Garcı́a-Castañon.
+Walker, Hannah L., and Marcela García-Castañon.
 2017.
-``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (04): 541--68.
-\url{https://doi.org/10.1017/s1743923x17000198}.
-
-\leavevmode\hypertarget{ref-Weaver2014}{}%
-Weaver, Vesla, and Amy E. Lerman. 2014. \emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}. Chicago: University of Chicago Press.
+``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (4): 541--68.
+\url{https://doi.org/10.1017/S1743923X17000198}.
 
 \leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
+Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/S0003055410000456}.
+
+\leavevmode\hypertarget{ref-White2019a}{}%
+White, Ariel.
+2019a.
+``Family Matters?
+Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
+\url{https://doi.org/10.1017/S0003055418000862}.
 
 \leavevmode\hypertarget{ref-White2019}{}%
-White, Ariel.
-2019a.
+---------.
+2019b.
 ``Misdemeanor Disenfranchisement?
 The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/s000305541800093x}.
-
-\leavevmode\hypertarget{ref-White2019a}{}%
----------.
-2019b.
-``Family Matters?
-Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
-\url{https://doi.org/10.1017/s0003055418000862}.
+\url{https://doi.org/10.1017/S000305541800093X}.
 
 \leavevmode\hypertarget{ref-Wilson1990}{}%
-Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
+Wilson, William J. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Paperback ed., {[}Nachdr.{]}. Chicago: Univ. of Chicago Press.
 

### 2020-03-04 markdown/part1/p1.Rmd
write

@@
     - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+    - \usepackage{setspace}\doublespacing
 abstract: |

### 2020-03-04 markdown/part1/p1_no_tract.Rmd
write

@@
 The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
+\begin{singlespace}
 ```{r shift-dobs-chunk, echo = F}
@@
 ```
+\end{singlespace}
 Testing for false negatives is more challenging.
@@
 More than two-thirds of block groups that lost any voters lost only one voter.
-
+\begin{singlespace}
 ```{r distribution-lost, echo = F, message = F, warning = F}
@@
 
-t <- t[, c(1,2)]
+t <- t[1:11, c(1,2)]
 
@@
 ```
-
+\end{singlespace}
 I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
@@
 As Table \ref{tab:blockgroup-table} demonstrates, the matching procedure results in an untreated group of neighborhoods that looks substantially like the treatment group.
-
+\begin{singlespace}
 ```{r block-table-chunk, echo = F}
@@
 ```
-
+\end{singlespace}
 In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
@@
 Models 2 and 4 include the covariates on which the matching was done.
+\begin{singlespace}
 ```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table1.tex"}
+\end{singlespace}
 
@@
 Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
-
+\begin{singlespace}
 ```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table2.tex"}
-
+\end{singlespace}
 The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model, though they are slightly smaller.
@@
 Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
-
+\begin{singlespace}
 ```{r block-table-chunk-did, echo = F}
@@
 ```
+\end{singlespace}
 In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups.
@@
 This coefficient is negative and statistically significant, indicating that in the first election after the removal of a voter a neighborhood's turnout rate decreased by 2.6 percentage points.
-
+\begin{singlespace}
 ```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table1111.tex"}
+\end{singlespace}
 

### 2020-03-04 markdown/part2/part2_standalone.Rmd
write

@@
     keep_tex: true
-bibliography: "../citations.bib"
+bibliography: "../welcome_home_vote.bib"
 link-citations: yes
@@
     - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+    - \usepackage{setspace}
 abstract: |
@@
 Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
-This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure black Americans’ view of the democratic process.
+This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans’ view of the democratic process.
 ---
@@
 \pagebreak
+\doublespacing
 ```{r setup, echo = F, message = F, warning = F}
@@
 The Supreme Court, in cases such as *Richardson v.
-Ramirez* (1974), has upheld states' right to do just that.
+Ramirez* (1974), has upheld this practice.
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
+Of these, the majority are no longer incarcerated[@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
 
@@
 
-On April 18^th^, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+In the spring of 2018,^[Although the executive order was signed on April 18^th^, it did not actually go into effect until May 18^th^.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
 Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
@@
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
 This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
@@
 
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer, rather than simply at the end of a sentence and communicated through the mail, Executive Order 181 offers us the opportunity to build on existing research.
+Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
+Or do parole officers exhibit biases against their stewards, making them ineffective conduits for restoration?
+
 # Turnout Among the Formerly Disenfranchised {-}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from @Uggen2002.
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the number of individuals who would have voted if they were not disenfranchised, often using survey data or interviews to construct their estimates [@Uggen2004; @Drucker2005].
 
@@
 @Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states [-@Burch2010; -@Burch2011].
+Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states [-@Burch2012; -@Burch2011].
 She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
@@
 
+# Notification, Re-enfranchisement, and Turnout {-}
+
 Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
@@
 
-The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
-I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
-
 Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
@@
 @Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
+Prior to 2005 (and after 2011), individuals with felony convictions were permanently disenfranchised unless they submitted an application to the governor.
+Executive Order 42 eliminated this requirement, instead re-enfranchising automatically upon being discharged from their sentence.
+Although all formerly disenfranchised individuals were re-enfranchised upon the signing of the Executive Order, not all re-enfranchised individuals were informed of their change in status.
+@Meredith2015 leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
 @Meredith2013, however, examines states where so-called notification laws went into effect.
@@
 @Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
-@Gerber2014 conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+
+@Gerber2015 built on the quasi-experimental design of @Meredith2015 and conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
+Like @Meredith2015, they find evidence that these reminders successfully increased turnout.
 "Whatever the participatory consequences of incarceration," they conclude, "they are not in large part impossible to overcome" (924).
@@
 
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life [@Pierson1993].
-@Weaver2010 argues that "contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual" (829).
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-Such repairs may lead to increased propensity to participate in elections in the future.
+@Meredith2015 and @Gerber2015 provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
+Nevertheless, the research to date has examined the effect of mail notification on post-supervision turnout.
+In the case of New York State, formerly incarcerated individuals were being informed of their voting rights through the mail when they finished their sentence prior to the executive order.
+This study tests, therefore, not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
+
+There is reason to believe that some of the negative treatment effects of incarceration arise from the very nature of person-to-person relationships in the criminal justice system.
+@Lerman2014 argues that contact with the criminal justice system can uniquely structure democratic participation more than other types of government contact.
+"It may also be," they write, "that social benefits [arising from non-criminal justice contact with the state] are less visible to citizens because they often occur through the mail, rather than through personal contact with government agencies or officials, making it easier to disconnect social benefits from government and the political system" (93).
+Because Executive Order 181 operates *not* through the mail but rather through "personal contact with government...
+officials," its effects may differ from the mail notification programs.
+The social benefits (restoration of voting rights) might be communicated more effectively, thus increasing turnout further.
+
+Of course, reliance on a human-mediated system to deliver social benefits poses potential practical and scientific complications.
+A notification sent in the mail is a binary treatment: the letter was either sent or not sent.
+Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if parolees have a strong relationship with their parole officer, they may be strongly encouraged to register and participate.
+On the other hand, if parole officers are biased against their parolees (due either to racial hostility, personal dislike, or expected partisan affiliation) they may not provide any effective treatment at all.
+Thus, while the communication of voting rights restoration in the context of a personal relationship has the potential to deliver much larger benefits, it also opens the door to potential discrepancies in treatment.
+
+I expect that restoring voting rights while an individual is still on parole increases that individual's later propensity to vote substantially.
+This increase will operate through two mechanisms: as the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a former parolee more sure of her voting rights than a letter in the mail.
+Secondly, the interpersonal nature of the parolee-parole officer relationship is expected to more effectively repair (some of) the damage done to the former parolees interpretation of the state.
+The individual, in-person invitation to rejoin the body politic was likely effective.
 
@@
 By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1^st^, 2018, and April 17^th^, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18^th^, 2018, and January 13^th^, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1 percent of former parolees are not in the online lookup.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-
+\begin{singlespace}
 ```{r shift-dobs-chunk, echo = F}
@@
 
+# Research Design {-}
+
+In this paper, I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+All individuals discharged from parole after the start of the executive order are assigned to the "treatment" group; individuals discharged earlier are assigned to the control group.
+When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
+The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
+
+The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually "treated" by having their voting rights restored while on parole in a meeting with their parole officer.
+In other words, our natural experiment has "imperfect compliance." Noncitizens, for instance, were not eligible for rights "restoration" while on parole because in New York State they were not allowed to vote even prior to their conviction.
+Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intend-to-treat effect.
+Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
+
+Although we can identify the individuals in the assigned-to-treatment group who do not comply with their treatment status, we do not know who among the assigned-to-control group would have gone untreated if they had been assigned to the other group.
+We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
+In an ideal world, we would limit both the treatment *and* the control group to individuals who were eligible to have their rights restored while on parole; we do not live in such a world.
+Worse yet, there is obviously a relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are, of course, both ineligible to have voting rights bestowed and cannot vote.
+
+The standard way of dealing with imperfect compliance in an experimental setting is to leverage the (random) assignment to treatment as an instrument for the (nonrandom) actual compliance [@Angrist1996].
+Such an approach has been widely used in the context of voter turnout [e.g.
+@Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
+@Gerber2000 provides a helpful example of why this two-step process is necessary.
+The authors test the efficacy of phone calls on increasing turnout.
+One group is assigned to be the control; the authors do not attempt to call them.
+Another set of voters receive a phone call, but only some of them are actually contacted, or "treated" with the call.
+The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
+Then "treatment" would simply be identifying a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
+
+The same process is likely at play here.
+The control group includes individuals that *would have* had their rights restored at an in-person meeting if they had been discharged later; it also includes individuals who would not have received the treatment thanks to their citizenship status.
+By comparing turnout among the actually treated to the control group, we may simply be comparing the turnout of a group of *citizens* to the turnout of a group that includes both citizens *and* noncitizens.
+Although a positive "treatment effect" is likely in this context, it would be unable to speak to causality.
+
+Figure \ref{fig:flow} demonstrates the buckets into which former parolees can fall.
+Although I argue that a parolee's assignment to the control or treatment categories is random and uncorrelated with turnout except via the treatment, whether a treated former parolee is treated or is a noncomplier is not random and is associated with their propensity to vote through mechanisms other than the treatment.
+
+```{r flow, echo = FALSE, message = F, warning = F, fig.width = 4, fig.height = 4, fig.cap = "\\label{fig:flow}Categorization of Former Parolees", fig.pos = "H", fig.align = "center"}
+knitr::include_graphics("../../temp/flowchart.pdf")
+```
+
+Using the natural experiment with imperfect compliance setup, I begin by estimating the intend-to-treat effect of the policy change.
+This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
+I use turnout in the 2018 election as the dependent variable in these logistic regressions.
+
+After estimating the intend-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
+As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
+To explore this potential, I also explore whether the treatment had different effects for different subgroups of the formerly incarcerated population.
+I finally explore whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+
 # Results {-}
@@
 Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
+This curve is fit on monthly data running from January 2010 through May 2016, and extended through October 2016.
 
@@
 Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
+This curve is fit on monthly data running from January 2012 through May 2018, and extended through October 2018.
 ```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
@@
 
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated for individuals discharged in 2017 or 2018 before the executive order was enacted.
-
-```{r to-18-recent-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18-recent}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart_recent.rds")
-print(j)
-```
-Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1^st^, 2017, and April 17^th^, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
-
+However, Table \ref{tab:to-18-logit-short} indicates that this relationship breaks down in the short-run.
+These logistic models include individuals last discharged from parole between January 1^st^, 2017, and May 18^th^, 2018.
+\begin{singlespace}
 ```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table3.tex"}
-
+\end{singlespace}
 The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[[Appendix A] provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
-
-Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-According to the rights restoration data, it is not until May 18^th^ that the majority of individuals discharged from parole had their rights restored prior to discharge.
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[The Supplemental Information provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
 
@@
 To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
-
+\begin{singlespace}
 ```{r rd-demo-chunk, echo = F, warning = F}
@@
 \input{"../../temp/table_whatever2.tex"}
-
+\end{singlespace}
 In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
-
+\begin{singlespace}
 ```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table4.tex"}
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
+\end{singlespace}
+Model 1 in Table \ref{tab:to-18-logit} controls only for whether an individual was discharged after Executive Order 181 went into effect.
 Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
@@
 Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model2[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
-The turnout rate for the control group was `r format(round(100*readRDS("./temp/control_to_18.rds"), digits = 1), nsmall = 1)` percent.
+The turnout rate for the control group was `r format(round(100*readRDS("./temp/control_to_18.rds"), digits = 1), nsmall = 1)` percent, indicating a treatment effect of `r round((exp(model3[["coefficients"]][["finished_postTRUE"]])-1) * readRDS("./temp/control_to_18.rds"), digits = 4) * 100` percentage points.
 
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
 To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+Not everyone discharged after the policy changed had their rights restored.
+In other words, not everyone "assigned" to the treatment group was actually treated.
+We therefore have a noncompliance problem.
 
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group "complied." Noncitizens and individuals who violated their terms of parole were not eligible for rights restoration, and therefore did not "comply" with the treatment.
-Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
-
-An individual's propensity to vote can be expressed using the following equation:
-$$ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} $$
-Where *Y~i~* is 1 if individual *i* cast a ballot, *X~1i~* is the probability that individual *i* was eligible to have his rights restored prior to discharge from parole, and *X~2i~* is 1 if individual *i* actually had his rights restored prior to discharge from parole.
-*Z* is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe *X~1i~*, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if *X~1i~* and *X~2i~* (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if *b~1~* is uncorrelated with an individual's propensity to vote (*Y~i~*).
-
-Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
-Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Table \ref{tab:demos-restorees} presents demographic differences between individuals who had their rights restored and those who did not among the intend-to-treat group (we cannot tell who would and would not have complied among the control group).
-
-```{r restoree-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_restorees.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos-restorees) Demographics of \"Compliers\" and \"Non-Compliers\"", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
-\\\\Shows demographics for individuals discharged from parole \\\\between May 18th, 2018, and October 12th, 2018.}}}"
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demos-restorees} Demographics of \"Compliers\" and \"Non-Compliers\"}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever.tex"}
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'') is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that *b~1~* is very probably correlated with *Y*.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration *caused* turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach [@Angrist1996].
-Such an approach has been widely used in the context of voter turnout [e.g.
-@Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the [0, 1] interval.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population.
+The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall between 0 and 1.
 Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
 Some research [@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2014].
-The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2007].
+The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2015].
+The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2008].
 This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
@@
 For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
+\begin{singlespace}
 \input{"../../temp/iv_clean.tex"}
+\end{singlespace}
 The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
@@
 In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only black former parolees (Models 3 and 6).
+These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
 Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
-
+\begin{singlespace}
 \input{"../../temp/iv_clean_race.tex"}
-
+\end{singlespace}
 ```{r nameless-chunk, echo = F, results = "asis", message = F}
@@
 We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
+Turnout among treated non-white and Black former parolees was 3.0 and 3.3 percent, respectively.
 
-Why would the intervention have increased turnout among white individuals and have had no effect on black individuals?
+Why would the intervention have increased turnout among white individuals and have had no effect on Black individuals?
 Some of this may be explained by different propensities to vote.
-@White2019, for instance, shows that brief periods of incarceration decreases black individuals' propensity to vote by substantially more than white individuals.
-She demonstrates that, prior to incarceration, black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
-The inverse may hold true here: if black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
+@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than white individuals.
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
+The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of black individuals to vote.
-
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Because Latinos are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
+\begin{singlespace}
 ```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
@@
 \input{"../../temp/table5.tex"}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were black.
-These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, black individuals were still a third more likely to cast a ballot.
+\end{singlespace}
+Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
+These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
+Even when Latinos are excluded from this group, Black individuals were still a third more likely to cast a ballot.
 In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for black individuals, but that does not mean they are less likely to vote than other former parolees.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.^[This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation, and to that end I use turnout — not registration — as the dependent variable of interest throughout.
+The Supplemental Information demonstrates that all treatment effects discussed here also hold when the dependent variable is registration, not turnout]
 
@@
 There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
+Figure \ref{fig:to-18-chart} suggests that individuals discharged later in the treatment period had a higher propensity to vote.
 This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
@@
 It is coded as 0 for individuals discharged prior to the implementation of the executive order.^[Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.]
-
+\begin{singlespace}
 \input{"../../temp/iv_clean_months.tex"}
-
+\end{singlespace}
 Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
@@
 It could also be a story of better information.
-As @Meredith2015 and @Gerber2014 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
+As @Meredith2015 and @Gerber2015 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
 Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
@@
 Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-@WHITE2014, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2016].
+@White2015, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2017].
 
-Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
+Parole officers may more enthusiastically encourage white parolees to register to vote and cast a ballot.
 This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
@@
 \newpage
-\beginsupplement
-# Appendix A {-}
-
-In the [Instrumental Variables Approach] section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
-
-```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_16.rdata")
-
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
-                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))
-          ))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table55.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table55.tex"}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
-
-\newpage
 # References {-}

### 2020-03-04 markdown/part2/part2_standalone.tex
write

@@
 \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+\usepackage{setspace}
 \usepackage{booktabs}
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{January 23, 2020}
+\date{March 04, 2020}
 
@@
 Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
-This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure black Americans' view of the democratic process.
+This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans' view of the democratic process.
 \end{abstract}
@@
 \pagebreak
+\doublespacing
 
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
 The Supreme Court, in cases such as \emph{Richardson v.
-Ramirez} (1974), has upheld states' right to do just that.
+Ramirez} (1974), has upheld this practice.
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
+Of these, the majority are no longer incarcerated(Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
 
@@
 
-On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+In the spring of 2018,\footnote{Although the executive order was signed on April 18\textsuperscript{th}, it did not actually go into effect until May 18\textsuperscript{th}.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
 Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
@@
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
+These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
 This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
 However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
+
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer, rather than simply at the end of a sentence and communicated through the mail, Executive Order 181 offers us the opportunity to build on existing research.
+Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
+Or do parole officers exhibit biases against their stewards, making them ineffective conduits for restoration?
 
@@
 
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the number of individuals who would have voted if they were not disenfranchised, often using survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-
 In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
 Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}).
+Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states (\protect\hyperlink{ref-Burch2012}{2012}, \protect\hyperlink{ref-Burch2011}{2011}).
 She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
@@
 
+\hypertarget{notification-re-enfranchisement-and-turnout}{%
+\section*{Notification, Re-enfranchisement, and Turnout}\label{notification-re-enfranchisement-and-turnout}}
+\addcontentsline{toc}{section}{Notification, Re-enfranchisement, and Turnout}
+
 Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
@@
 
-The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
-I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
-
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2006}).
+Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2008}).
 Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
 Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
+Prior to 2005 (and after 2011), individuals with felony convictions were permanently disenfranchised unless they submitted an application to the governor.
+Executive Order 42 eliminated this requirement, instead re-enfranchising automatically upon being discharged from their sentence.
+Although all formerly disenfranchised individuals were re-enfranchised upon the signing of the Executive Order, not all re-enfranchised individuals were informed of their change in status.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
 Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
@@
 Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
+
 Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
+(\protect\hyperlink{ref-Gerber2015}{2015}) built on the quasi-experimental design of Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
+Like Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}), they find evidence that these reminders successfully increased turnout.
 ``Whatever the participatory consequences of incarceration,'' they conclude, ``they are not in large part impossible to overcome'' (924).
@@
 
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that ``contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual'' (829).
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-Such repairs may lead to increased propensity to participate in elections in the future.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
+(\protect\hyperlink{ref-Gerber2015}{2015}) provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
+Nevertheless, the research to date has examined the effect of mail notification on post-supervision turnout.
+In the case of New York State, formerly incarcerated individuals were being informed of their voting rights through the mail when they finished their sentence prior to the executive order.
+This study tests, therefore, not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
+
+There is reason to believe that some of the negative treatment effects of incarceration arise from the very nature of person-to-person relationships in the criminal justice system.
+Lerman and Weaver (\protect\hyperlink{ref-Lerman2014}{2014}) argues that contact with the criminal justice system can uniquely structure democratic participation more than other types of government contact.
+``It may also be,'' they write, ``that social benefits {[}arising from non-criminal justice contact with the state{]} are less visible to citizens because they often occur through the mail, rather than through personal contact with government agencies or officials, making it easier to disconnect social benefits from government and the political system'' (93).
+Because Executive Order 181 operates \emph{not} through the mail but rather through ``personal contact with government\ldots{} officials,'' its effects may differ from the mail notification programs.
+The social benefits (restoration of voting rights) might be communicated more effectively, thus increasing turnout further.
+
+Of course, reliance on a human-mediated system to deliver social benefits poses potential practical and scientific complications.
+A notification sent in the mail is a binary treatment: the letter was either sent or not sent.
+Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if parolees have a strong relationship with their parole officer, they may be strongly encouraged to register and participate.
+On the other hand, if parole officers are biased against their parolees (due either to racial hostility, personal dislike, or expected partisan affiliation) they may not provide any effective treatment at all.
+Thus, while the communication of voting rights restoration in the context of a personal relationship has the potential to deliver much larger benefits, it also opens the door to potential discrepancies in treatment.
+
+I expect that restoring voting rights while an individual is still on parole increases that individual's later propensity to vote substantially.
+This increase will operate through two mechanisms: as the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a former parolee more sure of her voting rights than a letter in the mail.
+Secondly, the interpersonal nature of the parolee-parole officer relationship is expected to more effectively repair (some of) the damage done to the former parolees interpretation of the state.
+The individual, in-person invitation to rejoin the body politic was likely effective.
 
@@
 By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1 percent of former parolees are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 
@@
 
+\begin{singlespace}
 \begin{table}[H]
@@
 \end{table}
+\end{singlespace}
 
@@
 
+\hypertarget{research-design}{%
+\section*{Research Design}\label{research-design}}
+\addcontentsline{toc}{section}{Research Design}
+
+In this paper, I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+All individuals discharged from parole after the start of the executive order are assigned to the ``treatment'' group; individuals discharged earlier are assigned to the control group.
+When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
+The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
+
+The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually ``treated'' by having their voting rights restored while on parole in a meeting with their parole officer.
+In other words, our natural experiment has ``imperfect compliance.'' Noncitizens, for instance, were not eligible for rights ``restoration'' while on parole because in New York State they were not allowed to vote even prior to their conviction.
+Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intend-to-treat effect.
+Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
+
+Although we can identify the individuals in the assigned-to-treatment group who do not ``comply'' with their treatment status, we do not know who among the ``assigned-to-control'' group would have gone untreated if they had been assigned to the other group.
+We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
+In an ideal world, we would limit both the treatment \emph{and} the control group to individuals who were eligible to have their rights restored while on parole; we do not live in such a world.
+Worse yet, there is obviously a relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are, of course, both ineligible to have voting rights bestowed and cannot vote.
+
+The standard way of dealing with imperfect compliance in an experimental setting is to leverage the (random) assignment to treatment as an instrument for the (nonrandom) actual compliance (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
+Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
+Gerber and Green (\protect\hyperlink{ref-Gerber2000}{2000}) provides a helpful example of why this two-step process is necessary.
+The authors test the efficacy of phone calls on increasing turnout.
+One group is assigned to be the control; the authors do not attempt to call them.
+Another set of voters receive a phone call, but only some of them are actually contacted, or ``treated'' with the call.
+The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
+Then ``treatment'' would simply be identifying a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
+
+The same process is likely at play here.
+The control group includes individuals that \emph{would have} had their rights restored at an in-person meeting if they had been discharged later; it also includes individuals who would not have received the treatment thanks to their citizenship status.
+By comparing turnout among the actually treated to the control group, we may simply be comparing the turnout of a group of \emph{citizens} to the turnout of a group that includes both citizens \emph{and} noncitizens.
+Although a positive ``treatment effect'' is likely in this context, it would be unable to speak to causality.
+
+Figure \ref{fig:flow} demonstrates the buckets into which former parolees can fall.
+Although I argue that a parolee's assignment to the control or treatment categories is random and uncorrelated with turnout except via the treatment, whether a treated former parolee is treated or is a noncomplier is not random and is associated with their propensity to vote through mechanisms other than the treatment.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{../../temp/flowchart}
+
+}
+
+\caption{\label{fig:flow}Categorization of Former Parolees}\label{fig:flow}
+\end{figure}
+
+Using the natural experiment with imperfect compliance setup, I begin by estimating the intend-to-treat effect of the policy change.
+This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
+I turnout in the 2018 election as the dependent variable in these logistic regressions.
+
+After estimating the intend-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
+As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
+To explore this potential, I also explore whether the treatment had different effects for different subgroups of the formerly incarcerated population.
+I finally explore whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+
 \hypertarget{results}{%
@@
 Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
+This curve is fit on monthly data running from January 2010 through May 2016, and extended through October 2016.
 
@@
 Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
+This curve is fit on monthly data running from January 2012 through May 2018, and extended through October 2018.
 
@@
 
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated for individuals discharged in 2017 or 2018 before the executive order was enacted.
+However, Table \ref{tab:to-18-logit-short} indicates that this relationship breaks down in the short-run.
+These logistic models include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018.
 
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-recent-chart-1}
-
-}
-
-\caption{\label{fig:to-18-recent}Turnout in 2018 Midterm Election}\label{fig:to-18-recent-chart}
-\end{figure}
-
-Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
+\begin{singlespace}
 
 \input{"../../temp/table3.tex"}
+\end{singlespace}
 
@@
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
-
-Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-According to the rights restoration data, it is not until May 18\textsuperscript{th} that the majority of individuals discharged from parole had their rights restored prior to discharge.
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
 
@@
 
+\begin{singlespace}
+
 \input{"../../temp/table_whatever2.tex"}
+\end{singlespace}
 
@@
 
+\begin{singlespace}
+
 \input{"../../temp/table4.tex"}
+\end{singlespace}
 
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
+Model 1 in Table \ref{tab:to-18-logit} controls only for whether an individual was discharged after Executive Order 181 went into effect.
 Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
@@
 Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 34.5 and 35 percent.
-The turnout rate for the control group was 2.1 percent.
+The turnout rate for the control group was 2.1 percent, indicating a treatment effect of 0.74 percentage points.
 
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
 To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+Not everyone discharged after the policy changed had their rights restored.
+In other words, not everyone ``assigned'' to the treatment group was actually treated.
+We therefore have a noncompliance problem.
 
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Noncitizens and individuals who violated their terms of parole were not eligible for rights restoration, and therefore did not ``comply'' with the treatment.
-Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
-
-An individual's propensity to vote can be expressed using the following equation:
-\[ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} \]
-Where \emph{Y\textsubscript{i}} is 1 if individual \emph{i} cast a ballot, \emph{X\textsubscript{1i}} is the probability that individual \emph{i} was eligible to have his rights restored prior to discharge from parole, and \emph{X\textsubscript{2i}} is 1 if individual \emph{i} actually had his rights restored prior to discharge from parole.
-\emph{Z} is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe \emph{X\textsubscript{1i}}, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if \emph{X\textsubscript{1i}} and \emph{X\textsubscript{2i}} (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if \emph{b\textsubscript{1}} is uncorrelated with an individual's propensity to vote (\emph{Y\textsubscript{i}}).
-
-Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
-Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Table \ref{tab:demos-restorees} presents demographic differences between individuals who had their rights restored and those who did not among the intend-to-treat group (we cannot tell who would and would not have complied among the control group).
-
-\input{"../../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'\,') is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that \emph{b\textsubscript{1}} is very probably correlated with \emph{Y}.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration \emph{caused} turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'\,' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
+Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population.
+The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
+Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall between 0 and 1.
 Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
 Some research (Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
-The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2007}{2007}).
+The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2015}{2015}).
+The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2008}{2008}).
 This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
@@
 
+\begin{singlespace}
 \input{"../../temp/iv_clean.tex"}
+\end{singlespace}
 
@@
 In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only black former parolees (Models 3 and 6).
+These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
 Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
 
+\begin{singlespace}
 \input{"../../temp/iv_clean_race.tex"}
+\end{singlespace}
 
@@
 We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
+Turnout among treated non-white and Black former parolees was 3.0 and 3.3 percent, respectively.
 
-Why would the intervention have increased turnout among white individuals and have had no effect on black individuals?
+Why would the intervention have increased turnout among white individuals and have had no effect on Black individuals?
 Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases black individuals' propensity to vote by substantially more than white individuals.
-She demonstrates that, prior to incarceration, black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
+White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than white individuals.
+She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
+The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
 
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of black individuals to vote.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Because Latinos are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
+
+\begin{singlespace}
 
 \input{"../../temp/table5.tex"}
+\end{singlespace}
 
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were black.
-These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, black individuals were still a third more likely to cast a ballot.
+Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
+These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
+Even when Latinos are excluded from this group, Black individuals were still a third more likely to cast a ballot.
 In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
+Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for black individuals, but that does not mean they are less likely to vote than other former parolees.
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.\footnote{This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation, and to that end I use turnout --- not registration --- as the dependent variable of interest throughout.
+The Supplemental Information demonstrates that all treatment effects discussed here also hold when the dependent variable is registration, not turnout}
 
@@
 
+\begin{singlespace}
 \input{"../../temp/iv_clean_months.tex"}
+\end{singlespace}
 
@@
 As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
-Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+(\protect\hyperlink{ref-Gerber2015}{2015}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
+Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
 When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
@@
 Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-WHITE2014}{2014}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2016}{2016}).
+White, Nathan, and Faller (\protect\hyperlink{ref-White2015}{2015}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2017}{2017}).
 
-Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
+Parole officers may more enthusiastically encourage white parolees to register to vote and cast a ballot.
 This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
@@
 
+\begin{singlespace}
+
 \input{"../../temp/table55.tex"}
+\end{singlespace}
 
@@
 \leavevmode\hypertarget{ref-Angrist2008}{}%
-Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
+Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics: An Empiricist's Companion}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
 
@@
 
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
+\leavevmode\hypertarget{ref-Burch2011}{}%
+Burch, Traci.
+2011.
+``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law \& Society Review} 45 (3): 699--730.
+\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
 
-\leavevmode\hypertarget{ref-Burch2010}{}%
-Burch, Traci.
-2010.
+\leavevmode\hypertarget{ref-Burch2012}{}%
+---------.
+2012.
 ``Did Disfranchisement Laws Help Elect President Bush?
@@
 
-\leavevmode\hypertarget{ref-Dong2014}{}%
+\leavevmode\hypertarget{ref-Dong2015}{}%
 Dong, Yingying, and Arthur Lewbel.
-2014.
+2015.
 ``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
@@
 ``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
-Sentencing Project.
-\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
+The Sentencing Project.
+\url{https://www.prisonpolicy.\%20org/scans/sp/fd_studiesvotingbehavior.pdf}.
 
-\leavevmode\hypertarget{ref-Einstein2016}{}%
+\leavevmode\hypertarget{ref-Einstein2017}{}%
 Einstein, Katherine Levine, and David M.
 Glick.
-2016.
+2017.
 ``Does Race Affect Access to Government Services?
@@
 ``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
-\url{https://doi.org/10.1017/s0007123400000247}.
+\url{https://www.jstor.org/stable/194285}.
 
@@
 2009.
-``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data.'' \emph{Social Science Quarterly} 90 (2): 262--73.
+``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data*.'' \emph{Social Science Quarterly} 90 (2): 262--73.
 \url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
@@
 Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
-\url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
+\url{https://doi.org/10.1016/S0047-2352(00)00044-1}.
+
+\leavevmode\hypertarget{ref-bcj_laws}{}%
+Justice, Brennan Center for.
+2019.
+``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
+\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
 
 \leavevmode\hypertarget{ref-Lassen2004}{}%
-Lassen, David Dreyer. 2004. ``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' \emph{SSRN Electronic Journal}. \url{https://doi.org/10.2139/ssrn.475821}.
+Lassen, David Dreyer.
+2004.
+``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' SSRN Scholarly Paper ID 475821.
+Rochester, NY: Social Science Research Network.
+\url{https://papers.ssrn.com/abstract=475821}.
+
+\leavevmode\hypertarget{ref-Lerman2014}{}%
+Lerman, Amy E., and Vesla M.
+Weaver.
+2014.
+\emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}.
+Chicago Studies in American Politics.
+Chicago ; London: The University of Chicago Press.
 
 \leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
+Manza, Jeff, and Christopher Uggen. 2008. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. Studies in Crime and Public Policy. New York: Oxford University Press.
 
@@
 2013.
-``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 220--49.
+``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
 \url{https://doi.org/10.1177/0002716213502931}.
@@
 ``Does Education Improve Citizenship?
-Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9-10): 1667--95.
+Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9): 1667--95.
 \url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
-
-\leavevmode\hypertarget{ref-Pierson1993}{}%
-Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
 
@@
 
-\leavevmode\hypertarget{ref-Terza2007}{}%
+\leavevmode\hypertarget{ref-Terza2008}{}%
 Terza, Joseph V., W.
@@
 Dismuke.
-2007.
+2008.
 ``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
@@
 ``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-Sentencing Project.
+The Sentencing Project.
 \url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
 
-\leavevmode\hypertarget{ref-Uggen2004}{}%
-Uggen, Christopher, and Jeff Manza.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
-
-\leavevmode\hypertarget{ref-Uggen2002}{}%
----------.
-2002.
-``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
-\url{https://doi.org/10.2307/3088970}.
-
 \leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
+Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/S0003055410000456}.
 
@@
 The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/s000305541800093x}.
+\url{https://doi.org/10.1017/S000305541800093X}.
 
-\leavevmode\hypertarget{ref-WHITE2014}{}%
-White, Ariel, Noah Nathan, and Julie Faller.
-2014.
+\leavevmode\hypertarget{ref-White2015}{}%
+White, Ariel R., Noah L.
+Nathan, and Julie K.
+Faller.
+2015.
 ``What Do I Need to Vote?
 Bureaucratic Discretion and Discrimination by Local Election Officials.'' \emph{American Political Science Review} 109 (1): 129--42.
-\url{https://doi.org/10.1017/s0003055414000562}.
+\url{https://doi.org/10.1017/S0003055414000562}.
 
 \leavevmode\hypertarget{ref-Wooldridge2010}{}%
-Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
+Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. 2nd ed. Cambridge, Mass: MIT Press.
 \end{cslreferences}

### 2020-03-04 markdown/part2/registration_regs.Rmd
write

---
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
    latex_engine: xelatex
    keep_tex: true
bibliography: "../welcome_home_vote.bib"
link-citations: yes
fontsize: 12pt
header-includes:
    - \usepackage{rotating}
    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
    - \usepackage{setspace}
---
\doublespace
```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(root.dir = "../..", dev = "cairo_pdf")
library(systemfit)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(tidyverse)
library(extrafont)
library(kableExtra)
library(data.table)
library(stargazer)
loadfonts()
```
# Turnout Among Parolees in 2016 {-}

In the Instrumental Variables Approach section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote. Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals. Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
\begin{singlespace}
```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
load("./temp/individual_turnout_16.rdata")

stargazer(model1, model2, model3,
          header = F,
          type = "latex", notes.align = "l",
          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
                               "Time Spent on Parole (Years)"),
          dep.var.labels = c("Cast Ballot in 2016 Election"),
          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
          table.placement = "H",
          omit = c("race", "felony"),
          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
                         c("Felony Class FE", "", "", "X")),
          table.layout = "-dcm#-t-a-s-n",
          out = "./temp/test.tex",
          out.header = F,
          notes = "TO REPLACE",
          se = list(
            sqrt(diag(vcovHC(model1, type = "HC0"))),
            sqrt(diag(vcovHC(model2, type = "HC0"))),
            sqrt(diag(vcovHC(model3, type = "HC0")))
          ))
j <- fread("./temp/test.tex", header = F, sep = "+")

note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"

j <- j %>% 
  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
  filter(!grepl("Note:", V1)) %>% 
  mutate(n = row_number())

insert1 <- "\\fontsize{10}{12}\\selectfont"

j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
  arrange(n) %>% 
  select(-n)

write.table(j, "./temp/table55.tex", quote = F, col.names = F,
            row.names = F)
```
\input{"../../temp/table55.tex"}
\end{singlespace}
Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot. The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.

# Treatment Effect on Registration {-}

In the body of this paper, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote. The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occured when the individual was no longer on parole. Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.

In this Supplemental Appendix, I re-estimate the primary analyses found in Tables 5, 6, and 8 in the main body of this paper. Here, however, I use registration as the dependent variable, rather than turnout in 2018. As before, the dependent variable is binary: an individual is either registered or not registered. As such, the functional forms remain the same. As before, the marginal effects of the probit models are presented with other covariates held at their means.

\begin{singlespace}
\input{"../../temp/iv_clean_reg.tex"}
\end{singlespace}

The 2SLS model indicates that voting rights restoration increased the registration rate of former parolees by 1.37 percentange points, while the biprobit model estimates that the effect was between 1.25 and 1.57 percentage points. Given that 5.5 percent of parolees in the control group were registered to vote, this implies an increase of between 23 and 29 percent.

\begin{singlespace}
\input{"../../temp/iv_clean_race_reg.tex"}
\end{singlespace}

As with turnout, voting rights restoration appears to have increased registrations only among white former parolees. The estimated treatment effect is between 2.3 and 2.4 percentage points, off a registration rate among the control, white individuals of 4.4 percent. No significant effect is detected among non-white or Black former parolees, whose control groups had base registration rates of 5.9 and 7.2 percent, respectively. As with turnout, Black former parolees had much higher base registration rates than whites.

\begin{singlespace}
\input{"../../temp/iv_clean_months_reg.tex"}
\end{singlespace}

Finally, we see that the longer an individual was on parole after her voting rights were restored, the more likely she was to be registered. Each additional month on parole after voting rights were restored is associated with an increase in registration of between 0.5 and 0.9 percentage points.

### 2020-03-04 markdown/part2/registration_regs.tex
write

--- 2020-01-28 markdown/part2/part2_standalone.tex
@@
 \hypersetup{
-  pdftitle={Welcome Home --- Now Vote!},
-  pdfauthor={Kevin Morris},
   hidelinks,
@@
 \usepackage{xcolor}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
 
-\title{Welcome Home --- Now Vote!\thanks{Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
-The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
-All errors are my responsibility.}}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Voting Rights Restoration and Post-Supervision Participation}
-\author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{January 23, 2020}
+\author{}
+\date{\vspace{-2.5em}}
 
 \begin{document}
-\maketitle
-\begin{abstract}
-This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
-In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to parolees increases their post-supervision propensity to cast a ballot.
-I demonstrate that individuals whose rights were restored while still on parole were more likely to participate than those whose rights were automatically restored upon completion of parole.
-This group-level effect, however, masks race-specific effects.
-Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
-This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure black Americans' view of the democratic process.
-\end{abstract}
 
-\pagenumbering{gobble}
-\pagebreak
+\doublespace
 
-\pagenumbering{arabic}
+\hypertarget{turnout-among-parolees-in-2016}{%
+\section*{Turnout Among Parolees in 2016}\label{turnout-among-parolees-in-2016}}
+\addcontentsline{toc}{section}{Turnout Among Parolees in 2016}
 
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
-This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as \emph{Richardson v.
-Ramirez} (1974), has upheld states' right to do just that.
-Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated, but are living and working in their communities (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
-
-There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}; but see Gerber et al.
-\protect\hyperlink{ref-Gerber2017}{2017}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
-The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
-To address only the formal laws contributing to disenfranchisement without also interrogating efforts to boost post-supervision participation risks leaving much of the system of effective disenfranchisement undisturbed.
-New York State offers us the opportunity to test how the timing of the re-instation of voting rights structures post-supervision participation.
-
-Prior to 2018, New Yorkers convicted of felony offenses and sentenced to prison were disenfranchised until they had completed all terms of their sentence --- their period of incarceration as well as any parole term.
-For New Yorkers on life parole or sentenced to life in prison, this law resulted in effective lifetime disenfranchisement.
-New Yorkers sentenced to felony probation, on the other hand, did not lose their voting rights.
-
-On April 18\textsuperscript{th}, 2018, Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
-
-There is reason to believe that the executive order may also have increased the political participation of formerly disenfranchised individuals.
-Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
-In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
-Thus, the executive order demands not only that re-enfranchised individuals receive in-person notification of their voting rights, but also that parole officers prioritize their registration.
-
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absense of the executive order.
-This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
-
-\hypertarget{turnout-among-the-formerly-disenfranchised}{%
-\section*{Turnout Among the Formerly Disenfranchised}\label{turnout-among-the-formerly-disenfranchised}}
-\addcontentsline{toc}{section}{Turnout Among the Formerly Disenfranchised}
-
-In the aftermath of the 2000 presidential election, academic interest in the political implications of felony disenfranchisement was stirred thanks to a paper from Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}).
-George W.
-Bush's margin of victory in Florida in 2000 was famously just 537 votes.
-In their 2002 paper, Uggen and Manza estimate the likely partisan composition of the disenfranchised population with felony convictions in their past.
-They estimate that if this group had been allowed to vote they would have supported Al Gore by a wide margin.
-Their enfranchisement, Uggen and Manza argued, would have tipped the presidential contest and resulted in the election of Al Gore.
-They based their estimates on the voting patterns of eligible individuals who were demographically similar to the disenfranchised population.
-Though much of the research conducted since their 2002 study has pushed back against some of their key assumptions (namely, that formerly incarcerated individuals turn out at the same rate as demographically-similar individuals who have not been incarcerated), Uggen and Manza convincingly demonstrated that felony disenfranchisement can have material political consequences.
-In the years after Uggen and Manza published their paper, scholars sought to investigate the number of individuals who would have voted if they were not disenfranchised, often using survey data or interviews to construct their estimates (Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states (\protect\hyperlink{ref-Burch2010}{2010}, \protect\hyperlink{ref-Burch2011}{2011}).
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she estimated the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
-
-Much of this literature has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}) uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that ``it appears that spending time in prison does not have large negative effects on subsequent participation'' (1144).
-
-White (\protect\hyperlink{ref-White2019}{2019}), however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
-By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration \emph{causes} low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
-
-The literature on turnout among formerly incarcerated individuals leads me to hypothesize that rights restoration prior to parole discharge will boost turnout among formerly incarcerated individuals.
-I expect that this increased propensity to vote will arise due both thanks to better clarity about individual eligibility to participate, but also because rights restoration may repair some of the distrust of the state engendered by incarceration.
-
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2006}).
-Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
-They find that individuals who received letters explicitly informing them of their re-enfranchisement were more likely to cast ballots in the next election than those who did not.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
-Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) conducted a field experiment in Connecticut in advance of the 2012 presidential election, finding that sending mailers to individuals to remind them of their voting rights was successful at increasing turnout among this population.
-``Whatever the participatory consequences of incarceration,'' they conclude, ``they are not in large part impossible to overcome'' (924).
-Even if incarceration does not decrease individuals' propensity to vote, there is reason to believe that reminding formerly incarcerated individuals of their rights increases participation.
-
-Research also indicates that individuals who have negative interactions with the state are less likely to participate in civic life (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that ``contact with the institutions of criminal justice is important in structuring patterns of participation long assumed in the dominant literature to stem primarily from aspects of the individual'' (829).
-To the extent that parole officers are accurately informing their parolees of their newly restored voting rights, Executive Order 181 is likely to bring about a positive interaction between the parolee and the government.
-Rather than simply have one's rights restored upon completion of sentence, Executive Order 181 may lead parolees to feel individually invited back into the democratic process --- an invitation that may be successful at repairing some of the negative associations created through incarceration.
-Such repairs may lead to increased propensity to participate in elections in the future.
-
-\hypertarget{data}{%
-\section*{Data}\label{data}}
-\addcontentsline{toc}{section}{Data}
-
-\hypertarget{criminal-justice-data}{%
-\subsection*{Criminal Justice Data}\label{criminal-justice-data}}
-\addcontentsline{toc}{subsection}{Criminal Justice Data}
-
-The criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of commitment; and others.
-This analysis is limited to individuals incarcerated for felony offenses.
-Individuals convicted of misdemeanors are not disenfranchised in New York State.
-These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These data are used to determine when individuals were incarcerated or on parole, when they finished their parole supervision, and demographic information such as age and race.
-
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-For individuals who finished parole between January 1\textsuperscript{st}, 2018, and April 17\textsuperscript{th}, 2018, 1.0 percent are not in the lookup tool.
-For those discharged from parole between April 18\textsuperscript{th}, 2018, and January 13\textsuperscript{th}, 2019 (the latest date of the parole records), 1.2 percent of individuals are not found in the lookup tool.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
-
-\hypertarget{voter-file-data}{%
-\subsection*{Voter File Data}\label{voter-file-data}}
-\addcontentsline{toc}{subsection}{Voter File Data}
-
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from March 3\textsuperscript{rd}, 2019.
-
-\hypertarget{matching}{%
-\subsection*{Matching}\label{matching}}
-\addcontentsline{toc}{subsection}{Matching}
-
-Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a ``match,'' records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
-
-\begin{itemize}
-\tightlist
-\item
-  Middle names are identical. If neither set of records includes a middle name, this condition is met.
-\item
-  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-\item
-  A middle name or middle initial in one set of records, and a missing middle name in the other set.
-\end{itemize}
-
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
-
-There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the parole discharge dataset to create false records.
-Comparing the number of matches between these ``fake'' discharge records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
-This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 69,644\\
-Birthdate + 35 Days & 502\\
-Birthdate - 35 Days & 426\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
-
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
-
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.8 percent cast a ballot in the 2016 election, while just 4.3 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through April 2016, and extended through October 2016.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-16-chart-1}
-
-}
-
-\caption{\label{fig:to-16}Turnout in 2016 Presidential Election}\label{fig:to-16-chart}
-\end{figure}
-
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through April 2018, and extended through October 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-chart-1}
-
-}
-
-\caption{\label{fig:to-18}Turnout in 2018 Midterm Election}\label{fig:to-18-chart}
-\end{figure}
-
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- \emph{were} more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
-
-\hypertarget{individual-level-turnout-regressions}{%
-\subsection*{Individual-Level Turnout Regressions}\label{individual-level-turnout-regressions}}
-\addcontentsline{toc}{subsection}{Individual-Level Turnout Regressions}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being ``assigned'' to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, the relationship between time-off-parole and propensity to vote is far weaker in the short term.
-Figure \ref{fig:to-18-recent} indicates that parole discharge date and turnout rates in the 2018 midterm election are not correlated for individuals discharged in 2017 or 2018 before the executive order was enacted.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-recent-chart-1}
-
-}
-
-\caption{\label{fig:to-18-recent}Turnout in 2018 Midterm Election}\label{fig:to-18-recent-chart}
-\end{figure}
-
-Formalizing this chart into an individual-level logistic model demonstrates that time since discharge is not correlated with turnout in the short run.
-The models in Table \ref{tab:to-18-logit-short} include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and April 17\textsuperscript{th}, 2018 --- that is to say, all individuals discharged prior to the implementation of the executive order.
-
-\input{"../../temp/table3.tex"}
-
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017 and 2018 prior to the executive order as a control group is feasible.
-
-Although Governor Cuomo signed the executive order on April 18, 2018, an examination of the individuals whose rights were ultimately restored indicates that the program did not go into full effect until later in May.
-According to the rights restoration data, it is not until May 18\textsuperscript{th} that the majority of individuals discharged from parole had their rights restored prior to discharge.
-
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are assigned ``treatment'' by the policy change.
-Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
-
-\input{"../../temp/table_whatever2.tex"}
-
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18\textsuperscript{th}, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1\textsuperscript{st}, 2017, through October 10\textsuperscript{th}, 2018 (the registration deadline in New York State).
-
-\input{"../../temp/table4.tex"}
-
-Model 1 in Table \ref{tab:to-18-logit} formalizes the trend presented in Figure \ref{fig:to-18-recent-chart} by controlling only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 34.5 and 35 percent.
-The turnout rate for the control group was 2.1 percent.
-
-\hypertarget{instrumental-variables-approach}{%
-\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
-\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
-
-Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which rights restoration increased the propensity to vote for the individuals who actually received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-
-Investigating the causal effect of rights restoration using the natural experiment conceptualization introduces a complication: not all individuals who were discharged from parole after Executive Order 181 went into effect had their rights restored --- that is to say, not all individuals assigned to the treatment group ``complied.'' Noncitizens and individuals who violated their terms of parole were not eligible for rights restoration, and therefore did not ``comply'' with the treatment.
-Moreover, there is reason to suspect that compliance is correlated with propensity to vote.
-
-An individual's propensity to vote can be expressed using the following equation:
-\[ Y_i =  b_0 + b_{1}X_{1i} + b_{2}X_{2i} + b_{3}Z{i}+ \epsilon_{i} \]
-Where \emph{Y\textsubscript{i}} is 1 if individual \emph{i} cast a ballot, \emph{X\textsubscript{1i}} is the probability that individual \emph{i} was eligible to have his rights restored prior to discharge from parole, and \emph{X\textsubscript{2i}} is 1 if individual \emph{i} actually had his rights restored prior to discharge from parole.
-\emph{Z} is a vector of other factors (such as age and race) known to influence voter turnout.
-Given that we cannot observe \emph{X\textsubscript{1i}}, we could simply choose to ignore it.
-Doing so, however, will result in consistent regression coefficients only if \emph{X\textsubscript{1i}} and \emph{X\textsubscript{2i}} (that is, eligibility to have rights restored, and actual restoration) are uncorrelated or if \emph{b\textsubscript{1}} is uncorrelated with an individual's propensity to vote (\emph{Y\textsubscript{i}}).
-
-Neither of these assumptions are valid: firstly, an individual's eligibility to have her rights restored is certainly correlated with whether or not they actually were restored because the executive order required that all eligible individuals have their voting rights restored.
-Secondly, an individual's likelihood of having his rights restored is probably correlated with his natural propensity to vote.
-Individuals who were discharged from parole but are not U.S.
-citizens, for instance, did not having voting rights restored (or, in this case, granted) after the executive order went into effect.
-This ineligibility to have voting rights granted is of course correlated with these individuals' propensity to vote: non-citizens cannot cast ballots in elections in New York State.
-Table \ref{tab:demos-restorees} presents demographic differences between individuals who had their rights restored and those who did not among the intend-to-treat group (we cannot tell who would and would not have complied among the control group).
-
-\input{"../../temp/table_whatever.tex"}
-
-Table \ref{tab:demos-restorees} makes clear that the restoration of voting rights (``compliance'\,') is strongly correlated with certain demographics that are also linked to propensity to vote.
-Men, for instance, were less likely to have their rights restored than women, and Table \ref{tab:to-18-logit} above indicates that men who have been on probation are less likely to vote than women.
-Similarly, Latinos made up more than a third of individuals whose rights were not restored, but just 19 percent of individuals who did have voting rights restored.
-This is likely a reflection of citizenship status.
-Table \ref{tab:demos-restorees} indicates that \emph{b\textsubscript{1}} is very probably correlated with \emph{Y}.
-An approach in which is we regress turnout on a dummy indicating rights restoration is therefore inappropriate: such an approach cannot tell whether rights restoration \emph{caused} turnout to increase, or whether it simply identifies individuals with a higher propensity to vote.
-
-The standard way of dealing with such a problem is an instrumental variables approach (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-A valid instrumental variable must satisfy two criteria: it must be correlated with the right-hand-side (endogenous) variable of interest, and it must be uncorrelated with dependent variable (except via the endogenous variable).
-
-In this case, such a variable is readily at hand.
-The likelihood that an individual had his voting rights restored is (in part) a function of whether he was discharged from parole after Executive Order 181 took effect.
-As demonstrated above, date of discharge from parole is not correlated with propensity to vote when we limit our analysis to individuals discharged from parole in 2017 or 2018.
-A dummy variable indicating whether an individual was discharged after the executive order was implemented, therefore, satisfies the criteria for an instrumental variable.
-
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population --- what is often described as the ``local average treatment effect.'\,' The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall on the {[}0, 1{]} interval.
-Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research (Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2014}{2014}).
-The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2007}{2007}).
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
-\input{"../../temp/iv_clean.tex"}
-
-The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.84 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.96 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 36.1 and 43.5 percent.
-
-\hypertarget{racial-variation}{%
-\subsubsection*{Racial Variation}\label{racial-variation}}
-\addcontentsline{toc}{subsubsection}{Racial Variation}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only black former parolees (Models 3 and 6).
-Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
-
-\input{"../../temp/iv_clean_race.tex"}
-
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.8 percentage points.
-Given that just 3.5 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
-
-Why would the intervention have increased turnout among white individuals and have had no effect on black individuals?
-Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases black individuals' propensity to vote by substantially more than white individuals.
-She demonstrates that, prior to incarceration, black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether black individuals turned out at higher rates.
-Because Latino voters are more likely to be non-citizens (and therefore have a lower turnout rate than black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of black individuals to vote.
-
-\input{"../../temp/table5.tex"}
-
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were black.
-These black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, black individuals were still a third more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
-
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
-The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for black individuals, but that does not mean they are less likely to vote than other former parolees.
-
-\hypertarget{variable-treatment-intensity}{%
-\subsubsection*{Variable Treatment Intensity}\label{variable-treatment-intensity}}
-\addcontentsline{toc}{subsubsection}{Variable Treatment Intensity}
-
-There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figures \ref{fig:to-18-chart} and \ref{fig:to-18-recent-chart} suggest that, among individuals who finished parole after the Executive Order went into effect, individuals who were discharged later had a higher propensity to vote.
-This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
-If the individual did not have voting rights restored, this variable takes the value 0.
-This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.\footnote{Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.}
-
-\input{"../../temp/iv_clean_months.tex"}
-
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.29 percentage points (in the 2SLS model) and 0.50 percentage points (in the 2S probit model).
-
-Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
-
-\hypertarget{discussion}{%
-\subsection*{Discussion}\label{discussion}}
-\addcontentsline{toc}{subsection}{Discussion}
-
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
-Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
-That changed in 2018.
-On October 12\textsuperscript{th} --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
-Without the executive order, every single one of these individuals would have been barred from participating.
-Though turnout among this group was low (just 832, or 3.81 percent, of these individuals successfully cast a ballot), their re-enfranchisement marks an important milestone for New York State.
-
-As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least 36.1 percent.
-This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
-Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
-
-The mechanism through which the rules change increased turnout among these individuals is not clear.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
-Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
-
-It could also be a story of better information.
-As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2014}{2014}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
-Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
-In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
-
-Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
-Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
-The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
-This discrepancy could be caused by a number of different factors.
-Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
-The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
-
-We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
-Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
-Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-WHITE2014}{2014}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2016}{2016}).
-
-Parole officers may more enthusiastically encourage White parolees to register to vote and cast a ballot.
-This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
-
-Re-enfranchising voters while they are still under formal supervision is obviously beneficial to the individuals who are on parole on election day; such policies allow them to make their voices heard.
-The case of Executive Order 181 also indicates that restoring voting rights prior to parole discharge has further benefits.
-In 2018, it increased turnout among individuals who were formally discharged from parole prior to the registration deadline, and therefore would have been eligible to vote even if their rights were not restored until the completion of their sentence.
-This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
-By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
-
-\newpage
-\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}
-
-\hypertarget{appendix-a}{%
-\section*{Appendix A}\label{appendix-a}}
-\addcontentsline{toc}{section}{Appendix A}
-
-In the \protect\hyperlink{instrumental-variables-approach}{Instrumental Variables Approach} section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
+In the Instrumental Variables Approach section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
 Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
@@
 
-\newpage
+\hypertarget{treatment-effect-on-registration}{%
+\section*{Treatment Effect on Registration}\label{treatment-effect-on-registration}}
+\addcontentsline{toc}{section}{Treatment Effect on Registration}
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+In the body of this paper, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote.
+The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occured when the individual was no longer on parole.
+Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.
 
-\hypertarget{refs}{}
-\begin{cslreferences}
-\leavevmode\hypertarget{ref-Angrist2001}{}%
-Angrist, Joshua D.
-2001.
-``Estimation of Limited Dependent Variable Models with Dummy Endogenous Regressors.'' \emph{Journal of Business \& Economic Statistics} 19 (1): 2--28.
-\url{https://doi.org/10.1198/07350010152472571}.
+In this Supplemental Appendix, I re-estimate the primary analyses found in Tables 5, 6, and 8 in the main body of this paper.
+Here, however, I use registration as the dependent variable, rather than turnout in 2018.
+As before, the dependent variable is binary: an individual is either registered or not registered.
+As such, the functional forms remain the same.
+As before, the marginal effects of the probit models are presented with other covariates held at their means.
 
-\leavevmode\hypertarget{ref-Angrist1996}{}%
-Angrist, Joshua D., Guido W.
-Imbens, and Donald B.
-Rubin.
-1996.
-``Identification of Causal Effects Using Instrumental Variables.'' \emph{Journal of the American Statistical Association} 91 (434): 444--55.
-\url{https://doi.org/10.1080/01621459.1996.10476902}.
+\begin{singlespace}
+\input{"../../temp/iv_clean_reg.tex"}
+\end{singlespace}
 
-\leavevmode\hypertarget{ref-Angrist2008}{}%
-Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
+The 2SLS model indicates that voting rights restoration increased the registration rate of former parolees by 1.37 percentange points, while the biprobit model estimates that the effect was between 1.25 and 1.57 percentage points.
+Given that 5.5 percent of parolees in the control group were registered to vote, this implies an increase of between 23 and 29 percent.
 
-\leavevmode\hypertarget{ref-Ansolabehere1999}{}%
-Ansolabehere, Stephen D., Shanto Iyengar, and Adam Simon.
-1999.
-``Replicating Experiments Using Aggregate and Survey Data: The Case of Negative Advertising and Turnout.'' \emph{American Political Science Review} 93 (4): 901--9.
-\url{https://doi.org/10.2307/2586120}.
+\begin{singlespace}
+\input{"../../temp/iv_clean_race_reg.tex"}
+\end{singlespace}
 
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
+As with turnout, voting rights restoration appears to have increased registrations only among white former parolees.
+The estimated treatment effect is between 2.3 and 2.4 percentage points, off a registration rate among the control, white individuals of 4.4 percent.
+No significant effect is detected among non-white or Black former parolees, whose control groups had base registration rates of 5.9 and 7.2 percent, respectively.
+As with turnout, Black former parolees had much higher base registration rates than whites.
 
-\leavevmode\hypertarget{ref-Burch2010}{}%
-Burch, Traci.
-2010.
-``Did Disfranchisement Laws Help Elect President Bush?
-New Evidence on the Turnout Rates and Candidate Preferences of Florida's Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
-\url{https://doi.org/10.1007/s11109-010-9150-9}.
+\begin{singlespace}
+\input{"../../temp/iv_clean_months_reg.tex"}
+\end{singlespace}
 
-\leavevmode\hypertarget{ref-Burch2011}{}%
----------.
-2011.
-``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law and Society Review} 45 (3): 699--730.
-\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
-
-\leavevmode\hypertarget{ref-Butler2011}{}%
-Butler, Daniel M., and David E.
-Broockman.
-2011.
-``Do Politicians Racially Discriminate Against Constituents?
-A Field Experiment on State Legislators.'' \emph{American Journal of Political Science} 55 (3): 463--77.
-\url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
-
-\leavevmode\hypertarget{ref-Dong2014}{}%
-Dong, Yingying, and Arthur Lewbel.
-2014.
-``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
-\url{https://doi.org/10.1080/07474938.2014.944470}.
-
-\leavevmode\hypertarget{ref-Drucker2005}{}%
-Drucker, Ernest, and Ricardo Barreras.
-2005.
-``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
-Sentencing Project.
-\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
-
-\leavevmode\hypertarget{ref-Einstein2016}{}%
-Einstein, Katherine Levine, and David M.
-Glick.
-2016.
-``Does Race Affect Access to Government Services?
-An Experiment Exploring Street-Level Bureaucrats and Access to Public Housing.'' \emph{American Journal of Political Science} 61 (1): 100--116.
-\url{https://doi.org/10.1111/ajps.12252}.
-
-\leavevmode\hypertarget{ref-Gerber2000}{}%
-Gerber, Alan S., and Donald P.
-Green.
-2000.
-``The Effects of Canvassing, Telephone Calls, and Direct Mail on Voter Turnout: A Field Experiment.'' \emph{American Political Science Review} 94 (3): 653--63.
-\url{https://doi.org/10.2307/2585837}.
-
-\leavevmode\hypertarget{ref-Gerber2014}{}%
-Gerber, Alan S., Gregory A.
-Huber, Marc Meredith, Daniel R.
-Biggers, and David J.
-Hendry.
-2014.
-``Can Incarcerated Felons Be (Re)integrated into the Political System?
-Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
-\url{https://doi.org/10.1111/ajps.12166}.
-
-\leavevmode\hypertarget{ref-Gerber2017}{}%
----------.
-2017.
-``Does Incarceration Reduce Voting?
-Evidence About the Political Consequences of Spending Time in Prison.'' \emph{The Journal of Politics} 79 (4): 1130--46.
-\url{https://doi.org/10.1086/692670}.
-
-\leavevmode\hypertarget{ref-Green2000}{}%
-Green, Donald P., and Ron Shachar.
-2000.
-``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
-\url{https://doi.org/10.1017/s0007123400000247}.
-
-\leavevmode\hypertarget{ref-Haselswerdt2009}{}%
-Haselswerdt, Michael V.
-2009.
-``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data.'' \emph{Social Science Quarterly} 90 (2): 262--73.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
-
-\leavevmode\hypertarget{ref-Hemmens2000}{}%
-Hemmens, Craig, and James W.
-Marquart.
-2000.
-``Friend or Foe?
-Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
-\url{https://doi.org/10.1016/s0047-2352(00)00044-1}.
-
-\leavevmode\hypertarget{ref-Lassen2004}{}%
-Lassen, David Dreyer. 2004. ``The Effect of Information on Voter Turnout: Evidence from a Natural Experiment.'' \emph{SSRN Electronic Journal}. \url{https://doi.org/10.2139/ssrn.475821}.
-
-\leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
-
-\leavevmode\hypertarget{ref-Meredith2013}{}%
-Meredith, Marc, and Michael Morse.
-2013.
-``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 220--49.
-\url{https://doi.org/10.1177/0002716213502931}.
-
-\leavevmode\hypertarget{ref-Meredith2015}{}%
----------.
-2015.
-``The Politics of the Restoration of Ex-Felon Voting Rights: The Case of Iowa.'' \emph{Quarterly Journal of Political Science} 10 (1): 41--100.
-\url{https://doi.org/10.1561/100.00013026}.
-
-\leavevmode\hypertarget{ref-Milligan2004}{}%
-Milligan, Kevin, Enrico Moretti, and Philip Oreopoulos.
-2004.
-``Does Education Improve Citizenship?
-Evidence from the United States and the United Kingdom.'' \emph{Journal of Public Economics} 88 (9-10): 1667--95.
-\url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
-
-\leavevmode\hypertarget{ref-Pierson1993}{}%
-Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
-
-\leavevmode\hypertarget{ref-Sondheimer2010}{}%
-Sondheimer, Rachel Milstein, and Donald P.
-Green.
-2010.
-``Using Experiments to Estimate the Effects of Education on Voter Turnout.'' \emph{American Journal of Political Science} 54 (1): 174--89.
-\url{https://doi.org/10.1111/j.1540-5907.2009.00425.x}.
-
-\leavevmode\hypertarget{ref-Terza2007}{}%
-Terza, Joseph V., W.
-David Bradford, and Clara E.
-Dismuke.
-2007.
-``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
-\url{https://doi.org/10.1111/j.1475-6773.2007.00807.x}.
-
-\leavevmode\hypertarget{ref-sentencing_2016}{}%
-Uggen, Christopher, Ryan Larson, and Sarah Shannon.
-2016.
-``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-Sentencing Project.
-\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
-
-\leavevmode\hypertarget{ref-Uggen2004}{}%
-Uggen, Christopher, and Jeff Manza.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
-
-\leavevmode\hypertarget{ref-Uggen2002}{}%
----------.
-2002.
-``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
-\url{https://doi.org/10.2307/3088970}.
-
-\leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
-
-\leavevmode\hypertarget{ref-White2019}{}%
-White, Ariel.
-2019.
-``Misdemeanor Disenfranchisement?
-The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/s000305541800093x}.
-
-\leavevmode\hypertarget{ref-WHITE2014}{}%
-White, Ariel, Noah Nathan, and Julie Faller.
-2014.
-``What Do I Need to Vote?
-Bureaucratic Discretion and Discrimination by Local Election Officials.'' \emph{American Political Science Review} 109 (1): 129--42.
-\url{https://doi.org/10.1017/s0003055414000562}.
-
-\leavevmode\hypertarget{ref-Wooldridge2010}{}%
-Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. Cambridge, MA: The MIT Press.
-\end{cslreferences}
+Finally, we see that the longer an individual was on parole after her voting rights were restored, the more likely she was to be registered.
+Each additional month on parole after voting rights were restored is associated with an increase in registration of between 0.5 and 0.9 percentage points.
 

### 2020-03-05 markdown/part2/part2_standalone.Rmd
write

@@
 subtitle: "Voting Rights Restoration and Post-Supervision Participation"
-thanks: Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
-The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
-All errors are my responsibility.
-author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
-institute: "Brennan Center for Justice"
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+date: |
+  `r format(Sys.time(), '%B %d, %Y')`
+  \linebreak
+  \linebreak
+  \linebreak
+  \linebreak
+  Word Count: 9,553
 output:
@@
 By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to parolees increases their post-supervision propensity to cast a ballot.
-I demonstrate that individuals whose rights were restored while still on parole were more likely to participate than those whose rights were automatically restored upon completion of parole.
+I demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
 This group-level effect, however, masks race-specific effects.
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
+Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated[@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
+Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
 
@@
 To address only the formal laws contributing to disenfranchisement without also interrogating efforts to boost post-supervision participation risks leaving much of the system of effective disenfranchisement undisturbed.
-New York State offers us the opportunity to test how the timing of the re-instation of voting rights structures post-supervision participation.
+New York State offers us the opportunity to test how the timing and method of the re-instation of voting rights structures post-supervision participation.
 
@@
 
-In the spring of 2018,^[Although the executive order was signed on April 18^th^, it did not actually go into effect until May 18^th^.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+In the spring of 2018,^[Although the executive order was signed on April 18^th^, it did not go into effect until May 18^th^.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+Such a move was of course good for individuals who were still on parole on election day --- they would have been disenfranchised otherwise.
+The change in policy is also beneficial for felony probationers.
+Despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, it has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
-There is reason to believe that the executive order may also have increased the political participation of formerly disenfranchised individuals.
+The executive order may also have increased the political participation of *formerly* disenfranchised individuals.
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
@@
 Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[Link: https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
+Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
 In addition to these directives, the memorandum communicates that voter registration is to receive "high priority attention," and it separately calls the program "a **priority** initiative" [emphasis in the original].
@@
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
-This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
+I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
 
-By examining the effect of rights restoration in the context of a personal relationship with a parole officer, rather than simply at the end of a sentence and communicated through the mail, Executive Order 181 offers us the opportunity to build on existing research.
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility through the mail, Executive Order 181 offers us the opportunity to build on existing research.
 Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
-Or do parole officers exhibit biases against their stewards, making them ineffective conduits for restoration?
+Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration?
 
@@
 
-Much of this literature has established that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
+Much of this literature establishes that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
 The causal effect of incarceration on participation is the subject of some debate within the field.
@@
 A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
+Whether or not incarceration *causes* low turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
 
@@
 Prior to 2005 (and after 2011), individuals with felony convictions were permanently disenfranchised unless they submitted an application to the governor.
-Executive Order 42 eliminated this requirement, instead re-enfranchising automatically upon being discharged from their sentence.
+Executive Order 42 eliminated this requirement, instead re-enfranchising individuals automatically upon sentence completion.
 Although all formerly disenfranchised individuals were re-enfranchised upon the signing of the Executive Order, not all re-enfranchised individuals were informed of their change in status.
 @Meredith2015 leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
-@Meredith2013, however, examines states where so-called notification laws went into effect.
+@Meredith2013, on the other hand, examines states where so-called notification laws went into effect.
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
@@
 
-@Gerber2015 built on the quasi-experimental design of @Meredith2015 and conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+@Gerber2015 built on the quasi-experimental design of @Meredith2015, conducting a field experiment in Connecticut in advance of the 2012 presidential election.
 In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
@@
 Nevertheless, the research to date has examined the effect of mail notification on post-supervision turnout.
-In the case of New York State, formerly incarcerated individuals were being informed of their voting rights through the mail when they finished their sentence prior to the executive order.
-This study tests, therefore, not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
+In the case of New York State, formerly incarcerated individuals were already informed of their voting rights through the mail when they finished their sentence prior to the executive order.
+In other words, the treatment identified in past research is here incorporated into the status quo, or control, condition.
+This study therefore tests not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
 
@@
 officials," its effects may differ from the mail notification programs.
-The social benefits (restoration of voting rights) might be communicated more effectively, thus increasing turnout further.
+The social benefits (restoration of voting rights) might be communicated more effectively and in ways that improve the formerly incarcerated individuals' opinions of their democratic state, thus increasing turnout further.
 
@@
 A notification sent in the mail is a binary treatment: the letter was either sent or not sent.
-Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if parolees have a strong relationship with their parole officer, they may be strongly encouraged to register and participate.
+Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if parolees have a warm relationship with their parole officer, they may be strongly encouraged to register and participate.
 On the other hand, if parole officers are biased against their parolees (due either to racial hostility, personal dislike, or expected partisan affiliation) they may not provide any effective treatment at all.
@@
 Secondly, the interpersonal nature of the parolee-parole officer relationship is expected to more effectively repair (some of) the damage done to the former parolees interpretation of the state.
-The individual, in-person invitation to rejoin the body politic was likely effective.
+Individual, in-person invitations to rejoin the body politic are likely effective.
 
@@
 Roughly 1 percent of former parolees are not in the online lookup.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1,200 individuals who were discharged between the effective date of the executive order and the registration deadline did not have voting rights restored, due largely to their status as noncitizens.
 
@@
 In this paper, I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
-All individuals discharged from parole after the start of the executive order are assigned to the "treatment" group; individuals discharged earlier are assigned to the control group.
+All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
 When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
@@
 
-The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually "treated" by having their voting rights restored while on parole in a meeting with their parole officer.
-In other words, our natural experiment has "imperfect compliance." Noncitizens, for instance, were not eligible for rights "restoration" while on parole because in New York State they were not allowed to vote even prior to their conviction.
-Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intend-to-treat effect.
+The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually treated by having their voting rights restored while on parole in a meeting with their parole officer.
+In other words, our natural experiment has "imperfect compliance" [see, for instance, @Angrist1996].
+Noncitizens, for instance, were not eligible for rights "restoration" while on parole because in New York State they were not allowed to vote even prior to their conviction.
+Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intent-to-treat effect.
 Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
@@
 We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
-In an ideal world, we would limit both the treatment *and* the control group to individuals who were eligible to have their rights restored while on parole; we do not live in such a world.
-Worse yet, there is obviously a relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are, of course, both ineligible to have voting rights bestowed and cannot vote.
+In an ideal world, we would limit both the assigned-to-treatment *and* the assigned-to-control groups to voting-eligible individuals; we do not live in such a world.
+Worse yet, there is an obvious relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are both ineligible to have voting rights bestowed and cannot vote.
 
@@
 The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
-Then "treatment" would simply be identifying a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
+Then "treatment" would simply identify a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
 
@@
 Although a positive "treatment effect" is likely in this context, it would be unable to speak to causality.
+The instrumental variables approach allows us to mitigate this problem.
 
@@
 
-Using the natural experiment with imperfect compliance setup, I begin by estimating the intend-to-treat effect of the policy change.
+Using the natural experiment with imperfect compliance setup, I begin by estimating the intent-to-treat effect of the policy change.
 This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
@@
 
-After estimating the intend-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
+After estimating the intent-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
 As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
-To explore this potential, I also explore whether the treatment had different effects for different subgroups of the formerly incarcerated population.
-I finally explore whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+To explore this potential, I also test whether the treatment had different effects for different subgroups of the formerly incarcerated population.
+I finally ask whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+
+This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
+To that end I use turnout --— not registration —-- as the dependent variable of interest throughout.
+The Supplemental Information demonstrates that all treatment effects discussed here (including for subgroups within the population) also hold when the dependent variable is registration, not turnout.
 
@@
 When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[The Supplemental Information provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
+Because discharge date from parole is uncorrelated in the short term with an individual's propensity to vote in the short-run, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18^th^, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the "control group", while individuals discharged after are assigned "treatment" by the policy change.
-Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
+To further demonstrate the usefulness of discharge date as an intent-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intent-to-treat group (those discharged between May 18th and October 12th, 2018).
+With the exception of age, the control and intent-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote.
+To the extent our control group perhaps has a marginally higher propensity to vote than our intent-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
 \begin{singlespace}
@@
 t <- readRDS("./temp/demos_rd_time.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of \"Control\" and \"Intend-to-Treat\" Groups", linesep = "",
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of \"Control\" and \"intent-to-Treat\" Groups", linesep = "",
                   align = c("l", "c", "c"), "latex") %>% 
@@
 
-lab <- "\\caption{\\label{tab:demo-rd} Demographics of \"Control\" and \"Intend-to-Treat\" Groups}"
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of \"Control\" and \"intent-to-Treat\" Groups}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
@@
 
-Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model2[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
+Table \ref{tab:to-18-logit} also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
 The turnout rate for the control group was `r format(round(100*readRDS("./temp/control_to_18.rds"), digits = 1), nsmall = 1)` percent, indicating a treatment effect of `r round((exp(model3[["coefficients"]][["finished_postTRUE"]])-1) * readRDS("./temp/control_to_18.rds"), digits = 4) * 100` percentage points.
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+Although this intent-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
 To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-Not everyone discharged after the policy changed had their rights restored.
+As detailed in Figure \ref{fig:flow}, not everyone discharged after the policy changed had their rights restored.
 In other words, not everyone "assigned" to the treatment group was actually treated.
@@
 
-Two-stage least squares models allow us to leverage the random assignment of parole discharge date to identify the causal effect of rights restoration on the treated population.
+Two-stage least squares models allow us to leverage the random assignment to treatment to identify the causal effect of rights restoration on the treated population.
 The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
@@
 Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between `r round((0.84 / (3.1684 - 0.84)) * 100, digits = 1)` and `r round((0.96 / (3.1684 - 0.96)) * 100, digits = 1)` percent.
+
+Figure \ref{fig:sens} estimates Model 2 from Table \ref{tab:iv-models} while using different windows around the effective date of the executive order.
+The first window includes all individuals discharged between the start of 2017 and the registration deadline.
+The second window limits the control group to individuals discharged on or after June 1, 2017, while the third window limits the control group to individuals discharged in 2018.
+The final window limits the whole population to individuals discharged in the three months on either side of the executive order.
+
+```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/window_change_plot.rds")
+print(j)
+```
+
+In three of the four windows, the treatment effect remains significant at the 95 percent level.
+Only in the fourth window, where the sample includes just 30 percent as many observations as the largest window, is the treatment effect nonsignificant.
+Given that the treatment effect remains largely constant across the different windows, this is likely driven by the reduced sample size.
 
@@
 There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intend-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
+Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intent-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
 Because Latinos are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
@@
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+Table \ref{tab:black-to} also shows, however, that among the intent-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
 The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.^[This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation, and to that end I use turnout — not registration — as the dependent variable of interest throughout.
-The Supplemental Information demonstrates that all treatment effects discussed here also hold when the dependent variable is registration, not turnout]
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
 
@@
 When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-The executive order likely meant that more formerly disenfranchised individuals were sure that they could cast a ballot in the 2018 election.
+Of course, even voters in the control group were receiving notification of their eligibilty through the mail.
+If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
 In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
@@
 Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
-Limiting the analysis to just White parolees reveals that rights restoration increased turnout by more than 90 percent.
-The fact that rights restoration prior to discharge appears to have no such impact on non-White or Black parolees is surprising given the magnitude of the effect for Whites.
+Limiting the analysis to just white parolees reveals that rights restoration increased turnout by more than 90 percent.
+The fact that rights restoration prior to discharge appears to have no such impact on non-white or Black parolees is surprising given the magnitude of the effect for whites.
 This discrepancy could be caused by a number of different factors.
-Firstly, Black and White individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
+Firstly, Black and white individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
 As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
@@
 Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners [@Hemmens2000].
+Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than white prisoners [@Hemmens2000].
 Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-@White2015, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+@White2015, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino white aliases.
 There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2017].
@@
 Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such White individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among White former parolees; future research must investigate why this is the case.
+In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such white individuals were registered Democrats.
+Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among white former parolees; future research must investigate why this is the case.
 
@@
 By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
+Nevertheless, the success of the executive order is tempered by the racially disparate effects.
 

### 2020-03-05 markdown/part2/part2_standalone.tex
write

@@
   pdftitle={Welcome Home --- Now Vote!},
-  pdfauthor={Kevin Morris},
   hidelinks,
@@
 
-\title{Welcome Home --- Now Vote!\thanks{Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
-The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
-All errors are my responsibility.}}
+\title{Welcome Home --- Now Vote!}
 \usepackage{etoolbox}
@@
 \subtitle{Voting Rights Restoration and Post-Supervision Participation}
-\author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{March 04, 2020}
+\author{}
+\date{\vspace{-2.5em}March 05, 2020
+\linebreak
+\linebreak
+\linebreak
+\linebreak
+Word Count: 9,553}
 
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-In other states, notably Kentucky and Iowa, felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
+Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated(Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
+Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
 
@@
 
-In the spring of 2018,\footnote{Although the executive order was signed on April 18\textsuperscript{th}, it did not actually go into effect until May 18\textsuperscript{th}.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for individuals who were still on parole on election day, and would have been disenfranchised otherwise.
-The change in policy is also beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, the move has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+In the spring of 2018,\footnote{Although the executive order was signed on April 18\textsuperscript{th}, it did not go into effect until May 18\textsuperscript{th}.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+Such a move was of course good for individuals who were still on parole on election day --- they would have been disenfranchised otherwise.
+The change in policy is also beneficial for felony probationers.
+Despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, it has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
 
-There is reason to believe that the executive order may also have increased the political participation of formerly disenfranchised individuals.
+The executive order may also have increased the political participation of \emph{formerly} disenfranchised individuals.
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
@@
 Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
-Parole officers were also instructed to offer any assistance needed, including help filling out the registration form.
+Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
 In addition to these directives, the memorandum communicates that voter registration is to receive ``high priority attention,'' and it separately calls the program ``a \textbf{priority} initiative'' {[}emphasis in the original{]}.
@@
 In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
-These individuals, because they were discharged from parole before the registration deadline, would have been eligible to vote even in the absence of the executive order.
-This is only a subset of the individuals who were impacted by Executive Order 181 --- many of the individuals who had their voting rights restored were still on parole on election day, and therefore were eligible to vote only because of the rules change.
-However, by focusing on individuals who would have been eligible to vote either way, I can test whether rights restoration prior to parole discharge serves as an effective encouragement of post-supervision participation.
+I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
 
-By examining the effect of rights restoration in the context of a personal relationship with a parole officer, rather than simply at the end of a sentence and communicated through the mail, Executive Order 181 offers us the opportunity to build on existing research.
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility through the mail, Executive Order 181 offers us the opportunity to build on existing research.
 Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
-Or do parole officers exhibit biases against their stewards, making them ineffective conduits for restoration?
+Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration?
 
@@
 A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration \emph{causes} low turnout the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing turnout for this population.
+Whether or not incarceration \emph{causes} low turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
 
@@
 Prior to 2005 (and after 2011), individuals with felony convictions were permanently disenfranchised unless they submitted an application to the governor.
-Executive Order 42 eliminated this requirement, instead re-enfranchising automatically upon being discharged from their sentence.
+Executive Order 42 eliminated this requirement, instead re-enfranchising individuals automatically upon sentence completion.
 Although all formerly disenfranchised individuals were re-enfranchised upon the signing of the Executive Order, not all re-enfranchised individuals were informed of their change in status.
 Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, examines states where so-called notification laws went into effect.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), on the other hand, examines states where so-called notification laws went into effect.
 Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
@@
 Gerber et al.
-(\protect\hyperlink{ref-Gerber2015}{2015}) built on the quasi-experimental design of Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and conducted a field experiment in Connecticut in advance of the 2012 presidential election.
+(\protect\hyperlink{ref-Gerber2015}{2015}) built on the quasi-experimental design of Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}), conducting a field experiment in Connecticut in advance of the 2012 presidential election.
 In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
@@
 Because Executive Order 181 operates \emph{not} through the mail but rather through ``personal contact with government\ldots{} officials,'' its effects may differ from the mail notification programs.
-The social benefits (restoration of voting rights) might be communicated more effectively, thus increasing turnout further.
+The social benefits (restoration of voting rights) might be communicated more effectively and in ways that improve the formerly incarcerated individuals' opinions of their democratic state, thus increasing turnout further.
 
@@
 By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-Roughly 1 percent of former parolees are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1 percent of former parolees are not in the online lookup.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1,200 individuals who were discharged between the effective date of the executive order and the registration deadline did not have voting rights restored, due largely to their status as noncitizens.
 
@@
 In this paper, I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
-All individuals discharged from parole after the start of the executive order are assigned to the ``treatment'' group; individuals discharged earlier are assigned to the control group.
+All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
 When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
@@
 
-The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually ``treated'' by having their voting rights restored while on parole in a meeting with their parole officer.
-In other words, our natural experiment has ``imperfect compliance.'' Noncitizens, for instance, were not eligible for rights ``restoration'' while on parole because in New York State they were not allowed to vote even prior to their conviction.
-Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intend-to-treat effect.
+The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually treated by having their voting rights restored while on parole in a meeting with their parole officer.
+In other words, our natural experiment has ``imperfect compliance'' (see, for instance, Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
+Noncitizens, for instance, were not eligible for rights ``restoration'' while on parole because in New York State they were not allowed to vote even prior to their conviction.
+Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intent-to-treat effect.
 Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
 
-Although we can identify the individuals in the assigned-to-treatment group who do not ``comply'' with their treatment status, we do not know who among the ``assigned-to-control'' group would have gone untreated if they had been assigned to the other group.
+Although we can identify the individuals in the assigned-to-treatment group who do not comply with their treatment status, we do not know who among the assigned-to-control group would have gone untreated if they had been assigned to the other group.
 We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
-In an ideal world, we would limit both the treatment \emph{and} the control group to individuals who were eligible to have their rights restored while on parole; we do not live in such a world.
-Worse yet, there is obviously a relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are, of course, both ineligible to have voting rights bestowed and cannot vote.
+In an ideal world, we would limit both the assigned-to-treatment \emph{and} the assigned-to-control groups to voting-eligible individuals; we do not live in such a world.
+Worse yet, there is an obvious relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are both ineligible to have voting rights bestowed and cannot vote.
 
@@
 The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
-Then ``treatment'' would simply be identifying a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
+Then ``treatment'' would simply identify a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
 
@@
 Although a positive ``treatment effect'' is likely in this context, it would be unable to speak to causality.
+The instrumental variables approach allows us to mitigate this problem.
 
@@
 
-Using the natural experiment with imperfect compliance setup, I begin by estimating the intend-to-treat effect of the policy change.
+Using the natural experiment with imperfect compliance setup, I begin by estimating the intent-to-treat effect of the policy change.
 This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
-I turnout in the 2018 election as the dependent variable in these logistic regressions.
+I use turnout in the 2018 election as the dependent variable in these logistic regressions.
 
-After estimating the intend-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
+After estimating the intent-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
 As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
-To explore this potential, I also explore whether the treatment had different effects for different subgroups of the formerly incarcerated population.
-I finally explore whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+To explore this potential, I also test whether the treatment had different effects for different subgroups of the formerly incarcerated population.
+I finally ask whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+
+This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
+To that end I use turnout ----- not registration ----- as the dependent variable of interest throughout.
+The Supplemental Information demonstrates that all treatment effects discussed here (including for subgroups within the population) also hold when the dependent variable is registration, not turnout.
 
@@
 A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{\protect\hyperlink{appendix-a}{Appendix A} provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
+When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{The Supplemental Information provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
+Because discharge date from parole is uncorrelated in the short term with an individual's propensity to vote in the short-run, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
-Although it perhaps seems counterintuitive to throw out many of our observations, ignoring all individuals last discharged from parole prior to 2017 give us an important asset: it allows us to conceptualize Executive Order 181 as a natural experiment.
-Whether an individual was discharged from parole before or after May 18\textsuperscript{th}, 2018, is akin to a randomly assigned treatment.
-Individuals who were discharged from parole prior to the implementation of the executive order are part of the ``control group'', while individuals discharged after are assigned ``treatment'' by the policy change.
-Because Table \ref{tab:to-18-logit-short} shows that discharge date from parole is uncorrelated in the short term with an individual's propensity to vote, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intend-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intend-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intend-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them slightly more likely to vote.
-To the extent our control group perhaps has a slightly higher propensity to vote than our intend-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
+To further demonstrate the usefulness of discharge date as an intent-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intent-to-treat group (those discharged between May 18th and October 12th, 2018).
+With the exception of age, the control and intent-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote.
+To the extent our control group perhaps has a marginally higher propensity to vote than our intent-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
 
@@
 
-Models 2 and 3 also indicate that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 34.5 and 35 percent.
+Table \ref{tab:to-18-logit} also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
+Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 26.5 and 35 percent.
 The turnout rate for the control group was 2.1 percent, indicating a treatment effect of 0.74 percentage points.
@@
 Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intend-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
+Although this intent-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
 To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-Not everyone discharged after the policy changed had their rights restored.
+As detailed in Figure \ref{fig:flow}, not everyone discharged after the policy changed had their rights restored.
 In other words, not everyone ``assigned'' to the treatment group was actually treated.
@@
 Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 36.1 and 43.5 percent.
+
+Figure \ref{fig:sens} estimates Model 2 from Table \ref{tab:iv-models} while using different windows around the effective date of the executive order.
+The first window includes all individuals discharged between the start of 2017 and the registration deadline.
+The second window limits the control group to individuals discharged on or after June 1, 2017, while the third window limits the control group to individuals discharged in 2018.
+The final window limits the whole population to individuals discharged in the three months on either side of the executive order.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{part2_standalone_files/figure-latex/sensitivity-1}
+
+}
+
+\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
+\end{figure}
+
+In three of the four windows, the treatment effect remains significant at the 95 percent level.
+Only in the fourth window, where the sample includes just 30 percent as many observations as the largest window, is the treatment effect nonsignificant.
+Given that the treatment effect remains largely constant across the different windows, this is likely driven by the reduced sample size.
 
@@
 
-Table \ref{tab:black-to} also shows, however, that among the intend-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
+Table \ref{tab:black-to} also shows, however, that among the intent-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
 The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
 The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.\footnote{This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation, and to that end I use turnout --- not registration --- as the dependent variable of interest throughout.
-The Supplemental Information demonstrates that all treatment effects discussed here also hold when the dependent variable is registration, not turnout}
+Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
 
@@
 Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than White prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
+Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than white prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
 Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-White2015}{2015}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino White aliases.
+White, Nathan, and Faller (\protect\hyperlink{ref-White2015}{2015}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino white aliases.
 There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2017}{2017}).
@@
 By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
-
-\newpage
-\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}
-
-\hypertarget{appendix-a}{%
-\section*{Appendix A}\label{appendix-a}}
-\addcontentsline{toc}{section}{Appendix A}
-
-In the \protect\hyperlink{instrumental-variables-approach}{Instrumental Variables Approach} section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1\textsuperscript{st}, 2015, and October 14\textsuperscript{th}, 2016.
-
-\begin{singlespace}
-
-\input{"../../temp/table55.tex"}
-\end{singlespace}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
+Nevertheless, the success of the executive order is tempered by the racially disparate effects.
 
@@
 \url{https://doi.org/10.2307/2586120}.
+
+\leavevmode\hypertarget{ref-bcj_laws}{}%
+Brennan Center for Justice.
+2019.
+``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
+\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
 
@@
 
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Justice, Brennan Center for.
-2019.
-``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
-\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
-
 \leavevmode\hypertarget{ref-Lassen2004}{}%

### 2020-03-31 markdown/part1/UAR_submission.Rmd
uar r&r

@@
 In this paper, I redefine “lost voters” as disenfranchised individuals with a history of participating in elections.
-I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
+I map these individuals to their pre-incarceration addresses and use multiple approaches to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
 I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
@@
 \begin{singlespace}
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
 ```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
@@
 ```
-```{r distribution-lost, echo = F, message = F, warning = F}
-t <- readRDS("./temp/distribution_of_lost_voters.rds") %>% 
-  mutate_if(is.numeric, ~ comma(ifelse(is.na(.), 0, .)))
+```{r demos-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_nhoods.rds")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos) Comparing Neighborhoods with Lost Voters to Overall Neighborhoods", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
 
-t <- t[1:11, c(1,2)]
+write(j, "./temp/tempo.tex")
 
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:nhood-dist) Distribution of Lost Voters", linesep = "",
-                                                 align = "c") %>% 
-  row_spec(14, hline_after = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                font_size = 10)
-j
-```
-```{r block-table-chunk, echo = F}
-t <- readRDS("./temp/match_table_block_group.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:blockgroup-table) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"))
-j
-```
-```{r tract-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group.rdata")
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
 
-load("./temp/match_reg_block_group_nhb.rdata")
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-bg_reg_black <- reg_output_nhblack
-bg_reg_black_ses <- reg_output_nhblack_ses
-
-stargazer(bg_reg, reg_output2, bg_reg_black, reg_output2_nhblack,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses, bg_reg_black_ses, reg_output2_nhblack_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(Neighborhood Lost a Voter)", "Share Non-Hispanic Black", "D(Lost Voter) $\\times$ Share Non-Hispanic Black", "Constant"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          omit = c("median_income", "latino", "nh_white", "some_college",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "share_winner"),
-          title = "\\label{tab:match-reg} Matching Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F,
-          add.lines=list(c("Controls Included" , " ", "X", " ", "X")))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.}}}"
 
 j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
+  mutate(n = row_number())
 
+lab <- "\\caption{\\label{tab:demos} Neighborhood Demographics}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
 
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(17.5, 1.1))) %>% 
   arrange(n) %>% 
-  select(-n)
+  select(-n) %>% 
+  mutate(V1 = gsub("textbackslash\\{\\}", "", V1),
+         V1 = gsub("for Lost Voters", "\\&\\&for Lost Voters", V1))
 
-write.table(j, "./temp/table1.tex", quote = F, col.names = F,
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whateverrrrr2.tex", quote = F, col.names = F,
             row.names = F)
 ```
-\input{"../../temp/table1.tex"}
+\input{"../../temp/table_whateverrrrr2.tex"}
 ```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
@@
           se = list(bg_model_ses, bg_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
+          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black",
+                               "Stop-and-Frisk Encounters (Per 1,000 Residents)",
+                               "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
                                "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
@@
 
-stargazer(bg_reg, reg_output2,
+stargazer(bg_reg, reg_output2, reg_output3,
           header = F,
-          se = list(bg_reg_ses, reg_output2_ses),
+          se = list(bg_reg_ses, reg_output2_ses, reg_output3_ses),
           notes = c("TO REPLACE"),
@@
           out = "./temp/test_abc.tex",
-          omit = c("median_income", "latino", "nh_white", "some_college", "district",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black"),
+          omit = c("median_income", "latino", "nh_white", "some_college", "district", "sf",
+                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black", "control_GEOID"),
           out.header = F,
-          add.lines=list(c("Congressional District FE", "X", "X"),
-                         c("Controls Included", " ", "X")))
+          add.lines=list(c("Congressional District FEs", "X", "X", "X"),
+                         c("Controls Included", " ", "X", "X"),
+                         c("Block Group FEs", "", "", "X")))
 
@@
 
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
 
@@
 \input{"../../temp/table1111.tex"}
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
+true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
+plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
 \end{singlespace}
@@
 
-Due to economic and racial segregation, these effects are highly spatially concentrated.
+This paper furthers our understanding of the political implications of felony disenfranchisement at the neighborhood level.
+By redefining "lost voters" as disenfranchised individuals with a history of voting (rather than all individuals legally barred from voting) and by looking for effects at the neighborhood (Census block group level) I test the effects of felony disenfranchisement on local turnout.
+Using the 2017 mayoral election in New York City, I find that neighborhoods with lost voters have substantially lower turnout than other neighborhoods.
+This effect is concentrated in areas with large Black populations.
+
+Given the reality of economic and racial segregation, it should not be surprising that these effects are highly spatially concentrated.
 Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
@@
 
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
+Improving upon this literature is of great importance when we consider the spatial concentration of policing networks.
 If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
+If, however, the effects are both large enough to be detected at a statewide level (as some literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
 It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
 This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (8).
+As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (p 8).
 
-# Background {-}
+# Literature on the Indirect Effects of Felony Disenfranchisement {-}
 
-Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
-@Uggen2002 produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for disenfranchisement in Florida.
-Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
-Since @Uggen2002, other scholars have also investigated the direct effect of felony disenfranchisement [@Miles2004;@Uggen2004; @Drucker2005; @Ochs2006].
-
-A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+A number of recent papers have attempted to estimate the indirect effects of felony disenfranchisement on turnout.
+These papers have generally looked for effects at the state level, and have come to different conclusions.
 @King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
@@
 As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (724).
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p 724).
 
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the *state* level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-@Burch2013 is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that "at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods" (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of "lost voters."
+@Miles2004, however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the statewide level.
+This paper is more methodologically sophisticated than other research in this body of literature, and @Miles2004 finds that little-to-no effect from felony disenfranchisement on statewide turnout.
+The statewide research is in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
+Like the other literature looking for statewide disenfranchisement turnout effects, Miles uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
 
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods [e.g.
-@Sewell2016; @Clear2008; @Lerman2013].
-It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
-
-# Framing for Turnout Effects {-}
-
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with `r round(readRDS("./temp/active_voters_0319.rds")/1000000, digits = 1)` million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as *de facto* disenfranchisement.
-
-Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this *de facto* disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-@Burch2013, for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
-Moreover, voting is a social act, and social networks play an important role in predicting political participation [e.g.
+Research looking for the effect of felony disenfranchisement in statewide statistics is fundamentally flawed.
+The effects of incarceration --- and, therefore, felony disenfranchisement --- is not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
+Voting is a social act, and social networks play an important role in predicting political participation [e.g.
 @Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
@@
 It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-I therefore hypothesize that felony disenfranchisement does have locally bounded spillover effects, and that turnout in neighborhoods with lost voters systematically vote at a lower rate than others.
+This is also the level at which any effect from felony disenfranchisement likely operates.
 
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the indirect neighborhood effects of lost voters.
-@Walker2014 and @Walker2017, for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as "as having a loved one who is a custodial citizen without yourself having had contact" [@Walker2017, 542]) were not less likely to vote, but *were* more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
-These effects are particularly pronounced for women of color.
-Although Walker and others demonstrate that these actions largely take place outside of the voting booth, they may moderate spillover effects on turnout.
+The little research looking for turnout effects at the local level, however, is similarly conflicting.
+@Burch2013, for instance, demonstrates that block groups in North Carolina with higher levels of incarceration had lower turnout.
+@White2019a, on the other hand, finds that incarceration has only moderate indirect effects on voting.
+She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (p 607).
+Neither of these papers, however, account for felony disenfranchisement laws *per se*, but rather test the total indirect effects of incarceration (inclusive of disenfranchisement).
+Disenfranchised individuals who would not have voted in the absence of their incarceration are not directly affected by disenfranchisement policy; any indirect turnout effects associated with their incarceration, therefore, cannot be attributed to disenfranchisement policy.
 
-Similarly, @White2019a finds that incarceration has only moderate indirect effects on voting.
-She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (607).
-This study does not, however, interrogate whether the incarceration of a would-be voter has different indirect effects than the incarceration of an individual who would not have voted either way.
-It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect effects arising from other individuals.
-Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
-
-# Theoretical Implications for Reduced Turnout {-}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community [e.g.
-@King2016; @Bowers2009].
-With the exception of @Burch2013, however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: @Kinsella2015 examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes [@Martin2003].
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-"[H]igher citizen participation is rewarded," @Martin2013 concludes, "with enhanced policy responsiveness" (59).
-@Griffin2005 finds similar effects in the United States Senate, demonstrating that "voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not" (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
-
-Although research on the impact of local turnout on city-wide policy is scarce, @Anzia2019 examines the impact of senior turnout on "senior-friendly" policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials *are* responsive to senior turnout when seniors "are a more cohesive, meaningful group" (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of *neighborhood* turnout rates on city policy.
-
-@Hajnal2005 and @Hajnal2009, however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-@Hajnal2005 tells us that "Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils" (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-@Hajnal2005 and @Anzia2019 both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
+This study deals with both of these problems with the existing literature.
+I deal with geographic concentration by testing the relationship between felony disenfranchisement and turnout at the Census block group level.
+Secondly, I do not look for the indirect effects of *any* individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual with a demonstrated history of voting.
+As @Gerber2003 and others indicate, past turnout is a very strong predictor of future turnout.
+Thus, by defining lost voters as individuals who previously participated, I can identify the individuals most likely to vote in the absence of felony disenfranchisement laws --- and therefore the individuals (and neighborhoods) most directly impacted by these laws.
+Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood rates of stop-and-frisk police stops.
 
@@
 
-## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+Further discussion about how these administrative records are linked and geocoded can be found in the Supplemental Information.
 
-## Matching {-}
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a "match," records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
+# Research Design and Expectations {-}
 
-* Middle names are identical. If neither set of records includes a middle name, this condition is met.
-* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-* A middle name or middle initial in one set of records, and a missing middle name in the other set.
+I begin with an ordinary least squares regression estimated at the block group level.
+In this regression, neighborhood turnout is the dependent variable, while the number of lost voters is the independent variable of interest.
+I look at the 2017 mayoral election in New York City, and define a "lost voter" as an individual incarcerated or on felony parole on Election Day in 2017 who cast a ballot sometime between 2007 and 2016.
 
-Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andrew Doe" and "John Anthony Doe" would not.
+A regression that looks at turnout only in one year can shed little light on the causal effect of felony disenfranchisement.
+After investigating whether neighborhoods with lost voters turned out at lower rates in 2017, I use data from the 2016 presidential election to ask whether the loss of a voter leads to depressed turnout.
+I limit the pool of block groups to block groups with no lost voters in 2016, and test whether neighborhoods that had a lost voter one year later saw turnout decrease relative to their own history in 2016, and the 2017 turnout of neighborhoods that did not lose a voter in that period.
+Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods to ensure that treated and control observations were substantially similar prior to treatment.
 
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-@Meredith2013 offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-TABLE 1 ABOUT HERE
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
-
-# Identification of Lost Voters {-}
-
-Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
-By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
-Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
-
-In this analysis, I offer a different definition of "lost voter" than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- *potentially* lost voters --- and turnout.
-Such an approach is informative for understanding the impact of disenfranchisement.
-Many young men, for instance, are admitted to prison each year.
-If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have "aged in" to voting but for their incarceration, they will have no history of voting.
-This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
-Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
-
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the specific direct and indirect causal effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration [e.g.
-@Gerber2017].
-Research showing lower turnout where more individuals are formally disenfranchised might only be identifying the indirect effects of incarceration on turnout.
-Estimation of the effects of felony disenfranchisement above-and-beyond the effects of incarceration must account for the low propensity to vote among the formally disenfranchised.
-
-Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, "lost voters" are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote [@Gerber2003].
-As such, these are the individuals most likely to have been directly impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
-Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
-
-Of course, the incarceration of neighbors and family members who have *not* voted in the past might also have indirect effects on neighborhood turnout rates.
-Even if an eligible New Yorker's incarcerated brother would not have voted, for instance, the eligible voter might not cast a ballot due to a soured impression of the political process.
-Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
-I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
-These higher indirect (depressive) effects likely arise through three mechanisms.
+I expect that neighborhoods with lost voters turned out at lower rates in 2017, and that the loss of a voter to felony disenfranchisement causes neighborhood turnout to decline.
+I expect these effects to operate through three primary mechanisms:
 
@@
 
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7^th^, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7^th^, 2017, and had cast a ballot between 2007 and 2016.
-In many cases, these "lost voters" will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effect in its wake.
+# Results {-}
 
@@
 
-FIGURE 1 ABOUT HERE
+FIGURE \ref{fig:citymap} ABOUT HERE
 
@@
 
-# Testing for Neighborhood Turnout Effects {-}
+Table \ref{tab:demos} demonstrates just how different the neighborhoods with lost voters are from the rest of the city.
+The first column in Table \ref{tab:demos} weights the chacteristics of each block group in the city by its population; the second column weights the block groups by their number of lost voters.
+The median income in neighborhoods with lost voters is less than three-quarters that of unaffected neighborhoods.
+The average lost voter came from a neighborhood that was nearly twice as Black as the city as a whole; their neighborhoods were also home to a far larger share of Democratic voters than other neighborhoods.
+Table \ref{tab:demos} makes clear that neighborhoods with lost voters look substantially different than New York City in general.
 
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter.
+TABLE \ref{tab:demos} ABOUT HERE
 
-TABLE 2 ABOUT HERE
+## Testing for Neighborhood Turnout Differences {-}
 
-I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods are considered treated if they were home to *at least one* lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census block groups [@Sekhon2011], using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau's 5-year American Community Survey estimates.
+I begin by asking whether turnout was lower in the 2017 mayoral election in neighborhoods with lost voters.
+Block group turnout rates are calculated using the geocoded voter file.
+Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
+The number of ballots cast is divided by the neighborhood's citizen voting age population (CVAP, estimated by the Census Bureau).
+The independent variable of interest measures the number of lost voters in a neighborhood.
+
+I also control for other social and political indicators: neighborhood racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau's 2017 5-year American Community Survey estimates.
 Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
+Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the CVAP.
 I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups.
-Matching is done with replacement.
-Table \ref{tab:blockgroup-table} presents the results of these matches.
 
-## Match Output {-}
-As Table \ref{tab:blockgroup-table} demonstrates, the matching procedure results in an untreated group of neighborhoods that looks substantially like the treatment group.
+To completely decompose the effects of disenfranchisement from incarceration more generally, we would need to control for the number of incarcerated or paroled individuals in a neighborhood.
+If, after controlling for a neighborhoods general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
 
-TABLE 3 ABOUT HERE
-
-In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
-On average, the non-Hispanic Black share of a block group with a lost voter is more than *double* the share in block groups without lost voters.
-These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.^[Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.] The median income in neighborhoods without lost voters is more than 40 percent higher.
-Table \ref{tab:blockgroup-table} throws into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
-
-After matching neighborhoods, I use an ordinary least squares regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
+Unfortunately, such data does not exist; the State of New York does not make pre-incarceration addresses available to the public.
+In its place, I use data from the NYPD on stop-and-frisk encounters with the police.
+These models include the number of times residents were stopped by the police from January -- October of 2017.
+This is intended to control for general neighborhood exposure to the carceral state.
 
 Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match [@Abadie2019].
-Models 2 and 4 include the covariates on which the matching was done.
+Model 2, therefore, interacts the number of lost voters with the share of the block group that is non-Hispanic Black.
 
-TABLE 4 ABOUT HERE
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average `r -1*round(bg_reg[["coefficients"]][["treatTRUE"]]*100, digits = 1)` percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The average treated neighborhood was 38 percent Black, which means that the average block group with a lost voter saw turnout that was `r -1*round(bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]]*38, digits = 2)` percentage points lower than similar neighborhoods without lost voters.
-This is very close to the average treatment effect presented in Models 1 and 2; Models 3 and 4, however, makes clear that this effect is largely concentrated in Black communities.
-This disparate effect could have troubling ramifications for these neighborhoods' political representation.
-
-## Testing Intensity Effects {-}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on *how many* voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
-In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
+Table \ref{tab:trad-reg} presents the results of ordinary least squares regressions testing neighborhood turnout rates in 2017.
 Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
 
-TABLE 5 ABOUT HERE
+TABLE \ref{tab:trad-reg} ABOUT HERE
 
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model, though they are slightly smaller.
-This is unsurprising, because the average treated neighborhood had more than one lost voter.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points), but Model 2 again makes clear that this effect is concentrated in Black neighborhoods.
+The results presented in Table \ref{tab:trad-reg} indicate that neighborhoods with lost voters turned out at lower rates in the 2017 election.
+Each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points.
+Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
 In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
@@
 
-FIGURE 2 ABOUT HERE
+FIGURE \ref{fig:postest} ABOUT HERE
 
-## Toward a Causal Estimation {-}
+## Toward a Causal Estimate {-}
 
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
+As the analysis above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
 The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
+For instance, there is no reliable data available on incarceration rates at the neighborhood level.
+If rates of incarceration are higher in the block groups with lost voters than other similarly low-income and minority neighborhoods, causality cannot be ascribed to felony disenfranchisement without.
 
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
+In order to estimate the causal effect of felony disenfranchisement on neighborhood-level turnout, I now test whether losing a voter between the 2016 and 2017 elections resulted in decreased turnout in 2017.
+I begin by restricting the set of neighborhoods to only neighborhoods that had no lost voters on election day in 2016.
 This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
 These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were *newly* home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 *did* have a lost voter in 2016, and are thus excluded from this analysis.
+I then identify the neighborhoods that did have lost voters in 2017; these block groups are our "treated" block groups, while the neighborhoods with lost voters in neither 2016 nor 2017 serve as our controls.
 
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
+Of course, as demonstrated in Table \ref{tab:demos}, neighborhoods with lost voters are very different than neighborhoods without lost voters.
+Control and treatment neighborhoods might have different levels of turnout in 2017 even in the absence of felony disenfranchisement.
+To account for this possibility, I use a genetic matching model [@Sekhon2011] to match each treated block group to one untreated block group, using the same characteristics detailed in Table \ref{tab:trad-reg} but based on their 2016 characteristics.
+Matched ties are not broken, leaving the potential for some treated block groups to match to more than 1 controls; the observations are weighted accordingly.
+Only treated block groups and the untreated block groups with which they match --- the controls --- are retained.
 
-Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods. As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+Although it may seem counterintuitive to discard some potential control block groups, doing so gives us a major advantage: it allows us to conceptualize the "treatment" (the loss of a voter) as a natural experiment.
+Because the treatment and control groups look similar based on their observable 2016 characteristics, the treatment can be understood as quasi-randomly assigned.
 
-TABLE 6 ABOUT HERE
+Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods.
+The matching procedure substantially improves the similarly between treated and control block groups along all characteristics.
+Table \ref{tab:did-match} mirrors Table \ref{tab:demos} above: treated neighborhoods are far less White, had many more stop-and-frisk encounters, and had much lower median incomes.
+
+TABLE \ref{tab:did-match} ABOUT HERE
 
 In Table \ref{tab:did-reg}, I present the outcome of an OLS regression on this subset of block groups.
-The variable D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election, a difference that was not statistically significant.
-D(Treat) $\times$ D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-This coefficient is negative and statistically significant, indicating that in the first election after the removal of a voter a neighborhood's turnout rate decreased by 2.6 percentage points.
+The variable *D(2017)* tests whether turnout for control block groups declined in 2017 relative to 2016.
+I test whether there was a difference in turnout between control and treatment neighborhoods in 2016 with *D(Treat)*.
+Of most interest is the coefficient on *D(Treat) $\times$ D(2017)*, which tests whether turnout decreased more from 2016 to 2017 in treated block groups than in the controls.
+Model 1 includes only these covariates (and fixed effects for Congressional district to account for potential differences in competitiveness in 2016), while Model 2 also includes the variables that were used for the matching procedure.
+Since each neighborhood appears in the panel twice --- once in 2016, and once in 2017 --- each observation takes the characteristics of the year in which it appears.
+This is especially important for the stop-and-frisk variable; Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
+Model 3 adds block group-level fixed effects to Model 2 in order to control for time-invariant neighborhood characteristics.
+Robust standard errors are clustered at the level of the match [@Abadie2019].
 
-TABLE 7 ABOUT HERE
+TABLE \ref{tab:did-reg} ABOUT HERE
 
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
+Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates; in each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016.
 These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.]
+Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 32 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.]
 
@@
 Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
+Because these effects are so much larger than those shown in Table \ref{tab:trad-reg} it is possible that neighborhoods with consistently lost voters are less impacted by the marginal loss of a voter.
 
@@
 As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement impacts Black turnout at the state level, but this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
+Despite methodological problems, previous literature has hinted that felony disenfranchisement impacts Black turnout at the state level; this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
 As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
@@
 Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
+Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is nonsignificant.
 This means that lost voters have no spillover effects in neighborhoods with small Black populations.
@@
 In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.16 * -1*bg_reg_black[["coefficients"]][["treatTRUE:nh_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
 
@@
 Similarly, @Lerman2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
+It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
 
@@
 
-This finding mirrors @White2019, which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
-This is perhaps unsurprising.
 A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans [e.g.
@@
 There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
+
+Ultimately, the analysis at hand cannot answer the causal mechanism for these depressive effects, or why the depressive effect is concentrated in Black neighborhoods.
+As @Walker2020 notes, "contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact" (p 123).
+Administrative data can only get us so far.
+Other research designs are necessary to investigate how neighbors understand the incarceration of their neighbors, how they react on voting day, and whether these effects are racially moderated.
+Much of the work being done by Hannah Walker and coauthors [-@Walker2014; -@Walker2017; -@Walker2020] does point to racial differences in political responses to proximal contact with systems of incarceration.
+
+This paper uncovers surprisingly large effects.
+It is possible that, despite controlling for neighborhood exposure to the carceral state using stop-and-frisk data, the analysis cannot entirely decompose the effect of disenfranchisement from the bundled experience of seeing a neighbor go to prison.
+Future research must gather better data on local patterns of incarceration in order to better distinguish turnout patterns in neighborhoods where voters are sent to prison from neighborhoods where non-voters are incarcerated.
+Future work should similarly test whether these effects, identified in a relatively uncompetitive local election, are also present in competitive and national elections.
 
@@
 \newpage
+# Supplemental Information {-}
+
+I here present more information on how the full data set used in the body of this paper was constructed.
+
+## Geocoding {-}
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
+The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
+Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+
+## Matching {-}
+
+I construct the primary dataset by matching the administrative criminal supervision records with the registered voter file.
+I match using first, middle, and last names, as well as dates of birth.
+Capitalization in names in both datasets is standardized, and punctuation and spaces are removed.
+If a middle name is missing in one dataset but not in the other, I consider this a match.
+Similarly, if one dataset includes only a middle initial which matches to the first letter of the middle name the other data set, this is considered a match.
+First and last names, and dates of birth, must match exactly between the two datasets.
+
+To test for false positive matches, I employ the test developed in @Meredith2013.
+I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
+Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
+This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
+
+TABLE \ref{tab:change-dobs} ABOUT HERE
+
+Testing for false negatives is more challenging.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+
+\newpage
 # References {-}

### 2020-03-31 markdown/part2/part2_standalone.Rmd
write

@@
 subtitle: "Voting Rights Restoration and Post-Supervision Participation"
+thanks: Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
+The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
+All errors are my responsibility.
+author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
+institute: "Brennan Center for Justice"
 date: |
@@
   \linebreak
-  Word Count: 9,553
+  Word Count: 9,596
 output:
@@
 The instrumental variables approach allows us to mitigate this problem.
+Moreover, a very small number of individuals in the control group are marked as having their rights restored prior to the effective date of the executive order.
+While this likely reflects paperwork errors, I nonetheless treat them as a second set of non-compliers.
 

### 2020-03-31 markdown/part2/part2_standalone.tex
write

@@
   pdftitle={Welcome Home --- Now Vote!},
+  pdfauthor={Kevin Morris},
   hidelinks,
@@
 
-\title{Welcome Home --- Now Vote!}
+\title{Welcome Home --- Now Vote!\thanks{Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
+The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
+All errors are my responsibility.}}
 \usepackage{etoolbox}
@@
 \subtitle{Voting Rights Restoration and Post-Supervision Participation}
-\author{}
-\date{\vspace{-2.5em}March 05, 2020
+\author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
+\date{March 09, 2020
 \linebreak
@@
 \linebreak
-Word Count: 9,553}
+Word Count: 9,596}
 
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2019}).
+Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-Roughly 1 percent of former parolees are not in the online lookup.} There were 5,024 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Roughly 1 percent of former parolees are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 Roughly 1,200 individuals who were discharged between the effective date of the executive order and the registration deadline did not have voting rights restored, due largely to their status as noncitizens.
@@
 
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice.
-2019.
-``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
-\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
-
 \leavevmode\hypertarget{ref-Burch2011}{}%
@@
 
+\leavevmode\hypertarget{ref-bcj_laws}{}%
+Justice, Brennan Center for.
+2019.
+``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
+\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
+
 \leavevmode\hypertarget{ref-Lassen2004}{}%

### 2020-03-31 markdown/part2/registration_regs.Rmd
write

@@
 ---
+\newpage
 \doublespace

### 2020-03-31 markdown/part1/UAR_submission.Rmd
write

@@
 Other research designs are necessary to investigate how neighbors understand the incarceration of their neighbors, how they react on voting day, and whether these effects are racially moderated.
-Much of the work being done by Hannah Walker and coauthors [-@Walker2014; -@Walker2017; -@Walker2020] does point to racial differences in political responses to proximal contact with systems of incarceration.
+Much of the work being done by Hannah Walker and coauthors [@Walker2014; @Walker2017; @Walker2020] does point to racial differences in political responses to proximal contact with systems of incarceration.
 

### 2020-04-01 markdown/part1/UAR_submission.Rmd
uar re-submit

@@
 title: "Neighborhoods and Felony Disenfranchisement: The Case of New York City"
-# thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
-# author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
-# institute: "Brennan Center for Justice"
+thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, Peter Miller, participants at the 2020 Annual Meeting of the Southern Political Science Association, and several blind reviewers for their comments on this project.
+All errors are my responsibility.
+author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
+institute: "Brennan Center for Justice"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
     latex_engine: xelatex
-    keep_tex: false
+    keep_tex: true
 bibliography: "../nyc_neighborhoods.bib"
@@
     - \usepackage{setspace}\doublespacing
+    - \usepackage{endnotes}
+    - \let\footnote=\endnote
 abstract: |
@@
 This literature, however, has often overlooked both the geographic concentration of communities impacted by over-incarceration and the low propensity to vote exhibited by individuals convicted of felony crimes.
-In this paper, I redefine “lost voters” as disenfranchised individuals with a history of participating in elections.
+In this paper, I redefine "lost voters" as disenfranchised individuals with a history of participating in elections.
 I map these individuals to their pre-incarceration addresses and use multiple approaches to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
-I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
 These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
@@
 
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.}}}"
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95\\% confidence level.}}}"
 
@@
 
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(17.5, 1.1))) %>% 
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(18.5, 1.1))) %>% 
   arrange(n) %>% 
@@
 
-This paper furthers our understanding of the political implications of felony disenfranchisement at the neighborhood level.
-By redefining "lost voters" as disenfranchised individuals with a history of voting (rather than all individuals legally barred from voting) and by looking for effects at the neighborhood (Census block group level) I test the effects of felony disenfranchisement on local turnout.
-Using the 2017 mayoral election in New York City, I find that neighborhoods with lost voters have substantially lower turnout than other neighborhoods.
-This effect is concentrated in areas with large Black populations.
+This paper furthers our understanding of the political implications of felony disenfranchisement by investigating its indirect turnout effects at the local level in New York City's 2017 mayoral election.
+Using administrative data from the Department of Corrections and Community Supervision and the registered voter file, I redefine "lost voters" as disenfranchised individuals with a history of voting.
+I test the effect of lost voters on neighborhood turnout after controlling for general neighborhood exposure to the criminal justice system, allowing me to tease apart the effect of disenfranchisement from the bundled indirect effects of overpolicing.
+I demonstrate that the neighborhoods left behind by lost voters turned out at substantially lower rates than other similar neighborhoods --- and that there is likely a causal relationship at play.
 
-Given the reality of economic and racial segregation, it should not be surprising that these effects are highly spatially concentrated.
+Given the reality of economic and racial segregation, it should not be surprising that the effects of felony disenfranchisement are highly spatially concentrated.
 Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is an incarcerated population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100` percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.]
+The result is a disenfranchised population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100`% of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3% of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1% of the citizen voting age population and 22.9% of the incarcerated population.]
 
 Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
@@
 
-Improving upon this literature is of great importance when we consider the spatial concentration of policing networks.
+Improving upon this literature is of great importance when we consider the spatial concentration of incarceration --- and, therefore, disenfranchisement.
 If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as some literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
+If, however, the effects are both large enough to be detected at a statewide level (as the literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation.
+This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences [@Kinsella2015].
+It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas [@Martin2003; @Martin2013].
 This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (p 8).
+As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (p.
+8).
 
@@
 As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p 724).
+[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
+724).
 
-@Miles2004, however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the statewide level.
-This paper is more methodologically sophisticated than other research in this body of literature, and @Miles2004 finds that little-to-no effect from felony disenfranchisement on statewide turnout.
-The statewide research is in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
-Like the other literature looking for statewide disenfranchisement turnout effects, Miles uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
+@Miles2004, however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the state level.
+This paper is more methodologically sophisticated than other research in this body of literature, and @Miles2004 finds little-to-no effect from felony disenfranchisement on statewide turnout.
+The statewide research is thus in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
+Despite its methodological improves, @Miles2004 (like the other literature looking for statewide disenfranchisement turnout effects) uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
 
 Research looking for the effect of felony disenfranchisement in statewide statistics is fundamentally flawed.
-The effects of incarceration --- and, therefore, felony disenfranchisement --- is not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
+This approach is generally taken more for methodological reasons (disenfranchisement policies vary between but not within states) or practical reasons (national data at lower geographic levels about incarceration and voting patterns is difficult to find), but not for theoretical ones.
+The effects of incarceration --- and, therefore, felony disenfranchisement --- are not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
 Voting is a social act, and social networks play an important role in predicting political participation [e.g.
@@
 It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
-This is also the level at which any effect from felony disenfranchisement likely operates.
+This is also the level at which any effect from felony disenfranchisement should theoretically operate.
 
@@
 @White2019a, on the other hand, finds that incarceration has only moderate indirect effects on voting.
-She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (p 607).
-Neither of these papers, however, account for felony disenfranchisement laws *per se*, but rather test the total indirect effects of incarceration (inclusive of disenfranchisement).
+She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (p.
+607).
+Neither of these papers, however, account for felony disenfranchisement laws *per se*, but rather test the bundled indirect effects of proximal contact with incarceration (inclusive of disenfranchisement).
 Disenfranchised individuals who would not have voted in the absence of their incarceration are not directly affected by disenfranchisement policy; any indirect turnout effects associated with their incarceration, therefore, cannot be attributed to disenfranchisement policy.
 
-This study deals with both of these problems with the existing literature.
+This study addresses with both problems with the existing literature.
 I deal with geographic concentration by testing the relationship between felony disenfranchisement and turnout at the Census block group level.
-Secondly, I do not look for the indirect effects of *any* individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual with a demonstrated history of voting.
+Secondly, I do not look for the indirect effects of proximal contact with *any* individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual *with a demonstrated history of voting*.
 As @Gerber2003 and others indicate, past turnout is a very strong predictor of future turnout.
 Thus, by defining lost voters as individuals who previously participated, I can identify the individuals most likely to vote in the absence of felony disenfranchisement laws --- and therefore the individuals (and neighborhoods) most directly impacted by these laws.
-Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood rates of stop-and-frisk police stops.
+Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood police stop rates.
 
@@
 The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-They include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
+These records include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
 The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-These data are used to determine when individuals were incarcerated or on parole, the class of crimes for which they were incarcerated, and other demographic information.
 
@@
 I look at the 2017 mayoral election in New York City, and define a "lost voter" as an individual incarcerated or on felony parole on Election Day in 2017 who cast a ballot sometime between 2007 and 2016.
+Although the 2017 mayoral election was a low turnout contest, @Hajnal2009 argues that these low turnout local elections are precisely where differential turnout rates can have the biggest consequences.
 
@@
 I limit the pool of block groups to block groups with no lost voters in 2016, and test whether neighborhoods that had a lost voter one year later saw turnout decrease relative to their own history in 2016, and the 2017 turnout of neighborhoods that did not lose a voter in that period.
-Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods to ensure that treated and control observations were substantially similar prior to treatment.
+Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods.
+The matching procedure ensures that, prior to treatment, treated and control observations are substantially similar to one another based on observable characteristics.
 
@@
 
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Figure \ref{fig:citymap} shows where lost voters lived before going to prison, with city council districts also included.
 There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election.
@@
 
-Although `r comma(nrow(readRDS("./temp/in_2017.rds")))` individuals were formally disenfranchised as of the 2017 election, there were just `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` disenfranchised voters statewide who had participated in the past 10 years.^[Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.] Just `r round(100 * nrow(readRDS("./temp/ids_of_lost_voters.rds")) / nrow(readRDS("./temp/in_2017.rds")), digits = 1)` percent of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
+Although `r comma(nrow(readRDS("./temp/in_2017.rds")))` individuals were formally disenfranchised as of the 2017 election, there were just `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` disenfranchised voters statewide who had participated in the past 10 years.^[Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.] Just `r round(100 * nrow(readRDS("./temp/ids_of_lost_voters.rds")) / nrow(readRDS("./temp/in_2017.rds")), digits = 1)`% of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
 
@@
 In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
+In other communities, such as Harlem and Central Brooklyn, large numbers of past voters were not allowed to cast a ballot for mayor.
 
@@
 The average lost voter came from a neighborhood that was nearly twice as Black as the city as a whole; their neighborhoods were also home to a far larger share of Democratic voters than other neighborhoods.
-Table \ref{tab:demos} makes clear that neighborhoods with lost voters look substantially different than New York City in general.
+Unsurprisingly, lost voters also came from neighborhoods with 60% more police stops than the citywide average.
 
@@
 Block group turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
+Each voter's record indicates whether the voter participated in the 2017 general election.
+These records are then aggregated to estimate the number of ballots cast in each neighborhood.
 The number of ballots cast is divided by the neighborhood's citizen voting age population (CVAP, estimated by the Census Bureau).
@@
 To completely decompose the effects of disenfranchisement from incarceration more generally, we would need to control for the number of incarcerated or paroled individuals in a neighborhood.
-If, after controlling for a neighborhoods general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
+If, after controlling for a neighborhood's general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
 
@@
 The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-For instance, there is no reliable data available on incarceration rates at the neighborhood level.
-If rates of incarceration are higher in the block groups with lost voters than other similarly low-income and minority neighborhoods, causality cannot be ascribed to felony disenfranchisement without.
+Moreover, by looking at turnout from just one election, the best we could hope to uncover is a correlation between disenfranchisement and turnout.
+Uncovering the causal effect requires looking across elections (or the very strong assumptions that all relevant neighborhood characteristics are accounted for).
 
@@
 To account for this possibility, I use a genetic matching model [@Sekhon2011] to match each treated block group to one untreated block group, using the same characteristics detailed in Table \ref{tab:trad-reg} but based on their 2016 characteristics.
-Matched ties are not broken, leaving the potential for some treated block groups to match to more than 1 controls; the observations are weighted accordingly.
+Matched ties are not broken, leaving the potential for some treated block groups to match to more than one control; the observations are weighted accordingly.
 Only treated block groups and the untreated block groups with which they match --- the controls --- are retained.
@@
 Since each neighborhood appears in the panel twice --- once in 2016, and once in 2017 --- each observation takes the characteristics of the year in which it appears.
-This is especially important for the stop-and-frisk variable; Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
+This is especially important for the stop-and-frisk variable: Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
 Model 3 adds block group-level fixed effects to Model 2 in order to control for time-invariant neighborhood characteristics.
@@
 
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates; in each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016.
+Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
+In each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016 who lost one in the subsequent year.
 These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 32 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).^[Each of these differences are significant at the 99 percent confidence level.]
+The marginal effect of a lost voter could be lower in neighborhoods that consistently have lost voters and thus have less to "learn" about the state from the loss of any one individual to incarceration.
+Neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16% versus 21%), a lower share non-Hispanic Black (39% versus 32%), and a smaller share were registered as Democrats (76% versus 73%).^[Each of these differences are significant at the 99% confidence level.]
 
@@
 Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
-Because these effects are so much larger than those shown in Table \ref{tab:trad-reg} it is possible that neighborhoods with consistently lost voters are less impacted by the marginal loss of a voter.
 
@@
 However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
+Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large direct effects on turnout and representation.
 
@@
 Despite methodological problems, previous literature has hinted that felony disenfranchisement impacts Black turnout at the state level; this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
-Similarly, just 17 percent of individuals in the average block group with a lost voter were White, compared with 40 percent of the population in other block groups.
+As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40% higher than the median income of block groups with lost voters.
+Similarly, just 17% of individuals in the average block group with a lost voter were White, compared with 40% of the population in other block groups.
 Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
@@
 Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is nonsignificant.
-This means that lost voters have no spillover effects in neighborhoods with small Black populations.
+This means that lost voters have little spillover effects in neighborhoods with small Black populations.
 Understanding the magnitude of this finding is important.
-In New York City, there are `r readRDS("./temp/count_gt_90_black")` block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
+In New York City, there are `r comma(readRDS("./temp/count_gt_90_black"))` block groups where the Black community makes up more than 38% of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
+Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38% Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
+Conversely, in a block group with the same voting age population that is just 16% Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
 
@@
 @Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, @Lerman2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (16).
+Similarly, @Lerman2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (p.
+16).
 It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
@@
 Ultimately, the analysis at hand cannot answer the causal mechanism for these depressive effects, or why the depressive effect is concentrated in Black neighborhoods.
-As @Walker2020 notes, "contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact" (p 123).
+As @Walker2020 notes, "contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact" (p.
+123).
 Administrative data can only get us so far.
@@
 I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)` percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)` percent of these individuals were not geocoded.
+This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)`% of registered voters.
+The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)`% of these individuals were not geocoded.
 Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
@@
 Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+This analysis indicates that false positives account for between 0.4 and 0.5% of all matches, a share that is likely too small to have any material impact on the overall analysis.
 The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
@@
 If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6% of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "Alex" in one set of records and "Alexander" in the other).
+
+\theendnotes
 

### 2020-04-01 markdown/part1/UAR_submission.tex
uar re-submit

--- 2019-11-12 markdown/part1/p1.tex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+  \setmainfont[]{Times New Roman}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Neighborhoods and Felony Disenfranchisement: The Case of New York City},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Neighborhoods and Felony Disenfranchisement: The Case of New York City},
-            pdfauthor={Kevin Morris},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Kevin Morris\footnote{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{November 12, 2019}
-
+\usepackage{rotating}
+\usepackage{setspace}\doublespacing
+\usepackage{endnotes}
+\let\footnote=\endnote
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
-\usepackage{rotating}
-\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
+\title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
+\author{}
+\date{\vspace{-2.5em}April 01, 2020}
 
@@
 In this paper, I redefine ``lost voters'' as disenfranchised individuals with a history of participating in elections.
-I map these individuals to their pre-incarceration addresses and use matching and linear models to explore whether their home neighborhoods turned out at lower rates than other neighborhoods.
-I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+I map these individuals to their pre-incarceration addresses and use multiple approaches to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
+I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
 These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
@@
 \pagenumbering{gobble}
+\newpage
+\pagenumbering{arabic}
+
+\begin{singlespace}
+\begin{figure}[H]
+
+{\centering \includegraphics{UAR_submission_files/figure-latex/citywide-map-1}
+
+}
+
+\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
+\end{figure}
+
+\input{"../../temp/table_whateverrrrr2.tex"}
+
+\input{"../../temp/table2.tex"}
+\begin{figure}[H]
+
+{\centering \includegraphics{UAR_submission_files/figure-latex/postest-map-1}
+
+}
+
+\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
+\end{figure}
+\begin{table}[H]
+
+\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{10}{12}\selectfont
+\begin{tabular}[t]{lllllrrrr}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
+ & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
+\midrule
+Stop-and-Frisk Encounters (Per 1,000 Residents) & 1.52 & 1.22 & 1.52 & 1.40 & 59.18 & 80.84 & 74.64 & 64.80\\
+\% Latino & 0.35 & 0.24 & 0.35 & 0.35 & 98.18 & 93.37 & 91.90 & 83.66\\
+\% Non-Hispanic Black & 0.33 & 0.15 & 0.33 & 0.33 & 98.29 & 94.91 & 93.59 & 86.37\\
+\% Non-Hispanic White & 0.20 & 0.41 & 0.20 & 0.20 & 97.57 & 92.56 & 89.24 & 83.47\\
+Median Income & 52,352.43 & 70,311.87 & 52,352.43 & 52,177.85 & 99.03 & 92.14 & 87.32 & 68.43\\
+\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 96.85 & 90.55 & 86.81 & 69.31\\
+Median Age & 35.61 & 38.12 & 35.61 & 35.68 & 97.21 & 68.61 & 73.56 & 63.48\\
+Registration Rate & 0.98 & 0.93 & 0.98 & 0.96 & 49.13 & 72.80 & 64.99 & 31.69\\
+\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 94.29 & 94.58 & 91.69 & 84.47\\
+\% Noncitizen & 0.17 & 0.17 & 0.17 & 0.18 & 90.15 & 90.13 & 83.20 & 61.09\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\input{"../../temp/table1111.tex"}
+\begin{table}[H]
+
+\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
+\centering
+\begin{tabular}[t]{cc}
+\toprule
+Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
+\midrule
+Actual Birthdate & 20,955\\
+Birthdate + 35 Days & 105\\
+Birthdate - 35 Days & 92\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{singlespace}
 \pagebreak
-
 \pagenumbering{arabic}
@@
 Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2018}).
+In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
 
-Due to economic and racial segregation, these effects are highly spatially concentrated.
+This paper furthers our understanding of the political implications of felony disenfranchisement by investigating its indirect turnout effects at the local level in New York City's 2017 mayoral election.
+Using administrative data from the Department of Corrections and Community Supervision and the registered voter file, I redefine ``lost voters'' as disenfranchised individuals with a history of voting.
+I test the effect of lost voters on neighborhood turnout after controlling for general neighborhood exposure to the criminal justice system, allowing me to tease apart the effect of disenfranchisement from the bundled indirect effects of overpolicing.
+I demonstrate that the neighborhoods left behind by lost voters turned out at substantially lower rates than other similar neighborhoods --- and that there is likely a causal relationship at play.
+
+Given the reality of economic and racial segregation, it should not be surprising that the effects of felony disenfranchisement are highly spatially concentrated.
 Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
@@
 Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is an incarcerated population that looks far different than the rest of th state: according to data from the New York State Department of Corrections and Community Supervision, 49.5 percent of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3 percent of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1 percent of the citizen voting age population and 22.9 percent of the incarcerated population.}
+The result is a disenfranchised population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, 49.5\% of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3\% of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1\% of the citizen voting age population and 22.9\% of the incarcerated population.}
 
 Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The effects of disenfranchisement are concentrated in neighborhoods that already suffer from myriad disadvantages thanks to social and economic marginalization.
 The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}; Gerber et al.
-\protect\hyperlink{ref-Gerber2014}{2014}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
+A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{b}; Gerber et al.
+\protect\hyperlink{ref-Gerber2015}{2015}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
 Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
@@
 
-Filling this gap in the literature is of great importance when we consider the spatial concentration of policing networks.
+Improving upon this literature is of great importance when we consider the spatial concentration of incarceration --- and, therefore, disenfranchisement.
 If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature suggests) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences.
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas.
+If, however, the effects are both large enough to be detected at a statewide level (as the literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation.
+This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences (Kinsella, McTague, and Raleigh \protect\hyperlink{ref-Kinsella2015}{2015}).
+It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas (Martin \protect\hyperlink{ref-Martin2003}{2003}; Martin and Claibourn \protect\hyperlink{ref-Martin2013}{2013}).
 This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (8).
+As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (p.~8).
 
-\hypertarget{background}{%
-\section*{Background}\label{background}}
-\addcontentsline{toc}{section}{Background}
+\hypertarget{literature-on-the-indirect-effects-of-felony-disenfranchisement}{%
+\section*{Literature on the Indirect Effects of Felony Disenfranchisement}\label{literature-on-the-indirect-effects-of-felony-disenfranchisement}}
+\addcontentsline{toc}{section}{Literature on the Indirect Effects of Felony Disenfranchisement}
 
-Since the 2000 election, scholars have attempted to quantify the effect of felony disenfranchisement on the political representation of highly-incarcerated communities.
-Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}) produced the first estimates of felony disenfranchisement's impact on turnout, arguing that Al Gore would have won the presidency if not for for disenfranchisment in Florida.
-Their analysis estimated the direct effects of felony disenfranchisement on turnout by quantifying the number of actually disenfranchised individuals who would have participated if given the chance.
-Since Uggen and Manza (\protect\hyperlink{ref-Uggen2002}{2002}), other scholars have also investigated the direct effect of felony disenfranchisement (Miles \protect\hyperlink{ref-Miles2004}{2004}; Uggen and Manza \protect\hyperlink{ref-Uggen2004}{2004}; Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-
-A number of papers have also explored the indirect impact felony disenfranchisement policies have on turnout among non-disenfranchised residents.
+A number of recent papers have attempted to estimate the indirect effects of felony disenfranchisement on turnout.
+These papers have generally looked for effects at the state level, and have come to different conclusions.
 King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2006}) to explore the impact of these policies on eligible voters.
+They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) to explore the impact of these policies on eligible voters.
 They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
@@
 As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
-{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (724).
+{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (p.~724).
 
-Although scholars have established that felony disenfranchisement decreases turnout among Black voters at the \emph{state} level, relatively little research has been done on how felony disenfranchisement operates at the sub-state level.
-Though we know that Black voters are generally less likely to cast a ballot when they live in a state with strict disenfranchisement laws, little work has been done exploring the impact these laws might have at the local level.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}) is an exception to this.
-Burch explores the depressive effect of disenfranchisement laws at the local level in North Carolina by examining census block group level turnout and zip code level involvement with the criminal justice system, determining that ``at high concentrations, imprisonment and community supervision have an unequivocally demobilizing effect of neighborhoods'' (185).
-This project seeks to expand on her work by replicating her findings in New York City; by using a different estimation technique; and by using a different definition of ``lost voters.''
+Miles (\protect\hyperlink{ref-Miles2004}{2004}), however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the state level.
+This paper is more methodologically sophisticated than other research in this body of literature, and Miles (\protect\hyperlink{ref-Miles2004}{2004}) finds little-to-no effect from felony disenfranchisement on statewide turnout.
+The statewide research is thus in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
+Despite its methodological improves, Miles (\protect\hyperlink{ref-Miles2004}{2004}) (like the other literature looking for statewide disenfranchisement turnout effects) uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
 
-This paper begins by exploring the effect of felony disenfranchisement on neighborhood turnout in the New York City Mayoral election of 2017 using individual-level administrative data.
-Policing and incarceration patterns have historically targeted communities of color, damaging the social fabric of these neighborhoods (e.g.~Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}; Clear \protect\hyperlink{ref-Clear2008}{2008}; Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-It is possible, however, that these effects run deeper than the direct focus on policing and incarceration has acknowledged.
-We know that felony disenfranchisement systematically removes individuals from certain neighborhoods, but it is not clear that enough would-be voters are removed relative to the electorate to meaningfully distort neighborhood representation.
-To the extent that felony disenfranchisement has a spatially-concentrated depressive effect on eligible voters, it is possible that these policies are powerful enough to materially reduce the representation of certain parts of the city.
-
-\hypertarget{framing-for-turnout-effects}{%
-\section*{Framing for Turnout Effects}\label{framing-for-turnout-effects}}
-\addcontentsline{toc}{section}{Framing for Turnout Effects}
-
-Despite over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters.
-In New York, for instance, 46,232 individuals were imprisoned in New York State in early 2019, compared with 11.6 million actively registered voters.
-Despite the low share of residents who are directly disenfranchised, there is reason to believe the policy impacts more individuals than just those imprisoned.
-As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not suspended This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher, an effect often referred to as \emph{de facto} disenfranchisement.
-
-Though previous studies have focused largely on the state-level spill-over effects of felony disenfranchisement, there is reason to believe that this \emph{de facto} disenfranchisement is concentrated within the neighborhoods home to formally disenfranchised residents.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that neighborhoods in North Carolina with higher levels of incarceration had lower turnout.
-Moreover, voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
+Research looking for the effect of felony disenfranchisement in statewide statistics is fundamentally flawed.
+This approach is generally taken more for methodological reasons (disenfranchisement policies vary between but not within states) or practical reasons (national data at lower geographic levels about incarceration and voting patterns is difficult to find), but not for theoretical ones.
+The effects of incarceration --- and, therefore, felony disenfranchisement --- are not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
+Voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
 Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
 It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
-To the extent that felony disenfranchisement policies have depressive effects on turnout in the social and filial networks of the imprisoned and paroled, these effects are likely to be closely concentrated in the neighborhoods where the disenfranchised live.
-I therefore hypothesize that felony disenfranchisement does have locally bounded spillover effects, and that turnout in neighborhoods with lost voters systematically vote at a lower rate than others.
+This is also the level at which any effect from felony disenfranchisement should theoretically operate.
 
-Recent work from Hannah Walker and others, however, leads me to moderate my hypothesis about the indirect neighborhood effects of lost voters.
-Walker (\protect\hyperlink{ref-Walker2014}{2014}) and Walker and Garcı́a-Castañon (\protect\hyperlink{ref-Walker2017}{2017}), for instance, demonstrate that individuals who have proximal contact with the criminal justice system (defined as ``as having a loved one who is a custodial citizen without yourself having had contact'' (Walker and Garcı́a-Castañon \protect\hyperlink{ref-Walker2017}{2017}, 542)) were not less likely to vote, but \emph{were} more likely to participate in nonelectoral political events (such as signing a petition, attending a community meeting, or writing to an elected official).
-These effects are particularly pronounced for women of color.
-Although Walker and others demonstrate that these actions largely take place outside of the voting booth, they may moderate spillover effects on turnout.
+The little research looking for turnout effects at the local level, however, is similarly conflicting.
+Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that block groups in North Carolina with higher levels of incarceration had lower turnout.
+White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{a}), on the other hand, finds that incarceration has only moderate indirect effects on voting.
+She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (p.~607).
+Neither of these papers, however, account for felony disenfranchisement laws \emph{per se}, but rather test the bundled indirect effects of proximal contact with incarceration (inclusive of disenfranchisement).
+Disenfranchised individuals who would not have voted in the absence of their incarceration are not directly affected by disenfranchisement policy; any indirect turnout effects associated with their incarceration, therefore, cannot be attributed to disenfranchisement policy.
 
-Similarly, White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{b}) finds that incarceration has only moderate indirect effects on voting.
-She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (607).
-This study does not, however, interrogate whether the incarceration of a would-be voter has different indirect effects than the incarceration of an individual who would not have voted either way.
-It is possible that the indirect effects of the incarceration or jailing of a would-be voter are different than the indirect effects arising from other individuals.
-Nevertheless, the literature on the effects of proximal contact with the criminal justice system and felony disenfranchisement is hardly settled.
-
-\hypertarget{theoretical-implications-for-reduced-turnout}{%
-\section*{Theoretical Implications for Reduced Turnout}\label{theoretical-implications-for-reduced-turnout}}
-\addcontentsline{toc}{section}{Theoretical Implications for Reduced Turnout}
-
-As discussed above, it has been widely established that felony disenfranchisement reduces turnout even among eligible voters, particularly in the Black community (e.g.~King and Erickson \protect\hyperlink{ref-King2016}{2016}; Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, there has been little investigation into whether these depressive effects are narrowly concentrated in the neighborhoods home to disenfranchised individuals or are more widely dispersed.
-Understanding how these spillover effects are distributed is of practical importance.
-For instance, if neighborhoods home to disenfranchised individuals display different candidate preferences than the rest of the city, these demobilizing effects will undermine these neighborhoods' political representation.
-There is some evidence that spatial segregation leads to unique voting preferences among neighborhoods: Kinsella, McTague, and Raleigh (\protect\hyperlink{ref-Kinsella2015}{2015}) examines political clustering in the Greater Cincinnati Metropolitan Area from 1976 through 2008.
-Over this three-decade period, precinct-level presidential election results show trends toward increased polarization.
-
-Even if neighborhoods do not have unique preferences for candidates, lower turnout in impacted communities may reduce their allocation of public goods.
-At the Congressional level representatives direct resources to areas within their districts that provide the greatest political benefit --- that is to say, areas that will reward them with more votes (Martin \protect\hyperlink{ref-Martin2003}{2003}).
-Congressional representatives are also more responsive to the policy preferences of higher-turnout areas.
-``{[}H{]}igher citizen participation is rewarded,'' Martin and Claibourn (\protect\hyperlink{ref-Martin2013}{2013}) concludes, ``with enhanced policy responsiveness'' (59).
-Griffin and Newman (\protect\hyperlink{ref-Griffin2005}{2005}) finds similar effects in the United States Senate, demonstrating that ``voter preferences predict the aggregate roll-call behavior of Senators while nonvoter preferences do not'' (1206).
-If the neighborhoods most impacted by felony disenfranchisement turn out at lower rates, they may find that their elected representatives are less likely to support their needs and support their calls for greater public investment for amenities such as schools and parks.
-This may be true even if they support the same candidates as neighborhoods not touched by felony disenfranchisement.
-
-Although research on the impact of local turnout on city-wide policy is scarce, Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) examines the impact of senior turnout on ``senior-friendly'' policy at the city level.
-Anzia does not find that senior turnout in general increases the likelihood of senior-friendly policies, but that elected officials \emph{are} responsive to senior turnout when seniors ``are a more cohesive, meaningful group'' (1).
-American cities are highly segregated by race and by class, but less so by age.
-As such, Anzia's study does not speak directly to the impact of \emph{neighborhood} turnout rates on city policy.
-
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}), however, show that racial variation in turnout has real consequences for local elections and local political power.
-They estimate whether mayoral and city council races might have turned out differently if all races had turned out at the same rate, finding that three of the ten largest cities in America would likely have had different mayors if minorities had turned out as frequently as Whites.
-Of particular interest for the study at hand is their assertion that New York City would have elected a different mayor in 2001 if all racial groups had participated equally.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) tells us that ``Changes in the percentage of voters who turn out can and do alter mayoral election outcomes and racial representation on city councils'' (518).
-Neighborhoods most impacted by felony disenfranchisement are disproportionately home to residents of color, and they are likely to share concerns about criminal justice issue.
-Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) and Anzia (\protect\hyperlink{ref-Anzia2019}{2019}) both therefore indicate that depressed turnout from felony disenfranchisement is likely to materially reduce neighborhoods' political representation.
+This study addresses with both problems with the existing literature.
+I deal with geographic concentration by testing the relationship between felony disenfranchisement and turnout at the Census block group level.
+Secondly, I do not look for the indirect effects of proximal contact with \emph{any} individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual \emph{with a demonstrated history of voting}.
+As Gerber, Green, and Shachar (\protect\hyperlink{ref-Gerber2003}{2003}) and others indicate, past turnout is a very strong predictor of future turnout.
+Thus, by defining lost voters as individuals who previously participated, I can identify the individuals most likely to vote in the absence of felony disenfranchisement laws --- and therefore the individuals (and neighborhoods) most directly impacted by these laws.
+Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood police stop rates.
 
@@
 
-\hypertarget{geocoding}{%
-\subsection*{Geocoding}\label{geocoding}}
-\addcontentsline{toc}{subsection}{Geocoding}
+Further discussion about how these administrative records are linked and geocoded can be found in the Supplemental Information.
 
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1 percent of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6 percent of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+\hypertarget{research-design-and-expectations}{%
+\section*{Research Design and Expectations}\label{research-design-and-expectations}}
+\addcontentsline{toc}{section}{Research Design and Expectations}
 
-\hypertarget{matching}{%
-\subsection*{Matching}\label{matching}}
-\addcontentsline{toc}{subsection}{Matching}
+I begin with an ordinary least squares regression estimated at the block group level.
+In this regression, neighborhood turnout is the dependent variable, while the number of lost voters is the independent variable of interest.
+I look at the 2017 mayoral election in New York City, and define a ``lost voter'' as an individual incarcerated or on felony parole on Election Day in 2017 who cast a ballot sometime between 2007 and 2016.
+Although the 2017 mayoral election was a low turnout contest, Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) argues that these low turnout local elections are precisely where differential turnout rates can have the biggest consequences.
 
-Registered and formerly registered individuals who have been to prison in New York are identified by matching the NYSDOCCS records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a ``match,'' records must have the exact same birth date.
-The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
+A regression that looks at turnout only in one year can shed little light on the causal effect of felony disenfranchisement.
+After investigating whether neighborhoods with lost voters turned out at lower rates in 2017, I use data from the 2016 presidential election to ask whether the loss of a voter leads to depressed turnout.
+I limit the pool of block groups to block groups with no lost voters in 2016, and test whether neighborhoods that had a lost voter one year later saw turnout decrease relative to their own history in 2016, and the 2017 turnout of neighborhoods that did not lose a voter in that period.
+Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods.
+The matching procedure ensures that, prior to treatment, treated and control observations are substantially similar to one another based on observable characteristics.
 
-\begin{itemize}
-\tightlist
-\item
-  Middle names are identical. If neither set of records includes a middle name, this condition is met.
-\item
-  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-\item
-  A middle name or middle initial in one set of records, and a missing middle name in the other set.
-\end{itemize}
-
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
-
-There are two types of potential error in this methodology: a false positive will result when a formerly incarcerated individual's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 20,955\\
-Birthdate + 35 Days & 105\\
-Birthdate - 35 Days & 92\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
-A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
-Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
-
-\hypertarget{identification-of-lost-voters}{%
-\section*{Identification of Lost Voters}\label{identification-of-lost-voters}}
-\addcontentsline{toc}{section}{Identification of Lost Voters}
-
-Felony disenfranchisement is expected to lower neighborhood-level turnout through both direct and indirect mechanisms.
-By explicitly disenfranchising certain individuals, these laws prevent a certain number of ballots from being cast, though this number is likely small relative to the total electorate.
-Based on the aforementioned literature, I also expect disenfranchisement to have indirect effects on turnout --- that is to say, that eligible voters whose family members and neighbors are incarcerated are dissuaded from casting a ballot due because of their community member's incarceration.
-
-In this analysis, I offer a different definition of ``lost voter'' than much of the literature.
-Many recent papers have attempted to identify relationships between the number of disenfranchised residents --- \emph{potentially} lost voters --- and turnout.
-Such an approach is informative for understanding the impact of disenfranchisement.
-Many young men, for instance, are admitted to prison each year.
-If someone is incarcerated shortly after they turn 18 in an odd-numbered year, they may be incarcerated before they even have the opportunity to cast a ballot.
-Similarly, we know that older individuals are more likely to vote; if someone who is incarcerated would have ``aged in'' to voting but for their incarceration, they will not will have no history of voting.
-This analytical approach is also often taken due to data constraints: most states' registered voter files provide only a snapshot of currently registered voters, making it impossible to determine voting history for individuals who are currently disenfranchised (and therefore not currently registered).
-Similarly, state-level data does not report the number of incarcerated individuals with a history of voting; therefore, studies that leverage variation in laws between states to estimate the effect of felony disenfranchisement rely on estimates of the total disenfranchised population, not the number of disenfranchised individuals with a history of voting.
-
-Examining the impact of all disenfranchised voters, however, limits our ability to understand the specific direct and indirect causal effects of felony disenfranchisement.
-As previous research has detailed, turnout rates among individuals who are incarcerated are very low even prior to incarceration (e.g.~Gerber et al.
-\protect\hyperlink{ref-Gerber2017}{2017}).
-Research showing lower turnout where more individuals are formally disenfranchised might only be identifying the indirect effects of incarceration on turnout.
-Estimation of the effects of felony disenfranchisement above-and-beyond the effects of incarceration must account for the low propensity to vote among the formally disenfranchised.
-
-Here, I explore whether the disenfranchisement of residents with a history of participating in elections is related to neighborhood turnout.
-In this analysis, ``lost voters'' are individuals ineligible to cast a ballot on a given election day who have cast a ballot in the previous ten years.
-Although not all lost voters would have participated if they had not been disenfranchised, past participation is an extremely strong predictor of propensity to vote (Gerber, Green, and Shachar \protect\hyperlink{ref-Gerber2003}{2003}).
-As such, these are the individuals most likely to have been directly impacted not only by incarceration but specifically by felony disenfranchisement.
-Identifying the effects of felony disenfranchisement in neighborhoods that lost individuals with a record of voting --- individuals who would likely have cast a ballot had they been allowed --- provides insight into whether felony disenfranchisement reduces neighborhood turnout.
-Because New York State's voter file includes information on individuals who have been purged for felony convictions, I can reconstruct the vote history even for voters who are no longer eligible to vote.
-
-Of course, the incarceration of neighbors and family members who have \emph{not} voted in the past might also have indirect effects on neighborhood turnout rates.
-Even if an eligible New Yorker's incarcerated brother would not have voted, for instance, the eligible voter might not cast a ballot due to a soured impression of the political process.
-Such an indirect effect, however, would arise from the incarceration and not from the disenfranchisement of a voter.
-I expect that proximal contact with lost voters, which represent the most politically engaged subset of incarcerated individuals, will have much higher indirect effects than proximal contact with incarcerated individuals more generally.
-These higher indirect (depressive) effects likely arise through three mechanisms.
+I expect that neighborhoods with lost voters turned out at lower rates in 2017, and that the loss of a voter to felony disenfranchisement causes neighborhood turnout to decline.
+I expect these effects to operate through three primary mechanisms:
 
@@
 
-To study the effect of felony disenfranchisement on voting at the local level (and its potential implications for the distribution of political power at the local level), I use turnout in the most recent non-special election for city-wide office in New York --- the mayoral election which took place on November 7\textsuperscript{th}, 2017.
-Lost voters, therefore, are all individuals who were incarcerated or on parole on November 7\textsuperscript{th}, 2017, and had cast a ballot between 2007 and 2016.
-In many cases, these ``lost voters'' will no longer live in the neighborhoods that they are considered to be treating; they may be incarcerated, or have moved elsewhere after their incarceration.
-However, they need not be present and disenfranchised for their disenfranchisement to have indirect effects on turnout --- the forcible removal by the state of a would-be voter could leave a depressive effec in its wake.
+\hypertarget{results}{%
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
 
-Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included.
+Figure \ref{fig:citymap} shows where lost voters lived before going to prison, with city council districts also included.
 There were 2,518 such lost voters within New York City as of the 2017 general election.
 
-\begin{figure}[H]
+FIGURE \ref{fig:citymap} ABOUT HERE
 
-{\centering \includegraphics{p1_files/figure-latex/citywide-map-1}
-
-}
-
-\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
-\end{figure}
-
-Although 73,223 individuals were formally disenfranchised as of the 2017 election, there were just 6,166 disenfranchised voters statewide who had participated in the past 10 years.\footnote{Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.} Just 8.4 percent of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
+Although 73,223 individuals were formally disenfranchised as of the 2017 election, there were just 6,166 disenfranchised voters statewide who had participated in the past 10 years.\footnote{Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.} Just 8.4\% of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
 
@@
 In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of individuals with a demonstrated history of voting were not allowed to cast a ballot for mayor.
+In other communities, such as Harlem and Central Brooklyn, large numbers of past voters were not allowed to cast a ballot for mayor.
 
-\hypertarget{testing-for-neighborhood-turnout-effects}{%
-\section*{Testing for Neighborhood Turnout Effects}\label{testing-for-neighborhood-turnout-effects}}
-\addcontentsline{toc}{section}{Testing for Neighborhood Turnout Effects}
+Table \ref{tab:demos} demonstrates just how different the neighborhoods with lost voters are from the rest of the city.
+The first column in Table \ref{tab:demos} weights the chacteristics of each block group in the city by its population; the second column weights the block groups by their number of lost voters.
+The median income in neighborhoods with lost voters is less than three-quarters that of unaffected neighborhoods.
+The average lost voter came from a neighborhood that was nearly twice as Black as the city as a whole; their neighborhoods were also home to a far larger share of Democratic voters than other neighborhoods.
+Unsurprisingly, lost voters also came from neighborhoods with 60\% more police stops than the citywide average.
 
-Table \ref{tab:nhood-dist} shows the distribution of the number of lost voters by neighborhood.
-Most neighborhoods lost no voters at all; when a neighborhood was home to lost voters, they generally lost just one resident with a history of voting.
-More than two-thirds of block groups that lost any voters lost only one voter, while 45 percent of census tracts with a lost voter lost only one.
+TABLE \ref{tab:demos} ABOUT HERE
 
-\begin{table}[H]
+\hypertarget{testing-for-neighborhood-turnout-differences}{%
+\subsection*{Testing for Neighborhood Turnout Differences}\label{testing-for-neighborhood-turnout-differences}}
+\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Differences}
 
-\caption{\label{tab:distribution-lost}\label{tab:nhood-dist} Distribution of Lost Voters}
-\centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{ccc}
-\toprule
-Number of Lost Voters & Count of Block Groups & Count of Tracts\\
-\midrule
-0 & 4,272 & 1,008\\
-1 & 1,023 & 416\\
-2 & 301 & 212\\
-3 & 107 & 138\\
-4 & 30 & 51\\
-5 & 19 & 46\\
-6 & 9 & 23\\
-7 & 3 & 15\\
-8 & 5 & 13\\
-9 & 0 & 7\\
-10 & 0 & 2\\
-11 & 0 & 2\\
-12 & 0 & 2\\
-14 & 0 & 2\\
-\hline
-Total & 5,769 & 1,937\\
-\bottomrule
-\end{tabular}
-\end{table}
+I begin by asking whether turnout was lower in the 2017 mayoral election in neighborhoods with lost voters.
+Block group turnout rates are calculated using the geocoded voter file.
+Each voter's record indicates whether the voter participated in the 2017 general election.
+These records are then aggregated to estimate the number of ballots cast in each neighborhood.
+The number of ballots cast is divided by the neighborhood's citizen voting age population (CVAP, estimated by the Census Bureau).
+The independent variable of interest measures the number of lost voters in a neighborhood.
 
-I begin by testing the effect of lost voters on neighborhood turnout using a matching model.
-Neighborhoods (defined as census tracts and block groups) are considered treated if they were home to \emph{at least one} lost voter in the 2017 election; they are untreated if no voters were disqualified from the election because of a felony conviction.
-I use a genetic match to match treated to untreated census tracts and block groups (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}), using a series of demographic and political indicators.
-Estimates of racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau's 5-year American Community Survey estimates.
+I also control for other social and political indicators: neighborhood racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
+As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau's 2017 5-year American Community Survey estimates.
 Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the citizen voting age population (CVAP, estimated by the Census Bureau).
+Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the CVAP.
 I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
-Each treated block group is matched to 30 untreated block groups, and each treated census tract is matched with 10 other tracts.
-Matching is done with replacement.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} present the results of these matches.
 
-\hypertarget{match-output}{%
-\subsection*{Match Output}\label{match-output}}
-\addcontentsline{toc}{subsection}{Match Output}
+To completely decompose the effects of disenfranchisement from incarceration more generally, we would need to control for the number of incarcerated or paroled individuals in a neighborhood.
+If, after controlling for a neighborhood's general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
 
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk}\label{tab:blockgroup-table} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.34 & 93.02 & 89.13 & 85.79 & 79.01\\
-\% Non-Hispanic Black & 0.38 & 0.16 & 0.38 & 0.37 & 97.93 & 94.46 & 94.41 & 92.77\\
-\% Non-Hispanic White & 0.17 & 0.40 & 0.17 & 0.18 & 96.37 & 96.06 & 95.56 & 92.99\\
-Median Income & 51,792.52 & 72,612.36 & 51,792.52 & 52,768.97 & 95.31 & 93.47 & 88.62 & 74.49\\
-\% With Some College & 0.64 & 0.70 & 0.64 & 0.64 & 90.75 & 94.14 & 90.29 & 80.70\\
-Median Age & 35.99 & 38.34 & 35.99 & 36.25 & 89.18 & 90.84 & 85.93 & 78.56\\
-Registration Rate & 1.00 & 0.94 & 1.00 & 0.98 & 59.77 & 79.89 & 72.28 & 60.42\\
-\% Democrats & 0.75 & 0.65 & 0.75 & 0.75 & 96.22 & 96.52 & 94.92 & 91.02\\
-\% Noncitizen & 0.16 & 0.16 & 0.16 & 0.16 & 16.58 & 14.80 & -0.06 & -60.28\\
-\% Won by City Council Representative & 0.85 & 0.81 & 0.85 & 0.86 & 89.44 & 83.40 & 79.88 & 70.11\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\begin{table}[H]
-
-\caption{\label{tab:tract-table-chunk}\label{tab:tract-table} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.32 & 0.22 & 0.32 & 0.33 & 98.55 & 85.12 & 82.84 & 73.01\\
-\% Non-Hispanic Black & 0.36 & 0.14 & 0.36 & 0.36 & 99.77 & 88.74 & 88.36 & 83.30\\
-\% Non-Hispanic White & 0.19 & 0.42 & 0.19 & 0.19 & 97.50 & 96.32 & 91.29 & 79.95\\
-Median Income & 53,409.42 & 71,754.50 & 53,409.42 & 55,422.41 & 89.03 & 90.83 & 84.86 & 62.84\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 90.19 & 93.52 & 88.89 & 73.01\\
-Median Age & 35.76 & 38.04 & 35.76 & 35.80 & 98.12 & 74.51 & 80.56 & 71.91\\
-Registration Rate & 0.94 & 0.90 & 0.94 & 0.93 & 76.63 & 86.16 & 84.24 & 71.28\\
-\% Democrats & 0.74 & 0.62 & 0.74 & 0.73 & 94.58 & 94.93 & 92.05 & 82.37\\
-\% Noncitizen & 0.16 & 0.17 & 0.16 & 0.17 & -123.57 & 22.48 & 8.85 & -38.79\\
-\% Won by City Council Representative & 0.85 & 0.79 & 0.85 & 0.85 & 93.03 & 81.57 & 79.27 & 67.47\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-At both the tract and block group level, matching results in an untreated group of neighborhoods that looks substantially like the treatment group.
-
-In addition to demonstrating the quality of the genetic match, these tables detail the striking extent to which neighborhoods with lost voters differ from the rest of the city.
-On average, the non-Hispanic Black share of a block group with a lost voter is more than \emph{double} the share in block groups without lost voters.
-These impacted neighborhoods are also less than half as non-Hispanic white as unaffected neighborhoods.\footnote{Throughout the remainder of this project, I use White and non-Hispanic White interchangeably.
-Similarly, Black and non-Hispanic Black are used interchangeably, while Latino indicates Latinos of any race.} The median income in neighborhoods without lost voters is more than 40 percent higher.
-Tables \ref{tab:blockgroup-table} and \ref{tab:tract-table} throw into sharp relief the types of neighborhoods where likely would-be voters are being disenfranchised due to incarceration and parole.
-
-After matching neighborhoods, I use a simple regression to test whether neighborhood treatment status was associated with turnout in the 2017 mayoral election.
-Census tract and block group level turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's CVAP.
+Unfortunately, such data does not exist; the State of New York does not make pre-incarceration addresses available to the public.
+In its place, I use data from the NYPD on stop-and-frisk encounters with the police.
+These models include the number of times residents were stopped by the police from January -- October of 2017.
+This is intended to control for general neighborhood exposure to the carceral state.
 
 Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-I therefore include models which explore any potential difference in treatment effect in neighborhoods where a higher share of the population is Black.
-Table \ref{tab:match-reg} presents the results of these regression models.
-Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
+Model 2, therefore, interacts the number of lost voters with the share of the block group that is non-Hispanic Black.
 
-\input{"../../temp/table1.tex"}
-
-When neighborhoods are measured at the block group level, a lost voter is negatively associated with turnout in the 2017 election.
-In block groups with lost voters, turnout was on average 0.9 percentage points lower than in comparable block groups without lost voters.
-This decrease, however, appears to be entirely concentrated within Black neighborhoods.
-When the treatment indicator is interacted with the share of the neighborhood that is non-Hispanic Black, the basic treatment dummy becomes nonsignificant.
-The coefficient on the interaction between treatment and share Black indicates that treated neighborhoods that are largely Black saw turnout that was as much as 2.7 percentage points lower than similar neighborhoods without lost voters.
-Considering that the overall turnout rate in block groups with a lost voter was just 14.6 percent, this effect is alarmingly high.
-For every 100 votes cast in an entirely Black block group with a lost voter, as many as 18.5 votes went uncast.
-The effects are slightly larger when neighborhoods are measured as block groups than as census tracts, though turnout is still significantly depressed in tracts with lost voters.
-This is not surprising --- because the spillover effects are likely to operate through social networks, smaller geographical units are likely to be more affected.
-
-\hypertarget{testing-intensity-effects}{%
-\subsection*{Testing Intensity Effects}\label{testing-intensity-effects}}
-\addcontentsline{toc}{subsection}{Testing Intensity Effects}
-
-Matching methodologies, of course, only allow us to test the effect of being treated --- here, losing any voter for the 2017 election.
-The models above do not allow for different effects on turnout based on \emph{how many} voters a neighborhood lost, potentially understating the impact of felony disenfranchisement in the most hard-hit communities.
-In Table \ref{tab:trad-reg} below, I adopt a standard ordinary least squares regression to investigate whether lost voters are associated with lower turnout rates in the 2017 election.
-This regression uses the same covariates used in the matching procedure described above.
+Table \ref{tab:trad-reg} presents the results of ordinary least squares regressions testing neighborhood turnout rates in 2017.
 Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
 
-\input{"../../temp/table2.tex"}
+TABLE \ref{tab:trad-reg} ABOUT HERE
 
-The results presented in Table \ref{tab:trad-reg} align closely with the estimated effect from the matching model.
-Lost voters are generally associated with lower turnout (each missing voter in a block group reduces that neighborhood's turnout by about 0.82 percentage points), but Models 2 and 4 again make clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.87 percentage points.
+The results presented in Table \ref{tab:trad-reg} indicate that neighborhoods with lost voters turned out at lower rates in the 2017 election.
+Each missing voter in a block group reduces that neighborhood's turnout by about 0.8 percentage points.
+Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
+In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.93 percentage points.
 The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
@@
 
-\begin{figure}[H]
+FIGURE \ref{fig:postest} ABOUT HERE
 
-{\centering \includegraphics{p1_files/figure-latex/postest-map-1}
+\hypertarget{toward-a-causal-estimate}{%
+\subsection*{Toward a Causal Estimate}\label{toward-a-causal-estimate}}
+\addcontentsline{toc}{subsection}{Toward a Causal Estimate}
 
-}
+As the analysis above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
+The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
+Moreover, by looking at turnout from just one election, the best we could hope to uncover is a correlation between disenfranchisement and turnout.
+Uncovering the causal effect requires looking across elections (or the very strong assumptions that all relevant neighborhood characteristics are accounted for).
 
-\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
-\end{figure}
-
-\hypertarget{toward-a-causal-estimation}{%
-\subsection*{Toward a Causal Estimation}\label{toward-a-causal-estimation}}
-\addcontentsline{toc}{subsection}{Toward a Causal Estimation}
-
-As the analyses above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-If that were the case, we would expect that these treated neighborhoods would have lower turnout regardless of whether they lost a voter.
-As the analysis that follows shows, this is not the case: prior to losing a voter, treated neighborhoods did not have systematically lower turnout than other neighborhoods.
-
-I begin by restricting the set of neighborhoods to only neighborhoods that, on election day in 2016, had no lost voters.
+In order to estimate the causal effect of felony disenfranchisement on neighborhood-level turnout, I now test whether losing a voter between the 2016 and 2017 elections resulted in decreased turnout in 2017.
+I begin by restricting the set of neighborhoods to only neighborhoods that had no lost voters on election day in 2016.
 This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
 These individuals' home neighborhoods are removed from the subsequent analysis.
-I then use the genetic match algorithm discussed above to match neighborhoods with a lost voter in 2017 to neighborhoods with no lost voters in 2017.
-These neighborhoods, therefore, were \emph{newly} home to lost voters.
-If lost voters have no effect on neighborhood turnout, we would expect these treated neighborhoods to have lower turnout in both 2016 --- before they lost a voter --- and in 2017, after a voter was disenfranchised.
-In 2017, 1,497 neighborhoods had a lost voter; of these, 326 had no lost voters in 2016 and are included here.
-The original set of control block groups included 4,272 neighborhoods with no lost voters; of these, 211 \emph{did} have a lost voter in 2016, and are thus excluded from this analysis.
+I then identify the neighborhoods that did have lost voters in 2017; these block groups are our ``treated'' block groups, while the neighborhoods with lost voters in neither 2016 nor 2017 serve as our controls.
 
-Of course, felony disenfranchisement could structure neighborhood turnout differently for presidential and local elections.
-It is conceivable that the incarceration of a family member or neighbor would sour individuals on participating in local contests, but that the same individual would not read national politics through the same lens.
-Nevertheless, data limitations makes the 2016 election the only reasonable comparator for 2017 turnout.
+Of course, as demonstrated in Table \ref{tab:demos}, neighborhoods with lost voters are very different than neighborhoods without lost voters.
+Control and treatment neighborhoods might have different levels of turnout in 2017 even in the absence of felony disenfranchisement.
+To account for this possibility, I use a genetic matching model (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}) to match each treated block group to one untreated block group, using the same characteristics detailed in Table \ref{tab:trad-reg} but based on their 2016 characteristics.
+Matched ties are not broken, leaving the potential for some treated block groups to match to more than one control; the observations are weighted accordingly.
+Only treated block groups and the untreated block groups with which they match --- the controls --- are retained.
 
-Tables \ref{tab:did-match} and \ref{tab:did-match-tract} present the results of matching on these subsets of neighborhoods.
-As before, the match results in a highly balanced panel between treated and untreated neighborhoods.
+Although it may seem counterintuitive to discard some potential control block groups, doing so gives us a major advantage: it allows us to conceptualize the ``treatment'' (the loss of a voter) as a natural experiment.
+Because the treatment and control groups look similar based on their observable 2016 characteristics, the treatment can be understood as quasi-randomly assigned.
 
-\begin{table}[H]
+Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods.
+The matching procedure substantially improves the similarly between treated and control block groups along all characteristics.
+Table \ref{tab:did-match} mirrors Table \ref{tab:demos} above: treated neighborhoods are far less White, had many more stop-and-frisk encounters, and had much lower median incomes.
 
-\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.35 & 0.25 & 0.35 & 0.35 & 99.00 & 96.26 & 94.92 & 85.60\\
-\% Non-Hispanic Black & 0.33 & 0.16 & 0.33 & 0.33 & 99.02 & 96.57 & 95.67 & 90.10\\
-\% Non-Hispanic White & 0.21 & 0.41 & 0.21 & 0.21 & 99.61 & 94.66 & 92.61 & 82.94\\
-Median Income & 55,852.62 & 73,433.49 & 55,852.62 & 56,287.76 & 97.52 & 87.85 & 86.56 & 72.89\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 99.31 & 92.25 & 87.30 & 71.15\\
-Median Age & 36.17 & 38.47 & 36.17 & 36.04 & 94.37 & 62.35 & 72.27 & 69.84\\
-Registration Rate & 0.98 & 0.94 & 0.98 & 0.95 & 25.09 & 46.49 & 42.16 & 33.06\\
-\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 95.94 & 94.99 & 93.62 & 86.68\\
-\% Noncitizen & 0.17 & 0.16 & 0.17 & 0.17 & 63.46 & 75.01 & 66.85 & 45.71\\
-\% Won by City Council Representative & 0.85 & 0.80 & 0.85 & 0.85 & 88.19 & 83.99 & 78.23 & 67.54\\
-\bottomrule
-\end{tabular}}
-\end{table}
+TABLE \ref{tab:did-match} ABOUT HERE
 
-\begin{table}[H]
+In Table \ref{tab:did-reg}, I present the outcome of an OLS regression on this subset of block groups.
+The variable \emph{D(2017)} tests whether turnout for control block groups declined in 2017 relative to 2016.
+I test whether there was a difference in turnout between control and treatment neighborhoods in 2016 with \emph{D(Treat)}.
+Of most interest is the coefficient on \emph{D(Treat) \(\times\) D(2017)}, which tests whether turnout decreased more from 2016 to 2017 in treated block groups than in the controls.
+Model 1 includes only these covariates (and fixed effects for Congressional district to account for potential differences in competitiveness in 2016), while Model 2 also includes the variables that were used for the matching procedure.
+Since each neighborhood appears in the panel twice --- once in 2016, and once in 2017 --- each observation takes the characteristics of the year in which it appears.
+This is especially important for the stop-and-frisk variable: Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
+Model 3 adds block group-level fixed effects to Model 2 in order to control for time-invariant neighborhood characteristics.
+Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
 
-\caption{\label{tab:tract-chunk-did}\label{tab:did-match-tract} Results of Tract-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-\% Latino & 0.31 & 0.22 & 0.31 & 0.30 & 94.34 & 86.10 & 84.77 & 75.82\\
-\% Non-Hispanic Black & 0.25 & 0.12 & 0.25 & 0.25 & 99.28 & 89.75 & 86.62 & 77.90\\
-\% Non-Hispanic White & 0.29 & 0.44 & 0.29 & 0.28 & 97.02 & 87.45 & 77.73 & 61.67\\
-Median Income & 59,990.43 & 72,565.90 & 59,990.43 & 61,973.89 & 84.23 & 78.35 & 76.98 & 66.57\\
-\% With Some College & 0.67 & 0.71 & 0.67 & 0.66 & 96.86 & 86.51 & 83.73 & 76.55\\
-Median Age & 36.36 & 38.23 & 36.36 & 36.39 & 98.60 & 73.15 & 69.00 & 42.11\\
-Registration Rate & 0.92 & 0.90 & 0.92 & 0.91 & 10.68 & 64.27 & 63.46 & 56.35\\
-\% Democrats & 0.69 & 0.62 & 0.69 & 0.68 & 91.23 & 90.06 & 87.35 & 76.31\\
-\% Noncitizen & 0.18 & 0.17 & 0.18 & 0.18 & 78.09 & 73.55 & 62.03 & 48.36\\
-\% Won by City Council Representative & 0.83 & 0.79 & 0.83 & 0.83 & 99.91 & 75.59 & 65.98 & 38.91\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-In Table \ref{tab:did-reg}, I present the outcome of a simple least squares regression on this subset of block groups and tracts.
-In both models, D(2017) indicates that across the board turnout in 2017 was far lower than in 2016.
-D(Treat) can be understood as the difference between treated and control neighborhoods in the 2016 election.
-In Model 1, the treated and control neighborhoods did not have statistically significantly different turnout in the 2016 election; Model 2 indicates that treated census tracts had slightly \emph{higher} turnout prior to intervention.
-D(Treat) \(\times\) D(2017) is the coefficient of most interest here: it indicates whether, after treatment, treated neighborhoods turned out at lower rates than similar neighborhoods.
-In Model 1, this coefficient is negative and statistically significant, and is very similar to the coefficients reported in Tables \ref{tab:match-reg} and \ref{tab:trad-reg}.
-Model 1 in Table \ref{tab:did-reg} indicates that, in the first election after the removal of a voter, a neighborhood's turnout rate decreased by 2.6 percentage points.
-The relationship does not hold for tracts (suprisingly, treated tracts turned out at higher rates both before and after treatment than control tracts).
-
-\input{"../../temp/table1111.tex"}
+TABLE \ref{tab:did-reg} ABOUT HERE
 
 Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only neighborhoods with no lost voters in 2016.
+In each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
+It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016 who lost one in the subsequent year.
 These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-Indeed, neighborhoods that consistently have lost voters are more likely to be the places most impacted by policing patterns.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16 percent versus 21 percent), a lower share non-Hispanic Black (39 percent versus 33 percent), and a smaller share were registered as Democrats (76 percent versus 73 percent).\footnote{Each of these differences are significant at the 99 percent confidence level.} Each of these differences were even more magnified when comparing tracts included in this analysis to all tracts with lost voters in 2017.
+The marginal effect of a lost voter could be lower in neighborhoods that consistently have lost voters and thus have less to ``learn'' about the state from the loss of any one individual to incarceration.
+Neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
+Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16\% versus 21\%), a lower share non-Hispanic Black (39\% versus 32\%), and a smaller share were registered as Democrats (76\% versus 73\%).\footnote{Each of these differences are significant at the 99\% confidence level.}
 
@@
 However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large implications on its own.
+Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large direct effects on turnout and representation.
 
 As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Previous literature has established that felony disenfranchisement impacts Black turnout at the state level, but this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40 percent higher than the median income of block groups with lost voters.
-Similarly, just 17 percent of individuals in the average block group with a lost voter were White, compared with 40 percent of the population in other block groups.
+Despite methodological problems, previous literature has hinted that felony disenfranchisement impacts Black turnout at the state level; this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
+As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40\% higher than the median income of block groups with lost voters.
+Similarly, just 17\% of individuals in the average block group with a lost voter were White, compared with 40\% of the population in other block groups.
 Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
@@
 Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Both Tables \ref{tab:match-reg} and \ref{tab:trad-reg} indicate that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is either nonsignificant or significant and positive.
-This means that lost voters have no spillover effects in neighborhoods with small Black populations.
+Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is nonsignificant.
+This means that lost voters have little spillover effects in neighborhoods with small Black populations.
 Understanding the magnitude of this finding is important.
-In New York City, there are 1386 block groups where the Black community makes up more than 38 percent of the population (the average for treated block groups), with an average of 912 voting age citizens.
-Model 2 of Table \ref{tab:match-reg} indicates that, in a block group that is 38 percent Black with a citizen voting age population of 912, having any lost voter reduced turnout by 9.4 ballots.
-Conversely, in a block group with the same voting age population that is just 16 percent Black (the average for untreated block groups), the loss of any voter decreased turnout by just 3.8 ballots.
-
-The standard OLS regression results tell a similar story: in our average block group that is 38 percent Black, each lost voter cost the neighborhood 6.5 ballots.
-In block groups of the same size where the population is just 16 percent Black, lost voters cost the neighborhood just 1.6 ballots --- hardly more votes than the lost voter herself.
+In New York City, there are 1,386 block groups where the Black community makes up more than 38\% of the population (the average for treated block groups), with an average of 912 voting age citizens.
+Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38\% Black with a citizen voting age population of 912, having any lost voter reduced turnout by 6.7 ballots.
+Conversely, in a block group with the same voting age population that is just 16\% Black (the average for untreated block groups), the loss of any voter decreased turnout by just 1.7 ballots --- hardly more votes than the lost voter herself.
 
@@
 Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, Weaver and Lerman (\protect\hyperlink{ref-Weaver2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where police presence and criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
+Similarly, Lerman and Weaver (\protect\hyperlink{ref-Lerman2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (p.~16).
+It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
 
@@
 
-This finding mirrors White (\protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{a}), which finds that brief jail spells decrease future participation more for Black individuals than for White individuals.
-This is perhaps unsurprising.
 A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans (e.g.~Browning and Cao \protect\hyperlink{ref-Browning1992}{1992}; Hurwitz and Peffley \protect\hyperlink{ref-Hurwitz2005}{2005}; Henderson et al.
@@
 There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
+
+Ultimately, the analysis at hand cannot answer the causal mechanism for these depressive effects, or why the depressive effect is concentrated in Black neighborhoods.
+As Walker (\protect\hyperlink{ref-Walker2020}{2020}) notes, ``contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact'' (p.~123).
+Administrative data can only get us so far.
+Other research designs are necessary to investigate how neighbors understand the incarceration of their neighbors, how they react on voting day, and whether these effects are racially moderated.
+Much of the work being done by Hannah Walker and coauthors (Walker \protect\hyperlink{ref-Walker2014}{2014}, \protect\hyperlink{ref-Walker2020}{2020}; Walker and García-Castañon \protect\hyperlink{ref-Walker2017}{2017}) does point to racial differences in political responses to proximal contact with systems of incarceration.
+
+This paper uncovers surprisingly large effects.
+It is possible that, despite controlling for neighborhood exposure to the carceral state using stop-and-frisk data, the analysis cannot entirely decompose the effect of disenfranchisement from the bundled experience of seeing a neighbor go to prison.
+Future research must gather better data on local patterns of incarceration in order to better distinguish turnout patterns in neighborhoods where voters are sent to prison from neighborhoods where non-voters are incarcerated.
+Future work should similarly test whether these effects, identified in a relatively uncompetitive local election, are also present in competitive and national elections.
 
@@
 As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, politicing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
+The interlocking nature of racioeconomic segregation, policing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
+
+\newpage
+
+\hypertarget{supplemental-information}{%
+\section*{Supplemental Information}\label{supplemental-information}}
+\addcontentsline{toc}{section}{Supplemental Information}
+
+I here present more information on how the full data set used in the body of this paper was constructed.
+
+\hypertarget{geocoding}{%
+\subsection*{Geocoding}\label{geocoding}}
+\addcontentsline{toc}{subsection}{Geocoding}
+
+Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
+I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
+This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1\% of registered voters.
+The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6\% of these individuals were not geocoded.
+Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
+
+\hypertarget{matching}{%
+\subsection*{Matching}\label{matching}}
+\addcontentsline{toc}{subsection}{Matching}
+
+I construct the primary dataset by matching the administrative criminal supervision records with the registered voter file.
+I match using first, middle, and last names, as well as dates of birth.
+Capitalization in names in both datasets is standardized, and punctuation and spaces are removed.
+If a middle name is missing in one dataset but not in the other, I consider this a match.
+Similarly, if one dataset includes only a middle initial which matches to the first letter of the middle name the other data set, this is considered a match.
+First and last names, and dates of birth, must match exactly between the two datasets.
+
+To test for false positive matches, I employ the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}).
+I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
+Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
+This analysis indicates that false positives account for between 0.4 and 0.5\% of all matches, a share that is likely too small to have any material impact on the overall analysis.
+The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
+
+TABLE \ref{tab:change-dobs} ABOUT HERE
+
+Testing for false negatives is more challenging.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6\% of individuals who have been discharged from felony parole.
+Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``Alex'' in one set of records and ``Alexander'' in the other).
+
+\theendnotes
 
@@
 
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\noindent
-
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-Abadie2019}{}%
-Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \url{https://scholar.harvard.edu/spiess/publications/robust-post-matching-inference.}
-
-\leavevmode\hypertarget{ref-Anzia2019}{}%
-Anzia, Sarah F.
-2019.
-``When Does a Group of Citizens Influence Policy?
-Evidence from Senior Citizen Participation in City Politics.'' \emph{The Journal of Politics} 81 (1): 1--14.
-\url{https://doi.org/10.1086/699916}.
+Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \emph{Working Paper.}
 
@@
 2009.
-``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons.'' \emph{Social Science Quarterly} 90 (3): 722--43.
+``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons*.'' \emph{Social Science Quarterly} 90 (3): 722--43.
 \url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
-
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Brennan Center for Justice. 2018. ``Criminal Disenfranchisement Laws Across the United States.'' \url{https://www.brennancenter.org/criminal-disenfranchisement-laws-across-united-states}.
 
@@
 
-\leavevmode\hypertarget{ref-Drucker2005}{}%
-Drucker, Ernest, and Ricardo Barreras.
-2005.
-``Studies of Voting Behavior and Felony Disenfranchisement Among Individuals in the Criminal Justice System in New York, Connecticut, and Ohio.'' Research report.
-Sentencing Project.
-\url{https://www.prisonpolicy.org/scans/sp/fd_studiesvotingbehavior.pdf}.
-
 \leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516. \url{https://doi.org/10.2307/2146812}.
+Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516--29. \url{https://doi.org/10.2307/2146812}.
 
@@
 Hendry.
-2014.
-``Can Incarcerated Felons Be (Re)integrated into the Political System?
+2015.
+``Can Incarcerated Felons Be (Re)Integrated into the Political System?
 Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
 \url{https://doi.org/10.1111/ajps.12166}.
-
-\leavevmode\hypertarget{ref-Gerber2017}{}%
----------.
-2017.
-``Does Incarceration Reduce Voting?
-Evidence About the Political Consequences of Spending Time in Prison.'' \emph{The Journal of Politics} 79 (4): 1130--46.
-\url{https://doi.org/10.1086/692670}.
 
@@
 
-\leavevmode\hypertarget{ref-Griffin2005}{}%
-Griffin, John D., and Brian Newman. 2005. ``Are Voters Better Represented?'' \emph{The Journal of Politics} 67 (4): 1206--27. \url{https://doi.org/10.1111/j.1468-2508.2005.00357.x}.
-
 \leavevmode\hypertarget{ref-Guest1999}{}%
-Guest, Avery M., and Susan K. Wierzbicki. 1999. ``Social Ties at the Neighborhood Level.'' \emph{Urban Affairs Review} 35 (1): 92--111. \url{https://doi.org/10.1177/10780879922184301}.
+Guest, Avery M., and Susan K.
+Wierzbicki.
+1999.
+``Social Ties at the Neighborhood Level: Two Decades of GSS Evidence.'' \emph{Urban Affairs Review}, September.
+\url{https://doi.org/10.1177/10780879922184301}.
 
 \leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
+Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy: Race, Turnout, and Representation in City Politics}. Cambridge ; New York: Cambridge University Press.
 
@@
 
+\leavevmode\hypertarget{ref-bcj_laws}{}%
+Justice, Brennan Center for.
+2019.
+``Criminal Disenfranchisement Laws Across the United States.'' \emph{Brennan Center for Justice}, May.
+\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
+
 \leavevmode\hypertarget{ref-Kenny1992}{}%
-Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259. \url{https://doi.org/10.2307/2111432}.
+Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259--67. \url{https://doi.org/10.2307/2111432}.
 
 \leavevmode\hypertarget{ref-Keyssar2009}{}%
-Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. New York: Basic Books.
+Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. Revised edition. New York: Basic Books.
 
 \leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821. \url{https://doi.org/10.1177/0021934716659195}.
+King, Bridgett A., and Laura Erickson.
+2016.
+``Disenfranchising the Enfranchised: Exploring the Relationship Between Felony Disenfranchisement and African American Voter Turnout.'' \emph{Journal of Black Studies}, July.
+\url{https://doi.org/10.1177/0021934716659195}.
 
@@
 ``Staying Out of Sight?
-Concentrated Policing and Local Political Action.'' Edited by Christopher Wildeman, Jacob S.
-Hacker, and Vesla M.
-Weaver.
-\emph{The ANNALS of the American Academy of Political and Social Science} 651 (1): 202--19.
+Concentrated Policing and Local Political Action:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
 \url{https://doi.org/10.1177/0002716213503085}.
 
+\leavevmode\hypertarget{ref-Lerman2014}{}%
+Lerman, Amy E., and Vesla M.
+Weaver.
+2014.
+\emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}.
+Chicago Studies in American Politics.
+Chicago ; London: The University of Chicago Press.
+
 \leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2006. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. New York: Oxford University Press.
+Manza, Jeff, and Christopher Uggen. 2008. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. Studies in Crime and Public Policy. New York: Oxford University Press.
 
@@
 2003.
-``Votings Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
+``Voting's Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
 \url{https://doi.org/10.1111/1540-5907.00008}.
@@
 2011.
-``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (7).
+``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (1): 1--52.
 \url{https://doi.org/10.18637/jss.v042.i07}.
@@
 ``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-Sentencing Project.
+The Sentencing Project.
 \url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
-
-\leavevmode\hypertarget{ref-Uggen2002}{}%
-Uggen, Christopher, and Jeff Manza.
-2002.
-``Democratic Contraction?
-Political Consequences of Felon Disenfranchisement in the United States.'' \emph{American Sociological Review} 67 (6): 777.
-\url{https://doi.org/10.2307/3088970}.
-
-\leavevmode\hypertarget{ref-Uggen2004}{}%
----------.
-2004.
-``Lost Voices: The Civic and Political Views of Disfranchised Felons.'' In \emph{Imprisoning America: The Social Effects of Mass Incarceration}, edited by Mary Pattillo, David Weiman, and Bruce Western, 165--204.
-New York: Russell Sage Foundation.
 
@@
 2014.
-``Extending the Effects of the Carceral State.
-Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly} 67 (4): 809--22.
+``Extending the Effects of the Carceral State: Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly}, July.
 \url{https://doi.org/10.1177/1065912914542522}.
 
+\leavevmode\hypertarget{ref-Walker2020}{}%
+---------. 2020. ``Targeted: The Mobilizing Effect of Perceptions of Unfair Policing Practices.'' \emph{The Journal of Politics} 82 (1): 119--34. \url{https://doi.org/10.1086/705684}.
+
 \leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela Garcı́a-Castañon.
+Walker, Hannah L., and Marcela García-Castañon.
 2017.
-``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (04): 541--68.
-\url{https://doi.org/10.1017/s1743923x17000198}.
-
-\leavevmode\hypertarget{ref-Weaver2014}{}%
-Weaver, Vesla, and Amy E. Lerman. 2014. \emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}. Chicago: University of Chicago Press.
+``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (4): 541--68.
+\url{https://doi.org/10.1017/S1743923X17000198}.
 
 \leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/s0003055410000456}.
+Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/S0003055410000456}.
+
+\leavevmode\hypertarget{ref-White2019a}{}%
+White, Ariel.
+2019a.
+``Family Matters?
+Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
+\url{https://doi.org/10.1017/S0003055418000862}.
 
 \leavevmode\hypertarget{ref-White2019}{}%
-White, Ariel.
-2019a.
+---------.
+2019b.
 ``Misdemeanor Disenfranchisement?
 The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/s000305541800093x}.
-
-\leavevmode\hypertarget{ref-White2019a}{}%
----------.
-2019b.
-``Family Matters?
-Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
-\url{https://doi.org/10.1017/s0003055418000862}.
+\url{https://doi.org/10.1017/S000305541800093X}.
 
 \leavevmode\hypertarget{ref-Wilson1990}{}%
-Wilson, William Julius. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Chicago: University Of Chicago Press.
+Wilson, William J. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Paperback ed., {[}Nachdr.{]}. Chicago: Univ. of Chicago Press.
 

### 2020-04-01 markdown/part1/title.Rmd
uar re-submit

--- 2020-04-01 markdown/part1/UAR_submission.Rmd
@@
 output:
-  bookdown::pdf_document2:
-    toc: no
-    fig_caption: yes
-    latex_engine: xelatex
-    keep_tex: true
+  word_document
 bibliography: "../nyc_neighborhoods.bib"
@@
     - \usepackage{rotating}
-    - \usepackage{setspace}\doublespacing
     - \usepackage{endnotes}
@@
 I map these individuals to their pre-incarceration addresses and use multiple approaches to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
-I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+I find that neighborhoods that were home to lost voters turned out at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
 These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
 ---
-\pagenumbering{gobble}
-\newpage
-\pagenumbering{arabic}
-```{r setup, echo = F, message = F, warning = F}
-knitr::opts_knit$set(root.dir = "../..", dev = "cairo_pdf")
-library(systemfit)
-library(AER)
-library(sandwich)
-library(lmtest)
-library(scales)
-library(tidyverse)
-library(extrafont)
-library(kableExtra)
-library(stargazer)
-library(data.table)
-loadfonts()
-```
-\begin{singlespace}
-```{r citywide-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:citymap}Lost Voters on Election Day, 2017", out.extra = '', fig.pos = "H", fig.align = "center"}
-j <- readRDS("./output/city_map.RDS")
-print(j)
-```
-```{r demos-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_nhoods.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demos) Comparing Neighborhoods with Lost Voters to Overall Neighborhoods", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95\\% confidence level.}}}"
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demos} Neighborhood Demographics}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(18.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n) %>% 
-  mutate(V1 = gsub("textbackslash\\{\\}", "", V1),
-         V1 = gsub("for Lost Voters", "\\&\\&for Lost Voters", V1))
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whateverrrrr2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whateverrrrr2.tex"}
-```{r trad-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/bg_model_reg_ols.rdata")
-
-stargazer(bg_model, bg_model2,
-          header = F,
-          type = "latex",
-          notes.align = "l",
-          notes.append = T,
-          se = list(bg_model_ses, bg_model2_ses),
-          covariate.labels = c("Lost Voters", "Lost Voters $\\times$ Share Non-Hispanic Black",
-                               "Stop-and-Frisk Encounters (Per 1,000 Residents)",
-                               "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
-                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
-          dep.var.labels = c("Turnout Rate in 2017 General Election"),
-          title = "\\label{tab:trad-reg} Standard Regression",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          notes = "TO REPLACE",
-          omit = "boro",
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test.tex",
-          out.header = F)
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.7\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors (clustered by city council district) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table2.tex"}
-```{r postest-map, echo = FALSE, message = F, warning = F, fig.width = 5, fig.height = 5, fig.cap = "\\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/dec_block_map.rds")
-print(j)
-```
-```{r block-table-chunk-did, echo = F}
-t <- readRDS("./temp/match_table_block_group_did.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:did-match) Results of Block Group-Level Matching", linesep = "") %>% 
-  add_header_above(c(" " = 1, "Means: Unmatched Data" = 2, "Means: Matched Data" = 2, "Percent Improvement" = 4), align = "c") %>% 
-  kable_styling(latex_options = c("scale_down", "HOLD_position"),
-                font_size = 10)
-j
-```
-```{r tract-reg-chunkssss, echo = F, results = "hide", message = F, warning = F}
-load("./temp/match_reg_block_group_did.rdata")
-
-bg_reg <- reg_output
-bg_reg_ses <- reg_output_ses
-
-stargazer(bg_reg, reg_output2, reg_output3,
-          header = F,
-          se = list(bg_reg_ses, reg_output2_ses, reg_output3_ses),
-          notes = c("TO REPLACE"),
-          type = "latex", notes.align = "l",
-          notes.append = T,
-          covariate.labels = c("D(2017)", "D(Treat)", "D(2017) $\\times$ D(Treat)", "Constant"),
-
-          dep.var.labels = c("Turnout Rate"),
-          title = "\\label{tab:did-reg} Neighborhood Turnout in 2016 and 2017 General Elections",
-          table.placement = "H",
-          omit.stat = c("f", "ser"),
-          table.layout = "-dcm#-t-a-s-n",
-          out = "./temp/test_abc.tex",
-          omit = c("median_income", "latino", "nh_white", "some_college", "district", "sf",
-                   "median_age", "reg_rate", "share_dem", "share_non_citizen", "nh_black", "control_GEOID"),
-          out.header = F,
-          add.lines=list(c("Congressional District FEs", "X", "X", "X"),
-                         c("Controls Included", " ", "X", "X"),
-                         c("Block Group FEs", "", "", "X")))
-
-j <- fread("./temp/test_abc.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors (clustered at match level) in parentheses.}}}"
-
-j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table1111.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table1111.tex"}
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_m35.rds")))
-true <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17.rds")))
-plus_35 <- scales::comma(nrow(readRDS("./temp/matched_ids_in_17_p35.rds")))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-\end{singlespace}
-\pagebreak
-\pagenumbering{arabic}
-# Introduction {-}
-
-The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage [see, for instance, @Keyssar2009].
-At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections [@bcj_laws].
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated `r 6.1 - 1.4` million American citizens remain disenfranchised [@sentencing_2016].
-
-This paper furthers our understanding of the political implications of felony disenfranchisement by investigating its indirect turnout effects at the local level in New York City's 2017 mayoral election.
-Using administrative data from the Department of Corrections and Community Supervision and the registered voter file, I redefine "lost voters" as disenfranchised individuals with a history of voting.
-I test the effect of lost voters on neighborhood turnout after controlling for general neighborhood exposure to the criminal justice system, allowing me to tease apart the effect of disenfranchisement from the bundled indirect effects of overpolicing.
-I demonstrate that the neighborhoods left behind by lost voters turned out at substantially lower rates than other similar neighborhoods --- and that there is likely a causal relationship at play.
-
-Given the reality of economic and racial segregation, it should not be surprising that the effects of felony disenfranchisement are highly spatially concentrated.
-Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
-Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health [@Sewell2016] and are more likely to suffer from anxiety and exhibit symptoms of trauma [@Geller2014].
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted [@Clear2008], while concentrated policing has also been credited with having a "chilling effect" on neighborhoods' willingness to reach out for help to local governments [@Lerman2013].
-Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as @Gelman2007 shows, for instance, New York's "stop-and-frisk" policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is a disenfranchised population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, `r round(readRDS("./temp/share_black_incarcerated.rds"), digits = 3) * 100`% of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3% of the citizen voting age population in the state is non-Hispanic Black.^[Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1% of the citizen voting age population and 22.9% of the incarcerated population.]
-
-Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation [@White2019; @Gerber2015; @Burch2011].
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level [@Bowers2009; @King2016].
-With the exception of @Burch2013, however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Improving upon this literature is of great importance when we consider the spatial concentration of incarceration --- and, therefore, disenfranchisement.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation.
-This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences [@Kinsella2015].
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas [@Martin2003; @Martin2013].
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As @Hajnal2009 tells us: "low [and uneven] turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups" (p.
-8).
-
-# Literature on the Indirect Effects of Felony Disenfranchisement {-}
-
-A number of recent papers have attempted to estimate the indirect effects of felony disenfranchisement on turnout.
-These papers have generally looked for effects at the state level, and have come to different conclusions.
-@King2016, for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from @locked_out to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended [@Bowers2009; @Ochs2006].
-As @Bowers2009 sums up: "[I]t is not solely the direct vote of ex-felons that is denied through these laws.
-[Felony disenfranchisement] impacts the political power of communities that extends beyond felons' collateral penalty" (p.
-724).
-
-@Miles2004, however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the state level.
-This paper is more methodologically sophisticated than other research in this body of literature, and @Miles2004 finds little-to-no effect from felony disenfranchisement on statewide turnout.
-The statewide research is thus in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
-Despite its methodological improves, @Miles2004 (like the other literature looking for statewide disenfranchisement turnout effects) uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
-
-Research looking for the effect of felony disenfranchisement in statewide statistics is fundamentally flawed.
-This approach is generally taken more for methodological reasons (disenfranchisement policies vary between but not within states) or practical reasons (national data at lower geographic levels about incarceration and voting patterns is difficult to find), but not for theoretical ones.
-The effects of incarceration --- and, therefore, felony disenfranchisement --- are not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
-Voting is a social act, and social networks play an important role in predicting political participation [e.g.
-@Foladare1968; @Huckfeldt1979; @Kenny1992; @Mutz2002].
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods [@Guest1999; @Dawkins2006].
-It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics [@Gimpel2004; @Cho2006].
-This is also the level at which any effect from felony disenfranchisement should theoretically operate.
-
-The little research looking for turnout effects at the local level, however, is similarly conflicting.
-@Burch2013, for instance, demonstrates that block groups in North Carolina with higher levels of incarceration had lower turnout.
-@White2019a, on the other hand, finds that incarceration has only moderate indirect effects on voting.
-She finds "evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences" (p.
-607).
-Neither of these papers, however, account for felony disenfranchisement laws *per se*, but rather test the bundled indirect effects of proximal contact with incarceration (inclusive of disenfranchisement).
-Disenfranchised individuals who would not have voted in the absence of their incarceration are not directly affected by disenfranchisement policy; any indirect turnout effects associated with their incarceration, therefore, cannot be attributed to disenfranchisement policy.
-
-This study addresses with both problems with the existing literature.
-I deal with geographic concentration by testing the relationship between felony disenfranchisement and turnout at the Census block group level.
-Secondly, I do not look for the indirect effects of proximal contact with *any* individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual *with a demonstrated history of voting*.
-As @Gerber2003 and others indicate, past turnout is a very strong predictor of future turnout.
-Thus, by defining lost voters as individuals who previously participated, I can identify the individuals most likely to vote in the absence of felony disenfranchisement laws --- and therefore the individuals (and neighborhoods) most directly impacted by these laws.
-Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood police stop rates.
-
-# Data {-}
-
-## Criminal Justice Data {-}
-
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These records include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-## Voter File Data {-}
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from April 30^th^, 2018.
-
-The New York State Voter File is unique in its treatment of "purged" voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
-
-Further discussion about how these administrative records are linked and geocoded can be found in the Supplemental Information.
-
-# Research Design and Expectations {-}
-
-I begin with an ordinary least squares regression estimated at the block group level.
-In this regression, neighborhood turnout is the dependent variable, while the number of lost voters is the independent variable of interest.
-I look at the 2017 mayoral election in New York City, and define a "lost voter" as an individual incarcerated or on felony parole on Election Day in 2017 who cast a ballot sometime between 2007 and 2016.
-Although the 2017 mayoral election was a low turnout contest, @Hajnal2009 argues that these low turnout local elections are precisely where differential turnout rates can have the biggest consequences.
-
-A regression that looks at turnout only in one year can shed little light on the causal effect of felony disenfranchisement.
-After investigating whether neighborhoods with lost voters turned out at lower rates in 2017, I use data from the 2016 presidential election to ask whether the loss of a voter leads to depressed turnout.
-I limit the pool of block groups to block groups with no lost voters in 2016, and test whether neighborhoods that had a lost voter one year later saw turnout decrease relative to their own history in 2016, and the 2017 turnout of neighborhoods that did not lose a voter in that period.
-Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods.
-The matching procedure ensures that, prior to treatment, treated and control observations are substantially similar to one another based on observable characteristics.
-
-I expect that neighborhoods with lost voters turned out at lower rates in 2017, and that the loss of a voter to felony disenfranchisement causes neighborhood turnout to decline.
-I expect these effects to operate through three primary mechanisms:
-
-* Firstly, lost voters are likely members of families and communities where voting is valued.
-Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
-Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
-Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
-
-* Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
-The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
-
-* Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
-Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
-
-# Results {-}
-
-Figure \ref{fig:citymap} shows where lost voters lived before going to prison, with city council districts also included.
-There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election.
-
-FIGURE \ref{fig:citymap} ABOUT HERE
-
-Although `r comma(nrow(readRDS("./temp/in_2017.rds")))` individuals were formally disenfranchised as of the 2017 election, there were just `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` disenfranchised voters statewide who had participated in the past 10 years.^[Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.] Just `r round(100 * nrow(readRDS("./temp/ids_of_lost_voters.rds")) / nrow(readRDS("./temp/in_2017.rds")), digits = 1)`% of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
-
-The spatial concentration of lost voters is readily apparent.
-In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of past voters were not allowed to cast a ballot for mayor.
-
-Table \ref{tab:demos} demonstrates just how different the neighborhoods with lost voters are from the rest of the city.
-The first column in Table \ref{tab:demos} weights the chacteristics of each block group in the city by its population; the second column weights the block groups by their number of lost voters.
-The median income in neighborhoods with lost voters is less than three-quarters that of unaffected neighborhoods.
-The average lost voter came from a neighborhood that was nearly twice as Black as the city as a whole; their neighborhoods were also home to a far larger share of Democratic voters than other neighborhoods.
-Unsurprisingly, lost voters also came from neighborhoods with 60% more police stops than the citywide average.
-
-TABLE \ref{tab:demos} ABOUT HERE
-
-## Testing for Neighborhood Turnout Differences {-}
-
-I begin by asking whether turnout was lower in the 2017 mayoral election in neighborhoods with lost voters.
-Block group turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election.
-These records are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's citizen voting age population (CVAP, estimated by the Census Bureau).
-The independent variable of interest measures the number of lost voters in a neighborhood.
-
-I also control for other social and political indicators: neighborhood racial characteristics, median income, education, age, population, and share noncitizen^[The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.] come from the Census Bureau's 2017 5-year American Community Survey estimates.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the CVAP.
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;^[Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood] where city council district races were more competitive, I expect that more voters will have turned out.
-
-To completely decompose the effects of disenfranchisement from incarceration more generally, we would need to control for the number of incarcerated or paroled individuals in a neighborhood.
-If, after controlling for a neighborhood's general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
-
-Unfortunately, such data does not exist; the State of New York does not make pre-incarceration addresses available to the public.
-In its place, I use data from the NYPD on stop-and-frisk encounters with the police.
-These models include the number of times residents were stopped by the police from January -- October of 2017.
-This is intended to control for general neighborhood exposure to the carceral state.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-Model 2, therefore, interacts the number of lost voters with the share of the block group that is non-Hispanic Black.
-
-Table \ref{tab:trad-reg} presents the results of ordinary least squares regressions testing neighborhood turnout rates in 2017.
-Robust standard errors are clustered by city council district.^[Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.]
-
-TABLE \ref{tab:trad-reg} ABOUT HERE
-
-The results presented in Table \ref{tab:trad-reg} indicate that neighborhoods with lost voters turned out at lower rates in the 2017 election.
-Each missing voter in a block group reduces that neighborhood's turnout by about `r -1*round(bg_model[["coefficients"]][["lost_voters"]]*100, digits = 2)` percentage points.
-Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to `r -1*round(bg_model2[["coefficients"]][["lost_voters_black"]]*100, digits = 2)` percentage points.
-The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
-
-The block groups where these depressive effects are concentrated are not randomly distributed throughout the city.
-They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on $Lost\ Voters \times Share\ Black$ from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is $-0.019\times Lost\ Voters \times Share\ Black$.
-
-FIGURE \ref{fig:postest} ABOUT HERE
-
-## Toward a Causal Estimate {-}
-
-As the analysis above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-Moreover, by looking at turnout from just one election, the best we could hope to uncover is a correlation between disenfranchisement and turnout.
-Uncovering the causal effect requires looking across elections (or the very strong assumptions that all relevant neighborhood characteristics are accounted for).
-
-In order to estimate the causal effect of felony disenfranchisement on neighborhood-level turnout, I now test whether losing a voter between the 2016 and 2017 elections resulted in decreased turnout in 2017.
-I begin by restricting the set of neighborhoods to only neighborhoods that had no lost voters on election day in 2016.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8^th^, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then identify the neighborhoods that did have lost voters in 2017; these block groups are our "treated" block groups, while the neighborhoods with lost voters in neither 2016 nor 2017 serve as our controls.
-
-Of course, as demonstrated in Table \ref{tab:demos}, neighborhoods with lost voters are very different than neighborhoods without lost voters.
-Control and treatment neighborhoods might have different levels of turnout in 2017 even in the absence of felony disenfranchisement.
-To account for this possibility, I use a genetic matching model [@Sekhon2011] to match each treated block group to one untreated block group, using the same characteristics detailed in Table \ref{tab:trad-reg} but based on their 2016 characteristics.
-Matched ties are not broken, leaving the potential for some treated block groups to match to more than one control; the observations are weighted accordingly.
-Only treated block groups and the untreated block groups with which they match --- the controls --- are retained.
-
-Although it may seem counterintuitive to discard some potential control block groups, doing so gives us a major advantage: it allows us to conceptualize the "treatment" (the loss of a voter) as a natural experiment.
-Because the treatment and control groups look similar based on their observable 2016 characteristics, the treatment can be understood as quasi-randomly assigned.
-
-Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods.
-The matching procedure substantially improves the similarly between treated and control block groups along all characteristics.
-Table \ref{tab:did-match} mirrors Table \ref{tab:demos} above: treated neighborhoods are far less White, had many more stop-and-frisk encounters, and had much lower median incomes.
-
-TABLE \ref{tab:did-match} ABOUT HERE
-
-In Table \ref{tab:did-reg}, I present the outcome of an OLS regression on this subset of block groups.
-The variable *D(2017)* tests whether turnout for control block groups declined in 2017 relative to 2016.
-I test whether there was a difference in turnout between control and treatment neighborhoods in 2016 with *D(Treat)*.
-Of most interest is the coefficient on *D(Treat) $\times$ D(2017)*, which tests whether turnout decreased more from 2016 to 2017 in treated block groups than in the controls.
-Model 1 includes only these covariates (and fixed effects for Congressional district to account for potential differences in competitiveness in 2016), while Model 2 also includes the variables that were used for the matching procedure.
-Since each neighborhood appears in the panel twice --- once in 2016, and once in 2017 --- each observation takes the characteristics of the year in which it appears.
-This is especially important for the stop-and-frisk variable: Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
-Model 3 adds block group-level fixed effects to Model 2 in order to control for time-invariant neighborhood characteristics.
-Robust standard errors are clustered at the level of the match [@Abadie2019].
-
-TABLE \ref{tab:did-reg} ABOUT HERE
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-In each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016 who lost one in the subsequent year.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-The marginal effect of a lost voter could be lower in neighborhoods that consistently have lost voters and thus have less to "learn" about the state from the loss of any one individual to incarceration.
-Neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16% versus 21%), a lower share non-Hispanic Black (39% versus 32%), and a smaller share were registered as Democrats (76% versus 73%).^[Each of these differences are significant at the 99% confidence level.]
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
-
-# Discussion {-}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated `r comma(readRDS("./temp/lost_count_nyc.rds"))` voters from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
-However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large direct effects on turnout and representation.
-
-As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Despite methodological problems, previous literature has hinted that felony disenfranchisement impacts Black turnout at the state level; this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40% higher than the median income of block groups with lost voters.
-Similarly, just 17% of individuals in the average block group with a lost voter were White, compared with 40% of the population in other block groups.
-Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
-The effects are concentrated in neighborhoods where the most marginalized members of society live.
-These highly concentrated spillover effects are cause for major concern.
-
-Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is nonsignificant.
-This means that lost voters have little spillover effects in neighborhoods with small Black populations.
-Understanding the magnitude of this finding is important.
-In New York City, there are `r comma(readRDS("./temp/count_gt_90_black"))` block groups where the Black community makes up more than 38% of the population (the average for treated block groups), with an average of `r round(readRDS("./temp/vap_gt_90_black"))` voting age citizens.
-Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38% Black with a citizen voting age population of `r round(readRDS("./temp/vap_gt_90_black"))`, having any lost voter reduced turnout by `r round(0.38 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_gt_90_black"), digits = 1)` ballots.
-Conversely, in a block group with the same voting age population that is just 16% Black (the average for untreated block groups), the loss of any voter decreased turnout by just `r round(0.1 * -1*bg_model2[["coefficients"]][["lost_voters_black"]] * readRDS("./temp/vap_lt_10_black"), digits = 1)` ballots --- hardly more votes than the lost voter herself.
-
-Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large "interpretive effects" [@Pierson1993].
-@Lerman2013, for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-@Weaver2010 argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, @Lerman2014 tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather "as something to be actively avoided" (p.
-16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
-
-Individuals who live in neighborhoods where police activity is relatively limited may interpret the incarceration of a neighbor as a largely individual phenomenon.
-If it is understood as an isolated or individual event, voters in these neighborhoods who are not incarcerated are not likely to update their view of the state.
-They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
-In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
-It may, rather, be interpreted as another reminder of the government's unfairness.
-If a would-be voter finds herself soured on political participation because of her neighbor's incarceration, she may be less likely to cast a ballot.
-
-A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans [e.g.
-@Browning1992; @Hurwitz2005; @Henderson1997; @Wu2009].
-Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
-There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
-
-Ultimately, the analysis at hand cannot answer the causal mechanism for these depressive effects, or why the depressive effect is concentrated in Black neighborhoods.
-As @Walker2020 notes, "contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact" (p.
-123).
-Administrative data can only get us so far.
-Other research designs are necessary to investigate how neighbors understand the incarceration of their neighbors, how they react on voting day, and whether these effects are racially moderated.
-Much of the work being done by Hannah Walker and coauthors [@Walker2014; @Walker2017; @Walker2020] does point to racial differences in political responses to proximal contact with systems of incarceration.
-
-This paper uncovers surprisingly large effects.
-It is possible that, despite controlling for neighborhood exposure to the carceral state using stop-and-frisk data, the analysis cannot entirely decompose the effect of disenfranchisement from the bundled experience of seeing a neighbor go to prison.
-Future research must gather better data on local patterns of incarceration in order to better distinguish turnout patterns in neighborhoods where voters are sent to prison from neighborhoods where non-voters are incarcerated.
-Future work should similarly test whether these effects, identified in a relatively uncompetitive local election, are also present in competitive and national elections.
-
-For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods [@Wilson1990].
-More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
-Much of this work, however, has focused on either the impacts of living in marginalized communities (by measuring health and economic impacts) or has not accounted for the importance of physical space (by focusing only on proximal social, and not geographic, contact with directly impacted individuals).
-This analysis allows us to understand the implications for living in a neighborhood where individuals likely to cast a ballot are not allowed to because of a felony conviction.
-I find that neighborhoods that are home to lost voters --- and particularly neighborhoods with large Black populations --- systematically turn out for local elections at lower rates than otherwise similar neighborhoods.
-
-This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
-In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
-As @Hajnal2005 shows, racial turnout differentials can have real consequences for city politics.
-
-Moreover, we cannot conclude that depressed turnout in these neighborhoods in 2017 had no impact on their representation.
-City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
-Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals *held* and *used* their power.
-
-Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
-As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, policing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
-
-\newpage
-# Supplemental Information {-}
-
-I here present more information on how the full data set used in the body of this paper was constructed.
-
-## Geocoding {-}
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of `r round(readRDS("./temp/bad_geocode_nyc.rds") * 100, digits = 1)`% of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); `r round(readRDS("./temp/bad_geo_lost_voters.rds") * 100, digits = 1)`% of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
-
-## Matching {-}
-
-I construct the primary dataset by matching the administrative criminal supervision records with the registered voter file.
-I match using first, middle, and last names, as well as dates of birth.
-Capitalization in names in both datasets is standardized, and punctuation and spaces are removed.
-If a middle name is missing in one dataset but not in the other, I consider this a match.
-Similarly, if one dataset includes only a middle initial which matches to the first letter of the middle name the other data set, this is considered a match.
-First and last names, and dates of birth, must match exactly between the two datasets.
-
-To test for false positive matches, I employ the test developed in @Meredith2013.
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these "fake" records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5% of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-TABLE \ref{tab:change-dobs} ABOUT HERE
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6% of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "Alex" in one set of records and "Alexander" in the other).
-
-\theendnotes
-
-\newpage
-# References {-}

### 2020-04-01 markdown/part1/title.tex
uar re-submit

--- 2020-04-01 markdown/part1/UAR_submission.tex
@@
   pdftitle={Neighborhoods and Felony Disenfranchisement: The Case of New York City},
+  pdfauthor={Kevin Morris},
   hidelinks,
@@
 \usepackage{rotating}
-\usepackage{setspace}\doublespacing
 \usepackage{endnotes}
 \let\footnote=\endnote
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newenvironment{cslreferences}%
-  {\setlength{\parindent}{0pt}%
-  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
-  {\par}
 
-\title{Neighborhoods and Felony Disenfranchisement: The Case of New York City}
-\author{}
-\date{\vspace{-2.5em}April 01, 2020}
+\title{Neighborhoods and Felony Disenfranchisement: The Case of New York City\thanks{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, Peter Miller, participants at the 2020 Annual Meeting of the Southern Political Science Association, and several blind reviewers for their comments on this project.
+All errors are my responsibility.}}
+\author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
+\date{April 01, 2020}
 
@@
 I map these individuals to their pre-incarceration addresses and use multiple approaches to explore whether their home neighborhoods turned out at lower rates than other neighborhoods in the 2017 NYC mayoral election.
-I find that neighborhoods that were home to lost voters turnout at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
+I find that neighborhoods that were home to lost voters turned out at substantially lower rates than similar neighborhoods, and that Black neighborhoods are particularly impacted by the spillover effects of disenfranchisement.
 These indirect effects of the incarceration of would-be voters may have serious implications for the representation of impacted neighborhoods.
@@
 
-\pagenumbering{gobble}
-\newpage
-\pagenumbering{arabic}
-
-\begin{singlespace}
-\begin{figure}[H]
-
-{\centering \includegraphics{UAR_submission_files/figure-latex/citywide-map-1}
-
-}
-
-\caption{\label{fig:citymap}Lost Voters on Election Day, 2017}\label{fig:citywide-map}
-\end{figure}
-
-\input{"../../temp/table_whateverrrrr2.tex"}
-
-\input{"../../temp/table2.tex"}
-\begin{figure}[H]
-
-{\centering \includegraphics{UAR_submission_files/figure-latex/postest-map-1}
-
-}
-
-\caption{\label{fig:postest}Estimated Depressive Effect of Felony Disenfranchisement}\label{fig:postest-map}
-\end{figure}
-\begin{table}[H]
-
-\caption{\label{tab:block-table-chunk-did}\label{tab:did-match} Results of Block Group-Level Matching}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lllllrrrr}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Means: Unmatched Data} & \multicolumn{2}{c}{Means: Matched Data} & \multicolumn{4}{c}{Percent Improvement} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-9}
- & Treated & Control & Treated & Control & Mean Diff & eQQ Med & eQQ Mean & eQQ Max\\
-\midrule
-Stop-and-Frisk Encounters (Per 1,000 Residents) & 1.52 & 1.22 & 1.52 & 1.40 & 59.18 & 80.84 & 74.64 & 64.80\\
-\% Latino & 0.35 & 0.24 & 0.35 & 0.35 & 98.18 & 93.37 & 91.90 & 83.66\\
-\% Non-Hispanic Black & 0.33 & 0.15 & 0.33 & 0.33 & 98.29 & 94.91 & 93.59 & 86.37\\
-\% Non-Hispanic White & 0.20 & 0.41 & 0.20 & 0.20 & 97.57 & 92.56 & 89.24 & 83.47\\
-Median Income & 52,352.43 & 70,311.87 & 52,352.43 & 52,177.85 & 99.03 & 92.14 & 87.32 & 68.43\\
-\% With Some College & 0.65 & 0.71 & 0.65 & 0.65 & 96.85 & 90.55 & 86.81 & 69.31\\
-Median Age & 35.61 & 38.12 & 35.61 & 35.68 & 97.21 & 68.61 & 73.56 & 63.48\\
-Registration Rate & 0.98 & 0.93 & 0.98 & 0.96 & 49.13 & 72.80 & 64.99 & 31.69\\
-\% Democrats & 0.73 & 0.64 & 0.73 & 0.73 & 94.29 & 94.58 & 91.69 & 84.47\\
-\% Noncitizen & 0.17 & 0.17 & 0.17 & 0.18 & 90.15 & 90.13 & 83.20 & 61.09\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\input{"../../temp/table1111.tex"}
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 20,955\\
-Birthdate + 35 Days & 105\\
-Birthdate - 35 Days & 92\\
-\bottomrule
-\end{tabular}
-\end{table}
-\end{singlespace}
-\pagebreak
-\pagenumbering{arabic}
-
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-The political history of the United States has been characterized by a general, if nonlinear, trend toward universal suffrage (see, for instance, Keyssar \protect\hyperlink{ref-Keyssar2009}{2009}).
-At the time of the nation's founding, access to the ballot box was restricted to landed White men; over the following two centuries, the franchise was greatly expanded.
-Today, voting rights are considered foundational aspects of full citizenship (United Nations General Assembly Resolution 2200 (XXI)).
-Despite the United State's march toward ever-more-inclusive systems of democracy, however, one large group of American citizens is formally barred from voting.
-In most of the United States, citizens convicted of felonies are at least temporarily prohibited from casting ballots in elections (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
-Although some states such as Florida and Louisiana have gradually moved to dismantle their systems of felony disenfranchisement, an estimated 4.7 million American citizens remain disenfranchised (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).
-
-This paper furthers our understanding of the political implications of felony disenfranchisement by investigating its indirect turnout effects at the local level in New York City's 2017 mayoral election.
-Using administrative data from the Department of Corrections and Community Supervision and the registered voter file, I redefine ``lost voters'' as disenfranchised individuals with a history of voting.
-I test the effect of lost voters on neighborhood turnout after controlling for general neighborhood exposure to the criminal justice system, allowing me to tease apart the effect of disenfranchisement from the bundled indirect effects of overpolicing.
-I demonstrate that the neighborhoods left behind by lost voters turned out at substantially lower rates than other similar neighborhoods --- and that there is likely a causal relationship at play.
-
-Given the reality of economic and racial segregation, it should not be surprising that the effects of felony disenfranchisement are highly spatially concentrated.
-Data available from New York City shows that in 2017, 10 of the New York Police Department's 77 precincts were responsible for more than a quarter of all arrests for felony charges.
-Many scholars have detailed the impact of living in areas with high levels of police activity.
-Residents of such neighborhoods suffer from worse physical health (Sewell and Jefferson \protect\hyperlink{ref-Sewell2016}{2016}) and are more likely to suffer from anxiety and exhibit symptoms of trauma (Geller et al.
-\protect\hyperlink{ref-Geller2014}{2014}).
-The labor markets and social networks in neighborhoods with high levels of policing and incarceration are disrupted (Clear \protect\hyperlink{ref-Clear2008}{2008}), while concentrated policing has also been credited with having a ``chilling effect'' on neighborhoods' willingness to reach out for help to local governments (Lerman and Weaver \protect\hyperlink{ref-Lerman2013}{2013}).
-Perhaps most troubling of all, these effects are not concentrated in random neighborhoods; as Gelman, Fagan, and Kiss (\protect\hyperlink{ref-Gelman2007}{2007}) shows, for instance, New York's ``stop-and-frisk'' policy impacted Black and Latino New Yorkers at rates far higher than Whites, even after controlling for neighborhood variability and race-specific arrest rates.
-The result is a disenfranchised population that looks far different than the rest of the state: according to data from the New York State Department of Corrections and Community Supervision, 49.5\% of individuals who were incarcerated in December of 2018 were non-Hispanic Black, although the Census Bureau estimates that just 14.3\% of the citizen voting age population in the state is non-Hispanic Black.\footnote{Latinos are also over-represented among the incarcerated population, though not as dramatically: Latinos make up 14.1\% of the citizen voting age population and 22.9\% of the incarcerated population.}
-
-Felony disenfranchisement policies are part of a criminal justice system that disproportionately impacts Black Americans living in certain communities.
-The neighborhood-specific implications of felony disenfranchisement, however, remain largely unstudied.
-A number of studies have explored the effect of imprisonment and disenfranchisement on later political participation (White \protect\hyperlink{ref-White2019}{2019}\protect\hyperlink{ref-White2019}{b}; Gerber et al.
-\protect\hyperlink{ref-Gerber2015}{2015}; Burch \protect\hyperlink{ref-Burch2011}{2011}).
-Others have looked at the spillover effects of disenfranchisement on eligible Black voters at the state level (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; King and Erickson \protect\hyperlink{ref-King2016}{2016}).
-With the exception of Burch (\protect\hyperlink{ref-Burch2013}{2013}), however, little attention has been paid to the impact of felony disenfranchisement on political participation at the neighborhood level.
-
-Improving upon this literature is of great importance when we consider the spatial concentration of incarceration --- and, therefore, disenfranchisement.
-If the spillover effects of felony disenfranchisement are widely dispersed, the political consequences are likely similarly diffuse.
-If, however, the effects are both large enough to be detected at a statewide level (as the literature hints it might be) but also concentrated in a few neighborhoods, disenfranchisement laws likely severely undermine local political representation.
-This lowered participation can disadvantage neighborhoods insofar as they have unique candidate preferences (Kinsella, McTague, and Raleigh \protect\hyperlink{ref-Kinsella2015}{2015}).
-It could also lower the resources distributed to these neighborhoods, as politicians determine that too few votes can be won through investing in these areas (Martin \protect\hyperlink{ref-Martin2003}{2003}; Martin and Claibourn \protect\hyperlink{ref-Martin2013}{2013}).
-This is not merely of theoretical concern: research indicates that turnout differentials at the city level have material consequences.
-As Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) tells us: ``low {[}and uneven{]} turnout results in losses in mayoral elections, less equitable racial and ethnic representation on city councils, and spending policies that are less in line with the preferences of racial and ethnic minorities and other disadvantaged groups'' (p.~8).
-
-\hypertarget{literature-on-the-indirect-effects-of-felony-disenfranchisement}{%
-\section*{Literature on the Indirect Effects of Felony Disenfranchisement}\label{literature-on-the-indirect-effects-of-felony-disenfranchisement}}
-\addcontentsline{toc}{section}{Literature on the Indirect Effects of Felony Disenfranchisement}
-
-A number of recent papers have attempted to estimate the indirect effects of felony disenfranchisement on turnout.
-These papers have generally looked for effects at the state level, and have come to different conclusions.
-King and Erickson (\protect\hyperlink{ref-King2016}{2016}), for instance, leverages state-level variation in disenfranchisement laws to estimate the impact that felony disenfranchisement has on turnout among Black Americans.
-They use data from the 2004 Current Population Survey Voting and Registration Supplement to calculate statewide turnout rates, and include estimates of the share of Black Americans who are disenfranchised in each state from Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) to explore the impact of these policies on eligible voters.
-They conclude that disenfranchisement has large spillover effects for Black voters: where more Black residents are disenfranchised, eligible Black voters are less likely to cast a ballot.
-These findings are in line with other research that has explored whether the effects of disenfranchisement extend beyond those whose voting rights are directly suspended (Bowers and Preuhs \protect\hyperlink{ref-Bowers2009}{2009}; Ochs \protect\hyperlink{ref-Ochs2006}{2006}).
-As Bowers and Preuhs (\protect\hyperlink{ref-Bowers2009}{2009}) sums up: ``{[}I{]}t is not solely the direct vote of ex-felons that is denied through these laws.
-{[}Felony disenfranchisement{]} impacts the political power of communities that extends beyond felons' collateral penalty'' (p.~724).
-
-Miles (\protect\hyperlink{ref-Miles2004}{2004}), however, leverages a triple-differences framework to investigate whether felony disenfranchisement reduces participation at the state level.
-This paper is more methodologically sophisticated than other research in this body of literature, and Miles (\protect\hyperlink{ref-Miles2004}{2004}) finds little-to-no effect from felony disenfranchisement on statewide turnout.
-The statewide research is thus in conflict --- it is unclear that felony disenfranchisement has a big enough effect on turnout to be detected at the state level.
-Despite its methodological improves, Miles (\protect\hyperlink{ref-Miles2004}{2004}) (like the other literature looking for statewide disenfranchisement turnout effects) uses the total disenfranchised population in each state, without disentangling the disenfranchised voters who would have voted in the absence of the policy from those who would not have.
-
-Research looking for the effect of felony disenfranchisement in statewide statistics is fundamentally flawed.
-This approach is generally taken more for methodological reasons (disenfranchisement policies vary between but not within states) or practical reasons (national data at lower geographic levels about incarceration and voting patterns is difficult to find), but not for theoretical ones.
-The effects of incarceration --- and, therefore, felony disenfranchisement --- are not uniformly distributed throughout a given state; if felony disenfranchisement reduces turnout, it is likely to do so in geographically concentrated ways.
-Voting is a social act, and social networks play an important role in predicting political participation (e.g.~Foladare \protect\hyperlink{ref-Foladare1968}{1968}; Huckfeldt \protect\hyperlink{ref-Huckfeldt1979}{1979}; Kenny \protect\hyperlink{ref-Kenny1992}{1992}; Mutz \protect\hyperlink{ref-Mutz2002}{2002}).
-Literature from urban sociology has established that social networks are largely spatially bounded, and that local social ties are more important in lower-income neighborhoods (Guest and Wierzbicki \protect\hyperlink{ref-Guest1999}{1999}; Dawkins \protect\hyperlink{ref-Dawkins2006}{2006}).
-It should be no surprise, then, that neighborhoods have been shown to mobilize and demobilize voters through mechanisms above-and-beyond individual characteristics (Gimpel, Dyck, and Shaw \protect\hyperlink{ref-Gimpel2004}{2004}; Cho, Gimpel, and Dyck \protect\hyperlink{ref-Cho2006}{2006}).
-This is also the level at which any effect from felony disenfranchisement should theoretically operate.
-
-The little research looking for turnout effects at the local level, however, is similarly conflicting.
-Burch (\protect\hyperlink{ref-Burch2013}{2013}), for instance, demonstrates that block groups in North Carolina with higher levels of incarceration had lower turnout.
-White (\protect\hyperlink{ref-White2019a}{2019}\protect\hyperlink{ref-White2019a}{a}), on the other hand, finds that incarceration has only moderate indirect effects on voting.
-She finds ``evidence of a short-term demobilization effect for people who see household members convicted or jailed in the weeks before the election, but no evidence of a lasting turnout effect from these experiences'' (p.~607).
-Neither of these papers, however, account for felony disenfranchisement laws \emph{per se}, but rather test the bundled indirect effects of proximal contact with incarceration (inclusive of disenfranchisement).
-Disenfranchised individuals who would not have voted in the absence of their incarceration are not directly affected by disenfranchisement policy; any indirect turnout effects associated with their incarceration, therefore, cannot be attributed to disenfranchisement policy.
-
-This study addresses with both problems with the existing literature.
-I deal with geographic concentration by testing the relationship between felony disenfranchisement and turnout at the Census block group level.
-Secondly, I do not look for the indirect effects of proximal contact with \emph{any} individual incarcerated or paroled; rather, I test the relationship between neighborhood turnout and the disenfranchisement of an individual \emph{with a demonstrated history of voting}.
-As Gerber, Green, and Shachar (\protect\hyperlink{ref-Gerber2003}{2003}) and others indicate, past turnout is a very strong predictor of future turnout.
-Thus, by defining lost voters as individuals who previously participated, I can identify the individuals most likely to vote in the absence of felony disenfranchisement laws --- and therefore the individuals (and neighborhoods) most directly impacted by these laws.
-Although data on neighborhood incarceration rates is not available, I decompose the effect of disenfranchisement from the bundled treatments of exposure to the carceral state by controlling for neighborhood police stop rates.
-
-\hypertarget{data}{%
-\section*{Data}\label{data}}
-\addcontentsline{toc}{section}{Data}
-
-\hypertarget{criminal-justice-data}{%
-\subsection*{Criminal Justice Data}\label{criminal-justice-data}}
-\addcontentsline{toc}{subsection}{Criminal Justice Data}
-
-The primary criminal justice dataset comes from a public records request filed with the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These records include individual-level incarceration and parole records for individuals who have been incarcerated in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; sex; race; and others.
-
-The state makes records available only for individuals who have been incarcerated for felony offenses.
-It does not make information about individuals sentenced to probation or incarcerated for misdemeanors.
-Thus, while the data covers all individuals subject to felony disenfranchisement rules (only individuals incarcerated for felony offenses lose their voting rights), it limits the availability of a potentially helpful control group.
-It does not include individuals who are held in federal prisons; however, because the vast majority of incarcerated felons are held in state prisons, this is unlikely to affect the analysis.
-
-\hypertarget{voter-file-data}{%
-\subsection*{Voter File Data}\label{voter-file-data}}
-\addcontentsline{toc}{subsection}{Voter File Data}
-
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from April 30\textsuperscript{th}, 2018.
-
-The New York State Voter File is unique in its treatment of ``purged'' voters: although most states remove voters from their voter files once they are no longer eligible to vote, New York continues to include them in the file (but marks them as purged).
-I can therefore identify voters who were registered in the past but have since been purged due to a felony conviction.
-
-Further discussion about how these administrative records are linked and geocoded can be found in the Supplemental Information.
-
-\hypertarget{research-design-and-expectations}{%
-\section*{Research Design and Expectations}\label{research-design-and-expectations}}
-\addcontentsline{toc}{section}{Research Design and Expectations}
-
-I begin with an ordinary least squares regression estimated at the block group level.
-In this regression, neighborhood turnout is the dependent variable, while the number of lost voters is the independent variable of interest.
-I look at the 2017 mayoral election in New York City, and define a ``lost voter'' as an individual incarcerated or on felony parole on Election Day in 2017 who cast a ballot sometime between 2007 and 2016.
-Although the 2017 mayoral election was a low turnout contest, Hajnal (\protect\hyperlink{ref-Hajnal2009}{2009}) argues that these low turnout local elections are precisely where differential turnout rates can have the biggest consequences.
-
-A regression that looks at turnout only in one year can shed little light on the causal effect of felony disenfranchisement.
-After investigating whether neighborhoods with lost voters turned out at lower rates in 2017, I use data from the 2016 presidential election to ask whether the loss of a voter leads to depressed turnout.
-I limit the pool of block groups to block groups with no lost voters in 2016, and test whether neighborhoods that had a lost voter one year later saw turnout decrease relative to their own history in 2016, and the 2017 turnout of neighborhoods that did not lose a voter in that period.
-Treated neighborhoods --- neighborhoods that lost a voter between the 2016 and 2017 elections --- are genetically matched to untreated neighborhoods.
-The matching procedure ensures that, prior to treatment, treated and control observations are substantially similar to one another based on observable characteristics.
-
-I expect that neighborhoods with lost voters turned out at lower rates in 2017, and that the loss of a voter to felony disenfranchisement causes neighborhood turnout to decline.
-I expect these effects to operate through three primary mechanisms:
-
-\begin{itemize}
-\item
-  Firstly, lost voters are likely members of families and communities where voting is valued.
-Because voting is a social act, incarcerated individuals who have participated in the past likely have family members who also vote.
-Proximal contact with an incarcerated individual can only have a depressive effect among eligible individuals who would have voted in the absence of the contact.
-Lost voters are more likely to be in relationship with individuals who would likely vote --- and who therefore can have their turnout depressed.
-\item
-  Secondly, would-be voters can encourage their family members and neighbors to go to the polls on Election Day.
-The removal of an individual who would provide social pressure to vote can reduce the sociality of voting for others and reduce their likelihood of voting.
-\item
-  Thirdly, proximal contact with voters who are imprisoned might change attitudes toward the state more dramatically than proximal contact with nonvoting individuals who are incarcerated.
-Voters are widely considered to be upholding their civic duty through participation; the incarceration of an engaged, dutiful citizen has the potential to change an eligible voter's perception of the state and democratic participation more than the incarceration of an individual who was not actively engaged through voting.
-\end{itemize}
-
-\hypertarget{results}{%
-\section*{Results}\label{results}}
-\addcontentsline{toc}{section}{Results}
-
-Figure \ref{fig:citymap} shows where lost voters lived before going to prison, with city council districts also included.
-There were 2,518 such lost voters within New York City as of the 2017 general election.
-
-FIGURE \ref{fig:citymap} ABOUT HERE
-
-Although 73,223 individuals were formally disenfranchised as of the 2017 election, there were just 6,166 disenfranchised voters statewide who had participated in the past 10 years.\footnote{Because address prior to incarceration is not available for individuals who were not registered to vote, computing the total number of formally disenfranchised individuals in New York City is not possible.} Just 8.4\% of formally disenfranchised individuals, therefore, were individuals with a demonstrated history of participating.
-
-The spatial concentration of lost voters is readily apparent.
-In some communities, such as Greenwich Village and Brooklyn Heights, hardly any voters were disqualified from participating in the 2017 elections.
-In other communities, such as Harlem and Central Brooklyn, large numbers of past voters were not allowed to cast a ballot for mayor.
-
-Table \ref{tab:demos} demonstrates just how different the neighborhoods with lost voters are from the rest of the city.
-The first column in Table \ref{tab:demos} weights the chacteristics of each block group in the city by its population; the second column weights the block groups by their number of lost voters.
-The median income in neighborhoods with lost voters is less than three-quarters that of unaffected neighborhoods.
-The average lost voter came from a neighborhood that was nearly twice as Black as the city as a whole; their neighborhoods were also home to a far larger share of Democratic voters than other neighborhoods.
-Unsurprisingly, lost voters also came from neighborhoods with 60\% more police stops than the citywide average.
-
-TABLE \ref{tab:demos} ABOUT HERE
-
-\hypertarget{testing-for-neighborhood-turnout-differences}{%
-\subsection*{Testing for Neighborhood Turnout Differences}\label{testing-for-neighborhood-turnout-differences}}
-\addcontentsline{toc}{subsection}{Testing for Neighborhood Turnout Differences}
-
-I begin by asking whether turnout was lower in the 2017 mayoral election in neighborhoods with lost voters.
-Block group turnout rates are calculated using the geocoded voter file.
-Each voter's record indicates whether the voter participated in the 2017 general election.
-These records are then aggregated to estimate the number of ballots cast in each neighborhood.
-The number of ballots cast is divided by the neighborhood's citizen voting age population (CVAP, estimated by the Census Bureau).
-The independent variable of interest measures the number of lost voters in a neighborhood.
-
-I also control for other social and political indicators: neighborhood racial characteristics, median income, education, age, population, and share noncitizen\footnote{The Census Bureau does not make noncitizen estimates available at the block group level.
-As such, block groups are assigned their census tract's share noncitizen for matching purposes.} come from the Census Bureau's 2017 5-year American Community Survey estimates.
-Party affiliation rates come from the geocoded voter file.
-Registration rate is calculated by dividing the number of registered voters (calculated using the voter file) by the CVAP.
-I include voteshare won by the winning city council representative in 2017 as a proxy for the competitiveness of the local race;\footnote{Where neighborhoods cross council district lines, this measure is the mean competitiveness faced by each voter in the neighborhood} where city council district races were more competitive, I expect that more voters will have turned out.
-
-To completely decompose the effects of disenfranchisement from incarceration more generally, we would need to control for the number of incarcerated or paroled individuals in a neighborhood.
-If, after controlling for a neighborhood's general exposure to incarceration, there was a negative turnout effect from the loss of a voter, that loss could be attributed to disenfranchisement.
-
-Unfortunately, such data does not exist; the State of New York does not make pre-incarceration addresses available to the public.
-In its place, I use data from the NYPD on stop-and-frisk encounters with the police.
-These models include the number of times residents were stopped by the police from January -- October of 2017.
-This is intended to control for general neighborhood exposure to the carceral state.
-
-Much of the literature has discussed whether felony disenfranchisement is particularly demobilizing for eligible Black voters.
-Model 2, therefore, interacts the number of lost voters with the share of the block group that is non-Hispanic Black.
-
-Table \ref{tab:trad-reg} presents the results of ordinary least squares regressions testing neighborhood turnout rates in 2017.
-Robust standard errors are clustered by city council district.\footnote{Where neighborhoods cross city council district lines, they are assigned the district in which most of their voters live for clustering purposes.}
-
-TABLE \ref{tab:trad-reg} ABOUT HERE
-
-The results presented in Table \ref{tab:trad-reg} indicate that neighborhoods with lost voters turned out at lower rates in the 2017 election.
-Each missing voter in a block group reduces that neighborhood's turnout by about 0.8 percentage points.
-Model 2, however, makes clear that this effect is concentrated in Black neighborhoods.
-In neighborhoods where most residents are Black, each lost voter is associated with a turnout decrease of up to 1.93 percentage points.
-The neighborhoods most affected by felony disenfranchisement are neighborhoods where incarceration patterns overlap with Black communities.
-
-The block groups where these depressive effects are concentrated are not randomly distributed throughout the city.
-They are highly spatially concentrated in Central Brooklyn, Eastern Queens, and Harlem.
-Figure \ref{fig:postest} applies the coefficient on \(Lost\ Voters \times Share\ Black\) from Model 2 in Table \ref{tab:trad-reg} to the city's block groups.
-The estimated depressive effect is \(-0.019\times Lost\ Voters \times Share\ Black\).
-
-FIGURE \ref{fig:postest} ABOUT HERE
-
-\hypertarget{toward-a-causal-estimate}{%
-\subsection*{Toward a Causal Estimate}\label{toward-a-causal-estimate}}
-\addcontentsline{toc}{subsection}{Toward a Causal Estimate}
-
-As the analysis above make clear, neighborhoods with lost voters saw substantially lower turnout in the 2017 mayoral election than neighborhoods without lost voters.
-The possibility remains, however, that treated neighborhoods systematically differ from untreated neighborhoods in ways that cannot be accounted for using the available data.
-Moreover, by looking at turnout from just one election, the best we could hope to uncover is a correlation between disenfranchisement and turnout.
-Uncovering the causal effect requires looking across elections (or the very strong assumptions that all relevant neighborhood characteristics are accounted for).
-
-In order to estimate the causal effect of felony disenfranchisement on neighborhood-level turnout, I now test whether losing a voter between the 2016 and 2017 elections resulted in decreased turnout in 2017.
-I begin by restricting the set of neighborhoods to only neighborhoods that had no lost voters on election day in 2016.
-This process is the inverse of the process described above: I identify all individuals who were in prison or on parole on November 8\textsuperscript{th}, 2016, who had cast a ballot between 2006 and 2015.
-These individuals' home neighborhoods are removed from the subsequent analysis.
-I then identify the neighborhoods that did have lost voters in 2017; these block groups are our ``treated'' block groups, while the neighborhoods with lost voters in neither 2016 nor 2017 serve as our controls.
-
-Of course, as demonstrated in Table \ref{tab:demos}, neighborhoods with lost voters are very different than neighborhoods without lost voters.
-Control and treatment neighborhoods might have different levels of turnout in 2017 even in the absence of felony disenfranchisement.
-To account for this possibility, I use a genetic matching model (Sekhon \protect\hyperlink{ref-Sekhon2011}{2011}) to match each treated block group to one untreated block group, using the same characteristics detailed in Table \ref{tab:trad-reg} but based on their 2016 characteristics.
-Matched ties are not broken, leaving the potential for some treated block groups to match to more than one control; the observations are weighted accordingly.
-Only treated block groups and the untreated block groups with which they match --- the controls --- are retained.
-
-Although it may seem counterintuitive to discard some potential control block groups, doing so gives us a major advantage: it allows us to conceptualize the ``treatment'' (the loss of a voter) as a natural experiment.
-Because the treatment and control groups look similar based on their observable 2016 characteristics, the treatment can be understood as quasi-randomly assigned.
-
-Table \ref{tab:did-match} presents the results of matching on these subsets of neighborhoods.
-The matching procedure substantially improves the similarly between treated and control block groups along all characteristics.
-Table \ref{tab:did-match} mirrors Table \ref{tab:demos} above: treated neighborhoods are far less White, had many more stop-and-frisk encounters, and had much lower median incomes.
-
-TABLE \ref{tab:did-match} ABOUT HERE
-
-In Table \ref{tab:did-reg}, I present the outcome of an OLS regression on this subset of block groups.
-The variable \emph{D(2017)} tests whether turnout for control block groups declined in 2017 relative to 2016.
-I test whether there was a difference in turnout between control and treatment neighborhoods in 2016 with \emph{D(Treat)}.
-Of most interest is the coefficient on \emph{D(Treat) \(\times\) D(2017)}, which tests whether turnout decreased more from 2016 to 2017 in treated block groups than in the controls.
-Model 1 includes only these covariates (and fixed effects for Congressional district to account for potential differences in competitiveness in 2016), while Model 2 also includes the variables that were used for the matching procedure.
-Since each neighborhood appears in the panel twice --- once in 2016, and once in 2017 --- each observation takes the characteristics of the year in which it appears.
-This is especially important for the stop-and-frisk variable: Model 2 specifically controls for the possibility that police activity increased more in treated neighborhoods in 2017.
-Model 3 adds block group-level fixed effects to Model 2 in order to control for time-invariant neighborhood characteristics.
-Robust standard errors are clustered at the level of the match (Abadie and Spiess \protect\hyperlink{ref-Abadie2019}{2019}).
-
-TABLE \ref{tab:did-reg} ABOUT HERE
-
-Table \ref{tab:did-reg} makes clear that the removal of lost voters likely has a causal effect on neighborhoods' turnout rates.
-In each specification, turnout decreased from 2016 to 2017 in treated neighborhoods by at least 2.6 percentage points more than in untreated neighborhoods.
-It is worth noting that this is the treatment effect on the treated --- which is to say, it is estimated using only the 319 neighborhoods with no lost voters in 2016 who lost one in the subsequent year.
-These neighborhoods may differ from neighborhoods that had lost voters in both 2016 and 2017.
-The marginal effect of a lost voter could be lower in neighborhoods that consistently have lost voters and thus have less to ``learn'' about the state from the loss of any one individual to incarceration.
-Neighborhoods that consistently have lost voters are more likely to be the places most impacted by felony disenfranchisement.
-Among block groups that had a lost voter in 2017, those who had no lost voters in 2016 were higher-income (\$51k versus \$56k), had higher shares non-Hispanic White (16\% versus 21\%), a lower share non-Hispanic Black (39\% versus 32\%), and a smaller share were registered as Democrats (76\% versus 73\%).\footnote{Each of these differences are significant at the 99\% confidence level.}
-
-Of particular note is the lower share of the population in these neighborhoods that is Black.
-The analyses above indicate that neighborhoods with more Black residents respond the most strongly to felony disenfranchisement.
-The set of neighborhoods in this section, therefore, are likely among the least likely to see this relationship.
-Nevertheless, that this relationship can be detected at the block group level lends credence to the causal relationship between lost voters and depressed turnout.
-
-\hypertarget{discussion}{%
-\section*{Discussion}\label{discussion}}
-\addcontentsline{toc}{section}{Discussion}
-
-During the 2017 mayoral election, felony disenfranchisement laws were responsible for removing an estimated 2,518 voters from New York City neighborhoods.
-The spatial concentration of these lost voters is striking, as demonstrated in Figure \ref{fig:citywide-map}, and the systematic removal of voters from these neighborhoods is troubling.
-However, more than one million votes were cast in the general election of 2017.
-Although felony disenfranchisement rules have implications for the individuals targeted by them, the removal of such a small number of voters --- concentrated though they are --- is unlikely to have large direct effects on turnout and representation.
-
-As this analysis makes clear, however, felony disenfranchisement reaches beyond the individuals who are incarcerated.
-Despite methodological problems, previous literature has hinted that felony disenfranchisement impacts Black turnout at the state level; this analysis demonstrates that these demobilizing effects intersect with geographical space to systematically depress the vote in neighborhoods where voters are being sent to prison.
-As discussed above, these neighborhoods have far lower incomes than the rest of the city: the median income of block groups without any lost voters is more than 40\% higher than the median income of block groups with lost voters.
-Similarly, just 17\% of individuals in the average block group with a lost voter were White, compared with 40\% of the population in other block groups.
-Felony disenfranchisement has spillover effects on neighborhood turnout, and these neighborhoods systematically differ from the rest of the city's population.
-The effects are concentrated in neighborhoods where the most marginalized members of society live.
-These highly concentrated spillover effects are cause for major concern.
-
-Of equal concern is the apparent concentration of these effects in Black neighborhoods.
-Table \ref{tab:trad-reg} indicates that once the lost voter indicator is interacted with the share of a neighborhood that is non-Hispanic Black, the lost voter indicator is nonsignificant.
-This means that lost voters have little spillover effects in neighborhoods with small Black populations.
-Understanding the magnitude of this finding is important.
-In New York City, there are 1,386 block groups where the Black community makes up more than 38\% of the population (the average for treated block groups), with an average of 912 voting age citizens.
-Model 2 of Table \ref{tab:trad-reg} indicates that, in a block group that is 38\% Black with a citizen voting age population of 912, having any lost voter reduced turnout by 6.7 ballots.
-Conversely, in a block group with the same voting age population that is just 16\% Black (the average for untreated block groups), the loss of any voter decreased turnout by just 1.7 ballots --- hardly more votes than the lost voter herself.
-
-Why do lost voters in Black neighborhoods have such large spillover effects, when lost voters in predominantly non-Black neighborhoods do not?
-Much of the previous literature in this space establishes that individuals who have negative interactions with the government are less likely to choose to interact with the state in the future --- that these interactions have large ``interpretive effects'' (Pierson \protect\hyperlink{ref-Pierson1993}{1993}).
-Lerman and Weaver (\protect\hyperlink{ref-Lerman2013}{2013}), for instance, shows that neighborhoods where there are many police stops that involve searches or use of force use 311 services less frequently.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that interactions with the criminal justice system changes how individuals understand both their identities as citizens and the nature of governmental structures.
-Similarly, Lerman and Weaver (\protect\hyperlink{ref-Lerman2014}{2014}) tells us that those who have had contact with the criminal justice system consider political participation not just unfruitful but rather ``as something to be actively avoided'' (p.~16).
-It is perhaps unsurprising that the negative spillover effects are largest in neighborhoods where criminal justice involvement is most keenly (and unfairly) felt; namely, in plurality and majority Black neighborhoods.
-
-Individuals who live in neighborhoods where police activity is relatively limited may interpret the incarceration of a neighbor as a largely individual phenomenon.
-If it is understood as an isolated or individual event, voters in these neighborhoods who are not incarcerated are not likely to update their view of the state.
-They may draw no connections between their neighbor's imprisonment and their own efficacy as a voter.
-In the neighborhoods where policing is most prevalent --- often, lower-income Black communities --- the incarceration of a neighbor might not be interpreted so individualistically.
-It may, rather, be interpreted as another reminder of the government's unfairness.
-If a would-be voter finds herself soured on political participation because of her neighbor's incarceration, she may be less likely to cast a ballot.
-
-A large body of research indicates that, even after controlling for various sociodemographic characteristics and interactions with the police, Black Americans have far more negative views of the criminal justice system than White Americans (e.g.~Browning and Cao \protect\hyperlink{ref-Browning1992}{1992}; Hurwitz and Peffley \protect\hyperlink{ref-Hurwitz2005}{2005}; Henderson et al.
-\protect\hyperlink{ref-Henderson1997}{1997}; Wu, Sun, and Triplett \protect\hyperlink{ref-Wu2009}{2009}).
-Experience with the criminal justice system is less likely to be viewed in isolation (and therefore more likely to affect voting patterns) in Black neighborhoods.
-There is therefore strong reason to suspect that the concentrated depressive effects of felony disenfranchisement in Black neighborhoods arise from these distinct interpretations of the act of incarceration by the state.
-
-Ultimately, the analysis at hand cannot answer the causal mechanism for these depressive effects, or why the depressive effect is concentrated in Black neighborhoods.
-As Walker (\protect\hyperlink{ref-Walker2020}{2020}) notes, ``contextual measures of proximity do not accurately assess the relational aspect central to the concept of proximal contact'' (p.~123).
-Administrative data can only get us so far.
-Other research designs are necessary to investigate how neighbors understand the incarceration of their neighbors, how they react on voting day, and whether these effects are racially moderated.
-Much of the work being done by Hannah Walker and coauthors (Walker \protect\hyperlink{ref-Walker2014}{2014}, \protect\hyperlink{ref-Walker2020}{2020}; Walker and García-Castañon \protect\hyperlink{ref-Walker2017}{2017}) does point to racial differences in political responses to proximal contact with systems of incarceration.
-
-This paper uncovers surprisingly large effects.
-It is possible that, despite controlling for neighborhood exposure to the carceral state using stop-and-frisk data, the analysis cannot entirely decompose the effect of disenfranchisement from the bundled experience of seeing a neighbor go to prison.
-Future research must gather better data on local patterns of incarceration in order to better distinguish turnout patterns in neighborhoods where voters are sent to prison from neighborhoods where non-voters are incarcerated.
-Future work should similarly test whether these effects, identified in a relatively uncompetitive local election, are also present in competitive and national elections.
-
-For decades, scholars have detailed the problems associated with poor and segregated urban neighborhoods (Wilson \protect\hyperlink{ref-Wilson1990}{1990}).
-More recent work has begun to interrogate the ways in which the increasing reach of the carceral state shapes the economic and political behavior of individuals caught up in the criminal justice system and their community members.
-Much of this work, however, has focused on either the impacts of living in marginalized communities (by measuring health and economic impacts) or has not accounted for the importance of physical space (by focusing only on proximal social, and not geographic, contact with directly impacted individuals).
-This analysis allows us to understand the implications for living in a neighborhood where individuals likely to cast a ballot are not allowed to because of a felony conviction.
-I find that neighborhoods that are home to lost voters --- and particularly neighborhoods with large Black populations --- systematically turn out for local elections at lower rates than otherwise similar neighborhoods.
-
-This has major ramifications for how we understand the political positioning of the minority neighborhoods most impacted by overpolicing and incarceration.
-In the case of the 2017 election, there were likely no electoral consequences: Bill de Blasio won reelection handily, and neighborhoods with lost voters overwhelmingly supported his candidacy.
-A lack of electoral consequences in 2017, however, should not be interpreted to mean that the disparate and concentrated depressive effects of felony disenfranchisement never have implications for who is elected to local office.
-As Hajnal and Trounstine (\protect\hyperlink{ref-Hajnal2005}{2005}) shows, racial turnout differentials can have real consequences for city politics.
-
-Moreover, we cannot conclude that depressed turnout in these neighborhoods in 2017 had no impact on their representation.
-City council members representing these neighborhoods, for instance, may determine that pushing for policies popular with these constituents (such as stricter police oversight) will not garner enough votes to make such a fight worthwhile.
-Although spillover effects from felony disenfranchisement may not have changed who won power in 2017, it very possibly altered how those individuals \emph{held} and \emph{used} their power.
-
-Felony disenfranchisement laws, originally adopted during Jim Crow, have taken on a new life in the era of mass incarceration.
-Abundant research has demonstrated the pernicious ways in which overincarceration dogs the lives of poor and non-White Americans.
-As this study shows, however, the neighborhoods bearing the brunt of felony disenfranchisement are also seeing their political power diminished through reduced turnout.
-The interlocking nature of racioeconomic segregation, policing patterns, and felony disenfranchisement all combine to undermine the political power of marginalized communities in New York City.
-
-\newpage
-
-\hypertarget{supplemental-information}{%
-\section*{Supplemental Information}\label{supplemental-information}}
-\addcontentsline{toc}{section}{Supplemental Information}
-
-I here present more information on how the full data set used in the body of this paper was constructed.
-
-\hypertarget{geocoding}{%
-\subsection*{Geocoding}\label{geocoding}}
-\addcontentsline{toc}{subsection}{Geocoding}
-
-Voters' home addresses were converted to latitudes and longitudes using a geocoder provided by SmartyStreets.
-I then used the statistical software R to map these latitudes and longitudes to census block groups, census tracts, and city council districts using shapefiles publicly available from the Census Bureau and the City of New York.
-This geocoder is not perfect: among individuals registered to vote in New York City, the geocoder failed to determine the latitude and longitude of the addresses of 1\% of registered voters.
-The geocoder was slightly less successful when it came to lost voters (defined and discussed below); 1.6\% of these individuals were not geocoded.
-Voters who were not successfully geocoded are dropped from the dataset; however, because so few observations went uncoded, it is unlikely to affect the analysis.
-
-\hypertarget{matching}{%
-\subsection*{Matching}\label{matching}}
-\addcontentsline{toc}{subsection}{Matching}
-
-I construct the primary dataset by matching the administrative criminal supervision records with the registered voter file.
-I match using first, middle, and last names, as well as dates of birth.
-Capitalization in names in both datasets is standardized, and punctuation and spaces are removed.
-If a middle name is missing in one dataset but not in the other, I consider this a match.
-Similarly, if one dataset includes only a middle initial which matches to the first letter of the middle name the other data set, this is considered a match.
-First and last names, and dates of birth, must match exactly between the two datasets.
-
-To test for false positive matches, I employ the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}).
-I slightly alter the date of birth reported in the NYSDOCCS dataset to create false records.
-Comparing the number of matches between these ``fake'' records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as matches using a set of fake records created by adding or subtracting 35 days from an individual's birthdate.
-This analysis indicates that false positives account for between 0.4 and 0.5\% of all matches, a share that is likely too small to have any material impact on the overall analysis.
-The numbers in Table \ref{tab:change-dobs} are derived by matching (and modifying) all individuals who were incarcerated or on parole on Election Day in 2017 with the registered voter file from April of 2018.
-
-TABLE \ref{tab:change-dobs} ABOUT HERE
-
-Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
-For two reasons, however, false positives are unlikely to pose major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6\% of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``Alex'' in one set of records and ``Alexander'' in the other).
-
-\theendnotes
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{cslreferences}
-\leavevmode\hypertarget{ref-Abadie2019}{}%
-Abadie, Alberto, and Jann Spiess. 2019. ``Robust Post-Matching Inference.'' \emph{Working Paper.}
-
-\leavevmode\hypertarget{ref-Bowers2009}{}%
-Bowers, Melanie, and Robert R.
-Preuhs.
-2009.
-``Collateral Consequences of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement Laws on the Political Participation of Nonfelons*.'' \emph{Social Science Quarterly} 90 (3): 722--43.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
-
-\leavevmode\hypertarget{ref-Browning1992}{}%
-Browning, Sandra Lee, and Liqun Cao. 1992. ``The Impact of Race on Criminal Justice Ideology.'' \emph{Justice Quarterly} 9 (4): 685--701. \url{https://doi.org/10.1080/07418829200091611}.
-
-\leavevmode\hypertarget{ref-Burch2011}{}%
-Burch, Traci.
-2011.
-``Turnout and Party Registration Among Criminal Offenders in the 2008 General Election.'' \emph{Law \& Society Review} 45 (3): 699--730.
-\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
-
-\leavevmode\hypertarget{ref-Burch2013}{}%
-Burch, Traci R.
-2013.
-``Effects of Imprisonment and Community Supervision on Neighborhood Political Participation in North Carolina:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
-\url{https://doi.org/10.1177/0002716213503093}.
-
-\leavevmode\hypertarget{ref-Cho2006}{}%
-Cho, Wendy K.
-Tam, James G.
-Gimpel, and Joshua J.
-Dyck.
-2006.
-``Residential Concentration, Political Socialization, and Voter Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
-\url{https://doi.org/10.1111/j.1468-2508.2006.00377.x}.
-
-\leavevmode\hypertarget{ref-Clear2008}{}%
-Clear, Todd R. 2008. ``The Effects of High Imprisonment Rates on Communities.'' \emph{Crime and Justice} 37 (1): 97--132. \url{https://doi.org/10.1086/522360}.
-
-\leavevmode\hypertarget{ref-Dawkins2006}{}%
-Dawkins, Casey J. 2006. ``Are Social Networks the Ties That Bind Families to Neighborhoods?'' \emph{Housing Studies} 21 (6): 867--81. \url{https://doi.org/10.1080/02673030600917776}.
-
-\leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting Behavior.'' \emph{Political Science Quarterly} 83 (4): 516--29. \url{https://doi.org/10.2307/2146812}.
-
-\leavevmode\hypertarget{ref-Geller2014}{}%
-Geller, Amanda, Jeffrey Fagan, Tom Tyler, and Bruce G.
-Link.
-2014.
-``Aggressive Policing and the Mental Health of Young Urban Men.'' \emph{American Journal of Public Health} 104 (12): 2321--7.
-\url{https://doi.org/10.2105/AJPH.2014.302046}.
-
-\leavevmode\hypertarget{ref-Gelman2007}{}%
-Gelman, Andrew, Jeffrey Fagan, and Alex Kiss.
-2007.
-``An Analysis of the New York City Police Department's `Stop-and-Frisk' Policy in the Context of Claims of Racial Bias.'' \emph{Journal of the American Statistical Association} 102 (479): 813--23.
-\url{https://doi.org/10.1198/016214506000001040}.
-
-\leavevmode\hypertarget{ref-Gerber2003}{}%
-Gerber, Alan S., Donald P.
-Green, and Ron Shachar.
-2003.
-``Voting May Be Habit-Forming: Evidence from a Randomized Field Experiment.'' \emph{American Journal of Political Science} 47 (3): 540--50.
-\url{https://doi.org/10.1111/1540-5907.00038}.
-
-\leavevmode\hypertarget{ref-Gerber2015}{}%
-Gerber, Alan S., Gregory A.
-Huber, Marc Meredith, Daniel R.
-Biggers, and David J.
-Hendry.
-2015.
-``Can Incarcerated Felons Be (Re)Integrated into the Political System?
-Results from a Field Experiment.'' \emph{American Journal of Political Science} 59 (4): 912--26.
-\url{https://doi.org/10.1111/ajps.12166}.
-
-\leavevmode\hypertarget{ref-Gimpel2004}{}%
-Gimpel, James G., Joshua J.
-Dyck, and Daron R.
-Shaw.
-2004.
-``Registrants, Voters, and Turnout Variability Across Neighborhoods.'' \emph{Political Behavior} 26 (4): 343--75.
-\url{https://doi.org/10.1007/s11109-004-0900-4}.
-
-\leavevmode\hypertarget{ref-Guest1999}{}%
-Guest, Avery M., and Susan K.
-Wierzbicki.
-1999.
-``Social Ties at the Neighborhood Level: Two Decades of GSS Evidence.'' \emph{Urban Affairs Review}, September.
-\url{https://doi.org/10.1177/10780879922184301}.
-
-\leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy: Race, Turnout, and Representation in City Politics}. Cambridge ; New York: Cambridge University Press.
-
-\leavevmode\hypertarget{ref-Hajnal2005}{}%
-Hajnal, Zoltan, and Jessica Trounstine.
-2005.
-``Where Turnout Matters: The Consequences of Uneven Turnout in City Politics.'' \emph{The Journal of Politics} 67 (2): 515--35.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00327.x}.
-
-\leavevmode\hypertarget{ref-Henderson1997}{}%
-Henderson, Martha L., Francis T.
-Cullen, Liqun Cao, Sandra Lee Browning, and Renee Kopache.
-1997.
-``The Impact of Race on Perceptions of Criminal Injustice.'' \emph{Journal of Criminal Justice} 25 (6): 447--62.
-\url{https://doi.org/10.1016/S0047-2352(97)00032-9}.
-
-\leavevmode\hypertarget{ref-Huckfeldt1979}{}%
-Huckfeldt, R. Robert. 1979. ``Political Participation and the Neighborhood Social Context.'' \emph{American Journal of Political Science} 23 (3): 579--92. \url{https://doi.org/10.2307/2111030}.
-
-\leavevmode\hypertarget{ref-Hurwitz2005}{}%
-Hurwitz, Jon, and Mark Peffley.
-2005.
-``Explaining the Great Racial Divide: Perceptions of Fairness in the U.S.
-Criminal Justice System.'' \emph{The Journal of Politics} 67 (3): 762--83.
-\url{https://doi.org/10.1111/j.1468-2508.2005.00338.x}.
-
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Justice, Brennan Center for.
-2019.
-``Criminal Disenfranchisement Laws Across the United States.'' \emph{Brennan Center for Justice}, May.
-\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
-
-\leavevmode\hypertarget{ref-Kenny1992}{}%
-Kenny, Christopher B. 1992. ``Political Participation and Effects from the Social Environment.'' \emph{American Journal of Political Science} 36 (1): 259--67. \url{https://doi.org/10.2307/2111432}.
-
-\leavevmode\hypertarget{ref-Keyssar2009}{}%
-Keyssar, Alexander. 2009. \emph{The Right to Vote: The Contested History of Democracy in the United States}. Revised edition. New York: Basic Books.
-
-\leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson.
-2016.
-``Disenfranchising the Enfranchised: Exploring the Relationship Between Felony Disenfranchisement and African American Voter Turnout.'' \emph{Journal of Black Studies}, July.
-\url{https://doi.org/10.1177/0021934716659195}.
-
-\leavevmode\hypertarget{ref-Kinsella2015}{}%
-Kinsella, Chad, Colleen McTague, and Kevin N.
-Raleigh.
-2015.
-``Unmasking Geographic Polarization and Clustering: A Micro-Scalar Analysis of Partisan Voting Behavior.'' \emph{Applied Geography} 62 (August): 404--19.
-\url{https://doi.org/10.1016/j.apgeog.2015.04.022}.
-
-\leavevmode\hypertarget{ref-Lerman2013}{}%
-Lerman, Amy E., and Vesla Weaver.
-2013.
-``Staying Out of Sight?
-Concentrated Policing and Local Political Action:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
-\url{https://doi.org/10.1177/0002716213503085}.
-
-\leavevmode\hypertarget{ref-Lerman2014}{}%
-Lerman, Amy E., and Vesla M.
-Weaver.
-2014.
-\emph{Arresting Citizenship: The Democratic Consequences of American Crime Control}.
-Chicago Studies in American Politics.
-Chicago ; London: The University of Chicago Press.
-
-\leavevmode\hypertarget{ref-locked_out}{}%
-Manza, Jeff, and Christopher Uggen. 2008. \emph{Locked Out: Felon Disenfranchisement and American Democracy}. Studies in Crime and Public Policy. New York: Oxford University Press.
-
-\leavevmode\hypertarget{ref-Martin2003}{}%
-Martin, Paul S.
-2003.
-``Voting's Rewards: Voter Turnout, Attentive Publics, and Congressional Allocation of Federal Money.'' \emph{American Journal of Political Science} 47 (1): 110--27.
-\url{https://doi.org/10.1111/1540-5907.00008}.
-
-\leavevmode\hypertarget{ref-Martin2013}{}%
-Martin, Paul S., and Michele P.
-Claibourn.
-2013.
-``Citizen Participation and Congressional Responsiveness: New Evidence That Participation Matters.'' \emph{Legislative Studies Quarterly} 38 (1): 59--81.
-\url{https://doi.org/10.1111/lsq.12003}.
-
-\leavevmode\hypertarget{ref-Meredith2013}{}%
-Meredith, Marc, and Michael Morse.
-2013.
-``Do Voting Rights Notification Laws Increase Ex-Felon Turnout?:'' \emph{The ANNALS of the American Academy of Political and Social Science}, November.
-\url{https://doi.org/10.1177/0002716213502931}.
-
-\leavevmode\hypertarget{ref-Miles2004}{}%
-Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.'' \emph{The Journal of Legal Studies} 33 (1): 85--129. \url{https://doi.org/10.1086/381290}.
-
-\leavevmode\hypertarget{ref-Mutz2002}{}%
-Mutz, Diana C. 2002. ``The Consequences of Cross-Cutting Networks for Political Participation.'' \emph{American Journal of Political Science} 46 (4): 838--55. \url{https://doi.org/10.2307/3088437}.
-
-\leavevmode\hypertarget{ref-Ochs2006}{}%
-Ochs, Holona Leanne.
-2006.
-``\,`Colorblind' Policy in Black and White: Racial Consequences of Disenfranchisement Policy.'' \emph{Policy Studies Journal} 34 (1): 81--93.
-\url{https://doi.org/10.1111/j.1541-0072.2006.00146.x}.
-
-\leavevmode\hypertarget{ref-Pierson1993}{}%
-Pierson, Paul. 1993. ``When Effect Becomes Cause: Policy Feedback and Political Change.'' \emph{World Politics} 45 (4): 595--628. \url{https://doi.org/10.2307/2950710}.
-
-\leavevmode\hypertarget{ref-Sekhon2011}{}%
-Sekhon, Jasjeet S.
-2011.
-``Multivariate and Propensity Score Matching Software with Automated Balance Optimization: The Matching Package for R.'' \emph{Journal of Statistical Software} 42 (1): 1--52.
-\url{https://doi.org/10.18637/jss.v042.i07}.
-
-\leavevmode\hypertarget{ref-Sewell2016}{}%
-Sewell, Abigail A., and Kevin A.
-Jefferson.
-2016.
-``Collateral Damage: The Health Effects of Invasive Police Encounters in New York City.'' \emph{Journal of Urban Health} 93 (1): 42--67.
-\url{https://doi.org/10.1007/s11524-015-0016-7}.
-
-\leavevmode\hypertarget{ref-sentencing_2016}{}%
-Uggen, Christopher, Ryan Larson, and Sarah Shannon.
-2016.
-``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
-The Sentencing Project.
-\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
-
-\leavevmode\hypertarget{ref-Walker2014}{}%
-Walker, Hannah L.
-2014.
-``Extending the Effects of the Carceral State: Proximal Contact, Political Participation, and Race.'' \emph{Political Research Quarterly}, July.
-\url{https://doi.org/10.1177/1065912914542522}.
-
-\leavevmode\hypertarget{ref-Walker2020}{}%
----------. 2020. ``Targeted: The Mobilizing Effect of Perceptions of Unfair Policing Practices.'' \emph{The Journal of Politics} 82 (1): 119--34. \url{https://doi.org/10.1086/705684}.
-
-\leavevmode\hypertarget{ref-Walker2017}{}%
-Walker, Hannah L., and Marcela García-Castañon.
-2017.
-``For Love and Justice: The Mobilizing of Race, Gender, and Criminal Justice Contact.'' \emph{Politics \& Gender} 13 (4): 541--68.
-\url{https://doi.org/10.1017/S1743923X17000198}.
-
-\leavevmode\hypertarget{ref-Weaver2010}{}%
-Weaver, Vesla M., and Amy E. Lerman. 2010. ``Political Consequences of the Carceral State.'' \emph{American Political Science Review} 104 (4): 817--33. \url{https://doi.org/10.1017/S0003055410000456}.
-
-\leavevmode\hypertarget{ref-White2019a}{}%
-White, Ariel.
-2019a.
-``Family Matters?
-Voting Behavior in Households with Criminal Justice Contact.'' \emph{American Political Science Review} 113 (2): 607--13.
-\url{https://doi.org/10.1017/S0003055418000862}.
-
-\leavevmode\hypertarget{ref-White2019}{}%
----------.
-2019b.
-``Misdemeanor Disenfranchisement?
-The Demobilizing Effects of Brief Jail Spells on Potential Voters.'' \emph{American Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/S000305541800093X}.
-
-\leavevmode\hypertarget{ref-Wilson1990}{}%
-Wilson, William J. 1990. \emph{The Truly Disadvantaged: The Inner City, the Underclass, and Public Policy}. Paperback ed., {[}Nachdr.{]}. Chicago: Univ. of Chicago Press.
-
-\leavevmode\hypertarget{ref-Wu2009}{}%
-Wu, Yuning, Ivan Y.
-Sun, and Ruth A.
-Triplett.
-2009.
-``Race, Class or Neighborhood Context: Which Matters More in Measuring Satisfaction with Police?'' \emph{Justice Quarterly} 26 (1): 125--56.
-\url{https://doi.org/10.1080/07418820802119950}.
-\end{cslreferences}
-
 \end{document}

### 2020-07-06 markdown/part2/blind.tex
commit july 6

--- 2020-03-05 markdown/part2/part2_standalone.tex
@@
 \subtitle{Voting Rights Restoration and Post-Supervision Participation}
-\author{}
-\date{\vspace{-2.5em}March 05, 2020
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-Word Count: 9,553}
+
+\date{April 08, 2020}
 
@@
 The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
+Middle names must match exactly as well, except for if one record has only a middle initial, which is allowed to match to a full middle name in the other set of records.
 
-\begin{itemize}
-\tightlist
-\item
-  Middle names are identical. If neither set of records includes a middle name, this condition is met.
-\item
-  A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-\item
-  A middle name or middle initial in one set of records, and a missing middle name in the other set.
-\end{itemize}
-
-Thus, ``John Andrew Doe'' and ``John A Doe'' would count as matches. Similarly, ``John Andrew Doe'' and ``John Doe'' would count, while ``John Andrew Doe'' and ``John Anthony Doe'' would not.
-
-There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) offers one way to test their prevalence using placebo matching.
+I adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
 I slightly alter the date of birth reported in the parole discharge dataset to create false records.
@@
 
-{\centering \includegraphics{../../temp/flowchart}
+{\centering \includegraphics{flowchart}
 
@@
 
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-16-chart-1}
+{\centering \includegraphics{to-16-chart-1}
 
@@
 
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-chart-1}
+{\centering \includegraphics{to-18-chart-1}
 
@@
 
-\input{"../../temp/table3.tex"}
+\begin{table}[H] \centering
+\fontsize{10}{12}\selectfont
+\caption{\label{tab:to-18-logit-short} Individual-Level Logit Models}
+\label{}
+\begin{tabular}{@{\extracolsep{5pt}}lcc}
+\\[-1.8ex]\hline \\[-1.8ex]
+\\[-1.8ex] & \multicolumn{2}{c}{Cast Ballot in 2018 Election} \\
+\\[-1.8ex] & (1) & (2)\\
+\hline \\[-1.8ex]
+Days Since Discharged &  & 0.001 \\
+&  & (0.003) \\
+& & \\
+Days Since Discharged\textsuperscript{2} &  & $-$0.00000 \\
+&  & (0.00000) \\
+& & \\
+D(Male) & $-$0.510$^{***}$ & $-$0.511$^{***}$ \\
+& (0.183) & (0.183) \\
+& & \\
+Age (Years) & 0.044$^{***}$ & 0.044$^{***}$ \\
+& (0.005) & (0.005) \\
+& & \\
+Years on Parole & 0.041$^{**}$ & 0.041$^{**}$ \\
+& (0.019) & (0.019) \\
+& & \\
+Constant & $-$6.646$^{***}$ & $-$6.878$^{***}$ \\
+& (1.069) & (1.168) \\
+& & \\
+\hline \\[-1.8ex]
+Race / Ethnicity FE & X & X \\
+Felony Class FE & X & X \\
+\hline \\[-1.8ex]
+Observations & 14,155 & 14,155 \\
+Log Likelihood & $-$1,365.859 & $-$1,365.740 \\
+Akaike Inf. Crit. & 2,761.719 & 2,765.480 \\
+\hline \\[-1.8ex]
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\Robust standard errors in parentheses.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 
-\input{"../../temp/table_whatever2.tex"}
+\begin{table}[H]
+\fontsize{10}{12}\selectfont
+
+\caption{\label{tab:demo-rd} Demographics of "Control" and "Intent-to-Treat" Groups}
+\centering
+\begin{tabular}[t]{lcc}
+\toprule
+Variable & Control & Intend-to-Treat\\
+\midrule
+Percent Male & 90.3\% & 90.2\%\\
+Percent NH-Black & 42.9\% & 41.6\%\\
+Percent NH-White & 30.8\% & 31.1\%\\
+Percent Latino & 22.6\% & 23.1\%\\
+Age (Years)* & 41.5 & 40.4\\
+Time Spent on Parole (Years) & 2.2 & 2.1\\
+Percent Voted in 2016 & 1.2\% & 1.3\%\\
+\bottomrule
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
+\\Shows demographics for individuals discharged from parole \\between January 1st, 2017, and October 12th, 2018.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 \begin{singlespace}
+\begin{table}[H] \centering
+\fontsize{10}{12}\selectfont
+\caption{\label{tab:to-18-logit} Individual-Level Logit Models}
+\label{}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}
+\\[-1.8ex]\hline \\[-1.8ex]
+\\[-1.8ex] & \multicolumn{3}{c}{Cast Ballot in 2018 Election} \\
+\\[-1.8ex] & (1) & (2) & (3)\\
+\hline \\[-1.8ex]
+D(Discharged After EO 181) & 0.235$^{**}$ & 0.296$^{***}$ & 0.300$^{***}$ \\
+& (0.112) & (0.113) & (0.113) \\
+& & & \\
+D(Male) &  & $-$0.337$^{**}$ & $-$0.369$^{**}$ \\
+&  & (0.158) & (0.159) \\
+& & & \\
+Age (Years) &  & 0.051$^{***}$ & 0.045$^{***}$ \\
+&  & (0.004) & (0.004) \\
+& & & \\
+Years on Parole &  &  & 0.037$^{**}$ \\
+&  &  & (0.016) \\
+& & & \\
+Constant & $-$3.839$^{***}$ & $-$7.278$^{***}$ & $-$7.223$^{***}$ \\
+& (0.059) & (1.050) & (1.054) \\
+& & & \\
+\hline \\[-1.8ex]
+Race / Ethnicity FE &  & X & X \\
+Felony Class FE &  &  & X \\
+\hline \\[-1.8ex]
+Observations & 18,423 & 18,423 & 18,423 \\
+Log Likelihood & $-$1,967.190 & $-$1,872.755 & $-$1,859.151 \\
+Akaike Inf. Crit. & 3,938.380 & 3,765.509 & 3,750.302 \\
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\Robust standard errors in parentheses.}}}
+\end{tabular}
+\end{table}
 
-\input{"../../temp/table4.tex"}
 \end{singlespace}
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-chunk}\label{tab:iv-models} Rights Restoration and Turnout}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lccccc}
+\\[-1.8ex]\hline
+& \multicolumn{5}{c}{Cast Ballot in 2018 Election} \\
+& \multicolumn{2}{c}{2SLS} & \multicolumn{1}{c}{2S Probit} &   \multicolumn{2}{c}{Bivariate Probit}\\
+& & & \multicolumn{1}{c}{(Marg. Effects)} &   \multicolumn{2}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-6}
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\
+\hline
+D(Rights Restored) & 0.0075** & 0.0096** & 0.0169*** & 0.0063** & 0.0084*** \\
+ & (0.0038) & (0.0038) & (0.0036) & (0.0030) & (0.0031) \\
+D(Male) &  & -0.0085** & -0.0079** &  & -0.0081** \\
+ &  & (0.0042) & (0.0035) &  & (0.0035) \\
+Age (Years) &  & 0.0011*** & 0.0010*** &  & 0.0010*** \\
+ &  & (0.0001) & (0.0001) &  & (0.0001) \\
+Years on Parole &  & 0.0015** & 0.0011** &  & 0.0011** \\
+ &  & (0.0006) & (0.0004) &  & (0.0004) \\
+Constant & 0.0210*** & -0.0416*** &  &  &  \\
+ & (0.0012) & (0.0090) &  &  &  \\
+\hline
+Race / Ethnicity FE &  & X & X &  & X \\
+Felony Class FE &  & X & X &  & X \\
+\hline
+Observations & 18,423 & 18,423 & 18,423 & 18,423 & 18,423 \\
+Adjusted R$^2$ & 0.0007 & 0.0134 &  &  &  \\
+Wald $\chi^2$ &  &  & 210.2 & 1168 & 1571 \\
+ $\rho$ &  &  &  & 0.138 & 0.0803 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in Models 2 -- 3. \\
+Models 3 -- 5 present marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 
-{\centering \includegraphics{part2_standalone_files/figure-latex/sensitivity-1}
+{\centering \includegraphics{sensitivity-1}
 
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean_race.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-r-chunk}\label{tab:iv-models-r} Rights Restoration and Turnout}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lcccccc}
+\\[-1.8ex]\hline
+& \multicolumn{6}{c}{Cast Ballot in 2018 Election} \\
+& \multicolumn{3}{c}{2SLS} &  \multicolumn{3}{c}{Bivariate Probit}\\
+& & & &   \multicolumn{3}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\\[-1.8ex] & White & Non-White & Black& White & Non-White & Black\\
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
+\hline
+D(Rights Restored) & 0.0182*** & 0.0045 & 0.0032 & 0.0176*** & 0.0039 & 0.0033 \\
+ & (0.0067) & (0.0045) & (0.0057) & (0.0067) & (0.0034) & (0.0046) \\
+D(Male) & -0.0020 & -0.0165** & -0.0202** & -0.0009 & -0.0141*** & -0.0174*** \\
+ & (0.0052) & (0.0067) & (0.0091) & (0.0054) & (0.0045) & (0.0061) \\
+Age (Years) & 0.0016*** & 0.0009*** & 0.0010*** & 0.0013*** & 0.0008*** & 0.0010*** \\
+ & (0.0002) & (0.0001) & (0.0002) & (0.0002) & (0.0001) & (0.0002) \\
+Years on Parole & 0.0037** & 0.0008 & 0.0024** & 0.0023** & 0.0006 & 0.0018** \\
+ & (0.0018) & (0.0007) & (0.0011) & (0.0011) & (0.0005) & (0.0007) \\
+Constant & -0.0498*** & -0.0100 & -0.0143 &  &  &  \\
+ & (0.0113) & (0.0092) & (0.0127) &  &  &  \\
+\hline
+Felony Class FE & X & X & X & X & X & X \\
+\hline
+Observations & 5,691 & 12,732 & 7,843 & 5,691 & 12,732 & 7,843 \\
+Adjusted R$^2$ & 0.0211 & 0.0095 & 0.0140 &  &  &  \\
+Wald $\chi^2$ &  &  &  & 414.2 & 1207 & 984.2 \\
+ $\rho$ &  &  &  & -0.181 & 0.255 & 0.291 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in all models. \\
+Models 4 -- 6 present marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 \begin{singlespace}
+\begin{table}[H] \centering
+\fontsize{10}{12}\selectfont
+\caption{\label{tab:black-to} Turnout Rate Among Former Parolees}
+\label{}
+\begin{tabular}{@{\extracolsep{5pt}}lcccccc}
+\\[-1.8ex]\hline \\[-1.8ex]
+\\[-1.8ex] & \multicolumn{6}{c}{Cast Ballot in 2018 Election} \\
+& \multicolumn{2}{c}{Discharged Prior to EO} & \multicolumn{2}{c}{ITT Group} & \multicolumn{2}{c}{Treatment Group} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}  \cmidrule(l{3pt}r{3pt}){6-7}
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
+\hline \\[-1.8ex]
+D(Black) & 0.417$^{***}$ & 0.285$^{**}$ & 0.039 & $-$0.161 & 0.084 & $-$0.039 \\
+& (0.121) & (0.141) & (0.197) & (0.215) & (0.212) & (0.236) \\
+& & & & & & \\
+D(Male) & $-$0.563$^{***}$ & $-$0.486$^{**}$ & $-$0.053 & $-$0.119 & 0.053 & $-$0.019 \\
+& (0.178) & (0.195) & (0.327) & (0.332) & (0.364) & (0.369) \\
+& & & & & & \\
+Age (Years) & 0.045$^{***}$ & 0.045$^{***}$ & 0.050$^{***}$ & 0.052$^{***}$ & 0.048$^{***}$ & 0.051$^{***}$ \\
+& (0.005) & (0.005) & (0.007) & (0.008) & (0.008) & (0.009) \\
+& & & & & & \\
+Years on Parole & 0.033$^{*}$ & 0.049$^{**}$ & 0.017 & 0.075 & 0.133$^{**}$ & 0.135$^{**}$ \\
+& (0.019) & (0.023) & (0.033) & (0.046) & (0.054) & (0.057) \\
+& & & & & & \\
+Constant & $-$5.825$^{***}$ & $-$5.862$^{***}$ & $-$6.125$^{***}$ & $-$5.991$^{***}$ & $-$6.182$^{***}$ & $-$6.075$^{***}$ \\
+& (0.299) & (0.321) & (0.486) & (0.530) & (0.548) & (0.601) \\
+& & & & & & \\
+\hline \\[-1.8ex]
+Observations & 14,155 & 10,963 & 4,268 & 3,280 & 3,093 & 2,514 \\
+Log Likelihood & $-$1,371.825 & $-$1,133.731 & $-$492.408 & $-$402.919 & $-$402.966 & $-$341.008 \\
+Akaike Inf. Crit. & 2,763.649 & 2,287.462 & 1,004.817 & 825.838 & 825.932 & 702.015 \\
+\hline \\[-1.8ex]
+\multicolumn{7}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\Robust standard errors in parentheses.
+\\Latino individuals excluded from Models 2, 4, and 6.}}}
+\end{tabular}
+\end{table}
 
-\input{"../../temp/table5.tex"}
 \end{singlespace}
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean_months.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-chunk-time}\label{tab:iv-models-time} Rights Restoration and Turnout}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lcc}
+\\[-1.8ex]\hline
+& \multicolumn{2}{c}{Cast Ballot in 2018 Election} \\
+& \multicolumn{1}{c}{2SLS} & \multicolumn{1}{c}{2S Probit}\\
+& &  \multicolumn{1}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
+\\[-1.8ex] & (1) & (2) \\
+\hline
+Months Restored & 0.0029** & 0.0050*** \\
+ & (0.0011) & (0.0010) \\
+D(Male) & -0.0084** & -0.0080** \\
+ & (0.0042) & (0.0035) \\
+Age (Years) & 0.0011*** & 0.0010*** \\
+ & (0.0001) & (0.0001) \\
+Years on Parole & 0.0015** & 0.0011** \\
+ & (0.0006) & (0.0004) \\
+Constant & -0.0418*** &  \\
+ & (0.0090) &  \\
+\hline
+Race / Ethnicity FE & X & X \\
+Felony Class FE & X & X \\
+\hline
+Observations & 18,423 & 18,423 \\
+Adjusted R$^2$ & 0.0134 &  \\
+ Wald $\chi^2$ &  & 210.9 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in both models. \\
+Model 2 presents marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}

### 2020-07-06 markdown/part2/part2_standalone.Rmd
commit july 6

@@
 The first and last names must also be exact matches (conditional on the adjustments discussed below).
-The middle names must meet one of the following conditions in order to qualify:
+Middle names must match exactly as well, except for if one record has only a middle initial, which is allowed to match to a full middle name in the other set of records.
 
-* Middle names are identical. If neither set of records includes a middle name, this condition is met.
-* A full middle name in one set of records and only a middle initial in the other. The first letter of the full middle name must be the same as the middle initial in the other set of records.
-* A middle name or middle initial in one set of records, and a missing middle name in the other set.
-
-Thus, "John Andrew Doe" and "John A Doe" would count as matches. Similarly, "John Andrew Doe" and "John Doe" would count, while "John Andrew Doe" and "John Anthony Doe" would not.
-
-There are two types of potential error in this methodology: a false positive will result when a parolee's records matches the record of a voter who is a different individual but shares the same name and date of birth.
-False negatives will occur when an individual has a different name in the different sets of records, or when the birthdate is incorrectly reported in one of the sets of records
-
-Testing for the presence of false positive matches is fairly straightforward.
-@Meredith2013 offers one way to test their prevalence using placebo matching.
+I adopt the test developed in @Meredith2013 to test the prevalence of false positives.
 I slightly alter the date of birth reported in the parole discharge dataset to create false records.

### 2020-07-06 markdown/part2/supplemental.tex
commit july 6

--- 2020-03-04 markdown/part2/registration_regs.tex
@@
 
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
+\newpage
 \doublespace
@@
 \begin{singlespace}
+\begin{table}[H] \centering
+\fontsize{10}{12}\selectfont
+\caption{\label{tab:to-16-logit} Individual-Level Logit Model}
+\label{}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}
+\\[-1.8ex]\hline \\[-1.8ex]
+\\[-1.8ex] & \multicolumn{3}{c}{Cast Ballot in 2016 Election} \\
+\\[-1.8ex] & (1) & (2) & (3)\\
+\hline \\[-1.8ex]
+D(Discharged on or after May 18\textsuperscript{th}, 2016) & $-$0.121 & $-$0.093 & $-$0.103 \\
+& (0.093) & (0.094) & (0.094) \\
+& & & \\
+D(Male) &  & $-$0.490$^{***}$ & $-$0.524$^{***}$ \\
+&  & (0.117) & (0.118) \\
+& & & \\
+Age (Years) &  & 0.034$^{***}$ & 0.030$^{***}$ \\
+&  & (0.003) & (0.003) \\
+& & & \\
+Counts in Most Recent Sentence &  &  & 0.049$^{***}$ \\
+&  &  & (0.014) \\
+& & & \\
+Time Spent on Parole (Years) & $-$3.105$^{***}$ & $-$4.354$^{***}$ & $-$4.264$^{***}$ \\
+& (0.044) & (0.498) & (0.502) \\
+& & & \\
+\hline \\[-1.8ex]
+Race / Ethnicity FE &  & X & X \\
+Felony Class FE &  &  & X \\
+\hline \\[-1.8ex]
+Observations & 16,867 & 16,867 & 16,867 \\
+Log Likelihood & $-$2,927.513 & $-$2,846.416 & $-$2,829.391 \\
+Akaike Inf. Crit. & 5,859.027 & 5,712.831 & 5,690.782 \\
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\Robust standard errors in parentheses.}}}
+\end{tabular}
+\end{table}
 
-\input{"../../temp/table55.tex"}
 \end{singlespace}
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean_reg.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-chunk}\label{tab:iv-models} Rights Restoration and Registration}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lccccc}
+\\[-1.8ex]\hline
+& \multicolumn{5}{c}{Registered to Vote} \\
+& \multicolumn{2}{c}{2SLS} & \multicolumn{1}{c}{2S Probit} &   \multicolumn{2}{c}{Bivariate Probit}\\
+& & & \multicolumn{1}{c}{(Marg. Effects)} &   \multicolumn{2}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-6}
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5)\\
+\hline
+D(Rights Restored) & 0.0137** & 0.0164*** & 0.0253*** & 0.0125** & 0.0157*** \\
+ & (0.0058) & (0.0058) & (0.0055) & (0.0051) & (0.0051) \\
+D(Male) &  & -0.0295*** & -0.0268*** &  & -0.0269*** \\
+ &  & (0.0067) & (0.0054) &  & (0.0054) \\
+Age (Years) &  & 0.0013*** & 0.0013*** &  & 0.0013*** \\
+ &  & (0.0002) & (0.0001) &  & (0.0001) \\
+Years on Parole &  & -0.0000 & -0.0000 &  & -0.0001 \\
+ &  & (0.0009) & (0.0008) &  & (0.0008) \\
+Constant & 0.0545*** & -0.0053 &  &  &  \\
+ & (0.0019) & (0.0151) &  &  &  \\
+\hline
+Race / Ethnicity FE &  & X & X &  & X \\
+Felony Class FE &  & X & X &  & X \\
+\hline
+Observations & 18,423 & 18,423 & 18,423 & 18,423 & 18,423 \\
+Adjusted R$^2$ & 0.0008 & 0.0113 &  &  &  \\
+Wald $\chi^2$ &  &  & 214.6 & 1171 & 1581 \\
+ $\rho$ &  &  &  & 0.0855 & 0.0452 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in Models 2 -- 3. \\
+Models 3 -- 5 present marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean_race_reg.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-r-chunk}\label{tab:iv-models-r} Rights Restoration and Registration}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lcccccc}
+\\[-1.8ex]\hline
+& \multicolumn{6}{c}{Registered to Vote} \\
+& \multicolumn{3}{c}{2SLS} &  \multicolumn{3}{c}{Bivariate Probit}\\
+& & & &   \multicolumn{3}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\\[-1.8ex] & White & Non-White & Black& White & Non-White & Black\\
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
+\hline
+D(Rights Restored) & 0.0235*** & 0.0114 & 0.0030 & 0.0225*** & 0.0109* & 0.0039 \\
+ & (0.0090) & (0.0075) & (0.0093) & (0.0087) & (0.0063) & (0.0084) \\
+D(Male) & -0.0099 & -0.0523*** & -0.0581*** & -0.0093 & -0.0424*** & -0.0479*** \\
+ & (0.0078) & (0.0112) & (0.0146) & (0.0073) & (0.0074) & (0.0101) \\
+Age (Years) & 0.0021*** & 0.0010*** & 0.0008*** & 0.0018*** & 0.0010*** & 0.0008*** \\
+ & (0.0003) & (0.0002) & (0.0003) & (0.0002) & (0.0002) & (0.0003) \\
+Years on Parole & 0.0019 & -0.0010 & 0.0018 & 0.0014 & -0.0010 & 0.0016 \\
+ & (0.0022) & (0.0010) & (0.0016) & (0.0017) & (0.0010) & (0.0015) \\
+Constant & -0.0417*** & 0.0688*** & 0.0834*** &  &  &  \\
+ & (0.0150) & (0.0146) & (0.0197) &  &  &  \\
+\hline
+Felony Class FE & X & X & X & X & X & X \\
+\hline
+Observations & 5,691 & 12,732 & 7,843 & 5,691 & 12,732 & 7,843 \\
+Adjusted R$^2$ & 0.0172 & 0.0066 & 0.0073 &  &  &  \\
+Wald $\chi^2$ &  &  &  & 408.1 & 1163 & 923.7 \\
+ $\rho$ &  &  &  & -0.135 & 0.150 & 0.153 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in all models. \\
+Models 4 -- 6 present marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}
@@
 \begin{singlespace}
-\input{"../../temp/iv_clean_months_reg.tex"}
+\begin{table}[H]
+\caption{\label{tab:iv-models-chunk-time}\label{tab:iv-models-time} Rights Restoration and Registration}
+\fontsize{10}{12}\selectfont
+\centering
+
+\begin{tabular}{lcc}
+\\[-1.8ex]\hline
+& \multicolumn{2}{c}{Registered to Vote} \\
+& \multicolumn{1}{c}{2SLS} & \multicolumn{1}{c}{2S Probit}\\
+& &  \multicolumn{1}{c}{(Marg. Effects)}\\
+\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
+\\[-1.8ex] & (1) & (2) \\
+\hline
+Months Restored & 0.0052*** & 0.0088*** \\
+ & (0.0018) & (0.0016) \\
+D(Male) & -0.0295*** & -0.0267*** \\
+ & (0.0067) & (0.0054) \\
+Age (Years) & 0.0013*** & 0.0013*** \\
+ & (0.0002) & (0.0001) \\
+Years on Parole & -0.0000 & -0.0000 \\
+ & (0.0009) & (0.0008) \\
+Constant & -0.0058 &  \\
+ & (0.0151) &  \\
+\hline
+Race / Ethnicity FE & X & X \\
+Felony Class FE & X & X \\
+\hline
+Observations & 18,423 & 18,423 \\
+Adjusted R$^2$ & 0.0115 &  \\
+ Wald $\chi^2$ &  & 213.6 \\
+\hline
+\multicolumn{3}{l}{\scriptsize{\parbox{.5\linewidth}{\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\
+Robust standard errors in parentheses. \\
+Instrument is dummy indicating whether \\
+individual discharged after EO 181 went into effect. \\
+First-stage equations include covariates in both models. \\
+Model 2 presents marginal effects, not model coefficients.}}}
+\end{tabular}
+\end{table}
+
 \end{singlespace}

### 2020-07-06 presentations/mpsa.Rmd
commit july 6

---
title: "Welcome Home --- Now Vote!"
subtitle: "Voting Rights Restoration and Post-Supervision Turnout"
shorttitle: "Other"
author: Kevin Morris
abstract: kevin.morris@nyu.edu
institute: Brennan Center for Justice
date: Midwest Political Science Association, 2020
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "Berlin"
    keep_tex: true
bibliography: "../markdown/welcome_home_vote.bib"
---
```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(dev = "cairo_pdf")
library(systemfit)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(tidyverse)
library(extrafont)
library(kableExtra)
library(stargazer)
library(data.table)
loadfonts()
```

## Outline

>* Previous research indicates that notification laws increase post-supervision turnout

>* The notification laws are "baked in" to the control group here. In 2018, New York State began implementing rights restoration prior to parole discharge at in-person meetings.

>* So --- did rights restoration *prior* to parole discharge increase formerly incarcerated individuals' propensity to vote?

## Formerly Incarcerated Individuals Turn Out at Low Rates

>* The low turnout rates among the formerly disenfranchised have been shown in a number of studies [e.g. @Burch2011; @Burch2012; @White2019]

>* The *causal effect* of incarceration, however, is the subject of some debate. @Gerber2017, for instance, argues that most of the low turnout observed in the formerly incarcerated population is attributable not to incarceration, but to other factors predating incarceration.

>* Nevertheless, @Lerman2014 and others have convincingly shown through qualitative and quantitative work that criminal justice involvement structures individuals' relationship to the state.

## Notification Laws and Post-Supervision Turnout

>* In 2005, Iowa changed its disenfranchisement laws to allow formerly incarcerated individuals to vote. Not all formerly incarcerated individuals, however, were notified of the change in the mail. @Meredith2015 exploits this uneven treatment, demonstrating that mail notification increased turnout.

>* In 2012, @Gerber2015 expanded on the natural experiment setup of @Meredith2015. In a randomized experiment, they sent mail notifications to some formerly disenfranchised individuals in Connecticut. They too find that notification increases participation, arguing "Whatever the participatory consequences of incarceration they are not in large part impossible to overcome" (p. 924).

## Changes in New York State in 2018

>* In the spring of 2018, Governor Andrew Cuomo signs Executive Order 181 restoring voting rights to individuals *prior* to their discharge from parole.

>* The New York DOC issues guidance calling the program a *priority* initiative, telling parole officers to help their parolees register to vote.

>* Does in-person rights restoration prior to parole discharge increase post-supervision turnout?

## Methodology

>* I limit the pool of individuals to those discharged before the 2018 registration deadline in 2018. These are all individuals, therefore, who could have voted even if the executive order had not been implemented.

>* I compare turnout to those treated by the executive order to those who were untreated by the order afterwards.

>* Because of "imperfect compliance" with the treatment, I adopt an instrumental variables approach. Treatment is instrumented by whether someone came off parole after the executive order went into effect.

## Turnout in 2016 Among the Formerly Incarcerated

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
j <- readRDS("../temp/to16_chart.rds") +
  theme(text = element_text(size = 18)) +
  labs(y = "Turnout in 2016")
j
```

## Turnout in 2018 Among the Formerly Incarcerated

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
j <- readRDS("../temp/to18_chart.rds") +
  theme(text = element_text(size = 18)) +
  labs(y = "Turnout in 2018")
j
```

## Overall Treatment Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center', caption = "", warning = FALSE}
j <- readRDS("../temp/window_change_plot_mpsa.rds")
j
```

## Treatment Effects Vary by Race

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center', caption = "", warning = FALSE}
j <- readRDS("../temp/race_change_plot_mpsa.rds")
j
```

## We Made It!

Thanks!

kevin.morris@nyu.edu

## References {.allowframebreaks}

### 2020-07-06 presentations/mpsa.tex
commit july 6

--- 2020-01-28 presentations/spsa.tex
@@
 \hypersetup{
-  pdftitle={Felony Disenfranchisement and Neighborhood Turnout},
+  pdftitle={Welcome Home --- Now Vote!},
   pdfauthor={Kevin Morris},
@@
 
-\title{Felony Disenfranchisement and Neighborhood Turnout}
-\subtitle{The Case of New York City}
+\title{Welcome Home --- Now Vote!}
+\subtitle{Voting Rights Restoration and Post-Supervision Turnout}
 \author{Kevin Morris}
-\date{Southern Political Science Association, 2020}
+\date{Midwest Political Science Association, 2020}
 \institute{Brennan Center for Justice}
@@
 \item
-  Introducing space into the conversation
+  Previous research indicates that notification laws increase
+  post-supervision turnout
 \end{itemize}
@@
 \item
-  Redefining ``lost voters''
+  The notification laws are ``baked in'' to the control group here. In
+  2018, New York State began implementing rights restoration prior to
+  parole discharge at in-person meetings.
 \end{itemize}
@@
 \item
-  Identifying neighborhoods with lost voters
+  So --- did rights restoration \emph{prior} to parole discharge
+  increase formerly incarcerated individuals' propensity to vote?
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Formerly Incarcerated Individuals Turn Out at Low Rates}
+\protect\hypertarget{formerly-incarcerated-individuals-turn-out-at-low-rates}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  The low turnout rates among the formerly disenfranchised have been
+  shown in a number of studies (e.g.~Burch 2011, 2012; White 2019)
 \end{itemize}
@@
 \item
-  Testing turnout effects
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Introducing Space}
-\protect\hypertarget{introducing-space}{}
-\begin{itemize}[<+->]
-\tightlist
-\item
-  Most of the existing literature on felony disenfranchisement and
-  indirect turnout effects looks for effects at the state level (Miles
-  2004; Bowers and Preuhs 2009; King and Erickson 2016)
+  The \emph{causal effect} of incarceration, however, is the subject of
+  some debate. Gerber et al. (2017), for instance, argues that most of
+  the low turnout observed in the formerly incarcerated population is
+  attributable not to incarceration, but to other factors predating
+  incarceration.
 \end{itemize}
@@
 \item
-  But we \emph{know} that incarceration patterns aren't uniformly
-  distributed throughout a given state (see, for instance, Gelman,
-  Fagan, and Kiss 2007)
+  Nevertheless, Lerman and Weaver (2014) and others have convincingly
+  shown through qualitative and quantitative work that criminal justice
+  involvement structures individuals' relationship to the state.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Notification Laws and Post-Supervision Turnout}
+\protect\hypertarget{notification-laws-and-post-supervision-turnout}{}
+\begin{itemize}[<+->]
+\tightlist
+\item
+  In 2005, Iowa changed its disenfranchisement laws to allow formerly
+  incarcerated individuals to vote. Not all formerly incarcerated
+  individuals, however, were notified of the change in the mail.
+  Meredith and Morse (2015) exploits this uneven treatment,
+  demonstrating that mail notification increased turnout.
 \end{itemize}
@@
 \item
-  Turnout effects are likely to be socially mediated - and therefore
-  show up in the neighborhoods where disenfranchised individuals live
-  (e.g.~Foladare 1968; Huckfeldt 1979; Cho, Gimpel, and Dyck 2006)
+  In 2012, Gerber et al. (2015) expanded on the natural experiment setup
+  of Meredith and Morse (2015). In a randomized experiment, they sent
+  mail notifications to some formerly disenfranchised individuals in
+  Connecticut. They too find that notification increases participation,
+  arguing ``Whatever the participatory consequences of incarceration
+  they are not in large part impossible to overcome'' (p.~924).
 \end{itemize}
@@
 
-\begin{frame}{Redefining Lost Voters}
-\protect\hypertarget{redefining-lost-voters}{}
+\begin{frame}{Changes in New York State in 2018}
+\protect\hypertarget{changes-in-new-york-state-in-2018}{}
 \begin{itemize}[<+->]
@@
 \item
-  Much of the existing literature cannot identify individuals who
-  \emph{would have voted} if not for their disenfranchisement (Burch
-  2013)
+  In the spring of 2018, Governor Andrew Cuomo signs Executive Order 181
+  restoring voting rights to individuals \emph{prior} to their discharge
+  from parole.
 \end{itemize}
@@
 \item
-  This undermines our ability to distinguish the causal effects of
-  incarceration / probation from the effects of disenfranchisement
+  The New York DOC issues guidance calling the program a \emph{priority}
+  initiative, telling parole officers to help their parolees register to
+  vote.
 \end{itemize}
@@
 \item
-  This is especially likely given the low turnout propensity among the
-  formerly disenfranchised (e.g.~Meredith and Morse 2015; Gerber et al.
-  2017; White 2019)
+  Does in-person rights restoration prior to parole discharge increase
+  post-supervision turnout?
 \end{itemize}
@@
 
-\begin{frame}{This Project Addresses Both Problems}
-\protect\hypertarget{this-project-addresses-both-problems}{}
+\begin{frame}{Methodology}
+\protect\hypertarget{methodology}{}
 \begin{itemize}[<+->]
@@
 \item
-  Administrative data from the NYS DOC allows us to conduct
-  individual-level analysis
+  I limit the pool of individuals to those discharged before the 2018
+  registration deadline in 2018. These are all individuals, therefore,
+  who could have voted even if the executive order had not been
+  implemented.
 \end{itemize}
@@
 \item
-  Purge records in the NYS voter file allow us to construct vote
-  histories for disenfranchised individuals
+  I compare turnout to those treated by the executive order to those who
+  were untreated by the order afterwards.
 \end{itemize}
@@
 \item
-  Lost voters are all formally disenfranchised individuals who have
-  voted in the past 10 years
+  Because of ``imperfect compliance'' with the treatment, I adopt an
+  instrumental variables approach. Treatment is instrumented by whether
+  someone came off parole after the executive order went into effect.
 \end{itemize}
@@
 
-\begin{frame}{Identifying Neighborhoods with Lost Voters}
-\protect\hypertarget{identifying-neighborhoods-with-lost-voters}{}
-\includegraphics[width=1\linewidth]{../temp/lost_voters_map}
+\begin{frame}{Turnout in 2016 Among the Formerly Incarcerated}
+\protect\hypertarget{turnout-in-2016-among-the-formerly-incarcerated}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{mpsa_files/figure-beamer/unnamed-chunk-1-1} \end{center}
 \end{frame}
 
-\begin{frame}{Methodology}
-\protect\hypertarget{methodology}{}
-\begin{itemize}
-\item
-  A genetic match algorithm (Sekhon 2011) is used to match treated
-  census block groups to untreated ones.
-\item
-  Each block group is matched to 30 untreated block groups; matches are
-  done with replacement.
-\end{itemize}
+\begin{frame}{Turnout in 2018 Among the Formerly Incarcerated}
+\protect\hypertarget{turnout-in-2018-among-the-formerly-incarcerated}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{mpsa_files/figure-beamer/unnamed-chunk-2-1} \end{center}
 \end{frame}
 
-\begin{frame}{Matching Results}
-\protect\hypertarget{matching-results}{}
-\begin{center}\includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_white_spsa} \includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_black_spsa} \end{center}
+\begin{frame}{Overall Treatment Effects}
+\protect\hypertarget{overall-treatment-effects}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{mpsa_files/figure-beamer/unnamed-chunk-3-1} \end{center}
 \end{frame}
 
-\begin{frame}{Matching Results}
-\protect\hypertarget{matching-results-1}{}
-\begin{center}\includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/perc_latino_spsa} \includegraphics[width=0.49\linewidth,height=0.49\textheight]{../temp/income_spsa} \end{center}
-\end{frame}
-
-\begin{frame}{Testing Turnout Effects}
-\protect\hypertarget{testing-turnout-effects}{}
-\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot1} \end{center}
-\end{frame}
-
-\begin{frame}{Testing Turnout Effects}
-\protect\hypertarget{testing-turnout-effects-1}{}
-\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot2} \end{center}
-\end{frame}
-
-\begin{frame}{Testing Turnout Effects}
-\protect\hypertarget{testing-turnout-effects-2}{}
-\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{../temp/coef_plot3b} \end{center}
-\end{frame}
-
-\begin{frame}{Testing Turnout Effects}
-\protect\hypertarget{testing-turnout-effects-3}{}
-\includegraphics[width=1\linewidth]{../output/dep_map}
-\end{frame}
-
-\begin{frame}{Conclusions}
-\protect\hypertarget{conclusions}{}
-\begin{itemize}[<+->]
-\tightlist
-\item
-  The indirect turnout effects identified in past research appear to be
-  geographically concentrated where the lost voters live.
-\end{itemize}
-
-\begin{itemize}[<+->]
-\tightlist
-\item
-  Hajnal (2009) and others have demonstrated that turnout differentials
-  can have real political implications, especially in low-turnout
-  contests.
-\end{itemize}
-
-\begin{itemize}[<+->]
-\tightlist
-\item
-  Felony disenfranchisement undermines Black political representation at
-  the local level.
-\end{itemize}
+\begin{frame}{Treatment Effects Vary by Race}
+\protect\hypertarget{treatment-effects-vary-by-race}{}
+\begin{center}\includegraphics[width=0.85\linewidth,height=0.85\textheight]{mpsa_files/figure-beamer/unnamed-chunk-4-1} \end{center}
 \end{frame}
@@
 \begin{cslreferences}
-\leavevmode\hypertarget{ref-Bowers2009}{}%
-Bowers, Melanie, and Robert R. Preuhs. 2009. ``Collateral Consequences
-of a Collateral Penalty: The Negative Effect of Felon Disenfranchisement
-Laws on the Political Participation of Nonfelons.'' \emph{Social Science
-Quarterly} 90 (3): 722--43.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00640.x}.
+\leavevmode\hypertarget{ref-Burch2011}{}%
+Burch, Traci. 2011. ``Turnout and Party Registration Among Criminal
+Offenders in the 2008 General Election.'' \emph{Law \& Society Review}
+45 (3): 699--730.
+\url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
 
-\leavevmode\hypertarget{ref-Burch2013}{}%
-Burch, Traci. 2013. ``Effects of Imprisonment and Community Supervision
-on Neighborhood Political Participation in North Carolina.'' Edited by
-Christopher Wildeman, Jacob S. Hacker, and Vesla M. Weaver. \emph{The
-ANNALS of the American Academy of Political and Social Science} 651 (1):
-184--201. \url{https://doi.org/10.1177/0002716213503093}.
+\leavevmode\hypertarget{ref-Burch2012}{}%
+---------. 2012. ``Did Disfranchisement Laws Help Elect President Bush?
+New Evidence on the Turnout Rates and Candidate Preferences of Florida's
+Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
+\url{https://doi.org/10.1007/s11109-010-9150-9}.
 
-\leavevmode\hypertarget{ref-Cho2006}{}%
-Cho, Wendy K. Tam, James G. Gimpel, and Joshua J. Dyck. 2006.
-``Residential Concentration, Political Socialization, and Voter
-Turnout.'' \emph{The Journal of Politics} 68 (1): 156--67.
-\url{https://doi.org/10.1111/j.1468-2508.2006.00377.x}.
-
-\leavevmode\hypertarget{ref-Foladare1968}{}%
-Foladare, Irving S. 1968. ``The Effect of Neighborhood on Voting
-Behavior.'' \emph{Political Science Quarterly} 83 (4): 516.
-\url{https://doi.org/10.2307/2146812}.
-
-\leavevmode\hypertarget{ref-Gelman2007}{}%
-Gelman, Andrew, Jeffrey Fagan, and Alex Kiss. 2007. ``An Analysis of the
-New York City Police Departments ``Stop-and-Frisk'' Policy in the
-Context of Claims of Racial Bias.'' \emph{Journal of the American
-Statistical Association} 102 (479): 813--23.
-\url{https://doi.org/10.1198/016214506000001040}.
+\leavevmode\hypertarget{ref-Gerber2015}{}%
+Gerber, Alan S., Gregory A. Huber, Marc Meredith, Daniel R. Biggers, and
+David J. Hendry. 2015. ``Can Incarcerated Felons Be (Re)Integrated into
+the Political System? Results from a Field Experiment.'' \emph{American
+Journal of Political Science} 59 (4): 912--26.
+\url{https://doi.org/10.1111/ajps.12166}.
 
 \leavevmode\hypertarget{ref-Gerber2017}{}%
-Gerber, Alan S., Gregory A. Huber, Marc Meredith, Daniel R. Biggers, and
-David J. Hendry. 2017. ``Does Incarceration Reduce Voting? Evidence
-About the Political Consequences of Spending Time in Prison.'' \emph{The
-Journal of Politics} 79 (4): 1130--46.
-\url{https://doi.org/10.1086/692670}.
+---------. 2017. ``Does Incarceration Reduce Voting? Evidence About the
+Political Consequences of Spending Time in Prison.'' \emph{The Journal
+of Politics} 79 (4): 1130--46. \url{https://doi.org/10.1086/692670}.
 
-\leavevmode\hypertarget{ref-Hajnal2009}{}%
-Hajnal, Zoltan. 2009. \emph{America's Uneven Democracy}. Cambridge
-University Press. \url{https://doi.org/10.1017/cbo9780511800535}.
-
-\leavevmode\hypertarget{ref-Huckfeldt1979}{}%
-Huckfeldt, R. Robert. 1979. ``Political Participation and the
-Neighborhood Social Context.'' \emph{American Journal of Political
-Science} 23 (3): 579. \url{https://doi.org/10.2307/2111030}.
-
-\leavevmode\hypertarget{ref-King2016}{}%
-King, Bridgett A., and Laura Erickson. 2016. ``Disenfranchising the
-Enfranchised.'' \emph{Journal of Black Studies} 47 (8): 799--821.
-\url{https://doi.org/10.1177/0021934716659195}.
+\leavevmode\hypertarget{ref-Lerman2014}{}%
+Lerman, Amy E., and Vesla M. Weaver. 2014. \emph{Arresting Citizenship:
+The Democratic Consequences of American Crime Control}. Chicago Studies
+in American Politics. Chicago ; London: The University of Chicago Press.
 
@@
 
-\leavevmode\hypertarget{ref-Miles2004}{}%
-Miles, Thomas J. 2004. ``Felon Disenfranchisement and Voter Turnout.''
-\emph{The Journal of Legal Studies} 33 (1): 85--129.
-\url{https://doi.org/10.1086/381290}.
-
-\leavevmode\hypertarget{ref-Sekhon2011}{}%
-Sekhon, Jasjeet S. 2011. ``Multivariate and Propensity Score Matching
-Software with Automated Balance Optimization: The Matching Package for
-R.'' \emph{Journal of Statistical Software} 42 (7).
-\url{https://doi.org/10.18637/jss.v042.i07}.
-
 \leavevmode\hypertarget{ref-White2019}{}%
@@
 Political Science Review} 113 (2): 311--24.
-\url{https://doi.org/10.1017/s000305541800093x}.
+\url{https://doi.org/10.1017/S000305541800093X}.
 \end{cslreferences}

### 2020-07-06 code/00_initialize.R

@@ -5,6 +5,7 @@ library(DiagrammeRsvg)
 library(rsvg)
 library(lubridate)
 library(Hmisc)
+library(Greg)
 library(RStata)
 library(estimatr)
 library(Matching)

### 2020-07-06 code/part_2/04_match_parolees_voter.R

@@ -2,7 +2,8 @@
 
 ############# READ IN NYS VOTER FILES #############
 db <- dbConnect(SQLite(), "D:/rolls.db")
-nys_roll <- dbGetQuery(db, "select political_party, last_name, first_name, middle_name, dob, voter_status, nys_id from nys_roll_0319")
+nys_roll <- dbGetQuery(db, "select political_party, last_name, first_name, middle_name, dob,
+                            voter_status, nys_id from nys_roll_0319")
 
 nys_roll <- nys_roll %>% 
   mutate(a = as.integer(voter_status == "ACTIVE"),
@@ -27,7 +28,7 @@ doccs_to_rolls <- doccs_to_rolls %>%
             ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
   rename(dob = dob_parole)
 
-merge_list <- match_rolls_to_doc(doccs_to_rolls, din, nys_roll, nys_id)
+merge_list <- match_rolls_to_doc(doccs_to_rolls, "din", nys_roll, "nys_id")
 
 small <- merge_list[[1]] %>% 
   select(nys_id, din) %>% 

### 2020-07-06 code/part_2/05_turnout_trends.R

@@ -1,28 +1,28 @@
-
-elects <- fread("./raw_data/misc/elects.csv")
-
-############# READ IN NYS VOTER FILES #############
-db <- dbConnect(SQLite(), "D:/rolls.db")
-nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id, political_party from nys_roll_0319")
-
-nys_roll <- nys_roll[nys_roll$nys_id %in% readRDS("./temp/din_nys_parolees.rds")$nys_id, ]
-
-nys_roll <- left_join(nys_roll, readRDS("./temp/din_nys_parolees.rds"), by = "nys_id")
-
-nys_roll <- cSplit(nys_roll, "history", sep = ";", direction = "long", type.convert = F)
-
-nys_roll <- left_join(nys_roll, elects, by = "history")
-
-
-nys_roll <- nys_roll %>%
-  group_by(nys_id) %>%
-  mutate(year = ifelse(is.na(year), 0, year),
-         v2018 = max(year == 2018 & election_type == "general"),
-         v2016 = max(year == 2016 & election_type == "general"),
-         v2008 = max(year == 2008 & election_type == "general")) %>%
-  filter(row_number() == 1)
-
-saveRDS(nys_roll, "./temp/parolee_to_18.rds")
+# 
+# elects <- fread("./raw_data/misc/elects.csv")
+# 
+# ############# READ IN NYS VOTER FILES #############
+# db <- dbConnect(SQLite(), "D:/rolls.db")
+# nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id, political_party from nys_roll_0319")
+# 
+# nys_roll <- nys_roll[nys_roll$nys_id %in% readRDS("./temp/din_nys_parolees.rds")$nys_id, ]
+# 
+# nys_roll <- left_join(nys_roll, readRDS("./temp/din_nys_parolees.rds"), by = "nys_id")
+# 
+# nys_roll <- cSplit(nys_roll, "history", sep = ";", direction = "long", type.convert = F)
+# 
+# nys_roll <- left_join(nys_roll, elects, by = "history")
+# 
+# 
+# nys_roll <- nys_roll %>%
+#   group_by(nys_id) %>%
+#   mutate(year = ifelse(is.na(year), 0, year),
+#          v2018 = max(year == 2018 & election_type == "general"),
+#          v2016 = max(year == 2016 & election_type == "general"),
[…]

### 2020-07-06 code/part_2/12_window_sensitivity.R

@@ -4,9 +4,6 @@
 nys_roll <- readRDS("./temp/parolee_to_18.rds")
 parolees <- readRDS("./temp/parolees_with_restoration.rds")
 
-how_many_restored <- nrow(filter(parolees, year(parole_status_date) >= 2017, restored))
-saveRDS(how_many_restored, "./temp/how_many_restored.rds")
-
 parolees <- left_join(parolees, nys_roll, by = "din") %>% 
   select(-history, -year, -election_type) %>% 
   filter(parole_status == "DISCHARGED",
@@ -30,7 +27,8 @@ parolees <- left_join(parolees, nys_roll, by = "din") %>%
 iv1 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
                age +
                felony_a + felony_b + felony_c + felony_d + felony_e + parole_time | . -restored + finished_post,
-             data = filter(parolees, year(parole_status_date) >= 2017)))) %>% 
+             data = filter(parolees, year(parole_status_date) >= 2017)), level = 0.95,
+             HC_type = "HC1")) %>% 
   add_rownames(var = "coef") %>% 
   filter(coef == "restoredTRUE") %>% 
   rename(lb = `2.5 %`,
@@ -42,7 +40,8 @@ iv1 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
 iv2 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
                                      age +
                                      felony_a + felony_b + felony_c + felony_d + felony_e + parole_time | . -restored + finished_post,
-                                   data = filter(parolees, parole_status_date >= as.Date("2017-06-01"))))) %>% 
+                                   data = filter(parolees, parole_status_date >= as.Date("2017-06-01"))), level = 0.95,
+                                   HC_type = "HC1")) %>% 
   add_rownames(var = "coef") %>% 
   filter(coef == "restoredTRUE") %>% 
   rename(lb = `2.5 %`,
@@ -54,7 +53,8 @@ iv2 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
 iv3 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
                                      age +
                                      felony_a + felony_b + felony_c + felony_d + felony_e + parole_time | . -restored + finished_post,
-                                   data = filter(parolees, parole_status_date >= as.Date("2018-01-01"))))) %>% 
[…]

### 2020-07-06 code/presentations/windows_overall_mpsa.R

@@ -1,0 +1,202 @@
+#### individual level regressions
+
+## this is the treated group
+nys_roll <- readRDS("./temp/parolee_to_18.rds")
+parolees <- readRDS("./temp/parolees_with_restoration.rds")
+
+
+parolees <- left_join(parolees, nys_roll, by = "din") %>% 
+  select(-history, -year, -election_type) %>% 
+  filter(parole_status == "DISCHARGED",
+         parole_status_date >= "2012-01-01",
+         parole_status_date <= "2018-10-12",
+         !is.na(dob_parole)) %>% 
+  mutate(age = as.numeric((as.Date("2018-11-06") - dob_parole) / 365.25),
+         parole_time = as.numeric((parole_status_date - release_date_parole) / 365.25),
+         v2018 = ifelse(is.na(v2018), 0, v2018),
+         finished_post = parole_status_date >= "2018-05-18",
+         days_since_done = as.numeric(as.Date("2018-11-06") - parole_status_date),
+         days2 = days_since_done^2,
+         restored = ifelse(is.na(restored), F, restored),
+         days_since_m21 = parole_status_date - as.Date("2018-05-19"),
+         v2016 = ifelse(is.na(v2016), 0, v2016),
+         v2008 = ifelse(is.na(v2008), 0, v2008),
+         registered = voter_status %in% c("ACTIVE", "INACTIVE"),
+         male = sex == "MALE")
+
+
+iv1 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
+                                            age +
+                                            felony_a + felony_b + felony_c + felony_d + felony_e + parole_time | . -restored + finished_post,
+                                          data = filter(parolees, year(parole_status_date) >= 2017)), level = 0.95,
+                                    HC_type = "HC1")) %>% 
+  add_rownames(var = "coef") %>% 
+  #filter(coef == "restoredTRUE") %>% 
+  rename(lb = `2.5 %`,
+         ub = `97.5 %`) %>% 
+  mutate(estimate = (lb + ub) / 2,
+         window = "1/1/2017 - 10/12/2018")
+
+iv2 <- as.data.frame(confint_robust(ivreg(v2018 ~ restored + male + race +
+                                            age +
+                                            felony_a + felony_b + felony_c + felony_d + felony_e + parole_time | . -restored + finished_post,
+                                          data = filter(parolees, parole_status_date >= as.Date("2017-06-01"))), level = 0.95,
+                                    HC_type = "HC1")) %>% 
[…]

### 2020-07-06 markdown/part2/part2_standalone.Rmd
substantial re-write

@@
 In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to parolees increases their post-supervision propensity to cast a ballot.
+By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
 I demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
@@
 In the spring of 2018,^[Although the executive order was signed on April 18^th^, it did not go into effect until May 18^th^.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for individuals who were still on parole on election day --- they would have been disenfranchised otherwise.
-The change in policy is also beneficial for felony probationers.
-Despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, it has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+Such a move of course directly re-enfranchised individuals who were still on parole on election day.
+The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
 
@@
 
-# Turnout Among the Formerly Disenfranchised {-}
-
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-@Haselswerdt2009 matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states [-@Burch2012; -@Burch2011].
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she estimated the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy [e.g.
-@Meredith2013; @Meredith2015].
-
-Much of this literature establishes that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, @Gerber2017 uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that "it appears that spending time in prison does not have large negative effects on subsequent participation" (1144).
-
-@White2019, however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with @Gerber2017; as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
-By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-
 # Notification, Re-enfranchisement, and Turnout {-}
 
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration *causes* low turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Whether --- and to what extent --- incarceration causes lower turnout among citizens is the subject of some debate in the literature.
+@Gerber2017, for instance, argues that prison does not materially impact post-incarceration turnout, though @White2019 finds that jail time can decrease the future turnout of misdemeanants.
+Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, @Burch2011).
+
+Regardless of whether incarceration reduces turnout, the  state has a unique opportunity to craft policies that will impact individuals under formal supervision.
 Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
@@
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
-When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
+When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics.
 The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
+The basic design is therefore an interrupted time series, in which I test whether turnout for individuals discharged after the executive order took effect turned out at higher rates than those discharged before the policy change.
+This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall turnout among formerly incarcerated individuals.
 
-The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually treated by having their voting rights restored while on parole in a meeting with their parole officer.
-In other words, our natural experiment has "imperfect compliance" [see, for instance, @Angrist1996].
-Noncitizens, for instance, were not eligible for rights "restoration" while on parole because in New York State they were not allowed to vote even prior to their conviction.
-Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intent-to-treat effect.
-Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
-
-Although we can identify the individuals in the assigned-to-treatment group who do not comply with their treatment status, we do not know who among the assigned-to-control group would have gone untreated if they had been assigned to the other group.
-We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
-In an ideal world, we would limit both the assigned-to-treatment *and* the assigned-to-control groups to voting-eligible individuals; we do not live in such a world.
-Worse yet, there is an obvious relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are both ineligible to have voting rights bestowed and cannot vote.
-
-The standard way of dealing with imperfect compliance in an experimental setting is to leverage the (random) assignment to treatment as an instrument for the (nonrandom) actual compliance [@Angrist1996].
-Such an approach has been widely used in the context of voter turnout [e.g.
-@Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
-@Gerber2000 provides a helpful example of why this two-step process is necessary.
-The authors test the efficacy of phone calls on increasing turnout.
-One group is assigned to be the control; the authors do not attempt to call them.
-Another set of voters receive a phone call, but only some of them are actually contacted, or "treated" with the call.
-The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
-Then "treatment" would simply identify a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
-
-The same process is likely at play here.
-The control group includes individuals that *would have* had their rights restored at an in-person meeting if they had been discharged later; it also includes individuals who would not have received the treatment thanks to their citizenship status.
-By comparing turnout among the actually treated to the control group, we may simply be comparing the turnout of a group of *citizens* to the turnout of a group that includes both citizens *and* noncitizens.
-Although a positive "treatment effect" is likely in this context, it would be unable to speak to causality.
-The instrumental variables approach allows us to mitigate this problem.
-Moreover, a very small number of individuals in the control group are marked as having their rights restored prior to the effective date of the executive order.
-While this likely reflects paperwork errors, I nonetheless treat them as a second set of non-compliers.
-
-Figure \ref{fig:flow} demonstrates the buckets into which former parolees can fall.
-Although I argue that a parolee's assignment to the control or treatment categories is random and uncorrelated with turnout except via the treatment, whether a treated former parolee is treated or is a noncomplier is not random and is associated with their propensity to vote through mechanisms other than the treatment.
-
-```{r flow, echo = FALSE, message = F, warning = F, fig.width = 4, fig.height = 4, fig.cap = "\\label{fig:flow}Categorization of Former Parolees", fig.pos = "H", fig.align = "center"}
-knitr::include_graphics("../../temp/flowchart.pdf")
-```
-
-Using the natural experiment with imperfect compliance setup, I begin by estimating the intent-to-treat effect of the policy change.
-This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
-I use turnout in the 2018 election as the dependent variable in these logistic regressions.
-
-After estimating the intent-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
-As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
-To explore this potential, I also test whether the treatment had different effects for different subgroups of the formerly incarcerated population.
-I finally ask whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+As discussed above, not all formerly incarcerated individuals were eligible to have their rights restored; the ITT effect will therefore be biased towards zero, insofar as the behavior of ineligible parolees remains unchanged.
+After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
+In this setup, rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
+This instrumentation process is a common method for identifying the CACE [see, for instance, @Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
 
@@
 
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, `r round(readRDS("./temp/turnout_in_16_finished_10.rds") * 100, digits = 1)` percent cast a ballot in the 2016 election, while just `r round(readRDS("./temp/turnout_in_16_finished_15.rds") * 100, digits = 1)` percent of those last discharged from parole in 2015 did so.^[Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.] A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through May 2016, and extended through October 2016.
+Before estimating the interrupted time series model, we must establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
+It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote, regardless of whether their voting rights are restored before they finish their sentence.
 
-```{r to-16-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-16}Turnout in 2016 Presidential Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to16_chart.rds")
-print(j)
-```
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through May 2018, and extended through October 2018.
-```{r to-18-chart, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:to-18}Turnout in 2018 Midterm Election", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/to18_chart.rds")
-print(j)
-```
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- *were* more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
+In Table \ref{tab:to-18-logit-short}, *Days Since Discharged* indicates the number of days between a parolee's discharge date and the 2018 election.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout.
 
-## Individual-Level Turnout Regressions {-}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being "assigned" to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, Table \ref{tab:to-18-logit-short} indicates that this relationship breaks down in the short-run.
-These logistic models include individuals last discharged from parole between January 1^st^, 2017, and May 18^th^, 2018.
 \begin{singlespace}
@@
 
-stargazer(date_to_model_notime, date_to_model_time,
+stargazer(date_to_model_time,
           header = F,
-          type = "latex", notes.align = "l",
+          type = "text", notes.align = "l",
           covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
+                               "D(Voted in 2008)",
                                "D(Male)", "Age (Years)",
@@
           out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(date_to_model_notime, type = "HC0"))),
-            sqrt(diag(vcovHC(date_to_model_time, type = "HC0"))))
-          )
+          se = date_ses)
 
@@
 
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
 
@@
 t <- readRDS("./temp/demos_rd_time.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of \"Control\" and \"intent-to-Treat\" Groups", linesep = "",
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Individuals Discharged Before and After EO 181", linesep = "",
                   align = c("l", "c", "c"), "latex") %>% 
@@
 \end{singlespace}
+
 In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
 The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
-\begin{singlespace}
-```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
 
-stargazer(model1, model2, model3,
-          header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit} Individual-Level Logit Models",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))))
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table4.tex"}
-\end{singlespace}
 Model 1 in Table \ref{tab:to-18-logit} controls only for whether an individual was discharged after Executive Order 181 went into effect.

### 2020-07-06 markdown/part2/registration_regs.Rmd
substantial re-write

@@
 ```
-# Turnout Among Parolees in 2016 {-}
+# Treatment Effect on Registration {-}
 
-In the Instrumental Variables Approach section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18^th^ -- October 14^th^, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1^st^, 2015, and October 14^th^, 2016.
+In the body of this manuscript, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote.
+The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occurred when the individual was no longer on parole.
+Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.
+
+In this Supplemental Appendix, I re-estimate the primary analysis found in Table 4 and 5in the main body of this manuscript.
+Here, however, I use registration as the dependent variable, rather than turnout in 2018.
+
 \begin{singlespace}
-```{r logit-reg-chunk-16, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_16.rdata")
+```{r race-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/app_reg.rdata")
 
-stargazer(model1, model2, model3,
+stargazer(model1, model2,
           header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Discharged on or after May 18\\textsuperscript{th}, 2016)", 
-                               "D(Male)", "Age (Years)", "Counts in Most Recent Sentence",
-                               "Time Spent on Parole (Years)"),
-          dep.var.labels = c("Cast Ballot in 2016 Election"),
-          title = "\\label{tab:to-16-logit} Individual-Level Logit Model",
+          type = "text", notes.align = "l",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election",
+                               "Black $\\times$ D(Discharged After EO 181)"),
+          dep.var.labels = c("Registered to Vote"),
+          title = "\\label{tab:racevar} ITT By Race",
           table.placement = "H",
           omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X"),
-                         c("Felony Class FE", "", "", "X")),
-          table.layout = "-dcm#-t-a-s-n",
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
+                         c("Felony Class FE", "", "", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
           out = "./temp/test.tex",
+          notes = "TO REPLACE",
           out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(model1, type = "HC0"))),
-            sqrt(diag(vcovHC(model2, type = "HC0"))),
-            sqrt(diag(vcovHC(model3, type = "HC0")))
-          ))
+          se = list(m1_ses$V1, m2_ses$V1))
+
 j <- fread("./temp/test.tex", header = F, sep = "+")
 
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
 
@@
 
-write.table(j, "./temp/table55.tex", quote = F, col.names = F,
+write.table(j, "./temp/tablexxx.tex", quote = F, col.names = F,
             row.names = F)
+
 ```
-\input{"../../temp/table55.tex"}
-\end{singlespace}
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
-
-# Treatment Effect on Registration {-}
-
-In the body of this paper, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote.
-The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occured when the individual was no longer on parole.
-Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.
-
-In this Supplemental Appendix, I re-estimate the primary analyses found in Tables 5, 6, and 8 in the main body of this paper.
-Here, however, I use registration as the dependent variable, rather than turnout in 2018.
-As before, the dependent variable is binary: an individual is either registered or not registered.
-As such, the functional forms remain the same.
-As before, the marginal effects of the probit models are presented with other covariates held at their means.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_reg.tex"}
+\input{"../../temp/tablexxx.tex"}
 \end{singlespace}
 
-The 2SLS model indicates that voting rights restoration increased the registration rate of former parolees by 1.37 percentange points, while the biprobit model estimates that the effect was between 1.25 and 1.57 percentage points.
-Given that 5.5 percent of parolees in the control group were registered to vote, this implies an increase of between 23 and 29 percent.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_race_reg.tex"}
-\end{singlespace}
-
-As with turnout, voting rights restoration appears to have increased registrations only among white former parolees.
-The estimated treatment effect is between 2.3 and 2.4 percentage points, off a registration rate among the control, white individuals of 4.4 percent.
-No significant effect is detected among non-white or Black former parolees, whose control groups had base registration rates of 5.9 and 7.2 percent, respectively.
-As with turnout, Black former parolees had much higher base registration rates than whites.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_months_reg.tex"}
-\end{singlespace}
-
-Finally, we see that the longer an individual was on parole after her voting rights were restored, the more likely she was to be registered.
-Each additional month on parole after voting rights were restored is associated with an increase in registration of between 0.5 and 0.9 percentage points.
+As with turnout, there is a positive ITT of around 1 percentage point.
+However, as Model 2 makes clear, this was not so for Black voters, whose registration was apparently undisturbed by the treatment.

### 2020-07-06 markdown/part2/rewrite.Rmd
substantial re-write

--- 2020-07-06 markdown/part2/part2_standalone.Rmd
@@
 link-citations: yes
-fontsize: 12pt
+fontsize: 10pt
 header-includes:
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10^th^, 2018 (the registration deadline for 2018).
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for 2018).
 I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
@@
 
-In Table \ref{tab:to-18-logit-short}, *Days Since Discharged* indicates the number of days between a parolee's discharge date and the 2018 election.
+Table \ref{tab:to-18-logit-short} presents an ordinary least squares regression using only the control voters.
+The variable *Days off Parole Before Election* indicates the number of days between a parolee's discharge date and the 2018 election.
 As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout.
@@
           type = "text", notes.align = "l",
-          covariate.labels = c("Days Since Discharged", "Days Since Discharged\\textsuperscript{2}",
+          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
                                "D(Voted in 2008)",
@@
           dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit-short} Individual-Level Logit Models",
+          title = "\\label{tab:to-18-logit-short} 2018 Turnout for Control Group",
           table.placement = "H",
@@
 \end{singlespace}
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.^[The Supplemental Information provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.] This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
-Because discharge date from parole is uncorrelated in the short term with an individual's propensity to vote in the short-run, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
 
-To further demonstrate the usefulness of discharge date as an intent-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intent-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intent-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+To further demonstrate the validity of the interrupted time series model, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the ITT group (those discharged between May 18th and October 12th, 2018) with a difference-of-means t-test.
+With the exception of age, the control and ITT groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
 The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote.
-To the extent our control group perhaps has a marginally higher propensity to vote than our intent-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
+To the extent our control group perhaps has a marginally higher propensity to vote than our ITT group our setup is (slightly) biased against finding a significant increase due to the executive order.
+
 \begin{singlespace}
@@
 t <- readRDS("./temp/demos_rd_time.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Individuals Discharged Before and After EO 181", linesep = "",
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
                   align = c("l", "c", "c"), "latex") %>% 
@@
 
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^*$ Difference is significant at 95 percent confidence level.
-\\\\Shows demographics for individuals discharged from parole \\\\between January 1st, 2017, and October 12th, 2018.}}}"
+note.latex <- ""
 
@@
 
-lab <- "\\caption{\\label{tab:demo-rd} Demographics of \"Control\" and \"intent-to-Treat\" Groups}"
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
@@
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18^th^, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1^st^, 2017, through October 10^th^, 2018 (the registration deadline in New York State).
+Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
+Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect.
+Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, and race.
+Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
+Model 4 replicates Model 3 but uses 2016 characteristics and turnout.
+This "placebo" regression provides a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+Robust standard errors are clustered by control / intention-to-treat status [@Angrist2008].
 
-Model 1 in Table \ref{tab:to-18-logit} controls only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6^th^, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
+Models 1 --- 3 include all individuals discharged from parole between January 1st, 2017, through October 12th, 2018 (the registration deadline in New York State), while Model 4 includes individuals last discharged from parole between January 1, 2015, and October 14th, 2016.
+\begin{singlespace}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/individual_turnout_18_itt.rdata")
+load("./temp/individual_turnout_16.rdata")
+
+stargazer(model1, model2, model3, model4,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:to-18-logit} Turnout in General Election",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
+                         c("Felony Class FE", "", "", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(m1_ses$V1, m2_ses$V1, m3_ses$V1, m4_ses$V1))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table4.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table4.tex"}
+\end{singlespace}
+
  Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-Table \ref{tab:to-18-logit} also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between `r round((exp(model1[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` and `r round((exp(model3[["coefficients"]][["finished_postTRUE"]]) - 1) * 100, digits = 1)` percent.
-The turnout rate for the control group was `r format(round(100*readRDS("./temp/control_to_18.rds"), digits = 1), nsmall = 1)` percent, indicating a treatment effect of `r round((exp(model3[["coefficients"]][["finished_postTRUE"]])-1) * readRDS("./temp/control_to_18.rds"), digits = 4) * 100` percentage points.
+Table \ref{tab:to-18-logit} also indicates that, although being discharged in the summer of 2016 had no impact on 2016 turnout, being discharged after the executive order went into effect in 2018 *was* associated with increased turnout.
+After controlling for other covariates, Model 3 indicates that the intention-to-treat effect was 0.6 percentage points.
+Turnout among the control voters was 2.1 percent, indicating that the executive order raised turnout by roughly 29 percent.
 
-## Instrumental Variables Approach {-}
+The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 4 using different windows.
+At the furthest left, we use a one-month window on either side of the effective date of the executive order, and gradually increase the size of the window.
+The window never extends past the registration deadline in the 2018 and 2016 registration deadlines; the estimates at the far right, therefore, are only extending the window backwards in time.
 
-Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intent-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-As detailed in Figure \ref{fig:flow}, not everyone discharged after the policy changed had their rights restored.
-In other words, not everyone "assigned" to the treatment group was actually treated.
-We therefore have a noncompliance problem.
-
-Two-stage least squares models allow us to leverage the random assignment to treatment to identify the causal effect of rights restoration on the treated population.
-The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall between 0 and 1.
-Although @Angrist2001 and @Angrist2008 argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research [@Gerber2000; @Green2000; @Lassen2004] using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable [@Dong2015].
-The bivariate probit approach [see @Wooldridge2010] is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous [@Terza2008].
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument [@Angrist2001].
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-\begin{singlespace}
-\input{"../../temp/iv_clean.tex"}
-\end{singlespace}
-The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.84 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.96 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between `r round((0.84 / (3.1684 - 0.84)) * 100, digits = 1)` and `r round((0.96 / (3.1684 - 0.96)) * 100, digits = 1)` percent.
-
-Figure \ref{fig:sens} estimates Model 2 from Table \ref{tab:iv-models} while using different windows around the effective date of the executive order.
-The first window includes all individuals discharged between the start of 2017 and the registration deadline.
-The second window limits the control group to individuals discharged on or after June 1, 2017, while the third window limits the control group to individuals discharged in 2018.
-The final window limits the whole population to individuals discharged in the three months on either side of the executive order.
-
-```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 4, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/window_change_plot.rds")
+```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/reestimates.rds")
 print(j)
@@
 
-In three of the four windows, the treatment effect remains significant at the 95 percent level.
-Only in the fourth window, where the sample includes just 30 percent as many observations as the largest window, is the treatment effect nonsignificant.
-Given that the treatment effect remains largely constant across the different windows, this is likely driven by the reduced sample size.
+Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95% level in 13 of the 16 models.
+The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant or negative.
+This means that voters discharged in the summer before an election are perhaps *less* likely to vote --- implying that our estimated effect of the executive order is perhaps somewhat conservative.
 
-### Racial Variation {-}
+## Racial Variation {-}
 
 Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
+Table \ref{tab:racevar} re-estimates Model 3 from Table \ref{tab:to-18-logit}, but now interacts the treatment variable with a dummy variable indicating that a former parolee was Black.
+As before, robust standard errors are clustered by treatment assignment group.
+
 \begin{singlespace}
-\input{"../../temp/iv_clean_race.tex"}
-\end{singlespace}
-```{r nameless-chunk, echo = F, results = "asis", message = F}
-load("./temp/iv_individual_turnout_18.rdata")
-```
+```{r race-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/individual_turnout_18_itt.rdata")
 
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.8 percentage points.
-Given that just 3.5 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
-Turnout among treated non-white and Black former parolees was 3.0 and 3.3 percent, respectively.
-
-Why would the intervention have increased turnout among white individuals and have had no effect on Black individuals?
-Some of this may be explained by different propensities to vote.
-@White2019, for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than white individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying "a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)" (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intent-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latinos are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
-\begin{singlespace}
-```{r logit-black-to-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/black_to.rdata")
-
-stargazer(black_to1, black_to2, black_to1b, black_to2b, black_to3, black_to4,
-          column.separate = c(2, 2, 2),
-          column.labels = c("Discharged Prior to EO", "ITT Group", "Treatment Group"),
+stargazer(model5,
           header = F,
-          type = "latex", notes.align = "l",
-          covariate.labels = c("D(Black)", "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:black-to} Turnout Rate Among Former Parolees",
+          type = "text", notes.align = "l",
+          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election",
+                               "Black $\\times$ D(Discharged After EO 181)"),
+          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:racevar} ITT By Race",
           table.placement = "H",
-          omit.stat = c("f", "ser"),
-          omit = "felony",
-          table.layout = "-dcm#-t-a-s-n",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
+                         c("Felony Class FE", "", "", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
           out = "./temp/test.tex",
+          notes = "TO REPLACE",
           out.header = F,
-          notes = "TO REPLACE",
-          se = list(
-            sqrt(diag(vcovHC(black_to1, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to2, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to1b, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to2b, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to3, type = "HC0"))),
-            sqrt(diag(vcovHC(black_to4, type = "HC0")))
-          ))
+          se = list(m5_ses$V1))
 
@@
 
-note.latex <- "\\multicolumn{7}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.
-\\\\Latino individuals excluded from Models 2, 4, and 6.}}}"
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
 
 j <- j %>% 
-  mutate(n = row_number(),
-         V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
          V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1))
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
 
-insert <- "\\cmidrule(l{3pt}r{3pt}){2-3} \\cmidrule(l{3pt}r{3pt}){4-5}  \\cmidrule(l{3pt}r{3pt}){6-7}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
 
-j <- bind_rows(j, data.frame(V1 = c(insert, insert1), n = c(7.5, 1.1))) %>% 
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
   arrange(n) %>% 
@@
             row.names = F)
+
 ```
@@
 \end{singlespace}
-Of the 14,155 individuals discharged from parole between January 1^st^, 2017, and May 18^th^, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still a third more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
 
-Table \ref{tab:black-to} also shows, however, that among the intent-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
-The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
+The estimated treatment effect for non-Black voters in Table \ref{tab:racevar} is 1.0 percentage points, though the interaction effect --- *Black × D(Discharged After EO 181)* --- is negative 0.8 percentage points.
+The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
+This appears to have had a leveling effect: although Black former parolees in the control group turned out at higher rates than non-Black parolees in that group, the turnout rate among Black and non-Black voters in the ITT group in 2018 was virtually identical, as shown in Table \ref{tab:re}.
 
-### Variable Treatment Intensity {-}
+\begin{singlespace}
+```{r rd-dmeans, echo = F, warning = F}
+k <- readRDS("./temp/means_diff_black.rds")
+j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout by Treatment Group, Race", linesep = "",
+                  align = c("l", "c", "c", "c")) %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
 
-There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figure \ref{fig:to-18-chart} suggests that individuals discharged later in the treatment period had a higher propensity to vote.
-This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
-If the individual did not have voting rights restored, this variable takes the value 0.
-This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.^[Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.]
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- ""
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:re} Turnout by Treatment Group, Race}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table_whatever3.tex"}
+\end{singlespace}
+
+## Instrumental Variables Approach {-}
+
+The analyses in the previous section indicate that executive order was successful at increasing turnout among formerly incarcerated individuals as a group.
+They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
+To answer this question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
+We consider individuals whose rights were in fact restored prior to discharge from parole "compliers."
+
+Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
+To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via treatment [@Angrist1996].
+In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
+The nonsignificance of the coefficients *Days off Parole* in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit} indicate that this condition is met.
+
+Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
+Model 1 includes just the dummy for whether an individual's rights were restored (instrumented by treatment assignment) while Model 2 includes the additional covariates from past specifications.
+As before, robust standard errors are clustered by treatment assignment group.
+
 \begin{singlespace}
-\input{"../../temp/iv_clean_months.tex"}
+```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/iv_individual_turnout_18.rdata")
+
+stargazer(iv1, iv3,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("D(Rights Restored Before Discharge)", "D(Male)", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
+                         c("Felony Class FE", "", "", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(iv1_ses, iv3_ses))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors in parentheses.\\\\Instrument is dummy indicating whether\\\\individual discharged after EO 181 went into effect.\\\\First-stage equations include covariates in Model 2.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table6.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table6.tex"}
 \end{singlespace}
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.29 percentage points (in the 2SLS model) and 0.50 percentage points (in the 2S probit model).
 
-Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
+The two-stage least squares approach indicates that the CACE was roughly 0.9 percentage points.
+Because overall turnout for individuals whose rights were restored was 3.2 percent, this implies that rights restoration prior to discharge from parole increased turnout by 39 percent.
 
@@
 As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least `r round((0.84 / (3.1684 - 0.84)) * 100, digits = 1)` percent.
-This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
-Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
+In the case of New York State, rights restoration prior to discharge from parole is successful at boosting post-supervision participation rates.
+In this project we estimate that overall turnout among formerly disenfranchised individuals increased by about 0.6 percentage points (or 29 percent), and that individuals whose rights were restored saw turnout about 0.9 percentage points (or 39 percent) above where it would have been otherwise.
+Although the absolute impact of the policy change is small, it is meaningful given the low turnout among these individuals.
 
-The mechanism through which the rules change increased turnout among these individuals is not clear.
+The mechanism through which the rules change increased turnout among these individuals is not clear, and deserves futher study.
 @Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
@@
 When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-Of course, even voters in the control group were receiving notification of their eligibilty through the mail.
+Of course, even voters in the control group were receiving notification of their eligibility through the mail.
 If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
@@
 
-Puzzlingly, the causal effect of pre-discharge rights restoration seems to vary based on parolees' race.
-Limiting the analysis to just white parolees reveals that rights restoration increased turnout by more than 90 percent.
-The fact that rights restoration prior to discharge appears to have no such impact on non-white or Black parolees is surprising given the magnitude of the effect for whites.
+Troublingly, the effect of pre-discharge rights restoration seems to vary based on parolees' race.
 This discrepancy could be caused by a number of different factors.

### 2020-07-06 markdown/part2/rewrite.tex
substantial re-write

--- 2020-03-31 markdown/part2/part2_standalone.tex
@@
 \documentclass[
-  12pt,
+  10pt,
 ]{article}
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{March 09, 2020
+\date{July 06, 2020
 \linebreak
@@
 In the spring of 2018,\footnote{Although the executive order was signed on April 18\textsuperscript{th}, it did not go into effect until May 18\textsuperscript{th}.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move was of course good for individuals who were still on parole on election day --- they would have been disenfranchised otherwise.
-The change in policy is also beneficial for felony probationers.
-Despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
-The executive order is a promising step: by changing the policy to allow all New York citizens living in their communities to cast a ballot, it has the potential to both re-enfranchise the nearly 30,000 New Yorkers on parole living in the community and to clarify the rules about who is eligible to vote.
+Such a move of course directly re-enfranchised individuals who were still on parole on election day.
+The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
 
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10\textsuperscript{th}, 2018 (the registration deadline for 2018).
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for 2018).
 I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
@@
 
-\hypertarget{turnout-among-the-formerly-disenfranchised}{%
-\section*{Turnout Among the Formerly Disenfranchised}\label{turnout-among-the-formerly-disenfranchised}}
-\addcontentsline{toc}{section}{Turnout Among the Formerly Disenfranchised}
-
-In a series of papers between 2009 and 2011, researchers developed methods for directly estimating the turnout of formerly disenfranchised individuals.
-Haselswerdt (\protect\hyperlink{ref-Haselswerdt2009}{2009}) matched administrative release data and voter registration data from Erie County, NY, to estimate turnout among a small group of formerly incarcerated individuals.
-Traci Burch expanded upon this matching methodology to estimate the voting patterns of formerly disenfranchised individuals in a range of states (\protect\hyperlink{ref-Burch2012}{2012}, \protect\hyperlink{ref-Burch2011}{2011}).
-She used release data from states' Departments of Corrections and their registered voter files to identify formerly incarcerated individuals who went on to register to vote.
-Using the registered voter files, she estimated the party affiliation of formerly incarcerated individuals (in states with party registration) and their turnout rates.
-Her methodology has been used to investigate other questions surrounding the voting patterns of formerly incarcerated individuals under different circumstances and to examine the impact of changes in disenfranchisement policy (e.g.~Meredith and Morse \protect\hyperlink{ref-Meredith2013}{2013}, \protect\hyperlink{ref-Meredith2015}{2015}).
-
-Much of this literature establishes that formerly incarcerated individuals rarely vote, even when they are no longer formally barred from doing so.
-The causal effect of incarceration on participation is the subject of some debate within the field.
-Individuals who go to prison share many characteristics with lower propensity voters generally.
-Less educated citizens, for instance, turnout at low rates whether they have been to prison or not.
-In an attempt to disentangle sociodemographic characteristics from the experience of imprisonment, Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}) uses administrative data from Pennsylvania to estimate turnout rates prior to and after incarceration.
-They argue that the majority of the low turnout observed among formerly incarcerated individuals can be explained by observable characteristics, concluding that ``it appears that spending time in prison does not have large negative effects on subsequent participation'' (1144).
-
-White (\protect\hyperlink{ref-White2019}{2019}), however, comes to a different conclusion.
-Using administrative court and voter file records from Harris County, Texas, she finds that individuals assigned jail time for misdemeanor offenses are less likely to participate in future elections than similarly-situated misdemeanants who do not go to jail.
-This finding does not necessarily conflict with Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}); as the earlier paper explains, prison often occurs after many other interactions with the criminal justice system.
-By the time an individual is incarcerated, interactions with the criminal justice system may have already reduced his propensity to vote.
-Individuals arrested for misdemeanors, on the other hand, likely reflect a much broader swath of the population --- and, therefore, individuals who may have had fewer interactions with the criminal justice system.
-
 \hypertarget{notification-re-enfranchisement-and-turnout}{%
@@
 
-Regardless of the precise mechanism, the low turnout among formerly incarcerated individuals is cause for concern, particularly given the racialized aspects of the criminal justice system.
-A criminal justice system that inflicts serious consequences on a population that has relatively little political voice is problematic.
-Whether or not incarceration \emph{causes} low turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Whether --- and to what extent --- incarceration causes lower turnout among citizens is the subject of some debate in the literature.
+Gerber et al.
+(\protect\hyperlink{ref-Gerber2017}{2017}), for instance, argues that prison does not materially impact post-incarceration turnout, though White (\protect\hyperlink{ref-White2019}{2019}) finds that jail time can decrease the future turnout of misdemeanants.
+Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, Burch (\protect\hyperlink{ref-Burch2011}{2011})).
+
+Regardless of whether incarceration reduces turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
 Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
@@
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
-When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical.
+When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics.
 The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
+The basic design is therefore an interrupted time series, in which I test whether turnout for individuals discharged after the executive order took effect turned out at higher rates than those discharged before the policy change.
+This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall turnout among formerly incarcerated individuals.
 
-The analysis is complicated, however, by the fact that not all individuals discharged from parole after the policy change --- that is to say, assigned to the treatment group --- were actually treated by having their voting rights restored while on parole in a meeting with their parole officer.
-In other words, our natural experiment has ``imperfect compliance'' (see, for instance, Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Noncitizens, for instance, were not eligible for rights ``restoration'' while on parole because in New York State they were not allowed to vote even prior to their conviction.
-Because not all individuals assigned to the treatment group were in fact treated, simply comparing turnout between the groups discharged on either side of the cutpoint reveals not the treatment effect, but rather the intent-to-treat effect.
-Put differently, it allows us to estimate how much the executive order increased turnout among former parolees as a group --- including those ineligible to vote --- but not the individual-level effect of the policy change on propensity to vote.
-
-Although we can identify the individuals in the assigned-to-treatment group who do not comply with their treatment status, we do not know who among the assigned-to-control group would have gone untreated if they had been assigned to the other group.
-We cannot identify, for instance, noncitizens who were discharged from parole before the policy change.
-In an ideal world, we would limit both the assigned-to-treatment \emph{and} the assigned-to-control groups to voting-eligible individuals; we do not live in such a world.
-Worse yet, there is an obvious relationship between compliance (rights restoration) and the dependent variable of interest: noncitizens are both ineligible to have voting rights bestowed and cannot vote.
-
-The standard way of dealing with imperfect compliance in an experimental setting is to leverage the (random) assignment to treatment as an instrument for the (nonrandom) actual compliance (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-Such an approach has been widely used in the context of voter turnout (e.g.~Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-Gerber and Green (\protect\hyperlink{ref-Gerber2000}{2000}) provides a helpful example of why this two-step process is necessary.
-The authors test the efficacy of phone calls on increasing turnout.
-One group is assigned to be the control; the authors do not attempt to call them.
-Another set of voters receive a phone call, but only some of them are actually contacted, or ``treated'' with the call.
-The authors could simply test to see whether those contacted by phone turned out at higher rates --- but what if high propensity voters are more likely to answer the phone?
-Then ``treatment'' would simply identify a group of individuals with higher vote propensity than the control group without actually speaking to the efficacy of the treatment.
-
-The same process is likely at play here.
-The control group includes individuals that \emph{would have} had their rights restored at an in-person meeting if they had been discharged later; it also includes individuals who would not have received the treatment thanks to their citizenship status.
-By comparing turnout among the actually treated to the control group, we may simply be comparing the turnout of a group of \emph{citizens} to the turnout of a group that includes both citizens \emph{and} noncitizens.
-Although a positive ``treatment effect'' is likely in this context, it would be unable to speak to causality.
-The instrumental variables approach allows us to mitigate this problem.
-Moreover, a very small number of individuals in the control group are marked as having their rights restored prior to the effective date of the executive order.
-While this likely reflects paperwork errors, I nonetheless treat them as a second set of non-compliers.
-
-Figure \ref{fig:flow} demonstrates the buckets into which former parolees can fall.
-Although I argue that a parolee's assignment to the control or treatment categories is random and uncorrelated with turnout except via the treatment, whether a treated former parolee is treated or is a noncomplier is not random and is associated with their propensity to vote through mechanisms other than the treatment.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{../../temp/flowchart}
-
-}
-
-\caption{\label{fig:flow}Categorization of Former Parolees}\label{fig:flow}
-\end{figure}
-
-Using the natural experiment with imperfect compliance setup, I begin by estimating the intent-to-treat effect of the policy change.
-This approach simply asks whether, as a whole, turnout was higher among individuals discharged after the policy change than those discharged earlier.
-I use turnout in the 2018 election as the dependent variable in these logistic regressions.
-
-After estimating the intent-to-treat effect of the policy change, I implement the instrumental variables approach to estimate the individual-level effect of in-person voting rights restoration on subsequent turnout.
-As discussed above, the human nature of the treatment allows for discrepancies in how the treatment is administered.
-To explore this potential, I also test whether the treatment had different effects for different subgroups of the formerly incarcerated population.
-I finally ask whether the treatment effects were larger for individuals who spent more time on parole after having their rights restored.
+As discussed above, not all formerly incarcerated individuals were eligible to have their rights restored; the ITT effect will therefore be biased towards zero, insofar as the behavior of ineligible parolees remains unchanged.
+After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
+In this setup, rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
+This instrumentation process is a common method for identifying the CACE (see, for instance, Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
 
@@
 
-Before analyzing turnout in the 2018 midterms, I begin by examining turnout in the 2016 election.
-It is possible that individuals discharged from parole shortly before a federal election are more likely to cast a ballot than individuals discharged earlier, whether or not their voting rights were restored.
-However, as Figure \ref{fig:to-16} (which plots 2016 turnout rates by month of parole discharge) makes clear, individuals discharged from parole in the final months before the 2016 presidential election were not substantially more likely to cast a ballot in the election than individuals discharged earlier.
-The longer an individual has been off of parole, the more likely he is to cast a ballot.
-For instance, of the individuals last discharged from parole in 2010, 6.8 percent cast a ballot in the 2016 election, while just 4.3 percent of those last discharged from parole in 2015 did so.\footnote{Figure \ref{fig:to-16} plots individuals' turnout by the last date of discharge from parole.
-Therefore, individuals discharged from parole in 2010 who reoffended and were discharged from parole again in 2015 are included only in 2015.} A quadratic curve is fitted (weighted by the number of individuals discharged each month), along with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2010 through May 2016, and extended through October 2016.
+Before estimating the interrupted time series model, we must establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
+It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote, regardless of whether their voting rights are restored before they finish their sentence.
 
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-16-chart-1}
-
-}
-
-\caption{\label{fig:to-16}Turnout in 2016 Presidential Election}\label{fig:to-16-chart}
-\end{figure}
-
-Figure \ref{fig:to-18} plots month of parole discharge and turnout in the 2018 midterm elections.
-Once again, a weighted quadratic curve is fitted with a 95 percent confidence band.
-This curve is fit on monthly data running from January 2012 through May 2018, and extended through October 2018.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{part2_standalone_files/figure-latex/to-18-chart-1}
-
-}
-
-\caption{\label{fig:to-18}Turnout in 2018 Midterm Election}\label{fig:to-18-chart}
-\end{figure}
-
-Figure \ref{fig:to-16} does not indicate that individuals who were discharged from parole shortly before the 2016 presidential election were more likely to cast a ballot than individuals discharged earlier in the year.
-Figure \ref{fig:to-18}, on the other hand, indicates that New Yorkers discharged from parole in the months leading up to the 2018 election --- many of whom had their rights restored while they were still on parole --- \emph{were} more likely to participate than those discharged earlier in the year.
-However, Figures \ref{fig:to-16} and \ref{fig:to-18} are noisy and do not prove that the executive order increased turnout.
-
-\hypertarget{individual-level-turnout-regressions}{%
-\subsection*{Individual-Level Turnout Regressions}\label{individual-level-turnout-regressions}}
-\addcontentsline{toc}{subsection}{Individual-Level Turnout Regressions}
-
-When considered over a multi-year period, the enactment of Executive Order 181 cannot be understood as a natural experiment.
-The longer an individual has been off of parole, the more likely she is to cast a ballot, but only individuals recently discharged from parole were eligible to have their voting rights restored prior to discharge.
-For a true natural experiment to hold, an individual's probability of being ``assigned'' to treatment (here, discharged from parole after the executive order went into effect) must be uncorrelated with the outcome of interest (propensity to vote).
-Figures \ref{fig:to-16} and \ref{fig:to-18} indicate that this is not the case when considering individuals discharged from parole over multiple years.
-
-However, Table \ref{tab:to-18-logit-short} indicates that this relationship breaks down in the short-run.
-These logistic models include individuals last discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018.
+Table \ref{tab:to-18-logit-short} presents an ordinary least squares regression using only the control voters.
+The variable \emph{Days off Parole Before Election} indicates the number of days between a parolee's discharge date and the 2018 election.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout.
 
@@
 
-The inclusion of time controls in Model 2 in Table \ref{tab:to-18-logit-short} leaves the AIC unchanged.
-A Chi-squared test confirms that the model is not improved when controls for time are included.
-When we look only at individuals recently discharged from parole, the length of time an individual has been off parole is not associated with his propensity to vote.\footnote{The Supplemental Information provides further corroboration that being discharged from parole in the months before an election is uncorrelated with propensity to vote by exploring turnout rates in the 2016 presidential election.} This indicates that using individuals released in 2017, and 2018 prior to the executive order, as a control group is feasible.
-Because discharge date from parole is uncorrelated in the short term with an individual's propensity to vote in the short-run, any observed difference in turnout between individuals discharged before and after late May can therefore be attributed to the executive order.
-
-To further demonstrate the usefulness of discharge date as an intent-to-treat indicator, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the intent-to-treat group (those discharged between May 18th and October 12th, 2018).
-With the exception of age, the control and intent-to-treat groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
+To further demonstrate the validity of the interrupted time series model, Table \ref{tab:demo-rd} shows the demographic characteristics of individuals in the control group (those discharged between January 1st, 2017, and May 17th, 2018) and the ITT group (those discharged between May 18th and October 12th, 2018) with a difference-of-means t-test.
+With the exception of age, the control and ITT groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
 The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote.
-To the extent our control group perhaps has a marginally higher propensity to vote than our intent-to-treat group our setup is (slightly) biased against finding a significant increase due to the executive order.
+To the extent our control group perhaps has a marginally higher propensity to vote than our ITT group our setup is (slightly) biased against finding a significant increase due to the executive order.
 
@@
 
-In Table \ref{tab:to-18-logit}, I present the results of an individual-level logistic regression exploring whether individuals who were discharged on or after May 18\textsuperscript{th}, 2018, turned out at higher rates than those discharged earlier.
-The models include all individuals discharged from parole between January 1\textsuperscript{st}, 2017, through October 10\textsuperscript{th}, 2018 (the registration deadline in New York State).
+Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
+Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect.
+Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, and race.
+Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
+Model 4 replicates Model 3 but uses 2016 characteristics and turnout.
+This ``placebo'' regression provides a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+Robust standard errors are clustered by control / intention-to-treat status (Angrist and Pischke \protect\hyperlink{ref-Angrist2008}{2008}).
+
+Models 1 --- 3 include all individuals discharged from parole between January 1st, 2017, through October 12th, 2018 (the registration deadline in New York State), while Model 4 includes individuals last discharged from parole between January 1, 2015, and October 14th, 2016.
 
@@
 
-Model 1 in Table \ref{tab:to-18-logit} controls only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6\textsuperscript{th}, 2018, and race.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
 Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-Table \ref{tab:to-18-logit} also indicates that individuals discharged from parole after the executive order went into effect were more likely to cast a ballot than those discharged earlier.
-Exponentiating the coefficients on D(Discharged After EO 181) indicates that Executive Order 181 raised the turnout rate among all formerly disenfranchised individuals by between 26.5 and 35 percent.
-The turnout rate for the control group was 2.1 percent, indicating a treatment effect of 0.74 percentage points.
+Table \ref{tab:to-18-logit} also indicates that, although being discharged in the summer of 2016 had no impact on 2016 turnout, being discharged after the executive order went into effect in 2018 \emph{was} associated with increased turnout.
+After controlling for other covariates, Model 3 indicates that the intention-to-treat effect was 0.6 percentage points.
+Turnout among the control voters was 2.1 percent, indicating that the executive order raised turnout by roughly 29 percent.
 
-\hypertarget{instrumental-variables-approach}{%
-\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
-\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
-
-Table \ref{tab:to-18-logit} indicates that executive order was successful at increasing turnout among all formerly disenfranchised individuals.
-Although this intent-to-treat estimate is important information for policymakers and advocates hoping to increase the political representation of formerly incarcerated individuals as a whole, it does not shed light on the extent to which in-person rights restoration increased the propensity to vote for the individuals who actually received the treatment.
-To answer that question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-As detailed in Figure \ref{fig:flow}, not everyone discharged after the policy changed had their rights restored.
-In other words, not everyone ``assigned'' to the treatment group was actually treated.
-We therefore have a noncompliance problem.
-
-Two-stage least squares models allow us to leverage the random assignment to treatment to identify the causal effect of rights restoration on the treated population.
-The nature of our dependent variable (voting), instrumented variable (rights restoration), and instrument (discharge from parole before or after EO 181 went into effect) pose a challenge: each are binary values that can either be equal to 0 or 1.
-Linear models such as two-stage least squares do not ensure that the predicted probability of voting will fall between 0 and 1.
-Although Angrist (\protect\hyperlink{ref-Angrist2001}{2001}) and Angrist and Pischke (\protect\hyperlink{ref-Angrist2008}{2008}) argue that in practice this limitation is usually trivial, the possibility remains that the linear model results in unacceptable misspecification.
-Some research (Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Green and Shachar \protect\hyperlink{ref-Green2000}{2000}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}) using instrumental variables in the context of a binary vote-no vote framework has employed a two-stage probit (2S probit) model to avoid the constraints of the nonparametric framework.
-The 2S probit model specification, however, works best when the instrumented variable is continuous --- not, as in the case of rights restoration, a dichotomous variable (Dong and Lewbel \protect\hyperlink{ref-Dong2015}{2015}).
-The bivariate probit approach (see Wooldridge \protect\hyperlink{ref-Wooldridge2010}{2010}) is well suited for situations where the dependent variable, the instrumented variable, and instrument are all dichotomous (Terza, Bradford, and Dismuke \protect\hyperlink{ref-Terza2008}{2008}).
-This is especially true when the models exclude covariates and include only the dependent variable, the endogenous instrumented variable, and the instrument (Angrist \protect\hyperlink{ref-Angrist2001}{2001}).
-
-Table \ref{tab:iv-models} presents the results of these various approaches on the question at hand.
-Models 1 and 2 utilize the linear two-stage least squares approach, with and without covariates.
-Model 3 uses the 2S probit specification (with all covariates), while Models 4 and 5 employ the bivariate probit specification (with and without covariates).
-For ease of comparison, I show the marginal effects of the probit models (measured at the means of the other variables).
-
-\begin{singlespace}
-\input{"../../temp/iv_clean.tex"}
-\end{singlespace}
-
-The estimates of the bivariate probit model are modestly more conservative than the two-stage least squares model.
-After controlling for available demographics, the bivariate probit model indicates that rights restoration prior to discharge boosted turnout by individuals who had their rights restored by around 0.84 percentage points; the two-stage least squares model estimates that it increased turnout by around 0.96 percentage points.
-Though these numbers may seem small, they represent relatively large gains.
-Just 3.2 percent of individuals who had their rights restored cast a ballot in 2018, indicating that the executive order increased turnout by between 36.1 and 43.5 percent.
-
-Figure \ref{fig:sens} estimates Model 2 from Table \ref{tab:iv-models} while using different windows around the effective date of the executive order.
-The first window includes all individuals discharged between the start of 2017 and the registration deadline.
-The second window limits the control group to individuals discharged on or after June 1, 2017, while the third window limits the control group to individuals discharged in 2018.
-The final window limits the whole population to individuals discharged in the three months on either side of the executive order.
+The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 4 using different windows.
+At the furthest left, we use a one-month window on either side of the effective date of the executive order, and gradually increase the size of the window.
+The window never extends past the registration deadline in the 2018 and 2016 registration deadlines; the estimates at the far right, therefore, are only extending the window backwards in time.
 
@@
 
-{\centering \includegraphics{part2_standalone_files/figure-latex/sensitivity-1}
+{\centering \includegraphics{rewrite_files/figure-latex/sensitivity-1}
 
@@
 
-In three of the four windows, the treatment effect remains significant at the 95 percent level.
-Only in the fourth window, where the sample includes just 30 percent as many observations as the largest window, is the treatment effect nonsignificant.
-Given that the treatment effect remains largely constant across the different windows, this is likely driven by the reduced sample size.
+Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95\% level in 13 of the 16 models.
+The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant or negative.
+This means that voters discharged in the summer before an election are perhaps \emph{less} likely to vote --- implying that our estimated effect of the executive order is perhaps somewhat conservative.
 
 \hypertarget{racial-variation}{%
-\subsubsection*{Racial Variation}\label{racial-variation}}
-\addcontentsline{toc}{subsubsection}{Racial Variation}
+\subsection*{Racial Variation}\label{racial-variation}}
+\addcontentsline{toc}{subsection}{Racial Variation}
 
 Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-In Table \ref{tab:iv-models-r} I present the two-stage least squares and bivariate probit models on subsets of former parolees.
-These models include only white former parolees (Models 1 and 4), all non-white former parolees (Models 2 and 5), and only Black former parolees (Models 3 and 6).
-Once again, the marginal effects in the bivariate probit models are calculated at the means of the other variables.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_race.tex"}
-\end{singlespace}
-
-Table \ref{tab:iv-models-r} indicates that rights restoration prior to discharge boosted turnout among white individuals by around 1.8 percentage points.
-Given that just 3.5 percent of white former parolees with restored voting rights cast a ballot, this means that rights restoration \emph{doubled} these individuals' likelihood of casting a ballot.
-We cannot determine whether rights restoration had an effect on non-white individuals: the coefficient on these estimates are small and not statistically significant.
-Turnout among treated non-white and Black former parolees was 3.0 and 3.3 percent, respectively.
-
-Why would the intervention have increased turnout among white individuals and have had no effect on Black individuals?
-Some of this may be explained by different propensities to vote.
-White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that brief periods of incarceration decreases Black individuals' propensity to vote by substantially more than white individuals.
-She demonstrates that, prior to incarceration, Black individuals were more likely to vote, identifying ``a narrative in which targeted policing brings many black defendants into court, including some voters (so they can be deterred), while lower arrest rates among whites mean that the white defendant pool rarely includes voters (so there is little demobilization, because the people jailed were unlikely to vote anyway)'' (321).
-The inverse may hold true here: if Black individuals released from parole have a higher natural propensity to vote (even after accounting for potentially larger depressive effects from incarceration), they may be less susceptible to a policy intervention of this sort.
-
-There is reason to believe this may be the case.
-Table \ref{tab:black-to} presents a series of logistic models estimating turnout in 2018 for the control group (individuals discharged from parole before the executive order went into effect), the intent-to-treat group (those discharged after the executive order), and of the treatment group (individuals whose rights were restored), testing whether Black individuals turned out at higher rates.
-Because Latinos are more likely to be non-citizens (and therefore have a lower turnout rate than Black individuals), these individuals have been excluded from Models 2, 4, and 6 in case their inclusion overstates the relative propensity of Black individuals to vote.
+Table \ref{tab:racevar} re-estimates Model 3 from Table \ref{tab:to-18-logit}, but now interacts the treatment variable with a dummy variable indicating that a former parolee was Black.
+As before, robust standard errors are clustered by treatment assignment group.
 
@@
 
-Of the 14,155 individuals discharged from parole between January 1\textsuperscript{st}, 2017, and May 18\textsuperscript{th}, 2018, just over 6 thousand (43 percent) were Black.
-These Black voters were more than 50 percent more likely to vote in the midterms than the rest of the population.
-Even when Latinos are excluded from this group, Black individuals were still a third more likely to cast a ballot.
-In both specifications, this higher turnout rate is significant at the 95 percent level of confidence.
-
-Table \ref{tab:black-to} also shows, however, that among the intent-to-treat group (those discharged from parole after the executive order went into effect), Black individuals were no more likely to participate in the 2018 general election.
-The same is true for the treatment group --- Black individuals who had their rights restored were no more likely than other individuals to vote in 2018.
-The intervention, therefore, appears to have a leveling effect.
-Rights restoration prior to discharge seems to have no effect on future propensity to vote for Black individuals, but that does not mean they are less likely to vote than other former parolees.
-
-\hypertarget{variable-treatment-intensity}{%
-\subsubsection*{Variable Treatment Intensity}\label{variable-treatment-intensity}}
-\addcontentsline{toc}{subsubsection}{Variable Treatment Intensity}
-
-There is some reason to believe that assuming a constant treatment effect understates the true impact of Executive Order 181.
-Figure \ref{fig:to-18-chart} suggests that individuals discharged later in the treatment period had a higher propensity to vote.
-This may be due to a number of factors: for instance, parole officers may have had longer to understand and communicate the new rules.
-Similarly, individuals discharged later may have had more meetings with their parole officers after the policy change, giving the parole officers multiple times to encourage the individuals under study to cast a ballot.
-Table \ref{tab:iv-models-time} investigates turnout as a function of the number of months an individual spent on parole after having her rights restored.
-If the individual did not have voting rights restored, this variable takes the value 0.
-This variable is instrumented by the number of months an individual spent on parole after the executive order went into effect.
-It is coded as 0 for individuals discharged prior to the implementation of the executive order.\footnote{Because the bivariate probit specification is not appropriate when the endogenous variable is not binary, that specification is not included in Table \ref{tab:iv-models-time}.}
+The estimated treatment effect for non-Black voters in Table \ref{tab:racevar} is 1.0 percentage points, though the interaction effect --- \emph{Black × D(Discharged After EO 181)} --- is negative 0.8 percentage points.
+The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
+This appears to have had a leveling effect: although Black former parolees in the control group turned out at higher rates than non-Black parolees in that group, the turnout rate among Black and non-Black voters in the ITT group in 2018 was virtually identical, as shown in Table \ref{tab:re}.
 
 \begin{singlespace}
-\input{"../../temp/iv_clean_months.tex"}
+
+\input{"../../temp/table_whatever3.tex"}
 \end{singlespace}
 
-Table \ref{tab:iv-models-time} indicates that variable treatment intensity matters --- individuals who spent longer on parole after they had their rights restored were more likely to vote than those whose rights were restored shortly before discharge.
-Table \ref{tab:iv-models-time} shows that for each month individuals spent on parole after having their rights restored, turnout increased by between 0.29 percentage points (in the 2SLS model) and 0.50 percentage points (in the 2S probit model).
+\hypertarget{instrumental-variables-approach}{%
+\subsection*{Instrumental Variables Approach}\label{instrumental-variables-approach}}
+\addcontentsline{toc}{subsection}{Instrumental Variables Approach}
 
-Of course, with just a handful of months between the implementation of the executive order and the registration deadline for the 2018 midterms, it is impossible to know the full impact more months on parole after having one's rights restored has.
-One thing, however, is clear: focusing only on the individuals discharged from parole shortly after Executive Order 181 went into effect very likely underestimates the true impact it will have on parolee's propensity to vote when parolees spend more time under supervision with their rights formally restored.
+The analyses in the previous section indicate that executive order was successful at increasing turnout among formerly incarcerated individuals as a group.
+They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
+To answer this question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
+We consider individuals whose rights were in fact restored prior to discharge from parole ``compliers.''
+
+Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
+To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via treatment (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
+In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
+The nonsignificance of the coefficients \emph{Days off Parole} in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit} indicate that this condition is met.
+
+Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
+Model 1 includes just the dummy for whether an individual's rights were restored (instrumented by treatment assignment) while Model 2 includes the additional covariates from past specifications.
+As before, robust standard errors are clustered by treatment assignment group.
+
+\begin{singlespace}
+
+\input{"../../temp/table6.tex"}
+\end{singlespace}
+
+The two-stage least squares approach indicates that the CACE was roughly 0.9 percentage points.
+Because overall turnout for individuals whose rights were restored was 3.2 percent, this implies that rights restoration prior to discharge from parole increased turnout by 39 percent.
 
@@
 As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is broadly successful at boosting post-supervision participation rates.
-In this project I estimate that rights restoration increased individuals' propensity to vote by at least 36.1 percent.
-This estimate, however, should be taken as a lower bound: the data indicate that individuals who spent longer on parole after having their rights restored were more likely to vote than individuals who were discharged immediately following the implementation of the Executive Order.
-Notably, this upward trend showed no signs of leveling off prior to the 2018 election; individuals whose rights were restored that were discharged in the fifth month after implementation voted at higher rates than those discharged after three months; those discharged after three months turned out at higher rates than those discharged in the first month.
+In the case of New York State, rights restoration prior to discharge from parole is successful at boosting post-supervision participation rates.
+In this project we estimate that overall turnout among formerly disenfranchised individuals increased by about 0.6 percentage points (or 29 percent), and that individuals whose rights were restored saw turnout about 0.9 percentage points (or 39 percent) above where it would have been otherwise.
+Although the absolute impact of the policy change is small, it is meaningful given the low turnout among these individuals.
 
-The mechanism through which the rules change increased turnout among these individuals is not clear.
+The mechanism through which the rules change increased turnout among these individuals is not clear, and deserves futher study.
 Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
@@
 \begin{cslreferences}
-\leavevmode\hypertarget{ref-Angrist2001}{}%
-Angrist, Joshua D.
-2001.
-``Estimation of Limited Dependent Variable Models with Dummy Endogenous Regressors.'' \emph{Journal of Business \& Economic Statistics} 19 (1): 2--28.
-\url{https://doi.org/10.1198/07350010152472571}.
-
 \leavevmode\hypertarget{ref-Angrist1996}{}%
@@
 
-\leavevmode\hypertarget{ref-Burch2012}{}%
----------.
-2012.
-``Did Disfranchisement Laws Help Elect President Bush?
-New Evidence on the Turnout Rates and Candidate Preferences of Florida's Ex-Felons.'' \emph{Political Behavior} 34 (1): 1--26.
-\url{https://doi.org/10.1007/s11109-010-9150-9}.
-
 \leavevmode\hypertarget{ref-Butler2011}{}%
@@
 \url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
-
-\leavevmode\hypertarget{ref-Dong2015}{}%
-Dong, Yingying, and Arthur Lewbel.
-2015.
-``A Simple Estimator for Binary Choice Models with Endogenous Regressors.'' \emph{Econometric Reviews} 34 (1-2): 82--105.
-\url{https://doi.org/10.1080/07474938.2014.944470}.
 
@@
 \url{https://doi.org/10.1086/692670}.
-
-\leavevmode\hypertarget{ref-Green2000}{}%
-Green, Donald P., and Ron Shachar.
-2000.
-``Habit Formation and Political Behaviour: Evidence of Consuetude in Voter Turnout.'' \emph{British Journal of Political Science} 30 (4): 561--73.
-\url{https://www.jstor.org/stable/194285}.
-
-\leavevmode\hypertarget{ref-Haselswerdt2009}{}%
-Haselswerdt, Michael V.
-2009.
-``Con Job: An Estimate of Ex-Felon Voter Turnout Using Document-Based Data*.'' \emph{Social Science Quarterly} 90 (2): 262--73.
-\url{https://doi.org/10.1111/j.1540-6237.2009.00616.x}.
 
@@
 
-\leavevmode\hypertarget{ref-Terza2008}{}%
-Terza, Joseph V., W.
-David Bradford, and Clara E.
-Dismuke.
-2008.
-``The Use of Linear Instrumental Variables Methods in Health Services Research and Health Economics: A Cautionary Note.'' \emph{Health Services Research} 43 (3): 1102--20.
-\url{https://doi.org/10.1111/j.1475-6773.2007.00807.x}.
-
 \leavevmode\hypertarget{ref-sentencing_2016}{}%
@@
 \url{https://doi.org/10.1017/S0003055414000562}.
-
-\leavevmode\hypertarget{ref-Wooldridge2010}{}%
-Wooldridge, Jeffrey M. 2010. \emph{Econometric Analysis of Cross Section and Panel Data}. 2nd ed. Cambridge, Mass: MIT Press.
 \end{cslreferences}

### 2020-07-06 code/part_2/06_individual_turnout.R

@@ -14,13 +14,32 @@ parolees <- left_join(parolees, nys_roll, by = "din") %>%
          parole_time = as.numeric((parole_status_date - release_date_parole) / 365.25),
          v2018 = ifelse(is.na(v2018), 0, v2018),
          finished_post = parole_status_date >= "2018-05-18",
-         days_since_done = as.numeric(as.Date("2018-11-06") - parole_status_date),
-         days2 = days_since_done^2,
+         days_bef = as.Date("2018-10-12") - parole_status_date,
+         days2 = as.integer(days_bef)^2,
          restored = ifelse(is.na(restored), F, restored),
          days_since_m21 = parole_status_date - as.Date("2018-05-19"),
          v2016 = ifelse(is.na(v2016), 0, v2016),
          v2008 = ifelse(is.na(v2008), 0, v2008),
          registered = voter_status %in% c("ACTIVE", "INACTIVE"))
+###########################################
+pool <- filter(parolees, year(parole_status_date) >= 2017) %>% 
+  mutate(b = race == "BLACK")
+
+h <- t.test(v2018 ~ b, filter(pool, !finished_post))
+j <- t.test(v2018 ~ b, filter(pool, finished_post))
+
+k <- pool %>% 
+  group_by(b, finished_post) %>% 
+  summarize(turnout = percent(mean(v2018), accuracy = .01)) %>% 
+  pivot_wider(id_cols = "finished_post", values_from = "turnout", names_from = "b") %>% 
+  mutate(finished_post = ifelse(finished_post, "ITT", "Control"))
+
+k$p <- round(c(h$p.value, j$p.value), digits = 3)
+
+colnames(k) <- c("Group", "Non-Black", "Black", "p-value")
+saveRDS(k, "./temp/means_diff_black.rds")
+
+####################
 
 how_many_restored <- sum(parolees$restored)
 saveRDS(how_many_restored, "./temp/how_many_restored.rds")
@@ -44,39 +63,67 @@ save(model1, model2, model3, file = "./temp/individual_turnout_18.rdata")
 
 ## intend-to-treat models
 
-model1 <- glm(v2018 ~ finished_post,
-              family = "binomial", data = filter(parolees, year(parole_status_date) >= 2017))
+model1 <- lm(v2018 ~ finished_post,
+              data = filter(parolees, year(parole_status_date) >= 2017),
+              cluster = "finished_post")
 
-model2 <- glm(v2018 ~ finished_post + 
[…]

### 2020-07-06 code/part_2/09_demo_compare_rd.R

@@ -53,15 +53,15 @@ demos <- left_join(demos, rename_v)
 
 
 pvals <- unlist(lapply(demos$varnames, function(x){
-  t.test(demos1[, x] ~ demos1$finished_post)$p.value < 0.05
+  t.test(demos1[, x] ~ demos1$finished_post)$p.value
 }))
 
 demos$pvals <- pvals
 
 demos <- demos %>% 
-  mutate(names = ifelse(pvals, paste0(names, "*"), as.character(names))) %>% 
-  select(names, finished_before, finished_after)
+  mutate(pvals = round(pvals, digits = 3)) %>% 
+  select(names, finished_before, finished_after, pvals)
 
-colnames(demos) <- c("Variable", "Control", "Intend-to-Treat")
+colnames(demos) <- c("Variable", "Control", "Intend-to-Treat", "p-value")
 
 saveRDS(demos, "./temp/demos_rd_time.rds")

### 2020-07-06 code/part_2/11_registrations_regs.R

@@ -1,3 +1,74 @@
+
+#### individual level regressions
+
+## this is the treated group
+nys_roll <- readRDS("./temp/parolee_to_18.rds")
+parolees <- readRDS("./temp/parolees_with_restoration.rds")
+
+parolees <- left_join(parolees, nys_roll, by = "din") %>% 
+  select(-history, -year, -election_type) %>% 
+  filter(parole_status == "DISCHARGED",
+         parole_status_date >= "2012-01-01",
+         parole_status_date <= "2018-10-12",
+         !is.na(dob_parole)) %>% 
+  mutate(age = as.numeric((as.Date("2018-11-06") - dob_parole) / 365.25),
+         parole_time = as.numeric((parole_status_date - release_date_parole) / 365.25),
+         v2018 = ifelse(is.na(v2018), 0, v2018),
+         finished_post = parole_status_date >= "2018-05-18",
+         days_bef = as.Date("2018-10-12") - parole_status_date,
+         days2 = as.integer(days_bef)^2,
+         restored = ifelse(is.na(restored), F, restored),
+         days_since_m21 = parole_status_date - as.Date("2018-05-19"),
+         v2016 = ifelse(is.na(v2016), 0, v2016),
+         v2008 = ifelse(is.na(v2008), 0, v2008),
+         registered = voter_status %in% c("ACTIVE", "INACTIVE"))
+
+
+
+
+model1 <- lm(registered ~ finished_post + as.factor(race) +
+               as.factor(sex) + age + v2008 +
+               felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef,
+             data = filter(parolees, year(parole_status_date) >= 2017),
+             cluster = "finished_post")
+
+m1_ses <- data.table(summary(lm.cluster(registered ~ finished_post + as.factor(race) +
+                                          as.factor(sex) + age + v2008 +
+                                          felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef,
+                                        data = filter(parolees, year(parole_status_date) >= 2017),
+                                        cluster = "finished_post"))[,2])
+
+parolees$pb <- (parolees$race == "BLACK") * (parolees$finished_post == T)
+
+model2 <- lm(registered ~ finished_post + as.factor(race) +
+               as.factor(sex) + age + v2008 +
+               felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef + pb,
+             data = filter(parolees, year(parole_status_date) >= 2017),
[…]

### 2020-07-07 markdown/part2/registration_regs.tex
write
ssq

@@
 
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
+\newpage
 \doublespace
 
-\hypertarget{turnout-among-parolees-in-2016}{%
-\section*{Turnout Among Parolees in 2016}\label{turnout-among-parolees-in-2016}}
-\addcontentsline{toc}{section}{Turnout Among Parolees in 2016}
+\begin{figure}[H]
 
-In the Instrumental Variables Approach section of this paper, I argue that being discharged from parole in the final months leading up to an election is uncorrelated with propensity to vote.
-Table \ref{tab:to-16-logit} demonstrates that individuals discharged between May 18\textsuperscript{th} -- October 14\textsuperscript{th}, 2016, did not participate at different rates in the 2016 presidential election than other formerly incarcerated individuals.
-Table \ref{tab:to-16-logit} includes all individuals last discharged from parole between January 1\textsuperscript{st}, 2015, and October 14\textsuperscript{th}, 2016.
+{\centering \includegraphics{registration_regs_files/figure-latex/sensitivity-1}
 
-\begin{singlespace}
+}
 
-\input{"../../temp/table55.tex"}
-\end{singlespace}
-
-Although turnout was generally higher in 2016 than in 2018 (reflecting statewide higher turnout thanks to the presidential contest), there is no evidence that being discharged in the summer of 2016 was associated with an individual's propensity to cast a ballot.
-The nonsignificant results from 2016 provide strong corroboration that discharge date serves as an effective instrument for rights restoration.
-
-\hypertarget{treatment-effect-on-registration}{%
-\section*{Treatment Effect on Registration}\label{treatment-effect-on-registration}}
-\addcontentsline{toc}{section}{Treatment Effect on Registration}
-
-In the body of this paper, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote.
-The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occured when the individual was no longer on parole.
-Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.
-
-In this Supplemental Appendix, I re-estimate the primary analyses found in Tables 5, 6, and 8 in the main body of this paper.
-Here, however, I use registration as the dependent variable, rather than turnout in 2018.
-As before, the dependent variable is binary: an individual is either registered or not registered.
-As such, the functional forms remain the same.
-As before, the marginal effects of the probit models are presented with other covariates held at their means.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_reg.tex"}
-\end{singlespace}
-
-The 2SLS model indicates that voting rights restoration increased the registration rate of former parolees by 1.37 percentange points, while the biprobit model estimates that the effect was between 1.25 and 1.57 percentage points.
-Given that 5.5 percent of parolees in the control group were registered to vote, this implies an increase of between 23 and 29 percent.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_race_reg.tex"}
-\end{singlespace}
-
-As with turnout, voting rights restoration appears to have increased registrations only among white former parolees.
-The estimated treatment effect is between 2.3 and 2.4 percentage points, off a registration rate among the control, white individuals of 4.4 percent.
-No significant effect is detected among non-white or Black former parolees, whose control groups had base registration rates of 5.9 and 7.2 percent, respectively.
-As with turnout, Black former parolees had much higher base registration rates than whites.
-
-\begin{singlespace}
-\input{"../../temp/iv_clean_months_reg.tex"}
-\end{singlespace}
-
-Finally, we see that the longer an individual was on parole after her voting rights were restored, the more likely she was to be registered.
-Each additional month on parole after voting rights were restored is associated with an increase in registration of between 0.5 and 0.9 percentage points.
+\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
+\end{figure}
 

### 2020-07-07 markdown/part2/rewrite.Rmd
ssq

@@
 subtitle: "Voting Rights Restoration and Post-Supervision Participation"
-thanks: Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
-The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
-All errors are my responsibility.
+thanks: The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.
 author: Kevin Morris^[Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (kevin.morris@nyu.edu)]
@@
   `r format(Sys.time(), '%B %d, %Y')`
-  \linebreak
-  \linebreak
-  \linebreak
-  \linebreak
-  Word Count: 9,596
 output:
@@
 link-citations: yes
-fontsize: 10pt
+fontsize: 12pt
+fontface: Times New Roman
 header-includes:
@@
 In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
-I demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
+By leveraging randomness in parole discharge date, we interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
+We demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
 This group-level effect, however, masks race-specific effects.
-Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
+Although rights restoration prior to parole discharge increased turnout by non-Black former parolees by 1.4 percentage points, the treatment effect was substantially smaller for Black former parolees.
 This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans’ view of the democratic process.
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
+Iowa stands alone at the other end of the spectrum, where all felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the impact of Amendment 4 in Florida.]
+Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the potential impact of Amendment 4 in Florida.]
 
 There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future [@White2019; but see @Gerber2017].
-As @Burch2011 and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
+Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future [@Weaver2010; @White2019], though other research casts some doubt on this conclusion [@Gerber2017].
+One thing is clear: As @Burch2011 and others have shown, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to reduce the representation of minority and marginalized voters.
 The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
@@
 
-In the spring of 2018,^[Although the executive order was signed on April 18^th^, it did not go into effect until May 18^th^.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move of course directly re-enfranchised individuals who were still on parole on election day.
+In the spring of 2018,^[Although the executive order was signed on April 18th, it did not go into effect until May 18th.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+The policy change of course directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
 The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+By re-enfranchising anyone not actively incarcerated, the rules were substantially simplified.
 
@@
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that "upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration" (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[Link: https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Beginning in 2010, New York State began including information about voter eligibility in parole discharge packets, a change that @Meredith2013 indicates did not increase turnout.
+Executive Order 181 both changed the timing of rights restoration (upon release from prison, not the end of parole) and the manner of restoration (via an in-person meeting, rather than in a discharge packet).
+This study tests the bundled effect of these changes.
+
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[See https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
 Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for 2018).
-I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
+In the analysis below, we examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election).
+We exclude individuals who were still on parole and could only vote because of the policy change in order to isolate the effect of the timing and manner of rights restoration for individuals whose legal eligibility to vote was not changed.
 
-By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility through the mail, Executive Order 181 offers us the opportunity to build on existing research.
-Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
-Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration?
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility via brochures in a parole discharge packet, Executive Order 181 offers us the opportunity to build on existing research.
+Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state [@Lerman2014]?
+Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration [see, for instance, @White2015]?
 
@@
 Whether --- and to what extent --- incarceration causes lower turnout among citizens is the subject of some debate in the literature.
-@Gerber2017, for instance, argues that prison does not materially impact post-incarceration turnout, though @White2019 finds that jail time can decrease the future turnout of misdemeanants.
-Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, @Burch2011).
+Although @Gerber2017 argues that prison does not materially impact post-incarceration turnout, @White2019 finds that jail time can decrease the future turnout of misdemeanants.
+Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates [see, for instance, @Burch2011].
+Regardless of whether incarceration reduces turnout, the  state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
 
-Regardless of whether incarceration reduces turnout, the  state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
+@Meredith2013 examines the direct effect of changing rights restoration notification policies in New Mexico, New York, and North Carolina.
+In the case of New Mexico, formerly disenfranchised individuals began automatically receiving a certificate attesting to their eligibility to vote, a certificate they were previously required to affirmatively request.
+In New York and North Carolina, parole discharge packets began including explicit information about eligibility to vote.
+@Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
+In the three states studied in @Meredith2013, however, newly re-enfranchised individuals received this information as part of a discharge packet.
+Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform former parolees that their voting rights had been restored.
+Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out], and it is possible that these brochures, in-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
 
-Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out].
-Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
 @Meredith2015 examines the impact of ending permanent disenfranchisement in Iowa.
@@
 @Meredith2015 leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
-@Meredith2013, on the other hand, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-@Meredith2013 finds no effect on turnout from notification in the absence of eligibility changes.
 
 @Gerber2015 built on the quasi-experimental design of @Meredith2015, conducting a field experiment in Connecticut in advance of the 2012 presidential election.
-In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
+In their field experiment, some formerly disenfranchised (but eligible) residents were reminded of their eligibility; others were not.
 Like @Meredith2015, they find evidence that these reminders successfully increased turnout.
@@
 
-@Meredith2015 and @Gerber2015 provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
-Nevertheless, the research to date has examined the effect of mail notification on post-supervision turnout.
-In the case of New York State, formerly incarcerated individuals were already informed of their voting rights through the mail when they finished their sentence prior to the executive order.
-In other words, the treatment identified in past research is here incorporated into the status quo, or control, condition.
-This study therefore tests not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
+This research provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
+While information included in discharge packets seems ineffective at boosting turnout, there is some evidence that notification delivered on its own is successful.
+Nevertheless, the research to date has examined the effect of rights restoration notification outside a personal relationship on post-supervision turnout.
+It seems likely that an in-person meeting will be more effective at dispelling confusion about eligibility than a brochure included in a discharge packet.
 
@@
 "It may also be," they write, "that social benefits [arising from non-criminal justice contact with the state] are less visible to citizens because they often occur through the mail, rather than through personal contact with government agencies or officials, making it easier to disconnect social benefits from government and the political system" (93).
-Because Executive Order 181 operates *not* through the mail but rather through "personal contact with government...
+Because Executive Order 181 operates *not* through the mail (or discharge packet) but rather through "personal contact with government...
 officials," its effects may differ from the mail notification programs.
@@
 
-I expect that restoring voting rights while an individual is still on parole increases that individual's later propensity to vote substantially.
-This increase will operate through two mechanisms: as the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a former parolee more sure of her voting rights than a letter in the mail.
+We expect that restoring voting rights while an individual is still on parole increases that individual's later propensity to vote.
+This increase will operate through two mechanisms: as the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a former parolee more sure of her voting rights than a brochure.
 Secondly, the interpersonal nature of the parolee-parole officer relationship is expected to more effectively repair (some of) the damage done to the former parolees interpretation of the state.
@@
 Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
+By using the identification number provided from the parolee public records request and this website, we identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
 Roughly 1 percent of former parolees are not in the online lookup.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
@@
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from March 3^rd^, 2019.
+We use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
+We match individuals in each dataset using first name, middle name, last name, and date of birth.
 To be considered a "match," records must have the exact same birth date.
@@
 I adopt the test developed in @Meredith2013 to test the prevalence of false positives.
-I slightly alter the date of birth reported in the parole discharge dataset to create false records.
+We slightly alter the date of birth reported in the parole discharge dataset to create false records.
 Comparing the number of matches between these "fake" discharge records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when we construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
 This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
-\begin{singlespace}
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
-true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
-plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
 
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+TABLE 1 ABOUT HERE
 
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-\end{singlespace}
 Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+If an individual marries and changes her name after being discharged from parole, for instance, we will not identify her using this matching methodology.
 Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
+To reduce the likelihood of these false negatives we remove all punctuation from all names, and standardize capitalization.
 A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
@@
 
-In this paper, I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+In this paper, we exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
-When I limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics.
+When we limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics.
 The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
-The basic design is therefore an interrupted time series, in which I test whether turnout for individuals discharged after the executive order took effect turned out at higher rates than those discharged before the policy change.
+The basic design is therefore an interrupted time series, in which we test whether individuals discharged after the executive order took effect turned out at higher rates than those discharged before the policy change.
 This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall turnout among formerly incarcerated individuals.
@@
 This instrumentation process is a common method for identifying the CACE [see, for instance, @Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
-
-This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
-To that end I use turnout --— not registration —-- as the dependent variable of interest throughout.
-The Supplemental Information demonstrates that all treatment effects discussed here (including for subgroups within the population) also hold when the dependent variable is registration, not turnout.
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.^[This manuscript is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
+To that end we use turnout --— not registration —-- as the dependent variable of interest throughout.
+However, replacing "turnout" with "registration" in the models presented herein demonstrates that Executive Order 181 increased registration among former parolees.]
 
@@
 
-\begin{singlespace}
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
-load("./temp/short_term_to.rdata")
-
-stargazer(date_to_model_time,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
-                               "D(Voted in 2008)",
-                               "D(Male)", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit-short} 2018 Turnout for Control Group",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
-                         c("Felony Class FE", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = date_ses)
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table3.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table3.tex"}
-\end{singlespace}
+TABLE 2 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r rd-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_rd_time.rds")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- ""
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever2.tex"}
-\end{singlespace}
+TABLE 3 ABOUT HERE
 
 Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
-Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, and race.
+Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect (*Discharged After EO 181*).
+Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, race, and whether they voted in the 2008 general election.
 Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Model 4 replicates Model 3 but uses 2016 characteristics and turnout.
-This "placebo" regression provides a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
-Robust standard errors are clustered by control / intention-to-treat status [@Angrist2008].
+Model 4 interacts the ITT indicator with a dummy variable specifying whether a former parolee was Black; this allows us to test whether the ITT differed for Black and non-Black individuals.
 
-Models 1 --- 3 include all individuals discharged from parole between January 1st, 2017, through October 12th, 2018 (the registration deadline in New York State), while Model 4 includes individuals last discharged from parole between January 1, 2015, and October 14th, 2016.
-\begin{singlespace}
-```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
-load("./temp/individual_turnout_16.rdata")
+In Model 5 we estimate a "placebo" regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+Model 5 includes all individuals last discharged between January 1, 2015, and October 14th, 2016.
+The dummy *Discharged After EO 181* measures whether an individual was discharged after May 18, 2016, and the dependent variable is turnout in the 2016 general election.
 
-stargazer(model1, model2, model3, model4,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
-          title = "\\label{tab:to-18-logit} Turnout in General Election",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
-                         c("Felony Class FE", "", "", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(m1_ses$V1, m2_ses$V1, m3_ses$V1, m4_ses$V1))
-j <- fread("./temp/test.tex", header = F, sep = "+")
+In each model, robust standard errors are clustered by control / intention-to-treat status
 
-note.latex <- "\\multicolumn{5}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+TABLE 4 ABOUT HERE
 
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
+The coefficients on *Discharged After EO 181* in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed turned out at higher rates than those discharged earlier.
+As expected, the coefficient is nonsignificant in the placebo regression, implying that individuals discharged shortly before an election are not generally more likely to vote.
+After controlling for other covariates, Model 3 estimates an ITT effect of 0.6 percentage points.
+Turnout among the control voters was 2.1 percent, meaning the executive order raised turnout by roughly 29 percent.
 
-insert1 <- "\\fontsize{10}{12}\\selectfont"
+Table \ref{tab:to-18-logit} also makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table4.tex"}
-\end{singlespace}
-
- Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
-
-Table \ref{tab:to-18-logit} also indicates that, although being discharged in the summer of 2016 had no impact on 2016 turnout, being discharged after the executive order went into effect in 2018 *was* associated with increased turnout.
-After controlling for other covariates, Model 3 indicates that the intention-to-treat effect was 0.6 percentage points.
-Turnout among the control voters was 2.1 percent, indicating that the executive order raised turnout by roughly 29 percent.
-
-The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
-To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 4 using different windows.
-At the furthest left, we use a one-month window on either side of the effective date of the executive order, and gradually increase the size of the window.
-The window never extends past the registration deadline in the 2018 and 2016 registration deadlines; the estimates at the far right, therefore, are only extending the window backwards in time.
-
-```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/reestimates.rds")
-print(j)
-```
-
-Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95% level in 13 of the 16 models.
-The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant or negative.
-This means that voters discharged in the summer before an election are perhaps *less* likely to vote --- implying that our estimated effect of the executive order is perhaps somewhat conservative.
-
-## Racial Variation {-}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-Table \ref{tab:racevar} re-estimates Model 3 from Table \ref{tab:to-18-logit}, but now interacts the treatment variable with a dummy variable indicating that a former parolee was Black.
-As before, robust standard errors are clustered by treatment assignment group.
-
-\begin{singlespace}
-```{r race-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
-
-stargazer(model5,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election",
-                               "Black $\\times$ D(Discharged After EO 181)"),
-          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
-          title = "\\label{tab:racevar} ITT By Race",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
-                         c("Felony Class FE", "", "", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(m5_ses$V1))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table5.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table5.tex"}
-\end{singlespace}
-
-The estimated treatment effect for non-Black voters in Table \ref{tab:racevar} is 1.0 percentage points, though the interaction effect --- *Black × D(Discharged After EO 181)* --- is negative 0.8 percentage points.
+Model 4 estimates that the ITT for non-Black voters was 1.0 percentage points, though the interaction effect --- *Black × Discharged After EO 181* --- is negative 0.8 percentage points.
 The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
@@
 
-\begin{singlespace}
-```{r rd-dmeans, echo = F, warning = F}
-k <- readRDS("./temp/means_diff_black.rds")
-j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout by Treatment Group, Race", linesep = "",
-                  align = c("l", "c", "c", "c")) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
+TABLE 5 ABOUT HERE
 
-write(j, "./temp/tempo.tex")
+## Sensitivity Analysis {-}
 
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
+The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 using different windows.
+In the most conservative window, we include only parolees discharged one month before or after the executive order went into effect (and the same period in 2016).
+As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further distance between parole discharge dates for control and ITT observations.^[The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.]
 
-note.latex <- ""
+FIGURE 1 ABOUT HERE
 
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:re} Turnout by Treatment Group, Race}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever3.tex"}
-\end{singlespace}
+Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95% level in 13 of the 16 models.
+The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
+This means that voters discharged in the summer before an election are perhaps *less* likely to vote absent treatment.
+This implies our estimated effect of the executive order is perhaps somewhat conservative.
 
@@
 Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
-We consider individuals whose rights were in fact restored prior to discharge from parole "compliers."
+We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as "compliers."
 
 Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
-To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via treatment [@Angrist1996].
+To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via the treatment [@Angrist1996].
 In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
-The nonsignificance of the coefficients *Days off Parole* in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit} indicate that this condition is met.
+The nonsignificance of the coefficients *Days off Parole* in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit}, and the general nonsignificance of the placebo regressions in Figure \ref{fig:sens} indicate that this condition is met.
 
 Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
-Model 1 includes just the dummy for whether an individual's rights were restored (instrumented by treatment assignment) while Model 2 includes the additional covariates from past specifications.
+Model 1 includes just *Rights Restored Before Discharge* (instrumented by *Discharged After EO 181*) while Model 2 includes the additional covariates from past specifications.
+In Model 3, we test whether the CACE was different for Black former parolees by interacting *Rights Restored Before Discharge* with *Black* (this endogenous variable is instrumented by *Black × Discharged After EO 181*).
 As before, robust standard errors are clustered by treatment assignment group.
 
-\begin{singlespace}
-```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/iv_individual_turnout_18.rdata")
-
-stargazer(iv1, iv3,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("D(Rights Restored Before Discharge)", "D(Male)", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
-          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
-                         c("Felony Class FE", "", "", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(iv1_ses, iv3_ses))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{3}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.\\\\Instrument is dummy indicating whether\\\\individual discharged after EO 181 went into effect.\\\\First-stage equations include covariates in Model 2.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table6.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table6.tex"}
-\end{singlespace}
+TABLE 6 ABOUT HERE
 
@@
 Because overall turnout for individuals whose rights were restored was 3.2 percent, this implies that rights restoration prior to discharge from parole increased turnout by 39 percent.
+Once again, however, we see that the treatment effect differed sharply by race: the CACE for non-Black former parolees was 1.4 percentage points, but just 0.3 percentage points for Black former parolees.
 
@@
 That changed in 2018.
-On October 12^th^ --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
+On October 12th --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
 Without the executive order, every single one of these individuals would have been barred from participating.
@@
 As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
-In the case of New York State, rights restoration prior to discharge from parole is successful at boosting post-supervision participation rates.
+In the case of New York State, rights restoration prior to discharge from parole was successful at boosting post-supervision participation rates.
 In this project we estimate that overall turnout among formerly disenfranchised individuals increased by about 0.6 percentage points (or 29 percent), and that individuals whose rights were restored saw turnout about 0.9 percentage points (or 39 percent) above where it would have been otherwise.
@@
 
-The mechanism through which the rules change increased turnout among these individuals is not clear, and deserves futher study.
-@Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
-Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by @Weaver2010.
-
-It could also be a story of better information.
-As @Meredith2015 and @Gerber2015 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
-Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-Of course, even voters in the control group were receiving notification of their eligibility through the mail.
-If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+The mechanism through which the rules change increased turnout among these individuals is not clear.
+As discussed above, the literature indicates that this could be because of the relational nature of the restoration process here [@Lerman2014] or because of better information communication of eligibility [@Meredith2015; @Gerber2015].
 In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
+Future work should attempt to "unbundle" this treatment to understand the relative importance of the constituent mechanisms.
 
@@
 # References {-}
+<div id="refs"></div>
+
+\newpage
+
+\begin{singlespace}
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
+true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
+plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
+load("./temp/short_term_to.rdata")
+
+stargazer(date_to_model_time,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
+                               "Voted in 2008",
+                               "Male", "Age (Years)",
+                               "Years on Parole"),
+          dep.var.labels = c("Cast Ballot in 2018 Election"),
+          title = "\\label{tab:to-18-logit-short} 2018 Turnout for Control Group",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
+                         c("Felony Class FE", "X", "X")),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = date_ses)
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table3.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table3.tex"}
+```{r rd-demo-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_rd_time.rds")
+colnames(t) <- c("Variable", "Control", "ITT", "p-value")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- ""
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table_whatever2.tex"}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/individual_turnout_18_itt.rdata")
+load("./temp/individual_turnout_16.rdata")
+
+stargazer(model1, model2, model3, model5, model4,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Discharged After EO 181",
+                               "Black $\\times$ Discharged After EO 181",
+                               "Male", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:to-18-logit} Turnout in General Election",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X", "X"),
+                         c("Felony Class FE", "", "", "X", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          order = c("finished_post", "pb"),
+          se = list(m1_ses$V1, m2_ses$V1, m3_ses$V1, m5_ses$V1, m4_ses$V1))
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{6}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table4.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table4.tex"}
+```{r rd-dmeans, echo = F, warning = F}
+k <- readRDS("./temp/means_diff_black.rds")
+j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout by Treatment Group, Race", linesep = "",
+                  align = c("l", "c", "c", "c")) %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- ""
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:re} Turnout by Treatment Group, Race}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table_whatever3.tex"}
+```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/iv_individual_turnout_18.rdata")
+
+stargazer(iv1, iv3, iv4,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Rights Restored Before Discharge", 
+                               "Black $\\times$ Rights Restored Before Discharge",
+                               "Male", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
+                         c("Felony Class FE", "", "X", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(iv1_ses, iv3_ses, iv4_ses))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors in parentheses.\\\\First-stage equations include covariates in Models 2 and 3.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table6.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table6.tex"}
+\end{singlespace}
+
+\newpage
+```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/reestimates.rds")
+print(j)
+```

### 2020-07-07 markdown/part2/rewrite.tex
ssq

@@
 \documentclass[
-  10pt,
+  12pt,
 ]{article}
@@
 
-\title{Welcome Home --- Now Vote!\thanks{Prepared for the 2020 Annual Meeting of the Midwest Political Science Association.
-The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project.
-All errors are my responsibility.}}
+\title{Welcome Home --- Now Vote!\thanks{The author thanks Jacob Faber, Jeff Manza, Myrna Pérez, Ariel White, and Peter Miller for their comments on this project. All errors are my responsibility.}}
 \usepackage{etoolbox}
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{July 06, 2020
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-Word Count: 9,596}
+\date{July 07, 2020}
 
@@
 In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
-I demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
+By leveraging randomness in parole discharge date, we interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
+We demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
 This group-level effect, however, masks race-specific effects.
-Although rights restoration prior to parole discharge effectively doubled turnout among white former parolees, it had no measurable effect on the turnout of non-white former parolees.
+Although rights restoration prior to parole discharge increased turnout by non-Black former parolees by 1.4 percentage points, the treatment effect was substantially smaller for Black former parolees.
 This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans' view of the democratic process.
@@
 In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-Iowa stands alone at the other end of the spectrum, where felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Justice \protect\hyperlink{ref-bcj_laws}{2019}).
+Iowa stands alone at the other end of the spectrum, where all felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2019}).
 This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
@@
 Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the impact of Amendment 4 in Florida.}
+Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the potential impact of Amendment 4 in Florida.}
 
 There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
-As previous literature has established, interactions with the criminal justice system leaves residents less likely to vote in the future (White \protect\hyperlink{ref-White2019}{2019}; but see Gerber et al.
+Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future (Weaver and Lerman \protect\hyperlink{ref-Weaver2010}{2010}; White \protect\hyperlink{ref-White2019}{2019}), though other research casts some doubt on this conclusion (Gerber et al.
 \protect\hyperlink{ref-Gerber2017}{2017}).
-As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, moreover, turnout rates among the formerly incarcerated are extremely low.
-Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to disenfranchise minority and marginalized voters.
+One thing is clear: As Burch (\protect\hyperlink{ref-Burch2011}{2011}) and others have shown, turnout rates among the formerly incarcerated are extremely low.
+Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to reduce the representation of minority and marginalized voters.
 The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
@@
 
-In the spring of 2018,\footnote{Although the executive order was signed on April 18\textsuperscript{th}, it did not go into effect until May 18\textsuperscript{th}.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-Such a move of course directly re-enfranchised individuals who were still on parole on election day.
+In the spring of 2018,\footnote{Although the executive order was signed on April 18th, it did not go into effect until May 18th.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+The policy change of course directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
 The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+By re-enfranchising anyone not actively incarcerated, the rules were substantially simplified.
 
@@
 Prior to the policy change, formerly incarcerated individuals had their voting rights restored automatically upon the completion of their parole term.
-New York's correction code provides no more guidance other than that ``upon a person's discharge from community supervision, the department shall notify such person of his or her right to vote and provide such person with a form of application for voter registration'' (Section 75 of the New York State Correction Law).
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{Link: \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Beginning in 2010, New York State began including information about voter eligibility in parole discharge packets, a change that Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) indicates did not increase turnout.
+Executive Order 181 both changed the timing of rights restoration (upon release from prison, not the end of parole) and the manner of restoration (via an in-person meeting, rather than in a discharge packet).
+This study tests the bundled effect of these changes.
+
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{See \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
 Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
@@
 
-In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for 2018).
-I exclude individuals who were still on parole and could only vote because of the policy change in order to unpack how the mechanisms of rights restoration can shape turnout.
+In the analysis below, we examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election).
+We exclude individuals who were still on parole and could only vote because of the policy change in order to isolate the effect of the timing and manner of rights restoration for individuals whose legal eligibility to vote was not changed.
 
-By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility through the mail, Executive Order 181 offers us the opportunity to build on existing research.
-Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state?
-Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration?
+By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility via brochures in a parole discharge packet, Executive Order 181 offers us the opportunity to build on existing research.
+Do these human interactions repair damage done to the formerly incarcerated individual's relationship to the state (Lerman and Weaver \protect\hyperlink{ref-Lerman2014}{2014})?
+Or do parole officers exhibit biases against their stewards, making them ineffective or uneven conduits for restoration (see, for instance, White, Nathan, and Faller \protect\hyperlink{ref-White2015}{2015})?
 
@@
 Whether --- and to what extent --- incarceration causes lower turnout among citizens is the subject of some debate in the literature.
-Gerber et al.
-(\protect\hyperlink{ref-Gerber2017}{2017}), for instance, argues that prison does not materially impact post-incarceration turnout, though White (\protect\hyperlink{ref-White2019}{2019}) finds that jail time can decrease the future turnout of misdemeanants.
-Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, Burch (\protect\hyperlink{ref-Burch2011}{2011})).
+Although Gerber et al.
+(\protect\hyperlink{ref-Gerber2017}{2017}) argues that prison does not materially impact post-incarceration turnout, White (\protect\hyperlink{ref-White2019}{2019}) finds that jail time can decrease the future turnout of misdemeanants.
+Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, Burch \protect\hyperlink{ref-Burch2011}{2011}).
+Regardless of whether incarceration reduces turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
+Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
 
-Regardless of whether incarceration reduces turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) examines the direct effect of changing rights restoration notification policies in New Mexico, New York, and North Carolina.
+In the case of New Mexico, formerly disenfranchised individuals began automatically receiving a certificate attesting to their eligibility to vote, a certificate they were previously required to affirmatively request.
+In New York and North Carolina, parole discharge packets began including explicit information about eligibility to vote.
+Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
+In the three states studied in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, newly re-enfranchised individuals received this information as part of a discharge packet.
+Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform former parolees that their voting rights had been restored.
+Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2008}), and it is possible that these brochures, in-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
 
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2008}).
-Some research indicates that dispelling misinformation boosts turnout among the formerly disenfranchised.
 Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) examines the impact of ending permanent disenfranchisement in Iowa.
@@
 Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) leverages differences in notification to test whether the notification increased turnout, finding a strong positive effect.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), on the other hand, examines states where so-called notification laws went into effect.
-Although rules about eligibility did not change in these states, new policies required Departments of Corrections to notify formerly disenfranchised individuals of their re-instated voting rights.
-Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) finds no effect on turnout from notification in the absence of eligibility changes.
 
@@
 (\protect\hyperlink{ref-Gerber2015}{2015}) built on the quasi-experimental design of Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}), conducting a field experiment in Connecticut in advance of the 2012 presidential election.
-In their field experiment, some formerly convicted (but eligible) residents were reminded of their eligibility; others were not.
+In their field experiment, some formerly disenfranchised (but eligible) residents were reminded of their eligibility; others were not.
 Like Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}), they find evidence that these reminders successfully increased turnout.
@@
 
-Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2015}{2015}) provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
-Nevertheless, the research to date has examined the effect of mail notification on post-supervision turnout.
-In the case of New York State, formerly incarcerated individuals were already informed of their voting rights through the mail when they finished their sentence prior to the executive order.
-In other words, the treatment identified in past research is here incorporated into the status quo, or control, condition.
-This study therefore tests not the efficacy of notification laws but rather whether in-person restoration prior to discharge increases turnout above-and-beyond any increase associated with mail notification.
+This research provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
+While information included in discharge packets seems ineffective at boosting turnout, there is some evidence that notification delivered on its own is successful.
+Nevertheless, the research to date has examined the effect of rights restoration notification outside a personal relationship on post-supervision turnout.
+It seems likely that an in-person meeting will be more effective at dispelling confusion about eligibility than a brochure included in a discharge packet.
 
@@
 ``It may also be,'' they write, ``that social benefits {[}arising from non-criminal justice contact with the state{]} are less visible to citizens because they often occur through the mail, rather than through personal contact with government agencies or officials, making it easier to disconnect social benefits from government and the political system'' (93).
-Because Executive Order 181 operates \emph{not} through the mail but rather through ``personal contact with government\ldots{} officials,'' its effects may differ from the mail notification programs.
+Because Executive Order 181 operates \emph{not} through the mail (or discharge packet) but rather through ``personal contact with government\ldots{} officials,'' its effects may differ from the mail notification programs.
 The social benefits (restoration of voting rights) might be communicated more effectively and in ways that improve the formerly incarcerated individuals' opinions of their democratic state, thus increasing turnout further.
@@
 Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, I was able to identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
+By using the identification number provided from the parolee public records request and this website, we identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
 Roughly 1 percent of former parolees are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
@@
 The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-I use a snapshot of the registered voter file from March 3\textsuperscript{rd}, 2019.
+We use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
-I match individuals in each dataset using first name, middle name, last name, and date of birth.
+We match individuals in each dataset using first name, middle name, last name, and date of birth.
 To be considered a ``match,'' records must have the exact same birth date.
@@
 I adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
-I slightly alter the date of birth reported in the parole discharge dataset to create false records.
+We slightly alter the date of birth reported in the parole discharge dataset to create false records.
 Comparing the number of matches between these ``fake'' discharge records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when we construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
 This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
 
-\begin{singlespace}
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 69,644\\
-Birthdate + 35 Days & 502\\
-Birthdate - 35 Days & 426\\
-\bottomrule
-\end{tabular}
-\end{table}
-\end{singlespace}
+TABLE 1 ABOUT HERE
 
 Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using my matching methodology.
+If an individual marries and changes her name after being discharged from parole, for instance, we will not identify her using this matching methodology.
 Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
+To reduce the likelihood of these false negatives we remove all punctuation from all names, and standardize capitalization.
 A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
@@
 This instrumentation process is a common method for identifying the CACE (see, for instance, Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
-
-This paper is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
-To that end I use turnout ----- not registration ----- as the dependent variable of interest throughout.
-The Supplemental Information demonstrates that all treatment effects discussed here (including for subgroups within the population) also hold when the dependent variable is registration, not turnout.
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.\footnote{This manuscript is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
+To that end we use turnout ----- not registration ----- as the dependent variable of interest throughout.
+However, replacing ``turnout'' with ``registration'' in the models presented herein demonstrates that Executive Order 181 increased registration among former parolees.}
 
@@
 
-\begin{singlespace}
-
-\input{"../../temp/table3.tex"}
-\end{singlespace}
+TABLE 2 ABOUT HERE
 
@@
 
-\begin{singlespace}
-
-\input{"../../temp/table_whatever2.tex"}
-\end{singlespace}
+TABLE 3 ABOUT HERE
 
 Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
-Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect.
-Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, and race.
+Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect (\emph{Discharged After EO 181}).
+Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, race, and whether they voted in the 2008 general election.
 Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Model 4 replicates Model 3 but uses 2016 characteristics and turnout.
-This ``placebo'' regression provides a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
-Robust standard errors are clustered by control / intention-to-treat status (Angrist and Pischke \protect\hyperlink{ref-Angrist2008}{2008}).
+Model 4 interacts the ITT indicator with a dummy variable specifying whether a former parolee was Black; this allows us to test whether the ITT differed for Black and non-Black individuals.
 
-Models 1 --- 3 include all individuals discharged from parole between January 1st, 2017, through October 12th, 2018 (the registration deadline in New York State), while Model 4 includes individuals last discharged from parole between January 1, 2015, and October 14th, 2016.
+In Model 5 we estimate a ``placebo'' regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+Model 5 includes all individuals last discharged between January 1, 2015, and October 14th, 2016.
+The dummy \emph{Discharged After EO 181} measures whether an individual was discharged after May 18, 2016, and the dependent variable is turnout in the 2016 general election.
 
-\begin{singlespace}
+In each model, robust standard errors are clustered by control / intention-to-treat status
 
-\input{"../../temp/table4.tex"}
-\end{singlespace}
+TABLE 4 ABOUT HERE
 
-Table \ref{tab:to-18-logit} makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+The coefficients on \emph{Discharged After EO 181} in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed turned out at higher rates than those discharged earlier.
+As expected, the coefficient is nonsignificant in the placebo regression, implying that individuals discharged shortly before an election are not generally more likely to vote.
+After controlling for other covariates, Model 3 estimates an ITT effect of 0.6 percentage points.
+Turnout among the control voters was 2.1 percent, meaning the executive order raised turnout by roughly 29 percent.
 
-Table \ref{tab:to-18-logit} also indicates that, although being discharged in the summer of 2016 had no impact on 2016 turnout, being discharged after the executive order went into effect in 2018 \emph{was} associated with increased turnout.
-After controlling for other covariates, Model 3 indicates that the intention-to-treat effect was 0.6 percentage points.
-Turnout among the control voters was 2.1 percent, indicating that the executive order raised turnout by roughly 29 percent.
+Table \ref{tab:to-18-logit} also makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
 
-The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
-To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 4 using different windows.
-At the furthest left, we use a one-month window on either side of the effective date of the executive order, and gradually increase the size of the window.
-The window never extends past the registration deadline in the 2018 and 2016 registration deadlines; the estimates at the far right, therefore, are only extending the window backwards in time.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{rewrite_files/figure-latex/sensitivity-1}
-
-}
-
-\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
-\end{figure}
-
-Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95\% level in 13 of the 16 models.
-The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant or negative.
-This means that voters discharged in the summer before an election are perhaps \emph{less} likely to vote --- implying that our estimated effect of the executive order is perhaps somewhat conservative.
-
-\hypertarget{racial-variation}{%
-\subsection*{Racial Variation}\label{racial-variation}}
-\addcontentsline{toc}{subsection}{Racial Variation}
-
-Although these models demonstrate that rights restoration had a generally positive effect on participation in the 2018 election, the models hide substantial variation between races.
-Table \ref{tab:racevar} re-estimates Model 3 from Table \ref{tab:to-18-logit}, but now interacts the treatment variable with a dummy variable indicating that a former parolee was Black.
-As before, robust standard errors are clustered by treatment assignment group.
-
-\begin{singlespace}
-
-\input{"../../temp/table5.tex"}
-\end{singlespace}
-
-The estimated treatment effect for non-Black voters in Table \ref{tab:racevar} is 1.0 percentage points, though the interaction effect --- \emph{Black × D(Discharged After EO 181)} --- is negative 0.8 percentage points.
+Model 4 estimates that the ITT for non-Black voters was 1.0 percentage points, though the interaction effect --- \emph{Black × Discharged After EO 181} --- is negative 0.8 percentage points.
 The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
@@
 
-\begin{singlespace}
+TABLE 5 ABOUT HERE
 
-\input{"../../temp/table_whatever3.tex"}
-\end{singlespace}
+\hypertarget{sensitivity-analysis}{%
+\subsection*{Sensitivity Analysis}\label{sensitivity-analysis}}
+\addcontentsline{toc}{subsection}{Sensitivity Analysis}
+
+The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 using different windows.
+In the most conservative window, we include only parolees discharged one month before or after the executive order went into effect (and the same period in 2016).
+As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further distance between parole discharge dates for control and ITT observations.\footnote{The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.}
+
+FIGURE 1 ABOUT HERE
+
+Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95\% level in 13 of the 16 models.
+The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
+This means that voters discharged in the summer before an election are perhaps \emph{less} likely to vote absent treatment.
+This implies our estimated effect of the executive order is perhaps somewhat conservative.
 
@@
 Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
-We consider individuals whose rights were in fact restored prior to discharge from parole ``compliers.''
+We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as ``compliers.''
 
 Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
-To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via treatment (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
+To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via the treatment (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
 In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
-The nonsignificance of the coefficients \emph{Days off Parole} in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit} indicate that this condition is met.
+The nonsignificance of the coefficients \emph{Days off Parole} in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit}, and the general nonsignificance of the placebo regressions in Figure \ref{fig:sens} indicate that this condition is met.
 
 Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
-Model 1 includes just the dummy for whether an individual's rights were restored (instrumented by treatment assignment) while Model 2 includes the additional covariates from past specifications.
+Model 1 includes just \emph{Rights Restored Before Discharge} (instrumented by \emph{Discharged After EO 181}) while Model 2 includes the additional covariates from past specifications.
+In Model 3, we test whether the CACE was different for Black former parolees by interacting \emph{Rights Restored Before Discharge} with \emph{Black} (this endogenous variable is instrumented by \emph{Black × Discharged After EO 181}).
 As before, robust standard errors are clustered by treatment assignment group.
 
-\begin{singlespace}
-
-\input{"../../temp/table6.tex"}
-\end{singlespace}
+TABLE 6 ABOUT HERE
 
@@
 That changed in 2018.
-On October 12\textsuperscript{th} --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
+On October 12th --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
 Without the executive order, every single one of these individuals would have been barred from participating.
@@
 
-The mechanism through which the rules change increased turnout among these individuals is not clear, and deserves futher study.
-Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
-Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
-Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
-
-It could also be a story of better information.
-As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
-(\protect\hyperlink{ref-Gerber2015}{2015}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
-Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-Of course, even voters in the control group were receiving notification of their eligibility through the mail.
-If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+The mechanism through which the rules change increased turnout among these individuals is not clear.
+As discussed above, the literature indicates that this could be because of the relational nature of the restoration process here (Lerman and Weaver \protect\hyperlink{ref-Lerman2014}{2014}) or because of better information communication of eligibility (Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}; Gerber et al.
+\protect\hyperlink{ref-Gerber2015}{2015}).
 In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
+Future work should attempt to ``unbundle'' this treatment to understand the relative importance of the constituent mechanisms.
 
@@
 
-\leavevmode\hypertarget{ref-Angrist2008}{}%
-Angrist, Joshua D., and Jörn-Steffen Pischke. 2008. \emph{Mostly Harmless Econometrics: An Empiricist's Companion}. Princeton University Press. \url{https://doi.org/10.2307/j.ctvcm4j72}.
-
 \leavevmode\hypertarget{ref-Ansolabehere1999}{}%
@@
 \url{https://doi.org/10.1016/S0047-2352(00)00044-1}.
-
-\leavevmode\hypertarget{ref-bcj_laws}{}%
-Justice, Brennan Center for.
-2019.
-``Criminal Disenfranchisement Laws Across the United States.'' May 30, 2019.
-\url{https://www.brennancenter.org/our-work/research-reports/criminal-disenfranchisement-laws-across-united-states}.
 
@@
 
+\newpage
+
+\begin{singlespace}
+\begin{table}[H]
+
+\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
+\centering
+\begin{tabular}[t]{cc}
+\toprule
+Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
+\midrule
+Actual Birthdate & 69,644\\
+Birthdate + 35 Days & 502\\
+Birthdate - 35 Days & 426\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\input{"../../temp/table3.tex"}
+
+\input{"../../temp/table_whatever2.tex"}
+
+\input{"../../temp/table4.tex"}
+
+\input{"../../temp/table_whatever3.tex"}
+
+\input{"../../temp/table6.tex"}
+\end{singlespace}
+
+\newpage
+\begin{figure}[H]
+
+{\centering \includegraphics{rewrite_files/figure-latex/sensitivity-1}
+
+}
+
+\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
+\end{figure}
+
 \end{document}

### 2020-07-07 code/00_initialize.R

@@ -1,3 +1,5 @@
+library(miceadds)
+library(ivpack)
 library(AER)
 library(jtools)
 library(DiagrammeR)

### 2020-07-07 code/part_2/06_individual_turnout.R

@@ -224,8 +224,19 @@ iv3 <- ivreg(v2018 ~ restored + as.factor(race) +
 
 iv3_ses <- cluster.robust.se(iv3, filter(parolees, year(parole_status_date) >= 2017)$finished_post)[,2]
 
+parolees$rw <- with(parolees, (race == "BLACK") * restored == T)
+parolees$pw <- with(parolees, (race == "BLACK") * finished_post == T)
 
-save(iv1, iv3, iv1_ses, iv3_ses, file = "./temp/iv_individual_turnout_18.rdata")
+iv4<- ivreg(v2018 ~ restored + rw + as.factor(race) +
+               as.factor(sex) + age + v2008 +
+               felony_a + felony_b + felony_c + felony_d + felony_e +
+               parole_time + days_bef | . -restored - rw + finished_post + pw,
+             data = filter(parolees, year(parole_status_date) >= 2017))
+
+iv4_ses <- cluster.robust.se(iv4, filter(parolees, year(parole_status_date) >= 2017)$finished_post)[,2]
+stargazer(iv4, se = list(iv4_ses), type = "text")
+
+save(iv1, iv3, iv4, iv1_ses, iv3_ses, iv4_ses, file = "./temp/iv_individual_turnout_18.rdata")
 
 ### variable treatment effect
 stata("

### 2020-07-07 code/part_2/13_window_sensitivity_2.R

@@ -104,13 +104,13 @@ p1 <- ggplot(filter(cints)) +
             fill = "gray60", alpha = 0.5) +
   geom_errorbar(aes(x = months,
                     ymin = lower, ymax = upper), size = 0.2) +
-  geom_line(aes(x = months, y = estimate, linetype = type), size = 0.2) + 
-  geom_point(aes(x = months, y = estimate, shape = type), size = 1.5) + 
+  geom_line(aes(x = months, y = estimate), size = 0.2) + 
+  geom_point(aes(x = months, y = estimate), size = 1.5) + 
   theme_bw() +
-  labs(y = "Estimated Treatment Effect",
+  labs(y = "Estimated ITT Effect",
        x = "Months Around Cutpoint",
        caption = "Notes: 95% confidence bars shown.
-Regressions include covariates displayed in Table 4.
+Regressions include covariates from Models 3 and 5 of Table 4.
 Number of observations in each regression shown on right axis.",
        shape = "Year",
        linetype = "Year") +
@@ -120,9 +120,10 @@ Number of observations in each regression shown on right axis.",
                                          breaks = seq(0, 20000, 10000))) +
   theme(text = element_text(family = "LM Roman 10"),
         plot.caption = element_text(hjust = 0),
-        plot.title = element_text(hjust = 0.5)) +
+        plot.title = element_text(hjust = 0.5),
+        legend.position = NULL) +
   geom_hline(yintercept = 0, linetype = "dashed", size = 1)
-
+p1
 saveRDS(p1, "./temp/reestimates.rds")
 
 ####################################################################################

### 2020-10-09 markdown/part2/rewrite.Rmd
start ssq revisions

@@
 
-In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-Iowa stands alone at the other end of the spectrum, where all felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor [@bcj_laws].
-This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights "for participation in rebellion, or other crime." The definition of "other crime," left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as *Richardson v.
-Ramirez* (1974), has upheld this practice.
-Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the potential impact of Amendment 4 in Florida.]
-
-There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+Collectively, felony disenfranchisement laws actively disenfranchise as many as 4.7 million American citizens.
+Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the potential impact of Amendment 4 in Florida.] There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future [@Weaver2010; @White2019], though other research casts some doubt on this conclusion [@Gerber2017].
@@
 In the spring of 2018,^[Although the executive order was signed on April 18th, it did not go into effect until May 18th.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-The policy change of course directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
+The policy change directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
 The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
@@
 
-In the analysis below, we examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election).
-We exclude individuals who were still on parole and could only vote because of the policy change in order to isolate the effect of the timing and manner of rights restoration for individuals whose legal eligibility to vote was not changed.
-
+In the analysis below, we examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election) --- in other words, voters whose legal eligibility to vote was not changed, but whose behavior may nonetheless have been impacted.
 By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility via brochures in a parole discharge packet, Executive Order 181 offers us the opportunity to build on existing research.
@@
 
-This research provide valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
+This research provides valuable insight into mechanisms that can increase the turnout of formerly disenfranchised individuals.
 While information included in discharge packets seems ineffective at boosting turnout, there is some evidence that notification delivered on its own is successful.
@@
 
-The criminal justice dataset comes from a public records request filed by the author to obtain individual-level incarceration and parole records for individuals sentenced to incarceration in New York State since 1990.
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; county of commitment; and others.
+Individual-level parole records come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; race or ethnicity; and others.
 This analysis is limited to individuals incarcerated for felony offenses.
 Individuals convicted of misdemeanors are not disenfranchised in New York State.
-These data come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-These data are used to determine when individuals were incarcerated or on parole, when they finished their parole supervision, and demographic information such as age and race.
 
@@
 ## Voter File Data {-}
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-We use a snapshot of the registered voter file from March 3rd, 2019.
+In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. We use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 
-I adopt the test developed in @Meredith2013 to test the prevalence of false positives.
+We adopt the test developed in @Meredith2013 to test the prevalence of false positives.
 We slightly alter the date of birth reported in the parole discharge dataset to create false records.
@@
 
-In this paper, we exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+We exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
-When we limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics.
+When we limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics (see Table \ref{tab:demo-rd}).
 The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
-The basic design is therefore an interrupted time series, in which we test whether individuals discharged after the executive order took effect turned out at higher rates than those discharged before the policy change.
-This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall turnout among formerly incarcerated individuals.
+The basic design is therefore an interrupted time series, in which we test whether individuals discharged after the executive order took effect registered and turned out at higher rates than those discharged before the policy change.
+This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall registration and turnout among formerly incarcerated individuals.
 
@@
 After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
-In this setup, rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
+In this setup, actual rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
 This instrumentation process is a common method for identifying the CACE [see, for instance, @Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.^[This manuscript is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
-To that end we use turnout --— not registration —-- as the dependent variable of interest throughout.
-However, replacing "turnout" with "registration" in the models presented herein demonstrates that Executive Order 181 increased registration among former parolees.]
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
 
@@
 
-Before estimating the interrupted time series model, we must establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
+Before estimating the interrupted time series model, we establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
 It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote, regardless of whether their voting rights are restored before they finish their sentence.
@@
 The variable *Days off Parole Before Election* indicates the number of days between a parolee's discharge date and the 2018 election.
-As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout / registration.
 
@@
 With the exception of age, the control and ITT groups are statistically indistinguishable from one another, further demonstrating the validity of the natural experiment conceptualization.
-The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote.
-To the extent our control group perhaps has a marginally higher propensity to vote than our ITT group our setup is (slightly) biased against finding a significant increase due to the executive order.
+The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote and be registered.
+To the extent our control group perhaps has a marginally higher participation propensity than our ITT group our setup is (slightly) biased against finding a significant increase due to the executive order.
 
@@
 
-Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
-Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect (*Discharged After EO 181*).
-Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, race, and whether they voted in the 2008 general election.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Model 4 interacts the ITT indicator with a dummy variable specifying whether a former parolee was Black; this allows us to test whether the ITT differed for Black and non-Black individuals.
+Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, registered and turned out at higher rates than those discharged earlier.
+Models 1 and 2 test turnout in the 2018 midterm election, while Models 3 and 4 ask whether former parolees in the treatment group were more likely to be registered prior to the registration deadline.
+Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black former parolees than others.
 
@@
 The dummy *Discharged After EO 181* measures whether an individual was discharged after May 18, 2016, and the dependent variable is turnout in the 2016 general election.
-
-In each model, robust standard errors are clustered by control / intention-to-treat status
+In each model, robust standard errors are clustered by control / intention-to-treat status.
 
@@
 
-The coefficients on *Discharged After EO 181* in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed turned out at higher rates than those discharged earlier.
+The coefficients on *Discharged After EO 181* in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed registered and turned out at higher rates than those discharged earlier.
 As expected, the coefficient is nonsignificant in the placebo regression, implying that individuals discharged shortly before an election are not generally more likely to vote.
-After controlling for other covariates, Model 3 estimates an ITT effect of 0.6 percentage points.
-Turnout among the control voters was 2.1 percent, meaning the executive order raised turnout by roughly 29 percent.
+Individuals discharged after the executive order went into effect were 0.6 percentage points more likely to vote (Model 1) and 0.8 points more likely to register (Model 3).
 
-Table \ref{tab:to-18-logit} also makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+Models 2 and 4 demonstrate that there is large racial heterogeneity in these effects.
+The interaction effect --- *Black × Discharged After EO 181* --- is negative, large relative to the general treatment effect, and statistically significant in both models.
+The executive order was apparently substantially more effective at boosting registration and turnout for non-Black former parolees than Black parolees.
 
-Model 4 estimates that the ITT for non-Black voters was 1.0 percentage points, though the interaction effect --- *Black × Discharged After EO 181* --- is negative 0.8 percentage points.
-The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
-This appears to have had a leveling effect: although Black former parolees in the control group turned out at higher rates than non-Black parolees in that group, the turnout rate among Black and non-Black voters in the ITT group in 2018 was virtually identical, as shown in Table \ref{tab:re}.
+Why would the intervention have increased turnout and registration among non-Black individuals and have had no effect on Black individuals?
+This difference may be explained in part through the causal mechanisms we have argued are at play in the case of Executive Order 181.
+As discussed above, many formerly incarcerated individuals are unsure about their legal eligibility to cast a ballot.
+It is possible that prior to the policy change formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
+This seems plausible.
+@Morris2020 looks at the geographic distribution of disenfranchised individuals in New York City with a history of voting, finding that voters lost to disenfranchisement are clustered in neighborhoods with large Black populations.
+It is possible that Black individuals return home from prison to neighborhoods with more formerly incarcerated individuals, and therefore have access to communities with better information about voting eligibility.
+This would reduce the marginal efficacy of the information-communicating aspect of Executive Order 181.
+
+If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the pre-treatment period.
+And, indeed, that is precisely what we observe.
+Table \ref{tab:re} Black individuals in the control group were far more likely to be registered in time for and participate in the 2018 general election than other control individuals.
+Post-treatment, the gap in turnout disappears, but Black individuals discharged after the executive order went into effect were still more likely to be registered than others.
+Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which even those released before the policy change had access.
 
@@
 
-The analyses in the previous section indicate that executive order was successful at increasing turnout among formerly incarcerated individuals as a group.
+The analyses in the previous section indicate that executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
 They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
-To answer this question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
-We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as "compliers."
+To answer this question, we must specifically control for whether an individual *actually* had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as "compliers." As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
+Some individuals --- namely, noncitizens --- were not eligible for rights restoration, and were thus "noncompliers:" despite being assigned to the treatment group based on their discharge date, they nonetheless did not receive the treatment.
 
@@
 Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
-Model 1 includes just *Rights Restored Before Discharge* (instrumented by *Discharged After EO 181*) while Model 2 includes the additional covariates from past specifications.
-In Model 3, we test whether the CACE was different for Black former parolees by interacting *Rights Restored Before Discharge* with *Black* (this endogenous variable is instrumented by *Black × Discharged After EO 181*).
+Each model includes *Rights Restored Before Discharge* (instrumented by *Discharged After EO 181*) while Models 2 and 4 once again test whether the CACE was different for Black individuals than for non-Black individuals.
+*Black × Rights Restored Before Discharge* is instrumented by *Black × Discharged After EO 181*.
 As before, robust standard errors are clustered by treatment assignment group.
@@
 
-The two-stage least squares approach indicates that the CACE was roughly 0.9 percentage points.
-Because overall turnout for individuals whose rights were restored was 3.2 percent, this implies that rights restoration prior to discharge from parole increased turnout by 39 percent.
-Once again, however, we see that the treatment effect differed sharply by race: the CACE for non-Black former parolees was 1.4 percentage points, but just 0.3 percentage points for Black former parolees.
+The two-stage least squares approach indicates that the CACE was roughly 0.9 percentage points (or 39 percent) for turnout (Model 1) and 1.1 percentage points (or 18 percent) for registration (Model 3).
+These models once again indicate that the treatment effect was substantially smaller, if not nonexistent, for Black individuals.
 
@@
 The mechanism through which the rules change increased turnout among these individuals is not clear.
-As discussed above, the literature indicates that this could be because of the relational nature of the restoration process here [@Lerman2014] or because of better information communication of eligibility [@Meredith2015; @Gerber2015].
-In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
-Future work should attempt to "unbundle" this treatment to understand the relative importance of the constituent mechanisms.
+@Weaver2010 argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
+Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
+On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+Such interactions may undo some of the negative socialization identified by @Weaver2010.
+
+It could also be a story of better information.
+As @Meredith2015 and @Gerber2015 show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see @Meredith2013).
+Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+Of course, even voters in the control group were receiving notification of their eligibility through the mail.
+If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+In reality, the executive order's success at boosting turnout and registration likely operated through multiple mechanisms.
 
 Troublingly, the effect of pre-discharge rights restoration seems to vary based on parolees' race.
-This discrepancy could be caused by a number of different factors.
-Firstly, Black and white individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
-The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
-
-We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
-Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than white prisoners [@Hemmens2000].
-Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-@White2015, for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino white aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races [@Butler2011], and that public housing officials respond less to inquiries from Latinos [@Einstein2017].
-
-Parole officers may more enthusiastically encourage white parolees to register to vote and cast a ballot.
-This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such white individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among white former parolees; future research must investigate why this is the case.
+AS discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
+Nevertheless, the in-person meeting with parole officers should still have improved their political socialization and relationship with the government.
+Whether this is because the negative socialization of an incarceration is "stickier" for Black voters, because parole officers gave less affirmative encouragement to Black parolees (in other words, they received a weaker treatment), or for some other reason is not clear.
+Ultimately, the data at hand cannot answer why the intervention was unsuccessful at raising turnout and registration among Black former parolees; future research must investigate this.
 
@@
 
-stargazer(date_to_model_time,
+stargazer(date_to_model_time, date_reg_model_time,
           header = F,
@@
                                "Years on Parole"),
-          dep.var.labels = c("Cast Ballot in 2018 Election"),
-          title = "\\label{tab:to-18-logit-short} 2018 Turnout for Control Group",
+          dep.var.labels = c("Voted in 2018", "Registered in 2018"),
+          title = "\\label{tab:to-18-logit-short} 2018 Behavior for Control Group",
           table.placement = "H",
@@
           out.header = F,
-          se = date_ses)
+          se = c(date_ses, date_reg_ses))
 
@@
 
-stargazer(model1, model2, model3, model5, model4,
+ses_cl[["m4"]] <- m4_ses
+
+stargazer(models, model4,
           header = F,
@@
                                "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          dep.var.labels = c("Voted in 2018", "Registered in 2018", "Voted in 2016"),
           title = "\\label{tab:to-18-logit} Turnout in General Election",
@@
           keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X", "X"),
-                         c("Felony Class FE", "", "", "X", "X", "X")),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X", "X"),
+                         c("Felony Class FE", "X", "X", "X", "X", "X")),
           table.layout = "-dm#-t-a-s-n",
@@
           order = c("finished_post", "pb"),
-          se = list(m1_ses$V1, m2_ses$V1, m3_ses$V1, m5_ses$V1, m4_ses$V1))
+          se = ses_cl)
 j <- fread("./temp/test.tex", header = F, sep = "+")
@@
 ```{r rd-dmeans, echo = F, warning = F}
-k <- readRDS("./temp/means_diff_black.rds")
-j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout by Treatment Group, Race", linesep = "",
+k <- readRDS("./temp/means_diff_black.rds") %>% 
+  select(-Statistic)
+j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout and Registration by Treatment Group, Race", linesep = "",
                   align = c("l", "c", "c", "c")) %>% 
   kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
+                full_width = F) %>% 
+  group_rows(start_row = 1,
+             end_row = 2,
+             group_label = "Turnout") %>% 
+  group_rows(start_row = 3,
+             end_row = 4,
+             group_label = "Registration")
 
@@
 
-lab <- "\\caption{\\label{tab:re} Turnout by Treatment Group, Race}"
+lab <- "\\caption{\\label{tab:re} Turnout and Registration by Treatment Group, Race}"
 insert1 <- "\\fontsize{10}{12}\\selectfont"
@@
 
-stargazer(iv1, iv3, iv4,
+stargazer(iv3, iv4, iv3b, iv4b,
           header = F,
@@
                                "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Voted in 2016"),
+          dep.var.labels = c("Voted in 2018", "Registered in 2018"),
           title = "\\label{tab:iv-tab} Two-Stage Least Squares",
@@
           keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
-                         c("Felony Class FE", "", "X", "X", "X")),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X"),
+                         c("Felony Class FE", "X", "X", "X", "X")),
           table.layout = "-dm#-t-a-s-n",
@@
           out.header = F,
-          se = list(iv1_ses, iv3_ses, iv4_ses))
+          se = list(iv3_ses, iv4_ses, iv3_sesb, iv4_sesb))
 

### 2020-10-09 markdown/part2/rewrite.tex
start ssq revisions

@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{July 07, 2020}
+\date{October 09, 2020}
 
@@
 
-In all but two states (Maine and Vermont), felony disenfranchisement laws mean that American citizens convicted of felony offenses lose the right to vote for at least some period of time.
-In some states, such as Oregon and Massachusetts, individuals lose that right only for the period in which they are actively incarcerated.
-Iowa stands alone at the other end of the spectrum, where all felony convictions result in lifelong disenfranchisement unless a returned citizen receives an individual pardon from the state's governor (Brennan Center for Justice \protect\hyperlink{ref-bcj_laws}{2019}).
-This variation in laws arises from language in the Fourteenth Amendment which allows states to revoke individuals' voting rights ``for participation in rebellion, or other crime.'' The definition of ``other crime,'' left so vague in the Constitution, is now generally used by states to disenfranchise citizens for any felony offense.
-The Supreme Court, in cases such as \emph{Richardson v.
-Ramirez} (1974), has upheld this practice.
-Collectively, these laws disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the potential impact of Amendment 4 in Florida.}
-
-There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+Collectively, felony disenfranchisement laws actively disenfranchise as many as 4.7 million American citizens.
+Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the potential impact of Amendment 4 in Florida.} There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future (Weaver and Lerman \protect\hyperlink{ref-Weaver2010}{2010}; White \protect\hyperlink{ref-White2019}{2019}), though other research casts some doubt on this conclusion (Gerber et al.
@@
 In the spring of 2018,\footnote{Although the executive order was signed on April 18th, it did not go into effect until May 18th.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
-The policy change of course directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
+The policy change directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
 The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
@@
 
-Most states in the United States are required to maintain files with information on all registered voters.
-In New York, this information is publicly available from the Board of Elections.
-It includes information on all registered voters, including: first, middle, and last name; date of birth; vote history; and other information.
-The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense.
-We use a snapshot of the registered voter file from March 3rd, 2019.
+In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. We use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 
-I adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
+We adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
 We slightly alter the date of birth reported in the parole discharge dataset to create false records.
@@
 After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
-In this setup, rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
+In this setup, actual rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
 This instrumentation process is a common method for identifying the CACE (see, for instance, Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.\footnote{This manuscript is primarily interested in the effect of in-person rights restoration prior to parole discharge on ultimate political representation.
-To that end we use turnout ----- not registration ----- as the dependent variable of interest throughout.
-However, replacing ``turnout'' with ``registration'' in the models presented herein demonstrates that Executive Order 181 increased registration among former parolees.}
+The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
 
@@
 The variable \emph{Days off Parole Before Election} indicates the number of days between a parolee's discharge date and the 2018 election.
-As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout / registration.
 
@@
 
-Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, turned out at higher rates than those discharged earlier.
-Model 1 controls only for whether an individual was discharged after Executive Order 181 went into effect (\emph{Discharged After EO 181}).
-Model 2 also controls for individual-level characteristics: sex, age on November 6th, 2018, race, and whether they voted in the 2008 general election.
-Model 3 adds sentence-specific information to Model 2: the amount of time the individual spent on parole, and the class(es) of felony for which they were convicted.
-Model 4 interacts the ITT indicator with a dummy variable specifying whether a former parolee was Black; this allows us to test whether the ITT differed for Black and non-Black individuals.
+Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, registered and turned out at higher rates than those discharged earlier.
+Models 1 and 2 test turnout in the 2018 midterm election, while Models 3 and 4 ask whether former parolees in the treatment group were more likely to be registered prior to the registration deadline.
+Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black former parolees than others.
 
@@
 The dummy \emph{Discharged After EO 181} measures whether an individual was discharged after May 18, 2016, and the dependent variable is turnout in the 2016 general election.
-
-In each model, robust standard errors are clustered by control / intention-to-treat status
+In each model, robust standard errors are clustered by control / intention-to-treat status.
 
@@
 
-The coefficients on \emph{Discharged After EO 181} in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed turned out at higher rates than those discharged earlier.
+The coefficients on \emph{Discharged After EO 181} in Table \ref{tab:to-18-logit} indicate that individuals discharged after the policy changed registered and turned out at higher rates than those discharged earlier.
 As expected, the coefficient is nonsignificant in the placebo regression, implying that individuals discharged shortly before an election are not generally more likely to vote.
-After controlling for other covariates, Model 3 estimates an ITT effect of 0.6 percentage points.
-Turnout among the control voters was 2.1 percent, meaning the executive order raised turnout by roughly 29 percent.
+Individuals discharged after the executive order went into effect were 0.6 percentage points more likely to vote (Model 1) and 0.8 points more likely to register (Model 3).
 
-Table \ref{tab:to-18-logit} also makes clear that formerly incarcerated men were far less likely to vote than formerly incarcerated women; that older formerly incarcerated individuals were more likely to cast a ballot; and individuals who spent longer on parole were more likely to participate in the midterm election.
+Models 2 and 4 demonstrate that there is large racial heterogeneity in these effects.
+The interaction effect --- \emph{Black × Discharged After EO 181} --- is negative, large relative to the general treatment effect, and statistically significant in both models.
+The executive order was apparently substantially more effective at boosting registration and turnout for non-Black former parolees than Black parolees.
 
-Model 4 estimates that the ITT for non-Black voters was 1.0 percentage points, though the interaction effect --- \emph{Black × Discharged After EO 181} --- is negative 0.8 percentage points.
-The executive order was apparently substantially more effective at boosting turnout for non-Black former parolees than Black parolees.
-This appears to have had a leveling effect: although Black former parolees in the control group turned out at higher rates than non-Black parolees in that group, the turnout rate among Black and non-Black voters in the ITT group in 2018 was virtually identical, as shown in Table \ref{tab:re}.
+Why would the intervention have increased turnout and registration among non-Black individuals and have had no effect on Black individuals?
+This difference may be explained in part through the causal mechanisms we have argued are at play in the case of Executive Order 181.
+As discussed above, many formerly incarcerated individuals are unsure about their legal eligibility to cast a ballot.
+It is possible that prior to the policy change formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
+This seems plausible.
+Morris (\protect\hyperlink{ref-Morris2020}{2020}) looks at the geographic distribution of disenfranchised individuals in New York City with a history of voting, finding that voters lost to disenfranchisement are clustered in neighborhoods with large Black populations.
+It is possible that Black individuals return home from prison to neighborhoods with more formerly incarcerated individuals, and therefore have access to communities with better information about voting eligibility.
+This would reduce the marginal efficacy of the information-communicating aspect of Executive Order 181.
+
+If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the pre-treatment period.
+And, indeed, that is precisely what we observe.
+Table \ref{tab:re} Black individuals in the control group were far more likely to be registered in time for and participate in the 2018 general election than other control individuals.
+Post-treatment, the gap in turnout disappears, but Black individuals discharged after the executive order went into effect were still more likely to be registered than others.
+Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which even those released before the policy change had access.
 
@@
 
-The analyses in the previous section indicate that executive order was successful at increasing turnout among formerly incarcerated individuals as a group.
+The analyses in the previous section indicate that executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
 They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
-To answer this question, we must specifically control for whether an individual actually had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-Because our ITT pool in the previous section include individuals not eligible for treatment, we expect that the ITT effect will be lower than the treatment effect on the treated individuals.
-We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as ``compliers.''
+To answer this question, we must specifically control for whether an individual \emph{actually} had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
+We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as ``compliers.'' As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
+Some individuals --- namely, noncitizens --- were not eligible for rights restoration, and were thus ``noncompliers:'' despite being assigned to the treatment group based on their discharge date, they nonetheless did not receive the treatment.
 
@@
 Table \ref{tab:iv-tab} presents the results of the two-stage least squares regression.
-Model 1 includes just \emph{Rights Restored Before Discharge} (instrumented by \emph{Discharged After EO 181}) while Model 2 includes the additional covariates from past specifications.
-In Model 3, we test whether the CACE was different for Black former parolees by interacting \emph{Rights Restored Before Discharge} with \emph{Black} (this endogenous variable is instrumented by \emph{Black × Discharged After EO 181}).
+Each model includes \emph{Rights Restored Before Discharge} (instrumented by \emph{Discharged After EO 181}) while Models 2 and 4 once again test whether the CACE was different for Black individuals than for non-Black individuals.
+\emph{Black × Rights Restored Before Discharge} is instrumented by \emph{Black × Discharged After EO 181}.
 As before, robust standard errors are clustered by treatment assignment group.
@@
 The mechanism through which the rules change increased turnout among these individuals is not clear.
-As discussed above, the literature indicates that this could be because of the relational nature of the restoration process here (Lerman and Weaver \protect\hyperlink{ref-Lerman2014}{2014}) or because of better information communication of eligibility (Meredith and Morse \protect\hyperlink{ref-Meredith2015}{2015}; Gerber et al.
-\protect\hyperlink{ref-Gerber2015}{2015}).
-In reality, the executive order's success at boosting turnout likely operated through multiple mechanisms.
-Future work should attempt to ``unbundle'' this treatment to understand the relative importance of the constituent mechanisms.
+Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}) argues that contact with the criminal justice system restructures how individuals understand their relationship with the government and sours their desire to participate.
+Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
+On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
+
+It could also be a story of better information.
+As Meredith and Morse (\protect\hyperlink{ref-Meredith2015}{2015}) and Gerber et al.
+(\protect\hyperlink{ref-Gerber2015}{2015}) show, reminding formerly disenfranchised individuals of their restored voting rights can increase their participation (but see Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013})).
+Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
+When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+Of course, even voters in the control group were receiving notification of their eligibility through the mail.
+If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+In reality, the executive order's success at boosting turnout and registration likely operated through multiple mechanisms.
 
 Troublingly, the effect of pre-discharge rights restoration seems to vary based on parolees' race.
-This discrepancy could be caused by a number of different factors.
-Firstly, Black and white individuals on parole might differ in meaningful ways that impact the successfulness of the intervention.
-As discussed above, Black participation among individuals discharged from parole prior to the executive order voted at significantly higher rates in 2018.
-The intervention was likely successful for parolees who would not have voted otherwise, but needed only a small encouragement.
-It may be that fewer Black former parolees are susceptible to small encouragements: they may separate more cleanly into voters and nonvoters, with fewer individuals susceptible to parole officer encouragement.
-
-We cannot, however, rule out the possibility that racial bias among parole officers plays some role in the effectiveness of the intervention.
-Although parole officers may not be overtly or consciously biased toward their parolees, such bias is possible.
-Although there is limited literature examining racial variation in parolees' perceptions of their parole officers, there is some evidence that Black prisoners report worse inmate-staff relationships than white prisoners (Hemmens and Marquart \protect\hyperlink{ref-Hemmens2000}{2000}).
-Moreover, recent research indicates that street-level bureaucrats may exhibit some racial bias in the services they provide.
-White, Nathan, and Faller (\protect\hyperlink{ref-White2015}{2015}), for instance, shows that local election administrators are less likely to respond to email questions from Latino aliases than non-Latino white aliases.
-There is also evidence that politicians are less likely to respond to requests from constituents of different races (Butler and Broockman \protect\hyperlink{ref-Butler2011}{2011}), and that public housing officials respond less to inquiries from Latinos (Einstein and Glick \protect\hyperlink{ref-Einstein2017}{2017}).
-
-Parole officers may more enthusiastically encourage white parolees to register to vote and cast a ballot.
-This variation in encouragement would not necessarily be indicative of racial animus: it could arise from a parole officer's expectations about a parolee's political preferences.
-Officers are likely to provide greater encouragement to parolees they perceive to have similar political preferences.
-In the case of former parolees in New York State, race does serve as an effective proxy for partisanship: 82.5 percent of Black individuals discharged from parole since 2012 who participated in the 2018 election were registered Democrats; just 31.2 percent of such white individuals were registered Democrats.
-Ultimately, the data at hand cannot answer why the intervention succeeded at raising turnout only among white former parolees; future research must investigate why this is the case.
+AS discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
+Nevertheless, the in-person meeting with parole officers should still have improved their political socialization and relationship with the government.
+Whether this is because the negative socialization of an incarceration is ``stickier'' for Black voters, because parole officers gave less affirmative encouragement to Black parolees (in other words, they received a weaker treatment), or for some other reason is not clear.
+Ultimately, the data at hand cannot answer why the intervention was unsuccessful at raising turnout and registration among Black former parolees; future research must investigate this.
 
@@
 \url{https://doi.org/10.1111/j.1540-5893.2011.00448.x}.
-
-\leavevmode\hypertarget{ref-Butler2011}{}%
-Butler, Daniel M., and David E.
-Broockman.
-2011.
-``Do Politicians Racially Discriminate Against Constituents?
-A Field Experiment on State Legislators.'' \emph{American Journal of Political Science} 55 (3): 463--77.
-\url{https://doi.org/10.1111/j.1540-5907.2011.00515.x}.
 
@@
 \url{https://www.prisonpolicy.\%20org/scans/sp/fd_studiesvotingbehavior.pdf}.
-
-\leavevmode\hypertarget{ref-Einstein2017}{}%
-Einstein, Katherine Levine, and David M.
-Glick.
-2017.
-``Does Race Affect Access to Government Services?
-An Experiment Exploring Street-Level Bureaucrats and Access to Public Housing.'' \emph{American Journal of Political Science} 61 (1): 100--116.
-\url{https://doi.org/10.1111/ajps.12252}.
 
@@
 \url{https://doi.org/10.1086/692670}.
-
-\leavevmode\hypertarget{ref-Hemmens2000}{}%
-Hemmens, Craig, and James W.
-Marquart.
-2000.
-``Friend or Foe?
-Race, Age, and Inmate Perceptions of Inmate-Staff Relations.'' \emph{Journal of Criminal Justice} 28 (4): 297--312.
-\url{https://doi.org/10.1016/S0047-2352(00)00044-1}.
 
@@
 \url{https://doi.org/10.1016/j.jpubeco.2003.10.005}.
+
+\leavevmode\hypertarget{ref-Morris2020}{}%
+Morris, Kevin. 2020. ``Neighborhoods and Felony Disenfranchisement: The Case of New York City.'' \emph{Urban Affairs Review}, May, 1078087420921522. \url{https://doi.org/10.1177/1078087420921522}.
 

### 2020-10-09 code/00_initialize.R

@@ -28,27 +28,6 @@ library(stargazer)
 ## DOING SO TO WORK AROUND MEMORY CONSTRAINTS
 
 
-theme_bc <- function(){
-  library(ggplot2)
-  library(extrafont)
-  theme(plot.title = element_text(family = "Roboto Black", 
-                                  color = "black", hjust = 0.5, margin = margin(0, 0, 10, 0)),
-        axis.title = element_text(family = "Roboto", 
-                                  color = "#525353"),
-        axis.line = element_line(colour = "black"), 
-        text = element_text(family = "Roboto", color = "#525353"), 
-        plot.caption = element_text(family = "Roboto", size = 8),
-        panel.border = element_blank(),
-        panel.grid.major = element_blank(), 
-        panel.background = element_rect(fill = "#dddddb"),
-        plot.background = element_rect(fill = "#dddddb"), 
-        panel.grid.minor = element_blank(),
-        legend.key = element_blank(), 
-        legend.background = element_rect(color = "black", fill = "#dddddb"), 
-        plot.subtitle = element_text(hjust = 0.5, size = 10), 
-        plot.margin = margin(10, 10, 10, 10))
-}
-
 weighted.ttest.ci <- function(x, weights, conf.level = 0.95) {
   nx <- length(x)
   df <- nx - 1

### 2020-10-09 code/part_2/05_turnout_trends.R

@@ -3,7 +3,7 @@
 # 
 # ############# READ IN NYS VOTER FILES #############
 # db <- dbConnect(SQLite(), "D:/rolls.db")
-# nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id, political_party from nys_roll_0319")
+# nys_roll <- dbGetQuery(db, "select history, voter_status, nys_id, political_party, registration_date from nys_roll_0319")
 # 
 # nys_roll <- nys_roll[nys_roll$nys_id %in% readRDS("./temp/din_nys_parolees.rds")$nys_id, ]
 # 

### 2020-10-09 code/part_2/06_individual_turnout.R

@@ -20,23 +20,29 @@ parolees <- left_join(parolees, nys_roll, by = "din") %>%
          days_since_m21 = parole_status_date - as.Date("2018-05-19"),
          v2016 = ifelse(is.na(v2016), 0, v2016),
          v2008 = ifelse(is.na(v2008), 0, v2008),
-         registered = voter_status %in% c("ACTIVE", "INACTIVE"))
+         registered = (voter_status %in% c("ACTIVE", "INACTIVE")) & (registration_date <= 20181012))
 ###########################################
 pool <- filter(parolees, year(parole_status_date) >= 2017) %>% 
   mutate(b = race == "BLACK")
 
-h <- t.test(v2018 ~ b, filter(pool, !finished_post))
-j <- t.test(v2018 ~ b, filter(pool, finished_post))
+g <- t.test(v2018 ~ b, filter(pool, !finished_post))
+h <- t.test(v2018 ~ b, filter(pool, finished_post))
+
+i <- t.test(registered ~ b, filter(pool, !finished_post))
+j <- t.test(registered ~ b, filter(pool, finished_post))
 
 k <- pool %>% 
   group_by(b, finished_post) %>% 
-  summarize(turnout = percent(mean(v2018), accuracy = .01)) %>% 
-  pivot_wider(id_cols = "finished_post", values_from = "turnout", names_from = "b") %>% 
-  mutate(finished_post = ifelse(finished_post, "ITT", "Control"))
+  summarize(Turnout = percent(mean(v2018), accuracy = .01),
+            Registration = percent(mean(registered), accuracy = .01)) %>% 
+  pivot_longer(cols = c(Turnout, Registration)) %>% 
+  pivot_wider(id_cols = c(name, finished_post), values_from = c(value), names_from = "b") %>% 
+  mutate(finished_post = ifelse(finished_post, "ITT", "Control")) %>% 
+  arrange(desc(name), finished_post)
 
-k$p <- round(c(h$p.value, j$p.value), digits = 3)
+k$p <- round(c(g$p.value, h$p.value, i$p.value, j$p.value), digits = 3)
 
-colnames(k) <- c("Group", "Non-Black", "Black", "p-value")
+colnames(k) <- c("Statistic", "Group", "Non-Black", "Black", "p-value")
 saveRDS(k, "./temp/means_diff_black.rds")
 
 ####################
@@ -44,86 +50,64 @@ saveRDS(k, "./temp/means_diff_black.rds")
[…]

### 2020-10-09 code/part_2/13_window_sensitivity_2.R

@@ -110,7 +110,7 @@ p1 <- ggplot(filter(cints)) +
   labs(y = "Estimated ITT Effect",
        x = "Months Around Cutpoint",
        caption = "Notes: 95% confidence bars shown.
-Regressions include covariates from Models 3 and 5 of Table 4.
+Regressions include covariates from Models 1 and 5 of Table 4.
 Number of observations in each regression shown on right axis.",
        shape = "Year",
        linetype = "Year") +

### 2020-10-27 markdown/part2/rewrite.Rmd
ssq r&r

@@
     - \usepackage{setspace}
+    - \usepackage{lineno}
+    - \linenumbers
 abstract: |
   This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
-In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, we interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
-We demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
-This group-level effect, however, masks race-specific effects.
-Although rights restoration prior to parole discharge increased turnout by non-Black former parolees by 1.4 percentage points, the treatment effect was substantially smaller for Black former parolees.
-This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans’ view of the democratic process.
+In 2018, New York State began restoring voting rights to individuals on parole, after previously restoring voting rights only at the completion of parole.
+By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to register and cast a ballot.
+I demonstrate that in-person rights restoration prior to parole discharge significantly increased participation, but this group-level effect masks racially heterogeneous effects.
+Although rights restoration prior to parole discharge increased turnout by non-Black individuals by 1.4 percentage points, the treatment effect was substantially smaller for Black individuals.
+This raises important questions about how members of different communities can be brought back into our democratic processes.
 ---
@@
 
-Collectively, felony disenfranchisement laws actively disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated [@sentencing_2016].^[The figures reported in @sentencing_2016 have been adjusted to reflect the potential impact of Amendment 4 in Florida.] There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+Collectively, felony disenfranchisement laws actively disenfranchise as many as 5.2 million American citizens.
+Of these, the majority are no longer incarcerated, but are either on parole, probation, or are disenfranchised after completing their sentence [@Uggen2020].
+The effects of felony disenfranchisement, however, might be even broader: there is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future [@Weaver2010; @White2019], though other research casts some doubt on this conclusion [@Gerber2017].
@@
 Formal disenfranchisement policy, the literature has made clear, is just one piece of an interlocking system that serves to reduce the representation of minority and marginalized voters.
-The incarcerated population is drawn from a pool of individuals unlikely to vote even prior to their incarceration.
+The incarcerated population is drawn from a pool of individuals who rarely vote, both prior to and after their incarceration.
 To address only the formal laws contributing to disenfranchisement without also interrogating efforts to boost post-supervision participation risks leaving much of the system of effective disenfranchisement undisturbed.
@@
 
-In the spring of 2018,^[Although the executive order was signed on April 18th, it did not go into effect until May 18th.] Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+In the spring of 2018,^[Although the executive order was signed on April 18th, it did not go into effect until May 18th.] Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most New Yorkers on parole.
 The policy change directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
-The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
+For instance, the change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to *de facto* disenfranchisement among probationers [@Drucker2005].
 By re-enfranchising anyone not actively incarcerated, the rules were substantially simplified.
@@
 Beginning in 2010, New York State began including information about voter eligibility in parole discharge packets, a change that @Meredith2013 indicates did not increase turnout.
-Executive Order 181 both changed the timing of rights restoration (upon release from prison, not the end of parole) and the manner of restoration (via an in-person meeting, rather than in a discharge packet).
-This study tests the bundled effect of these changes.
+Executive Order 181 both changed the *timing* of rights restoration (upon release from prison, not the end of parole) and the *manner* of restoration (via an in-person meeting, rather than in a discharge packet).
+This study tests the bundled effect of these changes on individuals whose legal right to vote was not changed by the executive order because they were discharged from parole before the registration deadline for the 2018 midterm elections.
 
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[See https://nyassembly.gov/member_files/139/webdocs/82103.pdf] The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.^[See https://nyassembly.gov/member_files/139/webdocs/82103.pdf.] The memorandum directs all parole officers to present individuals on parole with voter registration forms and to explain their purpose.
 Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
@@
 
-In the analysis below, we examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election) --- in other words, voters whose legal eligibility to vote was not changed, but whose behavior may nonetheless have been impacted.
+In the analysis below, I examine the effect of Executive Order 181 on individuals who finished parole before October 10th, 2018 (the registration deadline for the 2018 general election) --- in other words, voters whose legal eligibility to vote was not changed, but whose behavior may nonetheless have been impacted.
 By examining the effect of rights restoration in the context of a personal relationship with a parole officer against a status quo in which formerly incarcerated individuals were informed of their eligibility via brochures in a parole discharge packet, Executive Order 181 offers us the opportunity to build on existing research.
@@
 Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates [see, for instance, @Burch2011].
-Regardless of whether incarceration reduces turnout, the  state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
+Thus, even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
 
@@
 In the three states studied in @Meredith2013, however, newly re-enfranchised individuals received this information as part of a discharge packet.
-Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform former parolees that their voting rights had been restored.
-Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out], and it is possible that these brochures, in-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
+Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform individuals that their voting rights had been restored.
+Formerly convicted individuals are very often confused about their eligibility to vote [@Drucker2005; @locked_out], and it is possible that these brochures, non-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
 
@@
 Nevertheless, the research to date has examined the effect of rights restoration notification outside a personal relationship on post-supervision turnout.
-It seems likely that an in-person meeting will be more effective at dispelling confusion about eligibility than a brochure included in a discharge packet.
+An in-person meeting might be more effective at dispelling confusion about eligibility than a brochure included in a discharge packet.
 
@@
 A notification sent in the mail is a binary treatment: the letter was either sent or not sent.
-Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if parolees have a warm relationship with their parole officer, they may be strongly encouraged to register and participate.
-On the other hand, if parole officers are biased against their parolees (due either to racial hostility, personal dislike, or expected partisan affiliation) they may not provide any effective treatment at all.
+Communicating the restoration of voting rights through human interaction means that the treatment might vary by individuals: if individuals have a warm relationship with their parole officer, they may be strongly encouraged to register and participate.
+On the other hand, if parole officers are biased against their charges (due either to racial hostility, personal dislike, or expected partisan affiliation) they may not provide any effective treatment at all.
 Thus, while the communication of voting rights restoration in the context of a personal relationship has the potential to deliver much larger benefits, it also opens the door to potential discrepancies in treatment.
 
-We expect that restoring voting rights while an individual is still on parole increases that individual's later propensity to vote.
-This increase will operate through two mechanisms: as the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a former parolee more sure of her voting rights than a brochure.
-Secondly, the interpersonal nature of the parolee-parole officer relationship is expected to more effectively repair (some of) the damage done to the former parolees interpretation of the state.
+I therefore expect that Executive Order 181 will increase the turnout of individuals even after they are discharged from parole through two causal mechanisms.
+As the previous literature has established, notification of voting rights can increase turnout, and being informed by a parole officer likely leaves a formerly paroled individual more sure of her voting rights than a brochure.
+Secondly, the interpersonal nature of the relationship is expected to more effectively repair (some of) the damage done to the formerly incarcerated individual's interpretation of the state.
 Individual, in-person invitations to rejoin the body politic are likely effective.
@@
 Individual-level parole records come from the New York State Department of Corrections and Community Supervision (NYSDOCCS).
-The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; race or ethnicity; and others.
+The data includes a host of information, including: first, middle, and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; race / ethnicity; and others.
 This analysis is limited to individuals incarcerated for felony offenses.
@@
 
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, we identify individuals who had their voting rights restored.^[Not all parolees listed in the public records request data are included in the lookup tool.
-Roughly 1 percent of former parolees are not in the online lookup.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether an individual had her voting rights restored.
+By using the identification number provided from the public records request and this website, I identify individuals who had their voting rights restored.^[Not all individuals listed in the public records request data are included in the lookup tool.
+Roughly 1 percent are not in the online lookup.] There were `r comma(readRDS("./temp/how_many_restored.rds"))` individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 Roughly 1,200 individuals who were discharged between the effective date of the executive order and the registration deadline did not have voting rights restored, due largely to their status as noncitizens.
@@
 ## Voter File Data {-}
-In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. We use a snapshot of the registered voter file from March 3rd, 2019.
+In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. I use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
-We match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a "match," records must have the exact same birth date.
+I match individuals in each dataset using first name, middle name, last name, and date of birth.
+To be considered a match, records must have the exact same birth date.
 The first and last names must also be exact matches (conditional on the adjustments discussed below).
@@
 
-We adopt the test developed in @Meredith2013 to test the prevalence of false positives.
-We slightly alter the date of birth reported in the parole discharge dataset to create false records.
+I adopt the test developed in @Meredith2013 to test the prevalence of false positives.
+I slightly alter the date of birth reported in the parole discharge dataset to create false records.
 Comparing the number of matches between these "fake" discharge records and the voter file with the number of matches between the "true" records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when we construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
-This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from an individual's birthdate.
+This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is probably too small to have any material impact on the overall analysis.
 
@@
 Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, we will not identify her using this matching methodology.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using this matching methodology.
 Similarly, "John Doe" and "Jonathan Doe" would not result in a match.
-To reduce the likelihood of these false negatives we remove all punctuation from all names, and standardize capitalization.
+To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
 A record with a last name of "O'Donnell" in one dataset, therefore, would match a last name of "O DONNELL" in the other (provided the other criteria are satisfied).
 Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
+For three reasons this is not expected to present major challenges: firstly, far more women change their last names than men, and women make up less than 10 percent of individuals who have been discharged from felony parole.
+Secondly, because both discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be "John" in one set of records and "Jonathan" in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the executive order went into effect.
 
@@
 
-We exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change, measured both using turnout in the 2018 election, and registration status as of the registration deadline for the same contest.
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
 When we limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics (see Table \ref{tab:demo-rd}).
-The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
-The basic design is therefore an interrupted time series, in which we test whether individuals discharged after the executive order took effect registered and turned out at higher rates than those discharged before the policy change.
-This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall registration and turnout among formerly incarcerated individuals.
+The assignment to treatment is as nearly random as can be hoped for in the context of a natural experiment.
+The basic design is therefore an interrupted time series, in which I test whether individuals discharged after the executive order took effect registered and turned out at higher rates than those discharged before the policy change.
+This allows me to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall registration and turnout among formerly incarcerated individuals.
 
-As discussed above, not all formerly incarcerated individuals were eligible to have their rights restored; the ITT effect will therefore be biased towards zero, insofar as the behavior of ineligible parolees remains unchanged.
-After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
-In this setup, actual rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
-This instrumentation process is a common method for identifying the CACE [see, for instance, @Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
+Not all formerly incarcerated individuals were eligible to have their rights restored.
+Noncitizens, for instance, and individuals who were re-arrested while on parole did not have their voting rights restored.
+Individuals who fall into these categories and were discharged after the executive order are nonetheless included in the ITT group, even though they were not in fact treated (put differently, they did not "comply" with being assigned to the treatment group).
+Because some individuals in the ITT group went untreated, the ITT effect will be biased towards zero.
+The ITT therefore cannot be understood as the causal effect of rights restoration on the turnout of the individuals whose rights were actually restored.
+
+Two-stage least squares approaches are commonly used to estimate the causal effect of treatment in a scenario with "imperfect compliance" [see, for instance, @Ansolabehere1999; @Gerber2000; @Milligan2004; @Lassen2004; @Sondheimer2010].
+By using the random assignment to treatment as an instrument for actual treatment, the complier average causal effect (CACE) can be estimated.
+Here, whether someone was discharged from parole after the executive order went into effect (that is, whether someone was assigned to the treatment group) instruments whether their rights were actually restored (whether they "complied" with being treated).
+Of course, using terms like "compliance" here is slightly misleading: the individuals who did not "comply" were legally barred from having voting rights restored (or, in the case of noncitizens, conferred) --- they had little influence over whether or not they were actually treated, after being assigned to the treatment group.
+Nevertheless, I use this nomenclature to make the ties to existing methodological research clear.
 
@@
 
-Before estimating the interrupted time series model, we establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
-It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote, regardless of whether their voting rights are restored before they finish their sentence.
+Before estimating the interrupted time series model, I establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
+It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote irrespective of policy intervention.
 
 Table \ref{tab:to-18-logit-short} presents an ordinary least squares regression using only the control voters.
-The variable *Days off Parole Before Election* indicates the number of days between a parolee's discharge date and the 2018 election.
-As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout / registration.
+The variable *Days off Parole Before Election* indicates the number of days between an individual's discharge date and the 2018 election.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout and registration.
 
@@
 The control group is, on average, slightly older, which Table \ref{tab:to-18-logit-short} indicates makes them more likely to vote and be registered.
-To the extent our control group perhaps has a marginally higher participation propensity than our ITT group our setup is (slightly) biased against finding a significant increase due to the executive order.
+To the extent the control group perhaps has a marginally higher participation propensity than the ITT group, this setup is (slightly) biased against finding a significant increase due to the executive order.
 
@@
 Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, registered and turned out at higher rates than those discharged earlier.
-Models 1 and 2 test turnout in the 2018 midterm election, while Models 3 and 4 ask whether former parolees in the treatment group were more likely to be registered prior to the registration deadline.
-Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black former parolees than others.
+Models 1 and 2 ask whether individuals in the treatment group were more likely to be registered prior to the registration deadline, while Models 3 and 4 test turnout in the 2018 midterm election.
+Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black individuals than others.
 
-In Model 5 we estimate a "placebo" regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+In Model 5 I estimate a "placebo" regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
 Model 5 includes all individuals last discharged between January 1, 2015, and October 14th, 2016.
@@
 As expected, the coefficient is nonsignificant in the placebo regression, implying that individuals discharged shortly before an election are not generally more likely to vote.
-Individuals discharged after the executive order went into effect were 0.6 percentage points more likely to vote (Model 1) and 0.8 points more likely to register (Model 3).
+Individuals discharged after the executive order went into effect were 0.8 points more likely to register (Model 1) and 0.6 percentage points more likely to vote (Model 3).
 
@@
 The interaction effect --- *Black × Discharged After EO 181* --- is negative, large relative to the general treatment effect, and statistically significant in both models.
-The executive order was apparently substantially more effective at boosting registration and turnout for non-Black former parolees than Black parolees.
+The executive order was apparently substantially more effective at boosting registration and turnout for non-Black than Black individuals.
 
 Why would the intervention have increased turnout and registration among non-Black individuals and have had no effect on Black individuals?
-This difference may be explained in part through the causal mechanisms we have argued are at play in the case of Executive Order 181.
+This difference may be explained in part through the causal mechanisms I have argued are at play in the case of Executive Order 181.
 As discussed above, many formerly incarcerated individuals are unsure about their legal eligibility to cast a ballot.
-It is possible that prior to the policy change formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
-This seems plausible.
-@Morris2020 looks at the geographic distribution of disenfranchised individuals in New York City with a history of voting, finding that voters lost to disenfranchisement are clustered in neighborhoods with large Black populations.
-It is possible that Black individuals return home from prison to neighborhoods with more formerly incarcerated individuals, and therefore have access to communities with better information about voting eligibility.
+It is possible that, prior to the policy change, formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
 This would reduce the marginal efficacy of the information-communicating aspect of Executive Order 181.
 
-If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the pre-treatment period.
+If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the control group.
 And, indeed, that is precisely what we observe.
-Table \ref{tab:re} Black individuals in the control group were far more likely to be registered in time for and participate in the 2018 general election than other control individuals.
+Table \ref{tab:re} shows that considerably more Black individuals in the control group were registered in time for and participated in the 2018 general election than other control individuals.
+This is in line with @Meredith2013 which similarly finds higher post-supervision turnout among Black individuals in New York State.
 Post-treatment, the gap in turnout disappears, but Black individuals discharged after the executive order went into effect were still more likely to be registered than others.
-Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which even those released before the policy change had access.
+Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which they would have otherwise still had access.
+Nevertheless, the success of Executive Order 181 is tempered by these disparate outcomes.
+Not only did the policy fail to boost turnout or registration for some 40 percent of the treatment group (see Table \ref{tab:demo-rd}); it was also precisely the racialized individuals who are most over-represented in the criminal justice system whose post-intervention behavior was left unchanged.
 
@@
 The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
-To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 using different windows.
-In the most conservative window, we include only parolees discharged one month before or after the executive order went into effect (and the same period in 2016).
-As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further distance between parole discharge dates for control and ITT observations.^[The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.]
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 in Table \ref{tab:to-18-logit} using different windows.
+In the most conservative window, I include only individuals discharged one month before or after the executive order went into effect (and the same period in 2016).
+As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further temporal distance between control and ITT observations.^[The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.]
 
@@
 Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95% level in 13 of the 16 models.
-The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
+The estimated effects in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
 This means that voters discharged in the summer before an election are perhaps *less* likely to vote absent treatment.
-This implies our estimated effect of the executive order is perhaps somewhat conservative.
+This implies our estimated effect of the executive order is, if anything, somewhat conservative.
 
@@
 
-The analyses in the previous section indicate that executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
+The analyses in the previous sections indicate that the executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
 They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
 To answer this question, we must specifically control for whether an individual *actually* had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as "compliers." As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
-Some individuals --- namely, noncitizens --- were not eligible for rights restoration, and were thus "noncompliers:" despite being assigned to the treatment group based on their discharge date, they nonetheless did not receive the treatment.
+I consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as "compliers." As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
 
-Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
+A two-stage least squares model allows me to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
 To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via the treatment [@Angrist1996].
-In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
+In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have registered and voted at the same rate in 2018.
 The nonsignificance of the coefficients *Days off Parole* in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit}, and the general nonsignificance of the placebo regressions in Figure \ref{fig:sens} indicate that this condition is met.
@@
 
-The two-stage least squares approach indicates that the CACE was roughly 0.9 percentage points (or 39 percent) for turnout (Model 1) and 1.1 percentage points (or 18 percent) for registration (Model 3).
+The two-stage least squares approach indicates that the CACE was roughly 1.1 percentage points (or 18 percent) for registration (Model 1) and 0.9 percentage points (or 39 percent) for turnout (Model 3).
 These models once again indicate that the treatment effect was substantially smaller, if not nonexistent, for Black individuals.
@@
 
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both *de jure* and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
-Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
+Restoring voting rights to individuals on parole is an important step toward dismantling the disenfranchisement (both *de jure* and *de facto*) of communities of color disproportionately caught up in the criminal justice system.
+Prior to New York's Executive Order 181, individuals were required to wait until they finished their parole term to register to vote.
 That changed in 2018.
-On October 12th --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` active parolees whose voting rights had been restored.
+On October 12th --- the registration deadline for the 2018 midterms --- there were `r comma(readRDS("./temp/new_eligible_18.rds"))` New Yorkers actively on parole whose voting rights had been restored.
 Without the executive order, every single one of these individuals would have been barred from participating.
@@
 
-As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been disenfranchised in its absence.
+As this analysis demonstrates, however, the impact of Executive Order 181 was not limited only to the individuals who would have been legally disenfranchised in its absence.
 In the case of New York State, rights restoration prior to discharge from parole was successful at boosting post-supervision participation rates.
-In this project we estimate that overall turnout among formerly disenfranchised individuals increased by about 0.6 percentage points (or 29 percent), and that individuals whose rights were restored saw turnout about 0.9 percentage points (or 39 percent) above where it would have been otherwise.
-Although the absolute impact of the policy change is small, it is meaningful given the low turnout among these individuals.
+Overall turnout among formerly disenfranchised individuals increased by about 0.6 percentage points (or 29 percent), and individuals whose rights were restored saw turnout about 0.9 percentage points (or 39 percent) above where it would have been otherwise.
+Although the absolute impact of the policy change is small, it is meaningful given the low baseline turnout and registration among these individuals.
 
@@
 Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+On the other hand, an individual whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
 Such interactions may undo some of the negative socialization identified by @Weaver2010.
@@
 Research such as @locked_out further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-Of course, even voters in the control group were receiving notification of their eligibility through the mail.
-If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+When an individual has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+Of course, even voters in the control group were receiving notification of their eligibility upon discharge.
+If the full scope of the causal effect of Executive Order 181 is informational, it seems that information is communicated more effectively through in-person meetings.
 In reality, the executive order's success at boosting turnout and registration likely operated through multiple mechanisms.
 
-Troublingly, the effect of pre-discharge rights restoration seems to vary based on parolees' race.
-AS discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
-Nevertheless, the in-person meeting with parole officers should still have improved their political socialization and relationship with the government.
-Whether this is because the negative socialization of an incarceration is "stickier" for Black voters, because parole officers gave less affirmative encouragement to Black parolees (in other words, they received a weaker treatment), or for some other reason is not clear.
-Ultimately, the data at hand cannot answer why the intervention was unsuccessful at raising turnout and registration among Black former parolees; future research must investigate this.
+Troublingly, the effect of pre-discharge rights restoration varies by race.
+As discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
+As @Morris2020 demonstrates, voters lost to disenfranchisement in New York City come from predominantly Black neighborhoods.
+These dense communities may have been effective at communicating information about eligibility in the pre-treatment period.
+
+Of course, I have hypothesized that the executive order would operate through two causal mechanisms: better information *and* improved relationship with the government.
+Therefore, even if the executive order did not communicate information that Black individuals would not have learned from other sources, the pre-discharge restoration of voting rights should have increased their participation via a different avenue.
+That their participation was unaffected joins other research indicating that incarceration differentially structures turnout for Black and non-Black individuals.
+@White2019, for instance, shows that incarceration reduces turnout more for Black voters, and @Walker2020a argues that Black individuals understand criminal justice involvement from a systemic --- and not individualistic --- perspective.
+If incarceration does more harm to Black individuals' relationship with the state, recouping their turnout may demand more than an in-person meeting to restore voting rights.
 
@@
 This is encouraging, demonstrating that the state has a unique opportunity to shape the future participation of individuals who are currently under their supervision.
-By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform their parolees of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
-Nevertheless, the success of the executive order is tempered by the racially disparate effects.
+By restoring voting rights before individuals have completed their sentence, and by requiring parole officers to inform individuals of their voting rights, the state can increase the political participation of a group of often-marginalized individuals, thereby increasing the democratic representation of our elections.
+Nevertheless, the success of the executive order is tempered by its racially disparate effects.
 
@@
 ```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
-load("./temp/short_term_to.rdata")
+load("temp/short_term_to.rdata")
 
-stargazer(date_to_model_time, date_reg_model_time,
+stargazer(date_reg_model_time, date_to_model_time,
           header = F,
@@
                                "Years on Parole"),
-          dep.var.labels = c("Voted in 2018", "Registered in 2018"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
           title = "\\label{tab:to-18-logit-short} 2018 Behavior for Control Group",
@@
           out.header = F,
-          se = c(date_ses, date_reg_ses))
+          se = c(date_reg_ses, date_ses))
 
@@
                                "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Registered in 2018", "Voted in 2016"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018", "Voted in 2016"),
           title = "\\label{tab:to-18-logit} Turnout in General Election",
@@
 ```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/iv_individual_turnout_18.rdata")
+load("temp/iv_individual_turnout_18.rdata")
 
-stargazer(iv3, iv4, iv3b, iv4b,
+stargazer(iv3b, iv4b, iv3, iv4,
           header = F,
@@
                                "Days off Parole Before Election"),
-          dep.var.labels = c("Voted in 2018", "Registered in 2018"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
           title = "\\label{tab:iv-tab} Two-Stage Least Squares",
@@
           out.header = F,
-          se = list(iv3_ses, iv4_ses, iv3_sesb, iv4_sesb))
+          se = list(iv3_sesb, iv4_sesb, iv3_ses, iv4_ses))
 

### 2020-10-27 markdown/part2/rewrite.tex
ssq r&r

@@
 \usepackage{setspace}
+\usepackage{lineno}
+\linenumbers
 \usepackage{booktabs}
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{October 09, 2020}
+\date{October 27, 2020}
 
@@
 This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
-In 2018, New York State began restoring voting rights to parolees, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, we interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to cast a ballot.
-We demonstrate that in-person rights restoration prior to parole discharge significantly increased turnout.
-This group-level effect, however, masks race-specific effects.
-Although rights restoration prior to parole discharge increased turnout by non-Black former parolees by 1.4 percentage points, the treatment effect was substantially smaller for Black former parolees.
-This raises serious questions about how rights restoration programs are implemented, and how incarceration might differently structure Black Americans' view of the democratic process.
+In 2018, New York State began restoring voting rights to individuals on parole, after previously restoring voting rights only at the completion of parole.
+By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to register and cast a ballot.
+I demonstrate that in-person rights restoration prior to parole discharge significantly increased participation, but this group-level effect masks racially heterogeneous effects.
+Although rights restoration prior to parole discharge increased turnout by non-Black individuals by 1.4 percentage points, the treatment effect was substantially smaller for Black individuals.
+This raises important questions about how members of different communities can be brought back into our democratic processes.
 \end{abstract}
@@
 
-Collectively, felony disenfranchisement laws actively disenfranchise as many as 4.7 million American citizens.
-Of these, the majority are no longer incarcerated (Uggen, Larson, and Shannon \protect\hyperlink{ref-sentencing_2016}{2016}).\footnote{The figures reported in Uggen, Larson, and Shannon (\protect\hyperlink{ref-sentencing_2016}{2016}) have been adjusted to reflect the potential impact of Amendment 4 in Florida.} There is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
+Collectively, felony disenfranchisement laws actively disenfranchise as many as 5.2 million American citizens.
+Of these, the majority are no longer incarcerated, but are either on parole, probation, or are disenfranchised after completing their sentence (Uggen et al.
+\protect\hyperlink{ref-Uggen2020}{2020}).
+The effects of felony disenfranchisement, however, might be even broader: there is some evidence that incarceration continues to structure political participation even after an individual is no longer legally disenfranchised.
 Some literature argues that interactions with the criminal justice system leaves residents less likely to vote in the future (Weaver and Lerman \protect\hyperlink{ref-Weaver2010}{2010}; White \protect\hyperlink{ref-White2019}{2019}), though other research casts some doubt on this conclusion (Gerber et al.
@@
 
-In the spring of 2018,\footnote{Although the executive order was signed on April 18th, it did not go into effect until May 18th.} Governor Andrew Cuomo signed Executive Order 181 which effectively ended the disenfranchisement of New Yorkers on parole.
+In the spring of 2018,\footnote{Although the executive order was signed on April 18th, it did not go into effect until May 18th.} Governor Andrew Cuomo signed Executive Order 181 which restored voting rights to most New Yorkers on parole.
 The policy change directly re-enfranchised individuals who were still on parole on election day, but the consequences might be further reaching.
-The change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
+For instance, the change in policy is also perhaps beneficial for felony probationers: despite the fact that probationers do not formally lose their voting rights, there is evidence that confusion around the law contributes to \emph{de facto} disenfranchisement among probationers (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}).
 By re-enfranchising anyone not actively incarcerated, the rules were substantially simplified.
@@
 Beginning in 2010, New York State began including information about voter eligibility in parole discharge packets, a change that Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) indicates did not increase turnout.
-Executive Order 181 both changed the timing of rights restoration (upon release from prison, not the end of parole) and the manner of restoration (via an in-person meeting, rather than in a discharge packet).
-This study tests the bundled effect of these changes.
+Executive Order 181 both changed the \emph{timing} of rights restoration (upon release from prison, not the end of parole) and the \emph{manner} of restoration (via an in-person meeting, rather than in a discharge packet).
+This study tests the bundled effect of these changes on individuals whose legal right to vote was not changed by the executive order because they were discharged from parole before the registration deadline for the 2018 midterm elections.
 
-Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{See \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}} The memorandum directs all parole officers to present parolees with voter registration forms and to explain their purpose.
+Shortly after the implementation of the executive order, Acting Deputy Commissioner for Community Supervision Ana Enright sent a memorandum to New York State parole officers detailing the Department of Corrections' new approach.\footnote{See \url{https://nyassembly.gov/member_files/139/webdocs/82103.pdf}.} The memorandum directs all parole officers to present individuals on parole with voter registration forms and to explain their purpose.
 Parole officers are instructed to offer any assistance needed, including help filling out the registration form.
@@
 Regardless of the causal impact of incarceration on turnout, however, it is well established that formerly incarcerated individuals turn out at relatively low rates (see, for instance, Burch \protect\hyperlink{ref-Burch2011}{2011}).
-Regardless of whether incarceration reduces turnout, the state has a unique opportunity to craft policies that will impact individuals under formal supervision.
-Even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
+Thus, even if incarceration does not lead to lower turnout, policies targeting individuals caught up in the criminal justice system might still be effective at increasing their turnout, both while serving their sentence and afterwards.
 
@@
 In the three states studied in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}), however, newly re-enfranchised individuals received this information as part of a discharge packet.
-Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform former parolees that their voting rights had been restored.
-Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2008}), and it is possible that these brochures, in-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
+Moreover, in the case of New York, the brochure included in the discharge packet did not explicitly inform individuals that their voting rights had been restored.
+Formerly convicted individuals are very often confused about their eligibility to vote (Drucker and Barreras \protect\hyperlink{ref-Drucker2005}{2005}; Manza and Uggen \protect\hyperlink{ref-locked_out}{2008}), and it is possible that these brochures, non-explicit and bundled with other documents as they were, were ineffective at dispelling misinformation.
 
@@
 
-Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether a parolee had her voting rights restored.
-By using the identification number provided from the parolee public records request and this website, we identify individuals who had their voting rights restored.\footnote{Not all parolees listed in the public records request data are included in the lookup tool.
-Roughly 1 percent of former parolees are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
+Following Executive Order 181, the Department of Corrections and Community Supervision began indicating on their online Parolee Lookup Tool whether an individual had her voting rights restored.
+By using the identification number provided from the public records request and this website, I identify individuals who had their voting rights restored.\footnote{Not all individuals listed in the public records request data are included in the lookup tool.
+Roughly 1 percent are not in the online lookup.} There were 3,093 individuals who were discharged from parole before the registration deadline whose rights were restored while still under supervision.
 Roughly 1,200 individuals who were discharged between the effective date of the executive order and the registration deadline did not have voting rights restored, due largely to their status as noncitizens.
@@
 
-In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. We use a snapshot of the registered voter file from March 3rd, 2019.
+In New York, the registered voter file includes first, middle, and last name; date of birth; vote history; and other information. I use a snapshot of the registered voter file from March 3rd, 2019.
 
@@
 Turnout in the 2018 midterm election is estimated by matching the parole records with the registered voter file.
-We match individuals in each dataset using first name, middle name, last name, and date of birth.
-To be considered a ``match,'' records must have the exact same birth date.
+I match individuals in each dataset using first name, middle name, last name, and date of birth.
+To be considered a match, records must have the exact same birth date.
 The first and last names must also be exact matches (conditional on the adjustments discussed below).
@@
 
-We adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
-We slightly alter the date of birth reported in the parole discharge dataset to create false records.
+I adopt the test developed in Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) to test the prevalence of false positives.
+I slightly alter the date of birth reported in the parole discharge dataset to create false records.
 Comparing the number of matches between these ``fake'' discharge records and the voter file with the number of matches between the ``true'' records and the voter file provides an estimate of how frequently false positives occur.
-Table \ref{tab:change-dobs} shows the results of true matches, as well as when we construct a set of fake records by adding or subtracting 35 days from a parolee's birthdate.
-This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is likely too small to have any material impact on the overall analysis.
+Table \ref{tab:change-dobs} shows the results of true matches, as well as when I construct a set of fake records by adding or subtracting 35 days from an individual's birthdate.
+This analysis indicates that false positives account for between 0.6 and 0.7 percent of all matches, a share that is probably too small to have any material impact on the overall analysis.
 
@@
 Testing for false negatives is more challenging.
-If an individual marries and changes her name after being discharged from parole, for instance, we will not identify her using this matching methodology.
+If an individual marries and changes her name after being discharged from parole, for instance, I will not identify her using this matching methodology.
 Similarly, ``John Doe'' and ``Jonathan Doe'' would not result in a match.
-To reduce the likelihood of these false negatives we remove all punctuation from all names, and standardize capitalization.
+To reduce the likelihood of these false negatives I remove all punctuation from all names, and standardize capitalization.
 A record with a last name of ``O'Donnell'' in one dataset, therefore, would match a last name of ``O DONNELL'' in the other (provided the other criteria are satisfied).
 Such standardizations, however, will miss individuals who change their names entirely.
-For three reasons, however, this is not likely to present major challenges: firstly, women are far more likely to change their last names than men, and women make up barely 6 percent of individuals who have been discharged from felony parole.
-Secondly, because both parolee discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
-Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the Executive Order went into effect.
+For three reasons this is not expected to present major challenges: firstly, far more women change their last names than men, and women make up less than 10 percent of individuals who have been discharged from felony parole.
+Secondly, because both discharge and voter registration are legal records, individuals are likely to be recorded using their full names (that is to say, an individual is unlikely to be ``John'' in one set of records and ``Jonathan'' in the other).
+Finally, rates of false negatives are likely to be constant within the state during the study period, and there is no reason to believe that these false negatives would be associated with being discharged from parole after the executive order went into effect.
 
@@
 
-We exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change.
+I exploit randomness in discharge date from parole in New York State to determine the efficacy of the policy change, measured both using turnout in the 2018 election, and registration status as of the registration deadline for the same contest.
 All individuals discharged from parole after the effective date of the executive order are assigned to the treatment group; individuals discharged earlier are assigned to the control group.
 When we limit the pool of formerly incarcerated individuals to those discharged from parole in 2017 and 2018, the control and treatment voters look virtually identical along observable characteristics (see Table \ref{tab:demo-rd}).
-The assignment to treatment, therefore, is as nearly random as can be hoped for in the context of a natural experiment.
-The basic design is therefore an interrupted time series, in which we test whether individuals discharged after the executive order took effect registered and turned out at higher rates than those discharged before the policy change.
-This allows us to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall registration and turnout among formerly incarcerated individuals.
+The assignment to treatment is as nearly random as can be hoped for in the context of a natural experiment.
+The basic design is therefore an interrupted time series, in which I test whether individuals discharged after the executive order took effect registered and turned out at higher rates than those discharged before the policy change.
+This allows me to test the intention-to-treat (ITT) effect of the executive order; in other words, the extent to which the policy change impacted overall registration and turnout among formerly incarcerated individuals.
 
-As discussed above, not all formerly incarcerated individuals were eligible to have their rights restored; the ITT effect will therefore be biased towards zero, insofar as the behavior of ineligible parolees remains unchanged.
-After estimating the ITT effect, we leverage the restoration records to estimate the complier average causal effect (CACE).
-In this setup, actual rights restoration prior to parole discharge is instrumented by whether an individual was assigned to the treatment group (that is to say, was discharged after the executive order went into effect).
-This instrumentation process is a common method for identifying the CACE (see, for instance, Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
-The CACE will provide the causal estimate of rights restoration prior to discharge from parole for the individuals actually affected by the policy change.
+Not all formerly incarcerated individuals were eligible to have their rights restored.
+Noncitizens, for instance, and individuals who were re-arrested while on parole did not have their voting rights restored.
+Individuals who fall into these categories and were discharged after the executive order are nonetheless included in the ITT group, even though they were not in fact treated (put differently, they did not ``comply'' with being assigned to the treatment group).
+Because some individuals in the ITT group went untreated, the ITT effect will be biased towards zero.
+The ITT therefore cannot be understood as the causal effect of rights restoration on the turnout of the individuals whose rights were actually restored.
+
+Two-stage least squares approaches are commonly used to estimate the causal effect of treatment in a scenario with ``imperfect compliance'' (see, for instance, Ansolabehere, Iyengar, and Simon \protect\hyperlink{ref-Ansolabehere1999}{1999}; Gerber and Green \protect\hyperlink{ref-Gerber2000}{2000}; Milligan, Moretti, and Oreopoulos \protect\hyperlink{ref-Milligan2004}{2004}; Lassen \protect\hyperlink{ref-Lassen2004}{2004}; Sondheimer and Green \protect\hyperlink{ref-Sondheimer2010}{2010}).
+By using the random assignment to treatment as an instrument for actual treatment, the complier average causal effect (CACE) can be estimated.
+Here, whether someone was discharged from parole after the executive order went into effect (that is, whether someone was assigned to the treatment group) instruments whether their rights were actually restored (whether they ``complied'' with being treated).
+Of course, using terms like ``compliance'' here is slightly misleading: the individuals who did not ``comply'' were legally barred from having voting rights restored (or, in the case of noncitizens, conferred) --- they had little influence over whether or not they were actually treated, after being assigned to the treatment group.
+Nevertheless, I use this nomenclature to make the ties to existing methodological research clear.
 
@@
 
-Before estimating the interrupted time series model, we establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
-It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote, regardless of whether their voting rights are restored before they finish their sentence.
+Before estimating the interrupted time series model, I establish that any change in turnout between those discharged before and after the policy went into effect is not due to underlying trends.
+It is possible, for instance, that individuals discharged from parole shortly before an election are more likely to vote irrespective of policy intervention.
 
 Table \ref{tab:to-18-logit-short} presents an ordinary least squares regression using only the control voters.
-The variable \emph{Days off Parole Before Election} indicates the number of days between a parolee's discharge date and the 2018 election.
-As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout / registration.
+The variable \emph{Days off Parole Before Election} indicates the number of days between an individual's discharge date and the 2018 election.
+As the table makes clear, for individuals discharged between January 1, 2017, and May 17, 2018 --- the day before the executive order went into effect --- there was no relationship between time-off-parole and 2018 turnout and registration.
 
@@
 Having established the validity of the interrupted time series model, Table \ref{tab:to-18-logit} presents the results of OLS specifications exploring whether individuals who were discharged on or after May 18th, 2018, registered and turned out at higher rates than those discharged earlier.
-Models 1 and 2 test turnout in the 2018 midterm election, while Models 3 and 4 ask whether former parolees in the treatment group were more likely to be registered prior to the registration deadline.
-Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black former parolees than others.
+Models 1 and 2 ask whether individuals in the treatment group were more likely to be registered prior to the registration deadline, while Models 3 and 4 test turnout in the 2018 midterm election.
+Models 1 and 3 estimate the overall treatment effect, while Models 2 and 4 investigate whether the treatment effect was different for Black individuals than others.
 
-In Model 5 we estimate a ``placebo'' regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
+In Model 5 I estimate a ``placebo'' regression as a final check against the possibility that individuals discharged in the summer of an election year turn out at higher rates.
 Model 5 includes all individuals last discharged between January 1, 2015, and October 14th, 2016.
@@
 The interaction effect --- \emph{Black × Discharged After EO 181} --- is negative, large relative to the general treatment effect, and statistically significant in both models.
-The executive order was apparently substantially more effective at boosting registration and turnout for non-Black former parolees than Black parolees.
+The executive order was apparently substantially more effective at boosting registration and turnout for non-Black than Black individuals.
 
 Why would the intervention have increased turnout and registration among non-Black individuals and have had no effect on Black individuals?
-This difference may be explained in part through the causal mechanisms we have argued are at play in the case of Executive Order 181.
+This difference may be explained in part through the causal mechanisms I have argued are at play in the case of Executive Order 181.
 As discussed above, many formerly incarcerated individuals are unsure about their legal eligibility to cast a ballot.
-It is possible that prior to the policy change formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
-This seems plausible.
-Morris (\protect\hyperlink{ref-Morris2020}{2020}) looks at the geographic distribution of disenfranchised individuals in New York City with a history of voting, finding that voters lost to disenfranchisement are clustered in neighborhoods with large Black populations.
-It is possible that Black individuals return home from prison to neighborhoods with more formerly incarcerated individuals, and therefore have access to communities with better information about voting eligibility.
+It is possible that, prior to the policy change, formerly incarcerated Black individuals were more aware of their post-supervision voting rights.
 This would reduce the marginal efficacy of the information-communicating aspect of Executive Order 181.
 
-If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the pre-treatment period.
+If this were the case, we would expect to see higher turnout and registration rates for Black individuals in the control group.
 And, indeed, that is precisely what we observe.
-Table \ref{tab:re} Black individuals in the control group were far more likely to be registered in time for and participate in the 2018 general election than other control individuals.
+Table \ref{tab:re} shows that considerably more Black individuals in the control group were registered in time for and participated in the 2018 general election than other control individuals.
+This is in line with Meredith and Morse (\protect\hyperlink{ref-Meredith2013}{2013}) which similarly finds higher post-supervision turnout among Black individuals in New York State.
 Post-treatment, the gap in turnout disappears, but Black individuals discharged after the executive order went into effect were still more likely to be registered than others.
-Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which even those released before the policy change had access.
+Although this does not rule out other potential explanations, it is possible that the executive order did not increase the information available to Black individuals because it was information to which they would have otherwise still had access.
+Nevertheless, the success of Executive Order 181 is tempered by these disparate outcomes.
+Not only did the policy fail to boost turnout or registration for some 40 percent of the treatment group (see Table \ref{tab:demo-rd}); it was also precisely the racialized individuals who are most over-represented in the criminal justice system whose post-intervention behavior was left unchanged.
 
@@
 The models in Table \ref{tab:to-18-logit} leverage a large pool of voters on either side of the effective date of the executive order; it is perhaps implausible that voters discharged in the early months of 2017 are truly good controls for those discharged in the summer of 2018, despite the above diagnostics.
-To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 using different windows.
-In the most conservative window, we include only parolees discharged one month before or after the executive order went into effect (and the same period in 2016).
-As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further distance between parole discharge dates for control and ITT observations.\footnote{The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.}
+To explore whether these results are sensitive to the window used around the effective date, Figure \ref{fig:sens} re-estimates Models 3 and 5 in Table \ref{tab:to-18-logit} using different windows.
+In the most conservative window, I include only individuals discharged one month before or after the executive order went into effect (and the same period in 2016).
+As the window increases, we gain more observations --- and therefore statistical power --- at the cost of further temporal distance between control and ITT observations.\footnote{The post-treatment window never extends past the 2018 and 2016 registration deadlines; the estimates at the far right of Figure \ref{fig:sens}, therefore, are only extending the window backwards in time.}
 
@@
 Although a small handful of the windows in the 2018 regressions are statistically nonsignificant, the size of the estimated treatment effect is stable and significant at the 95\% level in 13 of the 16 models.
-The estimated effect in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
+The estimated effects in the placebo regressions, on the other hand, are consistently nonsignificant (in 10 specifications) or negative (in the remaining 6 specifications).
 This means that voters discharged in the summer before an election are perhaps \emph{less} likely to vote absent treatment.
-This implies our estimated effect of the executive order is perhaps somewhat conservative.
+This implies our estimated effect of the executive order is, if anything, somewhat conservative.
 
@@
 
-The analyses in the previous section indicate that executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
+The analyses in the previous sections indicate that the executive order was successful at increasing turnout and registration among formerly incarcerated individuals as a group.
 They do not, however, provide us an estimate of the effect of rights restoration prior to parole discharge on the specific individuals who were eligible for treatment.
 To answer this question, we must specifically control for whether an individual \emph{actually} had his rights restored before he was discharged from parole --- not simply whether he was discharged after the policy change.
-We consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as ``compliers.'' As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
-Some individuals --- namely, noncitizens --- were not eligible for rights restoration, and were thus ``noncompliers:'' despite being assigned to the treatment group based on their discharge date, they nonetheless did not receive the treatment.
+I consider individuals discharged on or after May 18 whose rights were in fact restored prior to discharge from parole as ``compliers.'' As mentioned above, some 1,200 individuals who were discharged after the policy went into effect did not have their voting rights restored.
 
-Two-stage least squares models allow us to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
+A two-stage least squares model allows me to leverage the random assignment to treatment to identify the complier average causal effect (CACE).
 To use assignment to treatment as an instrument for compliance, assignment to treatment group must be uncorrelated with the outcome variable except via the treatment (Angrist, Imbens, and Rubin \protect\hyperlink{ref-Angrist1996}{1996}).
-In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have voted at the same rate in 2018.
+In other words, we can only uncover the CACE in this setup if, in the absence of the policy change, individuals discharged from parole before and after May 18 would have registered and voted at the same rate in 2018.
 The nonsignificance of the coefficients \emph{Days off Parole} in Tables \ref{tab:to-18-logit-short} and \ref{tab:to-18-logit}, and the general nonsignificance of the placebo regressions in Figure \ref{fig:sens} indicate that this condition is met.
@@
 
-Restoring voting rights to individuals on parole is an important step toward undermining the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
-Prior to New York's Executive Order 181, parolees were required to wait until they finished their parole term to register to vote.
+Restoring voting rights to individuals on parole is an important step toward dismantling the disenfranchisement (both \emph{de jure} and \emph{de facto}) of communities of color disproportionately caught up in the criminal justice system.
+Prior to New York's Executive Order 181, individuals were required to wait until they finished their parole term to register to vote.
 That changed in 2018.
-On October 12th --- the registration deadline for the 2018 midterms --- there were 21,863 active parolees whose voting rights had been restored.
+On October 12th --- the registration deadline for the 2018 midterms --- there were 21,863 New Yorkers actively on parole whose voting rights had been restored.
 Without the executive order, every single one of these individuals would have been barred from participating.
@@
 Automatic voting rights restoration upon the completion of a sentence likely does little to combat these negative perceptions of the government.
-On the other hand, a parolee whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
+On the other hand, an individual whose parole officer actively encourages them to register and participate may believe that the state is interested in their political participation.
 Such interactions may undo some of the negative socialization identified by Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010}).
@@
 Research such as Manza and Uggen (\protect\hyperlink{ref-locked_out}{2008}) further demonstrates that many formerly incarcerated individuals wrongly believe that they are ineligible to participate.
-When a parolee has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
-Of course, even voters in the control group were receiving notification of their eligibility through the mail.
-If the full scope of the causal effect of Executive Order 181 is informational, it is clear that information is communicated far more effectively through in-person meetings.
+When an individual has her voting rights restored prior to discharge --- and when her parole officer is required to inform her of that fact --- she is far more likely to be confident in her voting eligibility.
+Of course, even voters in the control group were receiving notification of their eligibility upon discharge.
+If the full scope of the causal effect of Executive Order 181 is informational, it seems that information is communicated more effectively through in-person meetings.
 In reality, the executive order's success at boosting turnout and registration likely operated through multiple mechanisms.
 
-Troublingly, the effect of pre-discharge rights restoration seems to vary based on parolees' race.
-AS discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
-Nevertheless, the in-person meeting with parole officers should still have improved their political socialization and relationship with the government.
-Whether this is because the negative socialization of an incarceration is ``stickier'' for Black voters, because parole officers gave less affirmative encouragement to Black parolees (in other words, they received a weaker treatment), or for some other reason is not clear.
-Ultimately, the data at hand cannot answer why the intervention was unsuccessful at raising turnout and registration among Black former parolees; future research must investigate this.
+Troublingly, the effect of pre-discharge rights restoration varies by race.
+As discussed in the body of this manuscript, this is possibly because one of the causal routes through which the executive order was expected to operate (namely, through improved information communication) had less purchase for Black voters, who may have already been well-informed about their rights.
+As Morris (\protect\hyperlink{ref-Morris2020}{2020}) demonstrates, voters lost to disenfranchisement in New York City come from predominantly Black neighborhoods.
+These dense communities may have been effective at communicating information about eligibility in the pre-treatment period.
+
+Of course, I have hypothesized that the executive order would operate through two causal mechanisms: better information \emph{and} improved relationship with the government.
+Therefore, even if the executive order did not communicate information that Black individuals would not have learned from other sources, the pre-discharge restoration of voting rights should have increased their participation via a different avenue.
+That their participation was unaffected joins other research indicating that incarceration differentially structures turnout for Black and non-Black individuals.
+White (\protect\hyperlink{ref-White2019}{2019}), for instance, shows that incarceration reduces turnout more for Black voters, and Walker (\protect\hyperlink{ref-Walker2020a}{2020}) argues that Black individuals understand criminal justice involvement from a systemic --- and not individualistic --- perspective.
+If incarceration does more harm to Black individuals' relationship with the state, recouping their turnout may demand more than an in-person meeting to restore voting rights.
 
@@
 
-\leavevmode\hypertarget{ref-sentencing_2016}{}%
-Uggen, Christopher, Ryan Larson, and Sarah Shannon.
-2016.
-``6 Million Lost Voters: State-Level Estimates of Felony Disenfranchisement, 2016.'' Research report.
+\leavevmode\hypertarget{ref-Uggen2020}{}%
+Uggen, Christopher, Ryan Larson, Sarah Shannon, and Arleth Pulido-Nava.
+2020.
+``Locked Out 2020: Estimates of People Denied Voting Rights Due to a Felony Conviction.'' Research report.
 The Sentencing Project.
-\url{https://www.sentencingproject.org/publications/6-million-lost-voters-state-level-estimates-felony-disenfranchisement-2016/}.
+\url{https://www.sentencingproject.org/wp-content/uploads/2020/10/Locked-Out-2020.pdf}.
+
+\leavevmode\hypertarget{ref-Walker2020a}{}%
+Walker, Hannah L.
+2020.
+\emph{Mobilized by Injustice: Criminal Justice Contact, Political Participation, and Race}.
+1st ed.
+Oxford University Press.
+\url{https://doi.org/10.1093/oso/9780190940645.001.0001}.
 

### 2020-10-27 code/part_2/06_individual_turnout.R

@@ -28,6 +28,8 @@ pool <- filter(parolees, year(parole_status_date) >= 2017) %>%
 g <- t.test(v2018 ~ b, filter(pool, !finished_post))
 h <- t.test(v2018 ~ b, filter(pool, finished_post))
 
+g1 <- t.test(v2008 ~ b, filter(pool))
+
 i <- t.test(registered ~ b, filter(pool, !finished_post))
 j <- t.test(registered ~ b, filter(pool, finished_post))
 
@@ -39,6 +41,13 @@ k <- pool %>%
   pivot_wider(id_cols = c(name, finished_post), values_from = c(value), names_from = "b") %>% 
   mutate(finished_post = ifelse(finished_post, "ITT", "Control")) %>% 
   arrange(desc(name), finished_post)
+
+k2 <- pool %>% 
+  group_by(b) %>% 
+  summarize(`Turnout 2008` = percent(mean(v2008), accuracy = .01)) %>% 
+  pivot_longer(cols = c(`Turnout 2008`)) %>% 
+  pivot_wider(id_cols = c(name), values_from = c(value), names_from = "b") %>% 
+  arrange(desc(name))
 
 k$p <- round(c(g$p.value, h$p.value, i$p.value, j$p.value), digits = 3)
 
@@ -56,19 +65,19 @@ parolees$pb <- (parolees$race == "BLACK") * (parolees$finished_post == T)
 
 rd <- filter(parolees, year(parole_status_date) >= 2017)
 
-f1 <- v2018 ~ finished_post + as.factor(race) +
+f1 <- registered ~ finished_post + as.factor(race) +
   as.factor(sex) + age + v2008 +
   felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef
 
-f2 <- v2018 ~ finished_post + as.factor(race) +
+f2 <- registered ~ finished_post + as.factor(race) +
   as.factor(sex) + age + v2008 +
   felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef + pb
 
-f3 <- registered ~ finished_post + as.factor(race) +
+f3 <- v2018 ~ finished_post + as.factor(race) +
   as.factor(sex) + age + v2008 +
   felony_a + felony_b + felony_c + felony_d + felony_e + parole_time + days_bef
 
-f4 <- registered ~ finished_post + as.factor(race) +
+f4 <- v2018 ~ finished_post + as.factor(race) +
   as.factor(sex) + age + v2008 +
[…]

### 2020-10-28 markdown/part2/table_embed.Rmd
ssq abstract

--- 2020-10-27 markdown/part2/rewrite.Rmd
@@
 abstract: |
-  This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
-In 2018, New York State began restoring voting rights to individuals on parole, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to register and cast a ballot.
-I demonstrate that in-person rights restoration prior to parole discharge significantly increased participation, but this group-level effect masks racially heterogeneous effects.
-Although rights restoration prior to parole discharge increased turnout by non-Black individuals by 1.4 percentage points, the treatment effect was substantially smaller for Black individuals.
-This raises important questions about how members of different communities can be brought back into our democratic processes.
+  Objective: Past research demonstrates that formerly incarcerated individuals turn out at very low rates.
+Executive Order 181 in New York State allows us to test whether restoring voting rights at an in-person meeting while an individual is still on parole can increase their post-supervision registration and turnout.
+
+  Methods: By linking administrative parole records with the registered voter file I estimate individual-level turnout.
+I use an interrupted time series and two-stage least squares approach to estimate the causal effect of Executive Order 181 on post-supervision participation.
+
+  Results: I find that Executive Order 181 increased post-supervision registration and turnout for individuals discharged from parole as a whole; these results, however, mask racial heterogeneity.
+Although the participation of non-Black individuals increased substantially, the treatment effect for Black individuals was either much smaller or nonexistent.
+
+  Conclusion: State policy can meaningfully increase the political participation of individuals even after they are no longer on parole.
+However, at least in the case of New York State, rights restoration prior to parole discharge appears less effective for Black individuals, who make up roughly 40 percent of the formerly incarcerated population.
 ---
@@
 
-TABLE 1 ABOUT HERE
+\begin{singlespace}
+```{r shift-dobs-chunk, echo = F}
+less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
+true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
+plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
+
+table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
+
+j <- table %>% 
+  mutate_all(linebreak) %>% 
+  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
+               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
+               align = "c") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                position = "center")
+j
+
+```
+\end{singlespace}
 
@@
 
-TABLE 2 ABOUT HERE
+\begin{singlespace}
+```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
+load("temp/short_term_to.rdata")
+
+stargazer(date_reg_model_time, date_to_model_time,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
+                               "Voted in 2008",
+                               "Male", "Age (Years)",
+                               "Years on Parole"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
+          title = "\\label{tab:to-18-logit-short} 2018 Behavior for Control Group",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
+                         c("Felony Class FE", "X", "X")),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = c(date_reg_ses, date_ses))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table3.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table3.tex"}
+\end{singlespace}
 
@@
 
-TABLE 3 ABOUT HERE
+\begin{singlespace}
+```{r rd-demo-chunk, echo = F, warning = F}
+t <- readRDS("./temp/demos_rd_time.rds")
+colnames(t) <- c("Variable", "Control", "ITT", "p-value")
+j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
+                  align = c("l", "c", "c"), "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F)
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- ""
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table_whatever2.tex"}
+\end{singlespace}
 
@@
 
-TABLE 4 ABOUT HERE
+\begin{singlespace}
+```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("./temp/individual_turnout_18_itt.rdata")
+load("./temp/individual_turnout_16.rdata")
+
+ses_cl[["m4"]] <- m4_ses
+
+stargazer(models, model4,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Discharged After EO 181",
+                               "Black $\\times$ Discharged After EO 181",
+                               "Male", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018", "Voted in 2016"),
+          title = "\\label{tab:to-18-logit} Turnout in General Election",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X", "X"),
+                         c("Felony Class FE", "X", "X", "X", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          order = c("finished_post", "pb"),
+          se = ses_cl)
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{6}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table4.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table4.tex"}
+\end{singlespace}
 
@@
 
-TABLE 5 ABOUT HERE
+\begin{singlespace}
+```{r rd-dmeans, echo = F, warning = F}
+k <- readRDS("./temp/means_diff_black.rds") %>% 
+  select(-Statistic)
+j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout and Registration by Treatment Group, Race", linesep = "",
+                  align = c("l", "c", "c", "c")) %>% 
+  kable_styling(latex_options = c("HOLD_position"),
+                full_width = F) %>% 
+  group_rows(start_row = 1,
+             end_row = 2,
+             group_label = "Turnout") %>% 
+  group_rows(start_row = 3,
+             end_row = 4,
+             group_label = "Registration")
+
+write(j, "./temp/tempo.tex")
+
+j <- fread("./temp/tempo.tex", sep = "+", header = F)
+
+note.latex <- ""
+
+j <- j %>% 
+  mutate(n = row_number())
+
+lab <- "\\caption{\\label{tab:re} Turnout and Registration by Treatment Group, Race}"
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+j$V1[4] <- lab
+
+write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
+            row.names = F)
+```
+\input{"../../temp/table_whatever3.tex"}
+\end{singlespace}
 
@@
 
-FIGURE 1 ABOUT HERE
+```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
+j <- readRDS("./temp/reestimates.rds")
+print(j)
+```
 
@@
 
-TABLE 6 ABOUT HERE
+\begin{singlespace}
+```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
+load("temp/iv_individual_turnout_18.rdata")
+
+stargazer(iv3b, iv4b, iv3, iv4,
+          header = F,
+          type = "text", notes.align = "l",
+          covariate.labels = c("Rights Restored Before Discharge", 
+                               "Black $\\times$ Rights Restored Before Discharge",
+                               "Male", "Age (Years)",
+                               "Voted in 2008",
+                               "Years on Parole",
+                               "Days off Parole Before Election"),
+          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
+          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
+          table.placement = "H",
+          omit = c("race", "felony"),
+          keep.stat = c("n", "rsq", "adj.rsq"),
+          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X"),
+                         c("Felony Class FE", "X", "X", "X", "X")),
+          table.layout = "-dm#-t-a-s-n",
+          out = "./temp/test.tex",
+          notes = "TO REPLACE",
+          out.header = F,
+          se = list(iv3_sesb, iv4_sesb, iv3_ses, iv4_ses))
+
+j <- fread("./temp/test.tex", header = F, sep = "+")
+
+note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
+\\\\Robust standard errors in parentheses.\\\\First-stage equations include covariates in Models 2 and 3.}}}"
+
+j <- j %>% 
+  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
+         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
+  filter(!grepl("Note:", V1)) %>% 
+  mutate(n = row_number())
+
+insert1 <- "\\fontsize{10}{12}\\selectfont"
+
+j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
+  arrange(n) %>% 
+  select(-n)
+
+write.table(j, "./temp/table6.tex", quote = F, col.names = F,
+            row.names = F)
+
+```
+\input{"../../temp/table6.tex"}
+\end{singlespace}
 
@@
 # References {-}
-<div id="refs"></div>
-
-\newpage
-
-\begin{singlespace}
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
-true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
-plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
-load("temp/short_term_to.rdata")
-
-stargazer(date_reg_model_time, date_to_model_time,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
-                               "Voted in 2008",
-                               "Male", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
-          title = "\\label{tab:to-18-logit-short} 2018 Behavior for Control Group",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
-                         c("Felony Class FE", "X", "X")),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = c(date_reg_ses, date_ses))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table3.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table3.tex"}
-```{r rd-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_rd_time.rds")
-colnames(t) <- c("Variable", "Control", "ITT", "p-value")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- ""
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever2.tex"}
-```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
-load("./temp/individual_turnout_16.rdata")
-
-ses_cl[["m4"]] <- m4_ses
-
-stargazer(models, model4,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Discharged After EO 181",
-                               "Black $\\times$ Discharged After EO 181",
-                               "Male", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018", "Voted in 2016"),
-          title = "\\label{tab:to-18-logit} Turnout in General Election",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X", "X"),
-                         c("Felony Class FE", "X", "X", "X", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          order = c("finished_post", "pb"),
-          se = ses_cl)
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{6}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table4.tex"}
-```{r rd-dmeans, echo = F, warning = F}
-k <- readRDS("./temp/means_diff_black.rds") %>% 
-  select(-Statistic)
-j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout and Registration by Treatment Group, Race", linesep = "",
-                  align = c("l", "c", "c", "c")) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F) %>% 
-  group_rows(start_row = 1,
-             end_row = 2,
-             group_label = "Turnout") %>% 
-  group_rows(start_row = 3,
-             end_row = 4,
-             group_label = "Registration")
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- ""
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:re} Turnout and Registration by Treatment Group, Race}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever3.tex"}
-```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("temp/iv_individual_turnout_18.rdata")
-
-stargazer(iv3b, iv4b, iv3, iv4,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Rights Restored Before Discharge", 
-                               "Black $\\times$ Rights Restored Before Discharge",
-                               "Male", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
-          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X"),
-                         c("Felony Class FE", "X", "X", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(iv3_sesb, iv4_sesb, iv3_ses, iv4_ses))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.\\\\First-stage equations include covariates in Models 2 and 3.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table6.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table6.tex"}
-\end{singlespace}
-
-\newpage
-```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/reestimates.rds")
-print(j)
-```

### 2020-10-28 markdown/part2/table_embed.tex
ssq abstract

--- 2020-10-27 markdown/part2/rewrite.tex
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{October 27, 2020}
+\date{October 28, 2020}
 
@@
 \begin{abstract}
-This paper presents causal estimates of the effect of voting rights restoration prior to discharge from parole on post-supervision participation.
-In 2018, New York State began restoring voting rights to individuals on parole, after previously restoring voting rights only at the completion of parole.
-By leveraging randomness in parole discharge date, I interrogate whether restoring voting rights to formerly incarcerated individuals while still on parole increases their post-supervision propensity to register and cast a ballot.
-I demonstrate that in-person rights restoration prior to parole discharge significantly increased participation, but this group-level effect masks racially heterogeneous effects.
-Although rights restoration prior to parole discharge increased turnout by non-Black individuals by 1.4 percentage points, the treatment effect was substantially smaller for Black individuals.
-This raises important questions about how members of different communities can be brought back into our democratic processes.
+Objective: Past research demonstrates that formerly incarcerated individuals turn out at very low rates.
+Executive Order 181 in New York State allows us to test whether restoring voting rights at an in-person meeting while an individual is still on parole can increase their post-supervision registration and turnout.
+
+Methods: By linking administrative parole records with the registered voter file I estimate individual-level turnout.
+I use an interrupted time series and two-stage least squares approach to estimate the causal effect of Executive Order 181 on post-supervision participation.
+
+Results: I find that Executive Order 181 increased post-supervision registration and turnout for individuals discharged from parole as a whole; these results, however, mask racial heterogeneity.
+Although the participation of non-Black individuals increased substantially, the treatment effect for Black individuals was either much smaller or nonexistent.
+
+Conclusion: State policy can meaningfully increase the political participation of individuals even after they are no longer on parole.
+However, at least in the case of New York State, rights restoration prior to parole discharge appears less effective for Black individuals, who make up roughly 40 percent of the formerly incarcerated population.
 \end{abstract}
@@
 
-TABLE 1 ABOUT HERE
+\begin{singlespace}
+\begin{table}[H]
+
+\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
+\centering
+\begin{tabular}[t]{cc}
+\toprule
+Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
+\midrule
+Actual Birthdate & 69,644\\
+Birthdate + 35 Days & 502\\
+Birthdate - 35 Days & 426\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{singlespace}
 
@@
 
-TABLE 2 ABOUT HERE
+\begin{singlespace}
+
+\input{"../../temp/table3.tex"}
+\end{singlespace}
 
@@
 
-TABLE 3 ABOUT HERE
+\begin{singlespace}
+
+\input{"../../temp/table_whatever2.tex"}
+\end{singlespace}
 
@@
 
-TABLE 4 ABOUT HERE
+\begin{singlespace}
+
+\input{"../../temp/table4.tex"}
+\end{singlespace}
 
@@
 
-TABLE 5 ABOUT HERE
+\begin{singlespace}
+
+\input{"../../temp/table_whatever3.tex"}
+\end{singlespace}
 
@@
 
-FIGURE 1 ABOUT HERE
+\begin{figure}[H]
+
+{\centering \includegraphics{table_embed_files/figure-latex/sensitivity-1}
+
+}
+
+\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
+\end{figure}
 
@@
 
-TABLE 6 ABOUT HERE
+\begin{singlespace}
+
+\input{"../../temp/table6.tex"}
+\end{singlespace}
 
@@
 
-\newpage
-
-\begin{singlespace}
-\begin{table}[H]
-
-\caption{\label{tab:shift-dobs-chunk}\label{tab:change-dobs} Results of Shifting Birthdates}
-\centering
-\begin{tabular}[t]{cc}
-\toprule
-Group & \makecell[l]{Number of Matches Between\\DOCCS and Voter File Records}\\
-\midrule
-Actual Birthdate & 69,644\\
-Birthdate + 35 Days & 502\\
-Birthdate - 35 Days & 426\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\input{"../../temp/table3.tex"}
-
-\input{"../../temp/table_whatever2.tex"}
-
-\input{"../../temp/table4.tex"}
-
-\input{"../../temp/table_whatever3.tex"}
-
-\input{"../../temp/table6.tex"}
-\end{singlespace}
-
-\newpage
-\begin{figure}[H]
-
-{\centering \includegraphics{rewrite_files/figure-latex/sensitivity-1}
-
-}
-
-\caption{\label{fig:sens}Estimated Treatment Effect by Window Size}\label{fig:sensitivity}
-\end{figure}
-
 \end{document}

### 2020-11-04 code/part_2/04_match_parolees_voter.R
typo in table note

@@ -17,7 +17,7 @@ nys_roll <- select(nys_roll, -a, -b, -c)
 
 nys_roll <- nys_roll %>%
   mutate_at(vars(last_name, first_name, middle_name),
-            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
+            ~ tolower(gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .)))) %>% 
   mutate(dob = as.Date(as.character(dob), "%Y%m%d"))
 
 #### MATCH DOC TO VOTER FILE ####
@@ -25,7 +25,7 @@ nys_roll <- nys_roll %>%
 doccs_to_rolls <- readRDS("./temp/parolees_with_restoration.rds")
 doccs_to_rolls <- doccs_to_rolls %>% 
   mutate_at(vars(last_name, first_name, middle_name),
-            ~ gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .))) %>% 
+            ~ tolower(gsub("[[:punct:]]| ", "", ifelse(. == "", NA, .)))) %>% 
   rename(dob = dob_parole)
 
 merge_list <- match_rolls_to_doc(doccs_to_rolls, "din", nys_roll, "nys_id")

### 2020-11-05 markdown/part2/rewrite.tex
apparently I don't know how to write anymore

@@
 \usepackage{lineno}
-\linenumbers
 \usepackage{booktabs}
@@
 \author{Kevin Morris\footnote{Researcher, Brennan Center for Justice at NYU School of Law, 120 Broadway Ste 1750, New York, NY 10271 (\href{mailto:kevin.morris@nyu.edu}{\nolinkurl{kevin.morris@nyu.edu}})}}
-\date{October 27, 2020}
+\date{November 05, 2020}
 

### 2020-11-05 markdown/part2/work.Rmd
apparently I don't know how to write anymore

--- 2020-10-28 markdown/part2/table_embed.Rmd
@@
 output:
-  bookdown::pdf_document2:
+  word_document:
     toc: no
     fig_caption: yes
-    latex_engine: xelatex
-    keep_tex: true
 bibliography: "../welcome_home_vote.bib"
@@
 
-\begin{singlespace}
-```{r shift-dobs-chunk, echo = F}
-less_35 <- scales::comma(readRDS("./temp/m35_2.rds"))
-true <- scales::comma(nrow(readRDS("./temp/din_nys_parolees.rds")))
-plus_35 <- scales::comma(readRDS("./temp/p35_2.rds"))
-
-table <- data.frame(group = c("Actual Birthdate", "Birthdate + 35 Days", "Birthdate - 35 Days"), count = c(true, plus_35, less_35))
-
-j <- table %>% 
-  mutate_all(linebreak) %>% 
-  knitr::kable(booktabs = T, caption = "(\\#tab:change-dobs) Results of Shifting Birthdates", linesep = "", escape = F,
-               col.names = linebreak(c("Group", "Number of Matches Between\nDOCCS and Voter File Records")),
-               align = "c") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                position = "center")
-j
-
-```
-\end{singlespace}
+TABLE 1 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r logit-reg-chunk-short, echo = F, results = "hide", message = F, warning = F}
-load("temp/short_term_to.rdata")
-
-stargazer(date_reg_model_time, date_to_model_time,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Days off Parole Before Election", "Days off Parole Before Election\\textsuperscript{2}",
-                               "Voted in 2008",
-                               "Male", "Age (Years)",
-                               "Years on Parole"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
-          title = "\\label{tab:to-18-logit-short} 2018 Behavior for Control Group",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X"),
-                         c("Felony Class FE", "X", "X")),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = c(date_reg_ses, date_ses))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table3.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table3.tex"}
-\end{singlespace}
+TABLE 2 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r rd-demo-chunk, echo = F, warning = F}
-t <- readRDS("./temp/demos_rd_time.rds")
-colnames(t) <- c("Variable", "Control", "ITT", "p-value")
-j <- knitr::kable(t, booktabs = T, caption = "(\\#tab:demo-rd) Demographics of Control and ITT Groups", linesep = "",
-                  align = c("l", "c", "c"), "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F)
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- ""
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:demo-rd} Demographics of Control and ITT Groups}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever2.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever2.tex"}
-\end{singlespace}
+TABLE 3 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r logit-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("./temp/individual_turnout_18_itt.rdata")
-load("./temp/individual_turnout_16.rdata")
-
-ses_cl[["m4"]] <- m4_ses
-
-stargazer(models, model4,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Discharged After EO 181",
-                               "Black $\\times$ Discharged After EO 181",
-                               "Male", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018", "Voted in 2016"),
-          title = "\\label{tab:to-18-logit} Turnout in General Election",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X", "X"),
-                         c("Felony Class FE", "X", "X", "X", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          order = c("finished_post", "pb"),
-          se = ses_cl)
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{6}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table4.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table4.tex"}
-\end{singlespace}
+TABLE 4 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r rd-dmeans, echo = F, warning = F}
-k <- readRDS("./temp/means_diff_black.rds") %>% 
-  select(-Statistic)
-j <- knitr::kable(k, booktabs = T, caption = "(\\#tab:re) Turnout and Registration by Treatment Group, Race", linesep = "",
-                  align = c("l", "c", "c", "c")) %>% 
-  kable_styling(latex_options = c("HOLD_position"),
-                full_width = F) %>% 
-  group_rows(start_row = 1,
-             end_row = 2,
-             group_label = "Turnout") %>% 
-  group_rows(start_row = 3,
-             end_row = 4,
-             group_label = "Registration")
-
-write(j, "./temp/tempo.tex")
-
-j <- fread("./temp/tempo.tex", sep = "+", header = F)
-
-note.latex <- ""
-
-j <- j %>% 
-  mutate(n = row_number())
-
-lab <- "\\caption{\\label{tab:re} Turnout and Registration by Treatment Group, Race}"
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(note.latex, insert1), n = c(16.5, 1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-j$V1[4] <- lab
-
-write.table(j, "./temp/table_whatever3.tex", quote = F, col.names = F,
-            row.names = F)
-```
-\input{"../../temp/table_whatever3.tex"}
-\end{singlespace}
+TABLE 5 ABOUT HERE
 
@@
 
-```{r sensitivity, echo = FALSE, message = F, warning = F, fig.width = 7.5, fig.height = 6, fig.cap = "\\label{fig:sens}Estimated Treatment Effect by Window Size", fig.pos = "H", fig.align = "center"}
-j <- readRDS("./temp/reestimates.rds")
-print(j)
-```
+FIGURE 1 ABOUT HERE
 
@@
 
-\begin{singlespace}
-```{r iv-reg-chunk, echo = F, results = "hide", message = F, warning = F}
-load("temp/iv_individual_turnout_18.rdata")
-
-stargazer(iv3b, iv4b, iv3, iv4,
-          header = F,
-          type = "text", notes.align = "l",
-          covariate.labels = c("Rights Restored Before Discharge", 
-                               "Black $\\times$ Rights Restored Before Discharge",
-                               "Male", "Age (Years)",
-                               "Voted in 2008",
-                               "Years on Parole",
-                               "Days off Parole Before Election"),
-          dep.var.labels = c("Registered in 2018", "Voted in 2018"),
-          title = "\\label{tab:iv-tab} Two-Stage Least Squares",
-          table.placement = "H",
-          omit = c("race", "felony"),
-          keep.stat = c("n", "rsq", "adj.rsq"),
-          add.lines=list(c("Race / Ethnicity FE", "X", "X", "X", "X"),
-                         c("Felony Class FE", "X", "X", "X", "X")),
-          table.layout = "-dm#-t-a-s-n",
-          out = "./temp/test.tex",
-          notes = "TO REPLACE",
-          out.header = F,
-          se = list(iv3_sesb, iv4_sesb, iv3_ses, iv4_ses))
-
-j <- fread("./temp/test.tex", header = F, sep = "+")
-
-note.latex <- "\\multicolumn{4}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$.
-\\\\Robust standard errors in parentheses.\\\\First-stage equations include covariates in Models 2 and 3.}}}"
-
-j <- j %>% 
-  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
-         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
-  filter(!grepl("Note:", V1)) %>% 
-  mutate(n = row_number())
-
-insert1 <- "\\fontsize{10}{12}\\selectfont"
-
-j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
-  arrange(n) %>% 
-  select(-n)
-
-write.table(j, "./temp/table6.tex", quote = F, col.names = F,
-            row.names = F)
-
-```
-\input{"../../temp/table6.tex"}
-\end{singlespace}
+TABLE 6 ABOUT HERE
 
@@
 # References {-}
+<div id="refs"></div>
